### 2016-06-02 report/report.tex
first draft

% -----------------------------------------------------------------------------
% ------------------------------- PREAMBLE ------------------------------------
% -----------------------------------------------------------------------------[
	\documentclass[english]{scrartcl}

	\title{Spatial Complexity and Scaling in Cities}
	\subtitle{v0.0}
	\author{Phil Chodrow}
	\date{May 29th, 2016}

	\usepackage[T1]{fontenc}
	\usepackage[utf8]{inputenc}
	\usepackage{babel}
	\usepackage{blindtext}
	\usepackage{amsmath}
	\usepackage{amsthm}
	\usepackage{amsfonts}
	\usepackage{amssymb}
	\usepackage{enumitem}
	\usepackage{graphicx}
	\usepackage{pgfplotstable,filecontents}
	\usepackage{booktabs}
	\usepackage{caption}
	\usepackage{subcaption}
	\usepackage{csvsimple}

	\pgfplotsset{compat=1.9}% supress warning

	\setkomafont{disposition}{\normalfont\bfseries}

	\newtheorem{thm}{Theorem}
	\newtheorem{lm}{Lemma}
	\newtheorem{cor}{Corrolary}
	\newtheorem{clm}{Claim}
	\newtheorem*{thm*}{Theorem}
	\newtheorem*{lm*}{Lemma}
	\newtheorem*{cor*}{Corrolary}
	\newtheorem*{clm*}{Claim}
	\newtheorem*{fct*}{Fact}

	\newcommand\abs[1]{\left|#1\right|}
	\newcommand\E[0]{\mathbb{E}}
	\newcommand\R[0]{\mathbb{R}}
	\newcommand\decision[2]{\underset{#2}{\overset{#1}{\gtreqless}}}
	\newcommand{\argmax}{\operatornamewithlimits{argmax}}
	\newcommand{\argmin}{\operatornamewithlimits{argmin}}
	\newcommand\prob[0]{\mathbb{P}}
	\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
% -----------------------------------------------------------------------------
% --------------------------------- BODY --------------------------------------
% -----------------------------------------------------------------------------
\begin{document}

\maketitle

\abstract{We formulate the \emph{mean local information}, a novel measure of spatial compositional complexity in terms of the localized mutual information between spatial and compositional variables. This measure is closely related to the Fisher information of the underlying joint distribution, and is therefore an intrinsic property of the phenomena studied. Using this measure, we study spatial variations in racial trends in major American cities. The results neatly distinguish cities with large, monoracial neighborhoods from those with more intricate, interleaved neighborhood structure. Furthermore, the mean local information scales logarithmically with population density, inviting future research into possible universal dynamical processes of neighborhood formation that might explain this relationship.}

\section{Introduction}

	Information theory provides one natural approach toward thinking about complex systems. It achieves this by mapping the physical concept of structure to the epistemic concept of predictability. A complex system has enough structure to be at least partially predictable, but enough disorder to make prediction challenging. In this paper, we apply information-theoretic tools to the measurement of \emph{spatial compositional complexity}, with a focus on the spatial variation of racial trends in American cities.

	Information-theoretic concepts have already found application in urban planning problems related to zoning and predicting population distributions \cite{Royal2014,Batty1974,Batty1976,Battya}. Substantial recent work has addressed the measurement of difference and disparity in cities through an information-theoretic lens \cite{Theil1971,Bettencourt2015,Roberto2015a,Roberto2015}. Attractive features of information-theoretic measures for this purpose include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to multiple demographic phenomena, and the fact that Theil's (\cite{Theil1971}) original index satisfaction of many (though not all) of the invariance properties desirable for the measurement of segregation \cite{Reardon2002}. \nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015}.

	While closely-related to segregation, the concept of complexity has received less attention among quantitative sociologists. This is natural, since complexity is both less operational than segregation and challenging to define. On the other hand, as we demonstrate below, a spatial information-theoretic measure can be defined which measures the granularity of neighborhood structure, distinguishing cities with large, monolithic tracts of constant racial composition from those with many adjacent, smaller, neighborhoods of varying racial profiles. In doing so, it therefore captures one natural aspect of complexity in spatial compositional phenomena, and may also be of interest as a segregation measure as well.

\section{Information Theory and Spatial Structure}

	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two. Let $X$ be a spatial variable, such as a set of coordinates in space or a neighborhood name. Let $Y$ be a compositional variable that we aim to study, defined on an alphabet $\mathcal{Y}$. In the simple case in which $\mathcal{Y}  = \{\text{Black, White}\}$, Figure \ref{fig:toy} illustrates a range of possible joint distributions $p(X,Y)$. Comparing the completely undiverse city (a) and the spatially uniformly diverse (b) motivates the first fundamental information measure, the entropy:
	\begin{equation}
		H(Y) \triangleq \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
	\end{equation}
	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$. It thereby clearly distinguishes cities (a) and (b): since (a) is completely uniform, $H_a(Y) = 0$, while $H_b(Y) \neq 0$. On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c). To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
	\begin{align}
		D[p(Z)\|q(Z)] &\triangleq \sum_z p(z) \log \frac{p(z)}{q(z)} \\
		I(X,Y) &\triangleq D[p(X,Y) \| p(X)p(Y)]
	\end{align}
	The divergence $D$ is (with some caveats) a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$. Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent. In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).  In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence.

	City (d) shares with city (c) the fact that residence and race are maximally dependent. However, city (d) embodies the ``checkerboard problem'': measures that use only the joint distribution $p(X,Y)$ without additionally considering the spatial information contained within $X$ will evaluate city (d) to be the same as city (c), despite their considerably different patterns of racial separation and potentially very different implications for planning and policy. A recent working paper \cite{Roberto2015} provides one highly operational approach to this problem, using road network topology and a weighting function that decays with distance to define a localized measure based on the Kullback-Leibler divergence. Here we pursue an alternative strategy, showing that a localization of the mutual information both measures spatial variation in race and corresponds to the estimation of a fundamental statistical property of the distribution $p(X,Y)$.

	% NOTES FOR EXPANDING THE INTRODUCTION
	% \begin{itemize}
	% 	\item Concentration: $H(X|Y)$ with appropriate areal weighting. 
	% 	\item Centralization: Probably not available. 
	% 	\item Clustering: $J(X,Y)$
	% 	\item Exposure: $I(X,Y)$? Needs to be read in conjunction with $H(Y)$. 
	% 	\item Evenness: $H(X|Y)$, low score makes highly uneven. Or $I(X,Y)$? 
	% \end{itemize}

	% Possibly useful to note that \cite{Press2009a} express some dissatisfaction with their available clustering measures.

	% \cite{Press2009a}
	% \begin{enumerate}
	% 	\item Multi-layered aspects of diversity and segregation. A limitation of our methods is that we aren't getting into historical or sociological questions; an advantage is that  they generalize fully to other phenomena, such as class, education level, or job type. Discuss other measures of diversity. 
	% 	\item Discuss the toy cities, existing information measures, what they mean, and why they need a complementary local measure. 
	% 	\item Operational roles in modeling urban phenomena
	% \end{enumerate}

	\begin{figure}
		\centering
		  \includegraphics[width=.7\linewidth]{figs/toy.png}
		  \begin{tabular}{c | c c c}
			  City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
			  \hline			
			  (a) & 0.0 & 0.0 & 0.0\\
			  (b) & 0.7 & 0.0 & 0.0\\
			  (c) & 0.7 & 0.7 & 0.6\\
			  (d) & 0.7 & 0.7 & 2.7\\
			  \hline  
			\end{tabular}
		\caption{Information theory and the checkboard problem: model cities with various kinds of spatial diversity can be distinguished through progressively more subtle spatial information measures.}
		\label{fig:toy}
		\end{figure}

\section{Methods}
	To motivate our methods, we first consider an idealized problem in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is intuitively the map on which we work.

	Fix a point $x_0 \in M$ and a radius $r > 0$. Let $B_r(x_0)$ denote the ball of radius $r$ about $x_0$. Then, define the \emph{local mutual information in radius $r$} as the mutual information between $X$ and $Y$, restricted to the small ball $B_r(x_0)$ about $x_0$:
	\begin{equation}
		I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r(x_0))]|X \in B_r(x_0)]
	\end{equation}
	Intuitively, $I_r(x_0)$ measures how much the probability field $p(y|x)$ varies with $x$ in a small neighborhood of $x_0$. It is therefore a natural measure of local complexity, aligned in approach with the global mutual information but designed to detect pattern in local variations. $I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$, corresponding to a complete lack of local compositional complexity.

	\begin{figure}
		\centering
		\includegraphics[width=.5\textwidth]{figs/method.png}
		\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston). We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density. }
		\label{fig:method}
	\end{figure}

	In this definition, $r$ may be thought of as the spatial resolution at which we conduct analysis, and $I_r(x_0)$ is highly dependent on $r$. However, it is possible to show that $I_r(x_0)$ is related to a fundamental statistical property of the probability field $p(y|x)$ that is resolution-independent. Under the stated conditions, the following approximation holds:

	\begin{equation}
		\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace} (J_Y(x))\;, \label{approx}
	\end{equation}
	where $J_Y(x)$ is the Fisher information matrix in $Y$ about $x$, defined as 
	\begin{equation}
		J_Y(x) \triangleq \E\left[ (\nabla_x \log p(y|x))(\nabla_x \log p(y|x))^T \right]\;.
	\end{equation}
	A more formal statement and proof of \eqref{approx} are provided in Appendix 2. The Fisher information $J_Y$ is a fundamental quantity in statistics and information theory. From a geometric perspective, $J_Y$ provides the natural intrinsic metric in the geometric space of probability distributions parameterized by the spatial variable $x$. Equation \eqref{approx} therefore expresses a relationship between the local mutual information $I_r(x)$ and the information geometry of the underlying probability field. Corresponding to the fact that $I_r(x_0)$ vanishes if and only if $p(y|x)$ is constant in $B_r(x_0)$, $J_Y(x_0) = 0$ if and only if $\nabla_x p(y|x_0) = 0$ for all $y$. This implies that $x_0$ is a stationary point, about which the probability field $p(y|x_0)$ exhibits only small (2nd order or smaller) changes with respect to changes in $x$.

	Since the Fisher information is a strictly local measure of statistical variability around $x$, we can aggregate the Fisher information to derive a measure of average local variability. The \emph{mean local information} is 
	\begin{equation}
	J(X,Y) \triangleq \E_X[\text{trace }J_Y(X)]
	\end{equation} 
	We propose the aggregate quantity $J(X,Y)$ as a third  measure--alongside the entropy $H(Y)$ and mutual information $I(X,Y)$-- as a tool for the information-theoretic structure of spatial compositional complexity.

\section{Results}
	We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S. Census Bureau, on race and ethnicity for counties corresponding to 28 large US cities. At this early stage, the counties used were chosen to correspond to the most densely populated urban cores; in future developments we could standardize to conduct analyses for Metropolitan Statistical Areas. We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'. For each city, we computed the entropy $H(Y)$ and the mutual information $I(X,Y)$. To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 1km. We then computed the estimated mutual information within each grid cell, and averaged the results weighted by population population. A technical specification of this approach is provided in Appendix 1, and a complete summary of cities and their associated information measures is provided in Appendix 3.

	\begin{figure}
		\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
		\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
		\label{fig:info_cross}
	\end{figure}	
	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$. An overall positive trend is evident, reflecting the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.  On the other hand, there are substantial variations in $J(X,Y)$ even in cities with comparable global variability $I(X,Y)$. For example, the cities of Detroit and Philadelphia provide a striking contrast. While they have comparable global variability $I(X,Y)$, Philadelphia's $J(X,Y)$ is substantially higher. This reflects the fact that Detroit is composed of large, highly-segregated, monoracial neighborhoods, whereas Philadelphia has a much more fine-grained, intricate neighborhood structure. These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities.

	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density. In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities. A clear trend is evident, with $J(X,Y)$ growing linearly with the logarithm of density. We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix. One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own.

	\begin{figure}
		\includegraphics[width=1\textwidth]{figs/density_fisher.png}
		\caption{The mean local information $J(X,Y)$ scales with the logarithm of population density.}
		\label{fig:density}
	\end{figure}

\section{Discussion}

	We have formulated a novel measure of urban spatial complexity, and applied it to an analysis of spatial distributions of race in American cities. We emphasize that the mean local spatial variability $J(X,Y)$ is novel piece of a more complete information-theoretic characterization that also includes the entropy $H(Y)$ and the global variability $I(X,Y)$. Our results suggest two major directions of further exploration:

	One major physical question raised by our results is the origin of the scaling relation seen in \ref{fig:density}. This scaling relation may suggest a dynamical process of neighborhood formation and growth common to many American cities. If so, a physical understanding of this process in terms of individual-level behavior is called for. It is tempting to view the scaling relationship as an interplay of spatial preferential attachment and intrinsic limits to neighborhood density due to housing availability, but this view is speculative at this point. A theory to explain this behavior would be most welcome.

	Another question raised is more operational. In many urban planning contexts, the amount of available data may make direct computations prohibitive. In such contexts, it may be useful to construct a simplifying model of the data the data (such as a clustering) in order to reduce dimensionality and make ``the big picture'' more visible. When a particular demographic phenomenon (such as race) is under consideration, it may be important to ensure that the model preserves the major patterns of variation in the granular data. We conjecture that the local variability $J(X,Y)$ measures the ``model-ability'' of such data sets. Intuitively, a city like Detroit with relatively low $J(X,Y)$ and large, monoracial neighborhoods should be easily representable with relatively few modeled clusters, one for each of the relatively few major neighborhoods. On the other hand, relatively speaking, a city like Philadelphia with very intricate neighborhood structure may require substantially greater model complexity to represent without large loss of information. A promising course of further study is to design an information-theoretic clustering algorithm designed to model urban patterns, and then compare this algorithm's performance to $J(X,Y)$.

	A third question and final question relates to the time-dependence of spatial structure. On one time scale, daily movement around a city for work or leisure activities has the impact of ``mixing'' separated residents in public spaces, potentially leading to very different spatial patterns of racial difference. On another time scale, these measures may track how the spatial structure of cities evolves over decades, potentially shedding further light on the dynamics of city formation.

\bibliography{/Users/phil/bibs/library.bib}{}
\bibliographystyle{plain}

\section*{Appendix 1: Computational Methods and Assumptions}
	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[J_Y(X)]$ using blockgroup level data from the U.S. Census.

	For fixed Census blockgroup $i$, let $P_i$ be the population, let $A_i$ be the area, let $\rho_i = P_i / A_i$ be the population density, and let $p^i_Y(y)$ be the observed proportion of racial group $y$. For hex $k$ in our hexagonal grid, let $N_k$ be the set of overlapping Census blockgroups. We also define $p^{k}_I(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$. This definition embodies a computationally-simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area. Finally, $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$ is the estimated overall racial composition of hex $k$. Then, we estimate the mutual information in hex $k$ as 
	\begin{equation}
		I(k) = \sum_{i \in N_k} p^k_I(i) D[p^i_Y(\cdot) \| p^k_Y(\cdot)]\;. 
	\end{equation}
	Using \eqref{approx}, the estimated Fisher information is 
	\begin{equation}
		J(k) \approx \frac{4 I(k)}{r^2}
	\end{equation}
	where $r$ is the grid radius. The estimated population in hex $k$ is $P_k = A_k\sum_{i \in N_k} \rho_i$, where $A_k$ is the cell area. We finally estimate $\E_X[J(X)]$ as 
	\begin{equation}
		J(X,Y) = \E_X[J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
	\end{equation}
\section*{Appendix 2: Relation of Local Mutual Information and Fisher Information}
	In this appendix, we present a rigorous mathematical justification of the claim that the local mutual information approximates a fundamental quantity intrinsic to the distribution $p(Y,X)$. The argument relies on an idealization in which, rather than have a finite, discrete set of observations for each tract, we have a continuous field of observations across the map. The precise statement of this result and its justification follow.

	Let $X$ be a continuous random variable taking values in $\R^n$, and let $Y$ be a discrete random variable define on finite alphabet $\mathcal{Y}$. Suppose further that $p(y|x) > 0$ and that $p(y|x)$ is differentiable as a function of $x$ for all $x,y \in \R^n \times \mathcal{Y}$. Fix $x_0 \in R^n$, and define $B_r \triangleq B_r(x_0) = \{ x \in R^n \;|\; \norm{x - x_0} \leq r \}$. Additionally, define the \emph{local mutual information} in $B_r$ as
	\begin{equation}
		I_r \triangleq I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;.
	\end{equation}
	where $D[p\|q] = \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$.

	\begin{thm} \label{thm1}
		Under the stated conditions, 
		\begin{equation}
			\lim_{r\rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{n}{2(n+2)} \emph{trace}\; J_Y(x_0)\;.
		\end{equation}
		where the Fisher information $J_Y(x_0)$ is 
		\begin{align}
		J_Y(x) &\triangleq \E\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
		S_y(x) &\triangleq \log p(y|x)\;.
	\end{align}
	\end{thm}
	The proof of Theorem \ref{thm1} proceeds by the application of a number of Taylor approximations, in tandem with a fundamental relationship of information geometry. We first expand out $I_r$ explicitly as
	\begin{equation}
		I_r = \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;. \label{eq:explicit}
	\end{equation}

	\begin{lm} \label{approximations}
		The following approximation relationships hold for the components of \eqref{eq:explicit}:
		\begin{enumerate}[label=\emph{(\alph*)}]
			\item $p(X\in B_r,Y) = p(x_0, Y) v(B_r) + O(r^{n+2})$
			\item $p(Y|X \in B_r) = p(Y|x_0) + e_y$ where the error terms $e_y$ satisfy $e_y \in O(r^{2})$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
			\item $p(x|X \in B_r) $
		\end{enumerate}
	\end{lm}
	\begin{proof}
		For each approximation,  we directly apply Taylor expansions about $X = x_0$. 
		\begin{enumerate}[label=(\alph*)]
			\item We have 
				\begin{align}
					p(X\in B_r, Y) &= \int_{B_r} p(x,Y) \; d^nx \\
					&= \int_{B_r} p(x_0,Y) + \frac{\partial p(x_0, Y)}{\partial x}(x - x_0) + O(\norm{x - x_0}^2) \; d^nx \\
					&= p(x_0, Y)v(B_r) +  \frac{\partial p(x_0, Y)}{\partial x} \int_{B_r} (x - x_0)\; d^nx  \\
					&\quad \quad+ O\left(\int_{B_r} \norm{x - x_0}^2\; d^nx\right) \\
					&= p(x_0, Y)v(B_r) + O(r^{n+2})\;,
				\end{align}
				where the middle term vanishes due to spherical symmetry. 
			\item The fact that the error terms $e_y$ must satisfy $\sum_{y \in \mathcal{Y}} e_y = 0$ follows from the fact that $p(Y|X \in B_r)$ must be a valid probability distribution over $\mathcal{Y}$. We'll now show that $e_y \in O(r^{ß2})$. First, 
				\begin{align}
					p(X \in B_r) &= \sum_{y\in \mathcal{Y}} p(X \in B_r, y)  \\
					&= \sum_{y\in \mathcal{Y}} \left[p(x_0, y)v(B_r) + O(r^{n+2}) \right] \\
					&= p(x_0)v(B_r) + O(r^{2});\,
				\end{align}
			from part (a). Next, 
				\begin{align}
					p(Y|X \in B_r) &= \frac{p(X\in B_r, Y)}{p(X \in B_r)} \\
					&= \frac{p(x_0, Y)v(B_r) + O(r^{n+2})}{p(x_0)v(B_r) + O(r^{n+2})} \\
					&= p(Y|x_0) + O(r^2)\;,
				\end{align}
			which completes this part of the argument. 
			\item First, 
				\begin{align}
					p(X \in B_r) &= \int_{B_r} p(x) \; d^nx  \\
					&= \int_{B_r} \left[p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)\right] \; d^nx \\ 
					&= p(x_0) v(B_r) + O(r^{n+2})\;,
				\end{align}
				where the middle term again vanishes through sphericla symmetry. Thus, for $x \in B_r$, we have
				\begin{align}
					p(x | X \in B_r) &= \frac{p(x)}{p(X \in B_r)} \\
					&= \frac{p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)}{p(x_0) v(B_r) + O(r^{n+2})} \\
					&= \frac{1 + O(r)}{v(B_r)}\;.
				\end{align}
		\end{enumerate}
	\end{proof}

	\begin{lm} The following approximation holds for the divergence factor in the integral \eqref{eq:explicit}
		\begin{equation}
			D[p(\cdot|x)\| p(\cdot|X \in B_r)] = D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})
		\end{equation}
	\end{lm}
	\begin{proof}
		We compute directly: 
		\begin{align}
			D[p(\cdot|x)\| p(\cdot|X \in B_r)] &= \sum_{y \in \mathcal{Y}} p(y|x) \log \frac{p(y|x)}{p(y|X \in B_r)} \\
			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|X \in B_r) \\
			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log \left(p(y|x_0) + e_y)\right) \tag{from Lemma \ref{approximations}}\\
			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \left[\log p(y|x_0) + \frac{e_y}{p(y|x_0)} + O(e_y^2)\right] \\
			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}  e_y \tag{quadratic terms negligible} \\
			&=D[p(\cdot|x)\| p(\cdot|x_0)] \\
			&\quad + \sum_{y \in \mathcal{Y}} \left(1 + \frac{1}{p(y|x_0)} \nabla p(y|x_0)(x - x_0) + O(r^{2})\right)    e_y \\
			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \left[e_y + O(r^{3})\right] \tag{$e_y \in O(r^2)$}\\
			&= D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^3) \tag{$\sum_{y \in \mathcal{Y}} e_y = 0$}
		\end{align}
	\end{proof}

	\begin{lm} For any positive-semidefinite matrix $A \in \R^{n\times n}$, 
		$$\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x = \frac{n}{n+2} r^{2} v(B_r) \text{\emph{trace}} (A)$$
	\end{lm}

	\begin{proof}
		Since $A$ is positive-semidefinite, there exist an orthonormal matrix $P$ and a diagonal matrix $D$ such that $A = P^TDP$. Furthermore, the entries of $D$ are the eigenvalues $\{\lambda_i\}$ of $A$. Then, 
		\begin{align}
			\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0, P^T DP(x-x_0) \right> d^n x\\
			&= \int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x \;.
		\end{align}
		We can regard $P$ as a reparameterization of $B_r$; since $\det P = 1$, we have 
		\begin{align}
			\int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0,  D(x-x_0) \right> d^n x \\
			&= r^n\int_{B_n} \left< rx, rDx \right> d^n x \\
			&= r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x \;,
		\end{align}
		where $B_n$ is the unit $n$-ball. We also let $S_n(r)$ be the $n$-sphere of radius $r$. Continuing, 
		\begin{align}
			r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x &= r^{n+2} \int_{B_n} \sum_{i= 1}^n x_i^2 \lambda_i \; d^nx \\
			&= r^{n+2} \sum_{i = 1}^n \lambda _i \int_{B_n} x_i^2 \; d^nx \\
			&= \frac{r^{n+2}}{n} \sum_{i = 1}^n \lambda _i \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
			&= \frac{r^{n+2}}{n} \text{trace}(A)  \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^2 S_{n-1}(\rho) d\rho \\
			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^{n+1} S_{n-1}(1) d\rho \\
			&= \frac{r^{n+2}}{n} \text{trace} (A)  \frac{1}{n+2} S_{n-1}(1) \\
			&= \frac{r^2}{n+2}  \text{trace} (A)  n r^{n}v(B_n(1)) \\
			&= \frac{n}{n+2} r^{2} v(B_r) \text{trace} (A)\;,
		\end{align}
		as was to be shown. 
	\end{proof}

	\begin{fct*}
		The Kullback-Leibler divergence and the Fisher information $J_Y$ are related according to the approximation 
		\begin{equation}
			D[p(\cdot|x)\| p(\cdot|x_0)] = \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3)
		\end{equation}
	\end{fct*}

	We are finally ready to prove Theorem \ref{thm1}. Computing directly, we have 
	\begin{align}
		I_r(x_0) &\triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;. \\
		&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \\
		&= \int_{B_r} \left[\frac{1 + O(r)}{v(B_r)}\right] D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \tag{Lemma 1(c)}\\ 
		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r}  \left(D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})\right) d^n x \tag{Lemma 2}\\
		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left( \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3) + O(r^{3})\right) d^n x \\
		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left(\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(r^{3})\right) d^n x \\
		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\left(\frac{n}{n+2} r^{2} v(B_r) \text{trace} (J_Y(x_0)) + v(B_r)O(r^{3})\right) \\
		&= r^2\frac{n}{2(n+2)}\left[1 + O(r)\right]\left(\text{trace} (J_Y(x_0)) + O(r^{3})\right)\;.
	\end{align}
	Dividing through by $r^2$ and computing the limit as $r \rightarrow 0$ proves the result. 
\section*{Appendix 3: Table of Information Measures for Cities Studied}
	The table below is sorted  by the population density of the city studied. 
	\begin{tabular}{l c c c c }%
	    \bfseries City & Density (km$^{-2}$) & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ 
	    \csvreader[head to column names]{figs/table.csv}{}% use head of csv as column names
	    {\\\hline\City & \Density & \HY & \IXY & \J}% specify your coloumns here
    \end{tabular}

\end{document}

### 2016-06-02 analysis.R

@@ -1,0 +1,299 @@
+library(sp)
+library(compx)
+library(acs)
+library(plyr)
+library(dplyr)
+library(readr)
+library(ggmap)
+library(tidyr)
+library(magrittr)
+library(ggrepel)
+library(grid)
+
+setwd("~/projects/spatial_complexity/tech_report")
+
+
+# Toy cities ----------
+
+expand.grid(x = 1:8, y = 1:8) %>% 
+	mutate(`(a)` = 1,
+		   `(b)` = .5,
+		   `(c)` = (x > 4)*1,
+		   `(d)` = (x + y) %% 2) %>%
+	gather(key = model, value = p, -x, -y) %>%
+	ggplot(aes(x = x, y = y)) + 
+	theme_minimal() +
+	theme(axis.ticks = element_blank(), 
+		  axis.text.x = element_blank(), 
+		  axis.text.y = element_blank(),
+		  panel.background = element_rect(),
+		  panel.grid.major = element_line(size = 0),
+		  panel.grid.minor = element_line(size = 0),
+		  plot.margin=unit(c(0,0,0,0),"mm")) + 
+	geom_tile(aes(fill = p)) + 
+	scale_fill_continuous(low = 'white', high = 'black ', limits=c(0,1)) + 
+	facet_grid(~model) +
+	xlab('') + 
+	ylab('') + 
+	guides(fill=FALSE)
+
+ggsave('report/figs/toy.png', width = 5, height = 1.9)
+
+
+# Methodological Illustration ----------------
+
+boston <- compx::get_race_data(state = 'MA', counties = c(25))
+boston@data$area <- 1:length(boston) %>%
+	as.matrix() %>%
+	apply(MARGIN = 1,FUN = function(i) boston@polygons[[i]]@Polygons[[1]]@area * 85 * 111) # lat/lon conversion (approximate)
+
+boston@data$density <- boston@data$total / boston@data$area
+
+races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
+
+xx = spsample(boston, type="hexagonal", cellsize=.01)
+xxpl = HexPoints2SpatialPolygons(xx)
+grid_radius = .01 * sqrt(85 * 111)
+
+# Define information measures on the grid 
+h <- function(i){
+	window <- boston[xxpl[i,],]@data[,c(races, 'total', 'area')]
+	window <- window / window$area # total becomes density
+	c(mean(window$total), 4 * mutual_info(window[,races])/grid_radius^2) # returns estimated density and mutual info
+}
+
[…]

### 2016-06-04 analysis.R
msa level

@@ -9,118 +9,46 @@ library(tidyr)
 library(magrittr)
 library(ggrepel)
 library(grid)
+library(rgdal)
 
 setwd("~/projects/spatial_complexity/tech_report")
 
 
 # Toy cities ----------
 
-expand.grid(x = 1:8, y = 1:8) %>% 
-	mutate(`(a)` = 1,
-		   `(b)` = .5,
-		   `(c)` = (x > 4)*1,
-		   `(d)` = (x + y) %% 2) %>%
-	gather(key = model, value = p, -x, -y) %>%
-	ggplot(aes(x = x, y = y)) + 
-	theme_minimal() +
-	theme(axis.ticks = element_blank(), 
-		  axis.text.x = element_blank(), 
-		  axis.text.y = element_blank(),
-		  panel.background = element_rect(),
-		  panel.grid.major = element_line(size = 0),
-		  panel.grid.minor = element_line(size = 0),
-		  plot.margin=unit(c(0,0,0,0),"mm")) + 
-	geom_tile(aes(fill = p)) + 
-	scale_fill_continuous(low = 'white', high = 'black ', limits=c(0,1)) + 
-	facet_grid(~model) +
-	xlab('') + 
-	ylab('') + 
-	guides(fill=FALSE)
-
+checkerboard_illustration()
 ggsave('report/figs/toy.png', width = 5, height = 1.9)
-
 
 # Methodological Illustration ----------------
 
-boston <- compx::get_race_data(state = 'MA', counties = c(25))
-boston@data$area <- 1:length(boston) %>%
-	as.matrix() %>%
-	apply(MARGIN = 1,FUN = function(i) boston@polygons[[i]]@Polygons[[1]]@area * 85 * 111) # lat/lon conversion (approximate)
+# method_illustration()
+# ggsave('report/figs/method.png', width = 3, height = 3)
 
-boston@data$density <- boston@data$total / boston@data$area
 
-races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
 
-xx = spsample(boston, type="hexagonal", cellsize=.01)
-xxpl = HexPoints2SpatialPolygons(xx)
-grid_radius = .01 * sqrt(85 * 111)
-
-# Define information measures on the grid 
-h <- function(i){
-	window <- boston[xxpl[i,],]@data[,c(races, 'total', 'area')]
-	window <- window / window$area # total becomes density
[…]

### 2016-06-04 get_data.R

@@ -1,0 +1,132 @@
+setwd("~/projects/spatial_complexity/tech_report")
+
+library(readr)
+library(dplyr)
+library(compx)
+library(tidyr)
+library(rgdal)
+
+
+cities <- read_csv('msas.csv')
+
+cities <- cities %>%
+	separate(name, c('name','state'), sep = ',') %>%
+	mutate(state = substr(county_name, start  = nchar(county_name) - 1, stop  = nchar(county_name)),
+		   county_num = as.integer(substr(county_num, start = 3, stop = 5))) %>%
+	select(-county_name)
+
+county_list <- aggregate(county_num ~ state, c, data = cities)
+
+f <- function(state){
+	
+	counties <- county_list$county_num[county_list$state == state][[1]]
+	
+	race <- acs::acs.fetch(endyear = 2014,
+						   span = 5,
+						   geography = acs::geo.make(state = state, county = counties, tract = '*', block.group = '*'),
+						   table.number = 'B03002',
+						   col.names = "pretty")
+	
+	race <- cbind(data.frame(race@geography),
+				  data.frame(race@estimate)) %>%
+		tbl_df() %>%
+		mutate(GEOID = paste0(str_pad(state, 2, 'left', pad = '0'),
+							  str_pad(county, 3, 'left', pad = '0'),
+							  str_pad(tract, 6, 'left', pad = '0'),
+							  str_pad(blockgroup, 1, 'left', pad = '0')))
+	
+	race$Hispanic <- race[names(race)[!grepl('.Not.Hispanic.or.Latino.',names(race))]] %>%
+		select(-(`NAME`:`Hispanic.or.Latino.by.Race..Hispanic.or.Latino.`)) %>%
+		select(-GEOID) %>%
+		rowSums()
+	
+	others <- c('Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..American.Indian.and.Alaska.Native.alone',
+				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Native.Hawaiian.and.Other.Pacific.Islander.alone',
+				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Some.other.race.alone',
+				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Two.or.more.races.',
+				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Two.or.more.races..Two.races.including.Some.other.race',
[…]

### 2016-06-05 analysis.R
ua level

@@ -16,8 +16,8 @@ setwd("~/projects/spatial_complexity/tech_report")
 
 # Toy cities ----------
 
-checkerboard_illustration()
-ggsave('report/figs/toy.png', width = 5, height = 1.9)
+# checkerboard_illustration()
+# ggsave('report/figs/toy.png', width = 5, height = 1.9)
 
 # Methodological Illustration ----------------
 
@@ -31,54 +31,58 @@ ggsave('report/figs/toy.png', width = 5, height = 1.9)
 
 # Prep city list 
 
-cities <- city_list('msas.csv')
-density_list <- read_csv('density_list.csv') %>% filter(density > 200)
+# cities <- city_list('msas.csv')
+# density_list <- read_csv('density_list.csv') %>% filter(density > 200)
 # cities <- cities %>% filter(city == 'New York City')
 
-f <- function(city){
-	
-	# Pull down race data 
-	sub <- cities[cities$name == city,]
-	
-	g <- function(i){
-		state    <- sub$state[i]
-		tracts <- readOGR(dsn = paste0('data/',state), layer = 'geo', verbose = FALSE)
-		tracts@data <- tracts@data %>%
-			mutate(county = substr(GEOID, start = 3, stop = 5),
-				   county = as.integer(county))
-		counties <- unlist(sub$county[i])
-		tracts <- tracts[is.element(tracts@data$county, counties),]
-		tracts   <- spChFIDs(tracts,as.character(paste(state,rownames(as(tracts,"data.frame")),sep="_")))
-	}
-	
-	tracts <- lapply(1:nrow(sub), g)
-	race <- do.call(rbind, tracts)
-	
-	# Modify the race data a bit  
-	
-	# conversion (approximate)
-	race@data$area <- 1:length(race) %>%
-		as.matrix() %>%
-		apply(MARGIN = 1,FUN = function(i) race@polygons[[i]]@Polygons[[1]]@area * 85 * 111) # lat/lon
-	
-	# Make the grid  
+
+# Cache logic
+cities <- list.files('cities') %>%
+	sort(decreasing = T)
+
+if(file.exists('cities_cache.csv')){
+	cache <- read_csv('cities_cache.csv')
+	cities <-  cities[!is.element(cities, cache$city)]
+}else{
+	cache <- data.frame(
+		city = character(),
+		area = numeric(),
+		population = numeric(),
+		density = numeric(),
[…]

### 2016-06-05 get_data.R

@@ -1,25 +1,30 @@
 setwd("~/projects/spatial_complexity/tech_report")
 
-library(readr)
-library(dplyr)
+library(sp)
 library(compx)
+library(acs)
+library(plyr)
+library(dplyr)
+library(readr)
+library(ggmap)
 library(tidyr)
+library(magrittr)
+library(ggrepel)
+library(grid)
 library(rgdal)
+library(tigris)
+library(stringr)
+library(acs)
 
 
-cities <- read_csv('msas.csv')
-
-cities <- cities %>%
-	separate(name, c('name','state'), sep = ',') %>%
-	mutate(state = substr(county_name, start  = nchar(county_name) - 1, stop  = nchar(county_name)),
-		   county_num = as.integer(substr(county_num, start = 3, stop = 5))) %>%
-	select(-county_name)
-
-county_list <- aggregate(county_num ~ state, c, data = cities)
+ua <- read_table('ua_list_all.txt')
+ua$state = word(ua$NAME, -1)
 
 f <- function(state){
-	
-	counties <- county_list$county_num[county_list$state == state][[1]]
+	counties <- acs.fetch(geography=geo.make(state=state, county="*"), 
+				   endyear = 2014,
+				   table.number="B01003")
+	counties <- as.numeric(geography(counties)[[3]])
 	
 	race <- acs::acs.fetch(endyear = 2014,
 						   span = 5,
@@ -54,79 +59,28 @@ f <- function(state){
 			   White = Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..White.alone,
 			   Black = Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Black.or.African.American.alone,
 			   Asian = Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Asian.alone,
-			   GEOID)
-	
+			   GEOID) %>%
+		mutate(total = Hispanic + White + Black + Asian + Other)
+
 	tracts <- tigris::block_groups(state = state, county = counties, cb=TRUE)
+	tracts@data$area <- tracts@data$ALAND / 1000^2
 	
 	tracts <- tigris::geo_join(tracts, race, "GEOID", "GEOID")
-	
-	tracts@data$area <- 1:length(tracts) %>%
-		as.matrix() %>%
-		apply(MARGIN = 1,FUN = function(i) tracts@polygons[[i]]@Polygons[[1]]@area * 85 * 111) # lat/lon
-	
-	tracts@data <- tracts@data %>%
[…]

### 2016-06-05 make_uas.R

@@ -1,0 +1,69 @@
+setwd("~/projects/spatial_complexity/tech_report")
+
+library(sp)
+library(compx)
+library(acs)
+library(plyr)
+library(dplyr)
+library(readr)
+library(ggmap)
+library(tidyr)
+library(magrittr)
+library(ggrepel)
+library(grid)
+library(rgdal)
+library(tigris)
+library(stringr)
+library(acs)
+
+
+
+ua_poly <- tigris::urban_areas(cb = TRUE)
+ua <- read_tsv('ua_list_all.txt')
+ua$state = word(ua$NAME, -1)
+
+cities <- ua %>% 
+	filter(POP > 1e6) 
+
+states <- cities$state %>% 
+	str_split('--') %>% 
+	do.call(c,.) %>% 
+	unique()
+
+cities <- cities$NAME
+
+shapes <- list()
+for(state in states){
+	shapes[state] <- readOGR(dsn = paste0('data/',state), layer = 'geo', verbose = FALSE)
+}
+
+
+for(city in cities){
+	print(city)
+	city_info <- ua %>%
+		filter(NAME == city) 
+	
+	city_poly <- ua_poly[ua_poly@data$NAME10 == city,]
+	
+	
+	city_components <- list()
+	for(state in str_split(city_info$state, '--')[[1]]){
+		state_sub <- shapes[[state]][city_poly,]
+		city_components[state] <-  spChFIDs(state_sub,
+											as.character(paste(state,rownames(as(state_sub,"data.frame")),
+															   sep="_")))
+	}
+	
+	city_polys <- do.call(rbind, city_components)
+
+	tryCatch(writeOGR(city_polys, paste0("cities/", city),'geo', driver = 'ESRI Shapefile', morphToESRI = TRUE),
+			 error = function(e) NULL)
+}
+
+
+
+
+
+
+
+

### 2016-06-16 5_figs.Rmd
reorganized

---
title: "Report and Figures"
author: "Phil Chodrow"
date: "June 15, 2016"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r warning = FALSE, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(compx)
library(ggplot2)
library(rgdal)
library(ggmap)
library(dplyr)
library(gridExtra)
library(readr)
library(ggrepel)
library(DT)
library(caTools)
library(tidyr)

columns = c('Asian', 'Black', 'Other', 'White', 'Hispanic')

```

# Purpose

The purpose of this document is to generate the figures used in this project. It is not a complete technical report; see the full writeup if that's what you're looking for.

# Methods

## Motivation

```{r, fig.height=3}
checkerboard_illustration()
ggsave('figs/checkerboard.png', width = 4, height = 1.5)
```

## Methodological illustration
```{r, warning = FALSE, message = FALSE, eval = FALSE}
columns <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
method_illustration <- function(city){
	hexgrid <- readOGR(dsn = paste0('throughput/grids/',city), layer = 'geo', verbose = FALSE)
	df <- hexgrid@data
	df <- df %>% mutate(id = as.character(row_number()))
	hexgrid <- fortify(hexgrid) %>%
		left_join(df, by = c('id' = 'id'))
	bbox <- c(min(hexgrid$long), min(hexgrid$lat), max(hexgrid$long), max(hexgrid$lat))
	map <- get_map(location = bbox, maptype = 'terrain-background')

	p <- ggmap(map,darken = .5)
	p <- p +
		geom_polygon(data = hexgrid, aes(x = long, y = lat, group = group, fill = info), alpha = .8) +
		scale_fill_continuous(low = 'white', high = 'steelblue', name= expression(J[Y](x)), limits = c(0, log(length(columns)) / .5^2)) +
		theme(axis.ticks = element_blank(),
			  axis.text.x = element_blank(),
			  axis.text.y = element_blank()) +
		xlab('') +
		ylab('') +
		theme(legend.justification=c(.8,0),
			  legend.position=c(1,0),
			  legend.background = element_rect(fill = alpha('blue', 0)),
			  legend.title = element_text(colour="white"),
			  legend.text = element_text(colour="white"),
			  legend.key.size = unit(2, "mm"),
			  legend.text = element_text(size = rel(.1))) +
		ggtitle(paste0('Local Mutual Information in ', city, ': Mean = ', round(weighted.mean(df$info, df$density), 2)))
	p
}

grid.arrange(method_illustration('Philadelphia'),
			 method_illustration('Detroit'),
			 ncol = 2)

ggsave('figs/method.png')
```

#  Mean Local Information

```{r}

cache <- read_csv('throughput/info_cache.csv')

```

## Relationship to mutual information

```{r}
highlight <- cache %>%
	filter(I_XY > .4)

cache %>%
	ggplot(aes(x = I_XY, y = weighted_J)) + 
	geom_point(color = 'grey') + 
	theme_minimal() + 
	ylab(expression(J(X,Y))) + 
	xlab(expression(I(X,Y))) + 
	ggtitle('Mutual information and mean local information in major US cities') +
	geom_point(data = highlight, color = 'firebrick') + 
	geom_text_repel(data = highlight, aes(label = city))

ggsave('figs/mutual_fisher.png', width = 8, height = 4)

```

## Relationship to population density
```{r}
lm_eqn <- function(df){
	m <- lm(weighted_J ~ log(density), data = df) 
	eq <- substitute(italic(y) == a + b %.% ~'log' ~italic(rho)*","~~italic(r)^2~"="~r2, 
					 list(a = format(coef(m)[1], digits = 2), 
					 	 b = format(coef(m)[2], digits = 2), 
					 	 r2 = format(summary(m)$adj.r.squared, digits = 3)))
	as.character(as.expression(eq))                 
}

cache %>% 
	# filter(!grepl('Los Angeles', city)) %>%
	ggplot(aes(x = density, y = weighted_J)) +
	geom_smooth(method = 'lm', se = FALSE, color = 'black', size = .4) +
	annotate('text', x = 3000, y = .2, label = lm_eqn(cache), parse = T) + 
	geom_point(color = 'grey') + 
	geom_point(color = 'firebrick', data = highlight) + 
	geom_text_repel(aes(label = city), size = 3, data = highlight) +
	scale_x_continuous(trans = 'log10') +
	# scale_y_continuous(trans = 'log10') +
	theme_minimal() +
	xlab(expression(rho~(population/km^{2}))) +
	ylab(expression(J(X,Y))) +
	annotation_logticks(size = .1, sides = 'b') +
	ggtitle('Dependence of spatial complexity on population density')

ggsave('figs/density_fisher.png', width = 8, height = 4)
```

## Raw measures

```{r}
# Table --------

tab <- cache %>% 
	select(area, population, H_Y, I_XY, J = weighted_J) %>% 
	mutate(Density = population / area) %>% 
	round(2) %>%
	cbind(cache$city, .) %>%
	rename(City = `cache$city`, 
		   `Area` = area,
		   Population = population,
		   HY = H_Y,
		   IXY = I_XY) %>%
	arrange(desc(Density)) %>% 
	select(City, `Area`, Population, Density, HY, IXY, J)

tab %>% write_csv('figs/table.csv')

tab$Population <- format(tab$Population, big.mark=",", scientific=FALSE)
tab$Area <- format(tab$Area, big.mark=",", scientific=FALSE)
tab$Density <- format(tab$Density, big.mark=",", scientific=FALSE)

datatable(tab, options = list(pageLength = 10), rownames = FALSE,
		  caption = 'Table 1: This is a simple caption for the table.')

```

# Clusterings

```{r}

loss_curves <- read_csv('throughput/loss_curves.csv')
info_summary <- read_csv('throughput/info_cache.csv')

join_cols <- c('area', 'population', 'density', 'H_X', 'H_Y', 'local.H_Y', 'weighted_J', 'city')
loss_curves <- loss_curves %>% left_join(info_summary[,join_cols])

```

## Example clustering

```{r}

cluster_map <- function(city){
	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
	tracts <- tracts[sum(tracts@data[,columns]) != 0,]
	clusters <- readRDS(paste0('throughput/clusterings/', city))

	tracts@data$cluster <- cutree(clusters,10)

	tracts@data <- tracts@data %>%
		mutate(id = as.character(row_number() - 1))

	to_plot <- fortify(tracts)

	to_plot <- to_plot %>% 
		left_join(tracts@data[,c('id', 'cluster')])

	to_plot %>% 
		ggplot(aes(x = long, y = lat, group = id, fill = cluster)) +
		geom_polygon() + 
		scale_fill_distiller(palette = 'Set3', guide = FALSE) + 
		ggtitle(paste0('Example clustering for ', city)) + 
		theme_minimal() + 
		theme(axis.ticks = element_blank(),
			  axis.text = element_blank(),
			  axis.title = element_blank(),
			  panel.grid = element_blank())

}

cluster_map('Detroit')

ggsave('figs/example_cluster_map.png')

```

```{r}

make_cluster_groups <- function(city){
	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
	tracts <- tracts[sum(tracts@data[,columns]) != 0,]
	clusters <- readRDS(paste0('throughput/clusterings/', city))

	tracts@data$cluster <- cutree(clusters,10)

	I = mutual_info(tracts@data[,c(columns, 'cluster')])

	tracts@data <- tracts@data %>%
		mutate(id = as.character(row_number() - 1))

	tracts@data[,c(columns, 'cluster')] %>% 
		group_by(cluster) %>% 
		summarise_each(funs(sum)) %>%
		gather(key = race, value = n, -cluster) %>% 
		# group_by(cluster) %>% 
		mutate(n = n / sum(n)) %>% 
		ungroup() %>% 
		spread(key = race, value = n, fill = 0) %>% 
		arrange(desc(White)) %>% 
		mutate(cluster = factor(1:10, levels = 1:10)) %>%  
		gather(key = race, value = p, -cluster) %>% 
		mutate(city = city,
			   I = I,
			   title = paste0(city, ": I(C,Y) = ", round(I,2)))

}

df <- rbind(make_cluster_groups('Detroit'), 
	  make_cluster_groups('Boston'),
	  make_cluster_groups('Dallas'),
	  make_cluster_groups('Atlanta'))

df %>% 
	ggplot(aes(y = race, x = cluster, fill = p)) + 
	geom_tile(color = 'black') + 
	geom_text(aes(
		label = ifelse(p < 0.005, "", 100 * round(p,2)))) +
	theme_minimal() + 
	theme(axis.ticks.x = element_blank(),
		  axis.title = element_blank(),
		  legend.position="bottom",
		  legend.title = element_blank()) +
	scale_fill_continuous(low = 'white', high = 'steelblue', limits = c(0, NA), guide = FALSE) +
	ggtitle(paste0('Example clusters')) + 
	facet_wrap(~title, ncol = 2)

ggsave('figs/example_clusters.png')

```

```{r}

```

## Loss curves

```{r}
end_points <- loss_curves %>%
	group_by(city) %>%
	filter(nclust == max(nclust)) %>% 
	filter(weighted_J > 1.25 | weighted_J < .75)

loss_curves %>% 
	ggplot() +
	aes(x = nclust, y = I_XY, group = city) +
	geom_path(aes(color = weighted_J)) +
	geom_point(aes(color = weighted_J), size = .3) +
	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
	# scale_y_continuous(trans = 'log10') +
	geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) +
	theme_minimal() +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	annotation_logticks(size = .1, sides = 'b') +
	ggtitle('Loss curves for spatially-constrained information-theoretic clustering')

ggsave('figs/loss_curves.png', width = 8, height = 4)

```

## Normalized loss curves:

```{r}

loss_curves %>%
	group_by(city) %>%
	mutate(I_XY = I_XY / max(I_XY)) %>%
	mutate(nclust = nclust / max(nclust)) %>%
	# filter(nclust < 100) %>%
	ggplot() +
	aes(x = nclust, y = I_XY, color = weighted_J, group = city) +
	geom_path() +
	scale_x_continuous(trans = 'log10') +
	scale_y_continuous(trans = 'log10') +
	theme_minimal() +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	annotation_logticks(size = .1, sides = 'b') + 
	ggtitle('Normalized loss curves for spatially-constrained information-theoretic clustering')

ggsave('figs/normed_loss_curves.png', width = 8, height = 4)
```

## Evaluating loss curves

### AUC illustration

```{r}

cities <- c('Atlanta','Baltimore')

summary <- loss_curves %>%
	mutate(dx = log(nclust / (nclust - 1))) %>%
	filter(nclust > 1) %>%
	group_by(city) %>%
	arrange(nclust) %>%
	summarise(corner = max(I_XY) * log(max(nclust)),
			  area = trapz(nclust, I_XY / nclust)) %>%
	mutate(AUC = area / corner) %>%
	arrange(desc(AUC)) %>%
	left_join(info_summary[,c('city', 'weighted_J', 'I_XY')]) %>%
	filter(I_XY > 0)

sub_summary <- summary %>% filter(city %in% cities)

sub_curves <- loss_curves %>%
	filter(city %in% cities)

end_points <- sub_curves %>%
	group_by(city) %>%
	filter(nclust == max(nclust))

rects <- sub_curves %>% group_by(city) %>% 
	summarise(xmin = 1, ymin = 0, xmax = max(nclust), ymax = max(I_XY),weighted_J = max(weighted_J))

filled <- sub_curves[,c('nclust','I_XY', 'city')]

corner <- filled %>% group_by(city) %>%
	summarise(nclust = max(nclust),
			  I_XY = 0)

filled <- rbind(corner, filled) %>%
	arrange(city, nclust, desc(I_XY))

sub_curves %>% 
	ggplot() +
	geom_path(aes(x = nclust, y = I_XY, group = city, color = city)) +
	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
	theme_minimal() +
	annotation_logticks(size = .1, sides = 'b') + 
	geom_rect(data = rects, aes(xmin = xmin,
								ymin = ymin,
								xmax = xmax,
								ymax = ymax,
								group = city),
			  color = 'grey',
			  alpha = 0) +
	# geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) + 
	geom_polygon(aes(nclust, I_XY, fill = city), data = filled, alpha = .2) + 
	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
	geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(weighted_J,2))), data = sub_summary) +
	facet_wrap(~city) + 
	scale_fill_discrete(guide = F) + 
	scale_color_discrete(guide = F) + 
	xlab('Number of clusters') + 
	ylab('I(X,Y)') + 
	ggtitle('Evaluation of clustering')

ggsave('figs/AUC_illustration.png', width = 8, height = 4)
```

```{r}

adjusted_summary <- loss_curves %>%
	group_by(city) %>%
	filter(I_XY - lag(I_XY) > 0.005 | row_number() > 10) %>%
	mutate(nclust = row_number()) %>%
	arrange(nclust) %>%
	summarise(adj_corner = max(I_XY) * log(max(nclust)),
			  adj_area = trapz(nclust, I_XY / nclust)) %>%
	mutate(adj_AUC = adj_area / adj_corner) %>%
	arrange(desc(adj_AUC))

summary <- summary %>% 
	left_join(adjusted_summary)
```

## Raw clustering performance

```{r}

summary[,c('I_XY', 'weighted_J', 'AUC', 'adj_AUC')] %>% 
	round(2) %>% 
	cbind(summary$city) %>% 
	select(city = `summary$city`, I_XY, weighted_J, AUC, adj_AUC) %>%
	datatable(options = list(pageLength = 10), rownames = FALSE,
		  caption = 'Table 1: This is a simple caption for the table.')
```

# Relationship to information measures

```{r}

summary %>%
	ggplot() +
	aes(x = I_XY, y = AUC) +
	geom_point(aes(color = weighted_J)) +
	geom_smooth(color = 'grey', se = F) +
	geom_text_repel(aes(label = city)) + 
	theme_minimal() + 
	ggtitle('Clustering performance and mutual information') +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	xlab('I(X,Y)') + 
	ylab('AUC')

ggsave('figs/info_performance.png', width = 8, height = 4)

```

```{r}

summary %>%
	ggplot() +
	aes(x = I_XY, y = adj_AUC) +
	geom_point(aes(color = weighted_J)) +
	geom_smooth(color = 'grey', se = F) +
	geom_text_repel(aes(label = city)) + 
	theme_minimal() + 
	ggtitle('Clustering performance and mutual information (geographically adjusted)') +
	xlab('I(X,Y)') + 
	ylab('AUC (adjusted)') + 
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2))

ggsave('figs/info_performance_adjed.png', width = 8, height = 4)

```

### Basic linear regression

```{r}
summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
		   I_XY = I_XY / max(I_XY),
		   weighted_J = weighted_J / max(weighted_J)) %>%
	filter(I_XY > 0) %>%
	lm(AUC ~ I_XY + weighted_J, data = .) %>%
	summary()
```

### Loess-based detrending

```{r}

normalized <- summary %>% mutate(AUC = AUC / max(AUC),
		   I_XY = I_XY / max(I_XY),
		   weighted_J = weighted_J / max(weighted_J)) %>%
	filter(I_XY > 0)

fit <- normalized %>%
	loess(AUC ~ I_XY + weighted_J, data = .)

normalized$pred <- predict(fit)
normalized$resid <- normalized$pred - normalized$adj_AUC

normalized  %>%
	filter(I_XY > 0) %>%
	lm(resid ~ weighted_J, data = .) %>%
	summary()

```

# Session info
```{r}
sessionInfo()
```

### 2016-06-16 tex/appendix.tex
reorganized

--- 2016-06-02 report/report.tex
@@
-% -----------------------------------------------------------------------------
-% ------------------------------- PREAMBLE ------------------------------------
-% -----------------------------------------------------------------------------[
-	\documentclass[english]{scrartcl}
-
-	\title{Spatial Complexity and Scaling in Cities}
-	\subtitle{v0.0}
-	\author{Phil Chodrow}
-	\date{May 29th, 2016}
-
-	\usepackage[T1]{fontenc}
-	\usepackage[utf8]{inputenc}
-	\usepackage{babel}
-	\usepackage{blindtext}
-	\usepackage{amsmath}
-	\usepackage{amsthm}
-	\usepackage{amsfonts}
-	\usepackage{amssymb}
-	\usepackage{enumitem}
-	\usepackage{graphicx}
-	\usepackage{pgfplotstable,filecontents}
-	\usepackage{booktabs}
-	\usepackage{caption}
-	\usepackage{subcaption}
-	\usepackage{csvsimple}
-
-	\pgfplotsset{compat=1.9}% supress warning
-
-	\setkomafont{disposition}{\normalfont\bfseries}
-
-	\newtheorem{thm}{Theorem}
-	\newtheorem{lm}{Lemma}
-	\newtheorem{cor}{Corrolary}
-	\newtheorem{clm}{Claim}
-	\newtheorem*{thm*}{Theorem}
-	\newtheorem*{lm*}{Lemma}
-	\newtheorem*{cor*}{Corrolary}
-	\newtheorem*{clm*}{Claim}
-	\newtheorem*{fct*}{Fact}
-
-	\newcommand\abs[1]{\left|#1\right|}
-	\newcommand\E[0]{\mathbb{E}}
-	\newcommand\R[0]{\mathbb{R}}
-	\newcommand\decision[2]{\underset{#2}{\overset{#1}{\gtreqless}}}
-	\newcommand{\argmax}{\operatornamewithlimits{argmax}}
-	\newcommand{\argmin}{\operatornamewithlimits{argmin}}
-	\newcommand\prob[0]{\mathbb{P}}
-	\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-% -----------------------------------------------------------------------------
-% --------------------------------- BODY --------------------------------------
-% -----------------------------------------------------------------------------
-\begin{document}
-
-\maketitle
-
-\abstract{We formulate the \emph{mean local information}, a novel measure of spatial compositional complexity in terms of the localized mutual information between spatial and compositional variables.
-This measure is closely related to the Fisher information of the underlying joint distribution, and is therefore an intrinsic property of the phenomena studied.
-Using this measure, we study spatial variations in racial trends in major American cities.
-The results neatly distinguish cities with large, monoracial neighborhoods from those with more intricate, interleaved neighborhood structure.
-Furthermore, the mean local information scales logarithmically with population density, inviting future research into possible universal dynamical processes of neighborhood formation that might explain this relationship.}
-
-\section{Introduction}
-
-	Information theory provides one natural approach toward thinking about complex systems.
-It achieves this by mapping the physical concept of structure to the epistemic concept of predictability.
-A complex system has enough structure to be at least partially predictable, but enough disorder to make prediction challenging.
-In this paper, we apply information-theoretic tools to the measurement of \emph{spatial compositional complexity}, with a focus on the spatial variation of racial trends in American cities.
-
-	Information-theoretic concepts have already found application in urban planning problems related to zoning and predicting population distributions \cite{Royal2014,Batty1974,Batty1976,Battya}.
-Substantial recent work has addressed the measurement of difference and disparity in cities through an information-theoretic lens \cite{Theil1971,Bettencourt2015,Roberto2015a,Roberto2015}.
-Attractive features of information-theoretic measures for this purpose include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to multiple demographic phenomena, and the fact that Theil's (\cite{Theil1971}) original index satisfaction of many (though not all) of the invariance properties desirable for the measurement of segregation \cite{Reardon2002}.
-\nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015}.
-
-	While closely-related to segregation, the concept of complexity has received less attention among quantitative sociologists.
-This is natural, since complexity is both less operational than segregation and challenging to define.
-On the other hand, as we demonstrate below, a spatial information-theoretic measure can be defined which measures the granularity of neighborhood structure, distinguishing cities with large, monolithic tracts of constant racial composition from those with many adjacent, smaller, neighborhoods of varying racial profiles.
-In doing so, it therefore captures one natural aspect of complexity in spatial compositional phenomena, and may also be of interest as a segregation measure as well.
-
-\section{Information Theory and Spatial Structure}
-
-	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two.
-Let $X$ be a spatial variable, such as a set of coordinates in space or a neighborhood name.
-Let $Y$ be a compositional variable that we aim to study, defined on an alphabet $\mathcal{Y}$.
-In the simple case in which $\mathcal{Y}  = \{\text{Black, White}\}$, Figure \ref{fig:toy} illustrates a range of possible joint distributions $p(X,Y)$.
-Comparing the completely undiverse city (a) and the spatially uniformly diverse (b) motivates the first fundamental information measure, the entropy:
-	\begin{equation}
-		H(Y) \triangleq \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
-	\end{equation}
-	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$.
-It thereby clearly distinguishes cities (a) and (b): since (a) is completely uniform, $H_a(Y) = 0$, while $H_b(Y) \neq 0$.
-On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
-To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
-	\begin{align}
-		D[p(Z)\|q(Z)] &\triangleq \sum_z p(z) \log \frac{p(z)}{q(z)} \\
-		I(X,Y) &\triangleq D[p(X,Y) \| p(X)p(Y)]
-	\end{align}
-	The divergence $D$ is (with some caveats) a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$.
-Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent.
-In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).
-In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence.
-
-	City (d) shares with city (c) the fact that residence and race are maximally dependent.
-However, city (d) embodies the ``checkerboard problem'': measures that use only the joint distribution $p(X,Y)$ without additionally considering the spatial information contained within $X$ will evaluate city (d) to be the same as city (c), despite their considerably different patterns of racial separation and potentially very different implications for planning and policy.
-A recent working paper \cite{Roberto2015} provides one highly operational approach to this problem, using road network topology and a weighting function that decays with distance to define a localized measure based on the Kullback-Leibler divergence.
-Here we pursue an alternative strategy, showing that a localization of the mutual information both measures spatial variation in race and corresponds to the estimation of a fundamental statistical property of the distribution $p(X,Y)$.
-
-	% NOTES FOR EXPANDING THE INTRODUCTION
-	% \begin{itemize}
-	% 	\item Concentration: $H(X|Y)$ with appropriate areal weighting. 
-	% 	\item Centralization: Probably not available. 
-	% 	\item Clustering: $J(X,Y)$
-	% 	\item Exposure: $I(X,Y)$? Needs to be read in conjunction with $H(Y)$. 
-	% 	\item Evenness: $H(X|Y)$, low score makes highly uneven. Or $I(X,Y)$? 
-	% \end{itemize}
-
-	% Possibly useful to note that \cite{Press2009a} express some dissatisfaction with their available clustering measures.
-
-	% \cite{Press2009a}
-	% \begin{enumerate}
-	% 	\item Multi-layered aspects of diversity and segregation.
-A limitation of our methods is that we aren't getting into historical or sociological questions; an advantage is that  they generalize fully to other phenomena, such as class, education level, or job type.
-Discuss other measures of diversity.
-	% 	\item Discuss the toy cities, existing information measures, what they mean, and why they need a complementary local measure. 
-	% 	\item Operational roles in modeling urban phenomena
-	% \end{enumerate}
-
-	\begin{figure}
-		\centering
-		  \includegraphics[width=.7\linewidth]{figs/toy.png}
-		  \begin{tabular}{c | c c c}
-			  City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
-			  \hline			
-			  (a) & 0.0 & 0.0 & 0.0\\
-			  (b) & 0.7 & 0.0 & 0.0\\
-			  (c) & 0.7 & 0.7 & 0.6\\
-			  (d) & 0.7 & 0.7 & 2.7\\
-			  \hline  
-			\end{tabular}
-		\caption{Information theory and the checkboard problem: model cities with various kinds of spatial diversity can be distinguished through progressively more subtle spatial information measures.}
-		\label{fig:toy}
-		\end{figure}
-
-\section{Methods}
-	To motivate our methods, we first consider an idealized problem in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is intuitively the map on which we work.
-
-	Fix a point $x_0 \in M$ and a radius $r > 0$.
-Let $B_r(x_0)$ denote the ball of radius $r$ about $x_0$.
-Then, define the \emph{local mutual information in radius $r$} as the mutual information between $X$ and $Y$, restricted to the small ball $B_r(x_0)$ about $x_0$:
-	\begin{equation}
-		I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r(x_0))]|X \in B_r(x_0)]
-	\end{equation}
-	Intuitively, $I_r(x_0)$ measures how much the probability field $p(y|x)$ varies with $x$ in a small neighborhood of $x_0$.
-It is therefore a natural measure of local complexity, aligned in approach with the global mutual information but designed to detect pattern in local variations.
-$I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$, corresponding to a complete lack of local compositional complexity.
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.5\textwidth]{figs/method.png}
-		\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
-We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density.
-}
-		\label{fig:method}
-	\end{figure}
-
-	In this definition, $r$ may be thought of as the spatial resolution at which we conduct analysis, and $I_r(x_0)$ is highly dependent on $r$.
-However, it is possible to show that $I_r(x_0)$ is related to a fundamental statistical property of the probability field $p(y|x)$ that is resolution-independent.
-Under the stated conditions, the following approximation holds:
-
-	\begin{equation}
-		\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace} (J_Y(x))\;, \label{approx}
-	\end{equation}
-	where $J_Y(x)$ is the Fisher information matrix in $Y$ about $x$, defined as 
-	\begin{equation}
-		J_Y(x) \triangleq \E\left[ (\nabla_x \log p(y|x))(\nabla_x \log p(y|x))^T \right]\;.
-	\end{equation}
-	A more formal statement and proof of \eqref{approx} are provided in Appendix 2.
-The Fisher information $J_Y$ is a fundamental quantity in statistics and information theory.
-From a geometric perspective, $J_Y$ provides the natural intrinsic metric in the geometric space of probability distributions parameterized by the spatial variable $x$.
-Equation \eqref{approx} therefore expresses a relationship between the local mutual information $I_r(x)$ and the information geometry of the underlying probability field.
-Corresponding to the fact that $I_r(x_0)$ vanishes if and only if $p(y|x)$ is constant in $B_r(x_0)$, $J_Y(x_0) = 0$ if and only if $\nabla_x p(y|x_0) = 0$ for all $y$.
-This implies that $x_0$ is a stationary point, about which the probability field $p(y|x_0)$ exhibits only small (2nd order or smaller) changes with respect to changes in $x$.
-
-	Since the Fisher information is a strictly local measure of statistical variability around $x$, we can aggregate the Fisher information to derive a measure of average local variability.
-The \emph{mean local information} is
-	\begin{equation}
-	J(X,Y) \triangleq \E_X[\text{trace }J_Y(X)]
-	\end{equation} 
-	We propose the aggregate quantity $J(X,Y)$ as a third  measure--alongside the entropy $H(Y)$ and mutual information $I(X,Y)$-- as a tool for the information-theoretic structure of spatial compositional complexity.
-
-\section{Results}
-	We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S.
-Census Bureau, on race and ethnicity for counties corresponding to 28 large US cities.
-At this early stage, the counties used were chosen to correspond to the most densely populated urban cores; in future developments we could standardize to conduct analyses for Metropolitan Statistical Areas.
-We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'.
-For each city, we computed the entropy $H(Y)$ and the mutual information $I(X,Y)$.
-To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 1km.
-We then computed the estimated mutual information within each grid cell, and averaged the results weighted by population population.
-A technical specification of this approach is provided in Appendix 1, and a complete summary of cities and their associated information measures is provided in Appendix 3.
-
-	\begin{figure}
-		\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
-		\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
-		\label{fig:info_cross}
-	\end{figure}	
-	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$.
-An overall positive trend is evident, reflecting the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.
-On the other hand, there are substantial variations in $J(X,Y)$ even in cities with comparable global variability $I(X,Y)$.
-For example, the cities of Detroit and Philadelphia provide a striking contrast.
-While they have comparable global variability $I(X,Y)$, Philadelphia's $J(X,Y)$ is substantially higher.
-This reflects the fact that Detroit is composed of large, highly-segregated, monoracial neighborhoods, whereas Philadelphia has a much more fine-grained, intricate neighborhood structure.
-These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities.
-
-	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density.
-In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities.
-A clear trend is evident, with $J(X,Y)$ growing linearly with the logarithm of density.
-We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix.
-One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own.
-
-	\begin{figure}
-		\includegraphics[width=1\textwidth]{figs/density_fisher.png}
-		\caption{The mean local information $J(X,Y)$ scales with the logarithm of population density.}
-		\label{fig:density}
-	\end{figure}
-
-\section{Discussion}
-
-	We have formulated a novel measure of urban spatial complexity, and applied it to an analysis of spatial distributions of race in American cities.
-We emphasize that the mean local spatial variability $J(X,Y)$ is novel piece of a more complete information-theoretic characterization that also includes the entropy $H(Y)$ and the global variability $I(X,Y)$.
-Our results suggest two major directions of further exploration:
-
-	One major physical question raised by our results is the origin of the scaling relation seen in \ref{fig:density}.
-This scaling relation may suggest a dynamical process of neighborhood formation and growth common to many American cities.
-If so, a physical understanding of this process in terms of individual-level behavior is called for.
-It is tempting to view the scaling relationship as an interplay of spatial preferential attachment and intrinsic limits to neighborhood density due to housing availability, but this view is speculative at this point.
-A theory to explain this behavior would be most welcome.
-
-	Another question raised is more operational.
-In many urban planning contexts, the amount of available data may make direct computations prohibitive.
-In such contexts, it may be useful to construct a simplifying model of the data the data (such as a clustering) in order to reduce dimensionality and make ``the big picture'' more visible.
-When a particular demographic phenomenon (such as race) is under consideration, it may be important to ensure that the model preserves the major patterns of variation in the granular data.
-We conjecture that the local variability $J(X,Y)$ measures the ``model-ability'' of such data sets.
-Intuitively, a city like Detroit with relatively low $J(X,Y)$ and large, monoracial neighborhoods should be easily representable with relatively few modeled clusters, one for each of the relatively few major neighborhoods.
-On the other hand, relatively speaking, a city like Philadelphia with very intricate neighborhood structure may require substantially greater model complexity to represent without large loss of information.
-A promising course of further study is to design an information-theoretic clustering algorithm designed to model urban patterns, and then compare this algorithm's performance to $J(X,Y)$.
-
-	A third question and final question relates to the time-dependence of spatial structure.
-On one time scale, daily movement around a city for work or leisure activities has the impact of ``mixing'' separated residents in public spaces, potentially leading to very different spatial patterns of racial difference.
-On another time scale, these measures may track how the spatial structure of cities evolves over decades, potentially shedding further light on the dynamics of city formation.
-
-\bibliography{/Users/phil/bibs/library.bib}{}
-\bibliographystyle{plain}
-
-\section*{Appendix 1: Computational Methods and Assumptions}
-	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[J_Y(X)]$ using blockgroup level data from the U.S. Census.
-
-	For fixed Census blockgroup $i$, let $P_i$ be the population, let $A_i$ be the area, let $\rho_i = P_i / A_i$ be the population density, and let $p^i_Y(y)$ be the observed proportion of racial group $y$.
-For hex $k$ in our hexagonal grid, let $N_k$ be the set of overlapping Census blockgroups.
-We also define $p^{k}_I(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$.
-This definition embodies a computationally-simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
-Finally, $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$ is the estimated overall racial composition of hex $k$.
-Then, we estimate the mutual information in hex $k$ as
-	\begin{equation}
-		I(k) = \sum_{i \in N_k} p^k_I(i) D[p^i_Y(\cdot) \| p^k_Y(\cdot)]\;. 
-	\end{equation}
-	Using \eqref{approx}, the estimated Fisher information is 
-	\begin{equation}
-		J(k) \approx \frac{4 I(k)}{r^2}
-	\end{equation}
-	where $r$ is the grid radius. The estimated population in hex $k$ is $P_k = A_k\sum_{i \in N_k} \rho_i$, where $A_k$ is the cell area. We finally estimate $\E_X[J(X)]$ as 
-	\begin{equation}
-		J(X,Y) = \E_X[J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
-	\end{equation}
-\section*{Appendix 2: Relation of Local Mutual Information and Fisher Information}
-	In this appendix, we present a rigorous mathematical justification of the claim that the local mutual information approximates a fundamental quantity intrinsic to the distribution $p(Y,X)$.
-The argument relies on an idealization in which, rather than have a finite, discrete set of observations for each tract, we have a continuous field of observations across the map.
-The precise statement of this result and its justification follow.
-
+\subsection*{Relation of mutual and Fisher informations}
 	Let $X$ be a continuous random variable taking values in $\R^n$, and let $Y$ be a discrete random variable define on finite alphabet $\mathcal{Y}$.
 Suppose further that $p(y|x) > 0$ and that $p(y|x)$ is differentiable as a function of $x$ for all $x,y \in \R^n \times \mathcal{Y}$.
-Fix $x_0 \in R^n$, and define $B_r \triangleq B_r(x_0) = \{ x \in R^n \;|\; \norm{x - x_0} \leq r \}$.
-Additionally, define the \emph{local mutual information} in $B_r$ as
-	\begin{equation}
-		I_r \triangleq I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;.
-	\end{equation}
-	where $D[p\|q] = \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$.
+Fix $x_0 \in \R^n$, and define $B_r \triangleq B_r(x_0) = \{ x \in R^n \;|\; \norm{x - x_0} \leq r \}$.
+Additionally, define the \emph{local mutual information} in $B_r$ as the mutual information between $X$ and $Y$ where $X$ is restricted to $B_r$:
+	\begin{align}
+		I_r(x_0) &\triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r] \\
+		&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;.
+	\end{align}
+	where $D[p\|q] \triangleq \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$.
 
@@
 		\end{equation}
-		where the Fisher information $J_Y(x_0)$ is 
+		where the Fisher information matrix $J_Y$ is given by 
 		\begin{align}
-		J_Y(x) &\triangleq \E\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
+		J_Y(x) &\triangleq \E_Y\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
 		S_y(x) &\triangleq \log p(y|x)\;.
@@
 	\end{thm}
+
 	The proof of Theorem \ref{thm1} proceeds by the application of a number of Taylor approximations, in tandem with a fundamental relationship of information geometry.
-We first expand out $I_r$ explicitly as
+We first expand out $I_r(x_0)$ explicitly as
 	\begin{equation}
-		I_r = \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;. \label{eq:explicit}
+		I_r(x_0) = \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;. \label{eq:explicit}
 	\end{equation}
@@
 			\item $p(Y|X \in B_r) = p(Y|x_0) + e_y$ where the error terms $e_y$ satisfy $e_y \in O(r^{2})$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
-			\item $p(x|X \in B_r) $
+			\item $p(x|X \in B_r) = \frac{1 + O(r)}{v(B_r)}$
 		\end{enumerate}
@@
 				\end{align}
-				where the middle term again vanishes through sphericla symmetry. Thus, for $x \in B_r$, we have
+				where the middle term again vanishes through spherical symmetry. Thus, for $x \in B_r$, we have
 				\begin{align}
@@
 	Dividing through by $r^2$ and computing the limit as $r \rightarrow 0$ proves the result.
-\section*{Appendix 3: Table of Information Measures for Cities Studied}
-	The table below is sorted  by the population density of the city studied. 
-	\begin{tabular}{l c c c c }%
-	    \bfseries City & Density (km$^{-2}$) & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ 
-	    \csvreader[head to column names]{figs/table.csv}{}% use head of csv as column names
-	    {\\\hline\City & \Density & \HY & \IXY & \J}% specify your coloumns here
-    \end{tabular}
 
-\end{document}
+\subsection*{Computational Methods and Assumptions}
+	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[J_Y(X)]$ using blockgroup level data from the U.S. Census.
+
+	For fixed Census blockgroup $i$, let $P_i$ be the population, let $A_i$ be the area, let $\rho_i = P_i / A_i$ be the population density, and let $p^i_Y(y)$ be the observed proportion of racial group $y$.
+For hex $k$ in our hexagonal grid, let $N_k$ be the set of overlapping Census blockgroups.
+We also define $p^{k}_I(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$.
+This definition embodies a computationally-simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
+Finally, $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$ is the estimated overall racial composition of hex $k$.
+Then, we estimate the mutual information in hex $k$ as
+	\begin{equation}
+		I(k) = \sum_{i \in N_k} p^k_I(i) D[p^i_Y(\cdot) \| p^k_Y(\cdot)]\;. 
+	\end{equation}
+	Using \eqref{approx}, the estimated Fisher information is 
+	\begin{equation}
+		J(k) \approx \frac{4 I(k)}{r^2}
+	\end{equation}
+	where $r$ is the grid radius. The estimated population in hex $k$ is $P_k = A_k\sum_{i \in N_k} \rho_i$, where $A_k$ is the cell area. We finally estimate $\E_X[J(X)]$ as 
+	\begin{equation}
+		J(X,Y) = \E_X[J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
+	\end{equation}

### 2016-06-16 tex/discussion.tex
reorganized

--- 2016-06-02 report/report.tex
@@
-% -----------------------------------------------------------------------------
-% ------------------------------- PREAMBLE ------------------------------------
-% -----------------------------------------------------------------------------[
-	\documentclass[english]{scrartcl}
-
-	\title{Spatial Complexity and Scaling in Cities}
-	\subtitle{v0.0}
-	\author{Phil Chodrow}
-	\date{May 29th, 2016}
-
-	\usepackage[T1]{fontenc}
-	\usepackage[utf8]{inputenc}
-	\usepackage{babel}
-	\usepackage{blindtext}
-	\usepackage{amsmath}
-	\usepackage{amsthm}
-	\usepackage{amsfonts}
-	\usepackage{amssymb}
-	\usepackage{enumitem}
-	\usepackage{graphicx}
-	\usepackage{pgfplotstable,filecontents}
-	\usepackage{booktabs}
-	\usepackage{caption}
-	\usepackage{subcaption}
-	\usepackage{csvsimple}
-
-	\pgfplotsset{compat=1.9}% supress warning
-
-	\setkomafont{disposition}{\normalfont\bfseries}
-
-	\newtheorem{thm}{Theorem}
-	\newtheorem{lm}{Lemma}
-	\newtheorem{cor}{Corrolary}
-	\newtheorem{clm}{Claim}
-	\newtheorem*{thm*}{Theorem}
-	\newtheorem*{lm*}{Lemma}
-	\newtheorem*{cor*}{Corrolary}
-	\newtheorem*{clm*}{Claim}
-	\newtheorem*{fct*}{Fact}
-
-	\newcommand\abs[1]{\left|#1\right|}
-	\newcommand\E[0]{\mathbb{E}}
-	\newcommand\R[0]{\mathbb{R}}
-	\newcommand\decision[2]{\underset{#2}{\overset{#1}{\gtreqless}}}
-	\newcommand{\argmax}{\operatornamewithlimits{argmax}}
-	\newcommand{\argmin}{\operatornamewithlimits{argmin}}
-	\newcommand\prob[0]{\mathbb{P}}
-	\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-% -----------------------------------------------------------------------------
-% --------------------------------- BODY --------------------------------------
-% -----------------------------------------------------------------------------
-\begin{document}
-
-\maketitle
-
-\abstract{We formulate the \emph{mean local information}, a novel measure of spatial compositional complexity in terms of the localized mutual information between spatial and compositional variables.
-This measure is closely related to the Fisher information of the underlying joint distribution, and is therefore an intrinsic property of the phenomena studied.
-Using this measure, we study spatial variations in racial trends in major American cities.
-The results neatly distinguish cities with large, monoracial neighborhoods from those with more intricate, interleaved neighborhood structure.
-Furthermore, the mean local information scales logarithmically with population density, inviting future research into possible universal dynamical processes of neighborhood formation that might explain this relationship.}
-
-\section{Introduction}
-
-	Information theory provides one natural approach toward thinking about complex systems.
-It achieves this by mapping the physical concept of structure to the epistemic concept of predictability.
-A complex system has enough structure to be at least partially predictable, but enough disorder to make prediction challenging.
-In this paper, we apply information-theoretic tools to the measurement of \emph{spatial compositional complexity}, with a focus on the spatial variation of racial trends in American cities.
-
-	Information-theoretic concepts have already found application in urban planning problems related to zoning and predicting population distributions \cite{Royal2014,Batty1974,Batty1976,Battya}.
-Substantial recent work has addressed the measurement of difference and disparity in cities through an information-theoretic lens \cite{Theil1971,Bettencourt2015,Roberto2015a,Roberto2015}.
-Attractive features of information-theoretic measures for this purpose include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to multiple demographic phenomena, and the fact that Theil's (\cite{Theil1971}) original index satisfaction of many (though not all) of the invariance properties desirable for the measurement of segregation \cite{Reardon2002}.
-\nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015}.
-
-	While closely-related to segregation, the concept of complexity has received less attention among quantitative sociologists.
-This is natural, since complexity is both less operational than segregation and challenging to define.
-On the other hand, as we demonstrate below, a spatial information-theoretic measure can be defined which measures the granularity of neighborhood structure, distinguishing cities with large, monolithic tracts of constant racial composition from those with many adjacent, smaller, neighborhoods of varying racial profiles.
-In doing so, it therefore captures one natural aspect of complexity in spatial compositional phenomena, and may also be of interest as a segregation measure as well.
-
-\section{Information Theory and Spatial Structure}
-
-	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two.
-Let $X$ be a spatial variable, such as a set of coordinates in space or a neighborhood name.
-Let $Y$ be a compositional variable that we aim to study, defined on an alphabet $\mathcal{Y}$.
-In the simple case in which $\mathcal{Y}  = \{\text{Black, White}\}$, Figure \ref{fig:toy} illustrates a range of possible joint distributions $p(X,Y)$.
-Comparing the completely undiverse city (a) and the spatially uniformly diverse (b) motivates the first fundamental information measure, the entropy:
-	\begin{equation}
-		H(Y) \triangleq \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
-	\end{equation}
-	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$.
-It thereby clearly distinguishes cities (a) and (b): since (a) is completely uniform, $H_a(Y) = 0$, while $H_b(Y) \neq 0$.
-On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
-To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
-	\begin{align}
-		D[p(Z)\|q(Z)] &\triangleq \sum_z p(z) \log \frac{p(z)}{q(z)} \\
-		I(X,Y) &\triangleq D[p(X,Y) \| p(X)p(Y)]
-	\end{align}
-	The divergence $D$ is (with some caveats) a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$.
-Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent.
-In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).
-In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence.
-
-	City (d) shares with city (c) the fact that residence and race are maximally dependent.
-However, city (d) embodies the ``checkerboard problem'': measures that use only the joint distribution $p(X,Y)$ without additionally considering the spatial information contained within $X$ will evaluate city (d) to be the same as city (c), despite their considerably different patterns of racial separation and potentially very different implications for planning and policy.
-A recent working paper \cite{Roberto2015} provides one highly operational approach to this problem, using road network topology and a weighting function that decays with distance to define a localized measure based on the Kullback-Leibler divergence.
-Here we pursue an alternative strategy, showing that a localization of the mutual information both measures spatial variation in race and corresponds to the estimation of a fundamental statistical property of the distribution $p(X,Y)$.
-
-	% NOTES FOR EXPANDING THE INTRODUCTION
-	% \begin{itemize}
-	% 	\item Concentration: $H(X|Y)$ with appropriate areal weighting. 
-	% 	\item Centralization: Probably not available. 
-	% 	\item Clustering: $J(X,Y)$
-	% 	\item Exposure: $I(X,Y)$? Needs to be read in conjunction with $H(Y)$. 
-	% 	\item Evenness: $H(X|Y)$, low score makes highly uneven. Or $I(X,Y)$? 
-	% \end{itemize}
-
-	% Possibly useful to note that \cite{Press2009a} express some dissatisfaction with their available clustering measures.
-
-	% \cite{Press2009a}
-	% \begin{enumerate}
-	% 	\item Multi-layered aspects of diversity and segregation.
-A limitation of our methods is that we aren't getting into historical or sociological questions; an advantage is that  they generalize fully to other phenomena, such as class, education level, or job type.
-Discuss other measures of diversity.
-	% 	\item Discuss the toy cities, existing information measures, what they mean, and why they need a complementary local measure. 
-	% 	\item Operational roles in modeling urban phenomena
-	% \end{enumerate}
-
-	\begin{figure}
-		\centering
-		  \includegraphics[width=.7\linewidth]{figs/toy.png}
-		  \begin{tabular}{c | c c c}
-			  City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
-			  \hline			
-			  (a) & 0.0 & 0.0 & 0.0\\
-			  (b) & 0.7 & 0.0 & 0.0\\
-			  (c) & 0.7 & 0.7 & 0.6\\
-			  (d) & 0.7 & 0.7 & 2.7\\
-			  \hline  
-			\end{tabular}
-		\caption{Information theory and the checkboard problem: model cities with various kinds of spatial diversity can be distinguished through progressively more subtle spatial information measures.}
-		\label{fig:toy}
-		\end{figure}
-
-\section{Methods}
-	To motivate our methods, we first consider an idealized problem in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is intuitively the map on which we work.
-
-	Fix a point $x_0 \in M$ and a radius $r > 0$.
-Let $B_r(x_0)$ denote the ball of radius $r$ about $x_0$.
-Then, define the \emph{local mutual information in radius $r$} as the mutual information between $X$ and $Y$, restricted to the small ball $B_r(x_0)$ about $x_0$:
-	\begin{equation}
-		I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r(x_0))]|X \in B_r(x_0)]
-	\end{equation}
-	Intuitively, $I_r(x_0)$ measures how much the probability field $p(y|x)$ varies with $x$ in a small neighborhood of $x_0$.
-It is therefore a natural measure of local complexity, aligned in approach with the global mutual information but designed to detect pattern in local variations.
-$I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$, corresponding to a complete lack of local compositional complexity.
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.5\textwidth]{figs/method.png}
-		\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
-We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density.
-}
-		\label{fig:method}
-	\end{figure}
-
-	In this definition, $r$ may be thought of as the spatial resolution at which we conduct analysis, and $I_r(x_0)$ is highly dependent on $r$.
-However, it is possible to show that $I_r(x_0)$ is related to a fundamental statistical property of the probability field $p(y|x)$ that is resolution-independent.
-Under the stated conditions, the following approximation holds:
-
-	\begin{equation}
-		\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace} (J_Y(x))\;, \label{approx}
-	\end{equation}
-	where $J_Y(x)$ is the Fisher information matrix in $Y$ about $x$, defined as 
-	\begin{equation}
-		J_Y(x) \triangleq \E\left[ (\nabla_x \log p(y|x))(\nabla_x \log p(y|x))^T \right]\;.
-	\end{equation}
-	A more formal statement and proof of \eqref{approx} are provided in Appendix 2.
-The Fisher information $J_Y$ is a fundamental quantity in statistics and information theory.
-From a geometric perspective, $J_Y$ provides the natural intrinsic metric in the geometric space of probability distributions parameterized by the spatial variable $x$.
-Equation \eqref{approx} therefore expresses a relationship between the local mutual information $I_r(x)$ and the information geometry of the underlying probability field.
-Corresponding to the fact that $I_r(x_0)$ vanishes if and only if $p(y|x)$ is constant in $B_r(x_0)$, $J_Y(x_0) = 0$ if and only if $\nabla_x p(y|x_0) = 0$ for all $y$.
-This implies that $x_0$ is a stationary point, about which the probability field $p(y|x_0)$ exhibits only small (2nd order or smaller) changes with respect to changes in $x$.
-
-	Since the Fisher information is a strictly local measure of statistical variability around $x$, we can aggregate the Fisher information to derive a measure of average local variability.
-The \emph{mean local information} is
-	\begin{equation}
-	J(X,Y) \triangleq \E_X[\text{trace }J_Y(X)]
-	\end{equation} 
-	We propose the aggregate quantity $J(X,Y)$ as a third  measure--alongside the entropy $H(Y)$ and mutual information $I(X,Y)$-- as a tool for the information-theoretic structure of spatial compositional complexity.
-
-\section{Results}
-	We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S.
-Census Bureau, on race and ethnicity for counties corresponding to 28 large US cities.
-At this early stage, the counties used were chosen to correspond to the most densely populated urban cores; in future developments we could standardize to conduct analyses for Metropolitan Statistical Areas.
-We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'.
-For each city, we computed the entropy $H(Y)$ and the mutual information $I(X,Y)$.
-To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 1km.
-We then computed the estimated mutual information within each grid cell, and averaged the results weighted by population population.
-A technical specification of this approach is provided in Appendix 1, and a complete summary of cities and their associated information measures is provided in Appendix 3.
-
-	\begin{figure}
-		\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
-		\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
-		\label{fig:info_cross}
-	\end{figure}	
-	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$.
-An overall positive trend is evident, reflecting the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.
-On the other hand, there are substantial variations in $J(X,Y)$ even in cities with comparable global variability $I(X,Y)$.
-For example, the cities of Detroit and Philadelphia provide a striking contrast.
-While they have comparable global variability $I(X,Y)$, Philadelphia's $J(X,Y)$ is substantially higher.
-This reflects the fact that Detroit is composed of large, highly-segregated, monoracial neighborhoods, whereas Philadelphia has a much more fine-grained, intricate neighborhood structure.
-These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities.
-
-	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density.
-In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities.
-A clear trend is evident, with $J(X,Y)$ growing linearly with the logarithm of density.
-We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix.
-One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own.
-
-	\begin{figure}
-		\includegraphics[width=1\textwidth]{figs/density_fisher.png}
-		\caption{The mean local information $J(X,Y)$ scales with the logarithm of population density.}
-		\label{fig:density}
-	\end{figure}
-
-\section{Discussion}
-
 We have formulated a novel measure of urban spatial complexity, and applied it to an analysis of spatial distributions of race in American cities.
@@
 
-\bibliography{/Users/phil/bibs/library.bib}{}
-\bibliographystyle{plain}
-
-\section*{Appendix 1: Computational Methods and Assumptions}
-	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[J_Y(X)]$ using blockgroup level data from the U.S. Census.
-
-	For fixed Census blockgroup $i$, let $P_i$ be the population, let $A_i$ be the area, let $\rho_i = P_i / A_i$ be the population density, and let $p^i_Y(y)$ be the observed proportion of racial group $y$.
-For hex $k$ in our hexagonal grid, let $N_k$ be the set of overlapping Census blockgroups.
-We also define $p^{k}_I(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$.
-This definition embodies a computationally-simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
-Finally, $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$ is the estimated overall racial composition of hex $k$.
-Then, we estimate the mutual information in hex $k$ as
-	\begin{equation}
-		I(k) = \sum_{i \in N_k} p^k_I(i) D[p^i_Y(\cdot) \| p^k_Y(\cdot)]\;. 
-	\end{equation}
-	Using \eqref{approx}, the estimated Fisher information is 
-	\begin{equation}
-		J(k) \approx \frac{4 I(k)}{r^2}
-	\end{equation}
-	where $r$ is the grid radius. The estimated population in hex $k$ is $P_k = A_k\sum_{i \in N_k} \rho_i$, where $A_k$ is the cell area. We finally estimate $\E_X[J(X)]$ as 
-	\begin{equation}
-		J(X,Y) = \E_X[J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
-	\end{equation}
-\section*{Appendix 2: Relation of Local Mutual Information and Fisher Information}
-	In this appendix, we present a rigorous mathematical justification of the claim that the local mutual information approximates a fundamental quantity intrinsic to the distribution $p(Y,X)$.
-The argument relies on an idealization in which, rather than have a finite, discrete set of observations for each tract, we have a continuous field of observations across the map.
-The precise statement of this result and its justification follow.
-
-	Let $X$ be a continuous random variable taking values in $\R^n$, and let $Y$ be a discrete random variable define on finite alphabet $\mathcal{Y}$.
-Suppose further that $p(y|x) > 0$ and that $p(y|x)$ is differentiable as a function of $x$ for all $x,y \in \R^n \times \mathcal{Y}$.
-Fix $x_0 \in R^n$, and define $B_r \triangleq B_r(x_0) = \{ x \in R^n \;|\; \norm{x - x_0} \leq r \}$.
-Additionally, define the \emph{local mutual information} in $B_r$ as
-	\begin{equation}
-		I_r \triangleq I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;.
-	\end{equation}
-	where $D[p\|q] = \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$.
-
-	\begin{thm} \label{thm1}
-		Under the stated conditions, 
-		\begin{equation}
-			\lim_{r\rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{n}{2(n+2)} \emph{trace}\; J_Y(x_0)\;.
-		\end{equation}
-		where the Fisher information $J_Y(x_0)$ is 
-		\begin{align}
-		J_Y(x) &\triangleq \E\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
-		S_y(x) &\triangleq \log p(y|x)\;.
-	\end{align}
-	\end{thm}
-	The proof of Theorem \ref{thm1} proceeds by the application of a number of Taylor approximations, in tandem with a fundamental relationship of information geometry.
-We first expand out $I_r$ explicitly as
-	\begin{equation}
-		I_r = \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;. \label{eq:explicit}
-	\end{equation}
-
-	\begin{lm} \label{approximations}
-		The following approximation relationships hold for the components of \eqref{eq:explicit}:
-		\begin{enumerate}[label=\emph{(\alph*)}]
-			\item $p(X\in B_r,Y) = p(x_0, Y) v(B_r) + O(r^{n+2})$
-			\item $p(Y|X \in B_r) = p(Y|x_0) + e_y$ where the error terms $e_y$ satisfy $e_y \in O(r^{2})$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
-			\item $p(x|X \in B_r) $
-		\end{enumerate}
-	\end{lm}
-	\begin{proof}
-		For each approximation,  we directly apply Taylor expansions about $X = x_0$. 
-		\begin{enumerate}[label=(\alph*)]
-			\item We have 
-				\begin{align}
-					p(X\in B_r, Y) &= \int_{B_r} p(x,Y) \; d^nx \\
-					&= \int_{B_r} p(x_0,Y) + \frac{\partial p(x_0, Y)}{\partial x}(x - x_0) + O(\norm{x - x_0}^2) \; d^nx \\
-					&= p(x_0, Y)v(B_r) +  \frac{\partial p(x_0, Y)}{\partial x} \int_{B_r} (x - x_0)\; d^nx  \\
-					&\quad \quad+ O\left(\int_{B_r} \norm{x - x_0}^2\; d^nx\right) \\
-					&= p(x_0, Y)v(B_r) + O(r^{n+2})\;,
-				\end{align}
-				where the middle term vanishes due to spherical symmetry. 
-			\item The fact that the error terms $e_y$ must satisfy $\sum_{y \in \mathcal{Y}} e_y = 0$ follows from the fact that $p(Y|X \in B_r)$ must be a valid probability distribution over $\mathcal{Y}$.
-We'll now show that $e_y \in O(r^{ß2})$.
-First,
-				\begin{align}
-					p(X \in B_r) &= \sum_{y\in \mathcal{Y}} p(X \in B_r, y)  \\
-					&= \sum_{y\in \mathcal{Y}} \left[p(x_0, y)v(B_r) + O(r^{n+2}) \right] \\
-					&= p(x_0)v(B_r) + O(r^{2});\,
-				\end{align}
-			from part (a). Next, 
-				\begin{align}
-					p(Y|X \in B_r) &= \frac{p(X\in B_r, Y)}{p(X \in B_r)} \\
-					&= \frac{p(x_0, Y)v(B_r) + O(r^{n+2})}{p(x_0)v(B_r) + O(r^{n+2})} \\
-					&= p(Y|x_0) + O(r^2)\;,
-				\end{align}
-			which completes this part of the argument. 
-			\item First, 
-				\begin{align}
-					p(X \in B_r) &= \int_{B_r} p(x) \; d^nx  \\
-					&= \int_{B_r} \left[p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)\right] \; d^nx \\ 
-					&= p(x_0) v(B_r) + O(r^{n+2})\;,
-				\end{align}
-				where the middle term again vanishes through sphericla symmetry. Thus, for $x \in B_r$, we have
-				\begin{align}
-					p(x | X \in B_r) &= \frac{p(x)}{p(X \in B_r)} \\
-					&= \frac{p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)}{p(x_0) v(B_r) + O(r^{n+2})} \\
-					&= \frac{1 + O(r)}{v(B_r)}\;.
-				\end{align}
-		\end{enumerate}
-	\end{proof}
-
-	\begin{lm} The following approximation holds for the divergence factor in the integral \eqref{eq:explicit}
-		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] = D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})
-		\end{equation}
-	\end{lm}
-	\begin{proof}
-		We compute directly: 
-		\begin{align}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] &= \sum_{y \in \mathcal{Y}} p(y|x) \log \frac{p(y|x)}{p(y|X \in B_r)} \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|X \in B_r) \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log \left(p(y|x_0) + e_y)\right) \tag{from Lemma \ref{approximations}}\\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \left[\log p(y|x_0) + \frac{e_y}{p(y|x_0)} + O(e_y^2)\right] \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}  e_y \tag{quadratic terms negligible} \\
-			&=D[p(\cdot|x)\| p(\cdot|x_0)] \\
-			&\quad + \sum_{y \in \mathcal{Y}} \left(1 + \frac{1}{p(y|x_0)} \nabla p(y|x_0)(x - x_0) + O(r^{2})\right)    e_y \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \left[e_y + O(r^{3})\right] \tag{$e_y \in O(r^2)$}\\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^3) \tag{$\sum_{y \in \mathcal{Y}} e_y = 0$}
-		\end{align}
-	\end{proof}
-
-	\begin{lm} For any positive-semidefinite matrix $A \in \R^{n\times n}$, 
-		$$\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x = \frac{n}{n+2} r^{2} v(B_r) \text{\emph{trace}} (A)$$
-	\end{lm}
-
-	\begin{proof}
-		Since $A$ is positive-semidefinite, there exist an orthonormal matrix $P$ and a diagonal matrix $D$ such that $A = P^TDP$.
-Furthermore, the entries of $D$ are the eigenvalues $\{\lambda_i\}$ of $A$.
-Then,
-		\begin{align}
-			\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0, P^T DP(x-x_0) \right> d^n x\\
-			&= \int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x \;.
-		\end{align}
-		We can regard $P$ as a reparameterization of $B_r$; since $\det P = 1$, we have 
-		\begin{align}
-			\int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0,  D(x-x_0) \right> d^n x \\
-			&= r^n\int_{B_n} \left< rx, rDx \right> d^n x \\
-			&= r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x \;,
-		\end{align}
-		where $B_n$ is the unit $n$-ball. We also let $S_n(r)$ be the $n$-sphere of radius $r$. Continuing, 
-		\begin{align}
-			r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x &= r^{n+2} \int_{B_n} \sum_{i= 1}^n x_i^2 \lambda_i \; d^nx \\
-			&= r^{n+2} \sum_{i = 1}^n \lambda _i \int_{B_n} x_i^2 \; d^nx \\
-			&= \frac{r^{n+2}}{n} \sum_{i = 1}^n \lambda _i \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace}(A)  \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^2 S_{n-1}(\rho) d\rho \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^{n+1} S_{n-1}(1) d\rho \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A)  \frac{1}{n+2} S_{n-1}(1) \\
-			&= \frac{r^2}{n+2}  \text{trace} (A)  n r^{n}v(B_n(1)) \\
-			&= \frac{n}{n+2} r^{2} v(B_r) \text{trace} (A)\;,
-		\end{align}
-		as was to be shown. 
-	\end{proof}
-
-	\begin{fct*}
-		The Kullback-Leibler divergence and the Fisher information $J_Y$ are related according to the approximation 
-		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|x_0)] = \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3)
-		\end{equation}
-	\end{fct*}
-
-	We are finally ready to prove Theorem \ref{thm1}. Computing directly, we have 
-	\begin{align}
-		I_r(x_0) &\triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;. \\
-		&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \\
-		&= \int_{B_r} \left[\frac{1 + O(r)}{v(B_r)}\right] D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \tag{Lemma 1(c)}\\ 
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r}  \left(D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})\right) d^n x \tag{Lemma 2}\\
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left( \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3) + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left(\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\left(\frac{n}{n+2} r^{2} v(B_r) \text{trace} (J_Y(x_0)) + v(B_r)O(r^{3})\right) \\
-		&= r^2\frac{n}{2(n+2)}\left[1 + O(r)\right]\left(\text{trace} (J_Y(x_0)) + O(r^{3})\right)\;.
-	\end{align}
-	Dividing through by $r^2$ and computing the limit as $r \rightarrow 0$ proves the result. 
-\section*{Appendix 3: Table of Information Measures for Cities Studied}
-	The table below is sorted  by the population density of the city studied. 
-	\begin{tabular}{l c c c c }%
-	    \bfseries City & Density (km$^{-2}$) & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ 
-	    \csvreader[head to column names]{figs/table.csv}{}% use head of csv as column names
-	    {\\\hline\City & \Density & \HY & \IXY & \J}% specify your coloumns here
-    \end{tabular}
-
-\end{document}
+A final question is....

### 2016-06-16 tex/intro.tex
reorganized

--- 2016-06-02 report/report.tex
@@
-% -----------------------------------------------------------------------------
-% ------------------------------- PREAMBLE ------------------------------------
-% -----------------------------------------------------------------------------[
-	\documentclass[english]{scrartcl}
-
-	\title{Spatial Complexity and Scaling in Cities}
-	\subtitle{v0.0}
-	\author{Phil Chodrow}
-	\date{May 29th, 2016}
-
-	\usepackage[T1]{fontenc}
-	\usepackage[utf8]{inputenc}
-	\usepackage{babel}
-	\usepackage{blindtext}
-	\usepackage{amsmath}
-	\usepackage{amsthm}
-	\usepackage{amsfonts}
-	\usepackage{amssymb}
-	\usepackage{enumitem}
-	\usepackage{graphicx}
-	\usepackage{pgfplotstable,filecontents}
-	\usepackage{booktabs}
-	\usepackage{caption}
-	\usepackage{subcaption}
-	\usepackage{csvsimple}
-
-	\pgfplotsset{compat=1.9}% supress warning
-
-	\setkomafont{disposition}{\normalfont\bfseries}
-
-	\newtheorem{thm}{Theorem}
-	\newtheorem{lm}{Lemma}
-	\newtheorem{cor}{Corrolary}
-	\newtheorem{clm}{Claim}
-	\newtheorem*{thm*}{Theorem}
-	\newtheorem*{lm*}{Lemma}
-	\newtheorem*{cor*}{Corrolary}
-	\newtheorem*{clm*}{Claim}
-	\newtheorem*{fct*}{Fact}
-
-	\newcommand\abs[1]{\left|#1\right|}
-	\newcommand\E[0]{\mathbb{E}}
-	\newcommand\R[0]{\mathbb{R}}
-	\newcommand\decision[2]{\underset{#2}{\overset{#1}{\gtreqless}}}
-	\newcommand{\argmax}{\operatornamewithlimits{argmax}}
-	\newcommand{\argmin}{\operatornamewithlimits{argmin}}
-	\newcommand\prob[0]{\mathbb{P}}
-	\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-% -----------------------------------------------------------------------------
-% --------------------------------- BODY --------------------------------------
-% -----------------------------------------------------------------------------
-\begin{document}
-
-\maketitle
-
-\abstract{We formulate the \emph{mean local information}, a novel measure of spatial compositional complexity in terms of the localized mutual information between spatial and compositional variables.
-This measure is closely related to the Fisher information of the underlying joint distribution, and is therefore an intrinsic property of the phenomena studied.
-Using this measure, we study spatial variations in racial trends in major American cities.
-The results neatly distinguish cities with large, monoracial neighborhoods from those with more intricate, interleaved neighborhood structure.
-Furthermore, the mean local information scales logarithmically with population density, inviting future research into possible universal dynamical processes of neighborhood formation that might explain this relationship.}
-
-\section{Introduction}
-
 	Information theory provides one natural approach toward thinking about complex systems.
@@
 
-\section{Information Theory and Spatial Structure}
-
+\subsection{Information Theory and Spatial Structure}
 	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two.
@@
 
-	% NOTES FOR EXPANDING THE INTRODUCTION
-	% \begin{itemize}
-	% 	\item Concentration: $H(X|Y)$ with appropriate areal weighting. 
-	% 	\item Centralization: Probably not available. 
-	% 	\item Clustering: $J(X,Y)$
-	% 	\item Exposure: $I(X,Y)$? Needs to be read in conjunction with $H(Y)$. 
-	% 	\item Evenness: $H(X|Y)$, low score makes highly uneven. Or $I(X,Y)$? 
-	% \end{itemize}
-
-	% Possibly useful to note that \cite{Press2009a} express some dissatisfaction with their available clustering measures.
-
-	% \cite{Press2009a}
-	% \begin{enumerate}
-	% 	\item Multi-layered aspects of diversity and segregation.
-A limitation of our methods is that we aren't getting into historical or sociological questions; an advantage is that  they generalize fully to other phenomena, such as class, education level, or job type.
-Discuss other measures of diversity.
-	% 	\item Discuss the toy cities, existing information measures, what they mean, and why they need a complementary local measure. 
-	% 	\item Operational roles in modeling urban phenomena
-	% \end{enumerate}
-
 	\begin{figure}
 		\centering
-		  \includegraphics[width=.7\linewidth]{figs/toy.png}
+		  \includegraphics[width=.7\linewidth]{figs/checkerboard.png}
 		  \begin{tabular}{c | c c c}
@@
 		\end{figure}
-
-\section{Methods}
-	To motivate our methods, we first consider an idealized problem in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is intuitively the map on which we work.
-
-	Fix a point $x_0 \in M$ and a radius $r > 0$.
-Let $B_r(x_0)$ denote the ball of radius $r$ about $x_0$.
-Then, define the \emph{local mutual information in radius $r$} as the mutual information between $X$ and $Y$, restricted to the small ball $B_r(x_0)$ about $x_0$:
-	\begin{equation}
-		I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r(x_0))]|X \in B_r(x_0)]
-	\end{equation}
-	Intuitively, $I_r(x_0)$ measures how much the probability field $p(y|x)$ varies with $x$ in a small neighborhood of $x_0$.
-It is therefore a natural measure of local complexity, aligned in approach with the global mutual information but designed to detect pattern in local variations.
-$I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$, corresponding to a complete lack of local compositional complexity.
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.5\textwidth]{figs/method.png}
-		\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
-We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density.
-}
-		\label{fig:method}
-	\end{figure}
-
-	In this definition, $r$ may be thought of as the spatial resolution at which we conduct analysis, and $I_r(x_0)$ is highly dependent on $r$.
-However, it is possible to show that $I_r(x_0)$ is related to a fundamental statistical property of the probability field $p(y|x)$ that is resolution-independent.
-Under the stated conditions, the following approximation holds:
-
-	\begin{equation}
-		\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace} (J_Y(x))\;, \label{approx}
-	\end{equation}
-	where $J_Y(x)$ is the Fisher information matrix in $Y$ about $x$, defined as 
-	\begin{equation}
-		J_Y(x) \triangleq \E\left[ (\nabla_x \log p(y|x))(\nabla_x \log p(y|x))^T \right]\;.
-	\end{equation}
-	A more formal statement and proof of \eqref{approx} are provided in Appendix 2.
-The Fisher information $J_Y$ is a fundamental quantity in statistics and information theory.
-From a geometric perspective, $J_Y$ provides the natural intrinsic metric in the geometric space of probability distributions parameterized by the spatial variable $x$.
-Equation \eqref{approx} therefore expresses a relationship between the local mutual information $I_r(x)$ and the information geometry of the underlying probability field.
-Corresponding to the fact that $I_r(x_0)$ vanishes if and only if $p(y|x)$ is constant in $B_r(x_0)$, $J_Y(x_0) = 0$ if and only if $\nabla_x p(y|x_0) = 0$ for all $y$.
-This implies that $x_0$ is a stationary point, about which the probability field $p(y|x_0)$ exhibits only small (2nd order or smaller) changes with respect to changes in $x$.
-
-	Since the Fisher information is a strictly local measure of statistical variability around $x$, we can aggregate the Fisher information to derive a measure of average local variability.
-The \emph{mean local information} is
-	\begin{equation}
-	J(X,Y) \triangleq \E_X[\text{trace }J_Y(X)]
-	\end{equation} 
-	We propose the aggregate quantity $J(X,Y)$ as a third  measure--alongside the entropy $H(Y)$ and mutual information $I(X,Y)$-- as a tool for the information-theoretic structure of spatial compositional complexity.
-
-\section{Results}
-	We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S.
-Census Bureau, on race and ethnicity for counties corresponding to 28 large US cities.
-At this early stage, the counties used were chosen to correspond to the most densely populated urban cores; in future developments we could standardize to conduct analyses for Metropolitan Statistical Areas.
-We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'.
-For each city, we computed the entropy $H(Y)$ and the mutual information $I(X,Y)$.
-To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 1km.
-We then computed the estimated mutual information within each grid cell, and averaged the results weighted by population population.
-A technical specification of this approach is provided in Appendix 1, and a complete summary of cities and their associated information measures is provided in Appendix 3.
-
-	\begin{figure}
-		\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
-		\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
-		\label{fig:info_cross}
-	\end{figure}	
-	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$.
-An overall positive trend is evident, reflecting the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.
-On the other hand, there are substantial variations in $J(X,Y)$ even in cities with comparable global variability $I(X,Y)$.
-For example, the cities of Detroit and Philadelphia provide a striking contrast.
-While they have comparable global variability $I(X,Y)$, Philadelphia's $J(X,Y)$ is substantially higher.
-This reflects the fact that Detroit is composed of large, highly-segregated, monoracial neighborhoods, whereas Philadelphia has a much more fine-grained, intricate neighborhood structure.
-These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities.
-
-	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density.
-In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities.
-A clear trend is evident, with $J(X,Y)$ growing linearly with the logarithm of density.
-We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix.
-One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own.
-
-	\begin{figure}
-		\includegraphics[width=1\textwidth]{figs/density_fisher.png}
-		\caption{The mean local information $J(X,Y)$ scales with the logarithm of population density.}
-		\label{fig:density}
-	\end{figure}
-
-\section{Discussion}
-
-	We have formulated a novel measure of urban spatial complexity, and applied it to an analysis of spatial distributions of race in American cities.
-We emphasize that the mean local spatial variability $J(X,Y)$ is novel piece of a more complete information-theoretic characterization that also includes the entropy $H(Y)$ and the global variability $I(X,Y)$.
-Our results suggest two major directions of further exploration:
-
-	One major physical question raised by our results is the origin of the scaling relation seen in \ref{fig:density}.
-This scaling relation may suggest a dynamical process of neighborhood formation and growth common to many American cities.
-If so, a physical understanding of this process in terms of individual-level behavior is called for.
-It is tempting to view the scaling relationship as an interplay of spatial preferential attachment and intrinsic limits to neighborhood density due to housing availability, but this view is speculative at this point.
-A theory to explain this behavior would be most welcome.
-
-	Another question raised is more operational.
-In many urban planning contexts, the amount of available data may make direct computations prohibitive.
-In such contexts, it may be useful to construct a simplifying model of the data the data (such as a clustering) in order to reduce dimensionality and make ``the big picture'' more visible.
-When a particular demographic phenomenon (such as race) is under consideration, it may be important to ensure that the model preserves the major patterns of variation in the granular data.
-We conjecture that the local variability $J(X,Y)$ measures the ``model-ability'' of such data sets.
-Intuitively, a city like Detroit with relatively low $J(X,Y)$ and large, monoracial neighborhoods should be easily representable with relatively few modeled clusters, one for each of the relatively few major neighborhoods.
-On the other hand, relatively speaking, a city like Philadelphia with very intricate neighborhood structure may require substantially greater model complexity to represent without large loss of information.
-A promising course of further study is to design an information-theoretic clustering algorithm designed to model urban patterns, and then compare this algorithm's performance to $J(X,Y)$.
-
-	A third question and final question relates to the time-dependence of spatial structure.
-On one time scale, daily movement around a city for work or leisure activities has the impact of ``mixing'' separated residents in public spaces, potentially leading to very different spatial patterns of racial difference.
-On another time scale, these measures may track how the spatial structure of cities evolves over decades, potentially shedding further light on the dynamics of city formation.
-
-\bibliography{/Users/phil/bibs/library.bib}{}
-\bibliographystyle{plain}
-
-\section*{Appendix 1: Computational Methods and Assumptions}
-	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[J_Y(X)]$ using blockgroup level data from the U.S. Census.
-
-	For fixed Census blockgroup $i$, let $P_i$ be the population, let $A_i$ be the area, let $\rho_i = P_i / A_i$ be the population density, and let $p^i_Y(y)$ be the observed proportion of racial group $y$.
-For hex $k$ in our hexagonal grid, let $N_k$ be the set of overlapping Census blockgroups.
-We also define $p^{k}_I(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$.
-This definition embodies a computationally-simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
-Finally, $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$ is the estimated overall racial composition of hex $k$.
-Then, we estimate the mutual information in hex $k$ as
-	\begin{equation}
-		I(k) = \sum_{i \in N_k} p^k_I(i) D[p^i_Y(\cdot) \| p^k_Y(\cdot)]\;. 
-	\end{equation}
-	Using \eqref{approx}, the estimated Fisher information is 
-	\begin{equation}
-		J(k) \approx \frac{4 I(k)}{r^2}
-	\end{equation}
-	where $r$ is the grid radius. The estimated population in hex $k$ is $P_k = A_k\sum_{i \in N_k} \rho_i$, where $A_k$ is the cell area. We finally estimate $\E_X[J(X)]$ as 
-	\begin{equation}
-		J(X,Y) = \E_X[J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
-	\end{equation}
-\section*{Appendix 2: Relation of Local Mutual Information and Fisher Information}
-	In this appendix, we present a rigorous mathematical justification of the claim that the local mutual information approximates a fundamental quantity intrinsic to the distribution $p(Y,X)$.
-The argument relies on an idealization in which, rather than have a finite, discrete set of observations for each tract, we have a continuous field of observations across the map.
-The precise statement of this result and its justification follow.
-
-	Let $X$ be a continuous random variable taking values in $\R^n$, and let $Y$ be a discrete random variable define on finite alphabet $\mathcal{Y}$.
-Suppose further that $p(y|x) > 0$ and that $p(y|x)$ is differentiable as a function of $x$ for all $x,y \in \R^n \times \mathcal{Y}$.
-Fix $x_0 \in R^n$, and define $B_r \triangleq B_r(x_0) = \{ x \in R^n \;|\; \norm{x - x_0} \leq r \}$.
-Additionally, define the \emph{local mutual information} in $B_r$ as
-	\begin{equation}
-		I_r \triangleq I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;.
-	\end{equation}
-	where $D[p\|q] = \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$.
-
-	\begin{thm} \label{thm1}
-		Under the stated conditions, 
-		\begin{equation}
-			\lim_{r\rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{n}{2(n+2)} \emph{trace}\; J_Y(x_0)\;.
-		\end{equation}
-		where the Fisher information $J_Y(x_0)$ is 
-		\begin{align}
-		J_Y(x) &\triangleq \E\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
-		S_y(x) &\triangleq \log p(y|x)\;.
-	\end{align}
-	\end{thm}
-	The proof of Theorem \ref{thm1} proceeds by the application of a number of Taylor approximations, in tandem with a fundamental relationship of information geometry.
-We first expand out $I_r$ explicitly as
-	\begin{equation}
-		I_r = \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;. \label{eq:explicit}
-	\end{equation}
-
-	\begin{lm} \label{approximations}
-		The following approximation relationships hold for the components of \eqref{eq:explicit}:
-		\begin{enumerate}[label=\emph{(\alph*)}]
-			\item $p(X\in B_r,Y) = p(x_0, Y) v(B_r) + O(r^{n+2})$
-			\item $p(Y|X \in B_r) = p(Y|x_0) + e_y$ where the error terms $e_y$ satisfy $e_y \in O(r^{2})$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
-			\item $p(x|X \in B_r) $
-		\end{enumerate}
-	\end{lm}
-	\begin{proof}
-		For each approximation,  we directly apply Taylor expansions about $X = x_0$. 
-		\begin{enumerate}[label=(\alph*)]
-			\item We have 
-				\begin{align}
-					p(X\in B_r, Y) &= \int_{B_r} p(x,Y) \; d^nx \\
-					&= \int_{B_r} p(x_0,Y) + \frac{\partial p(x_0, Y)}{\partial x}(x - x_0) + O(\norm{x - x_0}^2) \; d^nx \\
-					&= p(x_0, Y)v(B_r) +  \frac{\partial p(x_0, Y)}{\partial x} \int_{B_r} (x - x_0)\; d^nx  \\
-					&\quad \quad+ O\left(\int_{B_r} \norm{x - x_0}^2\; d^nx\right) \\
-					&= p(x_0, Y)v(B_r) + O(r^{n+2})\;,
-				\end{align}
-				where the middle term vanishes due to spherical symmetry. 
-			\item The fact that the error terms $e_y$ must satisfy $\sum_{y \in \mathcal{Y}} e_y = 0$ follows from the fact that $p(Y|X \in B_r)$ must be a valid probability distribution over $\mathcal{Y}$.
-We'll now show that $e_y \in O(r^{ß2})$.
-First,
-				\begin{align}
-					p(X \in B_r) &= \sum_{y\in \mathcal{Y}} p(X \in B_r, y)  \\
-					&= \sum_{y\in \mathcal{Y}} \left[p(x_0, y)v(B_r) + O(r^{n+2}) \right] \\
-					&= p(x_0)v(B_r) + O(r^{2});\,
-				\end{align}
-			from part (a). Next, 
-				\begin{align}
-					p(Y|X \in B_r) &= \frac{p(X\in B_r, Y)}{p(X \in B_r)} \\
-					&= \frac{p(x_0, Y)v(B_r) + O(r^{n+2})}{p(x_0)v(B_r) + O(r^{n+2})} \\
-					&= p(Y|x_0) + O(r^2)\;,
-				\end{align}
-			which completes this part of the argument. 
-			\item First, 
-				\begin{align}
-					p(X \in B_r) &= \int_{B_r} p(x) \; d^nx  \\
-					&= \int_{B_r} \left[p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)\right] \; d^nx \\ 
-					&= p(x_0) v(B_r) + O(r^{n+2})\;,
-				\end{align}
-				where the middle term again vanishes through sphericla symmetry. Thus, for $x \in B_r$, we have
-				\begin{align}
-					p(x | X \in B_r) &= \frac{p(x)}{p(X \in B_r)} \\
-					&= \frac{p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)}{p(x_0) v(B_r) + O(r^{n+2})} \\
-					&= \frac{1 + O(r)}{v(B_r)}\;.
-				\end{align}
-		\end{enumerate}
-	\end{proof}
-
-	\begin{lm} The following approximation holds for the divergence factor in the integral \eqref{eq:explicit}
-		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] = D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})
-		\end{equation}
-	\end{lm}
-	\begin{proof}
-		We compute directly: 
-		\begin{align}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] &= \sum_{y \in \mathcal{Y}} p(y|x) \log \frac{p(y|x)}{p(y|X \in B_r)} \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|X \in B_r) \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log \left(p(y|x_0) + e_y)\right) \tag{from Lemma \ref{approximations}}\\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \left[\log p(y|x_0) + \frac{e_y}{p(y|x_0)} + O(e_y^2)\right] \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}  e_y \tag{quadratic terms negligible} \\
-			&=D[p(\cdot|x)\| p(\cdot|x_0)] \\
-			&\quad + \sum_{y \in \mathcal{Y}} \left(1 + \frac{1}{p(y|x_0)} \nabla p(y|x_0)(x - x_0) + O(r^{2})\right)    e_y \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \left[e_y + O(r^{3})\right] \tag{$e_y \in O(r^2)$}\\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^3) \tag{$\sum_{y \in \mathcal{Y}} e_y = 0$}
-		\end{align}
-	\end{proof}
-
-	\begin{lm} For any positive-semidefinite matrix $A \in \R^{n\times n}$, 
-		$$\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x = \frac{n}{n+2} r^{2} v(B_r) \text{\emph{trace}} (A)$$
-	\end{lm}
-
-	\begin{proof}
-		Since $A$ is positive-semidefinite, there exist an orthonormal matrix $P$ and a diagonal matrix $D$ such that $A = P^TDP$.
-Furthermore, the entries of $D$ are the eigenvalues $\{\lambda_i\}$ of $A$.
-Then,
-		\begin{align}
-			\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0, P^T DP(x-x_0) \right> d^n x\\
-			&= \int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x \;.
-		\end{align}
-		We can regard $P$ as a reparameterization of $B_r$; since $\det P = 1$, we have 
-		\begin{align}
-			\int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0,  D(x-x_0) \right> d^n x \\
-			&= r^n\int_{B_n} \left< rx, rDx \right> d^n x \\
-			&= r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x \;,
-		\end{align}
-		where $B_n$ is the unit $n$-ball. We also let $S_n(r)$ be the $n$-sphere of radius $r$. Continuing, 
-		\begin{align}
-			r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x &= r^{n+2} \int_{B_n} \sum_{i= 1}^n x_i^2 \lambda_i \; d^nx \\
-			&= r^{n+2} \sum_{i = 1}^n \lambda _i \int_{B_n} x_i^2 \; d^nx \\
-			&= \frac{r^{n+2}}{n} \sum_{i = 1}^n \lambda _i \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace}(A)  \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^2 S_{n-1}(\rho) d\rho \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^{n+1} S_{n-1}(1) d\rho \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A)  \frac{1}{n+2} S_{n-1}(1) \\
-			&= \frac{r^2}{n+2}  \text{trace} (A)  n r^{n}v(B_n(1)) \\
-			&= \frac{n}{n+2} r^{2} v(B_r) \text{trace} (A)\;,
-		\end{align}
-		as was to be shown. 
-	\end{proof}
-
-	\begin{fct*}
-		The Kullback-Leibler divergence and the Fisher information $J_Y$ are related according to the approximation 
-		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|x_0)] = \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3)
-		\end{equation}
-	\end{fct*}
-
-	We are finally ready to prove Theorem \ref{thm1}. Computing directly, we have 
-	\begin{align}
-		I_r(x_0) &\triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;. \\
-		&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \\
-		&= \int_{B_r} \left[\frac{1 + O(r)}{v(B_r)}\right] D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \tag{Lemma 1(c)}\\ 
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r}  \left(D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})\right) d^n x \tag{Lemma 2}\\
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left( \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3) + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left(\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\left(\frac{n}{n+2} r^{2} v(B_r) \text{trace} (J_Y(x_0)) + v(B_r)O(r^{3})\right) \\
-		&= r^2\frac{n}{2(n+2)}\left[1 + O(r)\right]\left(\text{trace} (J_Y(x_0)) + O(r^{3})\right)\;.
-	\end{align}
-	Dividing through by $r^2$ and computing the limit as $r \rightarrow 0$ proves the result. 
-\section*{Appendix 3: Table of Information Measures for Cities Studied}
-	The table below is sorted  by the population density of the city studied. 
-	\begin{tabular}{l c c c c }%
-	    \bfseries City & Density (km$^{-2}$) & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ 
-	    \csvreader[head to column names]{figs/table.csv}{}% use head of csv as column names
-	    {\\\hline\City & \Density & \HY & \IXY & \J}% specify your coloumns here
-    \end{tabular}
-
-\end{document}

### 2016-06-16 tex/methods.tex
reorganized

--- 2016-06-02 report/report.tex
@@
-% -----------------------------------------------------------------------------
-% ------------------------------- PREAMBLE ------------------------------------
-% -----------------------------------------------------------------------------[
-	\documentclass[english]{scrartcl}
-
-	\title{Spatial Complexity and Scaling in Cities}
-	\subtitle{v0.0}
-	\author{Phil Chodrow}
-	\date{May 29th, 2016}
-
-	\usepackage[T1]{fontenc}
-	\usepackage[utf8]{inputenc}
-	\usepackage{babel}
-	\usepackage{blindtext}
-	\usepackage{amsmath}
-	\usepackage{amsthm}
-	\usepackage{amsfonts}
-	\usepackage{amssymb}
-	\usepackage{enumitem}
-	\usepackage{graphicx}
-	\usepackage{pgfplotstable,filecontents}
-	\usepackage{booktabs}
-	\usepackage{caption}
-	\usepackage{subcaption}
-	\usepackage{csvsimple}
-
-	\pgfplotsset{compat=1.9}% supress warning
-
-	\setkomafont{disposition}{\normalfont\bfseries}
-
-	\newtheorem{thm}{Theorem}
-	\newtheorem{lm}{Lemma}
-	\newtheorem{cor}{Corrolary}
-	\newtheorem{clm}{Claim}
-	\newtheorem*{thm*}{Theorem}
-	\newtheorem*{lm*}{Lemma}
-	\newtheorem*{cor*}{Corrolary}
-	\newtheorem*{clm*}{Claim}
-	\newtheorem*{fct*}{Fact}
-
-	\newcommand\abs[1]{\left|#1\right|}
-	\newcommand\E[0]{\mathbb{E}}
-	\newcommand\R[0]{\mathbb{R}}
-	\newcommand\decision[2]{\underset{#2}{\overset{#1}{\gtreqless}}}
-	\newcommand{\argmax}{\operatornamewithlimits{argmax}}
-	\newcommand{\argmin}{\operatornamewithlimits{argmin}}
-	\newcommand\prob[0]{\mathbb{P}}
-	\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-% -----------------------------------------------------------------------------
-% --------------------------------- BODY --------------------------------------
-% -----------------------------------------------------------------------------
-\begin{document}
-
-\maketitle
-
-\abstract{We formulate the \emph{mean local information}, a novel measure of spatial compositional complexity in terms of the localized mutual information between spatial and compositional variables.
-This measure is closely related to the Fisher information of the underlying joint distribution, and is therefore an intrinsic property of the phenomena studied.
-Using this measure, we study spatial variations in racial trends in major American cities.
-The results neatly distinguish cities with large, monoracial neighborhoods from those with more intricate, interleaved neighborhood structure.
-Furthermore, the mean local information scales logarithmically with population density, inviting future research into possible universal dynamical processes of neighborhood formation that might explain this relationship.}
-
-\section{Introduction}
-
-	Information theory provides one natural approach toward thinking about complex systems.
-It achieves this by mapping the physical concept of structure to the epistemic concept of predictability.
-A complex system has enough structure to be at least partially predictable, but enough disorder to make prediction challenging.
-In this paper, we apply information-theoretic tools to the measurement of \emph{spatial compositional complexity}, with a focus on the spatial variation of racial trends in American cities.
-
-	Information-theoretic concepts have already found application in urban planning problems related to zoning and predicting population distributions \cite{Royal2014,Batty1974,Batty1976,Battya}.
-Substantial recent work has addressed the measurement of difference and disparity in cities through an information-theoretic lens \cite{Theil1971,Bettencourt2015,Roberto2015a,Roberto2015}.
-Attractive features of information-theoretic measures for this purpose include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to multiple demographic phenomena, and the fact that Theil's (\cite{Theil1971}) original index satisfaction of many (though not all) of the invariance properties desirable for the measurement of segregation \cite{Reardon2002}.
-\nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015}.
-
-	While closely-related to segregation, the concept of complexity has received less attention among quantitative sociologists.
-This is natural, since complexity is both less operational than segregation and challenging to define.
-On the other hand, as we demonstrate below, a spatial information-theoretic measure can be defined which measures the granularity of neighborhood structure, distinguishing cities with large, monolithic tracts of constant racial composition from those with many adjacent, smaller, neighborhoods of varying racial profiles.
-In doing so, it therefore captures one natural aspect of complexity in spatial compositional phenomena, and may also be of interest as a segregation measure as well.
-
-\section{Information Theory and Spatial Structure}
-
-	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two.
-Let $X$ be a spatial variable, such as a set of coordinates in space or a neighborhood name.
-Let $Y$ be a compositional variable that we aim to study, defined on an alphabet $\mathcal{Y}$.
-In the simple case in which $\mathcal{Y}  = \{\text{Black, White}\}$, Figure \ref{fig:toy} illustrates a range of possible joint distributions $p(X,Y)$.
-Comparing the completely undiverse city (a) and the spatially uniformly diverse (b) motivates the first fundamental information measure, the entropy:
-	\begin{equation}
-		H(Y) \triangleq \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
-	\end{equation}
-	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$.
-It thereby clearly distinguishes cities (a) and (b): since (a) is completely uniform, $H_a(Y) = 0$, while $H_b(Y) \neq 0$.
-On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
-To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
-	\begin{align}
-		D[p(Z)\|q(Z)] &\triangleq \sum_z p(z) \log \frac{p(z)}{q(z)} \\
-		I(X,Y) &\triangleq D[p(X,Y) \| p(X)p(Y)]
-	\end{align}
-	The divergence $D$ is (with some caveats) a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$.
-Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent.
-In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).
-In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence.
-
-	City (d) shares with city (c) the fact that residence and race are maximally dependent.
-However, city (d) embodies the ``checkerboard problem'': measures that use only the joint distribution $p(X,Y)$ without additionally considering the spatial information contained within $X$ will evaluate city (d) to be the same as city (c), despite their considerably different patterns of racial separation and potentially very different implications for planning and policy.
-A recent working paper \cite{Roberto2015} provides one highly operational approach to this problem, using road network topology and a weighting function that decays with distance to define a localized measure based on the Kullback-Leibler divergence.
-Here we pursue an alternative strategy, showing that a localization of the mutual information both measures spatial variation in race and corresponds to the estimation of a fundamental statistical property of the distribution $p(X,Y)$.
-
-	% NOTES FOR EXPANDING THE INTRODUCTION
-	% \begin{itemize}
-	% 	\item Concentration: $H(X|Y)$ with appropriate areal weighting. 
-	% 	\item Centralization: Probably not available. 
-	% 	\item Clustering: $J(X,Y)$
-	% 	\item Exposure: $I(X,Y)$? Needs to be read in conjunction with $H(Y)$. 
-	% 	\item Evenness: $H(X|Y)$, low score makes highly uneven. Or $I(X,Y)$? 
-	% \end{itemize}
-
-	% Possibly useful to note that \cite{Press2009a} express some dissatisfaction with their available clustering measures.
-
-	% \cite{Press2009a}
-	% \begin{enumerate}
-	% 	\item Multi-layered aspects of diversity and segregation.
-A limitation of our methods is that we aren't getting into historical or sociological questions; an advantage is that  they generalize fully to other phenomena, such as class, education level, or job type.
-Discuss other measures of diversity.
-	% 	\item Discuss the toy cities, existing information measures, what they mean, and why they need a complementary local measure. 
-	% 	\item Operational roles in modeling urban phenomena
-	% \end{enumerate}
-
-	\begin{figure}
-		\centering
-		  \includegraphics[width=.7\linewidth]{figs/toy.png}
-		  \begin{tabular}{c | c c c}
-			  City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
-			  \hline			
-			  (a) & 0.0 & 0.0 & 0.0\\
-			  (b) & 0.7 & 0.0 & 0.0\\
-			  (c) & 0.7 & 0.7 & 0.6\\
-			  (d) & 0.7 & 0.7 & 2.7\\
-			  \hline  
-			\end{tabular}
-		\caption{Information theory and the checkboard problem: model cities with various kinds of spatial diversity can be distinguished through progressively more subtle spatial information measures.}
-		\label{fig:toy}
-		\end{figure}
-
-\section{Methods}
+\subsection{Measuring Socio-Spatial Complexity}
 		To motivate our methods, we first consider an idealized problem in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is intuitively the map on which we work.
@@
 
-	\begin{figure}
-		\centering
-		\includegraphics[width=.5\textwidth]{figs/method.png}
-		\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
+		% \begin{figure}
+		% 	\centering
+		% 	\includegraphics[width=.5\textwidth]{figs/method.png}
+		% 	\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
 We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density.
 }
-		\label{fig:method}
-	\end{figure}
+		% 	\label{fig:method}
+		% \end{figure}
 
@@
 		We propose the aggregate quantity $J(X,Y)$ as a third  measure--alongside the entropy $H(Y)$ and mutual information $I(X,Y)$-- as a tool for the information-theoretic structure of spatial compositional complexity.
+	\subsection{Identifying Natural Neighborhoods}
+		Suppose that we have a collection of tracts labeled with the random variable $X$, and suppose that each tract has an additional label $C$ denoting the \emph{cluster} in which it lies.
+Then, it is possible to show that
+		\begin{equation}
+			I(X,Y) = I(Y,C) + I(X,Y|C) \label{eq:information_decomp}
+		\end{equation}
+		Equation \eqref{eq:information_decomp} has a natural interpretation: the information I have about $Y$ if I know $X$ is equal to the  information I have if I know the cluster $C$, plus the amount of information I have if I subsequently learn the exact location $X$.
+Usefully, equation \eqref{eq:information_decomp} decomposes the system-wide mutual information $I(X,Y)$ into a \emph{between-cluster} component $I(Y,C)$ and an \emph{within-cluster} component $I(X,Y|C)$.
+It is therefore comparable to various ``sum of squares'' decompositions that frequently appear in classical statistics.
+Indeed, it is possible to show that, when the differences between locations are small, the mutual information can be interpreted as a variance, in which case \eqref{eq:information_decomp} expresses the sum of squares composition directly.
 
-\section{Results}
-	We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S.
-Census Bureau, on race and ethnicity for counties corresponding to 28 large US cities.
-At this early stage, the counties used were chosen to correspond to the most densely populated urban cores; in future developments we could standardize to conduct analyses for Metropolitan Statistical Areas.
-We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'.
-For each city, we computed the entropy $H(Y)$ and the mutual information $I(X,Y)$.
-To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 1km.
-We then computed the estimated mutual information within each grid cell, and averaged the results weighted by population population.
-A technical specification of this approach is provided in Appendix 1, and a complete summary of cities and their associated information measures is provided in Appendix 3.
+		Finally, equation \eqref{eq:information_decomp} motivates a natural, information-based scheme for identifying natural neighborhoods.
+If all I had access to were the cluster labels $C$ and not the locations $X$, then my information would be $I(Y,C)$.
+A ``good'' clustering therefore maximizes the between-cluster information $I(Y,C)$, which entails minimizing the within-cluster information $I(X,Y|C)$.
+Solving this problem exactly is a challenging discrete optimization problem, and may not be computationally tractable.
+We can, however, construct a greedy algorithm with good performance.
+Suppose that we face the problem of choosing a collection $I = \{1,2,\ldots\}$ of locations to cluster together.
+The reduction in information associated with aggregating the locations $I$ into a single cluster is
+		\begin{align}
+			d(I) \triangleq  \sum_{i \in I} p(X = i)D[p(Y|X = i)\| p(Y|X \in I)] - p(X\in I)D[p(Y|X\in I) \| p(Y)] \label{eq:info_dist}
+		\end{align}
+		where $p(Y) = \sum_{x} p(x,Y)$ is the global marginal distribution.
+If we choose $I = \{i,j\}$, then equation \eqref{eq:info_dist} defines a natural information distance between locations $i$ and $j$.
+Like the KL divergence, this distance is strictly nonnegative; unlike the KL divergence, this distance is symmetric.
+Importantly, we can therefore use the distance $d(i,j)$ as a dissimilarity measure for the purposes of clustering.
+Our greedy procedure is simple: at each stage, the algorithm aggregates together two locations $i$ and $j$ such that
+		\begin{enumerate}
+			\item Locations $i$ and $j$ are adjacent in space. This requirement ensures that the results may be naturally understood as neighborhoods, rather than disparate tracts with similar demographics.  
+			\item Among all adjacent pairs, $d(i,j)$ is the smallest possible. 
+		\end{enumerate}
+		This greedy procedure defines a form of agglomerative clustering distinguished by two characteristics: its spatial constraints and its pursuit of minimal information loss at each step.
+As a greedy algorith, it possesses no guarantees for optimal solutions, but in practice its performance leads to intuitive neighborhoods.
 
-	\begin{figure}
-		\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
-		\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
-		\label{fig:info_cross}
-	\end{figure}	
-	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$.
-An overall positive trend is evident, reflecting the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.
-On the other hand, there are substantial variations in $J(X,Y)$ even in cities with comparable global variability $I(X,Y)$.
-For example, the cities of Detroit and Philadelphia provide a striking contrast.
-While they have comparable global variability $I(X,Y)$, Philadelphia's $J(X,Y)$ is substantially higher.
-This reflects the fact that Detroit is composed of large, highly-segregated, monoracial neighborhoods, whereas Philadelphia has a much more fine-grained, intricate neighborhood structure.
-These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities.
-
-	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density.
-In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities.
-A clear trend is evident, with $J(X,Y)$ growing linearly with the logarithm of density.
-We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix.
-One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own.
-
-	\begin{figure}
-		\includegraphics[width=1\textwidth]{figs/density_fisher.png}
-		\caption{The mean local information $J(X,Y)$ scales with the logarithm of population density.}
-		\label{fig:density}
-	\end{figure}
-
-\section{Discussion}
-
-	We have formulated a novel measure of urban spatial complexity, and applied it to an analysis of spatial distributions of race in American cities.
-We emphasize that the mean local spatial variability $J(X,Y)$ is novel piece of a more complete information-theoretic characterization that also includes the entropy $H(Y)$ and the global variability $I(X,Y)$.
-Our results suggest two major directions of further exploration:
-
-	One major physical question raised by our results is the origin of the scaling relation seen in \ref{fig:density}.
-This scaling relation may suggest a dynamical process of neighborhood formation and growth common to many American cities.
-If so, a physical understanding of this process in terms of individual-level behavior is called for.
-It is tempting to view the scaling relationship as an interplay of spatial preferential attachment and intrinsic limits to neighborhood density due to housing availability, but this view is speculative at this point.
-A theory to explain this behavior would be most welcome.
-
-	Another question raised is more operational.
-In many urban planning contexts, the amount of available data may make direct computations prohibitive.
-In such contexts, it may be useful to construct a simplifying model of the data the data (such as a clustering) in order to reduce dimensionality and make ``the big picture'' more visible.
-When a particular demographic phenomenon (such as race) is under consideration, it may be important to ensure that the model preserves the major patterns of variation in the granular data.
-We conjecture that the local variability $J(X,Y)$ measures the ``model-ability'' of such data sets.
-Intuitively, a city like Detroit with relatively low $J(X,Y)$ and large, monoracial neighborhoods should be easily representable with relatively few modeled clusters, one for each of the relatively few major neighborhoods.
-On the other hand, relatively speaking, a city like Philadelphia with very intricate neighborhood structure may require substantially greater model complexity to represent without large loss of information.
-A promising course of further study is to design an information-theoretic clustering algorithm designed to model urban patterns, and then compare this algorithm's performance to $J(X,Y)$.
-
-	A third question and final question relates to the time-dependence of spatial structure.
-On one time scale, daily movement around a city for work or leisure activities has the impact of ``mixing'' separated residents in public spaces, potentially leading to very different spatial patterns of racial difference.
-On another time scale, these measures may track how the spatial structure of cities evolves over decades, potentially shedding further light on the dynamics of city formation.
-
-\bibliography{/Users/phil/bibs/library.bib}{}
-\bibliographystyle{plain}
-
-\section*{Appendix 1: Computational Methods and Assumptions}
-	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[J_Y(X)]$ using blockgroup level data from the U.S. Census.
-
-	For fixed Census blockgroup $i$, let $P_i$ be the population, let $A_i$ be the area, let $\rho_i = P_i / A_i$ be the population density, and let $p^i_Y(y)$ be the observed proportion of racial group $y$.
-For hex $k$ in our hexagonal grid, let $N_k$ be the set of overlapping Census blockgroups.
-We also define $p^{k}_I(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$.
-This definition embodies a computationally-simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
-Finally, $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$ is the estimated overall racial composition of hex $k$.
-Then, we estimate the mutual information in hex $k$ as
-	\begin{equation}
-		I(k) = \sum_{i \in N_k} p^k_I(i) D[p^i_Y(\cdot) \| p^k_Y(\cdot)]\;. 
-	\end{equation}
-	Using \eqref{approx}, the estimated Fisher information is 
-	\begin{equation}
-		J(k) \approx \frac{4 I(k)}{r^2}
-	\end{equation}
-	where $r$ is the grid radius. The estimated population in hex $k$ is $P_k = A_k\sum_{i \in N_k} \rho_i$, where $A_k$ is the cell area. We finally estimate $\E_X[J(X)]$ as 
-	\begin{equation}
-		J(X,Y) = \E_X[J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
-	\end{equation}
-\section*{Appendix 2: Relation of Local Mutual Information and Fisher Information}
-	In this appendix, we present a rigorous mathematical justification of the claim that the local mutual information approximates a fundamental quantity intrinsic to the distribution $p(Y,X)$.
-The argument relies on an idealization in which, rather than have a finite, discrete set of observations for each tract, we have a continuous field of observations across the map.
-The precise statement of this result and its justification follow.
-
-	Let $X$ be a continuous random variable taking values in $\R^n$, and let $Y$ be a discrete random variable define on finite alphabet $\mathcal{Y}$.
-Suppose further that $p(y|x) > 0$ and that $p(y|x)$ is differentiable as a function of $x$ for all $x,y \in \R^n \times \mathcal{Y}$.
-Fix $x_0 \in R^n$, and define $B_r \triangleq B_r(x_0) = \{ x \in R^n \;|\; \norm{x - x_0} \leq r \}$.
-Additionally, define the \emph{local mutual information} in $B_r$ as
-	\begin{equation}
-		I_r \triangleq I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;.
-	\end{equation}
-	where $D[p\|q] = \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$.
-
-	\begin{thm} \label{thm1}
-		Under the stated conditions, 
-		\begin{equation}
-			\lim_{r\rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{n}{2(n+2)} \emph{trace}\; J_Y(x_0)\;.
-		\end{equation}
-		where the Fisher information $J_Y(x_0)$ is 
-		\begin{align}
-		J_Y(x) &\triangleq \E\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
-		S_y(x) &\triangleq \log p(y|x)\;.
-	\end{align}
-	\end{thm}
-	The proof of Theorem \ref{thm1} proceeds by the application of a number of Taylor approximations, in tandem with a fundamental relationship of information geometry.
-We first expand out $I_r$ explicitly as
-	\begin{equation}
-		I_r = \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;. \label{eq:explicit}
-	\end{equation}
-
-	\begin{lm} \label{approximations}
-		The following approximation relationships hold for the components of \eqref{eq:explicit}:
-		\begin{enumerate}[label=\emph{(\alph*)}]
-			\item $p(X\in B_r,Y) = p(x_0, Y) v(B_r) + O(r^{n+2})$
-			\item $p(Y|X \in B_r) = p(Y|x_0) + e_y$ where the error terms $e_y$ satisfy $e_y \in O(r^{2})$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
-			\item $p(x|X \in B_r) $
-		\end{enumerate}
-	\end{lm}
-	\begin{proof}
-		For each approximation,  we directly apply Taylor expansions about $X = x_0$. 
-		\begin{enumerate}[label=(\alph*)]
-			\item We have 
-				\begin{align}
-					p(X\in B_r, Y) &= \int_{B_r} p(x,Y) \; d^nx \\
-					&= \int_{B_r} p(x_0,Y) + \frac{\partial p(x_0, Y)}{\partial x}(x - x_0) + O(\norm{x - x_0}^2) \; d^nx \\
-					&= p(x_0, Y)v(B_r) +  \frac{\partial p(x_0, Y)}{\partial x} \int_{B_r} (x - x_0)\; d^nx  \\
-					&\quad \quad+ O\left(\int_{B_r} \norm{x - x_0}^2\; d^nx\right) \\
-					&= p(x_0, Y)v(B_r) + O(r^{n+2})\;,
-				\end{align}
-				where the middle term vanishes due to spherical symmetry. 
-			\item The fact that the error terms $e_y$ must satisfy $\sum_{y \in \mathcal{Y}} e_y = 0$ follows from the fact that $p(Y|X \in B_r)$ must be a valid probability distribution over $\mathcal{Y}$.
-We'll now show that $e_y \in O(r^{ß2})$.
-First,
-				\begin{align}
-					p(X \in B_r) &= \sum_{y\in \mathcal{Y}} p(X \in B_r, y)  \\
-					&= \sum_{y\in \mathcal{Y}} \left[p(x_0, y)v(B_r) + O(r^{n+2}) \right] \\
-					&= p(x_0)v(B_r) + O(r^{2});\,
-				\end{align}
-			from part (a). Next, 
-				\begin{align}
-					p(Y|X \in B_r) &= \frac{p(X\in B_r, Y)}{p(X \in B_r)} \\
-					&= \frac{p(x_0, Y)v(B_r) + O(r^{n+2})}{p(x_0)v(B_r) + O(r^{n+2})} \\
-					&= p(Y|x_0) + O(r^2)\;,
-				\end{align}
-			which completes this part of the argument. 
-			\item First, 
-				\begin{align}
-					p(X \in B_r) &= \int_{B_r} p(x) \; d^nx  \\
-					&= \int_{B_r} \left[p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)\right] \; d^nx \\ 
-					&= p(x_0) v(B_r) + O(r^{n+2})\;,
-				\end{align}
-				where the middle term again vanishes through sphericla symmetry. Thus, for $x \in B_r$, we have
-				\begin{align}
-					p(x | X \in B_r) &= \frac{p(x)}{p(X \in B_r)} \\
-					&= \frac{p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)}{p(x_0) v(B_r) + O(r^{n+2})} \\
-					&= \frac{1 + O(r)}{v(B_r)}\;.
-				\end{align}
-		\end{enumerate}
-	\end{proof}
-
-	\begin{lm} The following approximation holds for the divergence factor in the integral \eqref{eq:explicit}
-		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] = D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})
-		\end{equation}
-	\end{lm}
-	\begin{proof}
-		We compute directly: 
-		\begin{align}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] &= \sum_{y \in \mathcal{Y}} p(y|x) \log \frac{p(y|x)}{p(y|X \in B_r)} \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|X \in B_r) \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log \left(p(y|x_0) + e_y)\right) \tag{from Lemma \ref{approximations}}\\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \left[\log p(y|x_0) + \frac{e_y}{p(y|x_0)} + O(e_y^2)\right] \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}  e_y \tag{quadratic terms negligible} \\
-			&=D[p(\cdot|x)\| p(\cdot|x_0)] \\
-			&\quad + \sum_{y \in \mathcal{Y}} \left(1 + \frac{1}{p(y|x_0)} \nabla p(y|x_0)(x - x_0) + O(r^{2})\right)    e_y \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \left[e_y + O(r^{3})\right] \tag{$e_y \in O(r^2)$}\\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^3) \tag{$\sum_{y \in \mathcal{Y}} e_y = 0$}
-		\end{align}
-	\end{proof}
-
-	\begin{lm} For any positive-semidefinite matrix $A \in \R^{n\times n}$, 
-		$$\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x = \frac{n}{n+2} r^{2} v(B_r) \text{\emph{trace}} (A)$$
-	\end{lm}
-
-	\begin{proof}
-		Since $A$ is positive-semidefinite, there exist an orthonormal matrix $P$ and a diagonal matrix $D$ such that $A = P^TDP$.
-Furthermore, the entries of $D$ are the eigenvalues $\{\lambda_i\}$ of $A$.
-Then,
-		\begin{align}
-			\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0, P^T DP(x-x_0) \right> d^n x\\
-			&= \int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x \;.
-		\end{align}
-		We can regard $P$ as a reparameterization of $B_r$; since $\det P = 1$, we have 
-		\begin{align}
-			\int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0,  D(x-x_0) \right> d^n x \\
-			&= r^n\int_{B_n} \left< rx, rDx \right> d^n x \\
-			&= r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x \;,
-		\end{align}
-		where $B_n$ is the unit $n$-ball. We also let $S_n(r)$ be the $n$-sphere of radius $r$. Continuing, 
-		\begin{align}
-			r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x &= r^{n+2} \int_{B_n} \sum_{i= 1}^n x_i^2 \lambda_i \; d^nx \\
-			&= r^{n+2} \sum_{i = 1}^n \lambda _i \int_{B_n} x_i^2 \; d^nx \\
-			&= \frac{r^{n+2}}{n} \sum_{i = 1}^n \lambda _i \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace}(A)  \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^2 S_{n-1}(\rho) d\rho \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^{n+1} S_{n-1}(1) d\rho \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A)  \frac{1}{n+2} S_{n-1}(1) \\
-			&= \frac{r^2}{n+2}  \text{trace} (A)  n r^{n}v(B_n(1)) \\
-			&= \frac{n}{n+2} r^{2} v(B_r) \text{trace} (A)\;,
-		\end{align}
-		as was to be shown. 
-	\end{proof}
-
-	\begin{fct*}
-		The Kullback-Leibler divergence and the Fisher information $J_Y$ are related according to the approximation 
-		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|x_0)] = \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3)
-		\end{equation}
-	\end{fct*}
-
-	We are finally ready to prove Theorem \ref{thm1}. Computing directly, we have 
-	\begin{align}
-		I_r(x_0) &\triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;. \\
-		&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \\
-		&= \int_{B_r} \left[\frac{1 + O(r)}{v(B_r)}\right] D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \tag{Lemma 1(c)}\\ 
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r}  \left(D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})\right) d^n x \tag{Lemma 2}\\
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left( \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3) + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left(\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\left(\frac{n}{n+2} r^{2} v(B_r) \text{trace} (J_Y(x_0)) + v(B_r)O(r^{3})\right) \\
-		&= r^2\frac{n}{2(n+2)}\left[1 + O(r)\right]\left(\text{trace} (J_Y(x_0)) + O(r^{3})\right)\;.
-	\end{align}
-	Dividing through by $r^2$ and computing the limit as $r \rightarrow 0$ proves the result. 
-\section*{Appendix 3: Table of Information Measures for Cities Studied}
-	The table below is sorted  by the population density of the city studied. 
-	\begin{tabular}{l c c c c }%
-	    \bfseries City & Density (km$^{-2}$) & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ 
-	    \csvreader[head to column names]{figs/table.csv}{}% use head of csv as column names
-	    {\\\hline\City & \Density & \HY & \IXY & \J}% specify your coloumns here
-    \end{tabular}
-
-\end{document}
+	\subsection{Complexity and Neighborhood Identification}

### 2016-06-16 tex/results.tex
reorganized

--- 2016-06-02 report/report.tex
@@
-% -----------------------------------------------------------------------------
-% ------------------------------- PREAMBLE ------------------------------------
-% -----------------------------------------------------------------------------[
-	\documentclass[english]{scrartcl}
-
-	\title{Spatial Complexity and Scaling in Cities}
-	\subtitle{v0.0}
-	\author{Phil Chodrow}
-	\date{May 29th, 2016}
-
-	\usepackage[T1]{fontenc}
-	\usepackage[utf8]{inputenc}
-	\usepackage{babel}
-	\usepackage{blindtext}
-	\usepackage{amsmath}
-	\usepackage{amsthm}
-	\usepackage{amsfonts}
-	\usepackage{amssymb}
-	\usepackage{enumitem}
-	\usepackage{graphicx}
-	\usepackage{pgfplotstable,filecontents}
-	\usepackage{booktabs}
-	\usepackage{caption}
-	\usepackage{subcaption}
-	\usepackage{csvsimple}
-
-	\pgfplotsset{compat=1.9}% supress warning
-
-	\setkomafont{disposition}{\normalfont\bfseries}
-
-	\newtheorem{thm}{Theorem}
-	\newtheorem{lm}{Lemma}
-	\newtheorem{cor}{Corrolary}
-	\newtheorem{clm}{Claim}
-	\newtheorem*{thm*}{Theorem}
-	\newtheorem*{lm*}{Lemma}
-	\newtheorem*{cor*}{Corrolary}
-	\newtheorem*{clm*}{Claim}
-	\newtheorem*{fct*}{Fact}
-
-	\newcommand\abs[1]{\left|#1\right|}
-	\newcommand\E[0]{\mathbb{E}}
-	\newcommand\R[0]{\mathbb{R}}
-	\newcommand\decision[2]{\underset{#2}{\overset{#1}{\gtreqless}}}
-	\newcommand{\argmax}{\operatornamewithlimits{argmax}}
-	\newcommand{\argmin}{\operatornamewithlimits{argmin}}
-	\newcommand\prob[0]{\mathbb{P}}
-	\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-% -----------------------------------------------------------------------------
-% --------------------------------- BODY --------------------------------------
-% -----------------------------------------------------------------------------
-\begin{document}
-
-\maketitle
-
-\abstract{We formulate the \emph{mean local information}, a novel measure of spatial compositional complexity in terms of the localized mutual information between spatial and compositional variables.
-This measure is closely related to the Fisher information of the underlying joint distribution, and is therefore an intrinsic property of the phenomena studied.
-Using this measure, we study spatial variations in racial trends in major American cities.
-The results neatly distinguish cities with large, monoracial neighborhoods from those with more intricate, interleaved neighborhood structure.
-Furthermore, the mean local information scales logarithmically with population density, inviting future research into possible universal dynamical processes of neighborhood formation that might explain this relationship.}
-
-\section{Introduction}
-
-	Information theory provides one natural approach toward thinking about complex systems.
-It achieves this by mapping the physical concept of structure to the epistemic concept of predictability.
-A complex system has enough structure to be at least partially predictable, but enough disorder to make prediction challenging.
-In this paper, we apply information-theoretic tools to the measurement of \emph{spatial compositional complexity}, with a focus on the spatial variation of racial trends in American cities.
-
-	Information-theoretic concepts have already found application in urban planning problems related to zoning and predicting population distributions \cite{Royal2014,Batty1974,Batty1976,Battya}.
-Substantial recent work has addressed the measurement of difference and disparity in cities through an information-theoretic lens \cite{Theil1971,Bettencourt2015,Roberto2015a,Roberto2015}.
-Attractive features of information-theoretic measures for this purpose include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to multiple demographic phenomena, and the fact that Theil's (\cite{Theil1971}) original index satisfaction of many (though not all) of the invariance properties desirable for the measurement of segregation \cite{Reardon2002}.
-\nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015}.
-
-	While closely-related to segregation, the concept of complexity has received less attention among quantitative sociologists.
-This is natural, since complexity is both less operational than segregation and challenging to define.
-On the other hand, as we demonstrate below, a spatial information-theoretic measure can be defined which measures the granularity of neighborhood structure, distinguishing cities with large, monolithic tracts of constant racial composition from those with many adjacent, smaller, neighborhoods of varying racial profiles.
-In doing so, it therefore captures one natural aspect of complexity in spatial compositional phenomena, and may also be of interest as a segregation measure as well.
-
-\section{Information Theory and Spatial Structure}
-
-	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two.
-Let $X$ be a spatial variable, such as a set of coordinates in space or a neighborhood name.
-Let $Y$ be a compositional variable that we aim to study, defined on an alphabet $\mathcal{Y}$.
-In the simple case in which $\mathcal{Y}  = \{\text{Black, White}\}$, Figure \ref{fig:toy} illustrates a range of possible joint distributions $p(X,Y)$.
-Comparing the completely undiverse city (a) and the spatially uniformly diverse (b) motivates the first fundamental information measure, the entropy:
-	\begin{equation}
-		H(Y) \triangleq \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
-	\end{equation}
-	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$.
-It thereby clearly distinguishes cities (a) and (b): since (a) is completely uniform, $H_a(Y) = 0$, while $H_b(Y) \neq 0$.
-On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
-To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
-	\begin{align}
-		D[p(Z)\|q(Z)] &\triangleq \sum_z p(z) \log \frac{p(z)}{q(z)} \\
-		I(X,Y) &\triangleq D[p(X,Y) \| p(X)p(Y)]
-	\end{align}
-	The divergence $D$ is (with some caveats) a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$.
-Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent.
-In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).
-In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence.
-
-	City (d) shares with city (c) the fact that residence and race are maximally dependent.
-However, city (d) embodies the ``checkerboard problem'': measures that use only the joint distribution $p(X,Y)$ without additionally considering the spatial information contained within $X$ will evaluate city (d) to be the same as city (c), despite their considerably different patterns of racial separation and potentially very different implications for planning and policy.
-A recent working paper \cite{Roberto2015} provides one highly operational approach to this problem, using road network topology and a weighting function that decays with distance to define a localized measure based on the Kullback-Leibler divergence.
-Here we pursue an alternative strategy, showing that a localization of the mutual information both measures spatial variation in race and corresponds to the estimation of a fundamental statistical property of the distribution $p(X,Y)$.
-
-	% NOTES FOR EXPANDING THE INTRODUCTION
-	% \begin{itemize}
-	% 	\item Concentration: $H(X|Y)$ with appropriate areal weighting. 
-	% 	\item Centralization: Probably not available. 
-	% 	\item Clustering: $J(X,Y)$
-	% 	\item Exposure: $I(X,Y)$? Needs to be read in conjunction with $H(Y)$. 
-	% 	\item Evenness: $H(X|Y)$, low score makes highly uneven. Or $I(X,Y)$? 
-	% \end{itemize}
-
-	% Possibly useful to note that \cite{Press2009a} express some dissatisfaction with their available clustering measures.
-
-	% \cite{Press2009a}
-	% \begin{enumerate}
-	% 	\item Multi-layered aspects of diversity and segregation.
-A limitation of our methods is that we aren't getting into historical or sociological questions; an advantage is that  they generalize fully to other phenomena, such as class, education level, or job type.
-Discuss other measures of diversity.
-	% 	\item Discuss the toy cities, existing information measures, what they mean, and why they need a complementary local measure. 
-	% 	\item Operational roles in modeling urban phenomena
-	% \end{enumerate}
-
-	\begin{figure}
-		\centering
-		  \includegraphics[width=.7\linewidth]{figs/toy.png}
-		  \begin{tabular}{c | c c c}
-			  City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
-			  \hline			
-			  (a) & 0.0 & 0.0 & 0.0\\
-			  (b) & 0.7 & 0.0 & 0.0\\
-			  (c) & 0.7 & 0.7 & 0.6\\
-			  (d) & 0.7 & 0.7 & 2.7\\
-			  \hline  
-			\end{tabular}
-		\caption{Information theory and the checkboard problem: model cities with various kinds of spatial diversity can be distinguished through progressively more subtle spatial information measures.}
-		\label{fig:toy}
-		\end{figure}
-
-\section{Methods}
-	To motivate our methods, we first consider an idealized problem in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is intuitively the map on which we work.
-
-	Fix a point $x_0 \in M$ and a radius $r > 0$.
-Let $B_r(x_0)$ denote the ball of radius $r$ about $x_0$.
-Then, define the \emph{local mutual information in radius $r$} as the mutual information between $X$ and $Y$, restricted to the small ball $B_r(x_0)$ about $x_0$:
-	\begin{equation}
-		I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r(x_0))]|X \in B_r(x_0)]
-	\end{equation}
-	Intuitively, $I_r(x_0)$ measures how much the probability field $p(y|x)$ varies with $x$ in a small neighborhood of $x_0$.
-It is therefore a natural measure of local complexity, aligned in approach with the global mutual information but designed to detect pattern in local variations.
-$I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$, corresponding to a complete lack of local compositional complexity.
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.5\textwidth]{figs/method.png}
-		\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
-We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density.
-}
-		\label{fig:method}
-	\end{figure}
-
-	In this definition, $r$ may be thought of as the spatial resolution at which we conduct analysis, and $I_r(x_0)$ is highly dependent on $r$.
-However, it is possible to show that $I_r(x_0)$ is related to a fundamental statistical property of the probability field $p(y|x)$ that is resolution-independent.
-Under the stated conditions, the following approximation holds:
-
-	\begin{equation}
-		\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace} (J_Y(x))\;, \label{approx}
-	\end{equation}
-	where $J_Y(x)$ is the Fisher information matrix in $Y$ about $x$, defined as 
-	\begin{equation}
-		J_Y(x) \triangleq \E\left[ (\nabla_x \log p(y|x))(\nabla_x \log p(y|x))^T \right]\;.
-	\end{equation}
-	A more formal statement and proof of \eqref{approx} are provided in Appendix 2.
-The Fisher information $J_Y$ is a fundamental quantity in statistics and information theory.
-From a geometric perspective, $J_Y$ provides the natural intrinsic metric in the geometric space of probability distributions parameterized by the spatial variable $x$.
-Equation \eqref{approx} therefore expresses a relationship between the local mutual information $I_r(x)$ and the information geometry of the underlying probability field.
-Corresponding to the fact that $I_r(x_0)$ vanishes if and only if $p(y|x)$ is constant in $B_r(x_0)$, $J_Y(x_0) = 0$ if and only if $\nabla_x p(y|x_0) = 0$ for all $y$.
-This implies that $x_0$ is a stationary point, about which the probability field $p(y|x_0)$ exhibits only small (2nd order or smaller) changes with respect to changes in $x$.
-
-	Since the Fisher information is a strictly local measure of statistical variability around $x$, we can aggregate the Fisher information to derive a measure of average local variability.
-The \emph{mean local information} is
-	\begin{equation}
-	J(X,Y) \triangleq \E_X[\text{trace }J_Y(X)]
-	\end{equation} 
-	We propose the aggregate quantity $J(X,Y)$ as a third  measure--alongside the entropy $H(Y)$ and mutual information $I(X,Y)$-- as a tool for the information-theoretic structure of spatial compositional complexity.
-
-\section{Results}
 We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S.
@@
 	\end{figure}
-
-\section{Discussion}
-
-	We have formulated a novel measure of urban spatial complexity, and applied it to an analysis of spatial distributions of race in American cities.
-We emphasize that the mean local spatial variability $J(X,Y)$ is novel piece of a more complete information-theoretic characterization that also includes the entropy $H(Y)$ and the global variability $I(X,Y)$.
-Our results suggest two major directions of further exploration:
-
-	One major physical question raised by our results is the origin of the scaling relation seen in \ref{fig:density}.
-This scaling relation may suggest a dynamical process of neighborhood formation and growth common to many American cities.
-If so, a physical understanding of this process in terms of individual-level behavior is called for.
-It is tempting to view the scaling relationship as an interplay of spatial preferential attachment and intrinsic limits to neighborhood density due to housing availability, but this view is speculative at this point.
-A theory to explain this behavior would be most welcome.
-
-	Another question raised is more operational.
-In many urban planning contexts, the amount of available data may make direct computations prohibitive.
-In such contexts, it may be useful to construct a simplifying model of the data the data (such as a clustering) in order to reduce dimensionality and make ``the big picture'' more visible.
-When a particular demographic phenomenon (such as race) is under consideration, it may be important to ensure that the model preserves the major patterns of variation in the granular data.
-We conjecture that the local variability $J(X,Y)$ measures the ``model-ability'' of such data sets.
-Intuitively, a city like Detroit with relatively low $J(X,Y)$ and large, monoracial neighborhoods should be easily representable with relatively few modeled clusters, one for each of the relatively few major neighborhoods.
-On the other hand, relatively speaking, a city like Philadelphia with very intricate neighborhood structure may require substantially greater model complexity to represent without large loss of information.
-A promising course of further study is to design an information-theoretic clustering algorithm designed to model urban patterns, and then compare this algorithm's performance to $J(X,Y)$.
-
-	A third question and final question relates to the time-dependence of spatial structure.
-On one time scale, daily movement around a city for work or leisure activities has the impact of ``mixing'' separated residents in public spaces, potentially leading to very different spatial patterns of racial difference.
-On another time scale, these measures may track how the spatial structure of cities evolves over decades, potentially shedding further light on the dynamics of city formation.
-
-\bibliography{/Users/phil/bibs/library.bib}{}
-\bibliographystyle{plain}
-
-\section*{Appendix 1: Computational Methods and Assumptions}
-	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[J_Y(X)]$ using blockgroup level data from the U.S. Census.
-
-	For fixed Census blockgroup $i$, let $P_i$ be the population, let $A_i$ be the area, let $\rho_i = P_i / A_i$ be the population density, and let $p^i_Y(y)$ be the observed proportion of racial group $y$.
-For hex $k$ in our hexagonal grid, let $N_k$ be the set of overlapping Census blockgroups.
-We also define $p^{k}_I(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$.
-This definition embodies a computationally-simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
-Finally, $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$ is the estimated overall racial composition of hex $k$.
-Then, we estimate the mutual information in hex $k$ as
-	\begin{equation}
-		I(k) = \sum_{i \in N_k} p^k_I(i) D[p^i_Y(\cdot) \| p^k_Y(\cdot)]\;. 
-	\end{equation}
-	Using \eqref{approx}, the estimated Fisher information is 
-	\begin{equation}
-		J(k) \approx \frac{4 I(k)}{r^2}
-	\end{equation}
-	where $r$ is the grid radius. The estimated population in hex $k$ is $P_k = A_k\sum_{i \in N_k} \rho_i$, where $A_k$ is the cell area. We finally estimate $\E_X[J(X)]$ as 
-	\begin{equation}
-		J(X,Y) = \E_X[J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
-	\end{equation}
-\section*{Appendix 2: Relation of Local Mutual Information and Fisher Information}
-	In this appendix, we present a rigorous mathematical justification of the claim that the local mutual information approximates a fundamental quantity intrinsic to the distribution $p(Y,X)$.
-The argument relies on an idealization in which, rather than have a finite, discrete set of observations for each tract, we have a continuous field of observations across the map.
-The precise statement of this result and its justification follow.
-
-	Let $X$ be a continuous random variable taking values in $\R^n$, and let $Y$ be a discrete random variable define on finite alphabet $\mathcal{Y}$.
-Suppose further that $p(y|x) > 0$ and that $p(y|x)$ is differentiable as a function of $x$ for all $x,y \in \R^n \times \mathcal{Y}$.
-Fix $x_0 \in R^n$, and define $B_r \triangleq B_r(x_0) = \{ x \in R^n \;|\; \norm{x - x_0} \leq r \}$.
-Additionally, define the \emph{local mutual information} in $B_r$ as
-	\begin{equation}
-		I_r \triangleq I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;.
-	\end{equation}
-	where $D[p\|q] = \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$.
-
-	\begin{thm} \label{thm1}
-		Under the stated conditions, 
-		\begin{equation}
-			\lim_{r\rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{n}{2(n+2)} \emph{trace}\; J_Y(x_0)\;.
-		\end{equation}
-		where the Fisher information $J_Y(x_0)$ is 
-		\begin{align}
-		J_Y(x) &\triangleq \E\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
-		S_y(x) &\triangleq \log p(y|x)\;.
-	\end{align}
-	\end{thm}
-	The proof of Theorem \ref{thm1} proceeds by the application of a number of Taylor approximations, in tandem with a fundamental relationship of information geometry.
-We first expand out $I_r$ explicitly as
-	\begin{equation}
-		I_r = \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;. \label{eq:explicit}
-	\end{equation}
-
-	\begin{lm} \label{approximations}
-		The following approximation relationships hold for the components of \eqref{eq:explicit}:
-		\begin{enumerate}[label=\emph{(\alph*)}]
-			\item $p(X\in B_r,Y) = p(x_0, Y) v(B_r) + O(r^{n+2})$
-			\item $p(Y|X \in B_r) = p(Y|x_0) + e_y$ where the error terms $e_y$ satisfy $e_y \in O(r^{2})$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
-			\item $p(x|X \in B_r) $
-		\end{enumerate}
-	\end{lm}
-	\begin{proof}
-		For each approximation,  we directly apply Taylor expansions about $X = x_0$. 
-		\begin{enumerate}[label=(\alph*)]
-			\item We have 
-				\begin{align}
-					p(X\in B_r, Y) &= \int_{B_r} p(x,Y) \; d^nx \\
-					&= \int_{B_r} p(x_0,Y) + \frac{\partial p(x_0, Y)}{\partial x}(x - x_0) + O(\norm{x - x_0}^2) \; d^nx \\
-					&= p(x_0, Y)v(B_r) +  \frac{\partial p(x_0, Y)}{\partial x} \int_{B_r} (x - x_0)\; d^nx  \\
-					&\quad \quad+ O\left(\int_{B_r} \norm{x - x_0}^2\; d^nx\right) \\
-					&= p(x_0, Y)v(B_r) + O(r^{n+2})\;,
-				\end{align}
-				where the middle term vanishes due to spherical symmetry. 
-			\item The fact that the error terms $e_y$ must satisfy $\sum_{y \in \mathcal{Y}} e_y = 0$ follows from the fact that $p(Y|X \in B_r)$ must be a valid probability distribution over $\mathcal{Y}$.
-We'll now show that $e_y \in O(r^{ß2})$.
-First,
-				\begin{align}
-					p(X \in B_r) &= \sum_{y\in \mathcal{Y}} p(X \in B_r, y)  \\
-					&= \sum_{y\in \mathcal{Y}} \left[p(x_0, y)v(B_r) + O(r^{n+2}) \right] \\
-					&= p(x_0)v(B_r) + O(r^{2});\,
-				\end{align}
-			from part (a). Next, 
-				\begin{align}
-					p(Y|X \in B_r) &= \frac{p(X\in B_r, Y)}{p(X \in B_r)} \\
-					&= \frac{p(x_0, Y)v(B_r) + O(r^{n+2})}{p(x_0)v(B_r) + O(r^{n+2})} \\
-					&= p(Y|x_0) + O(r^2)\;,
-				\end{align}
-			which completes this part of the argument. 
-			\item First, 
-				\begin{align}
-					p(X \in B_r) &= \int_{B_r} p(x) \; d^nx  \\
-					&= \int_{B_r} \left[p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)\right] \; d^nx \\ 
-					&= p(x_0) v(B_r) + O(r^{n+2})\;,
-				\end{align}
-				where the middle term again vanishes through sphericla symmetry. Thus, for $x \in B_r$, we have
-				\begin{align}
-					p(x | X \in B_r) &= \frac{p(x)}{p(X \in B_r)} \\
-					&= \frac{p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)}{p(x_0) v(B_r) + O(r^{n+2})} \\
-					&= \frac{1 + O(r)}{v(B_r)}\;.
-				\end{align}
-		\end{enumerate}
-	\end{proof}
-
-	\begin{lm} The following approximation holds for the divergence factor in the integral \eqref{eq:explicit}
-		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] = D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})
-		\end{equation}
-	\end{lm}
-	\begin{proof}
-		We compute directly: 
-		\begin{align}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] &= \sum_{y \in \mathcal{Y}} p(y|x) \log \frac{p(y|x)}{p(y|X \in B_r)} \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|X \in B_r) \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log \left(p(y|x_0) + e_y)\right) \tag{from Lemma \ref{approximations}}\\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \left[\log p(y|x_0) + \frac{e_y}{p(y|x_0)} + O(e_y^2)\right] \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}  e_y \tag{quadratic terms negligible} \\
-			&=D[p(\cdot|x)\| p(\cdot|x_0)] \\
-			&\quad + \sum_{y \in \mathcal{Y}} \left(1 + \frac{1}{p(y|x_0)} \nabla p(y|x_0)(x - x_0) + O(r^{2})\right)    e_y \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \left[e_y + O(r^{3})\right] \tag{$e_y \in O(r^2)$}\\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^3) \tag{$\sum_{y \in \mathcal{Y}} e_y = 0$}
-		\end{align}
-	\end{proof}
-
-	\begin{lm} For any positive-semidefinite matrix $A \in \R^{n\times n}$, 
-		$$\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x = \frac{n}{n+2} r^{2} v(B_r) \text{\emph{trace}} (A)$$
-	\end{lm}
-
-	\begin{proof}
-		Since $A$ is positive-semidefinite, there exist an orthonormal matrix $P$ and a diagonal matrix $D$ such that $A = P^TDP$.
-Furthermore, the entries of $D$ are the eigenvalues $\{\lambda_i\}$ of $A$.
-Then,
-		\begin{align}
-			\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0, P^T DP(x-x_0) \right> d^n x\\
-			&= \int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x \;.
-		\end{align}
-		We can regard $P$ as a reparameterization of $B_r$; since $\det P = 1$, we have 
-		\begin{align}
-			\int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0,  D(x-x_0) \right> d^n x \\
-			&= r^n\int_{B_n} \left< rx, rDx \right> d^n x \\
-			&= r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x \;,
-		\end{align}
-		where $B_n$ is the unit $n$-ball. We also let $S_n(r)$ be the $n$-sphere of radius $r$. Continuing, 
-		\begin{align}
-			r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x &= r^{n+2} \int_{B_n} \sum_{i= 1}^n x_i^2 \lambda_i \; d^nx \\
-			&= r^{n+2} \sum_{i = 1}^n \lambda _i \int_{B_n} x_i^2 \; d^nx \\
-			&= \frac{r^{n+2}}{n} \sum_{i = 1}^n \lambda _i \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace}(A)  \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^2 S_{n-1}(\rho) d\rho \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^{n+1} S_{n-1}(1) d\rho \\
-			&= \frac{r^{n+2}}{n} \text{trace} (A)  \frac{1}{n+2} S_{n-1}(1) \\
-			&= \frac{r^2}{n+2}  \text{trace} (A)  n r^{n}v(B_n(1)) \\
-			&= \frac{n}{n+2} r^{2} v(B_r) \text{trace} (A)\;,
-		\end{align}
-		as was to be shown. 
-	\end{proof}
-
-	\begin{fct*}
-		The Kullback-Leibler divergence and the Fisher information $J_Y$ are related according to the approximation 
-		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|x_0)] = \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3)
-		\end{equation}
-	\end{fct*}
-
-	We are finally ready to prove Theorem \ref{thm1}. Computing directly, we have 
-	\begin{align}
-		I_r(x_0) &\triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;. \\
-		&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \\
-		&= \int_{B_r} \left[\frac{1 + O(r)}{v(B_r)}\right] D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \tag{Lemma 1(c)}\\ 
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r}  \left(D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})\right) d^n x \tag{Lemma 2}\\
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left( \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3) + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left(\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\left(\frac{n}{n+2} r^{2} v(B_r) \text{trace} (J_Y(x_0)) + v(B_r)O(r^{3})\right) \\
-		&= r^2\frac{n}{2(n+2)}\left[1 + O(r)\right]\left(\text{trace} (J_Y(x_0)) + O(r^{3})\right)\;.
-	\end{align}
-	Dividing through by $r^2$ and computing the limit as $r \rightarrow 0$ proves the result. 
-\section*{Appendix 3: Table of Information Measures for Cities Studied}
-	The table below is sorted  by the population density of the city studied. 
-	\begin{tabular}{l c c c c }%
-	    \bfseries City & Density (km$^{-2}$) & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ 
-	    \csvreader[head to column names]{figs/table.csv}{}% use head of csv as column names
-	    {\\\hline\City & \Density & \HY & \IXY & \J}% specify your coloumns here
-    \end{tabular}
-
-\end{document}

### 2016-06-16 1_get_states.R

@@ -1,0 +1,96 @@
+library(sp, quietly = TRUE)
+library(compx, quietly = TRUE)
+library(acs, quietly = TRUE)
+library(plyr, quietly = TRUE)
+library(dplyr, quietly = TRUE)
+library(readr, quietly = TRUE)
+library(ggmap, quietly = TRUE)
+library(tidyr, quietly = TRUE)
+library(magrittr, quietly = TRUE)
+library(ggrepel, quietly = TRUE)
+library(grid, quietly = TRUE)
+library(rgdal, quietly = TRUE)
+library(tigris, quietly = TRUE)
+library(stringr, quietly = TRUE)
+library(acs, quietly = TRUE)
+
+
+# get list of states
+
+cities <- read_csv('assumptions/cities.csv')
+cities$state = word(cities$state, -1)
+
+
+if(!dir.exists('data/states')){
+	dir.create('data/states')
+}
+
+
+# Define data-getter
+
+f <- function(state){
+	counties <- acs.fetch(geography=geo.make(state=state, county="*"), 
+				   endyear = 2014,
+				   table.number="B01003")
+	counties <- as.numeric(geography(counties)[[3]])
+	
+	race <- acs::acs.fetch(endyear = 2014,
+						   span = 5,
+						   geography = acs::geo.make(state = state, county = counties, tract = '*', block.group = '*'),
+						   table.number = 'B03002',
+						   col.names = "pretty")
+	
+	race <- cbind(data.frame(race@geography),
+				  data.frame(race@estimate)) %>%
+		tbl_df() %>%
+		mutate(GEOID = paste0(str_pad(state, 2, 'left', pad = '0'),
+							  str_pad(county, 3, 'left', pad = '0'),
+							  str_pad(tract, 6, 'left', pad = '0'),
+							  str_pad(blockgroup, 1, 'left', pad = '0')))
+	
+	race$Hispanic <- race[names(race)[!grepl('.Not.Hispanic.or.Latino.',names(race))]] %>%
+		select(-(`NAME`:`Hispanic.or.Latino.by.Race..Hispanic.or.Latino.`)) %>%
+		select(-GEOID) %>%
+		rowSums()
+	
+	others <- c('Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..American.Indian.and.Alaska.Native.alone',
[…]

### 2016-06-16 2_make_cities.R

@@ -1,0 +1,56 @@
+library(sp, quietly = TRUE)
+library(compx, quietly = TRUE)
+library(plyr, quietly = TRUE)
+library(dplyr, quietly = TRUE)
+library(readr, quietly = TRUE)
+library(tidyr, quietly = TRUE)
+library(magrittr, quietly = TRUE)
+library(grid, quietly = TRUE)
+library(rgdal, quietly = TRUE)
+library(stringr, quietly = TRUE)
+
+
+city_df <- read_csv('assumptions/cities.csv')
+cities <- unique(city_df$name)
+cities <- cities[!is.element(cities, list.dirs('cities', full.names = FALSE))]
+
+states <- city_df$state %>% 
+	str_split('--') %>% 
+	do.call(c,.) %>% 
+	unique()
+
+shapes <- list()
+for(state in states){
+	print(paste('Loading', state))
+	shapes[state] <- readOGR(dsn = paste0('data/states/',state), layer = 'geo', verbose = FALSE)
+}
+
+if(!dir.exists('data/cities')){
+	dir.create('data/cities')
+}
+
+
+
+
+for(city in cities){
+	print(city)
+	city_info <- city_df %>%
+		filter(name == city) 
+	
+	city_components <- list()
+	for(i in 1:nrow(city_info)){
+		state <- city_info$state[i]
+		county <- city_info$county[i]
+		component <- shapes[[state]]
+		component <- component[as.integer(as.character(component@data$COUNTYFP)) == county,]
+		city_components[paste(state, county, sep = '_')] <- spChFIDs(component,
+																	  as.character(paste(state, county, rownames(as(component,"data.frame")))))
+	}
+	city_polys <- do.call(rbind, city_components)
+	city_polys <- city_polys[city_polys@data$total != 0,]
+	tryCatch(writeOGR(city_polys, paste0("data/cities/", city),'geo', driver = 'ESRI Shapefile', morphToESRI = TRUE),
+			 error = function(e) NULL)
+}
+	
+	
+	

### 2016-06-16 3_local_info.R

@@ -1,0 +1,116 @@
+library(sp, quietly = TRUE)
+library(compx, quietly = TRUE)
+library(acs, quietly = TRUE)
+library(plyr, quietly = TRUE)
+library(dplyr, quietly = TRUE)
+library(readr, quietly = TRUE)
+library(ggmap, quietly = TRUE)
+library(tidyr, quietly = TRUE)
+library(magrittr, quietly = TRUE)
+library(ggrepel, quietly = TRUE)
+library(grid, quietly = TRUE)
+library(rgdal, quietly = TRUE)
+
+# Toy cities ----------
+
+# checkerboard_illustration()
+# ggsave('report/figs/toy.png', width = 5, height = 1.9)
+
+# Methodological Illustration ----------------
+
+# method_illustration()
+# ggsave('report/figs/method.png', width = 3, height = 3)
+
+
+
+cities <- list.files('data/cities')
+
+
+
+if(!dir.exists('throughput/grids')){
+	dir.create('throughput/grids')
+}
+
+
+
+# MAIN COMPUTATION ---------------
+
+
+# Cache logic
+
+if(file.exists('throughput/info_cache.csv')){
+	cache <- read_csv('throughput/info_cache.csv')
+	cities <-  cities[!is.element(cities, cache$city)]
+}else{
+	cache <- data.frame(
+		city = character(),
+		area = numeric(),
+		population = numeric(),
+		density = numeric(),
+		weighted_density = numeric(),
+		n_per_hex = numeric(),
+		H_X = numeric(),
+		H_Y = numeric(),
+		I_XY = numeric(),
+		local.H_Y = numeric(),
+		J = numeric(),
+		weighted_J = numeric())
+}
+
+
+for(city in cities){
+	races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
+	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
+	tracts <- tracts[tracts@data$total > 0,]
+	tracts@data$area <- tracts@data$ALAND / 1000^2	
+	tracts <- tracts[tracts@data$total / tracts@data$area > 50,] # at least 50 people per km^2
+	
+	radius = 1/sqrt(85 * 111) * 1/2 # (roughly 1 km after lat-lon conversion)
+	xx = spsample(tracts, type="hexagonal", cellsize=radius)
[…]

### 2016-06-16 4_cluster.R

@@ -1,0 +1,47 @@
+library(dplyr, quietly = TRUE)
+library(compx, quietly = TRUE)
+library(rgdal, quietly = TRUE)
+library(rgeos, quietly = TRUE)
+library(ggplot2, quietly = TRUE)
+library(Matrix, quietly = TRUE)
+library(maptools, quietly = TRUE)
+library(profvis, quietly = TRUE)
+library(readr, quietly = TRUE)
+library(ggrepel, quietly = TRUE)
+library(caTools, quietly = TRUE)
+
+columns <- c('White', 'Black', 'Asian', 'Hispanic', 'Other')
+
+# --------------------
+
+summary <- read_csv('throughput/info_cache.csv')
+cities <- summary$city
+
+if(file.exists('throughput/loss_curves.csv')){
+	cache <- read_csv('throughput/loss_curves.csv')
+	cities <- cities[!is.element(cities, cache$city)]
+} else {
+	cache <- data.frame('I_XY' = numeric(), 'nclust' = numeric(), 'city' = character())
+}
+
+if(!dir.exists('throughput/clusterings')){
+	dir.create('throughput/clusterings')
+}
+
+
+for(city in cities){
+	print(city)
+	x <- numeric()
+	spdf <- readOGR(dsn = paste0('data/cities/', city), layer = 'geo')
+	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
+	a <- info_clust(spdf, columns)
+	saveRDS(a, file = paste0('throughput/clusterings/',city))
+	for(i in 1:nrow(spdf@data)){
+		spdf@data$cluster <- factor(cutree(a, i))
+		x[i] <- mutual_info(data = spdf@data[,c(columns, 'cluster')], group_col = 'cluster')
+	}
+
+	added <- data.frame('I_XY' = x, 'nclust' = 1:nrow(spdf), 'city' = city)
+	cache <- rbind(cache, added)
+	write_csv(cache,'throughput/loss_curves.csv')
+}

### 2016-06-16 analysis.R
housecleaning

@@ -1,199 +1,0 @@
-library(sp)
-library(compx)
-library(acs)
-library(plyr)
-library(dplyr)
-library(readr)
-library(ggmap)
-library(tidyr)
-library(magrittr)
-library(ggrepel)
-library(grid)
-library(rgdal)
-
-setwd("~/projects/spatial_complexity/tech_report")
-
-
-# Toy cities ----------
-
-# checkerboard_illustration()
-# ggsave('report/figs/toy.png', width = 5, height = 1.9)
-
-# Methodological Illustration ----------------
-
-# method_illustration()
-# ggsave('report/figs/method.png', width = 3, height = 3)
-
-
-
-
-# MAIN COMPUTATION ---------------
-
-# Prep city list 
-
-# cities <- city_list('msas.csv')
-# density_list <- read_csv('density_list.csv') %>% filter(density > 200)
-# cities <- cities %>% filter(city == 'New York City')
-
-
-# Cache logic
-cities <- list.files('cities') %>%
-	sort(decreasing = T)
-
-if(file.exists('cities_cache.csv')){
-	cache <- read_csv('cities_cache.csv')
-	cities <-  cities[!is.element(cities, cache$city)]
-}else{
-	cache <- data.frame(
-		city = character(),
-		area = numeric(),
-		population = numeric(),
-		density = numeric(),
-		weighted_density = numeric(),
-		n_per_hex = numeric(),
-		H_X = numeric(),
-		H_Y = numeric(),
-		I_XY = numeric(),
-		local.H_Y = numeric(),
-		J = numeric(),
-		weighted_J = numeric())
-}
-
-
-for(city in cities){
-	races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
-	tracts <- readOGR(dsn = paste0('cities/',city), layer = 'geo', verbose = FALSE)
-	tracts <- tracts[tracts@data$total > 0,]
-	tracts@data$area <- tracts@data$ALAND / 1000^2		
-	# tracts <- tracts[tracts@data$area < 20,] # careful with this
-	
-	xx = spsample(tracts, type="hexagonal", cellsize=.01)
-	print(paste0(city, ': ',  nrow(tracts@data), ' tracts || ', length(xx), ' grid cells'))
-	xxpl = HexPoints2SpatialPolygons(xx)
-	grid_radius = sqrt(85 * 111) * .01
-	
-	
-	# Define information measures on the grid 
-	h <- function(i){
[…]

### 2016-06-16 get_data.R

@@ -1,86 +1,0 @@
-setwd("~/projects/spatial_complexity/tech_report")
-
-library(sp)
-library(compx)
-library(acs)
-library(plyr)
-library(dplyr)
-library(readr)
-library(ggmap)
-library(tidyr)
-library(magrittr)
-library(ggrepel)
-library(grid)
-library(rgdal)
-library(tigris)
-library(stringr)
-library(acs)
-
-
-ua <- read_table('ua_list_all.txt')
-ua$state = word(ua$NAME, -1)
-
-f <- function(state){
-	counties <- acs.fetch(geography=geo.make(state=state, county="*"), 
-				   endyear = 2014,
-				   table.number="B01003")
-	counties <- as.numeric(geography(counties)[[3]])
-	
-	race <- acs::acs.fetch(endyear = 2014,
-						   span = 5,
-						   geography = acs::geo.make(state = state, county = counties, tract = '*', block.group = '*'),
-						   table.number = 'B03002',
-						   col.names = "pretty")
-	
-	race <- cbind(data.frame(race@geography),
-				  data.frame(race@estimate)) %>%
-		tbl_df() %>%
-		mutate(GEOID = paste0(str_pad(state, 2, 'left', pad = '0'),
-							  str_pad(county, 3, 'left', pad = '0'),
-							  str_pad(tract, 6, 'left', pad = '0'),
-							  str_pad(blockgroup, 1, 'left', pad = '0')))
-	
-	race$Hispanic <- race[names(race)[!grepl('.Not.Hispanic.or.Latino.',names(race))]] %>%
-		select(-(`NAME`:`Hispanic.or.Latino.by.Race..Hispanic.or.Latino.`)) %>%
-		select(-GEOID) %>%
-		rowSums()
-	
-	others <- c('Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..American.Indian.and.Alaska.Native.alone',
-				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Native.Hawaiian.and.Other.Pacific.Islander.alone',
-				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Some.other.race.alone',
-				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Two.or.more.races.',
-				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Two.or.more.races..Two.races.including.Some.other.race',
[…]

### 2016-06-16 make_uas.R

@@ -1,69 +1,0 @@
-setwd("~/projects/spatial_complexity/tech_report")
-
-library(sp)
-library(compx)
-library(acs)
-library(plyr)
-library(dplyr)
-library(readr)
-library(ggmap)
-library(tidyr)
-library(magrittr)
-library(ggrepel)
-library(grid)
-library(rgdal)
-library(tigris)
-library(stringr)
-library(acs)
-
-
-
-ua_poly <- tigris::urban_areas(cb = TRUE)
-ua <- read_tsv('ua_list_all.txt')
-ua$state = word(ua$NAME, -1)
-
-cities <- ua %>% 
-	filter(POP > 1e6) 
-
-states <- cities$state %>% 
-	str_split('--') %>% 
-	do.call(c,.) %>% 
-	unique()
-
-cities <- cities$NAME
-
-shapes <- list()
-for(state in states){
-	shapes[state] <- readOGR(dsn = paste0('data/',state), layer = 'geo', verbose = FALSE)
-}
-
-
-for(city in cities){
-	print(city)
-	city_info <- ua %>%
-		filter(NAME == city) 
-	
-	city_poly <- ua_poly[ua_poly@data$NAME10 == city,]
-	
-	
-	city_components <- list()
-	for(state in str_split(city_info$state, '--')[[1]]){
-		state_sub <- shapes[[state]][city_poly,]
-		city_components[state] <-  spChFIDs(state_sub,
-											as.character(paste(state,rownames(as(state_sub,"data.frame")),
-															   sep="_")))
-	}
-	
-	city_polys <- do.call(rbind, city_components)
-
-	tryCatch(writeOGR(city_polys, paste0("cities/", city),'geo', driver = 'ESRI Shapefile', morphToESRI = TRUE),
-			 error = function(e) NULL)
-}
-
-
-
-
-
-
-
-

### 2016-06-23 5_figs.Rmd
minor adjustments due to compx refactor

@@
     toc_float: true
-    theme: flatly
+    theme: simplex
 ---
@@
 library(tidyr)
+library(maptools)
 
@@
 
-loss_curves <- read_csv('throughput/loss_curves.csv')
+loss_curves <- read_csv('throughput/loss_curves.csv') %>% 
+	group_by(city) %>% 
+	arrange(desc(I_XY)) %>% 
+	mutate(I_XY = max(I_XY) - I_XY,
+		   nclust = row_number())
+
 info_summary <- read_csv('throughput/info_cache.csv')
@@
 
-cluster_map <- function(city){
+cluster_map <- function(city, nclusters = 10){
 	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
@@
 
-	tracts@data$cluster <- cutree(clusters,10)
+	tracts@data$cluster <- cutree(clusters,nclusters)
+	overlay <- unionSpatialPolygons(tracts, IDs = tracts@data$cluster)
 
@@
 	to_plot <- fortify(tracts)
+	to_overlay <- fortify(overlay)
 
@@
 	to_plot %>% 
-		ggplot(aes(x = long, y = lat, group = id, fill = cluster)) +
-		geom_polygon() + 
+		ggplot() +
+		geom_polygon(aes(x = long, y = lat, group = group, fill = cluster)) + 
 		scale_fill_distiller(palette = 'Set3', guide = FALSE) + 
@@
 			  axis.title = element_blank(),
-			  panel.grid = element_blank())
+			  panel.grid = element_blank()) +
+		geom_polygon(aes(x = long, y= lat, group = group), data = to_overlay, fill = NA, size = .3, color = 'white')
 
@@
 
-cluster_map('Detroit')
+cluster_map('Chicago', nclusters = 10)
 
@@
 
-```{r}
+```{r fig.cap='Example clusterings for 4 major cities.'}
 
-make_cluster_groups <- function(city){
+make_cluster_groups <- function(city, nclusters = 10){
 	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
@@
 
-	tracts@data$cluster <- cutree(clusters,10)
+	tracts@data$cluster <- cutree(clusters,nclusters)
 
-	I = mutual_info(tracts@data[,c(columns, 'cluster')])
+	I <- mutual_info(tracts@data[,c(columns, 'cluster')],group_col = 'cluster')
 
@@
 		arrange(desc(White)) %>% 
-		mutate(cluster = factor(1:10, levels = 1:10)) %>%  
+		mutate(cluster = factor(1:nclusters, levels = 1:nclusters)) %>%  
 		gather(key = race, value = p, -cluster) %>% 
@@
 
-df <- rbind(make_cluster_groups('Detroit'), 
-	  make_cluster_groups('Boston'),
-	  make_cluster_groups('Dallas'),
-	  make_cluster_groups('Atlanta'))
+nclusters = 5
+df <- rbind(make_cluster_groups('Detroit', 3), 
+	  make_cluster_groups('Boston', 5),
+	  make_cluster_groups('Chicago', 4),
+	  make_cluster_groups('Atlanta', 2))
 
@@
 
-```{r}
-
-```
-
 ## Loss curves
@@
 	geom_path(aes(color = weighted_J)) +
-	geom_point(aes(color = weighted_J), size = .3) +
+	# geom_point(aes(color = weighted_J), size = .3) +
 	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +

### 2016-06-23 4_cluster.R

@@ -28,20 +28,16 @@ if(!dir.exists('throughput/clusterings')){
 	dir.create('throughput/clusterings')
 }
 
-
 for(city in cities){
 	print(city)
-	x <- numeric()
 	spdf <- readOGR(dsn = paste0('data/cities/', city), layer = 'geo')
 	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
-	a <- info_clust(spdf, columns)
+	df <- spdf@data[,columns]
+	row.names(df) <- 1:nrow(df)
+	constraint <- gRelate(spdf, byid = TRUE, pattern = '****1****')
+	a <- info_clust(df, constraint)
 	saveRDS(a, file = paste0('throughput/clusterings/',city))
-	for(i in 1:nrow(spdf@data)){
-		spdf@data$cluster <- factor(cutree(a, i))
-		x[i] <- mutual_info(data = spdf@data[,c(columns, 'cluster')], group_col = 'cluster')
-	}
-
-	added <- data.frame('I_XY' = x, 'nclust' = 1:nrow(spdf), 'city' = city)
+	added <- data.frame('I_XY' = c(0,a$height), 'nclust' = 1:nrow(spdf), 'city' = city)
 	cache <- rbind(cache, added)
 	write_csv(cache,'throughput/loss_curves.csv')
 }

### 2016-06-25 tex/intro.tex
city-based, moving to cca

@@
 It achieves this by mapping the physical concept of structure to the epistemic concept of predictability.
-A complex system has enough structure to be at least partially predictable, but enough disorder to make prediction challenging.
-In this paper, we apply information-theoretic tools to the measurement of \emph{spatial compositional complexity}, with a focus on the spatial variation of racial trends in American cities.
+A complex system has enough structure to be at least partially predictable, but enough variability to make prediction challenging.
+Information theory's formalization of the concept of predictability is therefore tailor-made for the study of complex phenomena.
 
+	In this essay, we apply information-theoretic tools to the measurement of \emph{spatial compositional complexity}.
+A data set is \emph{spatial} when it consists of a series of measurements on a set of points $\mathcal{X}$ endowed with a measure of distance; formally, $\mathcal{X}$ is a metric space.
+A data set is \emph{compositional} when its observations consist in a set of counts or proportions across a set of fixed subcategories of a whole.\footnote{Our use of this term differs slightly from that used in the field of Compositional Data Analysis (CDA), in which compositional data is always proportions and never frequencies \cite{Aitchison1982,Aitchison2002}.
+Because our focus is on the compositional structure of these phenomena, we take the term to be appropriate in this context as well.} The alphabet of categories might be $\mathcal{Y} = \{\text{Humanities, Social Sciences, Natural Sciences}\}$, in which case an observation might consist in a tabulation of undergraduates majoring in each of these fields.
+
+	Our focus is on data sets that are both spatial and compositional, especially as they arise in urban theory and urban planning.
+Our working example is the spatial variation of racial trends in American cities.
 Information-theoretic concepts have already found application in urban planning problems related to zoning and predicting population distributions \cite{Royal2014,Batty1974,Batty1976,Battya}.
@@
 Attractive features of information-theoretic measures for this purpose include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to multiple demographic phenomena, and the fact that Theil's (\cite{Theil1971}) original index satisfaction of many (though not all) of the invariance properties desirable for the measurement of segregation \cite{Reardon2002}.
-\nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015}.
+\nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015,Webber1979,Bivand2014b,Bivand2014a,Bivand2014}.
 
-	While closely-related to segregation, the concept of complexity has received less attention among quantitative sociologists.
-This is natural, since complexity is both less operational than segregation and challenging to define.
-On the other hand, as we demonstrate below, a spatial information-theoretic measure can be defined which measures the granularity of neighborhood structure, distinguishing cities with large, monolithic tracts of constant racial composition from those with many adjacent, smaller, neighborhoods of varying racial profiles.
-In doing so, it therefore captures one natural aspect of complexity in spatial compositional phenomena, and may also be of interest as a segregation measure as well.
-
-\subsection{Information Theory and Spatial Structure}
-	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two.
-Let $X$ be a spatial variable, such as a set of coordinates in space or a neighborhood name.
-Let $Y$ be a compositional variable that we aim to study, defined on an alphabet $\mathcal{Y}$.
-In the simple case in which $\mathcal{Y}  = \{\text{Black, White}\}$, Figure \ref{fig:toy} illustrates a range of possible joint distributions $p(X,Y)$.
-Comparing the completely undiverse city (a) and the spatially uniformly diverse (b) motivates the first fundamental information measure, the entropy:
-	\begin{equation}
-		H(Y) \triangleq \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
-	\end{equation}
-	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$.
-It thereby clearly distinguishes cities (a) and (b): since (a) is completely uniform, $H_a(Y) = 0$, while $H_b(Y) \neq 0$.
-On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
-To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
-	\begin{align}
-		D[p(Z)\|q(Z)] &\triangleq \sum_z p(z) \log \frac{p(z)}{q(z)} \\
-		I(X,Y) &\triangleq D[p(X,Y) \| p(X)p(Y)]
-	\end{align}
-	The divergence $D$ is (with some caveats) a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$.
-Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent.
-In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).
-In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence.
-
-	City (d) shares with city (c) the fact that residence and race are maximally dependent.
-However, city (d) embodies the ``checkerboard problem'': measures that use only the joint distribution $p(X,Y)$ without additionally considering the spatial information contained within $X$ will evaluate city (d) to be the same as city (c), despite their considerably different patterns of racial separation and potentially very different implications for planning and policy.
-A recent working paper \cite{Roberto2015} provides one highly operational approach to this problem, using road network topology and a weighting function that decays with distance to define a localized measure based on the Kullback-Leibler divergence.
-Here we pursue an alternative strategy, showing that a localization of the mutual information both measures spatial variation in race and corresponds to the estimation of a fundamental statistical property of the distribution $p(X,Y)$.
+	Because an operationally satisfying, mathematically unified concept of segregation remains elusive, we focus on the related concepts of spatial complexity and neighborhood structure.
+While these concepts may appear both harder to define and less useful in practical social science, we demonstrate in what follows that a mathematically unified approach can both give content to these concepts and shed light on some of the concerns more traditionally related to segregation studies.
 
@@
 	\centering
-		  \includegraphics[width=.7\linewidth]{figs/checkerboard.png}
-		  \begin{tabular}{c | c c c}
-			  City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
-			  \hline			
-			  (a) & 0.0 & 0.0 & 0.0\\
-			  (b) & 0.7 & 0.0 & 0.0\\
-			  (c) & 0.7 & 0.7 & 0.6\\
-			  (d) & 0.7 & 0.7 & 2.7\\
-			  \hline  
-			\end{tabular}
-		\caption{Information theory and the checkboard problem: model cities with various kinds of spatial diversity can be distinguished through progressively more subtle spatial information measures.}
-		\label{fig:toy}
+		\includegraphics[width=.4\textwidth]{external_figs/detroit_dot.png}
+		\includegraphics[width=.4\textwidth]{external_figs/philly_dot.png}
+		\caption{Dotmaps of racial composition for Detroit (top) and Philadelphia (bottom) by blockgroup according to the 2010 US Census.
+Racial Dotmap created by Dustin Cable and the Weldon Cooper Center at the University of Virginia.}
+		\label{fig:info_cross}
 	\end{figure}
+
+\subsection{Structure of the Essay}
+	In Section 2, we motivate and develop the two core components of our mathematical framework.
+The first of these is the mean local information $J(X,Y)$, a measure of the intrinsic spatial complexity of a compositional phenomenon.
+In the context of racial trends in cities, the mean local information measures the granularity of racial neighborhood structure.
+It is low in a city like Detroit, which consists of a few large, monolithic tracts of constant racial composition.
+It is high in a city like Philadelphia, in which many (spatially) smaller neighborhoods are knit together in an intricate patchwork.
+The second component of our mathematical framework is a simple algorithm for identifying clusters that are both spatially and compositionally coherent through agglomerative hierarchical clustering and greedy information maximization.
+In the context of race in cities, this algorithm may be viewed as an automated means to identify ``natural'' neighborhoods.
+
+	In Section 3, we apply these techniques to the analysis of spatial trends in race across U.S.
+cities.
+We show that the mean local information $J(X,Y)$ and the mutual information $I(X,Y)$ jointly supply simple and intuitive measures of spatial complexity for various cities.
+We also note an intriguing scaling relationship between spatial complexity and population density.
+We next evaluate our information-theoretic clustering method, and show that the neighborhoods it identifies can usefully shed light on traditional concerns of quantitative sociology, such as racial affinities and spatial concentration.
+Next, by defining a global evaluation measure on each clustering, we show that the mean local information $J(X,Y)$ does indeed measure the ``clusterability'' of a spatial compositional data set.
+
+	Section 4 is a discussion of our findings and prospects for future work.
+Finally, we include an Appendix supplying some mathematical details not present in the text, including a proof of our assertion that the mean local information is related to a fundamental statistical property of spatial compositional phenomena, and a thorough specification of algorithms and computations.

### 2016-06-25 tex/methods.tex
city-based, moving to cca

@@
+	
+\subsection{Information Theory and Spatial Structure}
+	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two.
+Let $X$ be a spatial variable, such as a set of coordinates in space or a neighborhood name.
+Let $Y$ be a compositional variable that we aim to study, defined on an alphabet $\mathcal{Y}$.
+In the simple case in which $\mathcal{Y}  = \{\text{Black, White}\}$, Figure \ref{fig:toy} illustrates a range of possible joint distributions $p(X,Y)$.
+Comparing the completely undiverse city (a) and the spatially uniformly diverse (b) motivates the first fundamental information measure, the entropy:
+	\begin{equation}
+		H(Y) \triangleq \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
+	\end{equation}
+	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$.
+It thereby clearly distinguishes cities (a) and (b): since (a) is completely uniform, $H_a(Y) = 0$, while $H_b(Y) \neq 0$.
+On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
+To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
+	\begin{align}
+		D[p(Z)\|q(Z)] &\triangleq \sum_z p(z) \log \frac{p(z)}{q(z)} \\
+		I(X,Y) &\triangleq D[p(X,Y) \| p(X)p(Y)] = \E_X[D[p(Y|X)\|p(Y)]]
+	\end{align}
+	The divergence $D$ is (with some caveats) a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$.
+Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent.
+In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).
+In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence.
+
+	City (d) shares with city (c) the fact that residence and race are maximally dependent.
+However, city (d) embodies the ``checkerboard problem'': measures that use only the joint distribution $p(X,Y)$ without additionally considering the spatial information contained within $X$ will evaluate city (d) to be the same as city (c), despite their considerably different patterns of racial separation and potentially very different implications for planning and policy.
+A recent working paper \cite{Roberto2015} provides one highly operational approach to this problem, using road network topology and a weighting function that decays with distance to define a localized measure based on the Kullback-Leibler divergence.
+Here we pursue an alternative strategy, showing that a localization of the mutual information both measures spatial variation in race and corresponds to the estimation of a fundamental statistical property of the distribution $p(X,Y)$.
+
+	\begin{figure}
+		\centering
+		  \includegraphics[width=.7\linewidth]{figs/checkerboard.png}
+		  \begin{tabular}{c | c c c}
+			  City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
+			  \hline			
+			  (a) & 0.0 & 0.0 & 0.0\\
+			  (b) & 0.7 & 0.0 & 0.0\\
+			  (c) & 0.7 & 0.7 & 0.6\\
+			  (d) & 0.7 & 0.7 & 2.7\\
+			  \hline  
+			\end{tabular}
+		\caption{Information theory and the checkboard problem: model cities with various kinds of spatial diversity can be distinguished through progressively more subtle spatial information measures.}
+		\label{fig:toy}
+		\end{figure}
+
 \subsection{Measuring Socio-Spatial Complexity}
-		To motivate our methods, we first consider an idealized problem in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is intuitively the map on which we work.
+		To motivate our methods, we first consider an idealized scenario in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is interpretable the ``map'' on which we work.
 
@@
 		\begin{equation}
-			I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r(x_0))]|X \in B_r(x_0)]
+			I_r(x_0) \triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r(x_0))]|X \in B_r(x_0)] \label{eq:local}
 		\end{equation}
@@
 It is therefore a natural measure of local complexity, aligned in approach with the global mutual information but designed to detect pattern in local variations.
-$I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$, corresponding to a complete lack of local compositional complexity.
+As expected, $I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$; that is, if within $B_r(x_0)$ the field $p(y|x)$ resembles toy city (b) in Figure \ref{fig:toy}.
 
-		% \begin{figure}
-		% 	\centering
-		% 	\includegraphics[width=.5\textwidth]{figs/method.png}
-		% 	\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
+		\begin{figure}
+			\centering
+			\includegraphics[width=.5\textwidth]{figs/method_illustration.png}
+			\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
 We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density.
 }
-		% 	\label{fig:method}
-		% \end{figure}
+			\label{fig:method}
+		\end{figure}
 
-		In this definition, $r$ may be thought of as the spatial resolution at which we conduct analysis, and $I_r(x_0)$ is highly dependent on $r$.
+		In the definition \eqref{eq:local}, $r$ may be thought of as the spatial resolution at which we conduct analysis, and $I_r(x_0)$ is highly dependent on $r$.
 However, it is possible to show that $I_r(x_0)$ is related to a fundamental statistical property of the probability field $p(y|x)$ that is resolution-independent.
@@
 		\begin{equation}
-			J_Y(x) \triangleq \E\left[ (\nabla_x \log p(y|x))(\nabla_x \log p(y|x))^T \right]\;.
+			J_Y(x) \triangleq \E_Y\left[ (\nabla_x \log p(Y|x))(\nabla_x \log p(Y|x))^T \right]\;.
 		\end{equation}
@@
 		\end{enumerate}
-		This greedy procedure defines a form of agglomerative clustering distinguished by two characteristics: its spatial constraints and its pursuit of minimal information loss at each step.
+		This greedy procedure defines a form of agglomerative hierarchical clustering distinguished by two characteristics: its spatial constraints and its pursuit of minimal information loss at each step.
 As a greedy algorith, it possesses no guarantees for optimal solutions, but in practice its performance leads to intuitive neighborhoods.
-
-	\subsection{Complexity and Neighborhood Identification}

### 2016-06-25 tex/results.tex
city-based, moving to cca

@@
+\subsection{Information Measures}
+
 	We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S.
@@
 		\end{figure}
+
+\subsection{Complexity and Neighborhood Identification}
+
+	Figure \ref{fig:cluster_map} shows an illustrative clustering into ten regions according to equation \eqref{eq:info_dist} for the city of Chicago.
+
+		\begin{figure}
+			\centering
+			\includegraphics[width=\textwidth]{figs/example_cluster_map.png}
+			\caption{The regions determined by our hierarchical clustering algorithm reflect high-level racial trends across space.
+The clusters on the right sharply segment predominantly Black, Hispanic, and White neighborhoods in Central Chicago.}
+			\label{fig:cluster_map}
+		\end{figure}
+
+		We expect that spatially complex cities with high $J(X,Y)$ would require more complicated models in order to convey similar amounts of demographic information.
+Testing this expectation requires a quality measure defined on cluster models for each city.
+The natural measure of loss for a clustering is the mutual information $I(C,Y)$ between cluster and racial labels.
+Cities vary sharply in the level of cluster detail needed to convey equal amounts of information about race, as shown in figure \ref{fig:clusters}.
+Each clustering has approximately equal $I(C,Y)$.
+In sharply segregated Atlanta, just two clusters--one predominantly white, the other predominantly black--suffice to convey more information than do five in Boston, where the clusters are substantially more nuanced.
+Cluster 1 is predominantly white, cluster 4 majority Hispanic, and cluster 5 overwhelmingly black.
+Cluster 2 is again majority white, but is unique in that its minority citizens are almost all Hispanic, while cluster 3 is approximately equally black and Hispanic.
+
+		These clusterings can also reveal phenomena related to traditional concepts in segregation studies, such as affinity and concentration.
+In Figure \ref{fig:clusters}, across all cities, Asians are much more likely to be found in predominantly white clusters than they are in predominantly black or Hispanic ones, indicating significant spatial integration between these two groups.
+We can observe the phenomenon of concentration -- in which marginalized groups tend to be more highly concentrated in their regions than majority groups are in theirs -- in Atlanta and Detroit.
+In Atlanta, whites make up 61\% of cluster 1, while blacks make up 83\% of cluster 2.
+Similarly in Detroit, whites make up 70\% of cluster 1, while blacks make up a full 86\% of cluster 2.
+
+		\begin{figure}
+			\centering
+			\includegraphics[width=\textwidth]{figs/example_clusters.png}
+			\caption{Example clusters for different cities, and the associated mutual information contained in each.
+The number in each box reflects the percentage of the population in the labeled racial group and cluster.
+}
+			\label{fig:clusters}
+		\end{figure}
+
+		To develop a global evaluation of the hierarchical clustering of a city according to the method defined by \eqref{eq:info_dist}, we borrow the methodology of the ``area under the curve'' (AUC) used extensively in statistical learning.
+By plotting the information $I(C,Y)$ against the number of clusters $n$, we obtain a curve reflecting how the information evolves at varying scales of aggregation.
+Figure \ref{fig:AUC} illustrates these curves for Boston and Detroit, with $N$ plotted on a logarithmic scale.
+
+		\begin{figure}
+			\centering
+			\includegraphics[width=\textwidth]{figs/AUC_illustration.png}
+			\caption{Evaluation of clustering using an Area Under the Curve (AUC) metric.
+The AUC is the fraction of the bounding box lying under the information gain curve.
+Boston's AUC is smaller than Detroit's, showing that Boston's spatial complexity (measured by $J(X,Y)$) requires more complex models than Detroit's.}
+			\label{fig:AUC}
+		\end{figure}
+
+		Figure \ref{fig:AUC} also shows the bounding rectangle, whose top right corner is defined by the number N of tracts in the census data set and the full mutual information $I(X,Y)$.
+The AUC is defined as the ratio of the shaded area in Figure \ref{fig:AUC} to the bounding rectangle.
+An AUC of 1 indicates that just one region carries full information; this can only occur in a city with no spatial variation, such as city (b) in \ref{fig:toy}.
+A larger AUC indicates that sample models with few regions capture more information about spatial variation in the city.
+In Detroit, there exists a clear dividing line between a predominantly white region on the west and a predominantly black region in the east.
+A two-cluster model therefore captures much of the information in the city, giving a high AUC of 0.74.
+In Boston, on the other hand, no such clear divide exists, and more complex models are necessary to capture similar amounts of information.
+Thus, Boston has a lower AUC of 0.67.
+
+		\begin{figure}
+			\centering
+			\includegraphics[width=\textwidth]{figs/info_performance_adjed.png}
+			\input{figs/adjusted_regression.txt}
+
+			\caption{Relationship between information measures $I(X,Y)$, $J(X,Y)$, and the clustering AUC.
+The AUC has been geographically adjusted to account for cities like New York, whose Census blockgroups form eight disconnected components, which would mildly distort results without adjustment.
+Table produced using \cite{Marek2015}}.
+			\label{fig:info_and_clusters}
+		\end{figure}
+
+		Based on our discussion so far, we would expect that the AUC is related to the information measure $J(X,Y)$.
+Figure \ref{fig:info_and_clusters} confirms this expectation.
+Overall, two factors explain much about the ``clusterability'' of a city as measured by the AUC.
+The most important factor is the overall mutual information $I(X,Y)$.
+This reflects a simple intuition: when there is little spatial variation, there is not much to cluster.
+The second factor is the mean local information $J(X,Y)$, reflecting that spatial complexity makes clustering harder.
+A simple linear regression makes these insights precise.
+As predictors of the AUC, both $I(X,Y)$ and $J(X,Y)$ are highly significant, and the coefficient of $J(X,Y)$ is negative.
+The two measures jointly account for an adjusted 72\% of the variation in AUC.

### 2016-06-25 cca_trial.R

@@ -1,0 +1,51 @@
+library(CCA)
+library(rgdal)
+library(rgeos)
+library(leaflet)
+library(ggplot2)
+
+MA <- readOGR(dsn = 'data/states/NM', layer = 'geo', verbose = FALSE)
+MA <- MA[MA$total > 0,]
+MA <- MA[MA$total / MA$ALAND > .0000,] # density cutoff
+centroids <- MA %>% gCentroid(byid = TRUE) 
+MA@data <- cbind(MA@data,centroids@coords) 
+
+pop.list <- cca(MA@data[,c('x', 'y')],s=.02)
+
+MA@data <- MA@data %>% left_join(pop.list$cluster, by = c('x' = 'long', y = 'lat'))
+
+counts <- MA@data %>% 
+	group_by(cluster_id) %>% 
+	summarise(pop = sum(total),
+			  n = n()) %>% 
+	arrange(desc(pop)) %>% 
+	filter(pop > 5e5)
+
+MA@data <- MA@data %>% left_join(counts)
+
+MA <- MA[!is.na(MA@data$pop),] 
+
+# pop.list$cluster %>% 
+# 	ggplot() + 
+# 	aes(x = long, y = lat, color = cluster_id) + 
+# 	geom_point()
+# 
+# 
+# 
+# MA@data <- cbind(MA@data, lag(pop.list$cluster))
+
+colors <- MA@data$cluster_id
+pal <- colorNumeric(
+	palette = 'Set3',
+	domain = colors
+)
+
+map <-leaflet() %>%
+	addProviderTiles("CartoDB.Positron") %>%
+	addPolygons(data = MA, 
+				fillColor = ~pal(colors), 
+				color = "#b2aeae", # you need to use hex colors
+				fillOpacity = 0.7, 
+				weight = 1, 
+				smoothFactor = 0.2)
+map

### 2016-06-27 1_get_states.R
automated city generation with osc

@@ -1,7 +1,7 @@
 library(sp, quietly = TRUE)
 library(compx, quietly = TRUE)
 library(acs, quietly = TRUE)
-library(plyr, quietly = TRUE)
+# library(plyr, quietly = TRUE)
 library(dplyr, quietly = TRUE)
 library(readr, quietly = TRUE)
 library(ggmap, quietly = TRUE)
@@ -17,9 +17,8 @@ library(acs, quietly = TRUE)
 
 # get list of states
 
-cities <- read_csv('assumptions/cities.csv')
-cities$state = word(cities$state, -1)
-
+states <- read_csv('states.csv')
+states <- states$Abbreviation
 
 if(!dir.exists('data/states')){
 	dir.create('data/states')
@@ -78,11 +77,9 @@ f <- function(state){
 }
 
 
-
-
 # Get all the states
-cities <- c(unique(toupper(cities$state[nchar(cities$state) == 2])), 'DC')
-for(state in cities){
+
+for(state in states){
 	if(!is.element(state,list.files('data/states'))){
 		print(state)
 		tryCatch(f(state), error = function(e) NULL)

### 2016-06-27 2_make_cities.R

@@ -1,6 +1,9 @@
+
+clustering_distance <- .02
+population_threshold <- 5e5
+
 library(sp, quietly = TRUE)
 library(compx, quietly = TRUE)
-library(plyr, quietly = TRUE)
 library(dplyr, quietly = TRUE)
 library(readr, quietly = TRUE)
 library(tidyr, quietly = TRUE)
@@ -8,49 +11,68 @@ library(magrittr, quietly = TRUE)
 library(grid, quietly = TRUE)
 library(rgdal, quietly = TRUE)
 library(stringr, quietly = TRUE)
+library(rgeos)
+library(osc)
 
+states <- list.files('data/states')
 
-city_df <- read_csv('assumptions/cities.csv')
-cities <- unique(city_df$name)
-cities <- cities[!is.element(cities, list.dirs('cities', full.names = FALSE))]
-
-states <- city_df$state %>% 
-	str_split('--') %>% 
-	do.call(c,.) %>% 
-	unique()
-
-shapes <- list()
+state_polys <- list()
 for(state in states){
 	print(paste('Loading', state))
-	shapes[state] <- readOGR(dsn = paste0('data/states/',state), layer = 'geo', verbose = FALSE)
+	state_poly <- readOGR(dsn = paste0('data/states/',state), layer = 'geo', verbose = FALSE)
+	state_poly <- state_poly[state_poly$total > 0,]
+	state_poly@data$state <- state
+	
+	state_polys[state] <- spChFIDs(state_poly,as.character(paste(state, rownames(as(state_poly,"data.frame")))))
 }
+state_polys <- do.call(rbind, state_polys)
+centroids <- state_polys %>% gCentroid(byid = TRUE) 
+state_polys@data <- cbind(state_polys@data,centroids@coords) 
+
+clusters <- cca(state_polys@data[,c('x', 'y')],clustering_distance) 
+
+state_polys@data <- state_polys@data %>% left_join(clusters$cluster, by = c('x' = 'long', y = 'lat'))
+
+counts <- state_polys@data %>% 
+	group_by(cluster_id) %>% 
+	summarise(population = sum(total)) %>% 
+	arrange(desc(population)) %>% 
+	filter(population > population_threshold) 
+
+state_polys@data <- state_polys@data %>% left_join(counts)
+
+state_polys <- state_polys[!is.na(state_polys@data$population),] 
+
[…]

### 2016-06-27 3_local_info.R

@@ -24,7 +24,7 @@ library(rgdal, quietly = TRUE)
 
 
 cities <- list.files('data/cities')
-
+cities <- cities[length(cities):1]
 
 
 if(!dir.exists('throughput/grids')){
@@ -57,52 +57,53 @@ if(file.exists('throughput/info_cache.csv')){
 		weighted_J = numeric())
 }
 
+resolution <- 1
 
 for(city in cities){
 	races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
 	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
 	tracts <- tracts[tracts@data$total > 0,]
-	tracts@data$area <- tracts@data$ALAND / 1000^2	
+	tracts@data$area <- tracts@data$ALAND / 1000^2
 	tracts <- tracts[tracts@data$total / tracts@data$area > 50,] # at least 50 people per km^2
-	
-	radius = 1/sqrt(85 * 111) * 1/2 # (roughly 1 km after lat-lon conversion)
+
+	radius = 1/sqrt(85 * 111) * resolution # (roughly 1 km after lat-lon conversion)
 	xx = spsample(tracts, type="hexagonal", cellsize=radius)
 	print(paste0(city, ': ',  nrow(tracts@data), ' tracts || ', length(xx), ' grid cells'))
 	xxpl = HexPoints2SpatialPolygons(xx)
-	
-	# Define information measures on the grid 
+
+	# Define information measures on the grid
 	h <- function(i){
 		window <- tracts[xxpl[i,],]@data[,c(races, 'total', 'area')]
 		window <- window / (window$area) # total becomes density
-		c(mean(window$total), 
-		  4 * mutual_info(window[,races]) / (1/2)^2,  
-		  H(colSums(window[,races]) /sum(window[,races])), 
+		c(mean(window$total),
+		  4 * mutual_info(window[,races]) / resolution^2,
+		  H(colSums(window[,races]) /sum(window[,races])),
 		  nrow(window)) # returns estimated density and mutual info
 	}
-			
+
 	# Compute measures on the grid and collect as df
 	df <- 1:length(xxpl) %>%
 		as.matrix() %>%
-		apply(MARGIN = 1, FUN = h) %>% 
-		t %>% 
+		apply(MARGIN = 1, FUN = h) %>%
+		t %>%
 		as.data.frame()
[…]

### 2016-06-27 4_cluster.R

@@ -14,8 +14,9 @@ columns <- c('White', 'Black', 'Asian', 'Hispanic', 'Other')
 
 # --------------------
 
-summary <- read_csv('throughput/info_cache.csv')
-cities <- summary$city
+cities <- list.files('data/cities/')
+cities <- cities[length(cities):1]
+
 
 if(file.exists('throughput/loss_curves.csv')){
 	cache <- read_csv('throughput/loss_curves.csv')
@@ -34,6 +35,7 @@ for(city in cities){
 	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
 	df <- spdf@data[,columns]
 	row.names(df) <- 1:nrow(df)
+	print('Computing adjacency matrix')
 	constraint <- gRelate(spdf, byid = TRUE, pattern = '****1****')
 	a <- info_clust(df, constraint)
 	saveRDS(a, file = paste0('throughput/clusterings/',city))

### 2016-06-27 5_figs.Rmd
figs for automated city generation

@@
 library(maptools)
+library(grid)
+library(stargazer)
+library(magrittr)
 
@@
 ## Methodological illustration
-```{r, warning = FALSE, message = FALSE, eval = FALSE}
+```{r, warning = FALSE, message = FALSE}
 columns <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
-method_illustration <- function(city){
+method_illustration <- function(city, vmax){
 	hexgrid <- readOGR(dsn = paste0('throughput/grids/',city), layer = 'geo', verbose = FALSE)
@@
 	bbox <- c(min(hexgrid$long), min(hexgrid$lat), max(hexgrid$long), max(hexgrid$lat))
+
+	# squarify
+	max_range <- max(bbox[3] - bbox[1], bbox[4] - bbox[2])
+
+	long_mid <- (bbox[3] + bbox[1]) / 2
+	lat_mid  <- (bbox[2] + bbox[4]) / 2
+
+	bbox[1] <- long_mid - max_range / 2
+	bbox[3] <- long_mid + max_range / 2
+	bbox[2] <- lat_mid  - max_range / 2
+	bbox[4] <- lat_mid  + max_range / 2
+
 	map <- get_map(location = bbox, maptype = 'terrain-background')
@@
 		geom_polygon(data = hexgrid, aes(x = long, y = lat, group = group, fill = info), alpha = .8) +
-		scale_fill_continuous(low = 'white', high = 'steelblue', name= expression(J[Y](x)), limits = c(0, log(length(columns)) / .5^2)) +
+		scale_fill_continuous(low = 'white', high = 'steelblue', name= expression(J[Y](x)), limits = c(0, vmax)) +
 		theme(axis.ticks = element_blank(),
@@
 			  legend.text = element_text(size = rel(.1))) +
-		ggtitle(paste0('Local Mutual Information in ', city, ': Mean = ', round(weighted.mean(df$info, df$density), 2)))
+		ggtitle(paste0('Local Information in ', city, '\nJ(X,Y) = ', round(weighted.mean(df$info, df$density), 2)))
 	p
@@
 
-grid.arrange(method_illustration('Philadelphia'),
-			 method_illustration('Detroit'),
-			 ncol = 2)
+first <- method_illustration('Orlando, FL', vmax = 1.5)
+second <- method_illustration('Philadelphia, PA', vmax = 1.5)
 
-ggsave('figs/method.png')
+gA <- ggplotGrob(first)
+gB <- ggplotGrob(second)
+g <- gridExtra::cbind.gtable(gA, gB, size = "max")
+
+grid.draw(g)
+
+ggsave('figs/method_illustration.png', g, width = 8, height = 5)
+
 ```
@@
 
-cluster_map('Chicago', nclusters = 10)
+cluster_map('Chicago, IL', nclusters = 10)
 
@@
 		gather(key = race, value = n, -cluster) %>% 
-		# group_by(cluster) %>% 
 		mutate(n = n / sum(n)) %>% 
+		group_by(cluster) %>% 
+		filter(sum(n) > .01) %>% 
 		ungroup() %>% 
@@
 		arrange(desc(White)) %>% 
-		mutate(cluster = factor(1:nclusters, levels = 1:nclusters)) %>%  
+		mutate(cluster = factor(1:nrow(.), levels = 1:nrow(.))) %>%  
 		gather(key = race, value = p, -cluster) %>% 
@@
 nclusters = 5
-df <- rbind(make_cluster_groups('Detroit', 3), 
-	  make_cluster_groups('Boston', 5),
-	  make_cluster_groups('Chicago', 4),
-	  make_cluster_groups('Atlanta', 2))
+df <- rbind(make_cluster_groups('Detroit, MI', 3), 
+	  make_cluster_groups('Boston, MA', 5),
+	  make_cluster_groups('Chicago, IL', 4),
+	  make_cluster_groups('Atlanta, GA', 2))
 
@@
 
+```{r}
+end_points <- loss_curves %>%
+	group_by(city) %>%
+	filter(nclust == max(nclust))
+
+loss_curves %>% 
+	group_by(city) %>% 
+	mutate(I_max = max(I_XY)) %>% 
+	ggplot() +
+	aes(x = nclust, y = I_XY, group = city) +
+	geom_path(aes(color = I_max)) +
+	# geom_point(aes(color = weighted_J), size = .3) +
+	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
+	# scale_y_continuous(trans = 'log10') +
+	geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) +
+	theme_minimal() +
+	annotation_logticks(size = .1, sides = 'b') +
+	ggtitle('Loss curves for spatially-constrained information-theoretic clustering')
+
+```
+
 ## Normalized loss curves:
@@
 
-cities <- c('Atlanta','Baltimore')
+cities <- c('New Orleans, L','Boston, MA')
 
-summary <- loss_curves %>%
+compute_AUC <- function(loss_curves, adjust_thresh = 0){
+	loss_curves %>% 
+		group_by(city) %>%
+		mutate(exclude = (nclust != 1) & (cummax(I_XY - lag(I_XY,default = 0)) < adjust_thresh) & (nclust < 10)) %>% 
+		filter(!exclude) %>%
+		mutate(nclust = row_number()) %>% 
 		mutate(dx = log(nclust / (nclust - 1))) %>%
@@
 		mutate(AUC = area / corner) %>%
-	arrange(desc(AUC)) %>%
-	left_join(info_summary[,c('city', 'weighted_J', 'I_XY')]) %>%
-	filter(I_XY > 0)
+		arrange(desc(AUC)) 
+}
+
+summary <- compute_AUC(loss_curves, adjust_thresh = 0)
+
+summary <- summary %>% left_join(cache, by = c('city' = 'city'))
+
+# %>%
+# 	left_join(info_summary[,c('city', 'weighted_J', 'I_XY')]) %>%
+# 	filter(I_XY > 0)
 
@@
 rects <- sub_curves %>% group_by(city) %>% 
-	summarise(xmin = 1, ymin = 0, xmax = max(nclust), ymax = max(I_XY),weighted_J = max(weighted_J))
+	summarise(xmin = 1, ymin = 0, xmax = max(nclust), ymax = max(I_XY))
 
@@
 	geom_path(aes(x = nclust, y = I_XY, group = city, color = city)) +
-	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
+	scale_x_continuous(trans = 'log10', limit = c(1, NA)) +
+	scale_y_continuous(limit = c(0,.7)) +
 	theme_minimal() +
-	annotation_logticks(size = .1, sides = 'b') + 
+	annotation_logticks(size = .1, sides = 'b', scaled = T) + 
 	geom_rect(data = rects, aes(xmin = xmin,
@@
 	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
-	geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(weighted_J,2))), data = sub_summary) +
+	# geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(weighted_J,2))), data = sub_summary) +
 	facet_wrap(~city) + 
@@
 	ylab('I(X,Y)') + 
-	ggtitle('Evaluation of clustering')
+	ggtitle('AUC as Clustering Evaluation')
 
@@
 
-adjusted_summary <- loss_curves %>%
-	group_by(city) %>%
-	filter(I_XY - lag(I_XY) > 0.005 | row_number() > 10) %>%
-	mutate(nclust = row_number()) %>%
-	arrange(nclust) %>%
-	summarise(adj_corner = max(I_XY) * log(max(nclust)),
-			  adj_area = trapz(nclust, I_XY / nclust)) %>%
-	mutate(adj_AUC = adj_area / adj_corner) %>%
-	arrange(desc(adj_AUC))
+# dippy point that doesn't seem to be true here: if you don't have any excluded points on your curve, your AUC should be the same.
+
+adjusted_summary <- compute_AUC(loss_curves, adjust_thresh = .005) %>%
+	select(city, adj_AUC = AUC)
 
@@
 	left_join(adjusted_summary)
+```
+
+```{r}
+
+summary %>% 
+	ggplot() +
+	aes(x = AUC, y = adj_AUC) +
+	geom_point() + 
+	geom_text_repel(aes(label = city)) + 
+	geom_abline(slope = 1, intercept = 0)
+
 ```
@@
 
+highlight <- summary %>% 
+	filter(I_XY >= 0)
+
 summary %>%
@@
 	geom_smooth(color = 'grey', se = F) +
-	geom_text_repel(aes(label = city)) + 
+	geom_text_repel(aes(label = city), data = highlight) + 
 	theme_minimal() + 
 	ggtitle('Clustering performance and mutual information') +
-	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
+	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,.6)) +
 	xlab('I(X,Y)') + 
@@
 ```{r}
+highlight <- summary %>% filter(I_XY > .4)
 
@@
 	geom_smooth(color = 'grey', se = F) +
-	geom_text_repel(aes(label = city)) + 
+	geom_text_repel(aes(label = city), data = highlight) + 
 	theme_minimal() + 
-	ggtitle('Clustering performance and mutual information (geographically adjusted)') +
+	ggtitle('Clustering performance and mutual information') +
 	xlab('I(X,Y)') + 
-	ylab('AUC (adjusted)') + 
-	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2))
+	ylab('AUC (geographically adjusted)') + 
+	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(.3,.6), name = "J(X,Y)", guide = guide_colorbar(direction = 'horizontal')) +
+	theme(legend.justification=c(1.5,0),
+			  legend.position=c(1,0))
 
@@
 summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
-		   I_XY = I_XY / max(I_XY),
-		   weighted_J = weighted_J / max(weighted_J)) %>%
+		   `I(X,Y)` = I_XY / max(I_XY),
+		   `J(X,Y)` = weighted_J / max(weighted_J)) %>%
 	filter(I_XY > 0) %>%
-	lm(AUC ~ I_XY + weighted_J, data = .) %>%
-	summary()
+	lm(AUC ~ `I(X,Y)` + `J(X,Y)`, data = .) %>%
+	stargazer(flip = T, 
+			  # table.layout = '#tn', 
+			  style = 'all2',
+			  ci = T,covariate.labels = c('$I(X,Y)$','$J(X,Y)$', 'Intercept'),
+			  table.layout = '-=-!=!dtas',
+			  float = F) %>% 
+	write(file = 'figs/adjusted_regression.txt')
 ```

### 2016-06-27 tex/intro.tex
changes to intro

@@
 Information-theoretic concepts have already found application in urban planning problems related to zoning and predicting population distributions \cite{Royal2014,Batty1974,Batty1976,Battya}.
-Substantial recent work has addressed the measurement of difference and disparity in cities through an information-theoretic lens \cite{Theil1971,Bettencourt2015,Roberto2015a,Roberto2015}.
+Substantial recent work has addressed the measurement of difference and disparity in cities through an information-theoretic lens \cite{Theil1971,Bettencourt2015,Roberto2015a,Roberto2015}, though an operationally satisfying, mathematically unified concept of segregation remains eluse \cite{Press2009a}.
 Attractive features of information-theoretic measures for this purpose include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to multiple demographic phenomena, and the fact that Theil's (\cite{Theil1971}) original index satisfaction of many (though not all) of the invariance properties desirable for the measurement of segregation \cite{Reardon2002}.
-\nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015,Webber1979,Bivand2014b,Bivand2014a,Bivand2014}.
 
-	Because an operationally satisfying, mathematically unified concept of segregation remains elusive, we focus on the related concepts of spatial complexity and neighborhood structure.
+	Rather than considering the question of sociological segregation directly, we focus on the related concepts of spatial complexity and neighborhood structure.
 While these concepts may appear both harder to define and less useful in practical social science, we demonstrate in what follows that a mathematically unified approach can both give content to these concepts and shed light on some of the concerns more traditionally related to segregation studies.
+
+	Core ideas to knit together here: neighborhoods, information theory, segregation.
+
+	\nocite{Wong2004,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015,Webber1979,Bivand2014b,Bivand2014a,Bivand2014}.
+
+	Neighborhood effects are both important for human behavior and difficult to quantify due to a variety of methodological issues \cite{Sampson2002}.
+
+	Study of neighborhood effects often defined purely in terms of census data sets, reflecting data limitations rather than theoretically-principled approaches. \cite{Dietz2002}
 

### 2016-06-27 tex/methods.tex
changes to intro

@@
 \subsection{Information Theory and Spatial Structure}
-	Two of the most fundamental objects of information theory are the entropy of a single random and the mutual information of two.
+	We begin by developing two fundamental measures of information theory, the entropy of one random variable and the mutual information of two.
 Let $X$ be a spatial variable, such as a set of coordinates in space or a neighborhood name.
@@
 	\begin{equation}
-		H(Y) \triangleq \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
+		H(Y) \triangleq - \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
 	\end{equation}
 	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$.
-It thereby clearly distinguishes cities (a) and (b): since (a) is completely uniform, $H_a(Y) = 0$, while $H_b(Y) \neq 0$.
-On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
+$H(Y) = 0$ when $Y$ always takes a single value, such as in city (a).
+On the other hand, $H(Y)$ achieves its maximum of $H(Y) = \log \abs{\mathcal{Y}}$ when $Y$ is uniformly distributed on $\mathcal{Y}$, as in city (b).
+The entropy $H(Y)$ is this sufficient to distinguish the presence of global diversity from its absence.
+On the other hand, On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
 To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
@@
 	\end{align}
-	The divergence $D$ is (with some caveats) a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$.
+	Though not a true metric, the divergence $D$ is interpretable a measure of distance in the space of probability distributions, and so $I(X,Y)$ may be interpreted as the distance between the true joint distribution $p(X,Y)$ and the product of marginals $p(X)p(Y)$.
 Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent.
 In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).
-In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence.
+In city (c), on the other hand $X$ and $Y$ are completely dependent: if you know where someone lives, you know their race with 100\% confidence, and $I(X,Y)$ achieves its maximum.
 
-	City (d) shares with city (c) the fact that residence and race are maximally dependent.
+	City (d) shares with city (c) the fact that race is completely determined by residence.
 However, city (d) embodies the ``checkerboard problem'': measures that use only the joint distribution $p(X,Y)$ without additionally considering the spatial information contained within $X$ will evaluate city (d) to be the same as city (c), despite their considerably different patterns of racial separation and potentially very different implications for planning and policy.
@@
 \subsection{Measuring Socio-Spatial Complexity}
-		To motivate our methods, we first consider an idealized scenario in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where $M$ is interpretable the ``map'' on which we work.
+		We will now derive a measure that solves the ``checkerboard problem'' by distinguishing the toy cities (c) and (d).
+To motivate our methods, we  consider an idealized scenario in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where the metric space $M$ is interpretable the ``map'' on which we work.
 
@@
 		\end{equation}
-		Intuitively, $I_r(x_0)$ measures how much the probability field $p(y|x)$ varies with $x$ in a small neighborhood of $x_0$.
-It is therefore a natural measure of local complexity, aligned in approach with the global mutual information but designed to detect pattern in local variations.
+		Intuitively, $I_r(x_0)$ measures how much knowing the ``location'' $x$ adds to our information about the category $Y$, or, equivalently, how much the probability field $p(y|x)$ varies with $x$ in a small neighborhood of $x_0$.
+It is therefore a natural measure of local complexity, aligned in approach with the global mutual information but designed to detect local variation.
 As expected, $I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$; that is, if within $B_r(x_0)$ the field $p(y|x)$ resembles toy city (b) in Figure \ref{fig:toy}.
@@
 		\begin{equation}
-			\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace} (J_Y(x))\;, \label{approx}
+			\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace } J_Y(x)\;, \label{approx}
 		\end{equation}
@@
 		\end{equation}
-		A more formal statement and proof of \eqref{approx} are provided in Appendix 2.
+		A formal statement and proof of \eqref{approx} are provided in Appendix 2.
 The Fisher information $J_Y$ is a fundamental quantity in statistics and information theory.
@@
 		\end{equation} 
+		As demonstrated in Figure \ref{fig:toy}, $J(X,Y)$ distinguishes cities (c) and (d), thereby addressing the ``checkerboard problem'' head on.
 We propose the aggregate quantity $J(X,Y)$ as a third  measure--alongside the entropy $H(Y)$ and mutual information $I(X,Y)$-- as a tool for the information-theoretic structure of spatial compositional complexity.
+
+		We note that, since 
+		\begin{align}
+			\text{trace } J_Y(x) &=  \sum_i \E_Y\left[ \left(\frac{1}{p(Y|x)} \frac{\partial p(Y|x)}{\partial x_i}\right)\right]^2\;
+		\end{align}
+		may be viewed as a weighted norm the gradient $\nabla_x p(Y|x)$, the quantity 
+		\begin{align}
+			J(X,Y) &= \E_X[\text{trace } J_Y(X)] \\
+			&= \int_M \text{trace } J_Y(X) d\prob_X
+		\end{align}
+		may be viewed as a cousin to total variation measures often encountered in analysis.
+
+		The application of this methodology to a discrete data set is conceptually simple.
+For a given set of tracts, overlay an evenly spaced grid of radius $r$, and measure the mutual information $I_r(x)$ in each grid cell.
+Then, when data and grid resolutions are sufficiently high, $4I_r(x) / r^2$ approximates the quantity $\text{trace }J_Y(x)$, which can then be aggregated across the data set.
+We provide a more formal statement of this computation in the appendix.
+
 	\subsection{Identifying Natural Neighborhoods}
-		Suppose that we have a collection of tracts labeled with the random variable $X$, and suppose that each tract has an additional label $C$ denoting the \emph{cluster} in which it lies.
-Then, it is possible to show that
+		We now develop a simple information-theoretic approach to the identification of natural neighborhoods.
+For concreteness, suppose that we have a collection of Census tracts whose labels are given by the random variable $X$, and suppose that each tract has an additional label $C$ denoting the \emph{cluster} in which it lies.
+Then, the following relationship holds between mutual informations:
 		\begin{equation}
-			I(X,Y) = I(Y,C) + I(X,Y|C) \label{eq:information_decomp}
+			I(X,Y) = I(C,Y) + I(X,Y|C) \label{eq:information_decomp}
 		\end{equation}
@@
 
-		Finally, equation \eqref{eq:information_decomp} motivates a natural, information-based scheme for identifying natural neighborhoods.
-If all I had access to were the cluster labels $C$ and not the locations $X$, then my information would be $I(Y,C)$.
+		Finally, equation \eqref{eq:information_decomp} motivates a simple, information-based scheme for identifying natural neighborhoods.
+If all I had access to were the cluster labels $C$ and not the locations $X$, then the information I had about $Y$ would be $I(Y,C)$.
 A ``good'' clustering therefore maximizes the between-cluster information $I(Y,C)$, which entails minimizing the within-cluster information $I(X,Y|C)$.
 Solving this problem exactly is a challenging discrete optimization problem, and may not be computationally tractable.
-We can, however, construct a greedy algorithm with good performance.
-Suppose that we face the problem of choosing a collection $I = \{1,2,\ldots\}$ of locations to cluster together.
+We can, however, construct a greedy algorithm which leads to intuitive results.
+Suppose that we face the problem of choosing a pair of tracts $\{i*,j*\}$ to cluster together.
 The reduction in information associated with aggregating the locations $I$ into a single cluster is
 		\begin{align}
-			d(I) \triangleq  \sum_{i \in I} p(X = i)D[p(Y|X = i)\| p(Y|X \in I)] - p(X\in I)D[p(Y|X\in I) \| p(Y)] \label{eq:info_dist}
+			d(i,j) &\triangleq  \sum_{k \in \{i,j\}} p(X = k)D[p(Y|X = k)\| p(Y|X \in \{i,j\})] \\
+			&- p(X\in \{i,j\})D[p(Y|X\in \{i,j\}) \| p(Y)] \label{eq:info_dist}
 		\end{align}
 		where $p(Y) = \sum_{x} p(x,Y)$ is the global marginal distribution.
-If we choose $I = \{i,j\}$, then equation \eqref{eq:info_dist} defines a natural information distance between locations $i$ and $j$.
+the first term is the information associated with the two separate tracts, while the second is the information associated with a merged tract.
+Equation \eqref{eq:info_dist} defines a natural information distance between locations $i$ and $j$.
 Like the KL divergence, this distance is strictly nonnegative; unlike the KL divergence, this distance is symmetric.
 Importantly, we can therefore use the distance $d(i,j)$ as a dissimilarity measure for the purposes of clustering.
-Our greedy procedure is simple: at each stage, the algorithm aggregates together two locations $i$ and $j$ such that
-		\begin{enumerate}
-			\item Locations $i$ and $j$ are adjacent in space. This requirement ensures that the results may be naturally understood as neighborhoods, rather than disparate tracts with similar demographics.  
-			\item Among all adjacent pairs, $d(i,j)$ is the smallest possible. 
-		\end{enumerate}
-		This greedy procedure defines a form of agglomerative hierarchical clustering distinguished by two characteristics: its spatial constraints and its pursuit of minimal information loss at each step.
+Our greedy procedure is simple: at each iteration, determine
+		\begin{equation}
+			(i^*,j^*) = \argmin_{i \text{neighbors} j} d(i,j),
+		\end{equation}
+		and then combine $i^*$ amd $j^*$ into a single tract, repeating until only one cluster remains.
+This procedure defines a form of agglomerative hierarchical clustering distinguished by two characteristics: its spatial constraints and its pursuit of minimal information loss at each step.
 As a greedy algorith, it possesses no guarantees for optimal solutions, but in practice its performance leads to intuitive neighborhoods.

### 2016-06-29 1_get_states.R
back to manually-constructed cities

@@ -1,7 +1,7 @@
 library(sp, quietly = TRUE)
 library(compx, quietly = TRUE)
 library(acs, quietly = TRUE)
-# library(plyr, quietly = TRUE)
+library(plyr, quietly = TRUE)
 library(dplyr, quietly = TRUE)
 library(readr, quietly = TRUE)
 library(ggmap, quietly = TRUE)
@@ -17,8 +17,9 @@ library(acs, quietly = TRUE)
 
 # get list of states
 
-states <- read_csv('states.csv')
-states <- states$Abbreviation
+cities <- read_csv('assumptions/cities.csv')
+cities$state = word(cities$state, -1)
+
 
 if(!dir.exists('data/states')){
 	dir.create('data/states')
@@ -77,9 +78,11 @@ f <- function(state){
 }
 
 
-# Get all the states
 
-for(state in states){
+
+# Get all the states
+cities <- c(unique(toupper(cities$state[nchar(cities$state) == 2])), 'DC')
+for(state in cities){
 	if(!is.element(state,list.files('data/states'))){
 		print(state)
 		tryCatch(f(state), error = function(e) NULL)

### 2016-06-29 2_make_cities.R

@@ -1,9 +1,6 @@
-
-clustering_distance <- .02
-population_threshold <- 5e5
-
 library(sp, quietly = TRUE)
 library(compx, quietly = TRUE)
+library(plyr, quietly = TRUE)
 library(dplyr, quietly = TRUE)
 library(readr, quietly = TRUE)
 library(tidyr, quietly = TRUE)
@@ -11,68 +8,49 @@ library(magrittr, quietly = TRUE)
 library(grid, quietly = TRUE)
 library(rgdal, quietly = TRUE)
 library(stringr, quietly = TRUE)
-library(rgeos)
-library(osc)
 
-states <- list.files('data/states')
 
-state_polys <- list()
+city_df <- read_csv('assumptions/cities.csv')
+cities <- unique(city_df$name)
+cities <- cities[!is.element(cities, list.dirs('cities', full.names = FALSE))]
+
+states <- city_df$state %>% 
+	str_split('--') %>% 
+	do.call(c,.) %>% 
+	unique()
+
+shapes <- list()
 for(state in states){
 	print(paste('Loading', state))
-	state_poly <- readOGR(dsn = paste0('data/states/',state), layer = 'geo', verbose = FALSE)
-	state_poly <- state_poly[state_poly$total > 0,]
-	state_poly@data$state <- state
-	
-	state_polys[state] <- spChFIDs(state_poly,as.character(paste(state, rownames(as(state_poly,"data.frame")))))
+	shapes[state] <- readOGR(dsn = paste0('data/states/',state), layer = 'geo', verbose = FALSE)
 }
-state_polys <- do.call(rbind, state_polys)
-centroids <- state_polys %>% gCentroid(byid = TRUE) 
-state_polys@data <- cbind(state_polys@data,centroids@coords) 
-
-clusters <- cca(state_polys@data[,c('x', 'y')],clustering_distance) 
-
-state_polys@data <- state_polys@data %>% left_join(clusters$cluster, by = c('x' = 'long', y = 'lat'))
-
-counts <- state_polys@data %>% 
-	group_by(cluster_id) %>% 
-	summarise(population = sum(total)) %>% 
-	arrange(desc(population)) %>% 
-	filter(population > population_threshold) 
-
-state_polys@data <- state_polys@data %>% left_join(counts)
-
-state_polys <- state_polys[!is.na(state_polys@data$population),] 
-
[…]

### 2016-06-29 3_local_info.R

@@ -24,7 +24,7 @@ library(rgdal, quietly = TRUE)
 
 
 cities <- list.files('data/cities')
-cities <- cities[length(cities):1]
+
 
 
 if(!dir.exists('throughput/grids')){
@@ -57,53 +57,52 @@ if(file.exists('throughput/info_cache.csv')){
 		weighted_J = numeric())
 }
 
-resolution <- 1
 
 for(city in cities){
 	races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
 	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
 	tracts <- tracts[tracts@data$total > 0,]
-	tracts@data$area <- tracts@data$ALAND / 1000^2
+	tracts@data$area <- tracts@data$ALAND / 1000^2	
 	tracts <- tracts[tracts@data$total / tracts@data$area > 50,] # at least 50 people per km^2
-
-	radius = 1/sqrt(85 * 111) * resolution # (roughly 1 km after lat-lon conversion)
+	
+	radius = 1/sqrt(85 * 111) * 1/2 # (roughly 1 km after lat-lon conversion)
 	xx = spsample(tracts, type="hexagonal", cellsize=radius)
 	print(paste0(city, ': ',  nrow(tracts@data), ' tracts || ', length(xx), ' grid cells'))
 	xxpl = HexPoints2SpatialPolygons(xx)
-
-	# Define information measures on the grid
+	
+	# Define information measures on the grid 
 	h <- function(i){
 		window <- tracts[xxpl[i,],]@data[,c(races, 'total', 'area')]
 		window <- window / (window$area) # total becomes density
-		c(mean(window$total),
-		  4 * mutual_info(window[,races]) / resolution^2,
-		  H(colSums(window[,races]) /sum(window[,races])),
+		c(mean(window$total), 
+		  4 * mutual_info(window[,races]) / (1/2)^2,  
+		  H(colSums(window[,races]) /sum(window[,races])), 
 		  nrow(window)) # returns estimated density and mutual info
 	}
-
+			
 	# Compute measures on the grid and collect as df
 	df <- 1:length(xxpl) %>%
 		as.matrix() %>%
-		apply(MARGIN = 1, FUN = h) %>%
-		t %>%
+		apply(MARGIN = 1, FUN = h) %>% 
+		t %>% 
 		as.data.frame()
[…]

### 2016-06-29 4_cluster.R

@@ -14,9 +14,8 @@ columns <- c('White', 'Black', 'Asian', 'Hispanic', 'Other')
 
 # --------------------
 
-cities <- list.files('data/cities/')
-cities <- cities[length(cities):1]
-
+summary <- read_csv('throughput/info_cache.csv')
+cities <- summary$city
 
 if(file.exists('throughput/loss_curves.csv')){
 	cache <- read_csv('throughput/loss_curves.csv')
@@ -35,7 +34,6 @@ for(city in cities){
 	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
 	df <- spdf@data[,columns]
 	row.names(df) <- 1:nrow(df)
-	print('Computing adjacency matrix')
 	constraint <- gRelate(spdf, byid = TRUE, pattern = '****1****')
 	a <- info_clust(df, constraint)
 	saveRDS(a, file = paste0('throughput/clusterings/',city))

### 2016-06-29 2_make_cities.R
back to manually converted cities pt. 2

@@ -30,8 +30,6 @@ if(!dir.exists('data/cities')){
 }
 
 
-
-
 for(city in cities){
 	print(city)
 	city_info <- city_df %>%

### 2016-06-29 3_local_info.R

@@ -1,3 +1,5 @@
+resolution <- 1/2
+
 library(sp, quietly = TRUE)
 library(compx, quietly = TRUE)
 library(acs, quietly = TRUE)
@@ -65,7 +67,7 @@ for(city in cities){
 	tracts@data$area <- tracts@data$ALAND / 1000^2	
 	tracts <- tracts[tracts@data$total / tracts@data$area > 50,] # at least 50 people per km^2
 	
-	radius = 1/sqrt(85 * 111) * 1/2 # (roughly 1 km after lat-lon conversion)
+	radius = 1/sqrt(85 * 111) * resolution # (roughly 1 km after lat-lon conversion)
 	xx = spsample(tracts, type="hexagonal", cellsize=radius)
 	print(paste0(city, ': ',  nrow(tracts@data), ' tracts || ', length(xx), ' grid cells'))
 	xxpl = HexPoints2SpatialPolygons(xx)
@@ -75,7 +77,7 @@ for(city in cities){
 		window <- tracts[xxpl[i,],]@data[,c(races, 'total', 'area')]
 		window <- window / (window$area) # total becomes density
 		c(mean(window$total), 
-		  4 * mutual_info(window[,races]) / (1/2)^2,  
+		  4 * mutual_info(window[,races]) / resolution^2,  
 		  H(colSums(window[,races]) /sum(window[,races])), 
 		  nrow(window)) # returns estimated density and mutual info
 	}

### 2016-07-01 5_figs.Rmd
2nd paper draft

@@
 		geom_polygon(data = hexgrid, aes(x = long, y = lat, group = group, fill = info), alpha = .8) +
-		scale_fill_continuous(low = 'white', high = 'steelblue', name= expression(J[Y](x)), limits = c(0, vmax)) +
+		scale_fill_continuous(low = 'white', high = 'steelblue', name= expression(J[Y](x)), limits = c(0, vmax), breaks = c(0,2,4,6)) +
 		theme(axis.ticks = element_blank(),
@@
 
-first <- method_illustration('Orlando, FL', vmax = 1.5)
-second <- method_illustration('Philadelphia, PA', vmax = 1.5)
+first <- method_illustration('Detroit', vmax = 6)
+second <- method_illustration('Philadelphia', vmax = 6)
 
@@
 ggsave('figs/method_illustration.png', g, width = 8, height = 5)
+
+```
+
+```{r additive_decomposability}
+
+expand.grid(x = 1:8, y = 1:8) %>% 
+	mutate(`(e)` = .5,
+		   `(e)` = ifelse(x %in% 2:4 & y %in% 2:7, 0, `(e)`),
+		   `(e)` = ifelse(x %in% 5:7 & y %in% 2:7, 1, `(e)`),
+		   `(f)` = .5) %>%
+	gather(key = model, value = p, -x, -y) %>%
+	ggplot(aes(x = x, y = y)) +
+		theme_minimal() +
+		theme(axis.ticks = element_blank(),
+			  axis.text.x = element_blank(),
+			  axis.text.y = element_blank(),
+			  panel.background = element_rect(),
+			  panel.grid.major = element_line(size = 0),
+			  panel.grid.minor = element_line(size = 0),
+			  plot.margin=unit(c(0,0,0,0),"mm")) +
+		geom_tile(aes(fill = p)) +
+		scale_fill_continuous(low = 'white', high = 'black ', limits=c(0,1)) +
+		facet_grid(~model) +
+		xlab('') +
+		ylab('') +
+		guides(fill=FALSE)
+
+ggsave('figs/decomposability.png', width = 2.5, height = 1.5)
 
@@
 		ggplot() +
-		geom_polygon(aes(x = long, y = lat, group = group, fill = cluster)) + 
-		scale_fill_distiller(palette = 'Set3', guide = FALSE) + 
+		geom_polygon(aes(x = long, y = lat, group = group, fill = factor(cluster))) + 
+		scale_fill_brewer(palette = 'Set3', name = NULL, guide = guide_legend(nrow = 1), labels = LETTERS[1:nclusters]) + 
 		ggtitle(paste0('Example clustering for ', city)) + 
@@
 			  axis.title = element_blank(),
-			  panel.grid = element_blank()) +
-		geom_polygon(aes(x = long, y= lat, group = group), data = to_overlay, fill = NA, size = .3, color = 'white')
+			  panel.grid = element_blank(),
+			  legend.position = 'bottom') +
+		geom_polygon(aes(x = long, y= lat, group = group), data = to_overlay, fill = NA, size = .3, color = 'black')
 
@@
 
-cluster_map('Chicago, IL', nclusters = 10)
+cluster_map('Detroit', nclusters = 6)
 
@@
 		arrange(desc(White)) %>% 
-		mutate(cluster = factor(1:nrow(.), levels = 1:nrow(.))) %>%  
+		mutate(cluster = factor(LETTERS[1:nrow(.)])) %>%  
 		gather(key = race, value = p, -cluster) %>% 
@@
 nclusters = 5
-df <- rbind(make_cluster_groups('Detroit, MI', 3), 
-	  make_cluster_groups('Boston, MA', 5),
-	  make_cluster_groups('Chicago, IL', 4),
-	  make_cluster_groups('Atlanta, GA', 2))
+df <- rbind(make_cluster_groups('Detroit', 3), 
+	  make_cluster_groups('Boston', 5),
+	  make_cluster_groups('Chicago', 4),
+	  make_cluster_groups('Atlanta', 2))
 
@@
 ggsave('figs/example_clusters.png')
+
+```
+
+Follow up on Detroit by area (percent of area occupied by each cluster.)
+
+```{r}
+city = 'Detroit'
+nclusters = 3
+
+tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
+tracts <- tracts[sum(tracts@data[,columns]) != 0,]
+clusters <- readRDS(paste0('throughput/clusterings/', city))
+
+tracts@data$cluster <- cutree(clusters,nclusters)
+overlay <- unionSpatialPolygons(tracts, IDs = tracts@data$cluster)
+plot(overlay)
+
+sapply(slot(overlay, "polygons"), slot, "area")  %>% simplex_normalize()
 
@@
 
-cities <- c('New Orleans, L','Boston, MA')
+cities <- c('Atlanta','Boston')
 
@@
 	ggtitle('Clustering performance and mutual information') +
-	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,.6)) +
+	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
 	xlab('I(X,Y)') + 
@@
 	ylab('AUC (geographically adjusted)') + 
-	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(.3,.6), name = "J(X,Y)", guide = guide_colorbar(direction = 'horizontal')) +
+	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,.2), name = "J(X,Y)", guide = guide_colorbar(direction = 'horizontal')) +
 	theme(legend.justification=c(1.5,0),
@@
 
-### Loess-based detrending
-
-```{r}
-
-normalized <- summary %>% mutate(AUC = AUC / max(AUC),
-		   I_XY = I_XY / max(I_XY),
-		   weighted_J = weighted_J / max(weighted_J)) %>%
-	filter(I_XY > 0)
-
-fit <- normalized %>%
-	loess(AUC ~ I_XY + weighted_J, data = .)
-
-normalized$pred <- predict(fit)
-normalized$resid <- normalized$pred - normalized$adj_AUC
-
-normalized  %>%
-	filter(I_XY > 0) %>%
-	lm(resid ~ weighted_J, data = .) %>%
-	summary()
-
-```
-
 # Session info

### 2016-07-01 tex/intro.tex
2nd paper draft

--- 2016-06-25 tex/intro.tex
@@
-	Information theory provides one natural approach toward thinking about complex systems.
-It achieves this by mapping the physical concept of structure to the epistemic concept of predictability.
+\subsection{Neighborhoods, Segregation, and Information}
+
+	The neighborhood in which an individual lives has profound impact on that individual's mental and physical well-being \cite{Ludwig2012, Ludwig2013}.
+Urban planners may aim to revitalize neighborhoods, and an entire cottage industry in sociology and epidemiology has grown around the idea of ``neighborhood effects'' (see, for example, \cite{DiezRoux2001, Sampson2002} for overviews of older work).
+Despite the persistent import of neighborhoods in social sciences, little consensus exists on how to define or demarcate neighborhoods in a non-arbitrary way.
+Many studies define ``neighborhoods'' by the boundaries supplied in their data sets, such as those determined by the U.S.
+Census \cite{Dietz2002}.
+Such an approach follows the limitations of the data provided, rather than a theoretically-principled concept of neighborhood.
+However, recent progress in non-arbitrary demarcations of urban form serve as proofs of concept for the theoretically-principled measurement of heterogeneity in urban form \cite{Rozenfeld2008,Rozenfeld2011}.
+
+	We take up the question of neighborhoods in the context of segregation studies in sociology, and will argue that the measurement of segregation is intricately linked with the identification of spatial demographic structure.
+Many indices have been proposed in the last half-century to measure various aspects of segregation; see \cite{Massey1988, Reardon2002} for useful synthesizing overviews.
+Many such approaches have also taken Census-defined boundaries as constitutive of neighborhoods, but more recent work has focused ``spatialized'' measures that are independent of neighborhood demarcations\cite{Reardon2004, Roberto2015a, Roberto2015,Wong2004, Wong1999, Lee2008}.
+Other work has quantified the tendency of different social groups to cluster together \cite{Louf2015}.
+
+	The approach we develop is grounded in information measures of spatial demographic structure.
+Information theory maps the physical concept of structure to the epistemic concept of predictability.
 A complex system has enough structure to be at least partially predictable, but enough variability to make prediction challenging.
-Information theory's formalization of the concept of predictability is therefore tailor-made for the study of complex phenomena.
+Information-theoretic concepts have already been introduced in the study of cities, finding application in urban planning problems, in predicting population distributions, and in quantifying difference between neighborhoods \cite{Royal2014,Batty1974,Batty1976,Battya,Bettencourt2015,Webber1979}.
+Attractive features of information measures include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to a wide variety of phenomena, and their conceptual linkage of cities with the methods of statistical physics.
+In the context of the measurement of urban demographic structure, information measures provide answers to questions such as:
+	\begin{enumerate}
+		\item Supposing you choose an individual randomly from a city, how accurately could I guess their race? 
+		\item Supposing you then told me where that person lived within the city, how much would this new information increase my accuracy? 
+	\end{enumerate}
+	The following considerations connect these questions to the measurement of diversity and structure.
+If I can guess a random individual's race with 100\% accuracy given no information, then it must be the case that the city is monoracial; increased diversity reduces my ability to guess.
+If knowing where a person lives dramatically increased my ability to guess, then there must be substantial structural dependence of demographics on space, a prerequisite of segregation.
 
-	In this essay, we apply information-theoretic tools to the measurement of \emph{spatial compositional complexity}.
-A data set is \emph{spatial} when it consists of a series of measurements on a set of points $\mathcal{X}$ endowed with a measure of distance; formally, $\mathcal{X}$ is a metric space.
-A data set is \emph{compositional} when its observations consist in a set of counts or proportions across a set of fixed subcategories of a whole.\footnote{Our use of this term differs slightly from that used in the field of Compositional Data Analysis (CDA), in which compositional data is always proportions and never frequencies \cite{Aitchison1982,Aitchison2002}.
-Because our focus is on the compositional structure of these phenomena, we take the term to be appropriate in this context as well.} The alphabet of categories might be $\mathcal{Y} = \{\text{Humanities, Social Sciences, Natural Sciences}\}$, in which case an observation might consist in a tabulation of undergraduates majoring in each of these fields.
-
-	Our focus is on data sets that are both spatial and compositional, especially as they arise in urban theory and urban planning.
-Our working example is the spatial variation of racial trends in American cities.
-Information-theoretic concepts have already found application in urban planning problems related to zoning and predicting population distributions \cite{Royal2014,Batty1974,Batty1976,Battya}.
-Substantial recent work has addressed the measurement of difference and disparity in cities through an information-theoretic lens \cite{Theil1971,Bettencourt2015,Roberto2015a,Roberto2015}.
-Attractive features of information-theoretic measures for this purpose include their deep relationship to statistical inference \cite{Cover1991,Csiszzr2004}, their generalizability to multiple demographic phenomena, and the fact that Theil's (\cite{Theil1971}) original index satisfaction of many (though not all) of the invariance properties desirable for the measurement of segregation \cite{Reardon2002}.
-\nocite{Sampson2002,Dietz2002,Wong2004,Keeling1999,Anas1997,Ioannides2004a,Wong1999,Press2009a,Holloway2012,Lee2008,Louf2015,Webber1979,Bivand2014b,Bivand2014a,Bivand2014}.
-
-	Because an operationally satisfying, mathematically unified concept of segregation remains elusive, we focus on the related concepts of spatial complexity and neighborhood structure.
-While these concepts may appear both harder to define and less useful in practical social science, we demonstrate in what follows that a mathematically unified approach can both give content to these concepts and shed light on some of the concerns more traditionally related to segregation studies.
-
-	\begin{figure}
-	\centering
-		\includegraphics[width=.4\textwidth]{external_figs/detroit_dot.png}
-		\includegraphics[width=.4\textwidth]{external_figs/philly_dot.png}
-		\caption{Dotmaps of racial composition for Detroit (top) and Philadelphia (bottom) by blockgroup according to the 2010 US Census.
-Racial Dotmap created by Dustin Cable and the Weldon Cooper Center at the University of Virginia.}
-		\label{fig:info_cross}
-	\end{figure}
+	We make two primary contributions.
+The first is to the field of quantitative segregation studies in sociology.
+We formulate the \emph{mean local information} $J(X,Y)$ as a measure of spatiosocial structure.
+We then show that this novel measure, in concert with the mutual information $I(X,Y)$, comprise a unified, spatially-aware methodology for the measurement of diversity and segregation.
+These metrics possess many of the traditionally desirable properties of sociological indices, and can be easily computed from open data such as that provided by the U.S.
+Census.
+Our second contribution is to provide a simple, theoretically-motivated, and computationally-tractable method for agglomerating spatio-social data into ``natural neighborhoods'' across which demographic trends are approximately constant.
+This method has multiple applications.
+Sociologists can use this agglomeration procedure to study racially coherent neighborhoods, rather than the partially-arbitrary administrative boundaries such as those produced by the U.S.
+Census.
+Computational urban planners can use the method for \emph{dimension-reduction}, in which a large data set is made more computationally tractable while preserving spatio-social relationships of interest.
+Finally, urban physicists may be interested in the scaling behavior of the mutual information across different levels of aggregation in analogy with coarse-graining in statistical mechanics \cite{Bettencourt2015}.
+The hierarchical character of agglomerative clustering defines a non-arbitrary way to perform this aggregation.
 

### 2016-07-01 tex/methods.tex
2nd paper draft

@@
 	
-\subsection{Information Theory and Spatial Structure}
+\subsection{Information and the Checkerboard}
 	We begin by developing two fundamental measures of information theory, the entropy of one random variable and the mutual information of two.
@@
 	\begin{equation}
-		H(Y) \triangleq - \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
+		H(Y) \triangleq -\E_Y[\log p(Y)] = - \sum_{y \in \mathcal{Y}} p(y) \log p(y)\;.
 	\end{equation}
 	The entropy measures how evenly the global marginal distribution $p(Y)$ is distributed over the alphabet $\mathcal{Y}$.
-$H(Y) = 0$ when $Y$ always takes a single value, such as in city (a).
+Epistemically, $H(Y)$ measures the difficulty in guessing the random variable $Y$, given no further information.
+$H(Y) = 0$ when $Y$ always takes a single value, such as in city (a) -- perfect guessing is possible.
 On the other hand, $H(Y)$ achieves its maximum of $H(Y) = \log \abs{\mathcal{Y}}$ when $Y$ is uniformly distributed on $\mathcal{Y}$, as in city (b).
 The entropy $H(Y)$ is this sufficient to distinguish the presence of global diversity from its absence.
-On the other hand, On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
+On the other hand, the entropy $H$ is unable to distinguish between the spatial uniformity of (b) and the spatially variability of (c).
 To distinguish these two cities we may use the mutual information, which is defined in terms of the Kullback-Leibler divergence $D$:
@@
 Since the latter expresses statistical independence between $X$ and $Y$, $I(X,Y)$ measures the extent to which $X$ and $Y$ are dependent.
+Epistemically, $I(X,Y)$ measures the extent to which knowledge of $X$increases possible accuracy in guessing $Y$.
 In city (b), $X$ and $Y$ are independent: knowing $X$ (where an individual lives) conveys no information about that $Y$ (that individual's race).
@@
 \subsection{Measuring Socio-Spatial Complexity}
+
 	We will now derive a measure that solves the ``checkerboard problem'' by distinguishing the toy cities (c) and (d).
-To motivate our methods, we  consider an idealized scenario in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where the metric space $M$ is interpretable the ``map'' on which we work.
+To motivate our methods, we  consider an idealized scenario in which we have a differentiable field $p(y|x)$ of observed probability distributions for each $x \in M \subset \R^n$, where the metric space $M$ is interpretable as the ``map'' on which we work.
 
@@
 As expected, $I_r(x_0) = 0$ if and only if $p(y|x)$ is constant for each $y$ in the ball $B_r(x_0)$; that is, if within $B_r(x_0)$ the field $p(y|x)$ resembles toy city (b) in Figure \ref{fig:toy}.
-
-		\begin{figure}
-			\centering
-			\includegraphics[width=.5\textwidth]{figs/method_illustration.png}
-			\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
-We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density.
-}
-			\label{fig:method}
-		\end{figure}
-
 In the definition \eqref{eq:local}, $r$ may be thought of as the spatial resolution at which we conduct analysis, and $I_r(x_0)$ is highly dependent on $r$.
@@
 	\begin{equation}
-			\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace } J_Y(x)\;, \label{approx}
+		\frac{I_r(x)}{r^2} \cong \frac{1}{4} \text{trace } J_Y(x)\;, \label{eq:approx}
 	\end{equation}
@@
 	\end{equation}
-		A formal statement and proof of \eqref{approx} are provided in Appendix 2.
+	A formal statement and proof of \eqref{eq:approx} are provided in Appendix 2.
 The Fisher information $J_Y$ is a fundamental quantity in statistics and information theory.
@@
 	\begin{equation}
-		J(X,Y) \triangleq \E_X[\text{trace }J_Y(X)]
+	J(X,Y) \triangleq \E_X[\text{trace }J_Y(X)] \label{eq:def_J}
 	\end{equation} 
@@
 	\end{align}
-		may be viewed as a weighted norm the gradient $\nabla_x p(Y|x)$, the quantity 
+	may be viewed as a weighted norm of the gradient $\nabla_x p(Y|x)$, the quantity 
 	\begin{align}
@@
 
-	\subsection{Identifying Natural Neighborhoods}
-		We now develop a simple information-theoretic approach to the identification of natural neighborhoods.
-For concreteness, suppose that we have a collection of Census tracts whose labels are given by the random variable $X$, and suppose that each tract has an additional label $C$ denoting the \emph{cluster} in which it lies.
-Then, the following relationship holds between mutual informations:
+\subsection{Information Measures and Segregation Studies}
+
+	Considerable attention has been paid to relating segregation indices to intuitive concepts of diversity and segregation.
+We note here that the mutual information $I(X,Y)$ and mean local information $J(X,Y)$ correspond closely to the two dimensions of segregation formulated in \cite{Reardon2002}.
+The authors of \cite{Reardon2002} describe \emph{evenness} as ``the extent to which groups are similarly distributed in residential space'' (page 126), and \emph{exposure} as ``the extent that members of one group encounter members of another group...in their local spatial environments.'' The mutual information $I(X,Y)$ may be viewed as a measure of (lack of) evenness, since to say that groups are similarly distributed in residential space is to say that knowing a spatial location conveys little about the race of the people who live there.
+Large $I(X,Y)$ reflects highly uneven distributions of demographic groups.
+Complementarily, the mean local information $J(X,Y)$may be viewed as a measure of spatial exposure \emph{for a fixed level of} $I(X,Y)$, as is illustrated by cities (c) and (d) in \ref{fig:toy}.
+Though distinct, these dimensions are not independent.
+To see this, note that the most thorough kind of exposure is achieved when $I(X,Y) = J(X,Y) = 0$, as in city (b).
+Since the concept of exposure only applies in a city with spatial differences, $J(X,Y)$ must be considered jointly with $I(X,Y)$ as a measure of spatial exposure.
+
+	There has also been much work showing the desirable properties of various segregation indicies.
+To give a brief sampling, indices should be invariant to changes in overall population size; they should decrease when populations ``even themselves out,'' and they should behave predictably under aggregation.
+Since we are presenting $I(X,Y)$ and $J(X,Y)$ as a suite of complementary information measures, we consider each in turn.
+The author of \cite{Roberto2015a} shows that the mutual information $I(X,Y)$ (which she calls the ``Divergence Index'') satisfies these and other desirable properties, generally as well or better than existing alternatives.
+We highlight one property of $I(X,Y)$ for special note, as this property will be central to our development of natural neighborhood identification.
+The principle of ``additive decomposability'' \cite{Reardon2002} stipulates that, when the data is grouped along either racial or spatial axes, a good segregation index should split into ``within group'' and ``between group'' components.
+In the context of the mutual information $I(X,Y)$, additive decomposability is simply the familiar chain rule of mutual information.
+For concreteness, let $C$ be a random variable giving the cluster label of location $X$; importantly, $C$ is completely determined by $X$.
+Then, the chain rule expresses additive decomposability as
 	\begin{equation}
-			I(X,Y) = I(C,Y) + I(X,Y|C) \label{eq:information_decomp}
+		I(X,Y) = I(C,Y) + I(X,Y|C)\;; \label{eq:information_decomp}
 	\end{equation}
-		Equation \eqref{eq:information_decomp} has a natural interpretation: the information I have about $Y$ if I know $X$ is equal to the  information I have if I know the cluster $C$, plus the amount of information I have if I subsequently learn the exact location $X$.
-Usefully, equation \eqref{eq:information_decomp} decomposes the system-wide mutual information $I(X,Y)$ into a \emph{between-cluster} component $I(Y,C)$ and an \emph{within-cluster} component $I(X,Y|C)$.
+	i.e.
+the information I have about $Y$ given that I know $X$ is equal to the information I have if I first learn the cluster $C$, plus the amount of additional information I gain if I subsequently learn the exact location $X$ as well.
+The first term is interpretable as the between-group information, while the second is interpretable as the within-group information.
 It is therefore comparable to various ``sum of squares'' decompositions that frequently appear in classical statistics.
 Indeed, it is possible to show that, when the differences between locations are small, the mutual information can be interpreted as a variance, in which case \eqref{eq:information_decomp} expresses the sum of squares composition directly.
+A similar version of the chain rule expresses additive decomposability for aggregation of racial groups rather than spatial locations.
 
-		Finally, equation \eqref{eq:information_decomp} motivates a simple, information-based scheme for identifying natural neighborhoods.
-If all I had access to were the cluster labels $C$ and not the locations $X$, then the information I had about $Y$ would be $I(Y,C)$.
+	What of $J(X,Y)$?
+Below, we provide a brief overview of the properties of $J(X,Y)$.
+$J(X,Y)$ possesses a variety of intuitive properties, and those that it does not possess either underscore the importance of reading it in conjunction with $I(X,Y)$ or do not apply to explicitly spatial measures.
+See \cite{Reardon2002,Reardon2004} for further discussion of these criteria.
+The mean local information $J(X,Y)$ satisfies:
+	\begin{description}
+		\item[Organizational Equivalence:] Both the theoretical definition \eqref{eq:def_J} or the procedure to compute it from tract data remain unchanged when a tract is subdivided into smaller tracts, each of which with identical demographic structure.
+		\item[Size and Density Invariance:] Since $J(X,Y)$ is completely determined by the marginal distribution $p(X,Y)$, it is invariant under changes in population density. 
+		\item[Additive Group Decomposability:] When demographic groups are aggregated into super-groups, the chain rule of mutual information applied to the demographic variable $Y$ provides an an additive decomposition of the form $J(X,Y) = J(X,G) + J(X,Y|G)$, which can be interpreted as the sum of a between-groups term and a within-groups term as needed.
+		\item[Scale Interpretability:] We have $J(X,Y) = 0$ if and only if $p(Y|X)$ is constant on each connect component of the metric space $M$.
+When only one connected component exists, this implies that every locale has the same demographic structure as the global environment.
+When multiple connected components exist (e.g.
+the city is divided by a river), demographics must be constant in space on each side of the division.
+$J(X,Y)$ does not satisfy the additional condition of achieving its maximum value when all locales are monoracial, but this point only emphasizes that $J(X,Y)$ and $I(X,Y)$ should be read jointly.
+When all locales are monoracial, $I(X,Y)$ achieves its maximum value and $J(X,Y) = 0$.
+		\item[Boundary Independent:] $J(X,Y)$ is defined in terms of a continuous underlying probability distribution $p(X,Y)$, rather than arbitrarily-defined tracts.
+In computational practice, $J(X,Y)$ is indeed sensitive to the boundaries supplied with the data; however, \eqref{eq:approx} guarantees that this sensitivity vanishes as the resolution grows sufficiently small.
+		\item[Exchanges:] Exchanges that tend to ``smooth out'' demographic distributions in space reduce both $I(X,Y)$ and $J(X,Y)$.  
+	\end{description}
+	The mean local information $J(X,Y)$ fails to one of the criteria enumerated in \cite{Reardon2002,Reardon2004}; that of \textbf{additive spatial decomposability}.
+This criterion states that ``If X spatial subareas are aggregated into Y larger spatial areas, then a segregation measure should be decomposable into a sum of within- and between-area components'' \cite{Reardon2004}, page 136.
+However, considering that we have imposed additional spatial structure on our model, this criterion does not appear to be motivated.
+Consider, for example, Figure \ref{fig:decomposability}.
+Aggregating the two central regions transforms (e) into (f), erasing all spatial variability.
+Any measure should therefore have a between-groups component equal to 0.
+On the other hand, the within-groups component can only consider the middle white/black boundary between the central two regions.
+The sum of the between-groups and within-groups components must therefore consist only in information included in the middle boundary.
+However, the spatial variability in city (e) is not exhausted by this middle boundary; there are a total of six other frontiers of racial difference that should be considered in any spatial segregation measure.
+We therefore content that additive decomposability is a desideratum of nonspatial measures (it is satisfied by $I(X,Y)$), but not of explicitly spatial ones.
+	\begin{figure}
+		\centering
+		\includegraphics[width=.5\textwidth]{figs/decomposability.png}
+		\caption{A problem with additive spatial decomposability: aggregating the inner two regions transforms (e) into (f), smoothing out all spatial variability.
+A ``between groups'' term must therefore vanish, while a ``within-groups'' term considers only the inner boundary; not the outer ones.}
+		\label{fig:decomposability}
+	\end{figure}
+	We have not exhausted the full complement of desiderata for segregation measures.
+However, many of the remaining ones -- such as composition invariance -- are either vague or controversial, and we will not consider further.
+
+\subsection{Identifying Natural Neighborhoods}
+
+	Equation \eqref{eq:information_decomp} motivates a simple scheme for identifying natural neighborhoods based on maximizing the between-group term $I(C,Y)$.
+If all I had access to were the cluster labels $C$ and not the locations $X$, then the information I had about $Y$would be $I(C,Y)$.
 A ``good'' clustering therefore maximizes the between-cluster information $I(Y,C)$, which entails minimizing the within-cluster information $I(X,Y|C)$.
 Solving this problem exactly is a challenging discrete optimization problem, and may not be computationally tractable.
-We can, however, construct a greedy algorithm which leads to intuitive results.
+We can, however, construct a greedy algorithm which leads to satisfactory results.
 Suppose that we face the problem of choosing a pair of tracts $\{i*,j*\}$ to cluster together.
@@
 Equation \eqref{eq:info_dist} defines a natural information distance between locations $i$ and $j$.
-Like the KL divergence, this distance is strictly nonnegative; unlike the KL divergence, this distance is symmetric.
+Like the KL divergence, this distance is strictly nonnegative; unlike the KL divergence, it is symmetric, and defines an axiomatic metric on the space of tracts.
 Importantly, we can therefore use the distance $d(i,j)$ as a dissimilarity measure for the purposes of clustering.
@@
 This procedure defines a form of agglomerative hierarchical clustering distinguished by two characteristics: its spatial constraints and its pursuit of minimal information loss at each step.
-As a greedy algorith, it possesses no guarantees for optimal solutions, but in practice its performance leads to intuitive neighborhoods.
+As a greedy algorith, it possesses no guarantees for optimal solutions, but in practice its performance leads to intuitive, racially-coherent regions.
+It thereby enables a study of spatial difference using non-arbitrarily-defined regions.

### 2016-07-01 tex/results.tex
2nd paper draft

@@
+\nocite{Bivand2014b,Bivand2014a,Bivand2014} % need to cite the census in here too
+
+\subsection{Data Used}
+	We assembled block-group level data from the 2010-2014 American Community Survey (ACS), conducted by the U.S.
+Census Bureau, on race and ethnicity for counties housing 51 large US cities.
+We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'.
+
 \subsection{Information Measures}
 
-	We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S.
-Census Bureau, on race and ethnicity for counties corresponding to 28 large US cities.
-At this early stage, the counties used were chosen to correspond to the most densely populated urban cores; in future developments we could standardize to conduct analyses for Metropolitan Statistical Areas.
-We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'.
 	 For each city, we computed the entropy $H(Y)$ and the mutual information $I(X,Y)$.
-To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 1km.
+To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 0.5km.
 We then computed the estimated mutual information within each grid cell, and averaged the results weighted by population population.
-A technical specification of this approach is provided in Appendix 1, and a complete summary of cities and their associated information measures is provided in Appendix 3.
+A technical specification of this approach is provided in Appendix 1, and an illustration of it in Figure \ref{fig:method}.
 
 	 \begin{figure}
-			\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
-			\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
-			\label{fig:info_cross}
+		\centering
+		\includegraphics[width=.8\textwidth]{figs/method_illustration.png}
+		\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston).
+We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density.
+}
+		\label{fig:method}
 	\end{figure}
+
 	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$.
@@
 These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities.
-
-	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density.
-In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities.
-A clear trend is evident, with $J(X,Y)$ growing linearly with the logarithm of density.
-We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix.
-One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own.
+	\begin{figure}
+		\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
+		\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
+		\label{fig:info_cross}
+	\end{figure}
 
@@
 
-\subsection{Complexity and Neighborhood Identification}
+	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density.
+In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities.
+It appears that $J(X,Y)$ grows linearly with the logarithm of density.
+We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix.
+One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own.
+It may be of interest for later studies to consider why this trend might arise through dynamical processes of urban formation.
 
-	Figure \ref{fig:cluster_map} shows an illustrative clustering into ten regions according to equation \eqref{eq:info_dist} for the city of Chicago.
+\subsection{Learning From Natural Neighborhoods}
+
+	Figure \ref{fig:cluster_map} shows an illustrative clustering of Wayne County (housing the city of Detroit) into six regions using the greedy information maximization procedure defined by equation \eqref{eq:info_dist}.
+The procedure cleanly divides the city into racially coherent zones.
+Cluster A is predominantly white and B predominantly black, demarcating the major racial divide in the city.
+Cluster C is a small transitional zone in which the two races overlap.
+Clusters D and E are the demographically distinct independent cities of Hamtramck and Grosse Pointe, respectively, while Cluster F is a Hispanic community known as Mexicantown.
+The existence of clusters like C reflects that not every racially coherent tract is interpretable as a ``neighborhood''; some may be more naturally thought as transitions between neighborhoods.
 
@@
 		\includegraphics[width=\textwidth]{figs/example_cluster_map.png}
-			\caption{The regions determined by our hierarchical clustering algorithm reflect high-level racial trends across space.
-The clusters on the right sharply segment predominantly Black, Hispanic, and White neighborhoods in Central Chicago.}
+		\caption{Clustering of Detroit into 6 neighborhoods based on race.}
 		\label{fig:cluster_map}
@@
 
-		We expect that spatially complex cities with high $J(X,Y)$ would require more complicated models in order to convey similar amounts of demographic information.
-Testing this expectation requires a quality measure defined on cluster models for each city.
-The natural measure of loss for a clustering is the mutual information $I(C,Y)$ between cluster and racial labels.
-Cities vary sharply in the level of cluster detail needed to convey equal amounts of information about race, as shown in figure \ref{fig:clusters}.
-Each clustering has approximately equal $I(C,Y)$.
-In sharply segregated Atlanta, just two clusters--one predominantly white, the other predominantly black--suffice to convey more information than do five in Boston, where the clusters are substantially more nuanced.
-Cluster 1 is predominantly white, cluster 4 majority Hispanic, and cluster 5 overwhelmingly black.
-Cluster 2 is again majority white, but is unique in that its minority citizens are almost all Hispanic, while cluster 3 is approximately equally black and Hispanic.
+	Cities vary sharply in the level of cluster detail needed to convey equal amounts of demographic information.
+Figure \ref{fig:clusters} shows example cluster divisions for four major cities.
+Each clustering conveys approximately equal demographic information, as measured by the mutual information $I(C,Y)$ between cluster labels and racial distribution.
+In sharply segregated Atlanta, just two clusters--one predominantly white, the other predominantly black--suffice to convey as much information as five in Boston, where the clusters are substantially more nuanced.
+Boston's Cluster A is predominantly white, cluster D majority Hispanic, and cluster E majority black.
+Cluster B is again majority white, but is distinct in that its minority citizens are almost all Hispanic.
+Cluster C is approximately equally black, Hispanic, and white.
 
-		These clusterings can also reveal phenomena related to traditional concepts in segregation studies, such as affinity and concentration.
-In Figure \ref{fig:clusters}, across all cities, Asians are much more likely to be found in predominantly white clusters than they are in predominantly black or Hispanic ones, indicating significant spatial integration between these two groups.
-We can observe the phenomenon of concentration -- in which marginalized groups tend to be more highly concentrated in their regions than majority groups are in theirs -- in Atlanta and Detroit.
-In Atlanta, whites make up 61\% of cluster 1, while blacks make up 83\% of cluster 2.
-Similarly in Detroit, whites make up 70\% of cluster 1, while blacks make up a full 86\% of cluster 2.
+	Clusterings such as those shown in Figure \ref{fig:clusters} shed useful light on traditional concerns in the study of segregation.
+We have already argued that $I(X,Y)$ and $J(X,Y)$ naturally encode the high-level ideas of evenness and exposure endorsed by \cite{Reardon2004}.
+We now consider the further concepts of clustering and concentration.
+According to \cite{Press2009a}, ``\emph{clustering} measures the degree to which minority areas are located adjacent to one another'' and ``\emph{concentration} refers to the degree of a group's agglomeration in urban space'' (pages 309-310).
+When areas with similar racial composition are located close to one another, fewer spatial clusters are necessary to convey equivalent information about racial trends.
+Thus, \ref{fig:clusters} suggests that Atlanta is more clustered than Boston, a suggestion that we can quantify using the AUC methodology developed below.
+The concept of concentration invites us to consider the composition and density of each cluster.
+In Detroit, for example, whites make up 70\% of ``their'' Cluster A, while black residents make up a full 86\% of ``their'' Cluster B.
+Cluster B is also more concentrated in that it is more densely populated than Cluster A: it contains 34\% of the Detroit's population, but accounts for just 19\% of land in the analyzed area.
+Thus, in Detroit, ``black'' neighborhoods tend to be more racially homogeneous and more densely packed in urban space than ``white'' ones.
+Finally, we note that these clusterings allow us to examine the somewhat abstract concept of ``exposure'' in considerable detail.
+Figure \ref{fig:clusters} indicates that across all cities, Asians are much more likely to be found in predominantly white clusters than they are in predominantly black or Hispanic ones, indicating significant spatial integration between these two groups.
+In Chicago, some groups of Hispanics are exposed primarily to Hispanics and other whites (Cluster B), while other groups are exposed primarily to black residents (Cluster D).
 
@@
 	\end{figure}
+
+\subsection{Complexity and Neighborhood Structure}
+
+	We expect that spatially complex cities with high $J(X,Y)$ would require more complicated models in order to capture similar levels of spatiosocial structure.
+Testing this expectation requires a quality measure defined on cluster models for each city.
+The natural measure of loss for a clustering is the mutual information $I(C,Y)$ between cluster and racial labels.
 
@@
 		\centering
-			\includegraphics[width=\textwidth]{figs/info_performance_adjed.png}
+		% \includegraphics[width=\textwidth]{figs/info_performance_adjed.png}
 		\input{figs/adjusted_regression.txt}
@@
 The two measures jointly account for an adjusted 72\% of the variation in AUC.
+
+	Add in a note here about the applicability of this method.
+Generally, it makes sense on data sets where the resolution is high but gratuitous, allowing us to aggregate up in a non-arbitrary way.
+Thus, we need the census data to have greater resolution than the ``true'' scale of urban variability.

### 2016-07-01 5_figs.Rmd
minor revisions to AUC fig

@@
 	scale_x_continuous(trans = 'log10', limit = c(1, NA)) +
-	scale_y_continuous(limit = c(0,.7)) +
+	scale_y_continuous(limit = c(0,max(sub_curves$I_XY))) +
 	theme_minimal() +
@@
 	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
-	# geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(weighted_J,2))), data = sub_summary) +
+	geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(weighted_J,2))), data = sub_summary) +
 	facet_wrap(~city) + 
@@
 	xlab('Number of clusters') + 
-	ylab('I(X,Y)') + 
+	ylab('I(C,Y)') + 
 	ggtitle('AUC as Clustering Evaluation')

### 2016-07-01 tex/results.tex
minor revisions to AUC fig

@@
 
-	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$.
-An overall positive trend is evident, reflecting the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.
+	Figure \ref{fig:info_cross} shows the relationship of the mutual information (or evenness) $I(X,Y)$ and mean local information (or exposure) $J(X,Y)$.
+The global positive trend reflects the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.
 On the other hand, there are substantial variations in $J(X,Y)$ even in cities with comparable global variability $I(X,Y)$.
 For example, the cities of Detroit and Philadelphia provide a striking contrast.
-While they have comparable global variability $I(X,Y)$, Philadelphia's $J(X,Y)$ is substantially higher.
+While they have mutual information $I(X,Y)$, Philadelphia's mean local information $J(X,Y)$ is substantially higher.
 This reflects the fact that Detroit is composed of large, highly-segregated, monoracial neighborhoods, whereas Philadelphia has a much more fine-grained, intricate neighborhood structure.
 These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities.
+It may be useful summarise the two measures by calling cities close to the bottom-right ``most segregated'', but we recommend reporting both $I(X,Y)$ and $J(X,Y)$.
+It is best to compare $J(X,Y)$ only between cities with similar $I(X,Y)$; thus, while it may be right to say that Detroit has lower levels of exposure than Phildelphia, when comparing Detroit to Baltimore it should be kept in mind that Baltimore is more even in its overall sociospatial structure.
 	\begin{figure}
@@
 Cluster B is also more concentrated in that it is more densely populated than Cluster A: it contains 34\% of the Detroit's population, but accounts for just 19\% of land in the analyzed area.
+This makes Cluster B more than twice as dense as Cluster A.
 Thus, in Detroit, ``black'' neighborhoods tend to be more racially homogeneous and more densely packed in urban space than ``white'' ones.
@@
 
-	To develop a global evaluation of the hierarchical clustering of a city according to the method defined by \eqref{eq:info_dist}, we borrow the methodology of the ``area under the curve'' (AUC) used extensively in statistical learning.
+	To develop a global evaluation of the hierarchical clustering of a city according to the method defined by \eqref{eq:info_dist}, we borrow the methodology of the ``area under the curve'' (AUC) used extensively in the field of statistical learning.
 By plotting the information $I(C,Y)$ against the number of clusters $n$, we obtain a curve reflecting how the information evolves at varying scales of aggregation.
@@
 As predictors of the AUC, both $I(X,Y)$ and $J(X,Y)$ are highly significant, and the coefficient of $J(X,Y)$ is negative.
-The two measures jointly account for an adjusted 72\% of the variation in AUC.
-
-	Add in a note here about the applicability of this method.
-Generally, it makes sense on data sets where the resolution is high but gratuitous, allowing us to aggregate up in a non-arbitrary way.
-Thus, we need the census data to have greater resolution than the ``true'' scale of urban variability.
+The two measures jointly account for an adjusted 70\% of the variation in AUC.

### 2016-07-01 cca_trial.R
housecleaning -- remove old cca trial

@@ -1,51 +1,0 @@
-library(CCA)
-library(rgdal)
-library(rgeos)
-library(leaflet)
-library(ggplot2)
-
-MA <- readOGR(dsn = 'data/states/NM', layer = 'geo', verbose = FALSE)
-MA <- MA[MA$total > 0,]
-MA <- MA[MA$total / MA$ALAND > .0000,] # density cutoff
-centroids <- MA %>% gCentroid(byid = TRUE) 
-MA@data <- cbind(MA@data,centroids@coords) 
-
-pop.list <- cca(MA@data[,c('x', 'y')],s=.02)
-
-MA@data <- MA@data %>% left_join(pop.list$cluster, by = c('x' = 'long', y = 'lat'))
-
-counts <- MA@data %>% 
-	group_by(cluster_id) %>% 
-	summarise(pop = sum(total),
-			  n = n()) %>% 
-	arrange(desc(pop)) %>% 
-	filter(pop > 5e5)
-
-MA@data <- MA@data %>% left_join(counts)
-
-MA <- MA[!is.na(MA@data$pop),] 
-
-# pop.list$cluster %>% 
-# 	ggplot() + 
-# 	aes(x = long, y = lat, color = cluster_id) + 
-# 	geom_point()
-# 
-# 
-# 
-# MA@data <- cbind(MA@data, lag(pop.list$cluster))
-
-colors <- MA@data$cluster_id
-pal <- colorNumeric(
-	palette = 'Set3',
-	domain = colors
-)
-
-map <-leaflet() %>%
-	addProviderTiles("CartoDB.Positron") %>%
-	addPolygons(data = MA, 
-				fillColor = ~pal(colors), 
-				color = "#b2aeae", # you need to use hex colors
-				fillOpacity = 0.7, 
-				weight = 1, 
-				smoothFactor = 0.2)
-map

### 2016-07-03 5_figs.Rmd
added slides, including a few new figs

@@
 
+tab %>% select(City, HY, IXY, J) %>% 
+	filter(City %in% c( 'Albuquerque', 'Detroit', 'Philadelphia')) %>% 
+	arrange(City) %>%
+	mutate(HY = sprintf("%.2f", HY),
+		   IXY = sprintf("%.2f", IXY),
+		   J = sprintf("%.2f", J)) %>% 
+	write_csv('figs/mini_table.csv')
+
 tab$Population <- format(tab$Population, big.mark=",", scientific=FALSE)
@@
 		scale_fill_brewer(palette = 'Set3', name = NULL, guide = guide_legend(nrow = 1), labels = LETTERS[1:nclusters]) + 
-		ggtitle(paste0('Example clustering for ', city)) + 
+		# ggtitle(paste0('Example clustering for ', city)) + 
 		theme_minimal() +
@@
 
-ggsave('figs/example_cluster_map.png')
+ggsave('figs/example_cluster_map.png',  width = 8, height = 4)
 
@@
 		spread(key = race, value = n, fill = 0) %>% 
-		arrange(desc(White)) %>% 
+		# arrange(desc(White)) %>% 
 		mutate(cluster = factor(LETTERS[1:nrow(.)])) %>%  
@@
 
+```{r}
+make_cluster_groups('Detroit', 6) %>% 
+	mutate(race = factor(race, levels = c('Other', 'Asian', 'Hispanic', 'Black', 'White'))) %>% 
+	ggplot(aes(y = race, x = cluster, fill = p)) + 
+	geom_tile(color = 'black') + 
+	geom_text(aes(
+		label = ifelse(p < 0.005, "", 100 * round(p,2)))) +
+	theme_minimal() + 
+	theme(axis.ticks.x = element_blank(),
+		  axis.title = element_blank(),
+		  legend.position="bottom",
+		  legend.title = element_blank()) +
+	scale_fill_continuous(low = 'white', high = 'steelblue', limits = c(0, NA), guide = FALSE) 
+	# ggtitle(paste0('Example clusters'))
+
+ggsave('figs/example_clusters_detailed.png', width = 4, height = 2)
+
+```
+
 Follow up on Detroit by area (percent of area occupied by each cluster.)
@@
 city = 'Detroit'
-nclusters = 3
+nclusters = 6
 
@@
 		   `I(X,Y)` = I_XY / max(I_XY),
-		   `J(X,Y)` = weighted_J / max(weighted_J)) %>%
+		   `J(X,Y)` = weighted_J / max(weighted_J),
+		   `H(Y)` = H_Y / max(H_Y)) %>%
 	filter(I_XY > 0) %>%

### 2016-07-03 tex/paper/abstract.tex
added slides, including a few new figs

\abstract{The problem of identifying natural, socio-economically defined neighborhoods arises in applied contexts including Census reporting, measurement of segregation, and dimension reduction in urban computing. This problem is also of interest for urban theory, since the difficulty of identifying neighborhoods may be viewed as a measure of socio-spatial complexity.

We develop a rigorous, information-theoretic approach to this topic, using open data on race in American cities as a case study. First, we formulate the \emph{mean local information} $J(X,Y)$ as a localization of the mutual information between spatial and socio-economic variables. The measure $J(X,Y)$ is closely related to the Fisher information of the underlying joint distribution, and is therefore a measure of the intrinsic spatial complexity of an urban phenomenon. Unlike standard global information measures, the mean local information clearly distinguishes between cities like Detroit--which is dominated by a few huge, monoracial superclusters--and cities like Philadelphia--which is an intricate patchwork of small, racially-distinct neighborhoods. Second, we provide a practical algorithm for identifying natural neighborhoods through greedy information maximization, and relate this algorithm's behavior to the mean local information. Questions raised by this work include the social, economic, and policy determinants of socio-spatial complexity in cities, and the potential use of spatial information measures in quantifying temporal changes in socio-economic structure, on time scales ranging from days to decades.}

### 2016-07-03 tex/paper/discussion.tex
added slides, including a few new figs

--- 2016-06-16 tex/discussion.tex
@@
+It is worthwhile to emphasize the characteristics of the data that make useful clustering possible.
+The unclustered data as provided by the U.S.
+Census presents a drawback: its boundaries are arbitrary.
+However, it also has an opportunity: its resolution is much higher than the resolution of the phenomena we aim to investigate.
+While Detroit contains over 1,800 blockgroups, it is clear that there are not 1,800 distinctive neighborhoods; rather, cluster analysis suggests that there are perhaps ten.
+We can think of our clustering method as a means of exchanging excess resolution for meaningful boundaries, with the mutual information quantifying how much meaning is lost in the exchange.
+
 We have formulated a novel measure of urban spatial complexity, and applied it to an analysis of spatial distributions of race in American cities.

### 2016-07-03 tex/paper/methods.tex
added slides, including a few new figs

--- 2016-07-01 tex/methods.tex
@@
 		  \includegraphics[width=.7\linewidth]{figs/checkerboard.png}
+		\centering
 		  \begin{tabular}{c | c c c}
@@
 From a geometric perspective, $J_Y$ provides the natural intrinsic metric in the geometric space of probability distributions parameterized by the spatial variable $x$.
-Equation \eqref{approx} therefore expresses a relationship between the local mutual information $I_r(x)$ and the information geometry of the underlying probability field.
+Equation \eqref{eq:approx} therefore expresses a relationship between the local mutual information $I_r(x)$ and the information geometry of the underlying probability field.
 Corresponding to the fact that $I_r(x_0)$ vanishes if and only if $p(y|x)$ is constant in $B_r(x_0)$, $J_Y(x_0) = 0$ if and only if $\nabla_x p(y|x_0) = 0$ for all $y$.

### 2016-07-03 tex/slides/intro.tex
added slides, including a few new figs

\documentclass{beamer}
\usepackage{import}

\usetheme{metropolis}           % Use metropolis theme
\title{Information Measures and...}
\date{\today}
\author{Phil Chodrow}
\institute{MIT Human Mobility and Networks Laboratory \& Operations Research Center}

% ---------------
% ---------------
\begin{document}
	\maketitle
	\section{Introduction: Neighborhoods, Segregation, Information Theory}

		\begin{frame}{Why care about neighborhoods?}

		\end{frame}

		\begin{frame}{Three Dimensions of Spatial Diversity}

			\begin{description}
				\item[Global diversity] is the extent to which different groups are equally represented in the city as a whole.  
				\item[Spatial exposure] \textit{refers to the extent that members of one group encounter members of another group (or their own group, in the case of spatialisolation) in their local spatial environments.} \cite{Reardon2004} 
				\item[Spatial evenness] \textit{or clustering, refers to the extent to which groups are similarly distributed in residential space.} \cite{Reardon2004}
			\end{description}

			These concepts are distinct but hierarchical:

			No global diversity $\implies$

			perfect evenness $\implies$

			maximal exposure.

  		\end{frame}

  		\begin{frame}{Albuquerque: (Relatively) Even}
	  		\begin{figure}
	  			\includegraphics[width=\textwidth]{external_figs/albuquerque_dot.png}
	  			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
	  		\end{figure}
  		\end{frame}

		\begin{frame}{Detroit: Uneven, Low Exposure}
	  		\begin{figure}
	  			\includegraphics[width=\textwidth]{external_figs/detroit_dot.png}
	  			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
	  		\end{figure}
  		\end{frame}

  		\begin{frame}{Philadelphia: Uneven, High Exposure}
	  		\begin{figure}
	  			\includegraphics[width=\textwidth]{external_figs/philly_dot.png}
	  			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
	  		\end{figure}
  		\end{frame}

  		\begin{frame}{Entropy as Global Diversity}
  			Let $p(X,Y)$ be the joint distribution of location $X$ and race $Y$. The \textbf{entropy} is defined as 
  			\begin{equation}
  				H(X) \triangleq - \mathbb{E}_Y[\log(Y)] = -\sum_{y \in \mathcal{Y}} p(Y)\log p(Y)\;.
  			\end{equation}
  			\begin{itemize}
  				\item More entropy $\implies$ more global diversity.
  				\item $H(X) = 0 \implies$ monoracial city
  				\item $H(X) = \log \left|X\right| \implies$ all races are represented equally.    
  			\end{itemize} 
  		\end{frame}

  		\begin{frame}{Mutual Information as Evenness}
  			The \textbf{mutual information} between the location $X$ and race $Y$ is defined as 	
  			\begin{equation}
  				I(X,Y) \triangleq \sum_{x,y \in \mathcal{X} \times \mathcal{Y}}p(X,Y) \log \frac{p(X,Y)}{p(X)p(Y)}\;.
  			\end{equation}
  			\begin{itemize}
  				\item More information $\implies$ less evenness. 
  				\item $I(X,Y) = 0$ implies that location and race are independent (like a correlation coefficient, but better!)
  				\item $I(X,Y) = \log \max\{\left|X\right|,\left|Y\right|\}$ implies either location completely determines race (think Detroit) or race completely determines location. 
  			\end{itemize}

  		\end{frame}

\end{document}

### 2016-07-03 slides.tex
short and long versions of slide deck

%!TEX program = xelatex

\documentclass{beamer}
\usepackage{standalone}
\usepackage{import}
\usepackage{csquotes}
\usepackage{appendixnumberbeamer}
\usepackage{csvsimple}
\usepackage{bbding}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}

\usetheme{metropolis}
\metroset{block=fill}
\metroset{progressbar=frametitle}
\setbeamertemplate{bibliography item}{}
\setbeamertemplate{caption}{\raggedright\insertcaption\par}

\makeatletter
\def\beamer@framenotesbegin{% at beginning of slide
  \gdef\beamer@noteitems{}%
  \gdef\beamer@notes{{}}% used to be totally empty.
}
\makeatother

\title{Neighborhoods and Segregation: An Information-Theoretic Lens}
\date{\today}
\author{Phil Chodrow}
\institute{MIT Human Mobility and Networks Laboratory \& Operations Research Center}

% \includeonly{tex/slides/intro} will want this later.

% -------------------------------------------------------------------------------------------------
% -------------------------------------------------------------------------------------------------
% -------------------------------------------------------------------------------------------------

\begin{document}
\maketitle
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}<handout:\included|beamer:\included>\frametitle{Table of contents}
	  \setbeamertemplate{section in toc}[sections numbered]
	  \tableofcontents[hideallsubsections]
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Neighborhoods and Diversity}
		Two questions for today: 
		\stepcounter{beamerpauses}
		\begin{enumerate}[<+ ->]
			\item What is \alert<2>{\textbf{urban diversity}}, and how should we measure it? 
			\item What is a \alert<3>{\textbf{neighborhood}}, and how can we distinguish neighborhoods in principled ways?
		\end{enumerate}
		\onslide<4->{
			\begin{block}{Central Claim}
				These two questions are deeply related, and we should approach both using \emph{information theory}.
			\end{block}}
			\note{This is a pretty damn testy note!}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Three Dimensions of Urban Diversity}

		\begin{description}[<+ ->]
			\item[\alert<1>{Global diversity}] is the extent to which different groups are equally represented in the city as a whole.  
			\item[\alert<2>{Spatial evenness}] \textit{``or clustering, refers to the extent to which groups are similarly distributed in residential space.''} \cite{Reardon2004}
			\item[\alert<3>{Spatial exposure}] \textit{``refers to the extent that members of one group encounter members of another group (or their own group, in the case of spatial isolation) in their local spatial environments.''} \cite{Reardon2004}

		\end{description}

		\onslide<4>{No global diversity $\implies$ perfect evenness $\implies$ maximal exposure.}

	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[t]\frametitle{Neighborhoods, Exposure, and the Checkerboard}
	    \textbf{Claim}: Spatial exposure is about the number, scale, and pattern of neighborhoods.
	    \begin{figure}
	    	\centering
	    	\onslide<2->{\includegraphics[width=1\textwidth]{figs/checkerboard.png}}

		    \onslide<3->{
			    \begin{tabular}{c | c c c}
			    	City & Exposure & Neighborhoods \\
			    	\hline
			    	(c) & Low & Few, large \\
			    	(d) & High & Many, small
			    \end{tabular}
		    }
	    \end{figure}

	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Albuquerque: Low diversity, fairly even, high exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/albuquerque_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Detroit: High diversity, highly uneven, low exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/detroit_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Philadelphia: High diversity, highly uneven, high exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/philly_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diversity and Information Theory}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Why Information Theory?}
		Information theory quantifies concepts like uncertainty and predictability. Why use it for studying race? 
		\onslide<2->{
			\begin{block}{Core Insight}
				\begin{itemize}
					\item Complex systems allow \textbf{some} prediction...
					\item ...But good prediction is \textbf{hard}.
				\end{itemize}
			\end{block}
		}
		\onslide<3>{
			\textbf{Examples of prediction:}
			\begin{itemize}
				\item If you pick a random person from a city, is it hard to guess their race? ``Yes'' $\implies$ more diverse city. 
				\item Does it help if you tell me where they live? ``Yes'' $\implies$ more uneven city. 
			\end{itemize}
		}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Entropy as Global Diversity}
		Let $p(X,Y)$ be the joint distribution of location $X$ and race $Y$. The marginal distribution of race alone is $p(Y) = \sum_{x \in \mathcal{X}} p(x,Y)$. 
		\begin{block}{Entropy}
			\begin{equation*}
				H(Y) \triangleq - \mathbb{E}_Y[\log p(Y)] = -\sum_{y \in \mathcal{Y}} p(Y)\log p(Y)
			\end{equation*}
		\end{block}
		\onslide<2->{\begin{align*}
			H(Y) = 0 &\implies \text{ city is monoracial} \\
			H(Y) = \log \left|\mathcal{Y}\right| &\implies \text{ all races are represented equally}
		\end{align*}}
		\onslide<3->{\textbf{Entropy measures global diversity:} more entropy, more diversity.}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Mutual Information as Evenness}
		The mutual information quantifies the degree of dependence between two random variables.
		\begin{block}{Mutual Information}
			\begin{equation*}
			I(X,Y) \triangleq \sum_{x,y \in \mathcal{X} \times \mathcal{Y}}p(X,Y) \log \frac{p(X,Y)}{p(X)p(Y)}\;.
		\end{equation*}	
		\end{block}
		\onslide<2->{\begin{align*}
			I(X,Y) = 0 &\implies \text{ location and race are independent} \\
			I(X,Y) = \log \left|Y\right| &\implies \text{ location determines race}
		\end{align*}}
		\onslide<3->{\textbf{Mutual information measures evenness:} more information, the more uneven the spatial distribution of demographics.}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Local Information}
		Let $I_r(x_0)$ be local the mutual information between $X$ and $Y$, restricted to a small area of radius $r$ centered at $x_0$. Think of this as the \textbf{spatial variation of racial trends in a small area.} 
		\onslide<2->{
			\begin{block}{Theorem}
				Under appropriate regularity conditions, for small $r$,
				\begin{equation*}
					\frac{I_r(x_0)}{r^2} \cong \frac{1}{4} \text{trace } J_Y(x_0)\;, \tag{proof on request}
				\end{equation*}
				where $J_Y(x_0)$ is the Fisher information in $Y$ about $x_0$ \nocite{Amari2000}
			\end{block}
		}
		\onslide<3->{
			\alert{Significance}: the local information $I_r(x_0)$ is related to an intrinsic statistical measure $J_Y(x)$ \emph{that is independent of the resolution $r$}.
		}

	\end{frame}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Mean Local Information Measures Spatial Exposure}
		We're interested in the average value of local information:

		\onslide<2->{
			\begin{block}{Mean Local Information}
				\begin{equation*}
					J(X,Y) \triangleq \mathbb{E}_X[\text{trace } J_Y(X)]\;.
				\end{equation*}
			\end{block}
		}
		\onslide<3->{
			\begin{align*}
				J(X,Y) = 0 &\implies \text{No neighborhoods, low exposure} \\
				J(X,Y) \text{ large} &\implies \text{Lots of neighborhoods, high exposure}
			\end{align*}
		}
		\onslide<4->{
			This is one way of ``spatializing'' the mutual information $I(X,Y)$; for another see \cite{Roberto2015a,Roberto2015}. 
		}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Back to the checkerboard}
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/checkerboard.png}
	    \end{figure}
	    \centering
	    \begin{tabular}{l c c c}
			  \textbf{City} & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
			  \hline			
			  (a) & \alert{0.0} & 0.0 & 0.0\\
			  (b) & \alert{0.7} & \alert{0.0} & 0.0\\
			  (c) & 0.7 & \alert{0.7} & \alert{0.6}\\
			  (d) & 0.7 & 0.7 & \alert{2.7}\\
		\end{tabular}

	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{A Three-Dimensional Characterization of Urban Diversity}
		\centering
		\begin{tabular}{l  c c c }%
 		    \bfseries City &  $H(Y)$  &  $I(X,Y)$  &  $J(X,Y)$ \\
 		    \hline
 	    	\csvreader[head to column names]{figs/mini_table.csv}{}% use head of csv as column names
 	    {\City & \HY & \IXY & \J \\}
    	\end{tabular}

    	\textit{Data accessed from the American Community Survey of the U.S. Census \cite{CensusRace}.
	}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Visualizing Diversity Profiles in Major Cities}
		\centering
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/mutual_fisher.png}
	    \end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Urban Density and the Compression of Social Space}
		\centering
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/density_fisher.png}
	    \end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Information-Theoretic Identification of Neighborhoods}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{How can we use this?}
		$H(Y)$, $I(X,Y)$, and $J(X,Y)$ measure diversity, evenness, and spatial exposure.

		$J(X,Y)$ also measures ``neighborhoodiness.''

		\onslide<2->{Does can we use these measures to \textbf{find} neighborhoods?}

		\onslide<3->{Yes! 
			\begin{itemize}
				\item We can use $I(X,Y)$ to find neighborhoods
				\item We can use $I(X,Y)$ and $J(X,Y)$ to estimate how many we need to preserve key patterns. 
				\item So, we can view $H(Y)$, $I(X,Y)$, and $J(X,Y)$ as measurements of \textbf{complexity} in spatiosocial structure. 
			\end{itemize} 
		}

	\end{frame}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[t]\frametitle{Key Property of Mutual Information}
	    Suppose we cluster the locations $X$ into groups with labels $C$. 
	    \begin{block}{\only<1-2>{``Additive Organizational Decomposability'' (Chain Rule)} \only<3->{Recipe for Neighborhood Identification} }
		    \begin{equation*}
		    	\only<1-2>{I(X,Y) = \onslide<1-2>{I(C,Y)} + \onslide<1-2>{I(X,Y|C)}\;.}
		    	\only<3->{I(X,Y) = \underbrace{I(C,Y)}_\text{maximize this} + \overbrace{I(X,Y|C)}^\text{minimize this}\;.}
		    \end{equation*}
	    \end{block}

	    \nocite{Reardon2004}

	    \only<2>{We can use this to identify neighborhoods!} 
	    \only<4>{This optimization is hard to do exactly, but we can use a greedy approach based on hierarchical clustering.}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[t]\frametitle{Neighborhoods in Detroit}
	    \begin{figure}
	    	\centering
	    	\includegraphics[width=.8\textwidth]{figs/example_cluster_map.png}

	    	\includegraphics[width=.6\textwidth]{figs/example_clusters_detailed.png}
	    \end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[t]\frametitle{Information Measures and ``Clusterability''}
	    We would expect that cities with low spatial exposure $J(X,Y)$ are ``easy'' to cluster. Evaluate using ``Area Under the Curve'':

	    \begin{figure}
	    	\centering
	    	\includegraphics[width=.8\textwidth]{figs/AUC_illustration.png}
	    \end{figure}
	    Together, $H(Y)$, $I(X,Y)$ and $J(X,Y)$ explain 75\% of the variability in the clusterability (AUC) of cities. 
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Wrapping Up}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}{Learnings}
		Information theory provides a natural language for the conceptualization and measurement of sociospatial variability.

	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}{Open Directions}
		Let's learn about...
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[standout]
		Thank you!

		Questions?

		Feedback? 
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% -------------------------------------------------------------------------------------------------
% APPENDIX
% -------------------------------------------------------------------------------------------------
\appendix

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\begin{frame}[allowframebreaks]{References}

	    \bibliographystyle{plain}
	    \bibliography{/Users/phil/bibs/library.bib}

	\end{frame}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}{Relation of local and Fisher informations}
		Suppose that $p(y|x) > 0$ and that $p(y|x)$ is differentiable as a function of $x$ for all $x,y \in \mathbb{R}^n \times \mathcal{Y}$. Define $B_r \triangleq B_r(x_0) \triangleq \{ x \in \mathbb{R}^n \;|\; \lvert{x - x_0}\rvert \leq r \}$. Additionally, define the \emph{local mutual information} in $B_r$ as the mutual information between $X$ and $Y$ where $X$ is restricted to $B_r$:
		\begin{align}
			I_r(x_0) &\triangleq \mathbb{E}_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r] \\
			&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;.
		\end{align}
		where $D[p\|q] \triangleq \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$. 
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}{Relation of local and Fisher informations}
		\begin{block}{Theorem}
		Under the stated conditions, 
			\begin{equation}
				\lim_{r\rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{n}{2(n+2)} \text{trace}\; J_Y(x_0)\;.
			\end{equation}
		\end{block}
			The Fisher information matrix $J_Y$ is given by 
			\begin{align}
				J_Y(x) &\triangleq \mathbb{E}_Y\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
				S_y(x) &\triangleq \log p(y|x)\;.
			\end{align}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Information Measures Check the Boxes}

		\begin{tabular}{l | c | c | c}
			\textbf{Criterion} & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
			\hline
			Organizational Equivalence & \textcolor{mLightGreen}{\CheckmarkBold} & \textcolor{mLightGreen}{\CheckmarkBold} & \textcolor{mLightGreen}{\CheckmarkBold} \\ 
			Size/Density Invariance & \textcolor{mLightGreen}{\CheckmarkBold} & \textcolor{mLightGreen}{\CheckmarkBold} & \textcolor{mLightGreen}{\CheckmarkBold} \\ 
			Additive Group Decomposability & \textcolor{mLightGreen}{\CheckmarkBold} & \textcolor{mLightGreen}{\CheckmarkBold} & \textcolor{mLightGreen}{\CheckmarkBold} \\ 
			Additive Spatial Decomposability & NA & \textcolor{mLightGreen}{\CheckmarkBold} &  * \\
			Scale Interpretability & NA & \textcolor{mLightGreen}{\CheckmarkBold} & \textcolor{mLightGreen}{\CheckmarkBold}\\ 
			Boundary Independence & NA & \textcolor{mLightBrown}{\CheckmarkBold} & \textcolor{mLightBrown}{\CheckmarkBold} \\ 
			Exchanges & NA & \textcolor{mLightGreen}{\CheckmarkBold} & \textcolor{mLightGreen}{\CheckmarkBold}
		\end{tabular}

		* In the full paper, we argue that \cite{Reardon2004}'s version is not a workable criterion for spatial measures.

		\textcolor{mLightBrown}{\CheckmarkBold} In the theoretical definition; in practice we are reliant on the data as it is provided.

	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[t]\frametitle{Some Comparative Clusters of Equal Information}
	    \begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/example_clusters.png}
	    \end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2016-07-03 slides.tex
fixes to short/long branching structure

@@
 \usepackage{amssymb}
+\usepackage{etoolbox}
 
@@
 \begin{document}
+
+% \def\short{1}
+
+\newif\iflong
+\ifdefined\short
+	\longfalse
+\else
+	\longtrue
+\fi
+
 \maketitle
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\begin{frame}<handout:\included|beamer:\included>\frametitle{Table of contents}
+	\iflong
+		\begin{frame}\frametitle{Table of contents}
 		  \setbeamertemplate{section in toc}[sections numbered]
@@
 		\end{frame}
+	\fi
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 	\begin{frame}\frametitle{Neighborhoods and Diversity}
@@
 	\end{frame}
+
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
 	\begin{frame}\frametitle{Albuquerque: Low diversity, fairly even, high exposure}
@@
 	\end{frame}
+	\fi
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
 	\begin{frame}\frametitle{Detroit: High diversity, highly uneven, low exposure}
@@
 	\end{frame}
+	\fi
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
 	\begin{frame}\frametitle{Philadelphia: High diversity, highly uneven, high exposure}
@@
 	\end{frame}
+	\fi
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
 		\begin{frame}\frametitle{Why Information Theory?}
 			Information theory quantifies concepts like uncertainty and predictability. Why use it for studying race? 
-		\onslide<2->{
+			\only<2->{
 				\begin{block}{Core Insight}
@@
 		\end{frame}
+
+	\fi
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+	\iflong
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\fi
+	\iflong
+		\longfalse
+	\else
+		\longtrue
+	\fi
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
+	\begin{frame}\frametitle{Summary: Information Theory and Diversity}
+		\begin{block}{Entropy Measures Diversity}
+			$H(Y) \triangleq - \mathbb{E}_Y[\log p(Y)]$
+		\end{block}
+		\onslide<2->{
+			\begin{block}{Mutual Information Measures Evenness}	
+				$I(X,Y) \triangleq \mathbb{E}_{X,Y}\left[\log \frac{p(X,Y)}{p(X)p(Y)}\right]$
+			\end{block}
+		}
+		\onslide<3->{
+			\begin{block}{Mean Local Information Measures Exposure}
+				$J(X,Y) \triangleq \mathbb{E}_X [\text{trace } J_Y(X)]$
+			\end{block}	
+		}	
+	\end{frame}
+	\fi
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
+		\longfalse
+	\else
+		\longtrue
+	\fi
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
 	\begin{frame}\frametitle{Urban Density and the Compression of Social Space}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
 	\begin{frame}\frametitle{How can we use this?}
@@
 	\end{frame}
-
+	\fi
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
 	\begin{frame}{Open Directions}

### 2016-07-04 5_figs.Rmd
content revisions to paper and slides

@@
 lm_eqn <- function(df){
-	m <- lm(weighted_J ~ log(density), data = df) 
-	eq <- substitute(italic(y) == a + b %.% ~'log' ~italic(rho)*","~~italic(r)^2~"="~r2, 
+	m <- lm(weighted_J ~ log(density) + I_XY, data = df) 
+	eq <- substitute(italic(y) == a + b %.% 'I(X,Y)' + c %.% ~'log' ~italic(rho)*","~~italic(r)^2~"="~r2, 
 					 list(a = format(coef(m)[1], digits = 2), 
 					 	 b = format(coef(m)[2], digits = 2), 
+					 	 c = format(coef(m)[3], digits = 2), 
 					 	 r2 = format(summary(m)$adj.r.squared, digits = 3)))
@@
 cache %>% 
+	rename(`I(X,Y)` = I_XY) %>% 
 	# filter(!grepl('Los Angeles', city)) %>%
@@
 	filter(I_XY > 0) %>%
-	lm(AUC ~ `I(X,Y)` + `J(X,Y)`, data = .) %>%
+	lm(AUC ~ `H(Y)` + `I(X,Y)` + `J(X,Y)`, data = .) %>%
 	stargazer(flip = T, 
@@
 			  style = 'all2',
-			  ci = T,covariate.labels = c('$I(X,Y)$','$J(X,Y)$', 'Intercept'),
-			  table.layout = '-=-!=!dtas',
+			  ci = T,covariate.labels = c('$H(Y)$', '$I(X,Y)$','$J(X,Y)$', 'Intercept'),
+			  # table.layout = '-=-!=!dtas',
 			  float = F) %>% 
 	write(file = 'figs/adjusted_regression.txt')
+```
+
+```{r}
+
+summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
+		   `I(X,Y)` = I_XY / max(I_XY),
+		   `J(X,Y)` = weighted_J / max(weighted_J),
+		   `H(Y)` = H_Y / max(H_Y),
+		   density = density / max(density)) %>%
+	lm(`J(X,Y)` ~ log(density) + `I(X,Y)`, data = .) %>% 
+	summary()
+
 ```

### 2016-07-04 slides.tex
content revisions to paper and slides

@@
 \usepackage{etoolbox}
+\usepackage{hyperref}
 
@@
 
-\title{Neighborhoods and Segregation: An Information-Theoretic Lens}
+\title{Urban Diversity Through an Information-Theoretic Lens}
 \date{\today}
 \author{Phil Chodrow}
-\institute{MIT Human Mobility and Networks Laboratory \& Operations Research Center}
+\institute{MIT Human Mobility and Networks Laboratory \alert| Operations Research Center}
 
@@
 	\iflong
-		\begin{frame}\frametitle{Table of contents}
+		\begin{frame}\frametitle{Outline}
 		  \setbeamertemplate{section in toc}[sections numbered]
@@
 		\begin{enumerate}[<+ ->]
-			\item What is \alert<2>{\textbf{urban diversity}}, and how should we measure it? 
-			\item What is a \alert<3>{\textbf{neighborhood}}, and how can we distinguish neighborhoods in principled ways?
+			\item What is \alert<2>{\textbf{urban diversity}}, and can we measure it in spatially-aware ways? 
+			\item What are \alert<3>{\textbf{neighborhoods}}, and how can we distinguish neighborhoods from each other?
 		\end{enumerate}
@@
 		\begin{frame}\frametitle{Why Information Theory?}
-			Information theory quantifies concepts like uncertainty and predictability. Why use it for studying race? 
-			\only<2->{
+			Information theory quantifies concepts like uncertainty and predictability. Why use it for studying diversity? 
+			\onslide<2->{
 				\begin{block}{Core Insight}
@@
 		\end{align*}}
-		\onslide<3->{\textbf{Entropy measures global diversity:} more entropy, more diversity.}
 	\end{frame}
@@
 		\begin{block}{Mutual Information}
-			\begin{equation*}
-			I(X,Y) \triangleq \sum_{x,y \in \mathcal{X} \times \mathcal{Y}}p(X,Y) \log \frac{p(X,Y)}{p(X)p(Y)}\;.
-		\end{equation*}	
+			\begin{align*}
+				I(X,Y) &\triangleq \mathbb{E}_{X,Y}\left[\log \frac{p(X,Y)}{p(X)p(Y)}\right] \\
+				&= \int_{\mathbb{R}^n} \; d^nx  \sum_{y \in \mathcal{Y}} p(x,Y) \log \frac{p(x,Y)}{p(x)p(Y)}\;.
+			\end{align*}	
 		\end{block}
@@
 		\end{align*}}
-		\onslide<3->{\textbf{Mutual information measures evenness:} more information, the more uneven the spatial distribution of demographics.}
 	\end{frame}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\begin{frame}\frametitle{Local Information}
+	\begin{frame}\frametitle{Spatializing Mutual Information}
 		Let $I_r(x_0)$ be local the mutual information between $X$ and $Y$, restricted to a small area of radius $r$ centered at $x_0$.
@@
 		\onslide<3->{
-			\alert{Significance}: the local information $I_r(x_0)$ is related to an intrinsic statistical measure $J_Y(x)$ \emph{that is independent of the resolution $r$}.
+			\alert{Significance}: the local information $I_r(x_0)$ is related to an intrinsic statistical measure $J_Y(x)$ \emph{that is independent of the spatial resolution $r$}.
 		}
@@
 	    \end{figure}
+	    \begin{itemize}
+	    	\item Philadelphia has greater exposure / more neighborhood structure than Detroit. 
+	    \end{itemize}
 	\end{frame}
@@
 	    \end{figure}
+
+	    \begin{itemize}
+		    \item Denser cities have higher exposure and more neighborhood structure
+	    \end{itemize}
+
 	\end{frame}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Information-Theoretic Identification of Neighborhoods}
+\section{Identifying Natural Neighborhoods}
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-		$J(X,Y)$ also measures ``neighborhoodiness.''
+		Also, spatial exposure is equivalent to neighborhood structure.
 
-		\onslide<2->{Does can we use these measures to \textbf{find} neighborhoods?}
+		\onslide<2->{So, can we use these measures to \textbf{find} neighborhoods?}
 
-		\onslide<3->{Yes! 
+		\onslide<3->{\textbf{Yes!}
 			\begin{itemize}
 				\item We can use $I(X,Y)$ to find neighborhoods
-				\item We can use $I(X,Y)$ and $J(X,Y)$ to estimate how many we need to preserve key patterns. 
+				\item We can use $H(Y)$, $I(X,Y)$, and $J(X,Y)$ to estimate how neighborhoods many we need to preserve key patterns. 
 				\item So, we can view $H(Y)$, $I(X,Y)$, and $J(X,Y)$ as measurements of \textbf{complexity} in spatiosocial structure. 
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\begin{frame}[t]\frametitle{Neighborhoods in Detroit}
+	\begin{frame}[t]\frametitle{Practical Example: Racial Neighborhoods in Detroit}
 	    \begin{figure}
@@
 	    	\includegraphics[width=.8\textwidth]{figs/example_cluster_map.png}
-
-	    	\includegraphics[width=.6\textwidth]{figs/example_clusters_detailed.png}
+		\begin{columns}
+			\begin{column}{0.6\textwidth}
+				\includegraphics[width=\textwidth]{figs/example_clusters_detailed.png}
+			\end{column}
+			\begin{column}{0.3\textwidth}  
+			     \emph{Percentage of population by race and neighborhood.}
+			\end{column}
+		\end{columns}
 	    \end{figure}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\iflong
 	\begin{frame}[t]\frametitle{Information Measures and ``Clusterability''}
-	    We would expect that cities with low spatial exposure $J(X,Y)$ are ``easy'' to cluster. Evaluate using ``Area Under the Curve'':
-
+	    Evaluate clusterings using ``Area Under the Curve'':
 	    \begin{figure}
@@
 	    \end{figure}
-	    Together, $H(Y)$, $I(X,Y)$ and $J(X,Y)$ explain 75\% of the variability in the clusterability (AUC) of cities. 
+	    Together, $H(Y)$, $I(X,Y)$ and $J(X,Y)$ explain 75\% of the variability in the clusterability (AUC) of cities -- they measure neighborhood structure.  
 	\end{frame}
+	\fi
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\begin{frame}{Learnings}
-		Information theory provides a natural language for the conceptualization and measurement of sociospatial variability.
+	\begin{frame}{Generality and future work}
 
+		\begin{itemize}
+			\item \textbf{Other demographics}: education level, income, occupation type, etc.
+			\item \textbf{Other dimensions}: time
+				\begin{itemize}
+					\item Evolution of urban demographics (timescale: decades)
+					\item Dynamics of diversity in daily mobility (timescale: minutes)
+				\end{itemize}
+		\end{itemize}
+	\vspace{5mm}
+		\begin{columns}
+			\begin{column}{0.3\textwidth}
+				\includegraphics[height=100pt]{external_figs/fig_od_cropped.png}
+			\end{column}
+			\begin{column}{0.4\textwidth}  
+			     Daily commuting flows in Riyad, Saudi Arabia. \emph{Image credit: Shan Jiang and MIT HuMNetLab}
+			\end{column}
+		\end{columns}
 	\end{frame}
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\begin{frame}{Download our tools!}
+		\begin{block}{R Package for Information-Theoretic Analysis}
+			\url{https://github.com/PhilChodrow/compx}
+		\end{block}
+		\begin{block}{This project: analysis and presentations}
+			\url{https://github.com/PhilChodrow/spatial_complexity}
+		\end{block}
+	\end{frame}
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 	\iflong
-	\begin{frame}{Open Directions}
-		Let's learn about...
+	\begin{frame}{Recap}
+		Information theory offers:
+		\begin{itemize}
+			\item A \textbf{unified mathematical approach} to traditional concerns in segregation studies
+			\item A \textbf{practical toolbox} for applied analysis of spatio-social trends
+			\item A \textbf{theoretical platform} for formulating and asking new questions about human difference in space and time. 
+		\end{itemize}
 	\end{frame}
@@
 	\begin{frame}[standout]
-		Thank you!
-
-		Questions?
-
-		Feedback? 
+		THANK YOU! 
+		Questions? Feedback? 
 	\end{frame}
@@
 \appendix
-
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Supplementary Materials}
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+	\nocite{RCoreTeam2016,Bettencourt2015,Bivand2014a,Bivand2014b,Bivand2014,Cover1991,Csiszzr2004,Lee2008,Massey1988,Reardon2002,Reardon2004}
+
 	\begin{frame}[allowframebreaks]{References}
-
 		    \bibliographystyle{plain}
 		    \bibliography{/Users/phil/bibs/library.bib}
-
 	\end{frame}
@@
 
-		* In the full paper, we argue that \cite{Reardon2004}'s version is not a workable criterion for spatial measures.
+		* We contend that \cite{Reardon2004}'s version is not a workable criterion for explicitly spatial measures.
 
@@
 	    	\includegraphics[width=\textwidth]{figs/example_clusters.png}
+
 	    \end{figure}

### 2016-07-04 tex/paper/abstract.tex
content revisions to paper and slides

@@
-\abstract{The problem of identifying natural, socio-economically defined neighborhoods arises in applied contexts including Census reporting, measurement of segregation, and dimension reduction in urban computing.
-This problem is also of interest for urban theory, since the difficulty of identifying neighborhoods may be viewed as a measure of socio-spatial complexity.
+\abstract{We consider two related problems.
+First, the measurement of socio-spatial segregation is an ongoing concern of quantitative sociologists.
+Second, the identification of natural, socio-economically defined neighborhoods arises in Census reporting, applied sociological analysis, and dimension reduction in urban computing.
 
-We develop a rigorous, information-theoretic approach to this topic, using open data on race in American cities as a case study.
-First, we formulate the \emph{mean local information} $J(X,Y)$ as a localization of the mutual information between spatial and socio-economic variables.
-The measure $J(X,Y)$ is closely related to the Fisher information of the underlying joint distribution, and is therefore a measure of the intrinsic spatial complexity of an urban phenomenon.
-Unlike standard global information measures, the mean local information clearly distinguishes between cities like Detroit--which is dominated by a few huge, monoracial superclusters--and cities like Philadelphia--which is an intricate patchwork of small, racially-distinct neighborhoods.
-Second, we provide a practical algorithm for identifying natural neighborhoods through greedy information maximization, and relate this algorithm's behavior to the mean local information.
-Questions raised by this work include the social, economic, and policy determinants of socio-spatial complexity in cities, and the potential use of spatial information measures in quantifying temporal changes in socio-economic structure, on time scales ranging from days to decades.}
+We unify these problems by developing a rigorous, information-theoretic approach to both, using open data on race in American cities as a case study.
+We formulate a suite of three information measures, including \emph{mean local information} $J(X,Y)$.
+The local information is a novel measure of complexity in unevenly organized cities, and we prove that it is closely related to the Fisher information of the underlying joint distribution.
+We show that the local information measures both spatial exposure as defined in segregation studies and the fineness of neighborhood structure.
+Unlike aspatial information measures, the mean local information clearly distinguishes between cities like Detroit--which is dominated by a few huge, monoracial superclusters--and cities like Philadelphia--which is an intricate patchwork of small, racially-distinct neighborhoods.
+Second, we provide a practical algorithm for identifying natural neighborhoods through greedy information maximization, and show that the fineness of neighborhood structure as identified by this algorithm is closely related to these information measures.
+This work provides a unified methodological foundation for further research into sociospatial phenomena, the dynamics of diversity over decades or days, and into scaling relations between urban density and neighborhood size.}

### 2016-07-04 tex/paper/appendix.tex
added slides, including a few new figs
content revisions to paper and slides

--- 2016-06-16 tex/appendix.tex
@@
+\subsection*{Software}
+
+  We are pleased to make available two software repositories accompanying this analysis. 
+  \begin{itemize}
+    \item Package 
+    \texttt{compx} for the \texttt{R} programming language implements computation of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$, as well as a method for information-theoretic clustering.
+Access \texttt{compx} at
+    \begin{displayquote}
+      \texttt{https://github.com/PhilChodrow/compx}. 
+    \end{displayquote}
+    To install in \texttt{R}, install the package \texttt{devtools} at 
+    \begin{displayquote}
+    \texttt{https://cran.r-project.org/web/packages/devtools/index.html}
+    \end{displayquote}
+     and run the command 
+    \begin{displayquote}
+      \texttt{devtools::install\_github(``PhilChodrow/compx'')}
+    \end{displayquote}
+    in the \texttt{R} console. 
+    \item The analysis repository for this project, including data acquisition and processing; core computations; and figure generation.
+We hope that this repository will provide useful examples of how to use \texttt{compx} for others aiming to replicate and extend our results.
+Download the project files at
+    \begin{displayquote}
+      \texttt{https://github.com/PhilChodrow/spatial\_complexity}. 
+    \end{displayquote}
+  \end{itemize}
+
 \subsection*{Relation of mutual and Fisher informations}
@@
 	\end{equation}
-	Using \eqref{approx}, the estimated Fisher information is 
+	Using \eqref{eq:approx}, the estimated Fisher information is 
 	\begin{equation}
@@
 	\end{equation}
+
+\subsection{Specification of Spatially-Constrained Information-Theoretic Clustering}

### 2016-07-04 tex/paper/discussion.tex
content revisions to paper and slides

--- 2016-06-16 tex/discussion.tex
@@
-We have formulated a novel measure of urban spatial complexity, and applied it to an analysis of spatial distributions of race in American cities.
-We emphasize that the mean local spatial variability $J(X,Y)$ is novel piece of a more complete information-theoretic characterization that also includes the entropy $H(Y)$ and the global variability $I(X,Y)$.
-Our results suggest two major directions of further exploration:
+We have presented an information-theoretic framework for conceptualizing three core dimensions of urban diversity.
+Entropy measures global diversity; mutual information measures spatial evenness; and local information measures spatial exposure.
+Using information-based agglomerative clustering, we have also shown that these three measures are intimately related to the fineness of neighborhood structure.
+As \ref{fig:info_and_clusters} indicates, cities whose neighborhood structure can be easily characterized with just a few clusters tend to be those with either low diversity $H(Y)$ or low spatial exposure $J(X,Y)$, like Detroit.
+Cities with finer-grained spatial structure tend to have higher diversity $H(Y)$ and greater spatial exposure $J(X,Y)$, like Philadelphia.
+These tools are emminently practical; allowing both for the quantitative comparison of diversity profiles between cities (Figure \ref{fig:info_cross}) and more detailed analysis within a single city using naturally-defined neighborhoods (Figures \ref{fig:cluster_map} and \ref{fig:clusters}).
+We conclude by noting some areas for future work, and some intriguing system properties illuminated by our findings.
 
-One major physical question raised by our results is the origin of the scaling relation seen in \ref{fig:density}.
-This scaling relation may suggest a dynamical process of neighborhood formation and growth common to many American cities.
-If so, a physical understanding of this process in terms of individual-level behavior is called for.
-It is tempting to view the scaling relationship as an interplay of spatial preferential attachment and intrinsic limits to neighborhood density due to housing availability, but this view is speculative at this point.
-A theory to explain this behavior would be most welcome.
+While we have presented our information-theoretic framework in the context of racial residential segregation, these methods are substantially more general.
+Generalizing from race, information-theoretic methods applicable to any categorical variable or combinations thereof.
+It would be simple to conduct similar analysis for e.g.
+occupation type, or even to for a combination of race and occupation type.
+Information methods are somewhat less applicable in the case of continuous indicators \cite{Cover1991}.
+However, even in this case it would be possible to define interpretable qualitative categories such as \{``working class'', ``middle class'', ``upper class''\} and proceed with analysis.
 
-Another question raised is more operational.
-In many urban planning contexts, the amount of available data may make direct computations prohibitive.
-In such contexts, it may be useful to construct a simplifying model of the data the data (such as a clustering) in order to reduce dimensionality and make ``the big picture'' more visible.
-When a particular demographic phenomenon (such as race) is under consideration, it may be important to ensure that the model preserves the major patterns of variation in the granular data.
-We conjecture that the local variability $J(X,Y)$ measures the ``model-ability'' of such data sets.
-Intuitively, a city like Detroit with relatively low $J(X,Y)$ and large, monoracial neighborhoods should be easily representable with relatively few modeled clusters, one for each of the relatively few major neighborhoods.
-On the other hand, relatively speaking, a city like Philadelphia with very intricate neighborhood structure may require substantially greater model complexity to represent without large loss of information.
-A promising course of further study is to design an information-theoretic clustering algorithm designed to model urban patterns, and then compare this algorithm's performance to $J(X,Y)$.
+Generalizing from space, the remaining dimension of sociological interest is time.
+Many adults spend less than half of their waking hours at home \cite{employment_stats}, indicating that residential segregation is only a partial characterization of city-wide diversity.
+Fortunately, our ability to learn about daily patterns of human mobility is progressing at a rapid pace.
+Recent years have seen an enormous increase in the use of mobile devices, allowing the passive collection of digital traces.
+These traces can be processed, analyzed, and validated to derive insight into daily activity patterns \cite{Widhalm2015,Yang,Jiang2013,Jiang2012c}.
+In the context of these developments, the information-theoretic suite of tools is attractive in its generality.
+Given appropriate data, only minimal mathematical changes are necessary:
+\begin{itemize} 
+	\item The entropy $H(Y)$ remains unchanged. 
+	\item Spatiotemporal evenness is the degree to which spatiosocial distributions change throughout the day.
+It can be measured by the global spatiotemporal mutual information $I([T,X];Y)$ between time $T$, space $X$, and demographics $Y$.
+Also of interest is the quantity $I(X,Y|T=t)$, which measures spatial unevenness at a given time of day $t$.
+	\item The local information $J(T,X,Y)$ retains its primary definition, but now splits into two parts:  $J(T,X,Y) = J(T,Y) + J(X,Y)$.
+The first term on the right measures how rapidly racial trends at fixed locations change with time, while the second term is the familiar (spatial) local information.
+\end{itemize}
+It is also possible to define a greedily information-maximizing spatiotemporal clustering algorithm.
+Such an algorithm may be of special interest in temporal analysis on longer time-scales.
+For example, given many years of Census data, spatiotemporal clustering would allow the identification of neighborhoods and a characterization of their demographic and spatial evolutions.
 
-A third question and final question relates to the time-dependence of spatial structure.
-On one time scale, daily movement around a city for work or leisure activities has the impact of ``mixing'' separated residents in public spaces, potentially leading to very different spatial patterns of racial difference.
-On another time scale, these measures may track how the spatial structure of cities evolves over decades, potentially shedding further light on the dynamics of city formation.
-
-A final question is....
+The scaling relationship between the local information $J(X,Y)$ and the population density as shown in Figure \ref{fig:density} also deserves further investigation.
+Keeping in mind that $J(X,Y)$ measures the fineness of neighborhood structure, the upshot of Figure \ref{fig:density} is that, controlling for evenness, denser cities tend to have more racially-distinct neighborhoods per area.
+While further analysis is necessary to make this observation precise, it is intriguing to postulate a dynamical process of racial neighborhood formation and growth common to many American cities.
+It may be, for example, that an explanation is possible in terms of the interplay between spatial preferential attachment and infrastructure constraints.

### 2016-07-04 tex/paper/intro.tex
content revisions to paper and slides

--- 2016-07-01 tex/intro.tex
@@
-\subsection{Neighborhoods, Segregation, and Information}
-
 	The neighborhood in which an individual lives has profound impact on that individual's mental and physical well-being \cite{Ludwig2012, Ludwig2013}.
@@
 Many indices have been proposed in the last half-century to measure various aspects of segregation; see \cite{Massey1988, Reardon2002} for useful synthesizing overviews.
-Many such approaches have also taken Census-defined boundaries as constitutive of neighborhoods, but more recent work has focused ``spatialized'' measures that are independent of neighborhood demarcations\cite{Reardon2004, Roberto2015a, Roberto2015,Wong2004, Wong1999, Lee2008}.
+Many such approaches have also taken Census-defined boundaries as constitutive of neighborhoods, but more recent work has focused ``spatialized'' measures that are independent of neighborhood demarcations \cite{Reardon2004, Roberto2015a, Roberto2015,Wong2004, Wong1999, Lee2008}.
 Other work has quantified the tendency of different social groups to cluster together \cite{Louf2015}.
@@
 
-\subsection{Structure of the Essay}
+\subsection*{Structure of the Essay}
 	In Section 2, we motivate and develop the two core components of our mathematical framework.
@@
 	Section 4 is a discussion of our findings and prospects for future work.
-Finally, we include an Appendix supplying some mathematical details not present in the text, including a proof of our assertion that the mean local information is related to a fundamental statistical property of spatial compositional phenomena, and a thorough specification of algorithms and computations.
+We have also provided an Appendix, which includes links to software repositories, ommitted mathematical notes, and specifications of computational procedures used.

### 2016-07-04 tex/paper/methods.tex
content revisions to paper and slides

@@
 	
-\subsection{Information and the Checkerboard}
+\subsection*{Information and the Checkerboard}
 	We begin by developing two fundamental measures of information theory, the entropy of one random variable and the mutual information of two.
@@
 
-\subsection{Measuring Socio-Spatial Complexity}
+\subsection*{Measuring Socio-Spatial Complexity}
 
@@
 	\begin{align}
-		\text{trace } J_Y(x) &=  \sum_i \E_Y\left[ \left(\frac{1}{p(Y|x)} \frac{\partial p(Y|x)}{\partial x_i}\right)\right]^2\;
+		\text{trace } J_Y(x) &= \E_Y\left[ \norm{\nabla_x \log p(Y|x)}^2 \right]\;
 	\end{align}
-	may be viewed as a weighted norm of the gradient $\nabla_x p(Y|x)$, the quantity 
+	may be viewed as the average norm of the gradient of $\log p(Y|x)$, the quantity  
 	\begin{align}
 		J(X,Y) &= \E_X[\text{trace } J_Y(X)] \\
-		&= \int_M \text{trace } J_Y(X) d\prob_X
+		&= \E_{X,Y}\left[\norm{\nabla_x \log p(Y|X)}^2\right]
 	\end{align}
-	may be viewed as a cousin to total variation measures often encountered in analysis.
+	may be viewed as a cousin to total variation measures often encountered in mathematical analysis.
 
-	The application of this methodology to a discrete data set is conceptually simple.
+	While we have developed these methods in the context of a differential field of observed distributions $p(y|x)$, their application to a discrete data set is conceptually simple.
 For a given set of tracts, overlay an evenly spaced grid of radius $r$, and measure the mutual information $I_r(x)$ in each grid cell.
@@
 
-\subsection{Information Measures and Segregation Studies}
+\subsection*{Information Measures and Segregation Studies}
 
 	Considerable attention has been paid to relating segregation indices to intuitive concepts of diversity and segregation.
-We note here that the mutual information $I(X,Y)$ and mean local information $J(X,Y)$ correspond closely to the two dimensions of segregation formulated in \cite{Reardon2002}.
-The authors of \cite{Reardon2002} describe \emph{evenness} as ``the extent to which groups are similarly distributed in residential space'' (page 126), and \emph{exposure} as ``the extent that members of one group encounter members of another group...in their local spatial environments.'' The mutual information $I(X,Y)$ may be viewed as a measure of (lack of) evenness, since to say that groups are similarly distributed in residential space is to say that knowing a spatial location conveys little about the race of the people who live there.
-Large $I(X,Y)$ reflects highly uneven distributions of demographic groups.
-Complementarily, the mean local information $J(X,Y)$may be viewed as a measure of spatial exposure \emph{for a fixed level of} $I(X,Y)$, as is illustrated by cities (c) and (d) in \ref{fig:toy}.
-Though distinct, these dimensions are not independent.
-To see this, note that the most thorough kind of exposure is achieved when $I(X,Y) = J(X,Y) = 0$, as in city (b).
-Since the concept of exposure only applies in a city with spatial differences, $J(X,Y)$ must be considered jointly with $I(X,Y)$ as a measure of spatial exposure.
+In this section, we relate the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ to three standard dimensions of diversity and segregation in sociology.
+In brief,
+	\begin{itemize}
+		\item The entropy $H(Y)$ measures global diversity. 
+		\item The mutual information $I(X,Y)$ measures \emph{evenness}. 
+		\item The local information $J(X,Y)$ measures \emph{spatial exposure}. 
+	\end{itemize}
+	We will define the latter two terms and argue for these claims below. In doing so, we make a case for these measures as practical tools in the quantitative study of urban diversity.
 
-	There has also been much work showing the desirable properties of various segregation indicies.
+	One way to quantify diversity is to consider how closely the distribution of races resembles the uniform distribution.
+A ``maximally diverse'' city might be 20\% each Asian, Black, Hispanic, Other, and White.
+The entropy measures distance from the uniform distribution, as can be formalized through the identity:
+	\begin{equation}
+		H(Y) = \log \abs{\mathcal{Y}} - D[p(Y)\|u(Y)]
+	\end{equation}
+	where $u$ is the uniform distribution on $\mathcal{Y}$.
+The entropy is thus largest when the distribution $p(Y)$ is nearly uniform, and smallest when $p(Y)$ is ``farthest'' from the uniform distribution.
+Through properties of $D$, it is possible to show that these cases occur when $p(Y)$ is itself uniform and when $Y$ is a constant, respectively.
+Since entropy measures nearness to uniformity, it is an intuitive characterization of the idea of diversity.
+
+	We now turn to the relationship between the mutual information $I(X,Y)$ and sociospatial evenness.
+The authors of \cite{Reardon2002} define \emph{spatial evenness} as ``the extent to which groups are similarly distributed in residential space'' (page 126), and argue that this concept encapsulates those of clustering, concentration, and centralization used in \cite{Massey1988}.
+To see that $I(X,Y)$ measures (un)evenness, consider the model cities (b) and (c) in Figure \ref{fig:toy}.
+In city (b), knowing where a person lives carries no information about their race, in the sense that, if you are attempting to guess someone's race, learning their residence will be of no help to you.
+In contrast, in city (c), location determines race: if you learn a person's residence, you can guess their race with 100\% accuracy.
+This difference is cast in information terms, but it is precisely the difference between evenness and unevenness.
+In (c), location epistemically determines race precisely because races are unevenly distributed across locations.
+
+	Finally, we consider the local information $J(X,Y)$ and its connection to spatial exposure.
+The authors of \cite{Reardon2002} define \emph{spatial exposure} as ``the extent that members of one group encounter members of another group...in their local spatial environments'' (page 126).
+Our contention is that $J(X,Y)$ measures exposure conditional on a fixed level of evenness (as measured by $I(X,Y)$).
+To see this, consider cities (c) and (d) in Figure \ref{fig:toy}.
+These cities are equally diverse as measured by $H(Y)$ and equally uneven as measured by $I(X,Y)$.
+The difference between them is that (c) consists of two large, racially-distinct neighborhoods, of which only residents near the center are exposed to residents of the other.
+On the other hand, in (d), all residents are just one ``block'' away from residents who do not share their race.
+$J(X,Y)$ detects this difference because the localized mutual information $I_r$ in terms of which $J(X,Y)$ is defined is precisely the unevenness in a small area.
+We should emphasize, however, that $J(X,Y)$ should always be considered in the context of $I(X,Y)$.
+City (d) clearly has more exposure than (c), but less than the perfectly even (b).
+The difference is that (b) has no neighborhood structure at all.
+
+	This point highlights that, though the concepts of diversity, evenness, and exposure are conceptually distinct, they are not fully independent.
+There exists a hierarchical relationship among them.
+Complete lack of diversity ($H(Y) = 0$) implies perfect evenness ($I(X,Y) = 0$), as in city (a).
+In turn, perfect evenness implies maximal exposure, but also a lack of neighborhood structure $J(X,Y)$, as in city (b).
+While this point may appear theoretical, it is necessary to keep in mind in the context of practical data analysis.
+Interpreting whether or not $I(X,Y)$ is ``high'' in a given city requires contextualizing its value against $H(Y)$.
+Similarly, evaluating a value of $J(X,Y)$ requires keeping $I(X,Y)$ in mind.
+
+\subsection*{Desirable Properties of Information Measures}
+	How do these information measures There has also been much work showing the desirable properties of various segregation indicies.
 To give a brief sampling, indices should be invariant to changes in overall population size; they should decrease when populations ``even themselves out,'' and they should behave predictably under aggregation.
@@
 A similar version of the chain rule expresses additive decomposability for aggregation of racial groups rather than spatial locations.
+
+\color{gray} % this section is very WIP.
 
@@
 However, many of the remaining ones -- such as composition invariance -- are either vague or controversial, and we will not consider further.
+\color{black}
 
-\subsection{Identifying Natural Neighborhoods}
+\subsection*{Identifying Natural Neighborhoods}
 
@@
 	\begin{equation}
-		(i^*,j^*) = \argmin_{i \text{neighbors} j} d(i,j),
+		(i^*,j^*) = \argmin_{i \text{ neighbors } j} d(i,j),
 	\end{equation}

### 2016-07-04 tex/paper/results.tex
content revisions to paper and slides

--- 2016-07-01 tex/results.tex
@@
 
-\subsection{Data Used}
+\subsection*{Data Used}
 	We assembled block-group level data from the 2010-2014 American Community Survey (ACS), conducted by the U.S.
@@
 
-\subsection{Information Measures}
+\subsection*{Information Measures}
 
@@
 
-\subsection{Learning From Natural Neighborhoods}
+\subsection*{Learning From Natural Neighborhoods}
 
@@
 We now consider the further concepts of clustering and concentration.
-According to \cite{Press2009a}, ``\emph{clustering} measures the degree to which minority areas are located adjacent to one another'' and ``\emph{concentration} refers to the degree of a group's agglomeration in urban space'' (pages 309-310).
+According to \cite{Massey1988}, ``\emph{clustering} measures the degree to which minority areas are located adjacent to one another'' and ``\emph{concentration} refers to the degree of a group's agglomeration in urban space'' (pages 309-310).
 When areas with similar racial composition are located close to one another, fewer spatial clusters are necessary to convey equivalent information about racial trends.
@@
 
-\subsection{Complexity and Neighborhood Structure}
+\subsection*{Complexity and Neighborhood Structure}
 
@@
 
-	To develop a global evaluation of the hierarchical clustering of a city according to the method defined by \eqref{eq:info_dist}, we borrow the methodology of the ``area under the curve'' (AUC) used extensively in the field of statistical learning.
+	To develop a global evaluation of the hierarchical clustering of a city according to the method defined by \eqref{eq:info_dist}, we borrow the methodology of the ``area under the curve'' (AUC) used extensively in statistical learning.
 By plotting the information $I(C,Y)$ against the number of clusters $n$, we obtain a curve reflecting how the information evolves at varying scales of aggregation.
@@
 
-		\caption{Relationship between information measures $I(X,Y)$, $J(X,Y)$, and the clustering AUC.
+		\caption{Summary of regression of the clustering AUC on information measures $H(Y)$, $I(X,Y)$, $J(X,Y)$.
 The AUC has been geographically adjusted to account for cities like New York, whose Census blockgroups form eight disconnected components, which would mildly distort results without adjustment.

### 2016-08-29 tex/paper/appendix.tex
Revisions to main proof and tech spec of computations.

@@
 
-\subsection*{Relation of mutual and Fisher informations}
+\subsection*{Relationship of Mutual and Fisher Informations}
+
 	Let $X$ be a continuous random variable taking values in $\R^n$, and let $Y$ be a discrete random variable define on finite alphabet $\mathcal{Y}$.
@@
 
-	The proof of Theorem \ref{thm1} proceeds by the application of a number of Taylor approximations, in tandem with a fundamental relationship of information geometry.
-We first expand out $I_r(x_0)$ explicitly as
+	We first define the following functions. For set $A$, let 
 	\begin{equation}
-		I_r(x_0) = \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;. \label{eq:explicit}
+		f_A(x) \triangleq  D[p(y|x) \| p(y|x\in A)] \;.
+	\end{equation}
+	Additionally, let 
+	\begin{align}
+		q_r(x) &\triangleq p(x|X \in B_r)\;.
+	\end{align}
+	We can then write the local mutual information as 
+	\begin{equation}
+		I_r(x_0) = \int_{B_r} q_r f_{B_r} \; d\lambda\;,
+	\end{equation}
+	where $\lambda$ is the Lebesgue measure in $\R^n$. We also let $V_r \triangleq \int_{B_r} \; d\lambda$ be the volume of $B_r$. Finally, define 
+	\begin{equation}
+		a_r(y) \triangleq p(X \in B_r, Y). 
 	\end{equation}
 
-	\begin{lm} \label{approximations}
-		The following approximation relationships hold for the components of \eqref{eq:explicit}:
-		\begin{enumerate}[label=\emph{(\alph*)}]
-			\item $p(X\in B_r,Y) = p(x_0, Y) v(B_r) + O(r^{n+2})$
-			\item $p(Y|X \in B_r) = p(Y|x_0) + e_y$ where the error terms $e_y$ satisfy $e_y \in O(r^{2})$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
-			\item $p(x|X \in B_r) = \frac{1 + O(r)}{v(B_r)}$
-		\end{enumerate}
+	\begin{lm}
+		For all $x \in B_r$,  
+		\begin{equation}
+			q_r(x) = \frac{1 + O(r)}{V_r}\;.	
+		\end{equation}
 	\end{lm}
 	\begin{proof}
-		For each approximation,  we directly apply Taylor expansions about $X = x_0$. 
-		\begin{enumerate}[label=(\alph*)]
-			\item We have 
+		First, Taylor's theorem implies that there exists a linear map $T$ such that 
 		\begin{align}
-					p(X\in B_r, Y) &= \int_{B_r} p(x,Y) \; d^nx \\
-					&= \int_{B_r} p(x_0,Y) + \frac{\partial p(x_0, Y)}{\partial x}(x - x_0) + O(\norm{x - x_0}^2) \; d^nx \\
-					&= p(x_0, Y)v(B_r) +  \frac{\partial p(x_0, Y)}{\partial x} \int_{B_r} (x - x_0)\; d^nx  \\
-					&\quad \quad+ O\left(\int_{B_r} \norm{x - x_0}^2\; d^nx\right) \\
-					&= p(x_0, Y)v(B_r) + O(r^{n+2})\;,
+			p(x) &= p(x_0) + T(x - x_0) + O(\norm{x - x_0}^2) \\
+			&= p(x_0) + T(x - x_0) + O(r^2)\;. 
 		\end{align}
-				where the middle term vanishes due to spherical symmetry. 
-			\item The fact that the error terms $e_y$ must satisfy $\sum_{y \in \mathcal{Y}} e_y = 0$ follows from the fact that $p(Y|X \in B_r)$ must be a valid probability distribution over $\mathcal{Y}$.
-We'll now show that $e_y \in O(r^{ß2})$.
-First,
+		 Then, 
 		\begin{align}
-					p(X \in B_r) &= \sum_{y\in \mathcal{Y}} p(X \in B_r, y)  \\
-					&= \sum_{y\in \mathcal{Y}} \left[p(x_0, y)v(B_r) + O(r^{n+2}) \right] \\
-					&= p(x_0)v(B_r) + O(r^{2});\,
+			p(X\in B_r) &= \int_{B_r} p \; d\lambda \\
+			&= \int_{B_r} p(x_0) + T(x - x_0) + O(r^2) d^n x\;.
 		\end{align}
-			from part (a). Next, 
+		The middle term vanishes by spherical symmetry, leaving 
 		\begin{align}
-					p(Y|X \in B_r) &= \frac{p(X\in B_r, Y)}{p(X \in B_r)} \\
-					&= \frac{p(x_0, Y)v(B_r) + O(r^{n+2})}{p(x_0)v(B_r) + O(r^{n+2})} \\
-					&= p(Y|x_0) + O(r^2)\;,
+			p(X\in B_r) &= p(x_0)V_r  + O(r^{n+2})\;.
 		\end{align}
-			which completes this part of the argument. 
-			\item First, 
+		For $x \in B_r$, we therefore have 
 		\begin{align}
-					p(X \in B_r) &= \int_{B_r} p(x) \; d^nx  \\
-					&= \int_{B_r} \left[p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)\right] \; d^nx \\ 
-					&= p(x_0) v(B_r) + O(r^{n+2})\;,
+			p(x|X \in B_r) &= \frac{p(x,X \in B_r)}{p(X \in B_r)} \\ 
+			&= \frac{p(x)}{p(X \in B_r)} \\
+			&= \frac{p(x_0) + T(x - x_0) + O(r^2)}{p(x_0)V_r  + O(r^{n+2})} \\
+			&= \frac{p(x_0) + O(r) + O(r^2)}{p(x_0)V_r} \\
+			&= \frac{1 + O(r)}{V_r}\;,
 		\end{align}
-				where the middle term again vanishes through spherical symmetry. Thus, for $x \in B_r$, we have
-				\begin{align}
-					p(x | X \in B_r) &= \frac{p(x)}{p(X \in B_r)} \\
-					&= \frac{p(x_0) + \nabla p(x_0)(x - x_0) + O(r^2)}{p(x_0) v(B_r) + O(r^{n+2})} \\
-					&= \frac{1 + O(r)}{v(B_r)}\;.
-				\end{align}
-		\end{enumerate}
+		as was to be shown. 
 	\end{proof}
 
-	\begin{lm} The following approximation holds for the divergence factor in the integral \eqref{eq:explicit}
+	\begin{lm}
+		For all $y\in Y$,
 		\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] = D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})
+			a_r(y) = p(x_0,y)V_r + O(r^{n+2})\;.
+		\end{equation}
+	\end{lm}
+	\begin{proof}
+		Taylor's Theorem again implies that, for any $(x,y)\in B_r \times \mathcal{Y}$, there is a linear map $T$ such that
+		\begin{align}
+			p(x,y) &= p(x_0,y) + T(x - x_0) + O(\norm{x - x_0}^2) \\
+			&= p(x_0,y) + T(x - x_0) + O(r^2). 
+		\end{align}
+		We then have 
+		\begin{align}
+			p(X \in B_r, y) &= \int_{B_r} p(x,y) \; d^n x \\ 
+			&= \int_{B_r} p(x_0,y) + T(x - x_0) + O(r^2) \; d^n x \\
+			&= p(x_0,y)V_r + O(r^{n+2}) \; d^n x\;,
+		\end{align}
+		where the middle term has again vanished due to spherical symmetry. 
+	\end{proof}
+	\begin{lm}
+		For any $y \in \mathcal{Y}$, we have $p(y|X \in B_r) = p(y|x_0) + e_y$, where the error terms $e_y$ satisfy $e_y \in O(r^2)$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
+	\end{lm}
+	\begin{proof}
+		That the errors must satisfy $\sum_{y \in \mathcal{Y}} e_y = 0$ follows from the fact that $p(\cdot| X \in B_r)$ must be a valid probability distribution on $\mathcal{Y}$.
+We'll now show that $e_y \in O(r^2)$.
+We then have
+		\begin{align}
+			p(y|X \in B_r) &= \frac{p(X \in B_r, Y)}{p(X\in B_r)} \\
+			&= \frac{p(x_0, y)V_r + O(r^{n+2})}{p(x_0)V_r + O(r^{n+2})} \\
+			&= p(y|x_0)  O(r^{2}) \tag{$V_r \propto r^n$}\;,
+		\end{align}
+		as needed. 
+	\end{proof}
+	\begin{lm}
+		For any $x \in B_r$, 
+		\begin{equation}
+			f_{B_r}(x) = f_{x_0}(x) + O(r^3)\;.
 		\end{equation}
@@
 		\begin{align}
-			D[p(\cdot|x)\| p(\cdot|X \in B_r)] &= \sum_{y \in \mathcal{Y}} p(y|x) \log \frac{p(y|x)}{p(y|X \in B_r)} \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|X \in B_r) \\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \log \left(p(y|x_0) + e_y)\right) \tag{from Lemma \ref{approximations}}\\
-			&= - H[Y|X = x] - \sum_{y \in \mathcal{Y}} p(y|x) \left[\log p(y|x_0) + \frac{e_y}{p(y|x_0)} + O(e_y^2)\right] \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}  e_y \tag{quadratic terms negligible} \\
-			&=D[p(\cdot|x)\| p(\cdot|x_0)] \\
-			&\quad + \sum_{y \in \mathcal{Y}} \left(1 + \frac{1}{p(y|x_0)} \nabla p(y|x_0)(x - x_0) + O(r^{2})\right)    e_y \\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + \sum_{y \in \mathcal{Y}} \left[e_y + O(r^{3})\right] \tag{$e_y \in O(r^2)$}\\
-			&= D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^3) \tag{$\sum_{y \in \mathcal{Y}} e_y = 0$}
+			f_{B_r}(x) &= D[p(\cdot|x)\|p(\cdot|X \in B_r)] \\
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log \frac{p(y|x)}{p(y|X \in B_r)} \\
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log p(y|x) - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|X \in B_r)\; \\
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log p(y|x) - \sum_{y \in \mathcal{Y}} p(y|x) \log (p(y|x_0) + e_y) \\ 
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log p(y|x) - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|x_0) + \frac{e_y}{p(y|x_0)} + O(e_y^2)   \\
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log p(y|x) - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|x_0) + \sum_{y\in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}e_y + O(e_y^2) \\
+			&= D[p(\cdot|x)\|p(\cdot|x_0)] + \sum_{y\in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}e_y + O(e_y^2) \\
+			&= f_{x_0}(x) + \sum_{y\in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}e_y + O(e_y^2)\;.
 		\end{align}
+		It remains to show that the error term is of order $O(r^3)$. Invoking Taylor's Theorem again, there exists a linear map $T$ such that 
+		\begin{align}
+			\sum_{y\in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}e_y + O(e_y^2) &= \sum_{y\in \mathcal{Y}} \left(1 +  \frac{1}{p(y|x_0)} T(x - x_0) + O(\norm{x - x_0}^2) \right)e_y + O(e_y^2) \\
+			&= \sum_{y\in \mathcal{Y}} \left(1 +  O(r) + O(r^2) \right)e_y + O(e_y^2) \\
+			&= \sum_{y\in\mathcal{Y}}\left[e_y + O(r^3)\right] \\
+			&= O(r^3)\;, 
+		\end{align}
+		since $\sum_{y\in \mathcal{Y}} e_y = 0$ by the previous lemma. 
 	\end{proof}
 
+	We now invoke the following well-known theorem of information geometry; for reference see \cite{Amari2000}.
+
+	\begin{thm}
+		The Kullback-Leibler divergence and Fisher information $J_Y$ are related according to the following approximation: 
+		\begin{equation}
+			D[p(\cdot|x)\|p(\cdot|x_0)] = \frac{1}{2} \left<x - x_0, J_Y(x_0)(x - x_0)\right> + O(\norm{x - x_0}^3)\;. 
+		\end{equation}
+	\end{thm}
+
+	Theorem 2 allows us to write $I_r(x_0)$ as the following spherical integral:
+
+	\begin{align}
+		I_r(x_0) &= \int_{B_r} q_r f_{B_r} \; d\lambda \\ 
+		&= \int_{B_r} q_r(x) f_{B_r}(x_0)\; d^n x \\
+		&= \int_{B_r} \frac{1 + O(r)}{V_r} \left(f_{x_0}(x) + O(r^3)\right) \; d^n x \\
+		&= \frac{1 + O(r)}{V_r} \int_{B_r} \left(f_{x_0}(x) + O(r^3)\right) \; d^n x \\
+		&= \frac{1}{2}\frac{1 + O(r)}{V_r} \int_{B_r} \left( \left<x - x_0, J_Y(x_0)(x - x_0)\right>  + O(r^3)\right) \; d^n x \\ 
+		&= \frac{1}{2}\frac{1 + O(r)}{V_r} \left(\int_{B_r}  \left<x - x_0, J_Y(x_0)(x - x_0)\right> d^n x + V_rO(r^3)\right). 
+	\end{align}
+	It remains to compute the integral, which we do using the following lemma:
+
 	\begin{lm} For any positive-semidefinite matrix $A \in \R^{n\times n}$, 
-		$$\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x = \frac{n}{n+2} r^{2} v(B_r) \text{\emph{trace}} (A)$$
+			$$\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x = \frac{n}{n+2} r^{2} V_r \text{\emph{trace}} (A)$$
 	\end{lm}
@@
 			&= r^{n+2} \sum_{i = 1}^n \lambda _i \int_{B_n} x_i^2 \; d^nx \\
-			&= \frac{r^{n+2}}{n} \sum_{i = 1}^n \lambda _i \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
-			&= \frac{r^{n+2}}{n} \text{trace}(A)  \int_{B_n} \norm{x}^2 \; d^nx \\ \tag{spherical symmetry} \\
+			&= \frac{r^{n+2}}{n} \sum_{i = 1}^n \lambda _i \int_{B_n} \norm{x}^2 \; d^nx \tag{spherical symmetry} \\
+			&= \frac{r^{n+2}}{n} \text{trace}(A)  \int_{B_n} \norm{x}^2 \; d^nx \\ 
 			&= \frac{r^{n+2}}{n} \text{trace} (A) \int_{\rho \in [0,1]} \rho^2 S_{n-1}(\rho) d\rho \\
@@
 			&= \frac{r^2}{n+2}  \text{trace} (A)  n r^{n}v(B_n(1)) \\
-			&= \frac{n}{n+2} r^{2} v(B_r) \text{trace} (A)\;,
+			&= \frac{n}{n+2} r^{2} V_r \text{trace} (A)\;,
 		\end{align}
@@
 
-	\begin{fct*}
-		The Kullback-Leibler divergence and the Fisher information $J_Y$ are related according to the approximation 
+	We may therefore complete the proof:
+
+	\begin{align}
+		I_r(x_0) &= \frac{1}{2}\frac{1 + O(r)}{V_r} \left(\int_{B_r}  \left<x - x_0, J_Y(x_0)(x - x_0)\right> d^n x + V_rO(r^3)\right) \\ 
+		&= \frac{1}{2}\frac{1 + O(r)}{V_r} \left[\frac{n}{n+2} V_r \text{trace } J_Y(x_0) + V_rO(r^3)\right] \\
+		&= r^2 \frac{n}{2(n+2)} \left(1 + O(r)\right)\left[\text{trace }J_Y(x_0) + O(r^3) \right]\;.
+	\end{align}
+	Dividing through by $r^2$ and taking the limit as $r\rightarrow 0$ proves the theorem.
+
+\subsection*{Computation of Local Information}
+	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[\text{trace } J_Y(X)]$ using blockgroup level data from the U.S.
+Census.
+For each fixed Census blockgroup $i$, let $P_i$ be the population, $A_i$ be the area, $\rho_i = P_i / A_i$ be the population density, and $p^i(y)$ be the observed proportion of racial group $y$.
+
+	Over the map we impose a hexagonal grid of radius $r$ (see \ref{fig:method} for an example).
+For each hex $k$ in the grid, let $N_k$ be the set of Census blockgroups overlapping $k$.
+Define $p^{k}(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$.
+Defining $p^k(i)$ this way reflects a simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
+Finally, define the estimated overall racial composition of hex $k$ through the formula $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$.
+The mutual information in hex $k$ is then:
 	\begin{equation}
-			D[p(\cdot|x)\| p(\cdot|x_0)] = \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3)
+		I(k) = \sum_{i \in N_k} p^k(i) D[p^i \| p^k]\;. 
 	\end{equation}
-	\end{fct*}
-
-	We are finally ready to prove Theorem \ref{thm1}. Computing directly, we have 
-	\begin{align}
-		I_r(x_0) &\triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r]\;. \\
-		&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \\
-		&= \int_{B_r} \left[\frac{1 + O(r)}{v(B_r)}\right] D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x \tag{Lemma 1(c)}\\ 
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r}  \left(D[p(\cdot|x)\| p(\cdot|x_0)] + O(r^{3})\right) d^n x \tag{Lemma 2}\\
-		&= \left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left( \frac{1}{2}\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(\norm{x - x_0}^3) + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\int_{B_r} \left(\left<x - x_0, J_Y(x_0)(x - x_0) \right> + O(r^{3})\right) d^n x \\
-		&= \frac{1}{2}\left[\frac{1 + O(r)}{v(B_r)}\right]\left(\frac{n}{n+2} r^{2} v(B_r) \text{trace} (J_Y(x_0)) + v(B_r)O(r^{3})\right) \\
-		&= r^2\frac{n}{2(n+2)}\left[1 + O(r)\right]\left(\text{trace} (J_Y(x_0)) + O(r^{3})\right)\;.
-	\end{align}
-	Dividing through by $r^2$ and computing the limit as $r \rightarrow 0$ proves the result.
-
-\subsection*{Computational Methods and Assumptions}
-	In this section, we provide a specification of the computational procedure used to estimate $J(X,Y) = \E_x[J_Y(X)]$ using blockgroup level data from the U.S. Census.
-
-	For fixed Census blockgroup $i$, let $P_i$ be the population, let $A_i$ be the area, let $\rho_i = P_i / A_i$ be the population density, and let $p^i_Y(y)$ be the observed proportion of racial group $y$.
-For hex $k$ in our hexagonal grid, let $N_k$ be the set of overlapping Census blockgroups.
-We also define $p^{k}_I(i) = \rho_i / \sum_{i \in N_k} \rho_i$ as the estimated proportion of population within hex $k$ residing in blockgroup $i$.
-This definition embodies a computationally-simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
-Finally, $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$ is the estimated overall racial composition of hex $k$.
-Then, we estimate the mutual information in hex $k$ as
+	Using \eqref{eq:approx}, the estimated Fisher information at the centroid of hex $k$ is 
 	\begin{equation}
-		I(k) = \sum_{i \in N_k} p^k_I(i) D[p^i_Y(\cdot) \| p^k_Y(\cdot)]\;. 
+		\text{trace } J_Y(k) \approx \frac{4 I(k)}{r^2}\;.
 	\end{equation}
-	Using \eqref{eq:approx}, the estimated Fisher information is 
+	The estimated population in hex $k$ is $P_k = A^k\sum_{i \in N_k} \rho_i$, where $A^k$ is the cell area. We finally compute $\E_X[\text{trace} J_Y(X)]$ as 
 	\begin{equation}
-		J(k) \approx \frac{4 I(k)}{r^2}
+		J(X,Y) = \E_X[\text{trace }J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
 	\end{equation}
-	where $r$ is the grid radius. The estimated population in hex $k$ is $P_k = A_k\sum_{i \in N_k} \rho_i$, where $A_k$ is the cell area. We finally estimate $\E_X[J(X)]$ as 
-	\begin{equation}
-		J(X,Y) = \E_X[J_Y(X)] \approx \frac{1}{\sum_k P_k} \sum_k P_k J(k)
-	\end{equation}
-
-\subsection{Specification of Spatially-Constrained Information-Theoretic Clustering}

### 2016-08-31 5_figs.Rmd
minor updates to slides and paper

@@
 
-first <- method_illustration('Detroit', vmax = 6)
-second <- method_illustration('Philadelphia', vmax = 6)
-
-gA <- ggplotGrob(first)
-gB <- ggplotGrob(second)
-g <- gridExtra::cbind.gtable(gA, gB, size = "max")
-
-grid.draw(g)
-
-ggsave('figs/method_illustration.png', g, width = 8, height = 5)
+# first <- method_illustration('Detroit', vmax = 6)
+# second <- method_illustration('Philadelphia', vmax = 6)
+# 
+# gA <- ggplotGrob(first)
+# gB <- ggplotGrob(second)
+# g <- gridExtra::cbind.gtable(gA, gB, size = "max")
+# 
+# grid.draw(g)
+# 
+# ggsave('figs/method_illustration.png', g, width = 8, height = 5)
 

### 2016-08-31 slides.tex
tech revisions
minor updates to slides and paper

@@
 \documentclass{beamer}
-\usepackage{standalone}
-\usepackage{import}
-\usepackage{csquotes}
-\usepackage{appendixnumberbeamer}
-\usepackage{csvsimple}
-\usepackage{bbding}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{etoolbox}
-\usepackage{hyperref}
-
-\usetheme{metropolis}
-\metroset{block=fill}
-\metroset{progressbar=frametitle}
-\setbeamertemplate{bibliography item}{}
-\setbeamertemplate{caption}{\raggedright\insertcaption\par}
-
-\makeatletter
-\def\beamer@framenotesbegin{% at beginning of slide
-  \gdef\beamer@noteitems{}%
-  \gdef\beamer@notes{{}}% used to be totally empty.
-}
-\makeatother
+\usepackage{pc_slides}
 
@@
 \author{Phil Chodrow}
-\institute{MIT Human Mobility and Networks Laboratory \alert| Operations Research Center}
-
-% \includeonly{tex/slides/intro} will want this later.
+\institute{MIT Human Mobility and Networks Laboratory \alert|\alert| Operations Research Center}
 
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\iflong
+
 	\begin{frame}\frametitle{Albuquerque: Low diversity, fairly even, high exposure}
@@
 	\end{frame}
-	\fi
+
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\iflong
+
 	\begin{frame}\frametitle{Detroit: High diversity, highly uneven, low exposure}
@@
 	\end{frame}
-	\fi
+
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\iflong
+
 	\begin{frame}\frametitle{Philadelphia: High diversity, highly uneven, high exposure}
@@
 	\end{frame}
-	\fi
+
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 				I(X,Y) &\triangleq \mathbb{E}_{X,Y}\left[\log \frac{p(X,Y)}{p(X)p(Y)}\right] \\
-				&= \int_{\mathbb{R}^n} \; d^nx  \sum_{y \in \mathcal{Y}} p(x,Y) \log \frac{p(x,Y)}{p(x)p(Y)}\;.
+				&= \int_{\mathbb{R}^n} \; d^nx  \sum_{y \in \mathcal{Y}} p(x,y) \log \frac{p(x,y)}{p(x)p(y)}\;.
 			\end{align*}	
@@
 				\begin{equation*}
-					\frac{I_r(x_0)}{r^2} \cong \frac{1}{4} \text{trace } J_Y(x_0)\;, \tag{proof on request}
+					\frac{I_r(x_0)}{r^2} \cong \frac{1}{4} \text{trace } J_Y(x_0)\;, 
 				\end{equation*}
@@
 	\begin{frame}\frametitle{Mean Local Information Measures Spatial Exposure}
-		We're interested in the average value of local information:
+		We're interested in the average value of local information as captured by $J_Y(X)$:
 

### 2016-08-31 tex/paper/intro.tex
minor updates to slides and paper

@@
 	\end{enumerate}
-	The following considerations connect these questions to the measurement of diversity and structure.
-If I can guess a random individual's race with 100\% accuracy given no information, then it must be the case that the city is monoracial; increased diversity reduces my ability to guess.
-If knowing where a person lives dramatically increased my ability to guess, then there must be substantial structural dependence of demographics on space, a prerequisite of segregation.
+	These questions are intimately connected with the measurement of diversity and structure.
+If I can guess a random individual's race with 100\% accuracy given no additional information, then it must be the case that the city is monoracial; increased diversity reduces my ability to guess.
+If knowing where a person lives dramatically increased my ability to guess, then there must be substantial structural dependence of demographics on space, reflecting a kind of spatial segregation.
 
 	We make two primary contributions.
-The first is to the field of quantitative segregation studies in sociology.
+The first is to the theory of segregation measurement in sociology.
 We formulate the \emph{mean local information} $J(X,Y)$ as a measure of spatiosocial structure.
-We then show that this novel measure, in concert with the mutual information $I(X,Y)$, comprise a unified, spatially-aware methodology for the measurement of diversity and segregation.
-These metrics possess many of the traditionally desirable properties of sociological indices, and can be easily computed from open data such as that provided by the U.S.
+We then show that this novel measure, in concert with the mutual information $I(X,Y)$, provides a mathematically unified, operationally interpretable, and spatially-aware methodology for the measurement of diversity and segregation.
+These metrics possess many of the traditional virtues of segregation indices and can be easily computed from open data such as that provided by the U.S.
 Census.
@@
 It is high in a city like Philadelphia, in which many (spatially) smaller neighborhoods are knit together in an intricate patchwork.
+We develop the mathematics of the mean local information and describe its properties.
 The second component of our mathematical framework is a simple algorithm for identifying clusters that are both spatially and compositionally coherent through agglomerative hierarchical clustering and greedy information maximization.
@@
 	Section 4 is a discussion of our findings and prospects for future work.
-We have also provided an Appendix, which includes links to software repositories, ommitted mathematical notes, and specifications of computational procedures used.
+We have also provided an Appendix, which includes links to software repositories, mathematical proofs, and specifications of computational procedures used.

### 2016-08-31 tex/paper/methods.tex
minor updates to slides and paper

@@
 On the other hand, in (d), all residents are just one ``block'' away from residents who do not share their race.
-$J(X,Y)$ detects this difference because the localized mutual information $I_r$ in terms of which $J(X,Y)$ is defined is precisely the unevenness in a small area.
+$J(X,Y)$ detects this difference because the localized mutual information $I_r$ in terms of which $J(X,Y)$ is defined precisely as the unevenness in a small area.
 We should emphasize, however, that $J(X,Y)$ should always be considered in the context of $I(X,Y)$.
@@
 
-\subsection*{Desirable Properties of Information Measures}
-	How do these information measures There has also been much work showing the desirable properties of various segregation indicies.
-To give a brief sampling, indices should be invariant to changes in overall population size; they should decrease when populations ``even themselves out,'' and they should behave predictably under aggregation.
-Since we are presenting $I(X,Y)$ and $J(X,Y)$ as a suite of complementary information measures, we consider each in turn.
-The author of \cite{Roberto2015a} shows that the mutual information $I(X,Y)$ (which she calls the ``Divergence Index'') satisfies these and other desirable properties, generally as well or better than existing alternatives.
-We highlight one property of $I(X,Y)$ for special note, as this property will be central to our development of natural neighborhood identification.
+\subsection*{Properties of Information Diversity Measures}
+	There has been much work showing the desirable properties of various segregation indices.
+To give a brief sampling, many scholars agree that such indices should be invariant to changes in overall population size; they should decrease when populations ``even themselves out,'' and they should behave predictably under aggregation.
+How do the measures $I(X,Y)$ and $J(X,Y)$ compare?
+The author of \cite{Roberto2015a} shows that the mutual information $I(X,Y)$ (which she calls the ``Divergence Index'') satisfies the core desiderata of segregation indices, generally as well or better than existing alternatives.
+We highlight one property of $I(X,Y)$ for special note, as it will be central to our development of natural neighborhood identification.
 The principle of ``additive decomposability'' \cite{Reardon2002} stipulates that, when the data is grouped along either racial or spatial axes, a good segregation index should split into ``within group'' and ``between group'' components.
@@
 	\end{equation}
-	i.e.
-the information I have about $Y$ given that I know $X$ is equal to the information I have if I first learn the cluster $C$, plus the amount of additional information I gain if I subsequently learn the exact location $X$ as well.
+	This information identity has a simple epistemic interpretation: he information I have about someone's race $Y$ given that I know their residence $X$ is equal to the information I have if I first learn their broad neighborhood $C$, plus the amount of additional information I gain if I subsequently learn the exact residence $X$ as well.
+Mathematically, this expression decomposes $I(X,Y)$ into two useful components.
 The first term is interpretable as the between-group information, while the second is interpretable as the within-group information.
-It is therefore comparable to various ``sum of squares'' decompositions that frequently appear in classical statistics.
-Indeed, it is possible to show that, when the differences between locations are small, the mutual information can be interpreted as a variance, in which case \eqref{eq:information_decomp} expresses the sum of squares composition directly.
+It is therefore comparable to the familiar between-group/within-group ``sum of squares'' decompositions that frequently appear in classical statistical analysis.
 A similar version of the chain rule expresses additive decomposability for aggregation of racial groups rather than spatial locations.
+In the next subsection, we use \eqref{eq:information_decomp} to define a clustering procedure to identify natural neighborhoods in cities.
 
-\color{gray} % this section is very WIP.
-
-	What of $J(X,Y)$?
-Below, we provide a brief overview of the properties of $J(X,Y)$.
-$J(X,Y)$ possesses a variety of intuitive properties, and those that it does not possess either underscore the importance of reading it in conjunction with $I(X,Y)$ or do not apply to explicitly spatial measures.
-See \cite{Reardon2002,Reardon2004} for further discussion of these criteria.
-The mean local information $J(X,Y)$ satisfies:
+	We now turn to the properties of $J(X,Y)$.
+Our core argument is that $J(X,Y)$ possesses a variety of intuitive properties, and that those it does not either underscore the importance of reading it in conjunction with $I(X,Y)$ or need not apply to spatialized measures.
+See \cite{Reardon2002,Reardon2004} for detailed discussion of these criteria.
 	\begin{description}
-		\item[Organizational Equivalence:] Both the theoretical definition \eqref{eq:def_J} or the procedure to compute it from tract data remain unchanged when a tract is subdivided into smaller tracts, each of which with identical demographic structure.
+		\item[Organizational Equivalence:] Both the theoretical definition \eqref{eq:def_J} and the procedure to compute it from tract data remain unchanged when a tract is subdivided into smaller tracts, each of which with identical demographic structure.
 		\item[Size and Density Invariance:] Since $J(X,Y)$ is completely determined by the marginal distribution $p(X,Y)$, it is invariant under changes in population density. 
 		\item[Additive Group Decomposability:] When demographic groups are aggregated into super-groups, the chain rule of mutual information applied to the demographic variable $Y$ provides an an additive decomposition of the form $J(X,Y) = J(X,G) + J(X,Y|G)$, which can be interpreted as the sum of a between-groups term and a within-groups term as needed.
-		\item[Scale Interpretability:] We have $J(X,Y) = 0$ if and only if $p(Y|X)$ is constant on each connect component of the metric space $M$.
+		\item[Scale Interpretability:] We have $J(X,Y) = 0$ if and only if $p(Y|X)$ is constant on each connected component of the metric space $M$.
 When only one connected component exists, this implies that every locale has the same demographic structure as the global environment.
@@
 When all locales are monoracial, $I(X,Y)$ achieves its maximum value and $J(X,Y) = 0$.
-		\item[Boundary Independent:] $J(X,Y)$ is defined in terms of a continuous underlying probability distribution $p(X,Y)$, rather than arbitrarily-defined tracts.
+		\item[Boundary Independence:] $J(X,Y)$ is defined in terms of a continuous underlying probability distribution $p(X,Y)$, rather than arbitrarily-defined tracts.
 In computational practice, $J(X,Y)$ is indeed sensitive to the boundaries supplied with the data; however, \eqref{eq:approx} guarantees that this sensitivity vanishes as the resolution grows sufficiently small.
-		\item[Exchanges:] Exchanges that tend to ``smooth out'' demographic distributions in space reduce both $I(X,Y)$ and $J(X,Y)$.  
-	\end{description}
-	The mean local information $J(X,Y)$ fails to one of the criteria enumerated in \cite{Reardon2002,Reardon2004}; that of \textbf{additive spatial decomposability}.
-This criterion states that ``If X spatial subareas are aggregated into Y larger spatial areas, then a segregation measure should be decomposable into a sum of within- and between-area components'' \cite{Reardon2004}, page 136.
+No alternative measure can overcome limited data resolution, and, to our knowledge, no other measure comes with the same assurance of resolution-independence in the limit.
+		\item[Transfers and Exchanges:] The core idea of the principles of transfers and exchanges is that population permutations that tend to ``smooth out'' differences should reduce segregation measures.
+However, the concrete principles as stated in \cite{Reardon2002} do not hold of $J(X,Y)$.
+The core point is that $J(X,Y)$ compares distributions to those nearby, so whether or not a transfer or exchange reduces $J(X,Y)$depends not only on the two regions involved in the transfer or exchange, but also the regions surrounding them.
+$I(X,Y)$ does satisfy the two principles.
+% Consider trying to demonstrate the spatial transfer principle for the computational version.
+		\item[Additive Spatial Decomposability:] This criterion states that ``If X spatial subareas are aggregated into Y larger spatial areas, then a segregation measure should be decomposable into a sum of within- and between-area components'' \cite{Reardon2004}, page 136.
 However, considering that we have imposed additional spatial structure on our model, this criterion does not appear to be motivated.
@@
 We therefore content that additive decomposability is a desideratum of nonspatial measures (it is satisfied by $I(X,Y)$), but not of explicitly spatial ones.
+	\end{description}
+
 	\begin{figure}
@@
 	\end{figure}
-	We have not exhausted the full complement of desiderata for segregation measures.
-However, many of the remaining ones -- such as composition invariance -- are either vague or controversial, and we will not consider further.
-\color{black}
+
+	In summary, $J(X,Y)$ possesses a variety of useful, intuitive properties.
+The ones it fails to possess reflect the fact that these desiderata are typically formulated to apply to a single measure.
+While we have not exhausted the full complement of desiderata for segregation measures.
+However, many of the remaining ones -- such as composition invariance -- are either difficult to define mathematically or controversial within the sociological community, and we will not them consider further.
+
+	We close with a discussion of some of the theoretical and operational virtues of $I(X,Y)$ and $J(X,Y)$ not considered above. 
+	\begin{description}
+		\item[Interpretability:] In contrast to many previous works, we make no attempt to encapsulate every aspect of diversity in a single measure.
+The result of maintaining distinct measures is that each of our two separate measures are highly interpretable.
+The mutual information $I(X,Y)$has a standard interpretation in information theory: it is the epistemic value of knowing where an individual lives for the purposes of predicting their race, as measured using the logarithmic loss function; see \cite{Cover1991} for a detailed discussion of information measures and prediction.
+The local information $J(X,Y)$ is then interpretable as the epistemic value of knowing where an individual lives when restricting one's attention to a small area.
+The direct operational meaning of these metrics is the source of their desirable mathematical properties, and reflects the virtue of starting from a unified mathematical framework.
+		\item[Kernel Independence:] Many existing spatial information measures depend on a choice of spatial kernel function which serves as a functional definition of proximity.
+While providing flexibility, dependence on a spatial kernel function increases researcher degrees of freedom and makes it more difficult to compare studies.
+There is no theoretical consensus on the appropriate kernel.
+The mutual information $I(X,Y)$ is aspatial and requires no kernel.
+In practice, computing $J(X,Y)$ does require a choice of a grid radius (see Appendix for details), but does not require the specification of a functional kernel.
+		\item[Isotropy:] As a result of its kernel independence, $J(X,Y)$ is \emph{isotropic}, taking no consideration of natural boundaries or infrastructure.
+This property has both virtues and drawbacks.
+On the one hand, an isotropic measure is simple to compute, simple to interpret, and easily generalized.
+On the other, boundaries and infrastructure unquestionable shape interactions between people, and are clearly relevant in segregation studies.
+We note, however, that most urban anisotropies exist on the scale of a few hundred meters at most.
+Therefore, we expect that anisotropic measures will only be relevant at scales of analysis substantially smaller than the 500m used in this study.
+A detailed numerical comparison of isotropic and anisotropic measures would be most welcome.
+		\item[Connection to Data Analysis:] As we will see below, the mutual information $I(X,Y)$ can be used to define a practical algorithm for the identification of natural, demographically coherent neighborhoods, and the local information $J(X,Y)$ predicts its performance.
+These two measures are thus not only of theoretical interest, but also provide tools for the practical analysis of urban diversity.
+We are not aware of any other measures that share this helpful property.
+	\end{description}
 
@@
 We can, however, construct a greedy algorithm which leads to satisfactory results.
-Suppose that we face the problem of choosing a pair of tracts $\{i*,j*\}$ to cluster together.
+At each stage of the algorithm, we consider the problem of choosing a pair of tracts $\{i*,j*\}$ to cluster together.
 The reduction in information associated with aggregating the locations $I$ into a single cluster is

### 2016-08-31 tex/paper/results.tex
minor updates to slides and paper

@@
 	Figure \ref{fig:AUC} also shows the bounding rectangle, whose top right corner is defined by the number N of tracts in the census data set and the full mutual information $I(X,Y)$.
-The AUC is defined as the ratio of the shaded area in Figure \ref{fig:AUC} to the bounding rectangle.
-An AUC of 1 indicates that just one region carries full information; this can only occur in a city with no spatial variation, such as city (b) in \ref{fig:toy}.
-A larger AUC indicates that sample models with few regions capture more information about spatial variation in the city.
-In Detroit, there exists a clear dividing line between a predominantly white region on the west and a predominantly black region in the east.
-A two-cluster model therefore captures much of the information in the city, giving a high AUC of 0.74.
+The AUC is defined as the ratio of the shaded area in Figure \ref{fig:AUC} to the bounding rectangle, and is computed via the formula
+
+	\begin{equation}
+		A = \frac{1}{I(X,Y) \log N }\sum_{n = 1}^N I(C_n, Y) \log \frac{n + 1}{n}  ,
+	\end{equation}
+
+	where $N$ is the total number of Census blockgroups and $C_n$ is the cluster label at the level in which $n$ total clusters are defined.
+
+	When $A = 1$, it must hold that $I(C_1,Y) = I(X,Y)$.
+This implies that just one region carries full information; this can only occur in a city with no spatial unevenness, such as city (b) in \ref{fig:toy}.
+More generally, a large AUC indicates that simple models with few distinct regions capture much of the information about spatial variation in the city.
+In Atlanta, there exists a clear dividing line between a predominantly white region in the north and a predominantly black region in the south.
+A two-cluster model therefore captures much of the information in the city, giving a high AUC of 0.71.
 In Boston, on the other hand, no such clear divide exists, and more complex models are necessary to capture similar amounts of information.
-Thus, Boston has a lower AUC of 0.67.
+Thus, Boston has a lower AUC of 0.65.
 
@@
 
-	Based on our discussion so far, we would expect that the AUC is related to the information measure $J(X,Y)$.
-Figure \ref{fig:info_and_clusters} confirms this expectation.
-Overall, two factors explain much about the ``clusterability'' of a city as measured by the AUC.
-The most important factor is the overall mutual information $I(X,Y)$.
-This reflects a simple intuition: when there is little spatial variation, there is not much to cluster.
-The second factor is the mean local information $J(X,Y)$, reflecting that spatial complexity makes clustering harder.
-A simple linear regression makes these insights precise.
-As predictors of the AUC, both $I(X,Y)$ and $J(X,Y)$ are highly significant, and the coefficient of $J(X,Y)$ is negative.
-The two measures jointly account for an adjusted 70\% of the variation in AUC.
+	We argued previously that $H(Y)$, $I(X,Y)$, and $J(X,Y)$ provide a multi-dimensional characterization of spatial diversity in a city.
+Based on this discussion, we would expect that the performance of our clustering algorithm, as measured by the AUC, is closely related to these measures.
+Figure \ref{fig:info_and_clusters} confirms this expectation using simple linear regression of the AUC on the three measures.
+Overall, these three information measures explain approximately an adjusted 75\% of the ``clusterability'' of cities.
+The signs associated with $I(X,Y)$ and $J(X,Y)$are suggestive: it is easiest to identify natural neighborhoods when there is substantial spatial variation (high $I(X,Y)$) but relatively simple local structure (low $J(X,Y)$).

### 2016-10-05 1_get_states.R
updates: server versions, figs as .pdfs, new figs

@@ -1,35 +1,28 @@
-library(sp, quietly = TRUE)
-library(compx, quietly = TRUE)
+# library(sp, quietly = TRUE)
 library(acs, quietly = TRUE)
-library(plyr, quietly = TRUE)
 library(dplyr, quietly = TRUE)
-library(readr, quietly = TRUE)
-library(ggmap, quietly = TRUE)
 library(tidyr, quietly = TRUE)
-library(magrittr, quietly = TRUE)
-library(ggrepel, quietly = TRUE)
-library(grid, quietly = TRUE)
-library(rgdal, quietly = TRUE)
-library(tigris, quietly = TRUE)
 library(stringr, quietly = TRUE)
-library(acs, quietly = TRUE)
 
 
 # get list of states
 
-cities <- read_csv('assumptions/cities.csv')
-cities$state = word(cities$state, -1)
+cities <- readr::read_csv('assumptions/cities.csv')
+cities$state <- word(cities$state, -1)
+
+if(!dir.exists('data')){
+	dir.create('data')
+}
 
 
 if(!dir.exists('data/states')){
 	dir.create('data/states')
 }
 
-
 # Define data-getter
 
 f <- function(state){
-	counties <- acs.fetch(geography=geo.make(state=state, county="*"), 
+	counties <- acs::acs.fetch(geography=acs::geo.make(state=state, county="*"), 
 				   endyear = 2014,
 				   table.number="B01003")
 	counties <- as.numeric(geography(counties)[[3]])
@@ -74,11 +67,8 @@ f <- function(state){
 	tracts@data$area <- tracts@data$ALAND / 1000^2
 	
 	tracts <- tigris::geo_join(tracts, race, "GEOID", "GEOID")
-	writeOGR(tracts, paste0("data/states/", state),'geo', driver = 'ESRI Shapefile', morphToESRI = TRUE)
+	rgdal::writeOGR(tracts, paste0("data/states/", state),'geo', driver = 'ESRI Shapefile', morphToESRI = TRUE)
 }
-
-
-
 
 # Get all the states
 cities <- c(unique(toupper(cities$state[nchar(cities$state) == 2])), 'DC')
@@ -88,9 +78,3 @@ for(state in cities){
 		tryCatch(f(state), error = function(e) NULL)
 	}
 }
-
-
-
-
-
-

### 2016-10-05 2_make_cities.R

@@ -9,7 +9,6 @@ library(grid, quietly = TRUE)
 library(rgdal, quietly = TRUE)
 library(stringr, quietly = TRUE)
 
-
 city_df <- read_csv('assumptions/cities.csv')
 cities <- unique(city_df$name)
 cities <- cities[!is.element(cities, list.dirs('cities', full.names = FALSE))]
@@ -22,7 +21,9 @@ states <- city_df$state %>%
 shapes <- list()
 for(state in states){
 	print(paste('Loading', state))
-	shapes[state] <- readOGR(dsn = paste0('data/states/',state), layer = 'geo', verbose = FALSE)
+	shapes[state] <- readOGR(dsn = paste0('data/states/',state), 
+	                         layer = 'geo', 
+	                         verbose = FALSE)
 }
 
 if(!dir.exists('data/cities')){
@@ -42,13 +43,16 @@ for(city in cities){
 		component <- shapes[[state]]
 		component <- component[as.integer(as.character(component@data$COUNTYFP)) == county,]
 		city_components[paste(state, county, sep = '_')] <- spChFIDs(component,
-																	  as.character(paste(state, county, rownames(as(component,"data.frame")))))
+																	  as.character(paste(state, 
+																	                     county, 
+																	                     rownames(as(component,
+																	                                 "data.frame")))))
 	}
 	city_polys <- do.call(rbind, city_components)
 	city_polys <- city_polys[city_polys@data$total != 0,]
-	tryCatch(writeOGR(city_polys, paste0("data/cities/", city),'geo', driver = 'ESRI Shapefile', morphToESRI = TRUE),
+	tryCatch(writeOGR(city_polys, paste0("data/cities/", city),
+	                  'geo', 
+	                  driver = 'ESRI Shapefile', 
+	                  morphToESRI = TRUE),
 			 error = function(e) NULL)
 }
-	
-	
-	

### 2016-10-05 3_local_info.R

@@ -1,8 +1,5 @@
-resolution <- 1/2
-
 library(sp, quietly = TRUE)
 library(compx, quietly = TRUE)
-library(acs, quietly = TRUE)
 library(plyr, quietly = TRUE)
 library(dplyr, quietly = TRUE)
 library(readr, quietly = TRUE)
@@ -12,31 +9,23 @@ library(magrittr, quietly = TRUE)
 library(ggrepel, quietly = TRUE)
 library(grid, quietly = TRUE)
 library(rgdal, quietly = TRUE)
+library(data.table)
 
-# Toy cities ----------
-
-# checkerboard_illustration()
-# ggsave('report/figs/toy.png', width = 5, height = 1.9)
-
-# Methodological Illustration ----------------
-
-# method_illustration()
-# ggsave('report/figs/method.png', width = 3, height = 3)
-
-
+races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
+resolution <- 1
 
 cities <- list.files('data/cities')
-
-
+# cities <- c('Boston')
 
 if(!dir.exists('throughput/grids')){
 	dir.create('throughput/grids')
 }
 
-
+if(!dir.exists('throughput/grid_tracts')){
+  dir.create('throughput/grid_tracts')
+}
 
 # MAIN COMPUTATION ---------------
-
 
 # Cache logic
 
@@ -49,70 +38,50 @@ if(file.exists('throughput/info_cache.csv')){
 		area = numeric(),
 		population = numeric(),
 		density = numeric(),
-		weighted_density = numeric(),
-		n_per_hex = numeric(),
-		H_X = numeric(),
-		H_Y = numeric(),
 		I_XY = numeric(),
-		local.H_Y = numeric(),
-		J = numeric(),
-		weighted_J = numeric())
+		J_XY = numeric())
 }
 
-
 for(city in cities){
-	races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
-	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
-	tracts <- tracts[tracts@data$total > 0,]
-	tracts@data$area <- tracts@data$ALAND / 1000^2	
-	tracts <- tracts[tracts@data$total / tracts@data$area > 50,] # at least 50 people per km^2
 	
-	radius = 1/sqrt(85 * 111) * resolution # (roughly 1 km after lat-lon conversion)
-	xx = spsample(tracts, type="hexagonal", cellsize=radius)
[…]

### 2016-10-05 revision.tex

%!TEX program = xelatex

% -----------------------------------------------------------------------------
% ------------------------------- PREAMBLE ------------------------------------
% -----------------------------------------------------------------------------

\documentclass[english]{scrartcl}

\title{Segregation Measures for Data Analysis} % work on this
\author{\emph{Phil Chodrow}}
\date{\emph{\today}}

\usepackage{pc_writeup}
\usepackage{pc_math}

% -----------------------------------------------------------------------------
% --------------------------------- BODY --------------------------------------
% -----------------------------------------------------------------------------

\begin{document}
\setkomafont{disposition}{\mdseries\rmfamily}

\maketitle

\abstract{}

\section{Introduction} \label{sec:intro}
	\begin{itemize}
		\item While many measures are available for the measurement of spatial diversity, we are not aware of quantitative methods for \emph{identifying} spatial structure. 
		\item We endorse an information-theoretic approach to the study of diversity and segregation. In supporting this approach, we make two main contributions: 
		\begin{itemize}
			\item First, we argue that a simple suite of information measures -- entropy, mutual information, and local information -- intuitively measure global diversity, spatial unevenness, and spatial exposure. The mean local information is novel, and we develop some of its useful properties. This triplet of information measures readily lends itself to quantifying dimensions of segregation within and between cities. 
			\item Measures such as those we develop above can be viewed as measuring different kinds of sociospatial structure. To our knowledge, no such analysis methods have been proposed for \emph{identifying} that structure. Building on our information methods, we develop an algorithm for clustering urban areas into regions whose demographics are near constant in space. The result allows the analyst to both easily visualize spatial patterns of difference and conduct analysis using a natural partition of urban space. 
		\end{itemize}
	\end{itemize}
\section{Lit Review, Previous Work. Why our problem is new.} \label{sec:previous}
	Core points: 
	\begin{itemize}
		\item Previous work is missing unification
		\item Previous work is missing a practical component: how do you compute and learn about urban structure with it? 
	\end{itemize}

\section{Segregation and Information} \label{sec:information}
	Broadly introduce information measures as we motivate the corresponding sociological concepts. 
	\begin{itemize}
		\item The information-theoretic framework: relationship between structure and prediction
		\item Entropy measures global diversity
		\item Mutual information measures spatial unevenness
		\item Mean local information measures spatial exposure
		\item Example of how to compute
		\item Comparative metrics
	\end{itemize}

\section{Identifying Sociospatial Structure} \label{sec:id}
	\begin{itemize}
		\item Construct the algorithm. 
		\item Show algorithm results for one or two cities; show how knowing the clusters can lead to neat data analysis. 
		\item Relate algorithmic performance to suite of information measures above. Note that $J$ measures the ``natural fault lines'' where we would expect the algorithm to segment.  
	\end{itemize}

\section{Discussion} \label{sec:discussion}

\bibliography{/Users/phil/bibs/library.bib}{}
\bibliographystyle{plain}

\end{document}

### 2016-10-05 revision.tex
prep for merge

@@
 
-\title{Segregation Measures for Data Analysis} % work on this
+\title{An Information Lens on Demographic Diversity} % work on this
 \author{\emph{Phil Chodrow}}
@@
 \usepackage{pc_math}
+\usepackage[colorinlistoftodos]{todonotes}
 
@@
 \section{Segregation and Information} \label{sec:information}
-	Broadly introduce information measures as we motivate the corresponding sociological concepts. 
-	\begin{itemize}
-		\item The information-theoretic framework: relationship between structure and prediction
-		\item Entropy measures global diversity
-		\item Mutual information measures spatial unevenness
-		\item Mean local information measures spatial exposure
-		\item Example of how to compute
-		\item Comparative metrics
-	\end{itemize}
+	Information measures quantify structure through an epistemic framework.
+The core insight is that systemic structure enables \emph{prediction}, while systemic variability can make such prediction difficult.
+Two systemic features are related when knowledge of one enables better prediction of the other.
+
+	We can formalize the idea of prediction quality by considering the following ``guessing game.'' We discuss the weather, and I ask you to give me a \emph{predictive distribution} on the question of whether or not it will rain tomorrow; e.g.
+``70\% yes, 30\% no.'' Call this predictive distribution $p$.
+Tomorrow evening I give you a reward $R(p,Z)$ based on your predictive distribution $p$ and the observed outcome $Z$.
+For our game to be a true ``guessing game,'' we might expect the reward function $R$ to possess the following intuitive properties:
+	\begin{enumerate}
+		\item \textbf{Honesty:} The reward scheme $R$ encourages you to report your true beliefs about the rain.
+		\item \textbf{Fairness:} Your reward $R(p,Z)$ you receive depends only on the observed outcome, not ``what could have happened.''
+		\item \textbf{Smoothness:} Your reward increases smoothly as a function of the probability you assigned to the observed outcome. 
+	\end{enumerate}
+	It is possible to show (see e.g.
+\cite{Cover1991}) that these properties uniquely determine the logarithmic reward function $R(p,Z) = \log p(Z)$ up to an affine transformation.
+This logarithmic reward is the fundamental unit of information-theoretic analysis.
+It is customary to focus on the \emph{loss function} $L(p,Z) = - R(p,Z)$.
+Large loss corresponds to poor prediction.
+Systemic structure is present when knowledge of a system feature enables us to decrease the loss in such a guessing game.
+
+	\begin{figure}
+		\includegraphics[width=\textwidth]{figs/checkerboard.png}
+		\caption{Information and the checkerboard}. \label{fig:checkerboard}
+	\end{figure}
+
+	\subsection{Entropy and Global Diversity}
+
+		As an introductory example, consider the following guessing game: I pick a person at random from a city, and ask you to assign a predictive distribution on their race $Y$, where $Y$ takes possible values in the alphabet  $\mathcal{Y} = \{\text{Asian, Black, Hispanic, Other, White}\}$.
+There is a true distribution $p(Y)$ corresponding to the proportions of each race in the city in question.
+It is possible to show that your optimal guess is precisely this distribution $p$.
+The optimal expected loss is
+		\begin{equation}
+			\E_Y[L(p,Y)] = -\E_Y[R(p,Y)] = -\E_Y[\log p(Y)] \triangleq H(Y), 
+		\end{equation}
+		also known as the (Shannon) \emph{entropy} of $p$.
+
+		The Shannon entropy is a natural measure of the global diversity of a city.
+When $H(Y)$ is low, guessing a randomly-chosen individual's race is easy.
+An extreme example of this case is when $H(Y) = 0$; in this case, the true distribution $p(Y)$ is unity for one race and zero for the others; i.e.
+the city is monoracial.
+Figure \ref{fig:checkerboard}(a) gives an example of this case in a two-race city.
+Perfect prediction without loss is possible because this city lacks diversity.
+On the other hand, when $H(Y)$ is high, guessing a randomly-chosen individual's race is difficult.
+An extreme example of this case is when $H(Y)$ achieves its maximum value of $\log \abs{\mathcal{Y}}$, in which case each race is represented equally.
+The cities in Figure \ref{fig:checkerboard}(b)-(d) all have maximum demographic entropy.
+The underlying distribution gives you grounds on which to add probabilistic weight to one race over another; you may as well roll a fair (five-sided) die.
+Prediction is difficult because of the city's diversity.
+
+	\subsection{Mutual Information and Spatial (Un)evenness}
+
+		The authors of \cite{Reardon2004} define spatial evenness as ``the extent to which groups are similarly distributed in residential space.'' \emph{Un}evenness, then, is the extent to which these groups are dissimilarly distributed in space.
+This physical description has an epistemic analogue: knowing where a resident lives conveys information about their race, in the sense of enabling better prediction.
+To quantify this intuition, consider two versions of the guessing game above.
+In the first version, I simply choose a random resident and ask you to guess as before.
+In the second, I choose a random resident but then tell you their residence.
+In the presence of unevenness, this will generally enable you to improve your predictions.
+Let $X$ be a random variable denoting the spatial residence.
+Then, your expected loss in the first game is still $H(Y)$, and your expected loss in the second is $H(Y|X) \triangleq -\E_{Y}[\log p(Y)|X]$.
+The \emph{mutual information} $I(X,Y)$ is the predictive improvement associated with my telling you the residence:
+		\begin{equation}
+			 I(X,Y) \triangleq H(Y) - H(Y|X)\;.
+		\end{equation}
+		$H(Y)$ is the expected loss without any knowledge of $X$, while $H(Y|X)$ is the expected loss when you are provided the value of $X$ as well.
+Since knowing $X$ can only improve prediction, $I(X,Y) \geq 0$.
+Furthermore, it is possible to show that
+		\begin{equation}
+			I(X,Y) = D[p(X,Y)\|p(X)p(Y)]\;,
+		\end{equation}
+		where 
+		\begin{equation}
+			D[q\|r] \triangleq \sum_a q(a) \log \frac{q(a)}{r(a)}
+		\end{equation}
+		is the Kullback-Leibler (KL) divergence of $r$ from $q$.
+Though not an axiomatic metric, the KL divergence may be viewed as a measure of distance between two distributions.
+The mutual information $I(X,Y)$ can therefore also be viewed as the distance of the true distribution $p(X,Y)$ from the hypothetical distribution $p(X)p(Y)$ in which the variables $X$ and $Y$ are independent.
+$I(X,Y)$ therefore measures the degree of dependence between these variables.
+
+		The author of \cite{Roberto2015a} uses the mutual information $I(X,Y)$ -- which she refers to as the Divergence Index -- as a measure of (un)evenness.
+To see this within our epistemic framework, $I(X,Y) = 0$, $X$ and $Y$ are independent, indicating that knowing $X$ does not improve your ability to guess $Y$.
+City (b) in Figure \ref{fig:checkerboard} illustrates this situation; since ``everywhere is the same,'' location knowledge is of no predictive value.
+If we were playing in City (b), your performance in the second guessing game would then be equal to your performance in the first.
+Physically, this situation corresponds to perfect evenness: the distribution of races is $p(Y|X)$ is the same at every location $X$.
+In contrast, when $I(X,Y)$ achieves its maximum value of $H(Y)$, $X$ fully determines $Y$.
+This situation is illustrated by Figure \ref{fig:checkerboard}(c); if I tell you that a resident lives in the left- or right-hand side of the city, you can guess their race with perfect accuracy.
+This situation corresponds to one in which $H(Y|X) = 0$, i.e.
+every location is monoracial, reflecting maximal sociospatial unevenness.
+
+	\subsection{Spatial Exposure}
+
+		The authors of \cite{Reardon2004} define \emph{spatial exposure} as ``the extent that members of one group encounter members of another group...in their local spatial environments.'' We can construct an epistemic analogue of spatial exposure by considering the following pair of guessing games.
+In each, I pick a resident at random from a fixed ``local spatial environment'' that we both know.
+In the first game, I ask you to guess that resident's race directly.
+In the second, I tell you where \emph{within that environment} the resident lives.
+If the spatial environment in question is $R$, then your expected loss in the first game is $H(Y|X\in B)$ and in the second $H(Y|X, X\in B)$.
+The prediction improvement associated with knowing residence within this local environment is
+		\begin{equation}
+			I(X,Y | X \in B) = H(Y|X\in B) - H(Y|X, X\in B)\;.
+		\end{equation}
+		Our core claim is that $I(X,Y|X \in B)$ -- the \emph{local (mutual) information} -- is a measure of spatial exposure \emph{in unevenly distributed populations}.
+This qualification is important, since in a perfectly even population such as Figure \ref{fig:checkerboard}(b),  $I(X,Y) = 0$ and $I(X,Y|X\in B) = 0$ as well.
+Consider, however, the two unevenly distributed populations in Figure \ref{fig:checkerboard} (c) and (d).
+In Figure (c), if we play our two guessing games at a locale entirely within the left- or righthand halfs, the population is locally homogeneous and you can make perfect predictions regardless of the specific residence.
+Only if we play the game on a locale straddling the middle boundary between the two groups does knowing a resident's specific residence help you to make predictions.
+If our locale $B$ perfectly straddles the boundary, so that the population proportions are $(\frac{1}{2},\frac{1}{2})$, then we have $H(Y|X\in B) = \log 2$ and $H(Y|X,X\in B) = 0$, so $I(X,Y|X \in B) = \log 2$.
+
+		This example illustrates that the local information $I(X,Y|X\in B)$ is high when $B$ is an area of demographic transition.
+When two cities are similarly uneven (similar $I(X,Y)$), a city with more such transitional areas has more exposure between groups, since residents of one area are closer to demographically different residents of another area.
+To illustrate, the city of Figure \ref{fig:checkerboard}(d) has the same level of spatial unevenness as \ref{fig:checkerboard}(c), but greater spatial exposure.
+
+		For practical data analysis, it is necessary to understand how the value of $I(X,Y|X \in B)$ depends on the choice of the locale $B$.
+In particular, we should expect that when using fine-grained data and fine-grained spatial locales, our estimates of $I(X,Y|X \in B)$ should converge, and not vary wildly as we increase the resolution of our analysis.
+It is possible to prove mathematically that this is the case.
+The full proof is given in the appendix, but the idea of the proof is illustrative of the nature of the local information.
+It begins by noting that the local information is zero on areas of constant demographic structure, but achieves high values in transitional areas.
+In this sense it behaves like a mathematical derivative.
+Making this precise requires that we postulate a metric space $M$ and a \emph{probability field} $p(Y|X)$, $X \in M$ with the following properties:
+		\begin{enumerate}
+		 	\item $p(y|x) > 0$ for all $x \in M$
+		 	\item $p(y|x)$ is differentiable as a function of $x$ for all $x \in M$. 
+		\end{enumerate}
+		The probability field $p(y|x)$ corresponds to an idealization of our data; it can be viewed as an interpolation or approximation to the true data using a smooth curve.
+
+		Define $I_r(x_0)$ to be the local information in a locale of radius $r$ around the point $x_0$:
+		\begin{equation}
+			I_r(x_0) = I(X,Y| \norm{X - x_0}^2 \leq r^2). 
+		\end{equation}
+		We may view $r$ as the resolution of our analysis. Under the stated conditions, the following approximation relationship holds: 
+		\begin{equation}
+			\lim_{r \rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{1}{4} \text{tr } J_Y(x_0), \label{eq:fisher_approx}
+		\end{equation}
+		where $J_Y(x_0)$ is the \emph{Fisher information in $Y$ about $x_0$}. $J_Y(x_0)$ is a matrix defined by the formula
+		\begin{equation}
+			J_Y(x_0) \triangleq \E_Y\left[\nabla_x \log p(Y|x) \left(\nabla_x \log p(Y|x)\right)^T \right]\;. 
+		\end{equation}
+		The trace $\text{tr }J_Y(x_0)$ of $J_Y(x_0)$ is simply the sum of the diagonal entries of $J_Y(x_0)$.
+
+		Equation \eqref{eq:fisher_approx} implies that, for small $r$, the local information (when normalized by $r^2$) is approximately a constant that is independent of $r$.
+Furthermore, that constant is related to the derivatives of the probability field $p(Y|X)$.
+We can make this relationship more explicit by noting that
+		\begin{equation}
+			\text{tr }J_Y(x_0) = \E_Y\left[\norm{\nabla_x \log p(Y|x_0)}^2\right]\;, \label{eq:derivative}
+		\end{equation}
+		which together with \eqref{eq:fisher_approx} shows that $I_r(x_0)$ is proportional to the mean magnitude of the derivative of $p(Y|X)$ evaluated at $x = x_0$.
+
+		In the idealized mathematical framework, we may define the \emph{mean local information} as the average value of the trace of the Fisher information: 
+		\begin{equation}
+			J(X,Y) \triangleq \E_X[\text{tr } J_Y(X)].
+		\end{equation}
+		Using the relationship \eqref{eq:derivative}, we can recognize $J(X,Y)$ as the total variation of $\log p(Y|X)$ with respect to the probability measure induced by $X$.
+Intuitively, this is the average ``curviness'' of the field $p(Y|X)$.
+
+		It is simple to adopt this measure $J(X,Y)$ for practical data analysis. The procedure is as follows: 
+		\begin{enumerate}
+		 	\item Partition the city under analysis into spatial locales $\{B_i\}$.
+		 	\item Compute the local information $I(X,Y|X \in B_i)$ for each $i$. 
+		 	\item Average the results, weighted by the population of each $B_i$. 
+		\end{enumerate} 
+		The explicit formula is 
+		\begin{equation}
+			J(X,Y) = \sum_i p(X \in B_i) I(X,Y|X \in B_i). \label{eq:formula}
+		\end{equation}
+
+	\subsection{Illustrations}
+		\subsubsection{Comparative Analysis: Detroit and Philadelphia}
+			To illustrate these methods, we use them to compare the cities of Philadelphia and Detroit.
+Our data is blockgroup-level demographics from the American Community Survey \cite{CensusRace}.
+As we show, computing $H(Y)$, $I(X,Y)$, and $J(X,Y)$ together highlights similarities and differences between these cities.
+			\todo{We need to generate a nice summary plot here and then fill this section in.}
+			\begin{figure}
+				\caption{Spatial analysis of racial diversity in Detroit (left) and Philadelphia (right).
+Top: Percentage of black residents in Detroit (left) and Philadelphia (right).
+Middle: Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of $I(X,Y | X \in B_i)$.
+Bottom: Estimated contours of constant $I(X,Y | X \in B_i)$; high values correspond to spaces of demographic transition.
+Very bottom: Numerical summary, including the three information measures.} \label{fig:detroit_philly}
+			\end{figure}
+
+		\subsubsection{Comparative Diversity Profiles}
+			\todo{Is there a fun way to explore the density point? E.g. figure out people per ``neighborhood'' in some way? Think about what I would want the analysis here to be.}
+
+			We can achieve a broader picture by comparing $I(X,Y)$ and $J(X,Y)$ across a range of American cities, as we do in Figure \ref{fig:mutual_fisher}.
+The two measures $I(X,Y)$ and $J(X,Y)$ provide an intuitive sorting of major cities.
+Cities with small numbers of large, monoracial regions such as Atlanta and Detroit are sorted toward the bottom right while cities with more complicated patchworks of demographic variation like Philadelphia and New York are sorted toward the top right.
+
+			\begin{figure}
+				\includegraphics[width=\textwidth]{figs/mutual_fisher.png}
+				\caption{Comparative global diversity measures in major U.S.
+cities.
+Cities toward the bottom left are relatively spatially even.
+Cities toward the bottom right like Detroit and Atlanta are the most classically ``segregated'', with large, starkly separated monoracial regions.
+Cities toward the top right like Philadelphia and New York are complex ``patchworks'' of different, smaller regions.} \label{fig:mutual_fisher}
+			\end{figure}
+
+			\todo{Optional: plot of $J$ vs. density in here, but probably only if we can think of more to say, possibly a follow-up analysis.}
 
 \section{Identifying Sociospatial Structure} \label{sec:id}
+
+	We can view $H(Y)$, $I(X,Y)$, and $J(X,Y)$ as measures of the presence of sociospatial structure.
+It is somewhat unsatisfying to simply measure this structure without actually identifying it, since identifying this structure is of direct application in sociological data analysis.
+In this section, we use information-theoretic techniques to construct a procedure for finding subregions within cities that capture high-level demographic trends.
+
+	A useful place to start is with the property of ``additive spatial decomposition'' \cite{Reardon2004}.
+A diversity measure satisfies additive spatial decomposition when, if spatial units are aggregated into groups, the measure can be expressed as a sum of a ``within group'' term expressing variability within aggregations, and a ``between group'' term expressing variability across aggregations.
+
 	\begin{itemize}
@@
 
+% TODO:
+
+% - Think about whether we should also be looking at pairwise group exposure metrics?
+Is there an information-theoretic version of this?
+Maybe the information-loss associated with aggregating those two groups, that's not a bad guess.
+
 \end{document}

### 2016-10-05 3_local_info.R

@@ -1,4 +1,3 @@
-resolution <- 1/2
 
 library(sp, quietly = TRUE)
 library(compx, quietly = TRUE)
@@ -12,6 +11,10 @@ library(magrittr, quietly = TRUE)
 library(ggrepel, quietly = TRUE)
 library(grid, quietly = TRUE)
 library(rgdal, quietly = TRUE)
+
+races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
+resolution <- 1/2
+
 
 # Toy cities ----------
 
@@ -27,11 +30,18 @@ library(rgdal, quietly = TRUE)
 
 cities <- list.files('data/cities')
 
-
-
 if(!dir.exists('throughput/grids')){
 	dir.create('throughput/grids')
 }
+
+if(!dir.exists('throughput/grid_tracts')){
+	dir.create('throughput/grid_tracts')
+}
+
+
+
+
+
 
 
 
@@ -61,16 +71,36 @@ if(file.exists('throughput/info_cache.csv')){
 
 
 for(city in cities){
-	races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
 	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
-	tracts <- tracts[tracts@data$total > 0,]
+	
 	tracts@data$area <- tracts@data$ALAND / 1000^2	
 	tracts <- tracts[tracts@data$total / tracts@data$area > 50,] # at least 50 people per km^2
 	
-	radius = 1/sqrt(85 * 111) * resolution # (roughly 1 km after lat-lon conversion)
+	radius = 1/sqrt(85 * 111) * resolution # (roughly resolution km after lat-lon conversion)
 	xx = spsample(tracts, type="hexagonal", cellsize=radius)
 	print(paste0(city, ': ',  nrow(tracts@data), ' tracts || ', length(xx), ' grid cells'))
 	xxpl = HexPoints2SpatialPolygons(xx)
+	
+	local_information <- function(tracts, grid){
+		h <- function(i){
+			window <- tracts[xxpl[i,],]@data[,c(races, 'total', 'area')]
+			window <- window / (window$area) # total becomes density
+			c(mean(window$total), 
+			  4 * mutual_info(window[,races]) / resolution^2,  
+			  H(colSums(window[,races]) /sum(window[,races])), 
[…]

### 2016-10-06 5_figs.Rmd

--- 2016-07-03 5_figs.Rmd
@@
 checkerboard_illustration()
-ggsave('figs/checkerboard.png', width = 4, height = 1.5)
+ggsave('figs/checkerboard.pdf', width = 4, height = 1.5)
 ```
@@
 
-first <- method_illustration('Detroit', vmax = 6)
-second <- method_illustration('Philadelphia', vmax = 6)
-
-gA <- ggplotGrob(first)
-gB <- ggplotGrob(second)
-g <- gridExtra::cbind.gtable(gA, gB, size = "max")
-
-grid.draw(g)
-
-ggsave('figs/method_illustration.png', g, width = 8, height = 5)
+# first <- method_illustration('Detroit', vmax = 6)
+# second <- method_illustration('Philadelphia', vmax = 6)
+# 
+# gA <- ggplotGrob(first)
+# gB <- ggplotGrob(second)
+# g <- gridExtra::cbind.gtable(gA, gB, size = "max")
+# 
+# grid.draw(g)
+# 
+# ggsave('figs/method_illustration.pdf', g, width = 8, height = 5)
 
@@
 
-ggsave('figs/decomposability.png', width = 2.5, height = 1.5)
+ggsave('figs/decomposability.pdf', width = 2.5, height = 1.5)
 
@@
 cache %>%
-	ggplot(aes(x = I_XY, y = weighted_J)) + 
+	ggplot(aes(x = I_XY, y = J_XY)) + 
 	geom_point(color = 'grey') + 
-	theme_minimal() + 
+	theme_light() + 
 	ylab(expression(J(X,Y))) + 
@@
 
-ggsave('figs/mutual_fisher.png', width = 8, height = 4)
+ggsave('figs/mutual_fisher.pdf', width = 10, height = 6)
 
@@
 lm_eqn <- function(df){
-	m <- lm(weighted_J ~ log(density), data = df) 
-	eq <- substitute(italic(y) == a + b %.% ~'log' ~italic(rho)*","~~italic(r)^2~"="~r2, 
+	m <- lm(J_XY ~ log(density) + I_XY, data = df) 
+	eq <- substitute(italic(y) == a + b %.% 'I(X,Y)' + c %.% ~'log' ~italic(rho)*","~~italic(r)^2~"="~r2, 
 					 list(a = format(coef(m)[1], digits = 2), 
 					 	 b = format(coef(m)[2], digits = 2), 
+					 	 c = format(coef(m)[3], digits = 2), 
 					 	 r2 = format(summary(m)$adj.r.squared, digits = 3)))
@@
 cache %>% 
+	rename(`I(X,Y)` = I_XY) %>% 
 	# filter(!grepl('Los Angeles', city)) %>%
-	ggplot(aes(x = density, y = weighted_J)) +
+	ggplot(aes(x = density, y = J_XY)) +
 	geom_smooth(method = 'lm', se = FALSE, color = 'black', size = .4) +
@@
 
-ggsave('figs/density_fisher.png', width = 8, height = 4)
+ggsave('figs/density_fisher.pdf', width = 10, height = 6)
 ```
@@
 tab <- cache %>% 
-	select(area, population, H_Y, I_XY, J = weighted_J) %>% 
+	select(area, population, H_Y, I_XY, J = J_XY) %>% 
 	mutate(Density = population / area) %>% 
@@
 
-join_cols <- c('area', 'population', 'density', 'H_X', 'H_Y', 'local.H_Y', 'weighted_J', 'city')
+join_cols <- c('area', 'population', 'density',  'H_Y', 'J_XY', 'city')
 loss_curves <- loss_curves %>% left_join(info_summary[,join_cols])
@@
 
-ggsave('figs/example_cluster_map.png',  width = 8, height = 4)
+ggsave('figs/example_cluster_map.pdf',  width = 10, height = 6)
 
@@
 
-	I <- mutual_info(tracts@data[,c(columns, 'cluster')],group_col = 'cluster')
+	I <- mutual_info(tracts@data[,columns])
 
@@
 
-ggsave('figs/example_clusters.png')
+ggsave('figs/example_clusters.pdf', width = 10, height = 6)
 
@@
 
-ggsave('figs/example_clusters_detailed.png', width = 4, height = 2)
+ggsave('figs/example_clusters_detailed.pdf', width = 10, height = 6)
 
@@
 	filter(nclust == max(nclust)) %>% 
-	filter(weighted_J > 1.25 | weighted_J < .75)
+	filter(J_XY > 1.25 | J_XY < .75)
 
@@
 	aes(x = nclust, y = I_XY, group = city) +
-	geom_path(aes(color = weighted_J)) +
-	# geom_point(aes(color = weighted_J), size = .3) +
+	geom_path(aes(color = J_XY)) +
+	# geom_point(aes(color = J_XY), size = .3) +
 	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
@@
 
-ggsave('figs/loss_curves.png', width = 8, height = 4)
+ggsave('figs/loss_curves.pdf', width = 10, height = 6)
 
@@
 	geom_path(aes(color = I_max)) +
-	# geom_point(aes(color = weighted_J), size = .3) +
+	# geom_point(aes(color = J_XY), size = .3) +
 	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
@@
 	ggplot() +
-	aes(x = nclust, y = I_XY, color = weighted_J, group = city) +
+	aes(x = nclust, y = I_XY, color = J_XY, group = city) +
 	geom_path() +
@@
 
-ggsave('figs/normed_loss_curves.png', width = 8, height = 4)
+ggsave('figs/normed_loss_curves.pdf', width = 10, height = 6)
 ```
@@
 		group_by(city) %>%
-		arrange(nclust) %>%
+		dplyr::arrange(nclust) %>%
 		summarise(corner = max(I_XY) * log(max(nclust)),
@@
 		mutate(AUC = area / corner) %>%
-		arrange(desc(AUC)) 
+		dplyr::arrange(desc(AUC)) 
 }
@@
 summary <- compute_AUC(loss_curves, adjust_thresh = 0) 
-
 summary <- summary %>% left_join(cache, by = c('city' = 'city'))
@@
 # %>%
-# 	left_join(info_summary[,c('city', 'weighted_J', 'I_XY')]) %>%
+# 	left_join(info_summary[,c('city', 'J_XY', 'I_XY')]) %>%
 # 	filter(I_XY > 0)
@@
 
-filled <- sub_curves[,c('nclust','I_XY', 'city')]
+filled <- sub_curves[,c('nclust','I_XY', 'city')] %>% 
+  select(city, I_XY, nclust) %>% 
+  ungroup()
 
 corner <- filled %>% group_by(city) %>%
-	summarise(nclust = max(nclust),
-			  I_XY = 0)
+	summarise(I_XY = 0,
+	          nclust = max(nclust))
 
-filled <- rbind(corner, filled) %>%
-	arrange(city, nclust, desc(I_XY))
+filled <- rbind(corner, filled) %>%  # problem is right here, the columns don't add up
+	dplyr::arrange(city, nclust, desc(I_XY))
 
@@
 	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
-	geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(weighted_J,2))), data = sub_summary) +
+	geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
 	facet_wrap(~city) + 
@@
 
-ggsave('figs/AUC_illustration.png', width = 8, height = 4)
+ggsave('figs/AUC_illustration.pdf', width = 10, height = 6)
 ```
@@
 
-summary[,c('I_XY', 'weighted_J', 'AUC', 'adj_AUC')] %>% 
+summary[,c('I_XY', 'J_XY', 'AUC', 'adj_AUC')] %>% 
 	round(2) %>% 
 	cbind(summary$city) %>% 
-	select(city = `summary$city`, I_XY, weighted_J, AUC, adj_AUC) %>%
+	select(city = `summary$city`, I_XY, J_XY, AUC, adj_AUC) %>%
 	datatable(options = list(pageLength = 10), rownames = FALSE,
@@
 	aes(x = I_XY, y = AUC) +
-	geom_point(aes(color = weighted_J)) +
+	geom_point(aes(color = J_XY)) +
 	geom_smooth(color = 'grey', se = F) +
@@
 
-ggsave('figs/info_performance.png', width = 8, height = 4)
+ggsave('figs/info_performance.pdf', width = 10, height = 6)
 
@@
 	aes(x = I_XY, y = adj_AUC) +
-	geom_point(aes(color = weighted_J)) +
+	geom_point(aes(color = J_XY)) +
 	geom_smooth(color = 'grey', se = F) +
@@
 
-ggsave('figs/info_performance_adjed.png', width = 8, height = 4)
+ggsave('figs/info_performance_adjed.pdf', width = 10, height = 6)
 
@@
 		   `I(X,Y)` = I_XY / max(I_XY),
-		   `J(X,Y)` = weighted_J / max(weighted_J),
+		   `J(X,Y)` = J_XY / max(J_XY),
 		   `H(Y)` = H_Y / max(H_Y)) %>%
 	filter(I_XY > 0) %>%
-	lm(AUC ~ `I(X,Y)` + `J(X,Y)`, data = .) %>%
+	lm(AUC ~ `H(Y)` + `I(X,Y)` + `J(X,Y)`, data = .) %>% 
 	stargazer(flip = T, 
@@
 			  style = 'all2',
-			  ci = T,covariate.labels = c('$I(X,Y)$','$J(X,Y)$', 'Intercept'),
-			  table.layout = '-=-!=!dtas',
+			  ci = T,covariate.labels = c('$H(Y)$', '$I(X,Y)$','$J(X,Y)$', 'Intercept'),
+			  # table.layout = '-=-!=!dtas',
 			  float = F) %>% 
 	write(file = 'figs/adjusted_regression.txt')
+```
+
+```{r}
+
+summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
+		   `I(X,Y)` = I_XY / max(I_XY),
+		   `J(X,Y)` = J_XY / max(J_XY),
+		   `H(Y)` = H_Y / max(H_Y),
+		   density = density / max(density)) %>%
+	lm(`J(X,Y)` ~ log(density) + `I(X,Y)`, data = .) %>% 
+	summary()
+
+```
+
+```{r}
+cities <- c('Detroit', 'Philadelphia')
+
+tab <- data.frame(city = character(), percent_black = numeric(), H_Y = numeric(), I_XY = numeric(), J_XY = numeric())
+
+for(city in cities){
+  tracts <- readOGR(dsn = paste0('data/cities/',city), 
+  	                         layer = 'geo', 
+  	                         verbose = FALSE)
+
+  grid_tract <- read_csv(file = paste0('throughput/grid_tracts/', city, '.csv'))
+  grid <- readOGR(dsn = paste0('throughput/grids/',city), 
+  	                         layer = 'grid', 
+  	                         verbose = FALSE)
+
+  df <- tracts@data %>%
+    mutate(percent_black = Black / total) %>% 
+    tbl_df
+
+  tracts %>% 
+    fortify(region = 'GEOID') %>% 
+    tbl_df() %>% 
+    left_join(df, by = c('id' = 'GEOID')) %>% 
+    ggplot(aes(x = long, y = lat, fill = percent_black, group = group)) + 
+    geom_polygon() + 
+    viridis::scale_fill_viridis(name = 'Percent Black', labels = scales::percent) + 
+    ggtitle(city) + 
+    ggthemes::theme_map() + 
+    theme(legend.justification=c(0,0), legend.position=c(.7,0)) + 
+    theme(plot.title = element_text(size=10, 
+                                  vjust=1),
+          legend.title = element_text(size = 8))
+
+  ggsave(paste0('figs/', city, '_percent_black.pdf'), width = 10, height = 10, units = 'cm')
+
+  tracts@data <- tracts@data %>% 
+    mutate(non_black = Hispanic + Other + White + Asian)
+
+  columns = c('non_black', 'Black')
+  stuff <- compx::info_analysis(tracts, columns = columns, resolution = NULL, grid_tract = grid_tract, grid = grid)
+
+  stuff$grid %>% 
+    fortify(region = 'id') %>% 
+    left_join(stuff$grid@data[,c('id', 'pop', 'info')], by = c('id' = 'id')) %>% 
+    ggplot(aes(x = long, y = lat, fill =  info, group = group)) + 
+    geom_polygon() + 
+    viridis::scale_fill_viridis(name = expression(J[Y](x))) + 
+    ggthemes::theme_map() +
+    ggtitle(city) + 
+    theme(legend.justification=c(0,0), legend.position=c(.7,0),
+          legend.title = element_text(size = 8)) + 
+    theme(plot.title = element_text(size=10, 
+                                  vjust=1))
+
+  # paste0('figs/', city, '_grid.pdf')
+  ggsave(paste0('figs/', city, '_grid.pdf'), width = 10, height = 10, units = 'cm')
+
+  percent_black = sum(tracts@data$Black) / sum(tracts@data$total)
+
+  new_row <- data.frame(city = city, percent_black = percent_black, H_Y = stuff$H_Y, I_XY = stuff$I_XY, J_XY = stuff$J_XY)
+  tab <- rbind(tab, new_row)
+}
+
+tab[,-1] <- round(tab[,-1],2)
+
+tab %>% write_csv('figs/comparative_summary.csv')
+
 ```

### 2016-10-06 revision.tex

@@
 \usepackage{pc_math}
+\usepackage{csvsimple}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 	\begin{figure}
-		\includegraphics[width=\textwidth]{figs/checkerboard.png}
+		\includegraphics[width=\textwidth]{figs/checkerboard.pdf}
 		\caption{Information and the checkerboard}. \label{fig:checkerboard}
@@
 			\begin{figure}
-				\caption{Spatial analysis of racial diversity in Detroit (left) and Philadelphia (right).
-Top: Percentage of black residents in Detroit (left) and Philadelphia (right).
-Middle: Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of $I(X,Y | X \in B_i)$.
-Bottom: Estimated contours of constant $I(X,Y | X \in B_i)$; high values correspond to spaces of demographic transition.
-Very bottom: Numerical summary, including the three information measures.} \label{fig:detroit_philly}
+				\includegraphics[width = .5\textwidth]{figs/Detroit_percent_black.pdf}
+				\includegraphics[width = .5\textwidth]{figs/Detroit_grid.pdf} \\
+				\includegraphics[width = .5\textwidth]{figs/Philadelphia_percent_black.pdf}
+				\includegraphics[width = .5\textwidth]{figs/Philadelphia_grid.pdf} \\
+
+				\centering
+				\begin{tabular}{l | c c c c}
+					\bfseries City & Percent Black & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$  \\\hline
+					\csvreader[late after line=\\]{figs/comparative_summary.csv}{}
+					{\csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv}
+				\end{tabular}
+
+				\caption{Spatial analysis of racial diversity in Detroit (top) and Philadelphia (middle).
+Left: Percentage of black residents in Detroit (left) and Philadelphia (right).
+Right: Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of $I(X,Y | X \in B_i)$.
+Bottom: Numerical summary, including the three information measures.} \label{fig:detroit_philly}
 			\end{figure}
@@
 			\begin{figure}
-				\includegraphics[width=\textwidth]{figs/mutual_fisher.png}
+				\includegraphics[width=\textwidth]{figs/mutual_fisher.pdf}
 				\caption{Comparative global diversity measures in major U.S.
@@
 A diversity measure satisfies additive spatial decomposition when, if spatial units are aggregated into groups, the measure can be expressed as a sum of a ``within group'' term expressing variability within aggregations, and a ``between group'' term expressing variability across aggregations.
+The mutual information $I(X,Y)$ satisfies this property due to the following information-theoretic identity.
+Let $C$ be a random variable with the property that $H(C|X) = 0$; i.e.
+$X$ completely determines $C$.
+We can think of $C$ as group labels for the spatial locations $X$.
+Then, the chain rule of information implies that
+	\begin{equation}
+		I(X,Y) = \underbrace{I(C,Y)}_{\text{between-group}} + \underbrace{I(X,Y|C)}_{\text{within-group}}. \label{eq:chain_rule}
+	\end{equation}
+	Equation \eqref{eq:chain_rule} expresses the following intuitive information decomposition: the information you have if I tell you first a resident's approximate residence $C$ and subsequently their exact residence $X$ is just the same as the information you would have if I simply told you $X$ directly.
+
+	Equation \eqref{eq:chain_rule} also motivates a natural algorithm for identifying sociospatial structure.
+The core idea is that a clustering $C$ is ``good'' when it captures most of the intrinsic information in the system in the between-groups term $I(C,Y)$.
+This idea is analogous to between-group sum-of-squares maximization in classical ANOVA methods.
+We therefore seek to maximize $I(C,Y)$.
+In full generality, this is a challenging discrete optimization problem that may not be computationally tractable.
+We can, however, construct a ``greedy'' algorithm which, though suboptimal, leads to satisfactory results.
+At each stage of the algorithm, we consider the problem of choosing a pair of locales $\{i*, j*\}$ to aggregate together.
+A good choice of locales leads to minimal information loss, which we may quantify as follows.
+The mutual information may also be written:
+	\begin{align}
+	 	I(X,Y) &= \sum_k p(X = k) D[p(Y|X = k\|p(Y)] 
+	\end{align} 
+	The contribution of two locales $i$ and $j$ to this sum is 
+	\begin{align}
+		I_{i,j}(X,Y) &\triangleq \sum_{k \in \{i,j\}} p(X = k)D[p(Y|X = k\|p(Y)] 
+	\end{align}
+	If we replace $i$ and $j$ with an aggregated local $i\cup j$, its contribution to the mutual information is 
+	\begin{align}
+		I_{i\cup j}(X,Y) = p(X \in \{i,j\}) D[p(Y|X \in \{i,j\}) \| p(Y)]\;.
+	\end{align}
+	The  loss associated with aggregating $i$ and $j$ is the difference between these two terms: 
+	\begin{align}
+		d(i,j) \triangleq I_{i,j}(X,Y) - I_{i \cup j}(X,Y).  
+	\end{align}
+	Like the KL divergence $D$, the function $d$ can be viewed as a measure of distance between probability distributions $p(Y|X = i)$ and $p(Y|X = j)$.
+Unlike the KL divergence, $d$ is symmetric in $i$ and $j$.
 
@@
 		\item Construct the algorithm. 
+		\item Prove the triangle inequality
 		\item Show algorithm results for one or two cities; show how knowing the clusters can lead to neat data analysis. 

### 2016-10-06 3_make_grids.R
removed make_grids

@@ -1,75 +1,0 @@
-library(sp, quietly = TRUE)
-library(compx, quietly = TRUE)
-library(acs, quietly = TRUE)
-# library(plyr, quietly = TRUE)
-library(dplyr, quietly = TRUE)
-library(readr, quietly = TRUE)
-library(ggmap, quietly = TRUE)
-library(tidyr, quietly = TRUE)
-# library(magrittr, quietly = TRUE)
-
-# library(grid, quietly = TRUE)
-library(rgdal, quietly = TRUE)
-
-resolution <- 0.5
-
-cities <- list.files('data/cities')
-# cities <- c('Boston')
-
-
-if(!dir.exists('throughput')){
-  dir.create('throughput')
-}
-
-if(!dir.exists('throughput/grids')){
-	dir.create('throughput/grids')
-}
-
-if(!dir.exists('throughput/grid_tracts')){
-	dir.create('throughput/grid_tracts')
-}
-
-for(city in cities){
-	
-	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
-	
-	tracts@data$area <- tracts@data$ALAND / 1000^2	
-	tracts <- tracts[tracts@data$total / tracts@data$area > 50,] # at least 50 people per km^2
-	
-	radius = 1/sqrt(85 * 111) * resolution # (roughly resolution km after lat-lon conversion)
-	xx = spsample(tracts, type="hexagonal", cellsize=radius)
-	print(paste0(city, ': ',  nrow(tracts@data), ' tracts || ', length(xx), ' grid cells'))
-	xxpl = HexPoints2SpatialPolygons(xx)
-	xxpl <- as(xxpl, 'SpatialPolygonsDataFrame')
-	cell_area <- rgeos::gArea(xxpl[1])
-	
-	d <- data_frame(cell = character(), tract = numeric(), area = numeric())
-	
-	for(x in row.names(xxpl@data)){ 
-		window <- tracts[xxpl[x,],]
-		poly_i <- raster::intersect(xxpl[x,], window)
-		# areas <- rgeos::gArea(pi, byid = T)
-		areas <- sapply(poly_i@polygons, function(y) y@area)
-		full_areas <- sapply(window@polygons, function(y) y@area)
-		d <- rbind(d, data_frame(cell = x, tract = window@data$GEOID, area = areas, full_area = full_areas))
-	}
-	
[…]

### 2016-10-06 revision.tex
latest draft

@@
 	\end{align} 
-	The contribution of two locales $i$ and $j$ to this sum is 
+	The contribution of locale $i$ to this sum is 
 	\begin{align}
-		I_{i,j}(X,Y) &\triangleq \sum_{k \in \{i,j\}} p(X = k)D[p(Y|X = k\|p(Y)] 
+		I_{i}(X,Y) &\triangleq \sum_{k \in \{i\}} p(X = i)D[p(Y|X = i\|p(Y)] 
 	\end{align}
@@
 	\end{align}
-	The  loss associated with aggregating $i$ and $j$ is the difference between these two terms: 
+	The  loss associated with aggregating $i$ and $j$ is the difference between individual contributions and their aggregated contribution:
 	\begin{align}
-		d(i,j) \triangleq I_{i,j}(X,Y) - I_{i \cup j}(X,Y).  
+		d(i,j) \triangleq I_{i}(X,Y) + I_{j}(X,Y) - I_{i \cup j}(X,Y).  
 	\end{align}
 	Like the KL divergence $D$, the function $d$ can be viewed as a measure of distance between probability distributions $p(Y|X = i)$ and $p(Y|X = j)$.
-Unlike the KL divergence, $d$ is symmetric in $i$ and $j$.
+Unlike the KL divergence, $d$ is symmetric in $i$ and $j$; it is not a true metric because it does not satisfy the triangle inequality.
+It can nevertheless be used to motivate a form of agglomerative hierarchical clustering, where at each phase of the algorithm we compute
+	\begin{equation}
+		(i^*m, j^*) = \argmin_{i \text{neighbors} j} \; d(i,j)\;
+	\end{equation}
+	and combine $i^*$ and $j^*$ into a single new group.
+We repeat this procedure until only one group remains.
+The optimizatin over $i$ and $j$ ensures that the groupings so defined approximate maximum between-group information partitions, and the constraint that $i$ must neighbor $j$ ensures that the structure so determined is spatially contiguous.
+Since this is a greedy algorithm, it possesses no guarantees of solution optimality, but in practice its performance leads to intuitive, demographically-coherent regions.
+
+	To demonstrate the power of this approach for data analysis,
 

### 2016-10-06 5_figs.Rmd
merge prep

@@
 	xlab(expression(I(X,Y))) + 
-	ggtitle('Mutual information and mean local information in major US cities') +
+	# ggtitle('Mutual information and mean local information in major US cities') +
 	geom_point(data = highlight, color = 'firebrick') + 
-	geom_text_repel(data = highlight, aes(label = city))
+	geom_text_repel(data = highlight, aes(label = city)) + 
+  theme(legend.title = element_text(size = 8),
+        plot.title = element_text(size=10, vjust=1))
 
-ggsave('figs/mutual_fisher.pdf', width = 10, height = 6)
+ggsave('figs/mutual_fisher.pdf', width = 7, height = 4)
 
@@
 
-  ggsave(paste0('figs/', city, '_percent_black.pdf'), width = 10, height = 10, units = 'cm')
+  ggsave(paste0('figs/', city, '_percent_black.pdf'), width = 7, height = 7, units = 'cm')
 
@@
     geom_polygon() + 
-    viridis::scale_fill_viridis(name = expression(J[Y](x))) + 
+    viridis::scale_fill_viridis(name = expression(J[Y](x)), limits = c(0, .5)) + 
     ggthemes::theme_map() +
@@
     theme(legend.justification=c(0,0), legend.position=c(.7,0),
-          legend.title = element_text(size = 8)) + 
-    theme(plot.title = element_text(size=10, 
-                                  vjust=1))
+          legend.title = element_text(size = 8),
+          plot.title = element_text(size=10, vjust=1))
 
   # paste0('figs/', city, '_grid.pdf')
-  ggsave(paste0('figs/', city, '_grid.pdf'), width = 10, height = 10, units = 'cm')
+  ggsave(paste0('figs/', city, '_grid.pdf'), width = 7, height = 7, units = 'cm')
 
@@
 
+tab %>% mutate(percent_black = percent_black * 100)
+
 tab %>% write_csv('figs/comparative_summary.csv')

### 2016-10-06 4_cluster.R

@@ -5,7 +5,6 @@ library(rgeos, quietly = TRUE)
 library(ggplot2, quietly = TRUE)
 library(Matrix, quietly = TRUE)
 library(maptools, quietly = TRUE)
-# library(profvis, quietly = TRUE)
 library(readr, quietly = TRUE)
 library(ggrepel, quietly = TRUE)
 library(caTools, quietly = TRUE)

### 2016-10-06 tex/paper/methods.tex

@@
 	\begin{description}
-		\item[Organizational Equivalence:] Both the theoretical definition \eqref{eq:def_J} and the procedure to compute it from tract data remain unchanged when a tract is subdivided into smaller tracts, each of which with identical demographic structure.
-		\item[Size and Density Invariance:] Since $J(X,Y)$ is completely determined by the marginal distribution $p(X,Y)$, it is invariant under changes in population density. 
-		\item[Additive Group Decomposability:] When demographic groups are aggregated into super-groups, the chain rule of mutual information applied to the demographic variable $Y$ provides an an additive decomposition of the form $J(X,Y) = J(X,G) + J(X,Y|G)$, which can be interpreted as the sum of a between-groups term and a within-groups term as needed.
+		\item[Organizational Equivalence:] Both the theoretical definition \eqref{eq:def_J} and the procedure to compute it from discrete data remain unchanged when a tract is subdivided into smaller tracts, each of which with identical demographic structure.
+		\item[Size and Density Invariance:] Since $J(X,Y)$ is completely determined by the probability measure $p(X,Y)$, it is invariant under global changes in population density. 
+		\item[Additive Group Decomposability:] When demographic groups are aggregated into super-groups, the chain rule of mutual information applied to the demographic variable $Y$ provides an an additive decomposition of the form $J(X,Y) = J(X,G) + J(X,Y|G)$ for the mean local information.
+This can be interpreted as a sum of between-groups and within-gropus terms as required.
 		\item[Scale Interpretability:] We have $J(X,Y) = 0$ if and only if $p(Y|X)$ is constant on each connected component of the metric space $M$.
@@
 the city is divided by a river), demographics must be constant in space on each side of the division.
-$J(X,Y)$ does not satisfy the additional condition of achieving its maximum value when all locales are monoracial, but this point only emphasizes that $J(X,Y)$ and $I(X,Y)$ should be read jointly.
+$J(X,Y)$ does not achieve its maximum value when all locales are monoracial, but this point only emphasizes that $J(X,Y)$ and $I(X,Y)$ should be read jointly.
 When all locales are monoracial, $I(X,Y)$ achieves its maximum value and $J(X,Y) = 0$.
@@
 In computational practice, $J(X,Y)$ is indeed sensitive to the boundaries supplied with the data; however, \eqref{eq:approx} guarantees that this sensitivity vanishes as the resolution grows sufficiently small.
-No alternative measure can overcome limited data resolution, and, to our knowledge, no other measure comes with the same assurance of resolution-independence in the limit.
+No alternative measure can overcome limited data resolution, and, to our knowledge, no other measure comes with the same assurance of limiting resolution independence.
 		\item[Transfers and Exchanges:] The core idea of the principles of transfers and exchanges is that population permutations that tend to ``smooth out'' differences should reduce segregation measures.
-However, the concrete principles as stated in \cite{Reardon2002} do not hold of $J(X,Y)$.
-The core point is that $J(X,Y)$ compares distributions to those nearby, so whether or not a transfer or exchange reduces $J(X,Y)$depends not only on the two regions involved in the transfer or exchange, but also the regions surrounding them.
-$I(X,Y)$ does satisfy the two principles.
-% Consider trying to demonstrate the spatial transfer principle for the computational version.
+No strict such condition holds for $J(X,Y)$; however, this is comparatively not a large limitation, as most of the measures considered in \cite{Reardon2004} also fail to meet rigorous transfer and exchange principles.
+$J(X,Y)$ does satisfy the following useful and interpretable criterion: when a transfer makes all affected areas more closely resemble their immediate neighbors, $J(X,Y)$ decreases.
+Notably, these transfers can tend to either increase or reduce spatial evenness, and $I(X,Y)$ may increase or decrease accordingly.
 		\item[Additive Spatial Decomposability:] This criterion states that ``If X spatial subareas are aggregated into Y larger spatial areas, then a segregation measure should be decomposable into a sum of within- and between-area components'' \cite{Reardon2004}, page 136.
@@
 On the one hand, an isotropic measure is simple to compute, simple to interpret, and easily generalized.
-On the other, boundaries and infrastructure unquestionable shape interactions between people, and are clearly relevant in segregation studies.
+On the other, boundaries and infrastructure unquestionably shape interactions between people, and are clearly relevant in segregation studies.
 We note, however, that most urban anisotropies exist on the scale of a few hundred meters at most.
@@
 	\end{description}
+
+	\begin{figure}
+		\begin{tabular}{l | c c c}
+		Criterion & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
+		\hline
+		Scale Interpretability & \checkmark & \checkmark & \checkmark \\
+		Density Invariance & & & \\
+		Compositional Invariance & & & \\
+		Transfers & & & \\
+		Additive Organizational Decomposability & & & \\
+		Additive Spatial Decomposability & & & 
+		\end{tabular}
+	\end{figure}
 

### 2016-10-08 1_get_states.R
automated aggregation of racial categories

@@ -3,7 +3,7 @@ library(acs, quietly = TRUE)
 library(dplyr, quietly = TRUE)
 library(tidyr, quietly = TRUE)
 library(stringr, quietly = TRUE)
-
+library(readr)
 
 # get list of states
 
@@ -13,7 +13,6 @@ cities$state <- word(cities$state, -1)
 if(!dir.exists('data')){
 	dir.create('data')
 }
-
 
 if(!dir.exists('data/states')){
 	dir.create('data/states')
@@ -35,34 +34,25 @@ f <- function(state){
 	
 	race <- cbind(data.frame(race@geography),
 				  data.frame(race@estimate)) %>%
-		tbl_df() %>%
+		tbl_df() 
+	
+	race_lookup <- read_csv('assumptions/race_lookup.csv')
+	
+	race <- race %>% 
+		gather(key = 'original_race', value = 'n', -(NAME:blockgroup)) %>% 
+		left_join(race_lookup, by = c('original_race' = 'original')) %>% 
+		group_by(NAME, state, county, tract, blockgroup, aggregated) %>% 
+		summarise(n = sum(n)) %>% 
+		ungroup() %>% 
+		filter(!is.na(aggregated)) %>% 
+		spread(key = aggregated, value = n, fill = 0)
+	
+	race <- race %>% 
 		mutate(GEOID = paste0(str_pad(state, 2, 'left', pad = '0'),
 							  str_pad(county, 3, 'left', pad = '0'),
 							  str_pad(tract, 6, 'left', pad = '0'),
 							  str_pad(blockgroup, 1, 'left', pad = '0')))
 	
-	race$Hispanic <- race[names(race)[!grepl('.Not.Hispanic.or.Latino.',names(race))]] %>%
-		select(-(`NAME`:`Hispanic.or.Latino.by.Race..Hispanic.or.Latino.`)) %>%
-		select(-GEOID) %>%
-		rowSums()
-	
-	others <- c('Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..American.Indian.and.Alaska.Native.alone',
-				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Native.Hawaiian.and.Other.Pacific.Islander.alone',
-				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Some.other.race.alone',
-				'Hispanic.or.Latino.by.Race..Not.Hispanic.or.Latino..Two.or.more.races.',
[…]

### 2016-10-23 5_figs.Rmd
font update!

--- 2016-10-06 5_figs.Rmd
@@
 library(magrittr)
+library(rgeos)
+library(extrafont)
+loadfonts()
 
@@
 ```{r, fig.height=3}
-checkerboard_illustration()
+g <- checkerboard_illustration() 
+g + theme(text=element_text(size=16, family="Palatino"))
 ggsave('figs/checkerboard.pdf', width = 4, height = 1.5)
+embed_fonts("figs/checkerboard.pdf", outfile="figs/checkerboard.pdf")
 ```
@@
 		ylab('') +
-		theme(legend.justification=c(.8,0),
+		theme(text=element_text(size=16, family="Palatino"),
+		    legend.justification=c(.8,0),
 			  legend.position=c(1,0),
@@
 highlight <- cache %>%
-	filter(I_XY > .4)
+	filter(I_XY > .45 | I_XY < .22)
 
@@
 	xlab(expression(I(X,Y))) + 
-	ggtitle('Mutual information and mean local information in major US cities') +
+	# ggtitle('Mutual information and mean local information in major US cities') +
 	geom_point(data = highlight, color = 'firebrick') + 
-	geom_text_repel(data = highlight, aes(label = city))
+	geom_text_repel(data = highlight, aes(label = city), family = 'Palatino' ) + 
+  theme(text=element_text(size=16, family="Palatino"),
+        legend.title = element_text(size = 8),
+        plot.title = element_text(size=10, vjust=1))
 
-ggsave('figs/mutual_fisher.pdf', width = 10, height = 6)
+ggsave('figs/mutual_fisher.pdf', width = 7, height = 4)
+embed_fonts("figs/mutual_fisher.pdf", outfile="figs/mutual_fisher.pdf")
 
@@
 	geom_smooth(method = 'lm', se = FALSE, color = 'black', size = .4) +
-	annotate('text', x = 3000, y = .2, label = lm_eqn(cache), parse = T) + 
+	annotate('text', x = 300, y = .12, label = lm_eqn(cache), parse = T, family = 'Palatino') + 
 	geom_point(color = 'grey') + 
 	geom_point(color = 'firebrick', data = highlight) + 
-	geom_text_repel(aes(label = city), size = 3, data = highlight) +
+	geom_text_repel(aes(label = city), size = 3, data = highlight, family = 'Palatino' ) +
 	scale_x_continuous(trans = 'log10') +
 	# scale_y_continuous(trans = 'log10') +
-	theme_minimal() +
+	theme_light() +
+  theme(text=element_text(size=16, family="Palatino")) + 
 	xlab(expression(rho~(population/km^{2}))) +
 	ylab(expression(J(X,Y))) +
-	annotation_logticks(size = .1, sides = 'b') +
-	ggtitle('Dependence of spatial complexity on population density')
+	annotation_logticks(size = .1, sides = 'b')
 
-ggsave('figs/density_fisher.pdf', width = 10, height = 6)
+ggsave('figs/density_fisher.pdf', width = 7, height = 4)
+embed_fonts("figs/density_fisher.pdf", outfile="figs/density_fisher.pdf")
+
 ```
@@
 
-cluster_map <- function(city, nclusters = 10){
-	tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
+cluster_map <- function(tracts_name, city, nclusters = 10){
+	tracts <- readOGR(dsn = paste0('data/cities/',tracts_name), layer = 'geo', verbose = FALSE)
 	tracts <- tracts[sum(tracts@data[,columns]) != 0,]
@@
 		geom_polygon(aes(x = long, y = lat, group = group, fill = factor(cluster))) + 
-		scale_fill_brewer(palette = 'Set3', name = NULL, guide = guide_legend(nrow = 1), labels = LETTERS[1:nclusters]) + 
+		scale_fill_brewer(palette = 'Set3', name = NULL, guide = 'none') + 
 		# ggtitle(paste0('Example clustering for ', city)) + 
@@
 
-cluster_map('Detroit', nclusters = 6)
+columns = c('Black', 'White', 'Asian', 'Hispanic', 'Other')
 
-ggsave('figs/example_cluster_map.pdf',  width = 10, height = 6)
+cluster_map('Detroit', 'Detroit', nclusters = 6)
+
+ggsave('figs/example_cluster_map.pdf',  width = 7, height = 7)
 
@@
 	mutate(race = factor(race, levels = c('Other', 'Asian', 'Hispanic', 'Black', 'White'))) %>% 
-	ggplot(aes(y = race, x = cluster, fill = p)) + 
-	geom_tile(color = 'black') + 
+  group_by(cluster) %>% 
+  mutate(alpha = p / sum(p),
+         total = sum(p)) %>% 
+	ggplot(aes(y = race, x = cluster, fill = cluster)) + 
+  scale_fill_brewer(palette = 'Set3', name = NULL, guide = 'none', labels = LETTERS[1:nclusters]) + 
+	geom_tile(color = 'black',  aes(alpha = alpha), size = 1) + 
 	geom_text(aes(
-		label = ifelse(p < 0.005, "", 100 * round(p,2)))) +
+		label = ifelse(p < 0.005, "", 100 * round(p,2))), size = 8, family = 'Palatino') + 
+  scale_alpha_continuous(guide = 'none', limits = c(0, .3)) + 
 	theme_minimal() + 
-	theme(axis.ticks.x = element_blank(),
+	theme(text=element_text(size=16, family="Palatino"),
+		  axis.ticks.x = element_blank(),
 		  axis.title = element_blank(),
 		  legend.position="bottom",
-		  legend.title = element_blank()) +
-	scale_fill_continuous(low = 'white', high = 'steelblue', limits = c(0, NA), guide = FALSE) 
-	# ggtitle(paste0('Example clusters'))
+		  legend.title = element_blank(),
+		  line = element_blank(),
+		  text = element_text(size = 25))
 
-ggsave('figs/example_clusters_detailed.pdf', width = 10, height = 6)
+ggsave('figs/example_clusters_detailed.pdf', width = 8, height = 6)
+embed_fonts("figs/example_clusters_detailed.pdf", outfile="figs/example_clusters_detailed.pdf")
+
+make_cluster_groups('Detroit', 6) %>% 
+  spread(key = race, value = p, fill = 0) %>%
+  select(Asian, Black, Hispanic, Other, White) %>% 
+  as.matrix() %>% 
+  mutual_info()
+
+# 
+# 
+# tracts@data %>% group_by(cluster) %>% 
+#   summarise(Asian = sum(Asian), 
+#             Hispanic = sum(Hispanic), 
+#             Other = sum(Other), 
+#             White = sum(White), 
+#             Black = sum(Black)) %>%
+#   select(-cluster) %>% 
+#   as.matrix() %>% mutual_info()
 
@@
 
+process_loss_curves <- function(loss_curves){
   loss_curves %>%
@@
   	ggtitle('Normalized loss curves for spatially-constrained information-theoretic clustering')
-
+  }
 ggsave('figs/normed_loss_curves.pdf', width = 10, height = 6)
@@
 
-cities <- c('Atlanta','Boston')
+binary_loss_curves <- read_csv('throughput/loss_curves.csv') %>%
+  group_by(city) %>% 
+	arrange(desc(I_XY)) %>% 
+	mutate(I_XY = max(I_XY) - I_XY,
+		   nclust = row_number())
+
+cities <- c('Atlanta','Philadelphia', 'Detroit')
 
@@
 
-summary <- compute_AUC(loss_curves, adjust_thresh = 0) 
-summary <- summary %>% left_join(cache, by = c('city' = 'city'))
-
-# %>%
-# 	left_join(info_summary[,c('city', 'J_XY', 'I_XY')]) %>%
-# 	filter(I_XY > 0)
+summary <- compute_AUC(binary_loss_curves, adjust_thresh = 0.005) 
+summary <- summary %>% left_join(cache, by = c('city' = 'city')) %>%
+	left_join(info_summary[,c('city', 'J_XY', 'I_XY')]) %>%
+	filter(I_XY > 0)
 
@@
 	theme_minimal() +
+  theme(text=element_text(size=16, family="Palatino")) + 
 	annotation_logticks(size = .1, sides = 'b', scaled = T) + 
@@
 	geom_polygon(aes(nclust, I_XY, fill = city), data = filled, alpha = .2) + 
-	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
-	geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
+	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary, family = 'Palatino') + 
+	# geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
 	facet_wrap(~city) + 
@@
 
-ggsave('figs/AUC_illustration.pdf', width = 10, height = 6)
+ggsave('figs/AUC_illustration.pdf', width = 10, height = 4)
+embed_fonts('figs/AUC_illustration.pdf', outfile = 'figs/AUC_illustration.pdf')
 ```
@@
 ```{r}
-summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
+summary %>% mutate(AUC = AUC / max(AUC),
 		   `I(X,Y)` = I_XY / max(I_XY),
@@
 ```{r}
-cities <- c('Detroit', 'Philadelphia')
+cities <- c('Philadelphia', 'Atlanta')
 
@@
 
+  if(city == 'Atlanta'){
+    guide = 'none'
+  }else{
+    guide = guide_colorbar(barwidth = 1, barheight = 4, label.theme = element_text(size = 7, angle = 0, family = 'Palatino'))
+  }
+
+  this_city <- city
+  J_XY <- summary %>% filter(city == this_city) %>% 
+    select(J_XY) %>% 
+    as.numeric() %>% 
+    round(2)
+
   tracts %>% 
@@
     geom_polygon() + 
-    viridis::scale_fill_viridis(name = 'Percent Black', labels = scales::percent) + 
-    ggtitle(city) + 
+    viridis::scale_fill_viridis(name = 'Percent Black', 
+                                labels = scales::percent, 
+                                option = 'magma', 
+                                limits = c(0,1),
+                                guide = guide) + 
+    ggtitle(paste0(city, ': J(X,Y) = ', J_XY)) + 
     ggthemes::theme_map() + 
-    theme(legend.justification=c(0,0), legend.position=c(.7,0)) + 
-    theme(plot.title = element_text(size=10, 
+    theme(legend.justification=c(0,0), legend.position=c(.65,0)) + 
+    theme(text=element_text(size=16, family="Palatino"),
+          plot.title = element_text(size=10, 
                                   vjust=1),
@@
 
-  ggsave(paste0('figs/', city, '_percent_black.pdf'), width = 10, height = 10, units = 'cm')
+  ggsave(paste0('figs/', city, '_percent_black.pdf'), width = 7, height = 7, units = 'cm')
+  embed_fonts(paste0('figs/', city, '_percent_black.pdf'), outfile = paste0('figs/', city, '_percent_black.pdf'))
 
@@
     geom_polygon() + 
-    viridis::scale_fill_viridis(name = expression(J[Y](x))) + 
+    viridis::scale_fill_viridis(name = expression(J[Y](x)), 
+                                limits = c(0, .5), 
+                                option = 'magma',
+                                guide = guide) + 
     ggthemes::theme_map() +
     ggtitle(city) + 
-    theme(legend.justification=c(0,0), legend.position=c(.7,0),
-          legend.title = element_text(size = 8)) + 
-    theme(plot.title = element_text(size=10, 
-                                  vjust=1))
+    theme(text=element_text(size=16, family="Palatino"),
+          legend.justification=c(0,0), legend.position=c(.65,0),
+          legend.title = element_text(size = 8),
+          plot.title = element_text(size=10, vjust=1))
 
   # paste0('figs/', city, '_grid.pdf')
-  ggsave(paste0('figs/', city, '_grid.pdf'), width = 10, height = 10, units = 'cm')
+  ggsave(paste0('figs/', city, '_grid.pdf'), width = 7, height = 7, units = 'cm')
+  embed_fonts(paste0('figs/', city, '_grid.pdf'), outfile = paste0('figs/', city, '_grid.pdf'))
 
@@
 
-tab %>% write_csv('figs/comparative_summary.csv')
+tab %>% 
+  mutate(percent_black = percent_black * 100) %>%   
+  write_csv('figs/comparative_summary.csv')
+
+```
+
+```{r}
+
+cluster_map_percent_black <- function(tracts_name, city, nclusters = 10){
+	tracts <- readOGR(dsn = paste0('data/cities/',tracts_name), layer = 'geo', verbose = FALSE)
+	tracts <- tracts[sum(tracts@data[,columns]) != 0,]
+	clusters <- readRDS(paste0('throughput/clusterings/', city))
+
+	tracts@data$cluster <- cutree(clusters,nclusters)
+	tracts@data <- tracts@data %>%
+	  mutate(non_Black = White + Asian + Hispanic + Other) 
+	totals = tracts@data %>% 
+	  group_by(cluster) %>%
+	  summarise(non_Black = sum(non_Black),
+	            Black = sum(Black)) %>%
+	  mutate(percent_black = Black / (Black + non_Black))
+
+	info <- totals %>% select(non_Black, Black) %>% as.matrix() %>% mutual_info() %>% round(2)
+	total_info <- tracts@data %>% select(non_Black, Black) %>% as.matrix() %>% mutual_info() %>% round(2)
+
+	overlay <- unionSpatialPolygons(tracts, IDs = tracts@data$cluster)
+
+	tracts@data <- tracts@data %>% 
+	  left_join(totals, by = c('cluster' = 'cluster')) %>% 
+		mutate(id = as.character(row_number() - 1))
+
+	to_plot <- fortify(tracts)
+	to_overlay <- fortify(overlay)
+
+	to_plot <- to_plot %>% 
+		left_join(tracts@data[,c('id', 'cluster', 'percent_black')])
+
+	title <- paste0(tracts_name, ': ', nclusters, ' clusters')
+
+  if(tracts_name == 'Atlanta'){
+    guide = 'none'
+  }else{
+    guide = guide_colorbar(barwidth = 1, barheight = 4, label.theme = element_text(size = 7, angle = 0, family = 'Palatino'))
+  }
+
+	g <- to_plot %>% 
+		ggplot() +
+		geom_polygon(aes(x = long, y = lat, group = group, fill = percent_black)) + 
+		# ggtitle(paste0('Example clustering for ', city)) + 
+		ggthemes::theme_map() + 
+	  viridis::scale_fill_viridis(option = 'magma', 
+	                              name = 'Percent Black', 
+	                              labels = scales::percent, 
+	                              limits = c(0,1),
+	                              guide = guide) + 
+	  ggtitle(title) + 
+		geom_polygon(aes(x = long, y= lat, group = group), data = to_overlay, fill = NA, size = .3, color = 'black') + 
+	  theme(text=element_text(size=16, family="Palatino"),
+	        legend.justification=c(0,0), legend.position=c(.65,0),
+          legend.title = element_text(size = 8),
+          plot.title = element_text(size=10, vjust=1))
+
+	return(list(g = g, info = info, total_info = total_info))
+}
+
+```
+
+```{r}
+
+tab <- data.frame(city = character(), total_info = numeric(), info_6 = numeric(), info_20 = numeric())
+for(city in c('Philadelphia', 'Atlanta')){
+  if(!file.exists(paste0('throughput/clusterings/', city, '_binary'))){
+    print(city)
+  	spdf <- readOGR(dsn = paste0('data/cities/', city), layer = 'geo',verbose = FALSE)
+  	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
+  	df <- spdf@data[,columns]
+  	df <- df %>% 
+  	  mutate(non_Black = Asian + Hispanic + White + Other) %>% 
+  	  select(Black, non_Black)
+  	row.names(df) <- 1:nrow(df)
+  	constraint <- gRelate(spdf, byid = TRUE, pattern = '****1****')
+  	a <- info_clust(df, constraint)
+  	saveRDS(a, file = paste0('throughput/clusterings/',city, '_binary'))
+  }
+	columns = c('Black', 'White', 'Asian', 'Hispanic', 'Other')
+  new_row <- data.frame(city = city, total_info = 0, info_6 = 0, info_20 = 0)
+	for(nclusters in c(6, 20)){
+  	c <- cluster_map_percent_black(city, paste0(city, '_binary'), nclusters = nclusters)
+  	new_row$total_info = c$total_info
+  	new_row[,paste0('info_', nclusters)] <- c$info
+  	c$g
+  	ggsave(paste0('figs/', city, '_clusters_binary_', nclusters, '.pdf'), width = 7, height = 7, units = 'cm')
+  	embed_fonts(paste0('figs/', city, '_clusters_binary_', nclusters, '.pdf'),
+  	            outfile = paste0('figs/', city, '_clusters_binary_', nclusters, '.pdf'))
+  tab <- rbind(tab, new_row)
+	}
+}
+
+tab %>% 
+  write_csv('figs/clustering_summary.csv')
+
+```
+
+```{r}
+
+atlanta <- readOGR(dsn = paste0('data/cities/', 'Atlanta'), 
+  	                         layer = 'geo', 
+  	                         verbose = FALSE)
+
+pop <- atlanta@data %>% 
+  select(Asian, Black, Hispanic, Other, White) %>% 
+  summarise_each(funs(sum)) %>% 
+  as.numeric()
+
+pop / sum(pop)
 

### 2016-11-01 submission.tex
added basic PNAS articles

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Diversity, Segregation, and Complexity in Cities}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,c,1]{Author One}
\author[b,1,2]{Author Two} 
\author[a]{Author Three}

\affil[a]{Affiliation One}
\affil[b]{Affiliation Two}
\affil[c]{Affiliation Three}

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any conflict of interest here.}
\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format.  Instructions for use are provided below.

Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs.

\section*{Guide to using this template on Overleaf}

Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout. For more detailed information please see the \href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information for Authors}.

If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}. You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.

\subsection*{Author Affiliations}

Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. Authors with an ORCID ID may supply this information at submission.

\subsection*{Submitting Manuscripts}

All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

\subsection*{Format}

Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References. Other orders and headings are permitted.

\subsection*{Manuscript Length}

PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a Direct Submission research article is six pages and a PNAS PLUS research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.  When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for PNAS PLUS.

\subsection*{References}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}. All references, including for the SI, should be included in the main manuscript file. References appearing in both sections should not be duplicated.  SI references included in tables should be included with the main reference section.

\subsection*{Data Archival}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.

\subsection*{Language-Editing Services}
Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{frog}
\caption{Placeholder image of a frog with a long example caption to show justification setting.}
\label{fig:frog}
\end{figure}

\begin{SCfigure*}[\sidecaptionrelwidth][t]
\centering
\includegraphics[width=11.4cm,height=11.4cm]{frog}
\caption{This caption would be placed at the side of the figure, rather than below it.}\label{fig:side}
\end{SCfigure*}

\subsection*{Digital Figures}
\label{sec:figures}

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size. Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.  Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.

\subsection*{Single column equations}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, options are to use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures, or to use the \verb|\begin{widetext}...\end{widetext}| environment as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and footnotes, as mentioned in the \href{http://texdoc.net/pkg/cuted}{cuted package documentation}. In the case of problems with footnotes, it may be possible to correct the situation using commands \verb|\footnotemark| and \verb|\footnotetext|.

%% Do not use widetext if paper is in single column.
\begin{widetext}
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{align*}
\end{widetext}

\begin{table}%[tbhp]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsection*{Supporting Information (SI)}

The main text of the paper must stand on its own without the SI. Refer to SI in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc. Authors are limited to no more than 10 SI files, not including movie files. Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text. Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only.

\subsubsection*{SI Text}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).

\subsubsection*{SI Figures}

Provide a brief legend for each supporting figure after the supporting text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text. When saving TIFF files, use only LZW compression; do not use JPEG compression. Do not save figure numbers, legends, or author names as part of the image. Composite figures must be pre-assembled.

\subsubsection*{3D Figures}

Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.

\subsubsection*{SI Tables}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.

\subsubsection*{SI Datasets}

Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.

\subsubsection*{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.

\subsubsection*{Still images}

Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.

\subsubsection*{Appendices}

PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow{} % Display the acknowledgments section

% \pnasbreak splits and balances the columns before the references.
% Uncomment \pnasbreak to view the references in the PNAS-style
% If you see unexpected formatting errors, try commenting out \pnasbreak
% as it can run into problems with floats and footnotes on the final page.
%\pnasbreak

% Bibliography
\bibliography{pnas-sample}

\end{document}

### 2016-11-01 submission.tex
created outline with hipster ipsum

@@
+%!TEX program = xelatex
+
 \documentclass[9pt,twocolumn,twoside]{pnas-new}
@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,c,1]{Author One}
-\author[b,1,2]{Author Two} 
-\author[a]{Author Three}
+\author[a,b,1]{Philip Chodrow}
 
-\affil[a]{Affiliation One}
-\affil[b]{Affiliation Two}
-\affil[c]{Affiliation Three}
+\affil[a]{MIT Human Mobility and Networks Laboratory}
+\affil[b]{MIT Operations Research Center}
 
 % Please give the surname of the lead author for the running footer
-\leadauthor{Lead author last name}
+\leadauthor{Chodrow}
 
@@
 % Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any conflict of interest here.}
-\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}
+% \authorcontributions{Please provide details of author contributions here.}
+\authordeclaration{The author declares no competing interest.}
+% \equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: pchodrow@gmail.com}
 
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}
+\keywords{Information Theory $|$ Diversity and Segregation $|$ Urban Theory $|$ ...}
 
@@
 If this is the case, add \parshape=0 to the end of the list environment.
-\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format.  Instructions for use are provided below.
+\dropcap{I}nformation theory...blah blah blah.
 
-Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs.
+\section*{Results}
 
-\section*{Guide to using this template on Overleaf}
+Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas.
+Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party.
+Bicycle rights kinfolk man braid disrupt XOXO.
+Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch.
+Celiac 3 wolf moon kitsch tote bag bespoke pop-up.
+Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb.
+Paleo you probably haven't heard of them whatever typewriter normcore.
 
-Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout.
-For more detailed information please see the \href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information for Authors}.
+Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica.
+Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral.
+Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic.
+Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick.
+Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage.
+Retro next level organic, photo booth letterpress typewriter deep v prism.
+Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.
 
-If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}.
-You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.
+Fanny pack poutine trust fund cardigan.
+Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr.
+Pickled gastropub vegan etsy.
+Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi.
+Roof party fixie pok pok, cardigan prism next level woke lomo.
+Venmo butcher irony, single-origin coffee gastropub cliche tumblr.
+Fap pour-over banjo, freegan literally direct trade etsy.
 
-\subsection*{Author Affiliations}
+Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch.
+Kombucha whatever pug microdosing pitchfork.
+Activated charcoal ethical affogato echo park.
+Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine.
+Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole.
+Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr.
+Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.
 
-Include department, institution, and complete address, with the ZIP/postal code, for each author.
-Use lower case letters to match authors with institutions, as shown in the example.
-Authors with an ORCID ID may supply this information at submission.
+\subsection*{Using some damn math} 
+There's lots of damn math we could use.
 
-\subsection*{Submitting Manuscripts}
+\section*{Discussion}
 
-All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}.
-If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.
+Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas.
+Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party.
+Bicycle rights kinfolk man braid disrupt XOXO.
+Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch.
+Celiac 3 wolf moon kitsch tote bag bespoke pop-up.
+Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb.
+Paleo you probably haven't heard of them whatever typewriter normcore.
 
-\subsection*{Format}
+Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica.
+Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral.
+Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic.
+Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick.
+Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage.
+Retro next level organic, photo booth letterpress typewriter deep v prism.
+Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.
 
-Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References.
-Other orders and headings are permitted.
+Fanny pack poutine trust fund cardigan.
+Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr.
+Pickled gastropub vegan etsy.
+Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi.
+Roof party fixie pok pok, cardigan prism next level woke lomo.
+Venmo butcher irony, single-origin coffee gastropub cliche tumblr.
+Fap pour-over banjo, freegan literally direct trade etsy.
 
-\subsection*{Manuscript Length}
+Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch.
+Kombucha whatever pug microdosing pitchfork.
+Activated charcoal ethical affogato echo park.
+Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine.
+Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole.
+Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr.
+Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.
 
-PNAS generally uses a two-column format averaging 67 characters, including spaces, per line.
-The maximum length of a Direct Submission research article is six pages and a PNAS PLUS research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.
-When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for PNAS PLUS.
+Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas.
+Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party.
+Bicycle rights kinfolk man braid disrupt XOXO.
+Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch.
+Celiac 3 wolf moon kitsch tote bag bespoke pop-up.
+Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb.
+Paleo you probably haven't heard of them whatever typewriter normcore.
 
-\subsection*{References}
+Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica.
+Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral.
+Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic.
+Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick.
+Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage.
+Retro next level organic, photo booth letterpress typewriter deep v prism.
+Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.
 
-References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
-\cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}.
-All references, including for the SI, should be included in the main manuscript file.
-References appearing in both sections should not be duplicated.
-SI references included in tables should be included with the main reference section.
+Fanny pack poutine trust fund cardigan.
+Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr.
+Pickled gastropub vegan etsy.
+Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi.
+Roof party fixie pok pok, cardigan prism next level woke lomo.
+Venmo butcher irony, single-origin coffee gastropub cliche tumblr.
+Fap pour-over banjo, freegan literally direct trade etsy.
 
-\subsection*{Data Archival}
+Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch.
+Kombucha whatever pug microdosing pitchfork.
+Activated charcoal ethical affogato echo park.
+Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine.
+Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole.
+Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr.
+Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.
 
-PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.
+% \subsubsection*{Different table}
 
-\subsection*{Language-Editing Services}
-Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml).
-PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.
-
-\begin{figure}%[tbhp]
-\centering
-\includegraphics[width=.8\linewidth]{frog}
-\caption{Placeholder image of a frog with a long example caption to show justification setting.}
-\label{fig:frog}
-\end{figure}
-
-\begin{SCfigure*}[\sidecaptionrelwidth][t]
-\centering
-\includegraphics[width=11.4cm,height=11.4cm]{frog}
-\caption{This caption would be placed at the side of the figure, rather than below it.}\label{fig:side}
-\end{SCfigure*}
-
-\subsection*{Digital Figures}
-\label{sec:figures}
-
-Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size.
-Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
-
-Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
-
-Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
-To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.
-
-\subsection*{Single column equations}
-
-Authors may use 1- or 2-column equations in their article, according to their preference.
-
-To allow an equation to span both columns, options are to use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures, or to use the \verb|\begin{widetext}...\end{widetext}| environment as shown in equation \ref{eqn:example} below.
-
-Please note that this option may run into problems with floats and footnotes, as mentioned in the \href{http://texdoc.net/pkg/cuted}{cuted package documentation}.
-In the case of problems with footnotes, it may be possible to correct the situation using commands \verb|\footnotemark| and \verb|\footnotetext|.
-
-%% Do not use widetext if paper is in single column.
-\begin{widetext}
-\begin{align*}
-(x+y)^3&=(x+y)(x+y)^2\\
-       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
-       &=x^3+3x^2y+3xy^3+x^3. 
-\end{align*}
-\end{widetext}
-
-\begin{table}%[tbhp]
-\centering
-\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-\begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
-\midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-\bottomrule
-\end{tabular}
-
-\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
-\end{table}
-
-\subsection*{Supporting Information (SI)}
-
-The main text of the paper must stand on its own without the SI.
-Refer to SI in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.
-Authors are limited to no more than 10 SI files, not including movie files.
-Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
-Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only.
-
-\subsubsection*{SI Text}
-
-Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).
-
-\subsubsection*{SI Figures}
-
-Provide a brief legend for each supporting figure after the supporting text.
-Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text.
-When saving TIFF files, use only LZW compression; do not use JPEG compression.
-Do not save figure numbers, legends, or author names as part of the image.
-Composite figures must be pre-assembled.
-
-\subsubsection*{3D Figures}
-
-Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
-
-\subsubsection*{SI Tables}
-
-Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
+% Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
 Do not use tabs or spaces to separate columns in Word tables.
 
-\subsubsection*{SI Datasets}
+% \subsubsection*{Different table}
 
-Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
+% Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
+Do not use tabs or spaces to separate columns in Word tables.
 
-\subsubsection*{SI Movies}
+% \subsubsection*{SI Datasets}
 
-Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file.
+% Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
+
+% \subsubsection*{SI Movies}
+
+% Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file.
 All movies should be submitted at the desired reproduction size and length.
@@
 
-\subsubsection*{Still images}
+% \subsubsection{Still images}
 
-Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
+% Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
 
-\subsubsection*{Appendices}
+% \subsubsection*{Appendices}
 
-PNAS prefers that authors submit individual source files to ensure readability.
+% PNAS prefers that authors submit individual source files to ensure readability.
 If this is not possible, supply a single PDF file that contains all of the SI associated with the paper.
@@
 
-\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
+\acknow{P.C. is grateful to Marta C. Gonz\'{a}lez for helpful discussion, and to MIT-Saudi Arabia...}
 
@@
 % as it can run into problems with floats and footnotes on the final page.
-%\pnasbreak
+% \pnasbreak
 

### 2016-11-01 submission.tex
Added figure environments, removed ipsum

@@
 
-\title{Diversity, Segregation, and Complexity in Cities}
+\usepackage{csvsimple}
+% \usepackage{subcaption}
+% \usepackage{csquotes}
+
+\title{Diversity, Segregation, and Information in Cities}
 
@@
 
-\affil[a]{MIT Human Mobility and Networks Laboratory}
-\affil[b]{MIT Operations Research Center}
+\affil[a]{Human Mobility and Networks Laboratory, Massachusetts Institute of Technology, Cambridge, MA 02139}
+\affil[b]{Operations Research Center, Massachusetts Institute of Technology, Cambridge, MA 02139}
 
@@
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Information Theory $|$ Diversity and Segregation $|$ Urban Theory $|$ ...}
+\keywords{Information Theory $|$ Diversity and Segregation $|$ Urban Theory }
 
@@
 
+	\begin{figure}%[tbhp]
+		\centering % need to center this
+		\includegraphics[width=\linewidth]{figs/checkerboard.pdf}
+		\caption{
+			Schematic cities illustrating three dimensions of segregation. 
+			City (a) lacks global diversity, while Cities (b)-(d) are maximally globally diverse. 
+			City (b) is perfectly spatially even, whereas Cities (c) and (d) are maximally uneven. 
+			City (d) has higher spatial exposure than City (c) due to more contact points between racially distinct neighborhoods.
+		} \label{fig:checkerboard}
+	\end{figure}
+
 \section*{Results}
 
-Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas.
-Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party.
-Bicycle rights kinfolk man braid disrupt XOXO.
-Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch.
-Celiac 3 wolf moon kitsch tote bag bespoke pop-up.
-Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb.
-Paleo you probably haven't heard of them whatever typewriter normcore.
+	\begin{figure}
+			\includegraphics[width = .5\linewidth]{figs/Atlanta_percent_black.pdf}
+			\includegraphics[width = .5\linewidth]{figs/Philadelphia_percent_black.pdf} \\
+			\includegraphics[width = .5\linewidth]{figs/Atlanta_grid.pdf} 
+			\includegraphics[width = .5\linewidth]{figs/Philadelphia_grid.pdf} \\
 
-Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica.
-Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral.
-Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic.
-Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick.
-Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage.
-Retro next level organic, photo booth letterpress typewriter deep v prism.
-Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.
+			\centering
+			\begin{tabular}{l | c c c c}
+				\bfseries City & Percent Black & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$  \\\hline
+				\csvreader[late after line=\\]{figs/comparative_summary.csv}{}
+				{\csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv}
+			\end{tabular}
 
-Fanny pack poutine trust fund cardigan.
-Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr.
-Pickled gastropub vegan etsy.
-Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi.
-Roof party fixie pok pok, cardigan prism next level woke lomo.
-Venmo butcher irony, single-origin coffee gastropub cliche tumblr.
-Fap pour-over banjo, freegan literally direct trade etsy.
+			\caption{
+				Spatial analysis of residential segregation in Atlanta (left) and Philadelphia (right). 
+				(\emph{Top}): Percentage of black residents. 
+				(\emph{Middle}): Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of the local information $I(X,Y | X \in B_i)$. 
+				(\emph{Bottom}): Numerical summary, including the three information measures.
+			} \label{fig:Atlanta_philly}
+	\end{figure}
 
-Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch.
-Kombucha whatever pug microdosing pitchfork.
-Activated charcoal ethical affogato echo park.
-Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine.
-Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole.
-Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr.
-Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.
+	\begin{figure}
+		\includegraphics[width=\linewidth]{figs/mutual_fisher.pdf}
+		\caption{
+			Comparative global segregation measures in major U.S. cities, measured using racial categories $\mathcal{Y} = \{\text{Asian }, \text{Black }, \text{Hispanic }, \text{Other }, \text{White}\}$. 
+			We used a hexgrid of spatial resolution $r = 1$km for the computation of $J(X,Y)$.
+		} \label{fig:mutual_fisher}
+	\end{figure}
 
@@
 
+	\begin{figure}
+
+		\includegraphics[width = .5\linewidth]{figs/Atlanta_clusters_binary_6.pdf} 
+		\includegraphics[width = .5\linewidth]{figs/Philadelphia_clusters_binary_6.pdf} \\
+		\includegraphics[width = .5\linewidth]{figs/Atlanta_clusters_binary_20.pdf}
+		\includegraphics[width = .5\linewidth]{figs/Philadelphia_clusters_binary_20.pdf} \\
+
+		\centering
+		\begin{tabular}{l | c c c}
+			\bfseries City  & $I(C_6, Y)$ & $I(C_{20},Y)$ & $I(X,Y)$ \\\hline
+			\csvreader[late after line=\\]{figs/clustering_summary.csv}{}
+			{\csvcoli & \csvcoliii & \csvcoliv & \csvcolii}
+		\end{tabular}
+		\caption{
+			Example clusterings in Atlanta and Philadelphia using \eqref{eq:cluster_opt}, using $6$ and $20$ clusters. 
+			The table below summarizes the between-groups mutual information captured by each cluster and compares it to the mutual information in the unaggregated blockgroup level data.
+		} \label{fig:clusterings}
+
+	\end{figure}
+
+	\begin{figure}
+		\centering
+			% \begin{subfigure}[b]{0.45\linewidth}
+			\includegraphics[width = .4\linewidth]{figs/example_cluster_map.pdf}
+			% \caption{} \label{subfig:detroit_map}
+			% \end{subfigure}
+			% \begin{subfigure}[b]{0.45\linewidth}
+			\includegraphics[width = .4\linewidth]{figs/example_clusters_detailed.pdf}
+			% \caption{} \label{subfig:detroit_groups}
+			% \end{subfigure}
+			\caption{
+				Multigroup sociospatial structure in Detroit. 
+				\ref{subfig:detroit_map}: Map of six clusters. 
+				\ref{subfig:detroit_groups}: Joint distribution of demographics and clusters. 
+				Shadings reflect population proportions within clusters. 
+				Numerical labels give raw population proportions; for example, 29\% of the population is Black and resides in cluster B. 
+				Detroit has $H(Y) = 1.12$, $I(X,Y) = 0.57$, and $J(X,Y) = 0.08$. 
+				This clustering captures $I(C_6,Y) = 0.36$ nats of information.
+			} \label{fig:detroit}
+	\end{figure}
+
+	\begin{figure}
+			\includegraphics[width=\linewidth]{figs/AUC_illustration.pdf}
+			\caption{
+				Evaluation of clusterings using the area under the curve (AUC) for $\mathcal{Y} = \{\text{Black }, \text{Non-Black}\}$, computed according to the formula \eqref{eq:AUC_formula}. 
+				The bounding rectangle for each city is shown in grey. 
+				The AUC is the proportion of the rectangle contained in the shaded region. 
+				The AUC for Philadelphia has been adjusted to account for the data set's two disconnected geographic components.
+			} \label{fig:AUC}
+	\end{figure}
+	\begin{figure}
+		\centering
+		\input{figs/adjusted_regression.txt}
+		\caption{
+			Regression of the AUC computed by \eqref{eq:AUC_formula} on the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$. 
+			95\% confidence intervals for each of the determined coefficients are shown in parentheses. 
+			The three information measures jointly explain an adjusted 75\% of the variation in the AUC. 
+		}\label{fig:regression}
+	\end{figure}
+
 \section*{Discussion}
-
-Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas.
-Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party.
-Bicycle rights kinfolk man braid disrupt XOXO.
-Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch.
-Celiac 3 wolf moon kitsch tote bag bespoke pop-up.
-Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb.
-Paleo you probably haven't heard of them whatever typewriter normcore.
-
-Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica.
-Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral.
-Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic.
-Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick.
-Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage.
-Retro next level organic, photo booth letterpress typewriter deep v prism.
-Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.
-
-Fanny pack poutine trust fund cardigan.
-Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr.
-Pickled gastropub vegan etsy.
-Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi.
-Roof party fixie pok pok, cardigan prism next level woke lomo.
-Venmo butcher irony, single-origin coffee gastropub cliche tumblr.
-Fap pour-over banjo, freegan literally direct trade etsy.
-
-Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch.
-Kombucha whatever pug microdosing pitchfork.
-Activated charcoal ethical affogato echo park.
-Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine.
-Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole.
-Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr.
-Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.
-
-Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas.
-Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party.
-Bicycle rights kinfolk man braid disrupt XOXO.
-Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch.
-Celiac 3 wolf moon kitsch tote bag bespoke pop-up.
-Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb.
-Paleo you probably haven't heard of them whatever typewriter normcore.
-
-Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica.
-Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral.
-Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic.
-Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick.
-Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage.
-Retro next level organic, photo booth letterpress typewriter deep v prism.
-Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.
-
-Fanny pack poutine trust fund cardigan.
-Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr.
-Pickled gastropub vegan etsy.
-Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi.
-Roof party fixie pok pok, cardigan prism next level woke lomo.
-Venmo butcher irony, single-origin coffee gastropub cliche tumblr.
-Fap pour-over banjo, freegan literally direct trade etsy.
-
-Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch.
-Kombucha whatever pug microdosing pitchfork.
-Activated charcoal ethical affogato echo park.
-Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine.
-Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole.
-Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr.
-Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.
-
-% \subsubsection*{Different table}
-
-% Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
-Do not use tabs or spaces to separate columns in Word tables.
-
-% \subsubsection*{Different table}
-
-% Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
-Do not use tabs or spaces to separate columns in Word tables.
-
-% \subsubsection*{SI Datasets}
-
-% Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
-
-% \subsubsection*{SI Movies}
-
-% Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file.
-All movies should be submitted at the desired reproduction size and length.
-Movies should be no more than 10 MB in size.
-
-% \subsubsection{Still images}
-
-% Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
-
-% \subsubsection*{Appendices}
-
-% PNAS prefers that authors submit individual source files to ensure readability.
-If this is not possible, supply a single PDF file that contains all of the SI associated with the paper.
-This file type will be published in raw format and will not be edited or composed.
 
@@
 
+\nocite{Bettencourt2013}
+
 % Bibliography
-\bibliography{pnas-sample}
+\bibliography{/Users/phil/bibs/library.bib}
 

### 2016-11-01 5_figs.Rmd
centered checkerboard

@@
 ```{r, fig.height=3}
+
+checkerboard_illustration <- function(){
+	expand.grid(x = 1:8, y = 1:8) %>%
+		mutate(`(a)` = 1,
+			   `(b)` = .5,
+			   `(c)` = (x > 4)*1,
+			   `(d)` = (x + y) %% 2) %>%
+		gather(key = model, value = p, -x, -y) %>%
+		ggplot(aes(x = x, y = y)) +
+		ggthemes::theme_map() +
+		theme(axis.ticks = element_blank(),
+			  axis.text.x = element_blank(),
+			  axis.text.y = element_blank(),
+			  panel.margin = unit(1, "mm"),
+			  panel.background = element_rect(),
+			  panel.grid.major = element_line(size = 0),
+			  panel.grid.minor = element_line(size = 0),
+			  strip.background = element_rect(fill = "white", color = "white", size = 2),
+			  plot.margin=unit(c(0,1,0,0),"mm")) +
+		geom_tile(aes(fill = p)) +
+		scale_fill_continuous(low = 'white', high = 'black ', limits=c(0,1)) +
+		facet_grid(~model) +
+		xlab('') +
+		ylab('') +
+		guides(fill=FALSE)
+}
+
 g <- checkerboard_illustration() 
-g + theme(text=element_text(size=16, family="Palatino"))
-ggsave('figs/checkerboard.pdf', width = 4, height = 1.5)
+# g + theme(text=element_text(size=10, family="Palatino"))
+ggsave('figs/checkerboard.pdf', width = 4, height = 1.3)
 embed_fonts("figs/checkerboard.pdf", outfile="figs/checkerboard.pdf")
@@
 	theme_minimal() + 
-	theme(text=element_text(size=16, family="Palatino"),
+	theme(text=element_text(size=25, family="Palatino"),
 		  axis.ticks.x = element_blank(),
@@
 		  legend.title = element_blank(),
-		  line = element_blank(),
-		  text = element_text(size = 25))
+		  line = element_blank())
 

### 2016-11-02 submission.tex
major content included, needs trimming and refinement

@@
 	\usepackage{csvsimple}
+	% \usepackage{pc_writeup}
+	\usepackage{pc_math}
+	\usepackage[colorinlistoftodos]{todonotes}
+	\usepackage{csquotes}
+
 	% \usepackage{subcaption}
@@
 	% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Information Theory $|$ Diversity and Segregation $|$ Urban Theory }
+	\keywords{Information Theory $|$ Diversity and Segregation $|$ Urban Theory}
 
@@
 	\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
+% ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
+% ----------------------------------------------------------------------------
 
@@
 
+\section*{Results}
+	We model a city as a set of individuals distributed over a continuous spatial region $R$. 
+	Each individual can belong to one of a set $\mathcal{Y}$ of racial groups, such as $\mathcal{Y} = \{\text{Asian}, \text{ Black}, \text{ Hispanic}, \text{ Other}, \text{ White}\}$. 
+	The cross-tabulation of individuals and groups defines a probability $p(X,Y)$ that a randomly chosen individual from the city will live at location $X \in R$ and belong to group $Y$. 
+	The population density at $x$ is $p(x) = \sum_{y \in \mathcal{Y}} p(x,y)$. 
+	The global proportion of residents belonging to group $y$ is $p(y) = \int_R p(x,y)\; dx$. 
+	The proportion of residents at $x$ belonging to group $y$ is $p(y|x) = p(x,y) / p(x)$. 
+	To compare distributions to one another we use the Kullback-Leibler divergence 
+	\begin{equation}
+		D[q(A)\|r(A)] = \sum_{a \in \mathcal{A}} q(a) \log \frac{q(a)}{r(a)}\;.
+	\end{equation}
+	The divergence $D$ measures dissimilarity between distributions: it is zero if and only if $q = r$, and grows larger when $q$ and $r$ assign greater probability to different events.
+
+\subsection*{Entropy and Global Diversity}
+	\emph{Global} measures of the city depend only on $p(Y)$, the global composition of racial groups. 
+	The global Shannon entropy $H(Y)$ measures global diversity by comparing the global distribution $p(Y)$ to the uniform distribution $u(Y)$: 
+	\begin{equation}
+		H(Y) = -\sum_{y \in \mathcal{Y}}p(y) \log p(y) = \log \abs{\mathcal{Y}} - D[p(Y)\|u(Y)]\;.
+	\end{equation}
+
+	Local-to-global measures depend on comparisons between properties encoded by the local compositions $p(y|x)$ and the global composition $p(y)$.
+
+	Local measures depend only on comparisons between local compositions $p(y|x)$ and other nearby local compositions, without reference to the global composition $p(y)$. 
 \begin{figure}%[tbhp]
@@
 		\includegraphics[width=\linewidth]{figs/checkerboard.pdf}
+		\begin{tabular}{l | c c c} 
+		City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
+		\hline
+		(a) & $0$        & $0$        & $0$\\
+		(b) & $\log 2$ & $0$        & $0$\\
+		(c) & $\log 2$ & $\log 2$ & $\frac{1}{7} \log 2$\\
+		(d) & $\log 2$ & $\log 2$ & $\log 2$
+		\end{tabular}
 		\caption{
+			\todo[inline=true]{Illustrate computation of $J$ on this fig?}
 			Schematic cities illustrating three dimensions of segregation. 
-			City (a) lacks global diversity, while Cities (b)-(d) are maximally globally diverse. 
-			City (b) is perfectly spatially even, whereas Cities (c) and (d) are maximally uneven. 
-			City (d) has higher spatial exposure than City (c) due to more contact points between racially distinct neighborhoods.
+			City (a) lacks global diversity, while cities (b)-(d) are maximally globally diverse. 
+			City (b) is perfectly spatially even, whereas cities (c) and (d) are maximally uneven. 
+			City (d) has higher spatial exposure than city (c) due to more contact points between racially distinct neighborhoods.
 		} \label{fig:checkerboard}
 	\end{figure}
+\subsection*{Mutual Information and Spatial Evenness}
+	Local-to-global measures depend on comparisons between properties encoded by the local compositions $p(Y|x)$ and the global composition $p(Y)$. 
+	When each of the compositions $p(Y|x)$ is the same as $p(Y)$, the city is perfectly spatially even -- every racial group is represented at each location in proportion to their presence in the city overall.
+	When the local compositions $p(Y|x)$ differ from $p(Y)$, the city is spatially uneven. 
+	We can measure spatial evenness as the average difference between local and global compositions: 
+	\begin{equation}
+		I(X,Y) = \E_X[D[p(Y|X)\|p(Y)]\;.
+	\end{equation}
+	The operator $\E_X$ denotes average or expectation with respect to the probability distribution of $X$. 
+	$I(X,Y)$ is the \emph{mutual information}, a fixture of information theory. \cite{Bettencourt2015}
 
-\section*{Results}
+	When all groups are evenly distributed, such as in Figure \ref{fig:checkerboard}(b), each location has the same composition as the global composition, and $I(X,Y) = 0$. 
+	On the other hand, $I(X,Y)$ achieves its maximum value of $\log \abs{\mathcal{Y}}$ when every location is monogroup, such as in Figure \ref{fig:checkerboard}(c).
+
+\subsection*{Local Information and Spatial Complexity}
+
+	Global and local-global measures can illuminate important high-level urban structure, but miss an important class of phenomena. 
+	The model two-race cities (c) and (d) in Figure \ref{fig:checkerboard} each have the same values of $H(Y)$ and $I(X,Y)$. 
+	City (c), however, has a single dividing line of racial separation, while city (d) has smaller pockets of racial distinctiveness. 
+	In sociology or urban planning, this may be the difference between metastasized segregation and healthy variation. 
+	The difference is intrinsically local, defined in terms of relations between local compositions $p(Y|X)$ and nearby other local compositions.
+
+	To measure local variation, imagine segmenting the city into a grid of small neighborhoods. 
+	The \emph{local mutual information} in the neighborhood $B$ is the mutual information between $X$ and $Y$, restricted to $B$:
+	\begin{equation*}
+		I(X,Y | X \in B) = \E_{X|X \in B}[D[p(Y|X,X\in B)\|p(Y|X\in B)]\;.
+	\end{equation*}
+	The \emph{mean local information} measures the average local variation in group composition across the entire city:  
+	\begin{equation}
+		J(X,Y) = \E_B[I(X,Y|X \in B)]\;.
+	\end{equation}
+	If we regard the data as a discretization of a smooth probability field, then it is possible to show using arguments from information geometry \cite{Amari2000} that the local information in a neighborhood $B$ of radius $r$ is an approximation of the Fisher information $\mathcal{J}$:
+	\begin{equation}
+		I(X,Y|X \in B) \cong \frac{r^2}{4} \text{tr } \mathcal{J}_Y(x)\;, \label{eq:fisher_approx}
+	\end{equation}
+	where $[\mathcal{J}_Y(x)]_{ij} = \E_Y[-\partial_{x_i} \partial_{x_j} \log p(Y|x)]$ is the Fisher information matrix and $\text{tr } \mathcal{J}_Y(x)$ is the sum of its eigenvalues.
 
@@
 			\caption{
-				Spatial analysis of residential segregation in Atlanta (left) and Philadelphia (right). 
+				Spatial analysis of residential segregation in Atlanta and Philadelphia. 
 				(\emph{Top}): Percentage of black residents. 
-				(\emph{Middle}): Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of the local information $I(X,Y | X \in B_i)$. 
+				(\emph{Middle}): Hexgrid of spatial resolution 1km imposed over each city. Within each hex $B$ we compute the local information local information $J_Y(B) = I(X,Y | X \in B)$. 
 				(\emph{Bottom}): Numerical summary, including the three information measures.
@@
 
+\subsection*{Comparative Urban Analysis}
+
+	We can view Atlanta and Philadelphia in broader context by comparing $I(X,Y)$ and $J(X,Y)$ across a range of American cities in Figure \ref{fig:mutual_fisher}. 
+	For each city, we computed $J(X,Y)$ using a 1km hexgrid and the racial categories $\mathcal{Y} = \{\text{Asian}, \text{ Black}, \text{ Hispanic}, \text{ Other}, \text{ White}\}$. 
+	The mutual information and local information sort cities into three primary classes. 
+	Those toward the bottom left have low $I(X,Y)$ and therefore display relatively little spatial variation. 
+	Cities become more unevenly distributed as we move rightward. 
+	Those toward the bottom right, such as Atlanta and Detroit, are both highly uneven and have little local variation. 
+	These cities are highly segregated, with large, starkly separated monoracial regions. 
+	In contrast, cities toward the top right like Philadelphia and New York are more complex patchworks of smaller, ethnically distinct subregions.
+
 	\begin{figure}
@@
 		\caption{
-			Comparative global segregation measures in major U.S. cities, measured using racial categories $\mathcal{Y} = \{\text{Asian }, \text{Black }, \text{Hispanic }, \text{Other }, \text{White}\}$. 
+			Comparative global segregation measures in major U.S. cities, measured using racial categories $\mathcal{Y} = \{\text{Asian}, \text{ Black}, \text{ Hispanic}, \text{ Other}, \text{ White}\}$. 
 			We used a hexgrid of spatial resolution $r = 1$km for the computation of $J(X,Y)$.
@@
 
-\subsection*{Using some damn math} 
-There's lots of damn math we could use.
+\subsection*{Identifying Sociospatial Structure} 
+	The measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ quantify the presence of global, local-global, and purely local sociospatial structure.
+We can also use information methods to identify and visualize that structure directly.
+While approaches exist to this problem,\footnote{For three, see \cite{Logan2011}} these approaches tend to conceptualize this structure in terms of monoracial subregions.
+In practice, however, there may be coherent biracial or multiracial patterns in space.
+Fortunately, it is possible to motivate a structure-finding algorithm for identifying high-level demographic trends without restriction to monoracial regions.
 
@@
 	\end{figure}
+	Let $C$ be a random variable with the property that $H(C|X) = 0$; i.e.
+$X$ completely determines $C$.
+We can think of $C$ as group labels for the spatial locations $X$.
+Then, the chain rule of information\footnote{See \cite{Cover1991}} implies that
+	\begin{equation}
+		I(X,Y) = \underbrace{I(C,Y)}_{\text{between-group}} + \underbrace{I(X,Y|C)}_{\text{within-group}}. \label{eq:chain_rule}
+	\end{equation}
+
+	Equation \eqref{eq:chain_rule} motivates a natural algorithm for identifying sociospatial structure.
+The core idea is that a clustering $C$ is ``good'' when it captures most of the intrinsic information in the system in the between-groups term $I(C,Y)$.
+Mathematically, we therefore wish to maximize $I(C,Y)$ subject to a fixed number of clusters, an approach popularized by \cite{Dhillon2003,Banerjee2005}.
+In full generality, this is a challenging discrete optimization problem that may not be computationally tractable.
+Additionally, for interpretability, we need to apply a spatial constraint: locales should be clustered together only when they are adjacent in space.
+To handle these two problems, we use a greedy algorithm based on agglomerative clustering.
+At stage of the algorithm, we consider the problem of choosing a pair of locales $\{i*, j*\}$ to aggregate together.
+A good choice of locales leads to minimal information loss, which we may quantify as follows.
+The mutual information over the full complement $\mathcal{K}$ of locales may be written:
+	\begin{align*}
+	 	I(X,Y) &= \sum_{k \in \mathcal{K}} p(X = k) D[p(Y|X = k\|p(Y)] 
+	\end{align*} 
+	The contribution of locale $k$ to this sum is 
+	\begin{align*}
+		I_{k}(X,Y) &= p(X = k)D[p(Y|X = k)\|p(Y)]. 
+	\end{align*}
+	If we replace $i$ and $j$ with an aggregated locale $i\cup j$, the aggregated locale's contribution to the mutual information is 
+	\begin{align*}
+		I_{i\cup j}(X,Y) = p(X \in \{i,j\}) D[p(Y|X \in \{i,j\}) \| p(Y)]\;.
+	\end{align*}
+	The loss associated with aggregating $i$ and $j$ is the difference between their individual contributions and the aggregated contribution of $i\cup j$:
+	\begin{align*}
+		d(i,j) = I_{i}(X,Y) + I_{j}(X,Y) - I_{i \cup j}(X,Y).  
+	\end{align*}
+	The function $d$ may be regarded as a similarity measure.
+It satisfies $d(i,j) = 0$ if and only if $p(Y|X = i) = p(Y|X = j)$; i.e.
+the demographic profiles of locations $i$ and $j$ are identical.
+It is also symmetric in $i$ and $j$.
+It is not a true metric because it does not satisfy the triangle inequality.
+It can nevertheless be used to motivate a form of agglomerative hierarchical clustering, where at each phase of the algorithm we compute
+	\begin{equation}
+		(i^*, j^*) = \argmin_{i \text{ neighbors } j} \; d(i,j)\; \label{eq:cluster_opt}
+	\end{equation}
+	and combine $i^*$ and $j^*$ into a single new region.
+We repeat this procedure until only $N$ regions remain.
+The optimization over $i$ and $j$ ensures that the groupings so defined approximate maximum between-group information partitions, and the constraint that $i$ must neighbor $j$ ensures that the structure so determined is spatially contiguous.
+Equation \eqref{eq:cluster_opt} thus defines \emph{spatially constrained, information-maximizing agglomerative clustering}.
+Since the algorithm is greedy, it possesses no guarantees of optimality, but in practice its performance leads to intuitive, demographically-coherent regions.
+	\todo[inline=true]{Should we state the algorithm in a figure?}
 
@@
 				Multigroup sociospatial structure in Detroit. 
-				\ref{subfig:detroit_map}: Map of six clusters. 
-				\ref{subfig:detroit_groups}: Joint distribution of demographics and clusters. 
+				\emph{(Left)}: Map of six clusters. 
+				\emph{(Right)}: Joint distribution of demographics and clusters. 
 				Shadings reflect population proportions within clusters. 
@@
 	\end{figure}
+
+	This technique extends naturally to multi-group settings, though the visualization of results requires more subtlety.
+In Figure \ref{fig:detroit}, we conduct an illustrative analysis of multigroup segregation segregation in Detroit and the surrounding Wayne County.
+The prominent clusters A and B reflect the primary white/black division in the city, with A representing the white suburbs and B the denser black inner city.
+Cluster C is interpretable as a transitional area between these two populations, with both white and black residents represented.
+Cluster D coincides with the historically distinct, predominantly white town of Grosse Pointe.
+Cluster E coincides with the city of Hamtramck, whose history has been shaped by high levels of immigration from Europe and Asia.
+Finally, Cluster F coincides with the Hispanic neighborhood Mexicantown.
+
+	Aside from simply highlighting geographic trends, this analysis also enables the quantitative study of traditional concerns in segregation studies, such as concentration.
+According to \cite{Massey1988}, ``\emph{concentration} refers to the degree of a group's agglomeration in urban space.'' The concept of concentration invites us to consider the composition and density of each cluster.
+For example, whites make up 70\% of ``their'' cluster A, while black residents make up 86\% of ``their'' cluster B.
+Cluster B is also more densely populated than Cluster A: it contains 34\% of Wayne County's population, but just 19\% of the analyzed geographic area, making it more than twice as dense as Cluster A.
+Thus, in and around Detroit, black neighborhoods tend to be both more racially homogeneous and more densely packed in urban space than white ones.
+We can also learn about pairwise group exposure.
+For example, from Figure  \ref{subfig:detroit_groups}, we observe that Asian residents tend to be found alongside white residents, but more rarely alongside black or Hispanic residents.
+Hispanic residents tend to group in Mexicantown and in predominantly white neighborhoods -- not black ones.
+
+\subsection*{Evaluating Clusterings}
+	As discussed, the mean local information $J(X,Y)$ can be viewed as a measure of the number and character of transitional areas between regions of differing demographic structure.
+It is low in Atlanta because the city is dominated by one, large North-South boundary while it is higher in Philadelphia due to many points of contact between neighborhoods of differing composition.
+Sociospatial transitions are also precisely the ``fault lines'' along which we would expect the clustering algorithm \eqref{eq:cluster_opt} to split the city.
+It is therefore not a coincidence that Philadelphia both has a higher value of $J(X,Y)$ and requires a greater number of clusters to convey a similar amount of sociospatial information.
+
+	To quantify the relationship between the mean local information $J(X,Y)$ and the clustering method \eqref{eq:cluster_opt}, we must first quantify the performance of the latter.
+We can do this by considering the number of clusters necessary to convey a given fraction of overall mutual information.
+Consider the between-groups information $I(C_n,Y)$ plotted as a function of the number $n$ of clusters.
+The result is a curve reflecting how information scales at varying levels of aggregation.
+Figure \ref{fig:AUC} shows examples of these curves for Atlanta, Detroit, and Philadelphia, with $N$ plotted on a logarithmic scale.
+The area under the curve (AUC) serves a measure of how early information is ``captured'' by the clustering.
+When the AUC is high, small numbers of clusters suffice to capture most of the information; when it is low more complex clustering is necessary.
+The AUC can be computed explicitly via the formula
+	\begin{equation}
+		AUC = \frac{1}{I(X,Y) \log N} \sum_{n=1}^N I(C_n, Y) \log \frac{n+1}{n}, \label{eq:AUC_formula}
+	\end{equation}
+
+	AUC is therefore an algorithmic measure of sociospatial complexity, which is high when there are few sociospatial boundaries and low when there are many.
+Previously, we also argued that the quantities $H(Y)$, $I(X,Y)$, and $J(X,Y)$ may be viewed as information measures of sociospatial complexity, and particularly that $J(X,Y)$ measures the presence of sociospatial boundaries.
+We should therefore expect that these two sets of measures are related.
+A simple way to test for the relationship is via linear least squares regression.
+We aim to approximate the AUC via the formula
+		\begin{equation*}
+			AUC \approx m_1H(Y) + m_2I(X,Y) + m_3J(X,Y) + b\;.
+		\end{equation*}
+	The results of this approximation are shown in Figure \ref{fig:regression}.
+Collectively, the three information measures suffice to explain an adjusted 75\% of the variation in the AUC, confirming that both the information measures and the algorithmic AUC measure the same underlying sociospatial structure.
+The coefficients are also instructive.
+The largest magnitude by far is the coefficient of $I(X,Y)$; all else equal, clustering is more effective when groups are distributed unevenly.
+The negative sign on the coefficient of the mean local information $J(X,Y)$ confirms our specific expectations from above; for fixed $I(X,Y)$ and $H(Y)$, a city with higher $J(X,Y)$ (more sociospatial complexity) is more difficult to model or simplify than one with lower $J(X,Y)$.
 
@@
 
-\matmethods{Please describe your materials and methods here.
-This can be more than one paragraph, and may contain subsections and equations as required.
-Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.
+	\subsection*{Local Information as a Complexity Measure}
+		The local information $J(X,Y)$ is a measure of spatial exposure contextualized by the mutual information.
+It can also be viewed as a measure of \emph{intrinsic sociospatial complexity}.
+Cities with high mean local information, such as Philadelphia, have more complex spatial structure than those with low mean local information such as Atlanta.
+Urban complexity has been a lively topic of interest at the intersection of physics and urban planning.
+Physically, the complexity of cities can be viewed as a rough measure of the complexity of theories needed to explain their structure and dynamics \cite{Bettencourt2015}.
+From a planning perspective, the complexity of land use was famously theorized by Jane Jacobs \cite{Jacobs1992} to be a driver of the health of neighborhoods.
+This theory has received preliminary empirical support \cite{DeNadai2016a}, with more tests no doubt to come.
+The local information can capture local variations in land use as well as demographics.
+We hope that this measure will serve as a useful tool for scholars of urban spatial complexity.
+	\subsection*{Local Information as a Complexity Measure}
+		A major challenge for information theory in urban science is the development of principled dynamical models that can explain variations in observed information measures across cities \cite{Batty2014a}.
+One approach to this challenge is to observe how information changes over time at varying scales.
+On the timescale of decades, information measures may be used to analyze the evolution of the evolving sociospatial structure of cities.
+Information-maximizing agglomerative clustering could then be used to cluster coherent regions in time as well as space, enabling a view of shifting demographic boundaries.
 
-\subsection*{Subsection for Method}
-Example text for subsection.
+		Another avenue of exploration is on the much shorter timescale of daily mobility.
+Many adults spend less than half of their waking hours at home \cite{BureauofLaborStati2014}, indicating that residential segregation is only a partial characterization of city-wide diversity.
+Fortunately, our ability to learn about daily patterns of human mobility is progressing at a rapid pace.
+Recent years have seen an enormous increase in the use of mobile devices, allowing the passive collection of digital traces.
+These traces can be processed, analyzed, and validated to derive insight into daily activity patterns \cite{Widhalm2015,Yang,Jiang2013,Jiang2012c}.
+By combining these traces with residential demographics, it is possible to estimate flows of different demographic groups throughout the city on a daily basis.
+	\subsection*{Spatial Preprocessing}
+		The measures we have developed use their data (provided by the Census) as an unbiased approximation of the true distribution of racial groups across locations.
+This approach has two limitations.
+First, the demarcations of Census areal units are to some extent arbitrary: changing them would not change the underlying demographic phenomena, but may change the values of the mutual information or local information \cite{Openshaw1981}.
+Second, cities are anisotropic: two neighborhoods may be no more than 50 meters away, but if they are separated by a river, they may be functionally ``farther'' from each other than two other neighborhoods 500m apart.
+An approach to address both criticisms is to spatially preprocess the data using a smoothing kernel, as demonstrated in \cite{Roberto2015} for the case of the mutual information.
+Further experiments on the numerical impact of MAUP and anisotropy are called for in order to determine whether the benefit of preprocessing justifies the additional computational costs of this procedure.
+
+\matmethods{
+		\subsection*{Data Access}
+		All data used in this study are freely available from the U.S.
+Census.
+The \texttt{R} packages \texttt{acs} and \texttt{tigris} were used to programmatically download demographics and geographic shapefiles, respectively.
+All demographic data is from the Five-Year Estimates of the 2014 American Community Survey, Table B03002.
+		\subsection*{Software Repositories}
+			We are pleased to make available two software repositories accompanying this analysis.
+Package \texttt{compx} for the \texttt{R} programming language implements computation of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$, as well as a method for information-theoretic clustering.
+Access \texttt{compx} at \texttt{https://github.com/PhilChodrow/compx}.
+
+			The analysis repository for this project includes data acquisition and processing; core computations; and figure generation.
+We hope that this repository will provide useful examples of how to use \texttt{compx} for others aiming to replicate and extend our results.
+Download the project files at \texttt{https://github.com/PhilChodrow/spatial\_complexity}.
+
+		\subsection*{Delimiting Cities} 
+		The delimiting of cities based on non-arbitrary population densities or natural boundaries is an area of active discussion in urban planning.
+For recent advances, see \cite{Rozenfeld2008,Rozenfeld2011}.
+W took a simple approach for the purposes of this study.
+For each city considered, we analyzed the region composed of all (and only those) counties in which some or all of the city's municipal boundaries lie.
+For example, our region corresponding to Atlanta consists of Fulton and Dekalb Counties, our region corresponding to Philadelphia consists of Philadelphia County, and the region corresponding to Detroit consists of Wayne County.
+For independent cities such as Baltimore and Washington D.C., the region of analysis consists simply of the municipal city boundaries.
+		\subsection*{Aggregation of Racial Categories} 
+		The data provided by Table B03002 contains categories corresponding to two options for ethnicity (Hispanic or Latino, Not Hispanic or Latino) and nine corresponding to race, giving a total of 18 cross-tabulated categories.
+For the purposes of analysis, we aggregated these into just five categories, \{\text{Asian}, \text{Black}, \text{Hispanic}, \text{Other}, \text{White}\}.
+The correspondence used is given in the GitHub repository \texttt{PhilChodrow/spatial\_complexity} linked above, under \texttt{assumptions/race\_lookup.csv}.
+		\subsection*{Analysis and Visualization} 
+		All analysis and visualization was conducted in the \texttt{R} statistical programming language.
+We used \cite{Bivand2014b,Bivand2014a,Bivand2014} for geographic analysis, and \cite{Wickham} for visualization.
 }
@@
 
-\nocite{Bettencourt2013}
+% \nocite{Bettencourt2013}
 

### 2016-11-04 submission.tex
latest draft

@@
 			\caption{
+				\todo[inline=true]{comment in here that it is possible to walk a km in Atlanta and basically see no change, not so in Philly}
 				Spatial analysis of residential segregation in Atlanta and Philadelphia. 
@@
 	\todo[inline=true]{Should we state the algorithm in a figure?}
+
+	\todo[inline=true]{Consider plotting the loss-curves in a readable way.}
 

### 2016-11-04 figs.Rmd
fig revisions

--- 2016-10-23 5_figs.Rmd
@@
     theme: simplex
+    code_folding: hide
 ---
@@
 
+if(!dir.exists('figs')){
+  dir.create('figs')
+}
+
 ```
@@
 
+The checkerboard problem is a foundational issue in the spatial analysis of diversity; it demonstrates the modifiable areal units problem (maup)
+
 ```{r, fig.height=3}
-g <- checkerboard_illustration() 
-g + theme(text=element_text(size=16, family="Palatino"))
-ggsave('figs/checkerboard.pdf', width = 4, height = 1.5)
-embed_fonts("figs/checkerboard.pdf", outfile="figs/checkerboard.pdf")
-```
 
-## Methodological illustration
-```{r, warning = FALSE, message = FALSE}
-columns <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
-method_illustration <- function(city, vmax){
-	hexgrid <- readOGR(dsn = paste0('throughput/grids/',city), layer = 'geo', verbose = FALSE)
-	df <- hexgrid@data
-	df <- df %>% mutate(id = as.character(row_number()))
-	hexgrid <- fortify(hexgrid) %>%
-		left_join(df, by = c('id' = 'id'))
-	bbox <- c(min(hexgrid$long), min(hexgrid$lat), max(hexgrid$long), max(hexgrid$lat))
-
-	# squarify
-	max_range <- max(bbox[3] - bbox[1], bbox[4] - bbox[2])
-
-	long_mid <- (bbox[3] + bbox[1]) / 2
-	lat_mid  <- (bbox[2] + bbox[4]) / 2
-
-	bbox[1] <- long_mid - max_range / 2
-	bbox[3] <- long_mid + max_range / 2
-	bbox[2] <- lat_mid  - max_range / 2
-	bbox[4] <- lat_mid  + max_range / 2
-
-	map <- get_map(location = bbox, maptype = 'terrain-background')
-
-	p <- ggmap(map,darken = .5)
-	p <- p +
-		geom_polygon(data = hexgrid, aes(x = long, y = lat, group = group, fill = info), alpha = .8) +
-		scale_fill_continuous(low = 'white', high = 'steelblue', name= expression(J[Y](x)), limits = c(0, vmax), breaks = c(0,2,4,6)) +
-		theme(axis.ticks = element_blank(),
-			  axis.text.x = element_blank(),
-			  axis.text.y = element_blank()) +
+checkerboard_illustration <- function(){
+	expand.grid(x = 1:8, y = 1:8) %>%
+		mutate(`(a)` = 1,
+			   `(b)` = .5,
+			   `(c)` = (x > 4)*1,
+			   `(d)` = (x + y) %% 2) %>%
+		gather(key = model, value = p, -x, -y) %>%
+		ggplot(aes(x = x, y = y)) +
+		ggthemes::theme_map() +
+    geom_tile(aes(fill = p)) +
+		scale_fill_continuous(low = 'white', high = 'black ', limits=c(0,1)) +
+		facet_grid(~model, switch = 'x') +
 		xlab('') +
 		ylab('') +
-		theme(text=element_text(size=16, family="Palatino"),
-		    legend.justification=c(.8,0),
-			  legend.position=c(1,0),
-			  legend.background = element_rect(fill = alpha('blue', 0)),
-			  legend.title = element_text(colour="white"),
-			  legend.text = element_text(colour="white"),
-			  legend.key.size = unit(2, "mm"),
-			  legend.text = element_text(size = rel(.1))) +
-		ggtitle(paste0('Local Information in ', city, '\nJ(X,Y) = ', round(weighted.mean(df$info, df$density), 2)))
-	p
-}
-
-# first <- method_illustration('Detroit', vmax = 6)
-# second <- method_illustration('Philadelphia', vmax = 6)
-# 
-# gA <- ggplotGrob(first)
-# gB <- ggplotGrob(second)
-# g <- gridExtra::cbind.gtable(gA, gB, size = "max")
-# 
-# grid.draw(g)
-# 
-# ggsave('figs/method_illustration.pdf', g, width = 8, height = 5)
-
-```
-
-```{r additive_decomposability}
-
-expand.grid(x = 1:8, y = 1:8) %>% 
-	mutate(`(e)` = .5,
-		   `(e)` = ifelse(x %in% 2:4 & y %in% 2:7, 0, `(e)`),
-		   `(e)` = ifelse(x %in% 5:7 & y %in% 2:7, 1, `(e)`),
-		   `(f)` = .5) %>%
-	gather(key = model, value = p, -x, -y) %>%
-	ggplot(aes(x = x, y = y)) +
-		theme_minimal() +
+		guides(fill=FALSE) + 
 		theme(axis.ticks = element_blank(),
@@
 			  axis.text.y = element_blank(),
+			  panel.margin = unit(1, "mm"),
 			  panel.background = element_rect(),
@@
 			  panel.grid.minor = element_line(size = 0),
-			  plot.margin=unit(c(0,0,0,0),"mm")) +
-		geom_tile(aes(fill = p)) +
-		scale_fill_continuous(low = 'white', high = 'black ', limits=c(0,1)) +
-		facet_grid(~model) +
-		xlab('') +
-		ylab('') +
-		guides(fill=FALSE)
+			  strip.background = element_rect(fill = "white", color = "white", size = 2),
+			  plot.margin=unit(c(1,1,0,0),"mm"),
+			  strip.text=element_text(vjust=-100))
+}
 
-ggsave('figs/decomposability.pdf', width = 2.5, height = 1.5)
-
+g <- checkerboard_illustration() 
+g
+# g + theme(text=element_text(size=10, family="Palatino"))
+ggsave('figs/checkerboard.pdf', width = 4, height = 1.3)
+embed_fonts("figs/checkerboard.pdf", outfile="figs/checkerboard.pdf")
 ```
@@
 
+We'll read in some of the measures that we've computed:
 ```{r}
@@
 	ggplot(aes(x = I_XY, y = J_XY)) + 
-	geom_point(color = 'grey') + 
+	geom_point(color = 'firebrick', alpha = 0.2) + 
 	theme_light() + 
@@
 	xlab(expression(I(X,Y))) + 
+  geom_smooth(method = 'lm', se = FALSE, color = 'grey', linetype = 1, alpha = 0.2) + 
 	# ggtitle('Mutual information and mean local information in major US cities') +
 	geom_point(data = highlight, color = 'firebrick') + 
-	geom_text_repel(data = highlight, aes(label = city), family = 'Palatino' ) + 
-  theme(text=element_text(size=16, family="Palatino"),
+	geom_text_repel(data = highlight, aes(label = city)
+	                # , family = 'Palatino' 
+	                ) + 
+  theme(
+    # text=element_text(size=16, family="Palatino"),
         legend.title = element_text(size = 8),
@@
 ## Relationship to population density
+
+This is pretty interesting but we need to learn more
 ```{r}
@@
 	geom_smooth(method = 'lm', se = FALSE, color = 'black', size = .4) +
-	annotate('text', x = 300, y = .12, label = lm_eqn(cache), parse = T, family = 'Palatino') + 
+	annotate('text', x = 300, y = .12, label = lm_eqn(cache), parse = T) + 
 	geom_point(color = 'grey') + 
 	geom_point(color = 'firebrick', data = highlight) + 
-	geom_text_repel(aes(label = city), size = 3, data = highlight, family = 'Palatino' ) +
+	geom_text_repel(aes(label = city), size = 3, data = highlight) +
 	scale_x_continuous(trans = 'log10') +
@@
 	theme_light() +
-  theme(text=element_text(size=16, family="Palatino")) + 
+  theme(text=element_text(size=16)) + 
 	xlab(expression(rho~(population/km^{2}))) +
@@
 # Table --------
-
-tab <- cache %>% 
-	select(area, population, H_Y, I_XY, J = J_XY) %>% 
-	mutate(Density = population / area) %>% 
-	round(2) %>%
-	cbind(cache$city, .) %>%
-	rename(City = `cache$city`, 
-		   `Area` = area,
-		   Population = population,
-		   HY = H_Y,
-		   IXY = I_XY) %>%
-	arrange(desc(Density)) %>% 
-	select(City, `Area`, Population, Density, HY, IXY, J)
-
-tab %>% write_csv('figs/table.csv')
-
-tab %>% select(City, HY, IXY, J) %>% 
-	filter(City %in% c( 'Albuquerque', 'Detroit', 'Philadelphia')) %>% 
-	arrange(City) %>%
-	mutate(HY = sprintf("%.2f", HY),
-		   IXY = sprintf("%.2f", IXY),
-		   J = sprintf("%.2f", J)) %>% 
-	write_csv('figs/mini_table.csv')
-
-tab$Population <- format(tab$Population, big.mark=",", scientific=FALSE)
-tab$Area <- format(tab$Area, big.mark=",", scientific=FALSE)
-tab$Density <- format(tab$Density, big.mark=",", scientific=FALSE)
-
-datatable(tab, options = list(pageLength = 10), rownames = FALSE,
-		  caption = 'Table 1: This is a simple caption for the table.')
+# 
+# tab <- cache %>% 
+# 	select(area, population, H_Y, I_XY, J = J_XY) %>% 
+# 	mutate(Density = population / area) %>% 
+# 	round(2) %>%
+# 	cbind(cache$city, .) %>%
+# 	rename(City = `cache$city`, 
+# 		   `Area` = area,
+# 		   Population = population,
+# 		   HY = H_Y,
+# 		   IXY = I_XY) %>%
+# 	arrange(desc(Density)) %>% 
+# 	select(City, `Area`, Population, Density, HY, IXY, J)
+# 
+# tab %>% write_csv('figs/table.csv')
+# 
+# tab %>% select(City, HY, IXY, J) %>% 
+# 	filter(City %in% c( 'Albuquerque', 'Detroit', 'Philadelphia')) %>% 
+# 	arrange(City) %>%
+# 	mutate(HY = sprintf("%.2f", HY),
+# 		   IXY = sprintf("%.2f", IXY),
+# 		   J = sprintf("%.2f", J)) %>% 
+# 	write_csv('figs/mini_table.csv')
+# 
+# 
+# tab$Population <- format(tab$Population, big.mark=",", scientific=FALSE)
+# tab$Area <- format(tab$Area, big.mark=",", scientific=FALSE)
+# tab$Density <- format(tab$Density, big.mark=",", scientific=FALSE)
+# 
+# # datatable(tab, options = list(pageLength = 10), rownames = FALSE,
+# 		  # caption = 'Table 1: This is a simple caption for the table.')
 
@@
 		# ggtitle(paste0('Example clustering for ', city)) + 
-		theme_minimal() +
+		ggthemes::theme_map() +
 		theme(axis.ticks = element_blank(),
@@
 
-ggsave('figs/example_cluster_map.pdf',  width = 7, height = 7)
+ggsave('figs/example_cluster_map.pdf',  width = 7, height = 5)
 
@@
 	geom_text(aes(
-		label = ifelse(p < 0.005, "", 100 * round(p,2))), size = 8, family = 'Palatino') + 
+		label = ifelse(p < 0.005, "", 100 * round(p,2))), size = 8) + 
   scale_alpha_continuous(guide = 'none', limits = c(0, .3)) + 
 	theme_minimal() + 
-	theme(text=element_text(size=16, family="Palatino"),
+	theme(
+	  text=element_text(size=25),
 		  axis.ticks.x = element_blank(),
@@
 overlay <- unionSpatialPolygons(tracts, IDs = tracts@data$cluster)
-plot(overlay)
 
@@
 
-## Loss curves
-
-```{r}
-end_points <- loss_curves %>%
-	group_by(city) %>%
-	filter(nclust == max(nclust)) %>% 
-	filter(J_XY > 1.25 | J_XY < .75)
-
-loss_curves %>% 
-	ggplot() +
-	aes(x = nclust, y = I_XY, group = city) +
-	geom_path(aes(color = J_XY)) +
-	# geom_point(aes(color = J_XY), size = .3) +
-	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
-	# scale_y_continuous(trans = 'log10') +
-	geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) +
-	theme_minimal() +
-	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
-	annotation_logticks(size = .1, sides = 'b') +
-	ggtitle('Loss curves for spatially-constrained information-theoretic clustering')
-
-ggsave('figs/loss_curves.pdf', width = 10, height = 6)
-
-```
-
-```{r}
-end_points <- loss_curves %>%
-	group_by(city) %>%
-	filter(nclust == max(nclust))
-
-loss_curves %>% 
-	group_by(city) %>% 
-	mutate(I_max = max(I_XY)) %>% 
-	ggplot() +
-	aes(x = nclust, y = I_XY, group = city) +
-	geom_path(aes(color = I_max)) +
-	# geom_point(aes(color = J_XY), size = .3) +
-	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
-	# scale_y_continuous(trans = 'log10') +
-	geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) +
-	theme_minimal() +
-	annotation_logticks(size = .1, sides = 'b') +
-	ggtitle('Loss curves for spatially-constrained information-theoretic clustering')
-
-```
-
-## Normalized loss curves:
-
-```{r}
-
-process_loss_curves <- function(loss_curves){
-  loss_curves %>%
-  	group_by(city) %>%
-  	mutate(I_XY = I_XY / max(I_XY)) %>%
-  	mutate(nclust = nclust / max(nclust)) %>%
-  	# filter(nclust < 100) %>%
-  	ggplot() +
-  	aes(x = nclust, y = I_XY, color = J_XY, group = city) +
-  	geom_path() +
-  	scale_x_continuous(trans = 'log10') +
-  	scale_y_continuous(trans = 'log10') +
-  	theme_minimal() +
-  	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
-  	annotation_logticks(size = .1, sides = 'b') + 
-  	ggtitle('Normalized loss curves for spatially-constrained information-theoretic clustering')
-  }
-ggsave('figs/normed_loss_curves.pdf', width = 10, height = 6)
-```
-
-## Evaluating loss curves
-
 ### AUC illustration
@@
 
-binary_loss_curves <- read_csv('throughput/loss_curves.csv') %>%
+binary_loss_curves <- read_csv('throughput/loss_curves_binary.csv') %>%
   group_by(city) %>% 
@@
 
+adjust_thresh = 0.01
+highlights  <- c('Philadelphia', 'Atlanta', 'Detroit')
+curves <- binary_loss_curves %>% 
+  mutate(exclude = (nclust != 1) & (cummax(I_XY - lag(I_XY,default = 0)) < adjust_thresh) & (nclust < 10)) %>% 
+	filter(!exclude,
+	       max(I_XY > 0.0)) %>%
+	mutate(nclust = row_number())
+
+highlighted <- curves %>% 
+  filter(city %in% highlights)
+
+background <-  curves %>% 
+  filter(!(city %in% highlighted))
+
+end_points <- highlighted %>%
+	group_by(city) %>%
+	filter(nclust == max(nclust),
+	       city %in% highlights)
+
+# grey background
+
+background %>% 
+  ggplot(aes(x = nclust, y = I_XY)) +
+  geom_path(aes(group = city), color = 'grey') + 
+  geom_path(aes(group = city, color = city), data = highlighted) +
+  geom_text_repel(aes(label = city, color = city), data = end_points) +
+  scale_x_continuous(trans = 'log10') +
+  scale_y_continuous(limits = c(0,.5)) + 
+  theme_light() +
+  scale_color_brewer(palette = 'Set1', name = NULL, guide = 'none') +
+  xlab('Number of clusters (n)') + 
+	ylab(expression(I(C[n],Y))) +
+  # geom_vline(xintercept=2, alpha = .2) + 
+  annotation_logticks(sides = 'b', size = .2)
+
+  ggsave('figs/binary_loss_curves.pdf', width = 7, height = 4)
+  embed_fonts("figs/binary_loss_curves.pdf", outfile="figs/binary_loss_curves.pdf")
+
+```
+
+```{r}
+
 # dippy point that doesn't seem to be true here: if you don't have any excluded points on your curve, your AUC should be the same.
@@
 ```{r}
+summary <- compute_AUC(loss_curves, adjust_thresh = 0.005) 
+summary <- summary %>% left_join(cache, by = c('city' = 'city')) %>%
+	left_join(info_summary[,c('city', 'J_XY', 'I_XY')]) %>%
+	filter(I_XY > 0)
+
 summary %>% mutate(AUC = AUC / max(AUC),
@@
 
+# Core maps
+
+In this code chunk, we construct maps of binary (black/nonblack) segregation in Atlanta and Philadelphia, and then demonstrate the grid-based approach to computing the local information $J(X,Y)$.
+
 ```{r}
 
-summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
-		   `I(X,Y)` = I_XY / max(I_XY),
-		   `J(X,Y)` = J_XY / max(J_XY),
-		   `H(Y)` = H_Y / max(H_Y),
-		   density = density / max(density)) %>%
-	lm(`J(X,Y)` ~ log(density) + `I(X,Y)`, data = .) %>% 
-	summary()
-
-```
-
-```{r}
 cities <- c('Philadelphia', 'Atlanta')
@@
   }else{
-    guide = guide_colorbar(barwidth = 1, barheight = 4, label.theme = element_text(size = 7, angle = 0, family = 'Palatino'))
+    guide = guide_colorbar(barwidth = 1, barheight = 4, label.theme = element_text(size = 7, angle = 0))
   }
@@
                                 guide = guide) + 
-    ggtitle(paste0(city, ': J(X,Y) = ', J_XY)) + 
+    ggtitle(paste0(city)) + 
     ggthemes::theme_map() + 
     theme(legend.justification=c(0,0), legend.position=c(.65,0)) + 
-    theme(text=element_text(size=16, family="Palatino"),
+    theme(text=element_text(size=16),
           plot.title = element_text(size=10, 
@@
     geom_polygon() + 
-    viridis::scale_fill_viridis(name = expression(J[Y](x)), 
+    viridis::scale_fill_viridis(name = expression(J[Y](B)), 
                                 limits = c(0, .5), 
@@
     ggtitle(city) + 
-    theme(text=element_text(size=16, family="Palatino"),
+    theme(text=element_text(size=16),
           legend.justification=c(0,0), legend.position=c(.65,0),
@@
 
+# Binary cluster maps
+
+In this code chunk we produce maps based on the clusterings we found for Philadelphia and Atlanta.
+
 ```{r}
@@
 	tracts <- readOGR(dsn = paste0('data/cities/',tracts_name), layer = 'geo', verbose = FALSE)
+
 	tracts <- tracts[sum(tracts@data[,columns]) != 0,]
@@
   	spdf <- readOGR(dsn = paste0('data/cities/', city), layer = 'geo',verbose = FALSE)
+#   if(city == 'Philadelphia'){
+#   	   exclude <- c('421010098023','421010098022', '421010098021', '421010098011',  '421010098012')
+#     spdf <- spdf[!(spdf@data$GEOID %in% exclude),]   
+# 
+# 	}
   	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
@@
 
+# batch figs
+
+```{r}
+
+if(!dir.exists('figs/batch_figs')){
+  dir.create('figs/batch_figs')
+}
+
+guide = guide_colorbar(barwidth = 1, barheight = 4, label.theme = element_text(size = 7, angle = 0))
+cities <- cache$city
+
+tab <- data.frame(city = character(), percent_black = numeric(), H_Y = numeric(), I_XY = numeric(), J_XY = numeric())
+
+for(city in cities){
+  tracts <- readOGR(dsn = paste0('data/cities/',city), 
+  	                         layer = 'geo', 
+  	                         verbose = FALSE)
+
+  grid_tract <- read_csv(file = paste0('throughput/grid_tracts/', city, '.csv'))
+  grid <- readOGR(dsn = paste0('throughput/grids/',city), 
+  	                         layer = 'grid', 
+  	                         verbose = FALSE)
+
+  df <- tracts@data %>%
+    mutate(percent_black = Black / total) %>% 
+    tbl_df
+
+  this_city <- city
+  J_XY <- summary %>% filter(city == this_city) %>% 
+    select(J_XY) %>% 
+    as.numeric() %>% 
+    round(2)
+
+  tracts %>% 
+    fortify(region = 'GEOID') %>% 
+    tbl_df() %>% 
+    left_join(df, by = c('id' = 'GEOID')) %>% 
+    ggplot(aes(x = long, y = lat, fill = percent_black, group = group)) + 
+    geom_polygon() + 
+    viridis::scale_fill_viridis(name = 'Percent Black', 
+                                labels = scales::percent, 
+                                option = 'magma', 
+                                limits = c(0,1),
+                                guide = guide) + 
+    ggtitle(paste0(city)) + 
+    ggthemes::theme_map() + 
+    theme(legend.justification=c(0,0), legend.position=c(.65,0)) + 
+    theme(text=element_text(size=16),
+          plot.title = element_text(size=10, 
+                                  vjust=1),
+          legend.title = element_text(size = 8))
+
+  ggsave(paste0('figs/batch_figs/', city, '_percent_black.pdf'), width = 7, height = 7, units = 'cm')
+  embed_fonts(paste0('figs/batch_figs/', city, '_percent_black.pdf'), outfile = paste0('figs/batch_figs/', city, '_percent_black.pdf'))
+
+  tracts@data <- tracts@data %>% 
+    mutate(non_black = Hispanic + Other + White + Asian)
+
+  columns = c('non_black', 'Black')
+  stuff <- compx::info_analysis(tracts, columns = columns, resolution = NULL, grid_tract = grid_tract, grid = grid)
+
+  stuff$grid %>% 
+    fortify(region = 'id') %>% 
+    left_join(stuff$grid@data[,c('id', 'pop', 'info')], by = c('id' = 'id')) %>% 
+    ggplot(aes(x = long, y = lat, fill =  info, group = group)) + 
+    geom_polygon() + 
+    viridis::scale_fill_viridis(name = expression(J[Y](B)), 
+                                limits = c(0, .5), 
+                                option = 'magma',
+                                guide = guide) + 
+    ggthemes::theme_map() +
+    ggtitle(city) + 
+    theme(text=element_text(size=16),
+          legend.justification=c(0,0), legend.position=c(.65,0),
+          legend.title = element_text(size = 8),
+          plot.title = element_text(size=10, vjust=1))
+
+  # paste0('figs/', city, '_grid.pdf')
+  ggsave(paste0('figs/batch_figs/', city, '_grid.pdf'), width = 7, height = 7, units = 'cm')
+  embed_fonts(paste0('figs/batch_figs/', city, '_grid.pdf'), outfile = paste0('figs/batch_figs/', city, '_grid.pdf'))
+
+  percent_black = sum(tracts@data$Black) / sum(tracts@data$total)
+
+  new_row <- data.frame(city = city, percent_black = percent_black, H_Y = stuff$H_Y, I_XY = stuff$I_XY, J_XY = stuff$J_XY)
+  tab <- rbind(tab, new_row)
+}
+
+tab[,-1] <- round(tab[,-1],2)
+
+tab %>% 
+  mutate(percent_black = percent_black * 100) %>%   
+  write_csv('figs/comparative_summary.csv')
+
+```
+
+```{r}
+
+tab <- data.frame(city = character(), total_info = numeric(), info_6 = numeric(), info_20 = numeric())
+for(city in cities){
+  if(!file.exists(paste0('throughput/clusterings/', city, '_binary'))){
+    print(city)
+  	spdf <- readOGR(dsn = paste0('data/cities/', city), layer = 'geo',verbose = FALSE)
+  	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
+  	df <- spdf@data[,columns]
+  	df <- df %>% 
+  	  mutate(non_Black = Asian + Hispanic + White + Other) %>% 
+  	  select(Black, non_Black)
+  	row.names(df) <- 1:nrow(df)
+  	constraint <- gRelate(spdf, byid = TRUE, pattern = '****1****')
+  	a <- info_clust(df, constraint)
+  	saveRDS(a, file = paste0('throughput/clusterings/',city, '_binary'))
+  }
+	columns = c('Black', 'White', 'Asian', 'Hispanic', 'Other')
+  new_row <- data.frame(city = city, total_info = 0, info_6 = 0, info_20 = 0)
+	for(nclusters in c(6, 20)){
+  	c <- cluster_map_percent_black(city, paste0(city, '_binary'), nclusters = nclusters)
+  	new_row$total_info = c$total_info
+  	new_row[,paste0('info_', nclusters)] <- c$info
+  	c$g
+  	ggsave(paste0('figs/batch_figs/', city, '_clusters_binary_', nclusters, '.pdf'), width = 7, height = 7, units = 'cm')
+  	embed_fonts(paste0('figs/batch_figs/', city, '_clusters_binary_', nclusters, '.pdf'),
+  	            outfile = paste0('figs/batch_figs/', city, '_clusters_binary_', nclusters, '.pdf'))
+  tab <- rbind(tab, new_row)
+	}
+}
+
+tab %>% 
+  write_csv('figs/batch_figs/clustering_summary.csv')
+
+```
+
 # Session info

### 2016-11-04 5_make_figs.R

@@ -1,0 +1,2 @@
+Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio-server/bin/pandoc") 
+rmarkdown::render(input = "figs.Rmd")

### 2016-11-04 figs.Rmd
don't duplicate rows in clustering summary table

@@
   	            outfile = paste0('figs/', city, '_clusters_binary_', nclusters, '.pdf'))
+	}
   tab <- rbind(tab, new_row)
-	}
 }
@@
   mutate(percent_black = percent_black * 100) %>%   
-  write_csv('figs/comparative_summary.csv')
+  write_csv('figs/batch_figs/comparative_summary.csv')
 
@@
   	            outfile = paste0('figs/batch_figs/', city, '_clusters_binary_', nclusters, '.pdf'))
+	}
   tab <- rbind(tab, new_row)
-	}
 }

### 2016-11-04 make_figs.R

@@ -1,0 +1,2 @@
+Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio-server/bin/pandoc") 
+rmarkdown::render(input = "5_figs.Rmd")

### 2016-11-04 revision.tex
cleaning up

--- 2016-10-05 revision.tex
@@
 
-\title{An Information Lens on Demographic Diversity} % work on this
+\title{Segregation and Diversity Through an Information Lens} % work on this
 \author{\emph{Phil Chodrow}}
@@
 \usepackage{pc_math}
+\usepackage{csvsimple}
+\usepackage{subcaption}
+\usepackage{csquotes}
+
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \section{Introduction} \label{sec:intro}
-	\begin{itemize}
-		\item While many measures are available for the measurement of spatial diversity, we are not aware of quantitative methods for \emph{identifying} spatial structure. 
-		\item We endorse an information-theoretic approach to the study of diversity and segregation. In supporting this approach, we make two main contributions: 
-		\begin{itemize}
-			\item First, we argue that a simple suite of information measures -- entropy, mutual information, and local information -- intuitively measure global diversity, spatial unevenness, and spatial exposure.
-The mean local information is novel, and we develop some of its useful properties.
-This triplet of information measures readily lends itself to quantifying dimensions of segregation within and between cities.
-			\item Measures such as those we develop above can be viewed as measuring different kinds of sociospatial structure.
-To our knowledge, no such analysis methods have been proposed for \emph{identifying} that structure.
-Building on our information methods, we develop an algorithm for clustering urban areas into regions whose demographics are near constant in space.
-The result allows the analyst to both easily visualize spatial patterns of difference and conduct analysis using a natural partition of urban space.
-		\end{itemize}
-	\end{itemize}
-\section{Lit Review, Previous Work. Why our problem is new.} \label{sec:previous}
-	Core points: 
-	\begin{itemize}
-		\item Previous work is missing unification
-		\item Previous work is missing a practical component: how do you compute and learn about urban structure with it? 
-	\end{itemize}
 
-\section{Segregation and Information} \label{sec:information}
-	Information measures quantify structure through an epistemic framework.
+	% Residential segregation is a persistent fact of modern American cities.
+Spatial location within cities has been proposed as a generative force of disparity in education and income, as well as mental and physical health \cite{Sharkey2014,DiezRoux2001,Dietz2002,Ludwig2013,Ludwig2012}.
+Residential segregation -- the differential distribution of different racial groups across urban space -- would then be itself a driver of further inequalities.
+
+	The quantification of segregation has been a topic of lively conversation within sociology through the last seventy years.
+This conversation may be coarsely partitioned into two interlocking threads.
+The first thread concerns the \emph{conceptual dimensions} of segregation, with each dimension reflecting an aspect of the ways in which groups may be separated in space.
+Most modern authors prefer a subset of the dimensions proposed by Massey and Denton \cite{Massey1988}: evenness, exposure, concentration, centralization, and clustering.
+Recent writers \cite{Reardon2002,Reardon2004,Roberto2015a,Roberto2015} have tended to focus on the dimensions of evenness and exposure, with \cite{Reardon2004} arguing that the additional clustering and concentration may be considered special cases.
+
+	The second thread concerns the appropriate methods with which to quantify these dimensions.
+Most fall into one of two broad classes of mathematical structure.\footnote{These categories are related to but not identical to the standard classification introduced by \cite{Coulter1989}.} \emph{Combinatorial} measures are the most ``physical:'' they measure the probability that two individuals selected under specifiable conditions will fall into two specified racial categories \cite{Goodman1954,James1986}.
+Combinatorial measures may therefore be intepreted as probabilities of encounters between members of racial groups.
+They may be differentiated according to the conditions of selection and the choice of categories.
+Contrasting to combinatorial measures are \emph{information} measures \cite{Theil1971,Theil1967,Roberto2015a}.
+Information measures are nonphysical in the sense that they do not correspond to probabilities of events.
+Rather, they reflect an epistemic viewpoint, measuring the ability of an observer to make guesses about individuals' identities when supplied with various kinds of information.
+Combinatorial and information measures do not exhaust all possibilities, however.
+For example, \cite{Wong1999} shows that an parametric approach based on the standard deviational ellipses of different groups may also be used to characterize city-scale segregation patterns, and \cite{Osullivan2007} expands this approach using nonparametric methods.
+
+	Theorists have proposed multiple methods for each combination of dimension and mathematical framework, leading to a proliferation of segregation measures.
+While an abundance of options is in principle a boon to the analyst, relatively little guidance exists on which measures to use \emph{together}.\footnote{\cite{Massey1988} provides some guidance, but this is based on a statistical comparison of a variety measures on data, rather than from first principles.} Some authors use different mathematical paradigms for different dimensions; for example, \cite{Reardon2004} develops a combinatorial approach to evenness and an information theoretic approach to exposure.
+Considering the intrinsically multidimensional character of segregation itself, a mathematically coherent ensemble of measures is called for.
+This essay therefore advocates an approach to the study of residential segregation that is \emph{conceptually multidimensional}, \emph{computationally practical}, and \emph{mathematically unified}.
+
+	\textbf{Conceptually multidimensional.} ``Segregation'' is not a single quantifiable number; it is an intrinsically multidimensional phenomenon that therefore requires a coordinated ensemble of distinct measures.
+As far as possible, each measure should correspond to exactly one dimension of segregation.
+Furthermore, each measure should be interpreted in the context of each of the others.
+In this essay, we adopt the categories of evenness and exposure as emphasized by \cite{Reardon2004}, to which we add the dimension of global diversity of the city as a whole.
+
+	\textbf{Computationally practical.} Writing in 2004, the authors of \cite{Reardon2004} observe that 
+	\begin{displayquote}
+
+		[a]t present, few of the proposed spatial segregation measures have been used in published empirical segregation research.
+These measures have been ignored in part because they typically are more difficult to compute than the aspatial measures.
+
+	\end{displayquote}
+	Over a decade later, computing resources have expanded dramatically in capability and accessibility.
+While our focus in this essay is theoretical, our ultimate goal is to place effective tools in the hands of practitioners.
+To this end, we emphasize interpretability, performance, and ease-of-use in the methods we develop.
+All are freely available as a package for the \texttt{R} programming language, as described in the Appendix.
+
+	\textbf{Mathematically unified.} In the measurement of segregation, we typically treat a city as a collection of different types of individuals, distributed differentially in space.
+The distinct dimensions of segregation correspond to the distinct questions we can ask about these distributions.
+Information theory is the mathematical framework most suited for the comparison of distributions.
+Rather than developing  measures for each dimension separately, we show how that an information-theoretic framework leads naturally to an ensemble of conceptually coherent, interpretable metrics resting on a shared mathematical foundation.
+
+	Our original contribution to this program has three components.
+The first is a novel measure of spatial exposure -- the local information -- and prove some of its properties.
+As we show, the local information distinguishes cities with starkly separated monoracial regions from cities with more complex patchwork patterns of racial difference.
+We argue that the local information, together with global entropy and mutual information, constitute a coherent ensemble of measures for the multidimensional quantification of residential segregation.
+The second is a novel, information-based algorithm for identifying the spatial structure of residential segregation.
+The algorithm clusters urban areas into regions whose demographics are relatively constant in space, allowing the analyst to study organic, demographically coherent regions rather than the partially-arbitrary tract or blockgroup demarcations provided by the U.S.
+Census.
+The third contribution is to share the tools we develop freely with practitioners: our suite of information measures and our clustering tools are available as open-source software for the \texttt{R} statistical programming language.
+
+\section{Information and Segregation} \label{sec:information}
+	Information theory has long been a mainstay of engineering, computer science, and the natural sciences.
+Somewhat more recently, information theory has found application in the study of urban phenomena.
+Theorists have used information measures to have been used to estimate the distribution of populations and resources in space \cite{Webber1979}, guide urban zoning decisions \cite{Batty1972}, identify spatial hierarchies in population distributions \cite{Batty1974,Batty1976}, and measure the complexity of urban phenomena \cite{Bettencourt2015,Batty2014a}.
+In this section, we develop some key information theoretic concepts and relate them to the dimensions of residential segregation.
+
+	Information theory quantifies structure through an epistemic framework.
 The core insight is that systemic structure enables \emph{prediction}, while systemic variability can make such prediction difficult.
 Two systemic features are related when knowledge of one enables better prediction of the other.
-
-	We can formalize the idea of prediction quality by considering the following ``guessing game.'' We discuss the weather, and I ask you to give me a \emph{predictive distribution} on the question of whether or not it will rain tomorrow; e.g.
-``70\% yes, 30\% no.'' Call this predictive distribution $p$.
-Tomorrow evening I give you a reward $R(p,Z)$ based on your predictive distribution $p$ and the observed outcome $Z$.
-For our game to be a true ``guessing game,'' we might expect the reward function $R$ to possess the following intuitive properties:
+We can formalize the idea of prediction by considering the following ``guessing game.'' We discuss the weather, and I ask you to give me a \emph{predictive distribution} on the question of whether or not it will rain tomorrow.
+For example, you might tell me your predictive distribution $p$, where $p(\text{Rain}) = 70\%$ $p(\text{No Rain}) = 30\%$.
+Tomorrow evening I give you a reward $R(p,W)$ based on your predictive distribution $p$ and the observed weather $W$.
+For this game to be a true ``guessing game,'' we would expect the function $R$ by which I reward your guess to possess the following intuitive properties:
 	\begin{enumerate}
-		\item \textbf{Honesty:} The reward scheme $R$ encourages you to report your true beliefs about the rain.
-		\item \textbf{Fairness:} Your reward $R(p,Z)$ you receive depends only on the observed outcome, not ``what could have happened.''
+		\item \textbf{Honesty:} The reward function $R$ encourages you to report your true beliefs about the rain.
+		\item \textbf{Fairness:} The reward $R(p,W)$ you receive depends only on your prediction $p$ and the observed outcome, not ``what could have happened.''\footnote{An ``honest'' and ``fair'' reward function is more standardly called ``proper'' and ``local'' among information theorists.
+		}
 		\item \textbf{Smoothness:} Your reward increases smoothly as a function of the probability you assigned to the observed outcome. 
 	\end{enumerate}
-	It is possible to show (see e.g.
-\cite{Cover1991}) that these properties uniquely determine the logarithmic reward function $R(p,Z) = \log p(Z)$ up to an affine transformation.
-This logarithmic reward is the fundamental unit of information-theoretic analysis.
-It is customary to focus on the \emph{loss function} $L(p,Z) = - R(p,Z)$.
-Large loss corresponds to poor prediction.
+	It is possible to show that these properties uniquely determine the logarithmic reward function $R(p,W) = \log p(W)$ up to an affine transformation.\footnote{\cite{Cover1991,Csiszzr2004}} This logarithmic reward is therefore a fundamental unit of information-theoretic analysis.
+It is traditional to focus on the negative of the reward function -- the \emph{loss function} $L(p,W) = - R(p,W)$.
+Large loss reflects to poor prediction.
 Systemic structure is present when knowledge of a system feature enables us to decrease the loss in such a guessing game.
-
-	\begin{figure}
-		\includegraphics[width=\textwidth]{figs/checkerboard.png}
-		\caption{Information and the checkerboard}. \label{fig:checkerboard}
-	\end{figure}
 
@@
 
-		As an introductory example, consider the following guessing game: I pick a person at random from a city, and ask you to assign a predictive distribution on their race $Y$, where $Y$ takes possible values in the alphabet  $\mathcal{Y} = \{\text{Asian, Black, Hispanic, Other, White}\}$.
-There is a true distribution $p(Y)$ corresponding to the proportions of each race in the city in question.
-It is possible to show that your optimal guess is precisely this distribution $p$.
-The optimal expected loss is
-		\begin{equation}
+		Consider the following predictive guessing game.
+You and I are jointly studying residential segregation in Atlanta.
+I pick a person at random from the city, and ask you to give predictive distribution on their race $Y$, where $Y$ takes possible values in the alphabet  $\mathcal{Y} = \{\text{Asian, Black, Hispanic, Other, White}\}$.
+There is a true distribution $p(Y)$ corresponding to the proportions of each race in the city in question; in Atlanta in 2014 this true distribution is
+		\begin{align*}
+			p(\text{Asian})    &= 6\% \quad
+			p(\text{Black})    = 47\% \quad
+			p(\text{Hispanic}) = 9\% \\
+			p(\text{Other})    &= 4\% \quad
+			p(\text{White})    = 35\%\;. 
+		\end{align*}
+		It is possible to show that your optimal guess in our game is precisely this distribution $p$. Your optimal expected loss is therefore
+		\begin{equation*}
 			\E_Y[L(p,Y)] = -\E_Y[R(p,Y)] = -\E_Y[\log p(Y)] \triangleq H(Y), 
-		\end{equation}
-		also known as the (Shannon) \emph{entropy} of $p$.
+		\end{equation*}
+		where $\E_Y$ is the expectation or averaging operator with respect to $Y$.
+$H(Y)$ is the Shannon \emph{entropy} of $p$ \cite{Shannon1948}.
+Throughout this paper logarithms will use the natural base $e = 2.718\ldots$, in which case all information measures have dimensionless units of nats.
+
+		\begin{figure}
+			\includegraphics[width=\textwidth]{figs/checkerboard.pdf}
+			\caption{Schematic cities illustrating three dimensions of segregation.
+City (a) lacks global diversity, while Cities (b)-(d) are maximally globally diverse.
+City (b) is perfectly spatially even, whereas Cities (c) and (d) are maximally uneven.
+City (d) has higher spatial exposure than City (c) due to more contact points between racially distinct neighborhoods.}.
+\label{fig:checkerboard}
+		\end{figure}
 
@@
 the city is monoracial.
-Figure \ref{fig:checkerboard}(a) gives an example of this case in a two-race city.
+Figure \ref{fig:checkerboard}(a) illustrates this case in a schematic two-race city.
 Perfect prediction without loss is possible because this city lacks diversity.
@@
 The cities in Figure \ref{fig:checkerboard}(b)-(d) all have maximum demographic entropy.
-The underlying distribution gives you grounds on which to add probabilistic weight to one race over another; you may as well roll a fair (five-sided) die.
+The underlying distribution gives you no grounds on which to add probabilistic weight to one race over another; flip a fair coin.
 Prediction is difficult because of the city's diversity.
@@
 		The authors of \cite{Reardon2004} define spatial evenness as ``the extent to which groups are similarly distributed in residential space.'' \emph{Un}evenness, then, is the extent to which these groups are dissimilarly distributed in space.
-This physical description has an epistemic analogue: knowing where a resident lives conveys information about their race, in the sense of enabling better prediction.
-To quantify this intuition, consider two versions of the guessing game above.
+This physical description has an epistemic analogue.
+In an uneven city, knowing where a resident lives conveys information about their race, in the sense of enabling better prediction.
+To quantify this intuition, consider two versions of the guessing game in Atlanta.
 In the first version, I simply choose a random resident and ask you to guess as before.
-In the second, I choose a random resident but then tell you their residence.
+In the second, I choose a random resident but then tell you their location of residence.
 In the presence of unevenness, this will generally enable you to improve your predictions.
+For example, in Atlanta knowing that the resident lives in the predominantly white northern part of the city may substantially change your predictive distribution.
+We can quantify the value of this information as follows.
 Let $X$ be a random variable denoting the spatial residence.
@@
 The \emph{mutual information} $I(X,Y)$ is the predictive improvement associated with my telling you the residence:
+		\begin{equation*}
+			 I(X,Y) \triangleq H(Y) - H(Y|X)\;.
+		\end{equation*}
+		Since knowing $X$ can only improve prediction, $I(X,Y) \geq 0$. Furthermore, it is possible to show that 
 		\begin{equation}
-			 I(X,Y) \triangleq H(Y) - H(Y|X)\;.
-		\end{equation}
-		$H(Y)$ is the expected loss without any knowledge of $X$, while $H(Y|X)$ is the expected loss when you are provided the value of $X$ as well.
-Since knowing $X$ can only improve prediction, $I(X,Y) \geq 0$.
-Furthermore, it is possible to show that
-		\begin{equation}
-			I(X,Y) = D[p(X,Y)\|p(X)p(Y)]\;,
+			I(X,Y) = D[p(X,Y)\|p(X)p(Y)]\;, \label{eq:info_and_divergence}
 		\end{equation}
 		where 
-		\begin{equation}
+		\begin{equation*}
 			D[q\|r] \triangleq \sum_a q(a) \log \frac{q(a)}{r(a)}
-		\end{equation}
+		\end{equation*}
 		is the Kullback-Leibler (KL) divergence of $r$ from $q$.
 Though not an axiomatic metric, the KL divergence may be viewed as a measure of distance between two distributions.
-The mutual information $I(X,Y)$ can therefore also be viewed as the distance of the true distribution $p(X,Y)$ from the hypothetical distribution $p(X)p(Y)$ in which the variables $X$ and $Y$ are independent.
+Equation \eqref{eq:info_and_divergence} thus shows that the mutual information $I(X,Y)$ can be viewed as the distance of the true joint distribution $p(X,Y)$ of location and race from the hypothetical distribution $p(X)p(Y)$ in which location and race are independent.
 $I(X,Y)$ therefore measures the degree of dependence between these variables.
+In this sense, it resembles a correlation coefficient, but has the stronger property that $I(X,Y) = 0$ implies true statistical independence.
 
-		The author of \cite{Roberto2015a} uses the mutual information $I(X,Y)$ -- which she refers to as the Divergence Index -- as a measure of (un)evenness.
-To see this within our epistemic framework, $I(X,Y) = 0$, $X$ and $Y$ are independent, indicating that knowing $X$ does not improve your ability to guess $Y$.
+		The mutual information can be viewed as a measure of spatial unevenness (\cite{Roberto2015a}).
+To see why the mutual information should measure unevenness, consider first the case in which $I(X,Y) = 0$.
+Since $X$ and $Y$ are independent, knowledge of $X$ does not improve your ability to guess $Y$.
 City (b) in Figure \ref{fig:checkerboard} illustrates this situation; since ``everywhere is the same,'' location knowledge is of no predictive value.
@@
 In contrast, when $I(X,Y)$ achieves its maximum value of $H(Y)$, $X$ fully determines $Y$.
-This situation is illustrated by Figure \ref{fig:checkerboard}(c); if I tell you that a resident lives in the left- or right-hand side of the city, you can guess their race with perfect accuracy.
-This situation corresponds to one in which $H(Y|X) = 0$, i.e.
-every location is monoracial, reflecting maximal sociospatial unevenness.
+This case is illustrated by Figure \ref{fig:checkerboard}(c); if I tell you that a resident lives in the left- or right-hand side of the city, you can guess their race with perfect accuracy.
+Every location is monoracial, reflecting maximal sociospatial unevenness.
 
@@
 
-		The authors of \cite{Reardon2004} define \emph{spatial exposure} as ``the extent that members of one group encounter members of another group...in their local spatial environments.'' We can construct an epistemic analogue of spatial exposure by considering the following pair of guessing games.
-In each, I pick a resident at random from a fixed ``local spatial environment'' that we both know.
-In the first game, I ask you to guess that resident's race directly.
-In the second, I tell you where \emph{within that environment} the resident lives.
-If the spatial environment in question is $R$, then your expected loss in the first game is $H(Y|X\in B)$ and in the second $H(Y|X, X\in B)$.
-The prediction improvement associated with knowing residence within this local environment is
-		\begin{equation}
+		The authors of \cite{Reardon2004} define \emph{spatial exposure} as ``the extent that members of one group encounter members of another group...in their local spatial environments.'' We can construct an epistemic analogue of spatial exposure by considering the following pair of guessing games in Atlanta.
+In each, I pick a resident at random from a small neighborhood that we both know in advance.
+In the first game, I then ask you to guess that resident's race immediately.
+In the second, I tell you where \emph{within that neighborhood} the resident lives.
+If the neighborhood in question is $B$, then your expected loss in the first game is $H(Y|X\in B)$ and in the second $H(Y|X, X\in B)$.
+The predictive improvement associated with knowing residence within this neighborhood is therefore
+		\begin{equation*}
 			I(X,Y | X \in B) = H(Y|X\in B) - H(Y|X, X\in B)\;.
-		\end{equation}
-		Our core claim is that $I(X,Y|X \in B)$ -- the \emph{local (mutual) information} -- is a measure of spatial exposure \emph{in unevenly distributed populations}.
+		\end{equation*}
+		The measure $I(X,Y|X \in B)$ -- the \emph{local (mutual) information} -- is a measure of spatial exposure \emph{in unevenly distributed populations}.
 This qualification is important, since in a perfectly even population such as Figure \ref{fig:checkerboard}(b),  $I(X,Y) = 0$ and $I(X,Y|X\in B) = 0$ as well.
 Consider, however, the two unevenly distributed populations in Figure \ref{fig:checkerboard} (c) and (d).
-In Figure (c), if we play our two guessing games at a locale entirely within the left- or righthand halfs, the population is locally homogeneous and you can make perfect predictions regardless of the specific residence.
+In \ref{fig:checkerboard}(c), if we play our two guessing games at a locale entirely within the left- or right-hand halves, the population is locally homogeneous and you can make perfect predictions regardless of the specific residence.
 Only if we play the game on a locale straddling the middle boundary between the two groups does knowing a resident's specific residence help you to make predictions.
-If our locale $B$ perfectly straddles the boundary, so that the population proportions are $(\frac{1}{2},\frac{1}{2})$, then we have $H(Y|X\in B) = \log 2$ and $H(Y|X,X\in B) = 0$, so $I(X,Y|X \in B) = \log 2$.
+If $B$ perfectly straddles the boundary, so that the population proportions are $(\frac{1}{2},\frac{1}{2})$, then we have $H(Y|X\in B) = \log 2$ and $H(Y|X,X\in B) = 0$, so $I(X,Y|X \in B) = \log 2$.
 
 		This example illustrates that the local information $I(X,Y|X\in B)$ is high when $B$ is an area of demographic transition.
-When two cities are similarly uneven (similar $I(X,Y)$), a city with more such transitional areas has more exposure between groups, since residents of one area are closer to demographically different residents of another area.
+When two cities are similarly uneven (have similar $I(X,Y)$), a city with more such transitional areas has more exposure between groups, since residents of one area are closer to demographically different residents of another area.
 To illustrate, the city of Figure \ref{fig:checkerboard}(d) has the same level of spatial unevenness as \ref{fig:checkerboard}(c), but greater spatial exposure.
@@
 In particular, we should expect that when using fine-grained data and fine-grained spatial locales, our estimates of $I(X,Y|X \in B)$ should converge, and not vary wildly as we increase the resolution of our analysis.
-It is possible to prove mathematically that this is the case.
+It is possible to prove mathematically that this is indeed the case.
 The full proof is given in the appendix, but the idea of the proof is illustrative of the nature of the local information.
@@
 		 	\item $p(y|x) > 0$ for all $x \in M$
-		 	\item $p(y|x)$ is differentiable as a function of $x$ for all $x \in M$. 
+		 	\item $p(y|x)$ is differentiable as a function of $x$ for all $x \in M$ and $y \in \mathcal{Y}$. 
 		\end{enumerate}
-		The probability field $p(y|x)$ corresponds to an idealization of our data; it can be viewed as an interpolation or approximation to the true data using a smooth curve.
+		The probability field $p(y|x)$ corresponds to an idealization of our data; it can be viewed as an interpolation or approximation to the true data using a smooth function.
 
-		Define $I_r(x_0)$ to be the local information in a locale of radius $r$ around the point $x_0$:
-		\begin{equation}
-			I_r(x_0) = I(X,Y| \norm{X - x_0}^2 \leq r^2). 
-		\end{equation}
+		Define $I_r(x)$ to be the local information in a locale of radius $r$ around the point $x$:
+		\begin{equation*}
+			I_r(x) = I(X,Y| \norm{X - x}^2 \leq r^2). 
+		\end{equation*}
 		We may view $r$ as the resolution of our analysis. Under the stated conditions, the following approximation relationship holds: 
 		\begin{equation}
-			\lim_{r \rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{1}{4} \text{tr } J_Y(x_0), \label{eq:fisher_approx}
+			\lim_{r \rightarrow 0} \frac{I_r(x)}{r^2} = \frac{1}{4} \text{tr } J_Y(x), \label{eq:fisher_approx}
 		\end{equation}
-		where $J_Y(x_0)$ is the \emph{Fisher information in $Y$ about $x_0$}. $J_Y(x_0)$ is a matrix defined by the formula
-		\begin{equation}
-			J_Y(x_0) \triangleq \E_Y\left[\nabla_x \log p(Y|x) \left(\nabla_x \log p(Y|x)\right)^T \right]\;. 
-		\end{equation}
-		The trace $\text{tr }J_Y(x_0)$ of $J_Y(x_0)$ is simply the sum of the diagonal entries of $J_Y(x_0)$.
+		where $J_Y(x)$ is the \emph{Fisher information in $Y$ about $x$}. It is a matrix defined by the formula
+		\begin{equation*}
+			J_Y(x) \triangleq \E_Y\left[\nabla_x \log p(Y|x) \left(\nabla_x \log p(Y|x)\right)^T \right]\;. 
+		\end{equation*}
+		The trace $\text{tr }J_Y(x)$ is the sum of the diagonal entries of $J_Y(x)$.
 
-		Equation \eqref{eq:fisher_approx} implies that, for small $r$, the local information (when normalized by $r^2$) is approximately a constant that is independent of $r$.
+		Equation \eqref{eq:fisher_approx} implies that, for small $r$, the local information in a neighborhood of radius $r$, when normalized by $r^2$, is approximately a constant that is independent of $r$.
 Furthermore, that constant is related to the derivatives of the probability field $p(Y|X)$.
@@
 		\begin{equation}
-			\text{tr }J_Y(x_0) = \E_Y\left[\norm{\nabla_x \log p(Y|x_0)}^2\right]\;, \label{eq:derivative}
+			\text{tr }J_Y(x) = \E_Y\left[\norm{\nabla_x \log p(Y|x)}^2\right]\;, \label{eq:derivative}
 		\end{equation}
-		which together with \eqref{eq:fisher_approx} shows that $I_r(x_0)$ is proportional to the mean magnitude of the derivative of $p(Y|X)$ evaluated at $x = x_0$.
+		which together with \eqref{eq:fisher_approx} shows that $I_r(x)$ is proportional to the mean magnitude of the derivative of $p(Y|X)$ evaluated at $X = x$.
 
 		In the idealized mathematical framework, we may define the \emph{mean local information} as the average value of the trace of the Fisher information: 
-		\begin{equation}
+		\begin{equation*}
 			J(X,Y) \triangleq \E_X[\text{tr } J_Y(X)].
-		\end{equation}
+		\end{equation*}
 		Using the relationship \eqref{eq:derivative}, we can recognize $J(X,Y)$ as the total variation of $\log p(Y|X)$ with respect to the probability measure induced by $X$.
@@
 
-		It is simple to adopt this measure $J(X,Y)$ for practical data analysis. The procedure is as follows: 
+		The measure $J(X,Y)$ may be easily adapted for practical computation. The procedure is as follows:
 		\begin{enumerate}
@@
 		The explicit formula is 
-		\begin{equation}
-			J(X,Y) = \sum_i p(X \in B_i) I(X,Y|X \in B_i). \label{eq:formula}
-		\end{equation}
+		\begin{align}
+			J(X,Y) &= \E_B[I(X,Y|X \in B)] \\ 
+			       &= \sum_i p(X \in B_i) I(X,Y|X \in B_i). \label{eq:formula}
+		\end{align}
+		While the result \eqref{eq:fisher_approx} guarantees limiting resolution-independence in a mathematically idealized setting, in practice it is advisable to use locales $B_i$ of constant size and shape for all regions of analysis.
+
+	\subsection{The Spatial Hierarchy of Measures}
+		The information measures we have developed form a natural spatial hierarchy. We can write each using the KL divergence:
+		\begin{align}
+			H(Y) &= \log \mathcal{\abs{Y}} - D[p(Y) \| u(Y)] \;, \label{eq:H_div} \\ 
+			I(X,Y) &= \E_X[D[p(Y|X)]\| p(Y)] \;, \label{eq:I_div} \\ 
+			J(X,Y) &= \E_B,X[D[p(Y|X, X\in B) \| p(Y|X \in B)]] \;, \label{eq:J_div}
+		\end{align}
+		where $u(Y)$ is the uniform distribution on $\mathcal{Y}$.
+These formulae express each information measure as a divergence or average of divergences.
+Recalling that a divergence may be viewed as a distance in information space, the formulae make explicit the \emph{comparative} structure of the metrics.
+The entropy $H(Y)$ is a global measure because it compares the observed global distribution $p(Y)$ to a theoretical global measure, the uniform distribution $u(Y)$.
+The mutual information $I(X,Y)$ may similarly be viewed as a comparison between the local distribution $p(Y|X)$ and the global distribution $p(Y)$, making it a partially local and partially global metric.
+On the other hand, the mean local information is a strictly local measure, in the sense that its value depends only on comparisons between two local distributions $p(Y|X,X\in B)$ and $p(Y|X \in B)$, each of which express information only in the locale $B$.
+The three measures thus sort naturally into a spatial hierarchy, from purely global to purely local.
+A summary of this hierarchy is given in Figure \ref{fig:hierarchy}.
+
+		\begin{figure}
+			\centering
+			\begin{tabular}{l l l l l }
+				\textbf{Measure} & \textbf{Comparison type} & \textbf{Compares} & \textbf{To} \\
+				\hline 
+				 $H(Y)$ & Global-Global & $p(Y)$ & $u(Y)$ \\
+				 $I(X,Y)$ & Local-Global & $p(Y|X)$ & $p(Y)$ \\
+				 $J(X,Y)$ & Local-Local & $p(Y|X, X \in B)$ & $p(Y|X \in B)$ 
+			\end{tabular}
+			\caption{Spatial hierarchy of information measures.
+Each measure can be expressed as an average of comparisons of distributions via the KL divergence \eqref{eq:H_div}-\eqref{eq:J_div}.
+$u(Y)$ denotes the uniform distribution on $\mathcal{Y}$.} \label{fig:hierarchy}
+		\end{figure}
 
 	\subsection{Illustrations}
-		\subsubsection{Comparative Analysis: Detroit and Philadelphia}
-			To illustrate these methods, we use them to compare the cities of Philadelphia and Detroit.
-Our data is blockgroup-level demographics from the American Community Survey \cite{CensusRace}.
-As we show, computing $H(Y)$, $I(X,Y)$, and $J(X,Y)$ together highlights similarities and differences between these cities.
-			\todo{We need to generate a nice summary plot here and then fill this section in.}
+
 		\begin{figure}
-				\caption{Spatial analysis of racial diversity in Detroit (left) and Philadelphia (right).
-Top: Percentage of black residents in Detroit (left) and Philadelphia (right).
-Middle: Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of $I(X,Y | X \in B_i)$.
-Bottom: Estimated contours of constant $I(X,Y | X \in B_i)$; high values correspond to spaces of demographic transition.
-Very bottom: Numerical summary, including the three information measures.} \label{fig:detroit_philly}
+			\includegraphics[width = .5\textwidth]{figs/Atlanta_percent_black.pdf}
+			\includegraphics[width = .5\textwidth]{figs/Philadelphia_percent_black.pdf} \\
+			\includegraphics[width = .5\textwidth]{figs/Atlanta_grid.pdf} 
+			\includegraphics[width = .5\textwidth]{figs/Philadelphia_grid.pdf} \\
+
+			\centering
+			\begin{tabular}{l | c c c c}
+				\bfseries City & Percent Black & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$  \\\hline
+				\csvreader[late after line=\\]{figs/comparative_summary.csv}{}
+				{\csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv}
+			\end{tabular}
+
+			\caption{Spatial analysis of residential segregation in Atlanta (left) and Philadelphia (right).
+(\emph{Top}): Percentage of black residents.
+(\emph{Middle}): Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of the local information $I(X,Y | X \in B_i)$.
+(\emph{Bottom}): Numerical summary, including the three information measures.} \label{fig:Atlanta_philly}
 		\end{figure}
 
-		\subsubsection{Comparative Diversity Profiles}
-			\todo{Is there a fun way to explore the density point? E.g. figure out people per ``neighborhood'' in some way? Think about what I would want the analysis here to be.}
+		To illustrate these methods, we use them to compare spatial patterns of segregation in Atlanta and Philadelphia.
+Our data is blockgroup-level demographics from the American Community Survey \cite{CensusRace}.
+For visualization purposes, we use the racial categories $\mathcal{Y} = \{\text{Black}, \text{Non-Black}\}$, though we emphasize that these methods generalize to an arbitrary number of categories.
+Figure \ref{fig:Atlanta_philly} shows the stages of the analysis.
+First, we visualize the base demographics in space.
+We then partition each city using a hexagonal grid of radius 1km, and compute the local information within each hexagon.
+The mean local information $J(X,Y)$ is the average of these local informations.
+The entropy $H(Y)$ and $I(X,Y)$ do not require spatial computation and can be derived from the cross-tabulation of race by blockgroup.
 
-			We can achieve a broader picture by comparing $I(X,Y)$ and $J(X,Y)$ across a range of American cities, as we do in Figure \ref{fig:mutual_fisher}.
-The two measures $I(X,Y)$ and $J(X,Y)$ provide an intuitive sorting of major cities.
-Cities with small numbers of large, monoracial regions such as Atlanta and Detroit are sorted toward the bottom right while cities with more complicated patchworks of demographic variation like Philadelphia and New York are sorted toward the top right.
+		One quantitative difference is striking: while Atlanta and Philadelphia are approximately as globally diverse as measured by the entropy $H(Y)$, and approximately as uneven as measured by the mutual information $I(X,Y)$, they vary substantially in their degree of spatial exposure as measured by the mean local information.
+This quantitative difference reflects the qualitatively distinct patterns of clustering in each city.
+While Atlanta consists essentially of a northern, predominantly white cluster and a southern, predominantly black one, Philadelphia includes at least three visually distinct clusters with large proportions of black residents.
+This greater number of clusters generates more sociospatial transitions, which are highlighted by the local information.
+We will return to the relationship between the local information and clustering in Section \ref{sec:id}.
 
 		\begin{figure}
-				\includegraphics[width=\textwidth]{figs/mutual_fisher.png}
-				\caption{Comparative global diversity measures in major U.S.
-cities.
-Cities toward the bottom left are relatively spatially even.
-Cities toward the bottom right like Detroit and Atlanta are the most classically ``segregated'', with large, starkly separated monoracial regions.
-Cities toward the top right like Philadelphia and New York are complex ``patchworks'' of different, smaller regions.} \label{fig:mutual_fisher}
+			\includegraphics[width=\textwidth]{figs/mutual_fisher.pdf}
+			\caption{Comparative global segregation measures in major U.S.
+cities, measured using racial categories $\mathcal{Y} = \{\text{Asian }, \text{Black }, \text{Hispanic }, \text{Other }, \text{White}\}$.
+We used a hexgrid of spatial resolution $r = 1$km for the computation of $J(X,Y)$.} \label{fig:mutual_fisher}
 		\end{figure}
 
-			\todo{Optional: plot of $J$ vs. density in here, but probably only if we can think of more to say, possibly a follow-up analysis.}
+		We can view Atlanta and Philadelphia in broader context by comparing $I(X,Y)$ and $J(X,Y)$ across a range of American cities in Figure \ref{fig:mutual_fisher}.
+For each city, we computed $J(X,Y)$ using a 1km hexgrid and the racial categories $\mathcal{Y} = \{\text{Asian}, \text{ Black}, \text{ Hispanic}, \text{ Other}, \text{ White}\}$.
+The result provides an intuitive sorting of major cities into three overall profiles.
+Cities toward the bottom left of Figure \ref{fig:mutual_fisher} are relatively spatially even.
+Cities toward the bottom right, such as Atlanta and Detroit, are highly segregated, with large, starkly separated monoracial regions.
+Cities toward the top right like Philadelphia and New York are more complex patchworks of smaller, ethnically distinct subregions.
 
-\section{Identifying Sociospatial Structure} \label{sec:id}
+\section{The Spatial Structure of Segregation} \label{sec:id}
 
-	We can view $H(Y)$, $I(X,Y)$, and $J(X,Y)$ as measures of the presence of sociospatial structure.
-It is somewhat unsatisfying to simply measure this structure without actually identifying it, since identifying this structure is of direct application in sociological data analysis.
-In this section, we use information-theoretic techniques to construct a procedure for finding subregions within cities that capture high-level demographic trends.
+	The measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ quantify the presence of different forms of sociospatial structure.
+In practical analysis, we may wish not only to measure the quantity of structure, but to identify and visualize that structure directly.
+While approaches exist to this problem,\footnote{For three, see \cite{Logan2011}} these approaches tend to conceptualize this structure in terms of monoracial subregions.
+In practice, however, there may be coherent biracial or multiracial patterns in space.
+Fortunately, it is possible to motivate a structure-finding algorithm for identifying high-level demographic trends without restriction to monoracial regions.
 
-	A useful place to start is with the property of ``additive spatial decomposition'' \cite{Reardon2004}.
-A diversity measure satisfies additive spatial decomposition when, if spatial units are aggregated into groups, the measure can be expressed as a sum of a ``within group'' term expressing variability within aggregations, and a ``between group'' term expressing variability across aggregations.
+	The key property we will use is ``additive spatial decomposition'' \cite{Reardon2004}.
+A segregation measure satisfies additive spatial decomposition when, if spatial units are aggregated into groups, the measure can be expressed as a sum of a ``within group'' term expressing variability within aggregations, and a ``between group'' term expressing variability across aggregations.
+The mutual information $I(X,Y)$ satisfies this property due to the following information-theoretic identity.
+Let $C$ be a random variable with the property that $H(C|X) = 0$; i.e.
+$X$ completely determines $C$.
+We can think of $C$ as group labels for the spatial locations $X$.
+Then, the chain rule of information\footnote{See \cite{Cover1991}} implies that
+	\begin{equation}
+		I(X,Y) = \underbrace{I(C,Y)}_{\text{between-group}} + \underbrace{I(X,Y|C)}_{\text{within-group}}. \label{eq:chain_rule}
+	\end{equation}
+	Equation \eqref{eq:chain_rule} expresses the following intuitive information decomposition: the information you have if I tell you a person's residence about their race is just the same as the information you would have if I told you first their approximate neighborhood $C$ and then their exact residence $X$.
 
-	\begin{itemize}
-		\item Construct the algorithm. 
-		\item Show algorithm results for one or two cities; show how knowing the clusters can lead to neat data analysis. 
-		\item Relate algorithmic performance to suite of information measures above. Note that $J$ measures the ``natural fault lines'' where we would expect the algorithm to segment.  
-	\end{itemize}
+	Equation \eqref{eq:chain_rule} motivates a natural algorithm for identifying sociospatial structure.
+The core idea is that a clustering $C$ is ``good'' when it captures most of the intrinsic information in the system in the between-groups term $I(C,Y)$.
+This idea is analogous to between-group sum-of-squares maximization in classical k-means clustering.
+We can also think of it using the standard framework of predictive games discussed in Section \ref{sec:information}.
+Recall the mutual information game, in which I choose a random resident in Atlanta and tell you their blockgroup of residence.
+Let us modify the game: instead of telling you one of the blockgroups, I will instead tell you one of $N$ \emph{regions}, each consisting of many blockgroups, for some fixed $N$.
+Before playing, I give you the opportunity to choose what these regions will be.
+Some regional divisions will be more useful than others.
+Suppose, for example, that we are playing this game in Atlanta and $N = 2$.
+Clearly, knowing whether a resident comes from the northern or southern part of the city is much more predictively valuable than knowing whether that same resident comes from the eastern or western part.
+You may therefore optimize your performance in our predictive guessing game by choosing to aggregate blockgroups into northern and southern regions.
+
+	Mathematically, this corresponds to maximizing $I(C,Y)$ subject to a fixed number of clusters.
+In full generality, this is a challenging discrete optimization problem that may not be computationally tractable.
+We can, however, construct a greedy algorithm that produces useful results.
+We begin with the full set of blockgroups.
+At each stage of the algorithm, we consider the problem of choosing a pair of locales $\{i*, j*\}$ to aggregate together.
+A good choice of locales leads to minimal information loss, which we may quantify as follows.
+The mutual information over the full complement $\mathcal{K}$ of blockgroups may be written:
+	\begin{align*}
+	 	I(X,Y) &= \sum_{k \in \mathcal{K}} p(X = k) D[p(Y|X = k\|p(Y)] 
+	\end{align*} 
+	The contribution of locale $k$ to this sum is 
+	\begin{align*}
+		I_{k}(X,Y) &\triangleq p(X = k)D[p(Y|X = k)\|p(Y)], 
+	\end{align*}
+	with which notation we may write 
+	\begin{equation*}
+		I(X,Y) = \sum_{k} I_k(X,Y).
+	\end{equation*}
+	If we replace $i$ and $j$ with an aggregated local $i\cup j$, its contribution to the mutual information is 
+	\begin{align*}
+		I_{i\cup j}(X,Y) = p(X \in \{i,j\}) D[p(Y|X \in \{i,j\}) \| p(Y)]\;.
+	\end{align*}
+	The loss associated with aggregating $i$ and $j$ is the difference between individual contribution and  aggregated contributions:
+	\begin{align*}
+		d(i,j) \triangleq I_{i}(X,Y) + I_{j}(X,Y) - I_{i \cup j}(X,Y).  
+	\end{align*}
+	In the context of the predictive guessing game, the loss $d(i,j)$ measures how much worse it is for you to use the coarser region $i\cup j$ than keeping the two regions $i$ and $j$ separate.
+The function $d$ may be regarded as a similarity measure.
+It satisfies $d(i,j) = 0$ if and only if $p(Y|X = i) = p(Y|X = j)$; i.e.
+the demographic profiles of locations $i$ and $j$ are identical.
+It is also symmetric in $i$ and $j$.
+It is not a true metric because it does not satisfy the triangle inequality.
+It can nevertheless be used to motivate a form of agglomerative hierarchical clustering, where at each phase of the algorithm we compute
+	\begin{equation}
+		(i^*, j^*) = \argmin_{i \text{ neighbors } j} \; d(i,j)\; \label{eq:cluster_opt}
+	\end{equation}
+	and combine $i^*$ and $j^*$ into a single new region.
+We repeat this procedure until only $N$ regions remain.
+The optimization over $i$ and $j$ ensures that the groupings so defined approximate maximum between-group information partitions, and the constraint that $i$ must neighbor $j$ ensures that the structure so determined is spatially contiguous.
+Equation \eqref{eq:cluster_opt} thus defines \emph{spatially constrained, information-maximizing agglomerative clustering}.
+Since the algorithm is greedy, it possesses no guarantees of optimality, but in practice its performance leads to intuitive, demographically-coherent regions.
+
+	\todo{We should cite \cite{Dhillon2003,Banerjee2005} as inspirations for this algorithmic approach.}.
+
+	\begin{figure}
+
+		\includegraphics[width = .5\textwidth]{figs/Atlanta_clusters_binary_6.pdf} 
+		\includegraphics[width = .5\textwidth]{figs/Philadelphia_clusters_binary_6.pdf} \\
+		\includegraphics[width = .5\textwidth]{figs/Atlanta_clusters_binary_20.pdf}
+		\includegraphics[width = .5\textwidth]{figs/Philadelphia_clusters_binary_20.pdf} \\
+
+		\centering
+		\begin{tabular}{l | c c c}
+			\bfseries City  & $I(C_6, Y)$ & $I(C_{20},Y)$ & $I(X,Y)$ \\\hline
+			\csvreader[late after line=\\]{figs/clustering_summary.csv}{}
+			{\csvcoli & \csvcoliii & \csvcoliv & \csvcolii}
+		\end{tabular}
+		\caption{Example clusterings in Atlanta and Philadelphia using \eqref{eq:cluster_opt}, using $6$ and $20$ clusters.
+The table below summarizes the between-groups mutual information captured by each cluster and compares it to the mutual information in the unaggregated blockgroup level data.}
+
+	\end{figure}
+
+	\subsection{Binary Segregation Structure in Atlanta and Philadelphia}
+		To demonstrate this approach for data analysis, we again compare black and non-black sociospatial trends in Atlanta and Philadelphia.
+While these two cities have similar spatial unevenness as measured by the mutual information $I(X,Y)$, the structure of this unevenness is very different.
+Because of a stark North-South divide in Atlanta, a simple clustering into six regions captures the vast bulk of sociospatial structure, conveying 0.28 nats out of a possible .34.
+In contrast, capturing a similar proportion of sociospatial structure in Philadelphia requires 20 clusters, reflecting the substantially more complex pattern of sociospatial variation.
+
+		The clusters found by these method may be viewed as coarse-grained views of the city that discard small details and leave only the high-level patterns.
+Comparison with the unaggregated data in Figure \ref{fig:Atlanta_philly} shows that the aggregate pictures highlight patterns of variation that may not be visible to the naked eye.
+Not every such cluster corresponds to a sociologically meaningful area or phenomenon, and care by the analyst is required in choosing an appropriate number of clusters.
+
+	\subsection{Structure of Multigroup Segregation in Detroit}
+		\begin{figure}
+		\centering
+			\begin{subfigure}[b]{0.45\textwidth}
+			\includegraphics[width = \textwidth]{figs/example_cluster_map.pdf}
+			\caption{} \label{subfig:detroit_map}
+			\end{subfigure}
+			\begin{subfigure}[b]{0.45\textwidth}
+			\includegraphics[width = \textwidth]{figs/example_clusters_detailed.pdf}
+			\caption{} \label{subfig:detroit_groups}
+			\end{subfigure}
+			\caption{Multigroup sociospatial structure in Detroit.
+\ref{subfig:detroit_map}: Map of six clusters.
+\ref{subfig:detroit_groups}: Joint distribution of demographics and clusters.
+Shadings reflect population proportions within clusters.
+Numerical labels give raw population proportions; for example, 29\% of the population is Black and resides in cluster B.
+Detroit has $H(Y) = 1.12$, $I(X,Y) = 0.57$, and $J(X,Y) = 0.08$.
+This clustering captures $I(C_6,Y) = 0.36$ nats of information.} \label{fig:detroit}
+		\end{figure}
+
+		This technique extends naturally to multi-group settings, though the visualization of results requires more subtlety.
+In Figure \ref{fig:detroit}, we conduct an illustrative analysis of multigroup segregation segregation in Detroit and the surrounding Wayne County.
+The prominent clusters A and B reflect the primary white/black division in the city, with A representing the white suburbs and B the denser black inner city.
+Cluster C is interpretable as a transitional area between these two populations, with both white and black residents represented.
+Cluster D coincides with the historically distinct, predominantly white town of Grosse Pointe.
+Cluster E coincides with the city of Hamtramck, whose history has been shaped by high levels of immigration from Europe and Asia.
+Finally, Cluster F coincides with the Hispanic neighborhood Mexicantown.
+
+		Aside from simply highlighting geographic trends, this analysis also enables the quantitative study of traditional concerns in segregation studies, such as concentration.
+According to \cite{Massey1988}, ``\emph{concentration} refers to the degree of a group's agglomeration in urban space.'' The concept of concentration invites us to consider the composition and density of each cluster.
+For example, whites make up 70\% of ``their'' cluster A, while black residents make up 86\% of ``their'' cluster B.
+Cluster B is also more densely populated than Cluster A: it contains 34\% of Wayne County's population, but just 19\% of the analyzed geographic area, making it more than twice as dense as Cluster A.
+Thus, in and around Detroit, black neighborhoods tend to be both more racially homogeneous and more densely packed in urban space than white ones.
+We can also learn about pairwise group exposure.
+For example, from Figure  \ref{subfig:detroit_groups}, we observe that Asian residents tend to be found alongside white residents, but more rarely alongside black or Hispanic residents.
+Hispanic residents tend to group in Mexicantown and in predominantly white neighborhoods -- not black ones.
+
+	\subsection{Relation of Clusterings and Information Measures}
+
+		As discussed, the mean local information $J(X,Y)$ can be viewed as a measure of the number and character of transitional areas between regions of differing demographic structure.
+It is low in Atlanta because the city is dominated by one, large North-South boundary while it is higher in Philadelphia due to many points of contact between neighborhoods of differing composition.
+Sociospatial transitions are also precisely the ``fault lines'' along which we would expect the clustering algorithm \eqref{eq:cluster_opt} to split the city.
+It is therefore not a coincidence that Philadelphia both has a higher value of $J(X,Y)$ and requires a greater number of clusters to convey a similar amount of sociospatial information.
+
+		To quantify the relationship between the mean local information $J(X,Y)$ and the clustering method \eqref{eq:cluster_opt}, we must first quantify the performance of the latter.
+We can do this by considering the number of clusters necessary to convey a given fraction of overall mutual information.
+Consider the between-groups information $I(C_N,Y)$ plotted as a function of the number $n$ of clusters.
+The result is a curve reflecting how information changes at varying levels of aggregation.
+Figure \ref{fig:AUC} shows examples of these curves for Atlanta, Detroit, and Philadelphia, with $N$ plotted on a logarithmic scale.
+The area under the curve (AUC) serves a measure of how early information is ``captured'' by the clustering.
+When the AUC is high, small numbers of clusters suffice to capture most of the information; when it is low more complex clustering is necessary.
+The AUC can be computed explicitly via the formula
+		\begin{figure}
+			\includegraphics[width=\textwidth]{figs/AUC_illustration.pdf}
+			\caption{Evaluation of clusterings using the area under the curve (AUC) for $\mathcal{Y} = \{\text{Black }, \text{Non-Black}\}$, computed according to the formula \eqref{eq:AUC_formula}.
+The bounding rectangle for each city is shown in grey.
+The AUC is the proportion of the rectangle contained in the shaded region.
+The AUC for Philadelphia has been adjusted to account for the data set's two disconnected geographic components.} \label{fig:AUC}
+		\end{figure}
+		\begin{equation}
+			AUC = \frac{1}{I(X,Y) \log N} \sum_{n=1}^N I(C_n, Y) \log \frac{n+1}{n}, \label{eq:AUC_formula}
+		\end{equation}
+		and is shown overlaid in Figure \ref{fig:AUC}.
+
+		Comparing the three cities we have discussed so far, Atlanta and Detroit have relatively high AUC, indicating that that simple models with few clusters suffice to convey most of the sociospatial information in those cities.
+This can be interpreted physically as the existence of a small number of contiguous, starkly-separated regions, such as the North-South divide in Atlanta (cf.
+Figure \ref{fig:Atlanta_philly}).
+Philadelphia in contrast has a lower AUC, indicating that more complex models with more clusters are necessary to capture an equivalent amount of information.
+
+		\begin{figure}
+			\centering
+			\input{figs/adjusted_regression.txt}
+			\caption{Regression of the AUC computed by \eqref{eq:AUC_formula} on the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$.
+95\% confidence intervals for each of the determined coefficients are shown in parentheses.
+The three information measures jointly explain an adjusted 75\% of the variation in the AUC.
+Table produced using \cite{Marek2015}.}\label{fig:regression}
+		\end{figure}
+
+		AUC is therefore an algorithmic measure of sociospatial complexity, which is high when there are few sociospatial boundaries and low when there are many.
+Previously, we also argued that the quantities $H(Y)$, $I(X,Y)$, and $J(X,Y)$ may be viewed as information measures of sociospatial complexity, and particularly that $J(X,Y)$ measures the presence of sociospatial boundaries.
+We should therefore expect that these two sets of measures are related.
+A simple way to test for the relationship is via linear least squares regression.
+We aim to approximate the AUC via the formula
+		\begin{equation*}
+			AUC \approx m_1H(Y) + m_2I(X,Y) + m_3J(X,Y) + b\;.
+		\end{equation*}
+		The results of this approximation are shown in Figure \ref{fig:regression}.
+Collectively, the three information measures suffice to explain an adjusted 75\% of the variation in the AUC, confirming that both the information measures and the algorithmic AUC measure the same underlying sociospatial structure.
+The coefficients are also instructive.
+The largest magnitude by far is the coefficient of $I(X,Y)$; all else equal, clustering is more effective when groups are distributed unevenly.
+The negative sign on the coefficient of the mean local information $J(X,Y)$ confirms our specific expectations from above; for fixed $I(X,Y)$ and $H(Y)$, a city with higher $J(X,Y)$ (more sociospatial complexity) is more difficult to model or simplify than one with lower $J(X,Y)$.
 
 \section{Discussion} \label{sec:discussion}
+
+	We have developed a multidimensional, computationally practical, and mathematically unified approach to measuring residential segregation.
+An ensemble of three information measures -- the entropy $H(Y)$, mutual information $I(X,Y)$, and mean local information $J(X,Y)$ -- lie at the heart of this approach.
+These three measures jointly comprise an ensemble of measures corresponding to global diversity, spatial evenness, and spatial exposure.
+All three measures have operational interpretations in the context of predictive guessing games, and all may be computed easily from freely-available Census data.
+Furthermore, we have presented an information-based method for algorithmically identifying the high-level spatial structure of segregation, and related the performance of this algorithm to our core ensemble of information measures.
+We discuss two topics in this concluding section.
+First, we consider approaches to further improving these measures through spatial preprocessing of the underlying data.
+Second, we discuss the broader applicability of our information-theoretic framework to further problems both within and beyond segregation studies.
+
+	\subsection*{Areal Units, Anisotropy, and Spatial Preprocessing}
+
+		The information measures developed here take their data (provided by the Census) as ``given,'' and compute with it directly.
+Two criticisms may be justly leveled against this approach.
+First, the demarcations of Census areal units are to some extent arbitrary: changing them would not change the underlying demographic phenomena, but may change the values of the mutual information or local information.
+This is a case of what is often called the ``modifiable areal unit problem,'' or MAUP \cite{Openshaw1981}.
+Second, using these areal units directly takes no notice of the intrinsically anisotropic nature of urban systems: two neighborhoods may be no more than 50 meters away, but if they are separated by a river, they may be functionally ``farther'' from each other than two other neighborhoods 500m apart \cite{Roberto2015}.
+In \cite{Roberto2015}, Roberto suggests a way to address both of these criticisms.
+Instead of computing with the data distributions $p(Y|X)$ of demographics at locations $X$, use instead the processed distributions $\bar{p}(Y|X)$, where
+		\begin{equation}
+				\bar{p}(Y = y|X = x) \triangleq \frac{\int p(Y = y,X = \xi)\phi(x, \xi) d\xi}{\int p(X = \xi)\phi(x, \xi) d\xi}\;, \label{eq:divergence_smoother}
+		\end{equation} 
+		The function $\phi:M\times M \rightarrow \R_+$ is a spatial kernel that can be specified by the analyst. In the original formulation, the spatial kernel is the \emph{biweight proximity} given by 
+		\begin{equation*}
+			\phi(x, \xi) \triangleq 
+			\begin{cases} 
+				\left[1 - \frac{\delta(x, \xi)^2}{r^2} \right]^2 & \text{if } \delta(x, \xi) < r \\
+				0 & \text{otherwise\;},
+			\end{cases}
+		\end{equation*}
+		where $\delta:M\times M \rightarrow \R_+$ measures distance between two locations.
+
+		The processed distributions $\hat{p}(Y|X)$ may be regarded as having been \emph{smoothed} according to the kernel $\phi$.
+Appropriate smoothing can address both MAUP and anisotropy.
+With respect to MAUP, smoothing may be viewed as a way of robustifying measures by averaging out noise induced by arbitrarily demarcated areal units.
+With respect to anisotropy, the specific choice of the kernel $\phi$ is up to the analyst, and there is no requirement that it reflect the standard Euclidean distance.
+Roberto demonstrates proximity function that depends on road distance, implicitly reflecting anisotropy induced by infrastructure distributions and geographic boundaries.
+
+		Roberto goes on to compute a Divergence Index at each location $X$ according to the formula 
+		\begin{equation*}
+			DI(X) \triangleq D[\bar{p}(Y|X)\|p(Y)] 
+		\end{equation*}
+		and an overall Divergence Index 
+		\begin{equation}
+			DI \triangleq \E_X[DI(X)]. \label{eq:divergence_index} 
+		\end{equation}
+		Simple algebra may be used to express \eqref{eq:divergence_index} in the form 
+		\begin{equation*}
+			DI = D[\bar{p}(X,Y) \| p(X)\bar{p}(Y)]. 
+		\end{equation*}
+		Comparing to \eqref{eq:info_and_divergence}, we can recognize the the Divergence Index as a form of the mutual information with spatial preprocessing.
+A similar preprocessing approach may be used to calculate a modified local information; simply replace $p(Y|X)$ with $\hat{p}(Y|X)$ in all formulae.
+It may be that this preprocessing approach offers a substantial improvement to both measures.
+Further experiments on the numerical impact of MAUP and anisotropy are called for in order to determine whether these benefits justify the additional computational costs and analytical degrees of freedom associated with specifying $\phi$.
+
+	\subsection*{Local Information as a Complexity Measure}
+
+		We have developed the local information $J_Y(X)$ as a measure of spatial exposure contextualized by the mutual information.
+As discussed briefly, another way to view the local information is as a measure of \emph{intrinsic sociospatial complexity}.
+Cities with high mean local information, such as Philadelphia, have more complex spatial structure than those with low mean local information such as Atlanta.
+Urban complexity has been a lively topic of interest at the intersection of physics and urban planning.
+Physically, the complexity of cities can be viewed as a rough measure of the complexity of theories needed to explain their structure and dynamics \cite{Bettencourt2015}.
+From a planning perspective, the complexity of land use was famously theorized by Jane Jacobs \cite{Jacobs1992} to be a driver of the health of neighborhoods.
+This theory has received preliminary empirical support \cite{DeNadai2016a}, with more tests no doubt to come.
+The local information can capture local variations in land use as well as demographics.
+We hope that this measure will serve as a useful tool for scholars of urban spatial complexity.
+
+	\subsection*{Generalizations and Further Applications}
+
+		As noted above, the utility of our information-theoretic framework is not restricted to the study of residential racial segregation.
+The same methods may be used for studying other demographic variables distributed in space, such as income, education level, and occupation type.
+They may also be used for studying non-demographic phenomena, such as land use and infrastructure distributions.
+Considering this generality, it is tempting to also generalize these measures to handle two or more demographic variables simultaneously, enabling joint analysis of, for example, race and income.
+Unfortunately, this generalization is nontrivial.
+The core difficulty lies in generalizing the mutual information to multiple variables.
+Most formulations of multivariate mutual information have the challenging property that they may be negative, leading to difficulties of interpretation.
+Substantial progress has been made on this front in the three-variable case by Loet Leydesdorff and collaborators \cite{Leydesdorff2014,Leydesdorff2015,Leydesdorff2013}, but generalized intuition on how to measure relations between multiple random variables remains elusive.
+
+		A simpler and no less useful generalization relates to temporal dynamics.
+A major challenge for information theory in urban science is the development of principled dynamical models that can explain observed information measures \cite{Batty2014a}.
+One approach to this challenge is to observe how information changes over time at varying scales.
+On the timescale of decades, information measures may be used to analyze the evolution of the evolving sociospatial structure of cities.
+Information-maximizing agglomerative clustering could then be used to cluster coherent regions in time as well as space, enabling a view of shifting demographic boundaries.
+
+		Another avenue of exploration is on the much shorter timescale of daily mobility.
+Many adults spend less than half of their waking hours at home \cite{BureauofLaborStati2014}, indicating that residential segregation is only a partial characterization of city-wide diversity.
+Fortunately, our ability to learn about daily patterns of human mobility is progressing at a rapid pace.
+Recent years have seen an enormous increase in the use of mobile devices, allowing the passive collection of digital traces.
+These traces can be processed, analyzed, and validated to derive insight into daily activity patterns \cite{Widhalm2015,Yang,Jiang2013,Jiang2012c}.
+By combining these traces with residential demographics, it is possible to estimate flows of different demographic groups throughout the city on a daily basis.
 
@@
 
-% TODO:
+\section*{Appendix}
+	\subsection*{Software}
+		We are pleased to make available two software repositories accompanying this analysis. 
+		  \begin{itemize}
+		    \item Package \texttt{compx} for the \texttt{R} programming language implements computation of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$, as well as a method for information-theoretic clustering.
+Access \texttt{compx} at
+		    \begin{displayquote}
+		      \texttt{https://github.com/PhilChodrow/compx}. 
+		    \end{displayquote}
+		    To install in \texttt{R}, install the package \texttt{devtools} at 
+		    \begin{displayquote}
+		    \texttt{https://cran.r-project.org/web/packages/devtools/index.html}
+		    \end{displayquote}
+		     and run the command 
+		    \begin{displayquote}
+		      \texttt{devtools::install\_github(``PhilChodrow/compx'')}
+		    \end{displayquote}
+		    in the \texttt{R} console. 
+		    \item The analysis repository for this project, including data acquisition and processing; core computations; and figure generation.
+We hope that this repository will provide useful examples of how to use \texttt{compx} for others aiming to replicate and extend our results.
+Download the project files at
+		    \begin{displayquote}
+		      \texttt{https://github.com/PhilChodrow/spatial\_complexity}. 
+		    \end{displayquote}
+		  \end{itemize}
 
-% - Think about whether we should also be looking at pairwise group exposure metrics?
-Is there an information-theoretic version of this?
-Maybe the information-loss associated with aggregating those two groups, that's not a bad guess.
+	\subsection*{Data and Methods}
+
+	\begin{itemize}
+		\item \textbf{Access}: All data used in this study are freely available from the U.S.
+Census.
+The \texttt{R} packages \texttt{acs} and \texttt{tigris} were used to programmatically download demographics and geographic shapefiles, respectively.
+		\item \textbf{Specification:} All demographic data is from the Five-Year Estimates of the 2014 American Community Survey, Table B03002. 
+		\item \textbf{Selection of Geographic Areas:} The delimiting of cities based on non-arbitrary population densities or natural boundaries is an area of active discussion in urban planning.
+For recent advances, see \cite{Rozenfeld2008,Rozenfeld2011}.
+W took a simple approach for the purposes of this study.
+For each city considered, we analyzed the region composed of all (and only those) counties in which some or all of the city's municipal boundaries lie.
+For example, our region corresponding to Atlanta consists of Fulton and Dekalb Counties, our region corresponding to Philadelphia consists of Philadelphia County, and the region corresponding to Detroit consists of Wayne County.
+For independent cities such as Baltimore and Washington D.C., the region of analysis consists simply of the municipal city boundaries.
+		\item \textbf{Aggregation of Racial Categories:} The data provided by Table B03002 contains categories corresponding to two options for ethnicity (Hispanic or Latino, Not Hispanic or Latino) and nine corresponding to race, giving a total of 18 cross-tabulated categories.
+For the purposes of analysis, we aggregated these into just five categories, \{\text{Asian}, \text{Black}, \text{Hispanic}, \text{Other}, \text{White}\}.
+The correspondence used is given in the GitHub repository \texttt{PhilChodrow/spatial\_complexity} linked above, under \texttt{assumptions/race\_lookup.csv}.
+		\item \textbf{Analysis and Visualization:} All analysis and visualization was conducted in the \texttt{R} statistical programming language.
+We used \cite{Bivand2014b,Bivand2014a,Bivand2014} for geographic analysis, and \cite{Wickham} for visualization.
+	\end{itemize}
+
+	\subsection*{Relationship of Mutual and Fisher Informations}
+
+	Let $X$ be a continuous random variable taking values in $\R^n$, and let $Y$ be a discrete random variable define on finite alphabet $\mathcal{Y}$.
+Suppose further that $p(y|x) > 0$ and that $p(y|x)$ is differentiable as a function of $x$ for all $x,y \in \R^n \times \mathcal{Y}$.
+Fix $x_0 \in \R^n$, and define $B_r \triangleq B_r(x_0) = \{ x \in R^n \;|\; \norm{x - x_0} \leq r \}$.
+Additionally, define the \emph{local mutual information} in $B_r$ as the mutual information between $X$ and $Y$ where $X$ is restricted to $B_r$:
+		\begin{align*}
+			I_r(x_0) &\triangleq \E_X[D[p(\cdot|X)\| p(\cdot|X \in B_r)]|X \in B_r] \\
+			&= \int_{B_r} p(x|X \in B_r) D[p(\cdot|x)\| p(\cdot|X \in B_r)] d^n x\;.
+		\end{align*}
+		where $D[p\|q] \triangleq \sum_{y} p(y) \log \frac{p(y)}{q(y)}$ is the Kullback-Leibler divergence of $q$ from $p$.
+
+		\begin{thm} \label{thm1}
+			Under the stated conditions, 
+			\begin{equation*}
+				\lim_{r\rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{n}{2(n+2)} \emph{trace}\; J_Y(x_0)\;.
+			\end{equation*}
+			where the Fisher information matrix $J_Y$ is given by 
+			\begin{align*}
+			J_Y(x) &\triangleq \E_Y\left[\nabla_x S_{Y}(x)\nabla_x S_{Y}(x)^T \right] \\
+			S_y(x) &\triangleq \log p(y|x)\;.
+		\end{align*}
+		\end{thm}
+
+	We first define the following functions. For set $A$, let 
+	\begin{equation*}
+		f_A(x) \triangleq  D[p(y|x) \| p(y|x\in A)] \;.
+	\end{equation*}
+	Additionally, let 
+	\begin{align*}
+		q_r(x) &\triangleq p(x|X \in B_r)\;.
+	\end{align*}
+	We can then write the local mutual information as 
+	\begin{equation*}
+		I_r(x_0) = \int_{B_r} q_r f_{B_r} \; d\lambda\;,
+	\end{equation*}
+	where $\lambda$ is the Lebesgue measure in $\R^n$. We also let $V_r \triangleq \int_{B_r} \; d\lambda$ be the volume of $B_r$. Finally, define 
+	\begin{equation*}
+		a_r(y) \triangleq p(X \in B_r, Y). 
+	\end{equation*}
+
+	\begin{lm}
+		For all $x \in B_r$,  
+		\begin{equation*}
+			q_r(x) = \frac{1 + O(r)}{V_r}\;.	
+		\end{equation*}
+	\end{lm}
+	\begin{proof}
+		First, Taylor's theorem implies that there exists a linear map $T$ such that 
+		\begin{align*}
+			p(x) &= p(x_0) + T(x - x_0) + O(\norm{x - x_0}^2) \\
+			&= p(x_0) + T(x - x_0) + O(r^2)\;. 
+		\end{align*}
+		 Then, 
+		\begin{align*}
+			p(X\in B_r) &= \int_{B_r} p \; d\lambda \\
+			&= \int_{B_r} p(x_0) + T(x - x_0) + O(r^2) d^n x\;.
+		\end{align*}
+		The middle term vanishes by spherical symmetry, leaving 
+		\begin{align*}
+			p(X\in B_r) &= p(x_0)V_r  + O(r^{n+2})\;.
+		\end{align*}
+		For $x \in B_r$, we therefore have 
+		\begin{align*}
+			p(x|X \in B_r) &= \frac{p(x,X \in B_r)}{p(X \in B_r)} \\ 
+			&= \frac{p(x)}{p(X \in B_r)} \\
+			&= \frac{p(x_0) + T(x - x_0) + O(r^2)}{p(x_0)V_r  + O(r^{n+2})} \\
+			&= \frac{p(x_0) + O(r) + O(r^2)}{p(x_0)V_r} \\
+			&= \frac{1 + O(r)}{V_r}\;,
+		\end{align*}
+		as was to be shown. 
+	\end{proof}
+
+	\begin{lm}
+		For all $y\in Y$,
+		\begin{equation*}
+			a_r(y) = p(x_0,y)V_r + O(r^{n+2})\;.
+		\end{equation*}
+	\end{lm}
+	\begin{proof}
+		Taylor's Theorem again implies that, for any $(x,y)\in B_r \times \mathcal{Y}$, there is a linear map $T$ such that
+		\begin{align*}
+			p(x,y) &= p(x_0,y) + T(x - x_0) + O(\norm{x - x_0}^2) \\
+			&= p(x_0,y) + T(x - x_0) + O(r^2). 
+		\end{align*}
+		We then have 
+		\begin{align*}
+			p(X \in B_r, y) &= \int_{B_r} p(x,y) \; d^n x \\ 
+			&= \int_{B_r} p(x_0,y) + T(x - x_0) + O(r^2) \; d^n x \\
+			&= p(x_0,y)V_r + O(r^{n+2}) \; d^n x\;,
+		\end{align*}
+		where the middle term has again vanished due to spherical symmetry. 
+	\end{proof}
+	\begin{lm}
+		For any $y \in \mathcal{Y}$, we have $p(y|X \in B_r) = p(y|x_0) + e_y$, where the error terms $e_y$ satisfy $e_y \in O(r^2)$ and $\sum_{y \in \mathcal{Y}} e_y = 0$. 
+	\end{lm}
+	\begin{proof}
+		That the errors must satisfy $\sum_{y \in \mathcal{Y}} e_y = 0$ follows from the fact that $p(\cdot| X \in B_r)$ must be a valid probability distribution on $\mathcal{Y}$.
+We'll now show that $e_y \in O(r^2)$.
+We then have
+		\begin{align*}
+			p(y|X \in B_r) &= \frac{p(X \in B_r, Y)}{p(X\in B_r)} \\
+			&= \frac{p(x_0, y)V_r + O(r^{n+2})}{p(x_0)V_r + O(r^{n+2})} \\
+			&= p(y|x_0)  O(r^{2}) \tag{$V_r \propto r^n$}\;,
+		\end{align*}
+		as needed. 
+	\end{proof}
+	\begin{lm}
+		For any $x \in B_r$, 
+		\begin{equation*}
+			f_{B_r}(x) = f_{x_0}(x) + O(r^3)\;.
+		\end{equation*}
+	\end{lm}
+	\begin{proof}
+		We compute directly: 
+		\begin{align*}
+			f_{B_r}(x) &= D[p(\cdot|x)\|p(\cdot|X \in B_r)] \\
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log \frac{p(y|x)}{p(y|X \in B_r)} \\
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log p(y|x) - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|X \in B_r)\; \\
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log p(y|x) - \sum_{y \in \mathcal{Y}} p(y|x) \log (p(y|x_0) + e_y) \\ 
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log p(y|x) - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|x_0) + \frac{e_y}{p(y|x_0)} + O(e_y^2)   \\
+			&= \sum_{y\in \mathcal{Y}} p(y|x) \log p(y|x) - \sum_{y \in \mathcal{Y}} p(y|x) \log p(y|x_0) + \sum_{y\in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}e_y + O(e_y^2) \\
+			&= D[p(\cdot|x)\|p(\cdot|x_0)] + \sum_{y\in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}e_y + O(e_y^2) \\
+			&= f_{x_0}(x) + \sum_{y\in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}e_y + O(e_y^2)\;.
+		\end{align*}
+		It remains to show that the error term is of order $O(r^3)$. Invoking Taylor's Theorem again, there exists a linear map $T$ such that 
+		\begin{align*}
+			\sum_{y\in \mathcal{Y}} \frac{p(y|x)}{p(y|x_0)}e_y + O(e_y^2) &= \sum_{y\in \mathcal{Y}} \left(1 +  \frac{1}{p(y|x_0)} T(x - x_0) + O(\norm{x - x_0}^2) \right)e_y + O(e_y^2) \\
+			&= \sum_{y\in \mathcal{Y}} \left(1 +  O(r) + O(r^2) \right)e_y + O(e_y^2) \\
+			&= \sum_{y\in\mathcal{Y}}\left[e_y + O(r^3)\right] \\
+			&= O(r^3)\;, 
+		\end{align*}
+		since $\sum_{y\in \mathcal{Y}} e_y = 0$ by the previous lemma. 
+	\end{proof}
+
+	We now invoke the following well-known theorem of information geometry; for reference see \cite{Amari2000}.
+
+	\begin{thm}
+		The Kullback-Leibler divergence and Fisher information $J_Y$ are related according to the following approximation: 
+		\begin{equation*}
+			D[p(\cdot|x)\|p(\cdot|x_0)] = \frac{1}{2} \left<x - x_0, J_Y(x_0)(x - x_0)\right> + O(\norm{x - x_0}^3)\;. 
+		\end{equation*}
+	\end{thm}
+
+	Theorem 2 allows us to write $I_r(x_0)$ as the following spherical integral:
+
+	\begin{align*}
+		I_r(x_0) &= \int_{B_r} q_r f_{B_r} \; d\lambda \\ 
+		&= \int_{B_r} q_r(x) f_{B_r}(x_0)\; d^n x \\
+		&= \int_{B_r} \frac{1 + O(r)}{V_r} \left(f_{x_0}(x) + O(r^3)\right) \; d^n x \\
+		&= \frac{1 + O(r)}{V_r} \int_{B_r} \left(f_{x_0}(x) + O(r^3)\right) \; d^n x \\
+		&= \frac{1}{2}\frac{1 + O(r)}{V_r} \int_{B_r} \left( \left<x - x_0, J_Y(x_0)(x - x_0)\right>  + O(r^3)\right) \; d^n x \\ 
+		&= \frac{1}{2}\frac{1 + O(r)}{V_r} \left(\int_{B_r}  \left<x - x_0, J_Y(x_0)(x - x_0)\right> d^n x + V_rO(r^3)\right). 
+	\end{align*}
+	It remains to compute the integral, which we do using the following lemma:
+
+	\begin{lm} For any positive-semidefinite matrix $A \in \R^{n\times n}$, 
+			$$\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x = \frac{n}{n+2} r^{2} V_r \text{\emph{trace}} (A)$$
+	\end{lm}
+
+	\begin{proof}
+		Since $A$ is positive-semidefinite, there exist an orthonormal matrix $P$ and a diagonal matrix $D$ such that $A = P^TDP$.
+Furthermore, the entries of $D$ are the eigenvalues $\{\lambda_i\}$ of $A$.
+Then,
+		\begin{align*}
+			\int_{B_r} \left< x - x_0, A(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0, P^T DP(x-x_0) \right> d^n x\\
+			&= \int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x \;.
+		\end{align*}
+		We can regard $P$ as a reparameterization of $B_r$; since $\det P = 1$, we have 
+		\begin{align*}
+			\int_{B_r} \left< P(x - x_0),  DP(x-x_0) \right> d^n x &= \int_{B_r} \left< x - x_0,  D(x-x_0) \right> d^n x \\
+			&= r^n\int_{B_n} \left< rx, rDx \right> d^n x \\
+			&= r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x \;,
+		\end{align*}
+		where $B_n$ is the unit $n$-ball. We also let $S_n(r)$ be the $n$-sphere of radius $r$. Continuing, 
+		\begin{align*}
+			r^{n+2}\int_{B_n} \left< x, Dx \right> d^n x &= r^{n+2} \int_{B_n} \sum_{i= 1}^n x_i^2 \lambda_i \; d^nx \\
+			&= r^{n+2} \sum_{i = 1}^n \lambda _i \int_{B_n} x_i^2 \; d^nx \\
+			&= \frac{r^{n+2}}{n} \sum_{i = 1}^n \lambda _i \int_{B_n} \norm{x}^2 \; d^nx \tag{spherical symmetry} \\
+			&= \frac{r^{n+2}}{n} \text{tr}(A)  \int_{B_n} \norm{x}^2 \; d^nx \\ 
+			&= \frac{r^{n+2}}{n} \text{tr} (A) \int_{\rho \in [0,1]} \rho^2 S_{n-1}(\rho) d\rho \\
+			&= \frac{r^{n+2}}{n} \text{tr} (A) \int_{\rho \in [0,1]} \rho^{n+1} S_{n-1}(1) d\rho \\
+			&= \frac{r^{n+2}}{n} \text{tr} (A)  \frac{1}{n+2} S_{n-1}(1) \\
+			&= \frac{r^2}{n+2}  \text{tr} (A)  n r^{n}v(B_n(1)) \\
+			&= \frac{n}{n+2} r^{2} V_r \text{tr} (A)\;,
+		\end{align*}
+		as was to be shown. 
+	\end{proof}
+
+	We may therefore complete the proof:
+
+	\begin{align*}
+		I_r(x_0) &= \frac{1}{2}\frac{1 + O(r)}{V_r} \left(\int_{B_r}  \left<x - x_0, J_Y(x_0)(x - x_0)\right> d^n x + V_rO(r^3)\right) \\ 
+		&= \frac{1}{2}\frac{1 + O(r)}{V_r} \left[\frac{n}{n+2} V_r \text{tr } J_Y(x_0) + V_rO(r^3)\right] \\
+		&= r^2 \frac{n}{2(n+2)} \left(1 + O(r)\right)\left[\text{tr }J_Y(x_0) + O(r^3) \right]\;.
+	\end{align*}
+	Dividing through by $r^2$ and taking the limit as $r\rightarrow 0$ proves the theorem.
 

### 2016-11-04 tex/paper/appendix.tex
minor updates to slides and paper
cleaning up

@@
   \begin{itemize}
-    \item Package 
-    \texttt{compx} for the \texttt{R} programming language implements computation of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$, as well as a method for information-theoretic clustering.
+    \item Package \texttt{compx} for the \texttt{R} programming language implements computation of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$, as well as a method for information-theoretic clustering.
 Access \texttt{compx} at
@@
 Defining $p^k(i)$ this way reflects a simplifying assumption that each blockgroup in $N_k$ overlaps hex $k$ with equal area.
-Finally, define the estimated overall racial composition of hex $k$ through the formula $p^k_Y(y) = \sum_{i \in N_k} p^{k}_I(i) p^i_Y(y)$.
+Finally, define the estimated overall racial composition of hex $k$ through the formula $p^k(y) = \sum_{i \in N_k} p^{k}(i) p^i(y)$.
 The mutual information in hex $k$ is then:
@@
 	\begin{equation}
-		\text{trace } J_Y(k) \approx \frac{4 I(k)}{r^2}\;.
+		\text{trace } J_Y(k) \approx \frac{4 I(k)}{r^2} \triangleq J(k)\;.
 	\end{equation}
@@
 	\end{equation}
+
+% \subsection*{Proof that $J(X,Y)$ Satisfies the Transfer Principle}
+
+% 	As above, we consider the formula for $J(X,Y)$ in computational practice as 
+% 	\begin{equation} 
+% 		J(X,Y) \approx  \frac{1}{\sum_k P_k} \sum_k P_k J(k) = \frac{4}{r^2T} \sum_k P_k I(k), \label{comp}
+% 	\end{equation}
+% 	where we have defined $T$ as the total population, and where $k$ indexes the grid cells.
+
+% 	Suppose that we transfer an individual of race $y$ from blockgroup $i$ to blockgroup $j$ such that the following relation holds: 
+% 	\begin{equation}
+% 		p(y|X = i') \geq p(y|X = j')
+% 	\end{equation}
+% 	for all blockgroups $i'$ within radius $r$ of $i$ and all blockgroups $j'$ within radius $r$ of $j$.
+The transfer principle states that, in this case, $J'(X,Y) < J(X,Y)$, where $J'(X,Y)$ is the local information post-transfer.
+Using \eqref{comp}, it suffices to show that
+
+% 	\begin{equation}
+% 		(P_k + 1) I'(k) + (P_\ell - 1) I'(\ell) < P_k I(k) + P_\ell I(\ell),
+% 	\end{equation}
+% 	where $I'(k)$ is the mutual information in hex $k$ post-transfer, for any grid cells $k$ overlapping $i$ and $\ell$ overlapping $j$.

### 2016-11-04 3_local_info.R

@@ -58,39 +58,39 @@ if(file.exists('throughput/info_cache.csv')){
 		I_XY = numeric(),
 		J_XY = numeric())
 }
-
+city = 'Boston'
 for(city in cities){
-	tracts <- readOGR(dsn = paste0('data/cities/',city), 
-	                  layer = 'geo', 
+	tracts <- readOGR(dsn = paste0('data/cities/',city),
+	                  layer = 'geo',
 	                  verbose = FALSE)
-	grid_dir        <- paste0('throughput/grids/',city)	
+	grid_dir        <- paste0('throughput/grids/',city)
 	grid_tract_path <- paste0('throughput/grid_tracts/', city, '.csv')
-  
+
 	if(file.exists(grid_tract_path) & dir.exists(grid_dir)){
 	  grid_tract <- read_csv(grid_tract_path)
-	  grid <- readOGR(dsn = paste0('throughput/grids/',city), 
-	                  layer = 'grid', 
+	  grid <- readOGR(dsn = paste0('throughput/grids/',city),
+	                  layer = 'geo',
 	                  verbose = FALSE)
-	  analysis <- info_analysis(tracts, 
-	                            columns = races, 
+	  analysis <- info_analysis(tracts,
+	                            columns = races,
 	                            resolution = resolution,
-	                            grid_polys = grid, 
+	                            grid_polys = grid,
 	                            grid_tract = grid_tract)
 	}else{
-	  analysis <- info_analysis(tracts, 
-	                            columns = races, 
+	  analysis <- info_analysis(tracts,
+	                            columns = races,
 	                            resolution = resolution)
 	  write_csv(analysis$grid_tract, paste0('throughput/grid_tracts/',city, '.csv'))
 	  tryCatch({writeOGR(analysis$grid, paste0("throughput/grids/", city),
-	                     'grid', 
-	                     driver = 'ESRI Shapefile', 
+	                     'grid',
+	                     driver = 'ESRI Shapefile',
 	                     morphToESRI = TRUE)},
 	           error = function(e) e)
 	}
-	
+
 	# summarized output
 	out <- data.frame(
-	  city       = city, 
+	  city       = city,
 	  area       = sum(tracts@data$area),
 	  population = sum(tracts@data$total),
 	  density    = sum(tracts@data$total) / sum(tracts@data$area),
@@ -98,7 +98,7 @@ for(city in cities){
[…]

### 2016-11-04 figs.Rmd
cluster into 4 instead of six

@@
 
-tab <- data.frame(city = character(), total_info = numeric(), info_6 = numeric(), info_20 = numeric())
+tab <- data.frame(city = character(), total_info = numeric(), info_4 = numeric(), info_20 = numeric())
 for(city in c('Philadelphia', 'Atlanta')){
@@
 	columns = c('Black', 'White', 'Asian', 'Hispanic', 'Other')
-  new_row <- data.frame(city = city, total_info = 0, info_6 = 0, info_20 = 0)
-	for(nclusters in c(6, 20)){
+  new_row <- data.frame(city = city, total_info = 0, info_4 = 0, info_20 = 0)
+	for(nclusters in c(4, 20)){
   	c <- cluster_map_percent_black(city, paste0(city, '_binary'), nclusters = nclusters)

### 2016-11-04 figs.Rmd

@@
 	theme_minimal() +
-  theme(text=element_text(size=16, family="Palatino")) + 
+  theme(text=element_text(size=16)) + 
 	annotation_logticks(size = .1, sides = 'b', scaled = T) + 
@@
 	geom_polygon(aes(nclust, I_XY, fill = city), data = filled, alpha = .2) + 
-	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary, family = 'Palatino') + 
+	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
 	# geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
 	facet_wrap(~city) + 
-	scale_fill_discrete(guide = F) + 
-	scale_color_discrete(guide = F) + 
+  scale_fill_brewer(palette = 'Set1', name = NULL, guide = 'none') +
+  scale_color_brewer(palette = 'Set1', name = NULL, guide = 'none') + 
+	# scale_color_discrete(guide = F) + 
 	xlab('Number of clusters') + 
@@
   ggplot(aes(x = nclust, y = I_XY)) +
-  geom_path(aes(group = city), color = 'grey') + 
+  geom_path(aes(group = city), color = 'grey', alpha = 0.6) + 
   geom_path(aes(group = city, color = city), data = highlighted) +
   geom_text_repel(aes(label = city, color = city), data = end_points) +
-  scale_x_continuous(trans = 'log10') +
+  theme_light() +
+  scale_x_continuous(trans = 'log10', limits = c(1,NA)) +
   scale_y_continuous(limits = c(0,.5)) + 
-  theme_light() +
   scale_color_brewer(palette = 'Set1', name = NULL, guide = 'none') +
@@
 		geom_polygon(aes(x = long, y= lat, group = group), data = to_overlay, fill = NA, size = .3, color = 'black') + 
-	  theme(text=element_text(size=16, family="Palatino"),
+	  theme(text=element_text(size=16),
 	        legend.justification=c(0,0), legend.position=c(.65,0),

### 2016-11-04 figs.Rmd

@@
 	geom_path(aes(x = nclust, y = I_XY, group = city, color = city)) +
-	scale_x_continuous(trans = 'log10', limit = c(1, NA)) +
-	scale_y_continuous(limit = c(0,max(sub_curves$I_XY))) +
-	theme_minimal() +
-  theme(text=element_text(size=16)) + 
+	scale_x_continuous(trans = 'log10', limit = c(1, NA), expand = c(0, 0)) +
+	scale_y_continuous(limit = c(0,max(sub_curves$I_XY)), expand = c(0, 0)) +
+	theme_light() +
+  theme(text=element_text(size=16),
+        panel.margin = unit(2, "lines"),
+        panel.border = element_blank(),
+        strip.background = element_rect(fill = "white", color = "white", size = 2),
+        strip.text.x = element_text(size = 16, color = 'black')) + 
 	annotation_logticks(size = .1, sides = 'b', scaled = T) + 
@@
 	# scale_color_discrete(guide = F) + 
-	xlab('Number of clusters') + 
-	ylab('I(C,Y)') + 
-	ggtitle('AUC as Clustering Evaluation')
+	xlab('Number of clusters (n)') + 
+	ylab(expression(I(C[n],Y)))
 
@@
   theme_light() +
-  scale_x_continuous(trans = 'log10', limits = c(1,NA)) +
-  scale_y_continuous(limits = c(0,.5)) + 
+  scale_x_continuous(trans = 'log10', limits = c(1,NA), expand = c(0, 0)) +
+  scale_y_continuous(limits = c(0,.5), expand = c(0, 0)) + 
   scale_color_brewer(palette = 'Set1', name = NULL, guide = 'none') +

### 2016-11-04 figs.Rmd
label adjustments to binary loss curves

--- 2016-11-04 figs.Rmd
@@
 	geom_path(aes(x = nclust, y = I_XY, group = city, color = city)) +
-	scale_x_continuous(trans = 'log10', limit = c(1, NA)) +
-	scale_y_continuous(limit = c(0,max(sub_curves$I_XY))) +
-	theme_minimal() +
-  theme(text=element_text(size=16)) + 
+	scale_x_continuous(trans = 'log10', limit = c(1, NA), expand = c(0, 0)) +
+	scale_y_continuous(limit = c(0,max(sub_curves$I_XY + 0.05)), expand = c(0, 0)) +
+	theme_light() +
+  theme(text=element_text(size=16),
+        panel.margin = unit(2, "lines"),
+        panel.border = element_blank(),
+        strip.background = element_rect(fill = "white", color = "white", size = 2),
+        strip.text.x = element_text(size = 16, color = 'black')) + 
 	annotation_logticks(size = .1, sides = 'b', scaled = T) + 
@@
 								ymax = ymax,
-								group = city),
-			  color = 'grey',
-			  alpha = 0) +
+								group = city,
+								color = city),
+			  # color = 'grey',
+			  alpha = 0,
+			  size = 0.25) +
 	# geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) + 
 	geom_polygon(aes(nclust, I_XY, fill = city), data = filled, alpha = .2) + 
-	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
+	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary, size = 6) + 
 	# geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
@@
 	group_by(city) %>%
-	filter(nclust == max(nclust),
+	filter(nclust == 10,
 	       city %in% highlights)
+
+end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit'),
+                         nclust = c(50, 7, 8),
+                         I_XY = c(.35, .18, .37))
 
@@
   geom_path(aes(group = city, color = city), data = highlighted) +
-  geom_text_repel(aes(label = city, color = city), data = end_points) +
+  geom_text(aes(label = city), data = end_points) +
   theme_light() +
-  scale_x_continuous(trans = 'log10', limits = c(1,NA)) +
-  scale_y_continuous(limits = c(0,.5)) + 
+  scale_x_continuous(trans = 'log10', limits = c(1,NA), expand = c(0, 0)) +
+  scale_y_continuous(limits = c(0,.5), expand = c(0, 0)) + 
   scale_color_brewer(palette = 'Set1', name = NULL, guide = 'none') +

### 2016-11-04 figs.Rmd
auc example consistent with binary loss curves

@@
 
-sub_curves <- loss_curves %>%
+sub_curves <- binary_loss_curves %>%
 	filter(city %in% cities)
@@
 	scale_x_continuous(trans = 'log10', limit = c(1, NA), expand = c(0, 0)) +
-	scale_y_continuous(limit = c(0,max(sub_curves$I_XY + 0.05)), expand = c(0, 0)) +
+	scale_y_continuous(limit = c(0,max(sub_curves$I_XY + 0.02)), expand = c(0, 0)) +
 	theme_light() +
@@
 	geom_polygon(aes(nclust, I_XY, fill = city), data = filled, alpha = .2) + 
-	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary, size = 6) + 
+	geom_text(aes(x = 50, y = .15, label = paste('AUC = ', round(AUC,2))), data = sub_summary, size = 6) + 
 	# geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
@@
                          nclust = c(50, 7, 8),
-                         I_XY = c(.35, .18, .37))
+                         I_XY = c(.34, .18, .37))
 

### 2016-11-04 figs.Rmd
checkerboard highlighted

@@
 
-checkerboard_illustration <- function(){
+overlay <- data.frame(x = 6.5, y = 4.5)
+
 expand.grid(x = 1:8, y = 1:8) %>%
@@
 		  plot.margin=unit(c(1,1,0,0),"mm"),
-			  strip.text=element_text(vjust=-100))
-}
+		  strip.text=element_text(vjust=-100)) + 
+  geom_tile(aes(x = x, y = y), color = 'red', alpha = 0, size = 0.5, data = overlay)
 
-g <- checkerboard_illustration() 
-g
+# g <- checkerboard_illustration() 
+# g
 # g + theme(text=element_text(size=10, family="Palatino"))

### 2016-11-10 figs.Rmd
non-binary loss curves

@@
 
+adjust_thresh = 0.01
+highlights  <- c('Philadelphia', 'Atlanta', 'Detroit')
+curves <- loss_curves %>% 
+  mutate(exclude = (nclust != 1) & (cummax(I_XY - lag(I_XY,default = 0)) < adjust_thresh) & (nclust < 10)) %>% 
+	filter(!exclude,
+	       max(I_XY > 0.0)) %>%
+	mutate(nclust = row_number())
+
+highlighted <- curves %>% 
+  filter(city %in% highlights)
+
+background <-  curves %>% 
+  filter(!(city %in% highlighted))
+
+end_points <- highlighted %>%
+	group_by(city) %>%
+	filter(nclust == 10,
+	       city %in% highlights)
+
+end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit'),
+                         nclust = c(50, 7, 8),
+                         I_XY = c(.34, .18, .37))
+
+# grey background
+
+background %>% 
+  ggplot(aes(x = nclust, y = I_XY)) +
+  geom_path(aes(group = city), color = 'grey', alpha = 0.6) + 
+  geom_path(aes(group = city, color = city), data = highlighted) +
+  geom_text(aes(label = city), data = end_points) +
+  theme_light() +
+  scale_x_continuous(trans = 'log10', limits = c(1,NA), expand = c(0, 0)) +
+  scale_y_continuous(limits = c(0,.5), expand = c(0, 0)) + 
+  scale_color_brewer(palette = 'Set1', name = NULL, guide = 'none') +
+  xlab('Number of clusters (n)') + 
+	ylab(expression(I(C[n],Y))) +
+  # geom_vline(xintercept=2, alpha = .2) + 
+  annotation_logticks(sides = 'b', size = .2)
+
+  ggsave('figs/loss_curves.pdf', width = 7, height = 4)
+  embed_fonts("figs/loss_curves.pdf", outfile="figs/loss_curves.pdf")
+
+```
+
+```{r}
+
 # dippy point that doesn't seem to be true here: if you don't have any excluded points on your curve, your AUC should be the same.

### 2016-11-10 3_local_info.R
cleaning up

@@ -13,20 +13,6 @@ library(rgdal, quietly = TRUE)
 library(data.table)
 
 races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
-resolution <- 1/2
-
-
-# Toy cities ----------
-
-# checkerboard_illustration()
-# ggsave('report/figs/toy.png', width = 5, height = 1.9)
-
-# Methodological Illustration ----------------
-
-# method_illustration()
-# ggsave('report/figs/method.png', width = 3, height = 3)
-
-
 
 cities <- list.files('data/cities')
 resolution <- 1
@@ -41,6 +27,7 @@ if(!dir.exists('throughput/grids')){
 if(!dir.exists('throughput/grid_tracts')){
 	dir.create('throughput/grid_tracts')
 }
+
 
 # MAIN COMPUTATION ---------------
 
@@ -71,17 +58,21 @@ for(city in cities){
 	  grid <- readOGR(dsn = paste0('throughput/grids/',city), 
 	                  layer = 'grid', 
 	                  verbose = FALSE)
+	  print(city)
 	  analysis <- info_analysis(tracts, 
 	                            columns = races, 
 	                            resolution = resolution,
 	                            grid_polys = grid, 
 	                            grid_tract = grid_tract)
 	}else{
-	  analysis <- info_analysis(tracts, 
-	                            columns = races, 
-	                            resolution = resolution)
-	  write_csv(analysis$grid_tract, paste0('throughput/grid_tracts/',city, '.csv'))
-	  tryCatch({writeOGR(analysis$grid, paste0("throughput/grids/", city),
+	  
+	  tryCatch(
+	    {print(city)
+	    analysis <- info_analysis(tracts, 
+	                              columns = races, 
+	                              resolution = resolution)
+	    write_csv(analysis$grid_tract, paste0('throughput/grid_tracts/',city, '.csv'))
+	    writeOGR(analysis$grid, paste0("throughput/grids/", city),
 	                     'grid', 
 	                     driver = 'ESRI Shapefile', 
 	                     morphToESRI = TRUE)},

### 2016-11-10 revision.tex
cleaning up

--- 2016-10-06 revision.tex
@@
 	\subsection{Illustrations}
-		\subsubsection{Comparative Analysis: Detroit and Philadelphia}
+		% \subsubsection{Comparative Analysis: Detroit and Philadelphia}
 			To illustrate these methods, we use them to compare the cities of Philadelphia and Detroit.
@@
 As we show, computing $H(Y)$, $I(X,Y)$, and $J(X,Y)$ together highlights similarities and differences between these cities.
-			\todo{We need to generate a nice summary plot here and then fill this section in.}
+			% \todo{We need to generate a nice summary plot here and then fill this section in.}
 			\begin{figure}
@@
 
-		\subsubsection{Comparative Diversity Profiles}
-			\todo{Is there a fun way to explore the density point? E.g. figure out people per ``neighborhood'' in some way? Think about what I would want the analysis here to be.}
+		% \subsubsection{Comparative Diversity Profiles}
+			% \todo{Is there a fun way to explore the density point? E.g. figure out people per ``neighborhood'' in some way? Think about what I would want the analysis here to be.}
 
@@
 \section{Discussion} \label{sec:discussion}
-
-\bibliography{/Users/phil/bibs/library.bib}{}
-\bibliographystyle{plain}
+BLAH BLAH BLAH
+% \bibliography{/Users/phil/bibs/library.bib}{}
+% \bibliographystyle{plain}
 

### 2016-11-10 submission.tex

--- 2016-11-02 submission.tex
@@
 If this is the case, add \parshape=0 to the end of the list environment.
-\dropcap{I}nformation theory...blah blah blah.
+\dropcap{T}his is a story about the hierarchical structure of demographic and spatial form.
+We need to tell this story because traditional methods only give us a single number, whereas what we really want to see is spatial structure at varying levels of aggregation.
+This is a story about practical computation for seeing the hierarchical structure of social variation.
+This is a story about finding the smallest unit of spatial variation -- the local information -- and then scaling up in a smart way.
+It's a story about quantifying the intrinsic structure and then going into the world of algorithms.
+It's a story about computing with the hierarchical structure.
+It's a story about explicitly incorporating space on different levels.
+It's a story about not-too-novel algorithms applied in new ways.
+It's a story about finding a principled role for information theory in the study of urban form.
+It's a story about mathematical coherence.
+It's a story about love.
+It's a story about non-convexity and choosing to be ok with that.
+It's a story about finding new ways to view what's already visible, to organize what's there in a useful way.
+It's a story about computation and open data.
+It's open science.
+It's a story about not everything needing to be statistics.
+It's a story about taking the data as it is.
+It's a story about how to learn from the Census.
+It's a story about how we can improve on sociological methodology using results from physics and mathematics.
+It's a story about how physics and mathematics are missing insights that you only get from the social sciences.
+It's a story about collaboration.
+It's a story about how one day we'll get the dynamical picture too but we haven't coded that up yet or understood the mathematics.
+It's not yet a story about the quadratic structure of local variation through derived dispersions, but we'll get there one day even if that's just on arXiv.
+It's a story about the hierarchical structure, about different ways that information scales in different places.
+
+Multiple scales.
+How can we learn to see them all?
+Well, we can consider properties of information scaling at various levels.
+But it won't really do to organize these levels in arbitrary ways, we should actually seek the structure as best we can.
+Not optimally, but as best we can.
+Try to see what's already there, what's in front of us.
+Nothing is hidden.
+We can do some motivation around fractical structure, maybe cite something from batty or PNAS on fractal structure of cities and fractal structure of demographics, little pockets of variability.
+Just look at the US, where you have both global trends across the whole country and then smaller trends within each area and then within each neighborhood.
+Maybe New York is an example, there are different little pockets within each borough as well as trends between boroughs.
+So we can tell a bit of a story that way.
+Recall Roberto's question: is Detroit segregated?
+
+The agenda is to flexibly study demographic separation in space on varying scales.
+To do that we first need to understand demographic separation on the lowest level, and this requires calculus.
+So we do the calculus and we find that we get the trace of the Fisher information out at the end, which is pleasing.
+Ok, so that's the local picture, hooray!
+One way to go from the local picture to the global picture is just to integrate over everything, after which we find that we get a nice averaged measure that captures the high-level differences between cities.
+But this is in some sense a blunt instrument; we don't exactly learn too much about in what ways they are different.
+A more subtle instrument is to use an algorithm -- the information-maximizing agglomerative clustering -- to find the structure itself, and then we can quantify the performance of the algorithm and take that as a measure of complexity directly.
+What's nice about this is that our complexity measure is actually directly related to our ability to efficiently summarise the structure of urban form.
+Perhaps we should state the algorithm explicitly as a figure.
+Our ultimate goal is to show that this material is of relevance to a fairly broad audience, hopefully including some sociologists, and we can do some evangelizing at conferences etc.
 
@@
 	Each individual can belong to one of a set $\mathcal{Y}$ of racial groups, such as $\mathcal{Y} = \{\text{Asian}, \text{ Black}, \text{ Hispanic}, \text{ Other}, \text{ White}\}$. 
-	The cross-tabulation of individuals and groups defines a probability $p(X,Y)$ that a randomly chosen individual from the city will live at location $X \in R$ and belong to group $Y$. 
+	The cross-tabulation of individuals and groups defines a probability $p(X,Y)$ that a randomly chosen individual from the city will live at location $X \in R$ and belong to group $Y \in \mathcal{Y}$. 
 	The population density at $x$ is $p(x) = \sum_{y \in \mathcal{Y}} p(x,y)$. 
@@
 \subsection*{Entropy and Global Diversity}
-	\emph{Global} measures of the city depend only on $p(Y)$, the global composition of racial groups. 
+	A global measure depends only on $p(Y)$, the global composition of racial groups. 
 	The global Shannon entropy $H(Y)$ measures global diversity by comparing the global distribution $p(Y)$ to the uniform distribution $u(Y)$: 
@@
 	\end{equation}
+	Since $H(Y)$ is a global metric, it gives a high-level summary but is insensitive to spatial structure.
 
-	Local-to-global measures depend on comparisons between properties encoded by the local compositions $p(y|x)$ and the global composition $p(y)$.
-
-	Local measures depend only on comparisons between local compositions $p(y|x)$ and other nearby local compositions, without reference to the global composition $p(y)$. 
 \begin{figure}%[tbhp]
@@
 		\begin{tabular}{l | c c c} 
-		City & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
+		\textbf{City} & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
 		\hline
@@
 		\caption{
-			\todo[inline=true]{Illustrate computation of $J$ on this fig?}
-			Schematic cities illustrating three dimensions of segregation. 
+			\emph{(Top)}: Schematic cities illustrating three dimensions of segregation. 
 			City (a) lacks global diversity, while cities (b)-(d) are maximally globally diverse. 
 			City (b) is perfectly spatially even, whereas cities (c) and (d) are maximally uneven. 
-			City (d) has higher spatial exposure than city (c) due to more contact points between racially distinct neighborhoods.
+			City (c) has lower spatial exposure than city (d), since a typical small area of (c) (shown in highlight) is monoracial.
+			In contrast, a typical small neighborhood of (d) has multiple contact points between racially distinct areas. 
+			\emph{(Bottom)}: Information measures of each city. Entropy separates cities (a) and (b); mutual information separates (b) and (c), and mean local information separates (c) and (d). 
 		} \label{fig:checkerboard}
@@
 	The operator $\E_X$ denotes average or expectation with respect to the probability distribution of $X$. 
-	$I(X,Y)$ is the \emph{mutual information}, a fixture of information theory. \cite{Bettencourt2015}
-
+	The quantity $I(X,Y)$ is called the \emph{mutual information} in information theory \cite{Cover1991,Csiszzr2004,Bettencourt2015}.  
 	When all groups are evenly distributed, such as in Figure \ref{fig:checkerboard}(b), each location has the same composition as the global composition, and $I(X,Y) = 0$. 
@@
 
-	Global and local-global measures can illuminate important high-level urban structure, but miss an important class of phenomena. 
+	Global and local-to-global measures can illuminate important high-level urban structure, but miss an important class of phenomena. 
 	The model two-race cities (c) and (d) in Figure \ref{fig:checkerboard} each have the same values of $H(Y)$ and $I(X,Y)$. 
@@
 	In sociology or urban planning, this may be the difference between metastasized segregation and healthy variation. 
-	The difference is intrinsically local, defined in terms of relations between local compositions $p(Y|X)$ and nearby other local compositions.
+	The difference is intrinsically local: while the typical small neighborhood of city (c) is monoracial, the typical small neighborhood of (d) contains multiple contact points between different racial groups.
 
-	To measure local variation, imagine segmenting the city into a grid of small neighborhoods. 
-	The \emph{local mutual information} in the neighborhood $B$ is the mutual information between $X$ and $Y$, restricted to $B$:
+	To measure local variation, imagine segmenting the city into a grid of small cells. 
+	The \emph{local mutual information} in the cell $B$ is the mutual information between $X$ and $Y$, restricted to $B$:
 	\begin{equation*}
@@
 	\end{equation*}
+	The local information can be interpreted as the degree of change in demographics one would observe if one started in $B$ and then walked a small distance away.
+
+	If we regard the data as a discretization of a smooth probability field, then it is possible to show using arguments from information geometry \cite{Amari2000} that the local information in a cell $B$ of radius $r$ is an approximation of the Fisher information $\mathcal{J}$:
+	\begin{equation}
+		I(X,Y|X \in B) \cong \frac{r^2}{4} \text{tr } \mathcal{J}_Y(x)\;, \label{eq:fisher_approx}
+	\end{equation}
+	where $[\mathcal{J}_Y(x)]_{ij} = \E_Y[-\partial_{x_i} \partial_{x_j} \log p(Y|x)]$ is the Fisher information matrix and $\text{tr } \mathcal{J}_Y(x)$ is the sum of its eigenvalues.
+
 	The \emph{mean local information} measures the average local variation in group composition across the entire city:  
@@
 	\end{equation}
-	If we regard the data as a discretization of a smooth probability field, then it is possible to show using arguments from information geometry \cite{Amari2000} that the local information in a neighborhood $B$ of radius $r$ is an approximation of the Fisher information $\mathcal{J}$:
-	\begin{equation}
-		I(X,Y|X \in B) \cong \frac{r^2}{4} \text{tr } \mathcal{J}_Y(x)\;, \label{eq:fisher_approx}
-	\end{equation}
-	where $[\mathcal{J}_Y(x)]_{ij} = \E_Y[-\partial_{x_i} \partial_{x_j} \log p(Y|x)]$ is the Fisher information matrix and $\text{tr } \mathcal{J}_Y(x)$ is the sum of its eigenvalues.
+	Figure \ref{fig:Atlanta_philly} shows the computation of $J(X,Y)$ for Atlanta and Philadelphia.
+While the two cities have similar entropies $H(Y)$ and mutual informations $I(X,Y)$, their spatial structures are substantially different.
+Atlanta is characterized by a stark North-South divide between white and black residents.
+As a result, there are many areas in Atlanta where the local information is near $0$.
+These are areas in which one can walk for $1km$ or more without experiencing significant demographic change.
+Such areas are relatively rare in Philadelphia, which consists of a finer patchwork of demographically varying neighborhoods.
+Philadelphia therefore has a higher value of mean local information, reflecting greater spatial complexity in demographic variation.
 
@@
 			\caption{
-				Spatial analysis of residential segregation in Atlanta and Philadelphia. 
 				(\emph{Top}): Percentage of black residents. 
 				(\emph{Middle}): Hexgrid of spatial resolution 1km imposed over each city. Within each hex $B$ we compute the local information local information $J_Y(B) = I(X,Y | X \in B)$. 
-				(\emph{Bottom}): Numerical summary, including the three information measures.
+				(\emph{Bottom}): Numerical summary including entropy, mutual information, and mean local information. 
 			} \label{fig:Atlanta_philly}
@@
 
-\subsection*{Comparative Urban Analysis}
+\subsection*{Separation Profiles in the US}
 
@@
 			Comparative global segregation measures in major U.S. cities, measured using racial categories $\mathcal{Y} = \{\text{Asian}, \text{ Black}, \text{ Hispanic}, \text{ Other}, \text{ White}\}$. 
-			We used a hexgrid of spatial resolution $r = 1$km for the computation of $J(X,Y)$.
+			A hexgrid of spatial resolution $r = 1$km was for the computation of $J(X,Y)$. The grey line is the linear least-squares regression line provided as a visual aid. 
 		} \label{fig:mutual_fisher}
@@
 In practice, however, there may be coherent biracial or multiracial patterns in space.
-Fortunately, it is possible to motivate a structure-finding algorithm for identifying high-level demographic trends without restriction to monoracial regions.
-
+Information methods define a useful approach.
 	\begin{figure}
 
-		\includegraphics[width = .5\linewidth]{figs/Atlanta_clusters_binary_6.pdf} 
-		\includegraphics[width = .5\linewidth]{figs/Philadelphia_clusters_binary_6.pdf} \\
+		\includegraphics[width = .5\linewidth]{figs/Atlanta_clusters_binary_4.pdf} 
+		\includegraphics[width = .5\linewidth]{figs/Philadelphia_clusters_binary_4.pdf} \\
 		\includegraphics[width = .5\linewidth]{figs/Atlanta_clusters_binary_20.pdf}
@@
 		\begin{tabular}{l | c c c}
-			\bfseries City  & $I(C_6, Y)$ & $I(C_{20},Y)$ & $I(X,Y)$ \\\hline
+			\bfseries City  & $I(C_4, Y)$ & $I(C_{20},Y)$ & $I(X,Y)$ \\\hline
 			\csvreader[late after line=\\]{figs/clustering_summary.csv}{}
@@
 		\caption{
-			Example clusterings in Atlanta and Philadelphia using \eqref{eq:cluster_opt}, using $6$ and $20$ clusters. 
-			The table below summarizes the between-groups mutual information captured by each cluster and compares it to the mutual information in the unaggregated blockgroup level data.
+			\emph{(Top)}:Example clusterings in Atlanta and Philadelphia using \eqref{eq:cluster_opt}, using $4$ and $20$ clusters. 
+			\emph{(Bottom)}: Comparison of information content of clusterings and total mutual information in each city. 
 		} \label{fig:clusterings}
-
 	\end{figure}
-	Let $C$ be a random variable with the property that $H(C|X) = 0$; i.e.
-$X$ completely determines $C$.
-We can think of $C$ as group labels for the spatial locations $X$.
-Then, the chain rule of information\footnote{See \cite{Cover1991}} implies that
+	Let $C$ be a grouping of the locations $X$.
+Mathematically, $C$ is a random variable with the property that that $H(C|X) = 0$.
+The chain rule of information\footnote{See \cite{Cover1991}} implies that
 	\begin{equation}
@@
 	\end{equation}
-
-	Equation \eqref{eq:chain_rule} motivates a natural algorithm for identifying sociospatial structure.
-The core idea is that a clustering $C$ is ``good'' when it captures most of the intrinsic information in the system in the between-groups term $I(C,Y)$.
-Mathematically, we therefore wish to maximize $I(C,Y)$ subject to a fixed number of clusters, an approach popularized by \cite{Dhillon2003,Banerjee2005}.
+	Analogous to sum-of-squares maximization in statistics, we wish to maximize $I(C,Y)$ subject to a fixed number of clusters, an approach popularized by \cite{Dhillon2003,Banerjee2005}.
 In full generality, this is a challenging discrete optimization problem that may not be computationally tractable.
-Additionally, for interpretability, we need to apply a spatial constraint: locales should be clustered together only when they are adjacent in space.
+Additionally, for interpretability, should be clustered together only when they are adjacent in space.
 To handle these two problems, we use a greedy algorithm based on agglomerative clustering.
-At stage of the algorithm, we consider the problem of choosing a pair of locales $\{i*, j*\}$ to aggregate together.
-A good choice of locales leads to minimal information loss, which we may quantify as follows.
+At stage of the algorithm, we consider the problem of choosing a pair of locales $\{i*, j*\}$ such that aggregating them minimizes information loss.
+We quantify the loss as follows.
 The mutual information over the full complement $\mathcal{K}$ of locales may be written:
@@
 It is also symmetric in $i$ and $j$.
-It is not a true metric because it does not satisfy the triangle inequality.
-It can nevertheless be used to motivate a form of agglomerative hierarchical clustering, where at each phase of the algorithm we compute
+We can use it to motivate a form of agglomerative hierarchical clustering, where at each phase of the algorithm we compute
 	\begin{equation}
@@
 	\end{equation}
-	and combine $i^*$ and $j^*$ into a single new region.
-We repeat this procedure until only $N$ regions remain.
-The optimization over $i$ and $j$ ensures that the groupings so defined approximate maximum between-group information partitions, and the constraint that $i$ must neighbor $j$ ensures that the structure so determined is spatially contiguous.
-Equation \eqref{eq:cluster_opt} thus defines \emph{spatially constrained, information-maximizing agglomerative clustering}.
+	and combine $i^*$ and $j^*$ into a single new locale, repeating until only a specified number $n$ locales remain. 
+	Equation \eqref{eq:cluster_opt} defines \emph{spatially constrained, information-maximizing agglomerative clustering}. 
 	Since the algorithm is greedy, it possesses no guarantees of optimality, but in practice its performance leads to intuitive, demographically-coherent regions. 
-	\todo[inline=true]{Should we state the algorithm in a figure?}
+	The resulting regions may be viewed as coarse-grainings of the spatial structure, which emphasize high-level patterns and abstract away details.
+
+	Figure \ref{fig:clusterings} demonstrates this information-maximizing agglomerative clustering for Atlanta and Philadelphia, using 4 and 20 clusters.
+In Atlanta, the algorithm immediately highlights the prominent North-South divide, while in Philadelphia it segments two predominantly Black regions on either side of the Schuykill river from the rest of the city.
+As we allow the algorithm to use progressively more clusters, it highlights progressively more subtle patterns of variation.
+Strikingly, in Atlanta just 4 clusters are sufficient to capture 0.27 nats of information, where in Philadelphia 20 clusters are needed.
+Philadelphia has more complex sociospatial trends: compared to Atlanta, more complex models are necessary to model it to a similar degree of accuracy.
 
@@
 			% \begin{subfigure}[b]{0.45\linewidth}
-			\includegraphics[width = .4\linewidth]{figs/example_cluster_map.pdf}
+			\includegraphics[width = .5\linewidth]{figs/example_cluster_map.pdf}
 			% \caption{} \label{subfig:detroit_map}
@@
 			% \begin{subfigure}[b]{0.45\linewidth}
-			\includegraphics[width = .4\linewidth]{figs/example_clusters_detailed.pdf}
+			\includegraphics[width = .45\linewidth]{figs/example_clusters_detailed.pdf}
 			% \caption{} \label{subfig:detroit_groups}
@@
 
-	This technique extends naturally to multi-group settings, though the visualization of results requires more subtlety.
-In Figure \ref{fig:detroit}, we conduct an illustrative analysis of multigroup segregation segregation in Detroit and the surrounding Wayne County.
+	Information-maximizing agglomerative clustering also applies to multigroup analysis.
+Figure \ref{fig:detroit} shows a multigroup clustering in Detroit and the surrounding Wayne County.
 The prominent clusters A and B reflect the primary white/black division in the city, with A representing the white suburbs and B the denser black inner city.
@@
 
-	Aside from simply highlighting geographic trends, this analysis also enables the quantitative study of traditional concerns in segregation studies, such as concentration.
+	This technique also enables novel quantitative methods for studying traditional concerns in segregation studies.
 According to \cite{Massey1988}, ``\emph{concentration} refers to the degree of a group's agglomeration in urban space.'' The concept of concentration invites us to consider the composition and density of each cluster.
@@
 We can also learn about pairwise group exposure.
-For example, from Figure  \ref{subfig:detroit_groups}, we observe that Asian residents tend to be found alongside white residents, but more rarely alongside black or Hispanic residents.
-Hispanic residents tend to group in Mexicantown and in predominantly white neighborhoods -- not black ones.
+For example, from Figure  \ref{fig:detroit}, Asian residents tend to be found alongside white residents, but more rarely alongside black or Hispanic residents.
+Hispanic residents tend to group in Mexicantown and in predominantly white neighborhoods, but not black ones.
 
-\subsection*{Evaluating Clusterings}
-	As discussed, the mean local information $J(X,Y)$ can be viewed as a measure of the number and character of transitional areas between regions of differing demographic structure.
-It is low in Atlanta because the city is dominated by one, large North-South boundary while it is higher in Philadelphia due to many points of contact between neighborhoods of differing composition.
-Sociospatial transitions are also precisely the ``fault lines'' along which we would expect the clustering algorithm \eqref{eq:cluster_opt} to split the city.
-It is therefore not a coincidence that Philadelphia both has a higher value of $J(X,Y)$ and requires a greater number of clusters to convey a similar amount of sociospatial information.
+\subsection*{Local Information and Spatial Complexity}
+	The mean local information measures the number and character of spatial transitions between demographically distinct regions.
+When there are many, we would expect that information-maximizing agglomerative clustering would require more clusters to capture similar amounts of information.
+We can empirically quantify the relationship between $J(X,Y)$ and the performance of information-maximizing clustering as follows.
+Consider the between-groups information $I(C_n,Y)$ plotted as a function of the number $n$ of clusters.
+The result is a curve reflecting how information scales at varying levels of aggregation, as shown in Figure \ref{fig:AUC} shows examples of these curves for Atlanta, Detroit, and Philadelphia.
 
-	To quantify the relationship between the mean local information $J(X,Y)$ and the clustering method \eqref{eq:cluster_opt}, we must first quantify the performance of the latter.
-We can do this by considering the number of clusters necessary to convey a given fraction of overall mutual information.
-Consider the between-groups information $I(C_n,Y)$ plotted as a function of the number $n$ of clusters.
-The result is a curve reflecting how information scales at varying levels of aggregation.
-Figure \ref{fig:AUC} shows examples of these curves for Atlanta, Detroit, and Philadelphia, with $N$ plotted on a logarithmic scale.
-The area under the curve (AUC) serves a measure of how early information is ``captured'' by the clustering.
-When the AUC is high, small numbers of clusters suffice to capture most of the information; when it is low more complex clustering is necessary.
-The AUC can be computed explicitly via the formula
+	The area under these curves is a measure of how early information is ``captured'' by the clustering.
+When the AUC is high, small numbers of clusters suffice to capture most of the information.
+Atlanta and Detroit illustrate this case, as both can be starkly divided into monolithic Black and non-Black regions.
+In Philadelphia, in contrast, the AUC is lower and more complicated models are necessary, as was illustrated in \ref{fig:clusterings}.
+The explicit formula for the AUC is
 	\begin{equation}
@@
 	\end{equation}
-
-	AUC is therefore an algorithmic measure of sociospatial complexity, which is high when there are few sociospatial boundaries and low when there are many.
-Previously, we also argued that the quantities $H(Y)$, $I(X,Y)$, and $J(X,Y)$ may be viewed as information measures of sociospatial complexity, and particularly that $J(X,Y)$ measures the presence of sociospatial boundaries.
-We should therefore expect that these two sets of measures are related.
-A simple way to test for the relationship is via linear least squares regression.
-We aim to approximate the AUC via the formula
+	To show the relationship between $J(X,Y)$, we conduct a simple linear least-squares fit of the form
 		\begin{equation*}
@@
 		\end{equation*}
-	The results of this approximation are shown in Figure \ref{fig:regression}.
-Collectively, the three information measures suffice to explain an adjusted 75\% of the variation in the AUC, confirming that both the information measures and the algorithmic AUC measure the same underlying sociospatial structure.
-The coefficients are also instructive.
+	Figure \ref{fig:regression} gives the summary table for this fit.
+Collectively, the three information measures suffice to explain an adjusted 75\% of the variation in the AUC, showing that the AUC is highly dependent on the information structure of sociospatial variation.
+The signs of the coefficients are also interpretable.
 The largest magnitude by far is the coefficient of $I(X,Y)$; all else equal, clustering is more effective when groups are distributed unevenly.
-The negative sign on the coefficient of the mean local information $J(X,Y)$ confirms our specific expectations from above; for fixed $I(X,Y)$ and $H(Y)$, a city with higher $J(X,Y)$ (more sociospatial complexity) is more difficult to model or simplify than one with lower $J(X,Y)$.
+The negative sign on the coefficient of the mean local information $J(X,Y)$ shows that for fixed $H(Y)$ and $I(X,Y)$, higher values of $J(X,Y)$ (more sociospatial complexity) indicate cities that are harder to capture using coarse-grained models.
+	\begin{figure} 
+			\includegraphics[width=\linewidth]{figs/binary_loss_curves.pdf}
 
-	\begin{figure} 
+			\vspace{0.5cm}
+
 			\includegraphics[width=\linewidth]{figs/AUC_illustration.pdf}
 			\caption{
-				Evaluation of clusterings using the area under the curve (AUC) for $\mathcal{Y} = \{\text{Black }, \text{Non-Black}\}$, computed according to the formula \eqref{eq:AUC_formula}. 
-				The bounding rectangle for each city is shown in grey. 
-				The AUC is the proportion of the rectangle contained in the shaded region. 
-				The AUC for Philadelphia has been adjusted to account for the data set's two disconnected geographic components.
+				\emph{(Top)}: Hierarchical scaling of information for Black/Non-black segregation patterns in major U.S. cities.  
+				Each curve is a complexity profile for sociospatial structure in a city. 
+				Atlanta and Detroit's curves rise sharply and then level off, indicating that relatively few clusters suffice to express most of the sociospatial structure. 
+				Philadelphia's curve rises more slowly, indicating the need for more complex models. 
+				\emph{(Bottom)}:
+				Complexity quantification using the area under the curve (AUC) for $\mathcal{Y} = \{\text{Black }, \text{Non-Black}\}$, computed according to the formula \eqref{eq:AUC_formula}. 
+				The AUC is the proportion of the bounding rectangle contained within the shaded region. 
+				Lower values of the AUC correspond to more complex structure. 
 			} \label{fig:AUC}
 	\end{figure}
+
 	\begin{figure}

### 2016-11-10 template-preparing-research/PNAS-template-main.tex

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Template for preparing your research report submission to PNAS using Overleaf}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,c,1]{Author One}
\author[b,1,2]{Author Two} 
\author[a]{Author Three}

\affil[a]{Affiliation One}
\affil[b]{Affiliation Two}
\affil[c]{Affiliation Three}

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any conflict of interest here.}
\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format.  Instructions for use are provided below.

Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs.

\section*{Guide to using this template on Overleaf}

Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout. For more detailed information please see the \href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information for Authors}.

If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}. You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.

\subsection*{Author Affiliations}

Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. Authors with an ORCID ID may supply this information at submission.

\subsection*{Submitting Manuscripts}

All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

\subsection*{Format}

Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References. Other orders and headings are permitted.

\subsection*{Manuscript Length}

PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a Direct Submission research article is six pages and a PNAS PLUS research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.  When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for PNAS PLUS.

\subsection*{References}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}. All references, including for the SI, should be included in the main manuscript file. References appearing in both sections should not be duplicated.  SI references included in tables should be included with the main reference section.

\subsection*{Data Archival}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.

\subsection*{Language-Editing Services}
Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{frog}
\caption{Placeholder image of a frog with a long example caption to show justification setting.}
\label{fig:frog}
\end{figure}

\begin{SCfigure*}[\sidecaptionrelwidth][t]
\centering
\includegraphics[width=11.4cm,height=11.4cm]{frog}
\caption{This caption would be placed at the side of the figure, rather than below it.}\label{fig:side}
\end{SCfigure*}

\subsection*{Digital Figures}
\label{sec:figures}

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size. Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.  Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.

\subsection*{Single column equations}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, options are to use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures, or to use the \verb|\begin{widetext}...\end{widetext}| environment as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and footnotes, as mentioned in the \href{http://texdoc.net/pkg/cuted}{cuted package documentation}. In the case of problems with footnotes, it may be possible to correct the situation using commands \verb|\footnotemark| and \verb|\footnotetext|.

%% Do not use widetext if paper is in single column.
\begin{widetext}
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{align*}
\end{widetext}

\begin{table}%[tbhp]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsection*{Supporting Information (SI)}

The main text of the paper must stand on its own without the SI. Refer to SI in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc. Authors are limited to no more than 10 SI files, not including movie files. Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text. Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only.

\subsubsection*{SI Text}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).

\subsubsection*{SI Figures}

Provide a brief legend for each supporting figure after the supporting text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text. When saving TIFF files, use only LZW compression; do not use JPEG compression. Do not save figure numbers, legends, or author names as part of the image. Composite figures must be pre-assembled.

\subsubsection*{3D Figures}

Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.

\subsubsection*{SI Tables}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.

\subsubsection*{SI Datasets}

Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.

\subsubsection*{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.

\subsubsection*{Still images}

Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.

\subsubsection*{Appendices}

PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow{} % Display the acknowledgments section

% \pnasbreak splits and balances the columns before the references.
% Uncomment \pnasbreak to view the references in the PNAS-style
% If you see unexpected formatting errors, try commenting out \pnasbreak
% as it can run into problems with floats and footnotes on the final page.
%\pnasbreak

% Bibliography
\bibliography{pnas-sample}

\end{document}

### 2016-11-10 5_figs.Rmd
labels on non-binary loss curves

--- 2016-10-23 5_figs.Rmd
@@
 ```{r, fig.height=3}
-g <- checkerboard_illustration() 
-g + theme(text=element_text(size=16, family="Palatino"))
-ggsave('figs/checkerboard.pdf', width = 4, height = 1.5)
-embed_fonts("figs/checkerboard.pdf", outfile="figs/checkerboard.pdf")
-```
 
-## Methodological illustration
-```{r, warning = FALSE, message = FALSE}
-columns <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
-method_illustration <- function(city, vmax){
-	hexgrid <- readOGR(dsn = paste0('throughput/grids/',city), layer = 'geo', verbose = FALSE)
-	df <- hexgrid@data
-	df <- df %>% mutate(id = as.character(row_number()))
-	hexgrid <- fortify(hexgrid) %>%
-		left_join(df, by = c('id' = 'id'))
-	bbox <- c(min(hexgrid$long), min(hexgrid$lat), max(hexgrid$long), max(hexgrid$lat))
-
-	# squarify
-	max_range <- max(bbox[3] - bbox[1], bbox[4] - bbox[2])
-
-	long_mid <- (bbox[3] + bbox[1]) / 2
-	lat_mid  <- (bbox[2] + bbox[4]) / 2
-
-	bbox[1] <- long_mid - max_range / 2
-	bbox[3] <- long_mid + max_range / 2
-	bbox[2] <- lat_mid  - max_range / 2
-	bbox[4] <- lat_mid  + max_range / 2
-
-	map <- get_map(location = bbox, maptype = 'terrain-background')
-
-	p <- ggmap(map,darken = .5)
-	p <- p +
-		geom_polygon(data = hexgrid, aes(x = long, y = lat, group = group, fill = info), alpha = .8) +
-		scale_fill_continuous(low = 'white', high = 'steelblue', name= expression(J[Y](x)), limits = c(0, vmax), breaks = c(0,2,4,6)) +
-		theme(axis.ticks = element_blank(),
-			  axis.text.x = element_blank(),
-			  axis.text.y = element_blank()) +
+checkerboard_illustration <- function(){
+	expand.grid(x = 1:8, y = 1:8) %>%
+		mutate(`(a)` = 1,
+			   `(b)` = .5,
+			   `(c)` = (x > 4)*1,
+			   `(d)` = (x + y) %% 2) %>%
+		gather(key = model, value = p, -x, -y) %>%
+		ggplot(aes(x = x, y = y)) +
+		ggthemes::theme_map() +
+    geom_tile(aes(fill = p)) +
+		scale_fill_continuous(low = 'white', high = 'black ', limits=c(0,1)) +
+		facet_grid(~model, switch = 'x') +
 		xlab('') +
 		ylab('') +
-		theme(text=element_text(size=16, family="Palatino"),
-		    legend.justification=c(.8,0),
-			  legend.position=c(1,0),
-			  legend.background = element_rect(fill = alpha('blue', 0)),
-			  legend.title = element_text(colour="white"),
-			  legend.text = element_text(colour="white"),
-			  legend.key.size = unit(2, "mm"),
-			  legend.text = element_text(size = rel(.1))) +
-		ggtitle(paste0('Local Information in ', city, '\nJ(X,Y) = ', round(weighted.mean(df$info, df$density), 2)))
-	p
-}
-
-# first <- method_illustration('Detroit', vmax = 6)
-# second <- method_illustration('Philadelphia', vmax = 6)
-# 
-# gA <- ggplotGrob(first)
-# gB <- ggplotGrob(second)
-# g <- gridExtra::cbind.gtable(gA, gB, size = "max")
-# 
-# grid.draw(g)
-# 
-# ggsave('figs/method_illustration.pdf', g, width = 8, height = 5)
-
-```
-
-```{r additive_decomposability}
-
-expand.grid(x = 1:8, y = 1:8) %>% 
-	mutate(`(e)` = .5,
-		   `(e)` = ifelse(x %in% 2:4 & y %in% 2:7, 0, `(e)`),
-		   `(e)` = ifelse(x %in% 5:7 & y %in% 2:7, 1, `(e)`),
-		   `(f)` = .5) %>%
-	gather(key = model, value = p, -x, -y) %>%
-	ggplot(aes(x = x, y = y)) +
-		theme_minimal() +
+		guides(fill=FALSE) + 
 		theme(axis.ticks = element_blank(),
@@
 ```{r}
+
 cities <- c('Philadelphia', 'Atlanta')
@@
 
+```
+
+```{r}
+
 cluster_map_percent_black <- function(tracts_name, city, nclusters = 10){
 	tracts <- readOGR(dsn = paste0('data/cities/',tracts_name), layer = 'geo', verbose = FALSE)
+
 	tracts <- tracts[sum(tracts@data[,columns]) != 0,]

### 2016-11-11 figs.Rmd
multigroup maps

@@
 embed_fonts("figs/checkerboard.pdf", outfile="figs/checkerboard.pdf")
+```
+
+```{r, fig.height=3}
+
+expand.grid(x = 1:8, y = 1:8) %>%
+	mutate(`(a)` = .5,
+		     `(b)` = (x > 4)*1,
+		     `(c)` = ifelse(x == 7 & y %in% c(4,5), 0, `(b)`),
+		     `(c)` = ifelse(x == 2 & y %in% c(4,5), 1, `(c)`),
+		     `(d)` = (x + y) %% 2) %>%
+	gather(key = model, value = p, -x, -y) %>%
+	ggplot(aes(x = x, y = y)) +
+	ggthemes::theme_map() +
+  geom_tile(aes(fill = p)) +
+	scale_fill_continuous(low = 'white', high = 'black ', limits=c(0,1)) +
+	facet_grid(~model, switch = 'x') +
+	xlab('') +
+	ylab('') +
+	guides(fill=FALSE) + 
+	theme(axis.ticks = element_blank(),
+		  axis.text.x = element_blank(),
+		  axis.text.y = element_blank(),
+		  panel.margin = unit(1, "mm"),
+		  panel.background = element_rect(),
+		  panel.grid.major = element_line(size = 0),
+		  panel.grid.minor = element_line(size = 0),
+		  strip.background = element_rect(fill = "white", color = "white", size = 2),
+		  plot.margin=unit(c(1,1,0,0),"mm"),
+		  strip.text=element_text(vjust=-100))
+
+# g <- checkerboard_illustration() 
+# g
+# g + theme(text=element_text(size=10, family="Palatino"))
+out_dir = 'figs/checkerboard_b.pdf'
+ggsave(out_dir, width = 4, height = 1.3)
+embed_fonts(out_dir, outfile=out_dir)
 ```
@@
 
-summary <- compute_AUC(binary_loss_curves, adjust_thresh = 0.005) 
+summary <- compute_AUC(loss_curves, adjust_thresh = 0.005) 
 summary <- summary %>% left_join(cache, by = c('city' = 'city')) %>%
@@
                          nclust = c(50, 7, 8),
-                         I_XY = c(.34, .18, .37))
+                         I_XY = c(.36, .19, .44))
 
@@
   scale_x_continuous(trans = 'log10', limits = c(1,NA), expand = c(0, 0)) +
-  scale_y_continuous(limits = c(0,.5), expand = c(0, 0)) + 
+  scale_y_continuous(limits = c(0,.7), expand = c(0, 0)) + 
   scale_color_brewer(palette = 'Set1', name = NULL, guide = 'none') +
@@
 
+```{r}
+
+races <- c('Asian', 'Black', 'Hispanic', 'Other', 'White')
+city <- 'Atlanta'
+
+prominence_map <- function(city, races){
+  tracts <- readOGR(dsn = paste0('data/cities/',city), layer = 'geo', verbose = FALSE)
+
+  overall_dist <- tracts@data %>% 
+  	select_(.dots = races) %>% 
+  	colSums() %>% 
+  	simplex_normalize()
+
+  f <- function(row){
+  here <- tracts@data[row,races] %>% 
+  	simplex_normalize
+  index <- which.max(here * log(here / overall_dist))
+  return(races[index])
+  }
+
+  g <- function(row){
+  	here <- tracts@data[row,races] %>% 
+  		simplex_normalize
+  	return(max(here * log(here / overall_dist), na.rm = T))
+  }
+
+  tracts@data$prominent_race <- 1:nrow(tracts@data) %>% 
+  as.matrix() %>% 
+  apply(FUN = f, MARGIN = 1)
+
+  tracts@data$prominence <- 1:nrow(tracts@data) %>% 
+  	as.matrix() %>% 
+  	apply(FUN = g, MARGIN = 1)
+
+  df <- tracts@data
+
+  g <- tracts %>% fortify(region = 'GEOID') %>% 
+      tbl_df() %>% 
+      left_join(df, by = c('id' = 'GEOID')) %>% 
+      ggplot() + 
+      geom_polygon(aes(x = long, y = lat, fill = prominent_race, group = group, alpha = prominence))  +
+      ggthemes::theme_map() + 
+      scale_fill_brewer(palette = 'Set1') + 
+      scale_alpha_continuous(guide = 'none', limits = c(0,1)) + 
+      ggtitle(city) + 
+      theme(text=element_text(size=16),
+            plot.title = element_text(size=20, vjust = -5),
+            legend.title = element_blank(),
+            legend.position="bottom")
+  g
+}
+
+cluster_map <- function(map, city_name, races, nclusters){
+
+  tracts <- readOGR(dsn = paste0('data/cities/',city_name), layer = 'geo', verbose = FALSE)
+  tracts <- tracts[sum(tracts@data[,races]) != 0,]
+  clusters <- readRDS(paste0('throughput/clusterings/', city_name))
+  loss_curves <- readr::read_csv('throughput/loss_curves.csv') %>%
+  	group_by_('city') %>% 
+  	arrange(desc(I_XY)) %>% 
+  	mutate(I_XY = max(I_XY) - I_XY,
+  		   nclust = row_number(),
+  		   I_diff = I_XY - lag(I_XY))
+    for(ncluster in nclusters){
+
+  		weight <- loss_curves %>% filter(city == city_name, 
+  					   nclust == ncluster) %>% 
+  		ungroup() %>% 
+  		select(I_diff) %>% 
+  		as.numeric()
+
+  		tracts@data$cluster <- cutree(clusters,ncluster)
+  		overlay <- unionSpatialPolygons(tracts, IDs = tracts@data$cluster) %>% fortify()
+
+  		g <- g + geom_polygon(aes(x = long, y = lat, group = group),data = overlay,
+  		                      alpha = 0, fill = NA, size = 5*weight, color = 'black') 
+    }
+
+  max_weight <- loss_curves %>% filter(city == city_name) %>% 
+  		ungroup() %>% 
+  		summarise(max_diff = max(I_diff, na.rm = T)) %>% 
+  		as.numeric()
+
+  tracts@data$cluster <- 1
+  overlay <- unionSpatialPolygons(tracts, IDs = tracts@data$cluster) %>% fortify()
+  g <- g + geom_polygon(aes(x = long, y = lat, group = group),data = overlay,
+  		                      alpha = 0, fill = NA, size = 5*max_weight, color = 'white')
+
+  g
+}
+
+```
+
+```{r}
+
+for(city in info_summary$city){
+  print(city)
+
+  g <- prominence_map(city, races)
+
+  out_dir <- paste0('figs/batch_figs/prominence_map_', city, '.pdf')
+
+  ggsave(out_dir, width = 14, height = 14, units = 'cm')
+  embed_fonts(out_dir, outfile=out_dir)
+
+  nclusters <- 1:6
+  g <- g %>% cluster_map(city, races, nclusters)
+
+  out_dir <- paste0('figs/batch_figs/cluster_map_', city, '.pdf')
+  ggsave(out_dir, width = 14, height = 14, units = 'cm')
+  embed_fonts(out_dir, outfile=out_dir)
+}
+
+```
+
+```{r}
+
+city <- 'Detroit'
+print(city)
+nclusters <- 1:6
+g <- prominence_map(city, races)
+g <- g %>% cluster_map(city, races, nclusters)
+
+out_dir <- paste0('figs/prominence_map_Detroit.pdf')
+ggsave(out_dir, width = 14, height = 14, units = 'cm')
+embed_fonts(out_dir, outfile=out_dir)
+
+labels <- data_frame(label = c('A', 'B', 'C', 'D', 'E', 'F'),
+                     long   = c(-83.35, -83.1558, -83.164, -82.905, -83.055, -83.1258),
+                     lat  = c(42.37, 42.39, 42.28, 42.4, 42.4, 42.3214))
+
+g +
+  geom_point(aes(x = long, y = lat), data = labels, size = 6, pch=21, color = 'black', fill = 'white', alpha = .6) + 
+  geom_text(aes(x = long, y = lat, label = label), 
+                  data = labels, 
+                  size = 4, fontface = 'bold')
+
+out_dir <- paste0('figs/cluster_map_Detroit.pdf')
+
+ggsave(out_dir, width = 14, height = 14, units = 'cm')
+embed_fonts(out_dir, outfile=out_dir)
+
+```
+
+```{r}
+
+```
+
+```{r}
+pal <- RColorBrewer::brewer.pal(name = 'Set1', 5)
+
+make_cluster_groups('Detroit', 6) %>% 
+	mutate(race = factor(race, levels = c('Asian', 'Other', 'Hispanic', 'Black', 'White'))) %>% 
+  group_by(cluster) %>% 
+  mutate(alpha = p / sum(p),
+         total = sum(p)) %>% 
+	ggplot(aes(y = race, x = cluster, fill = race)) + 
+  scale_fill_manual(values = c(pal[1], pal[4], pal[3], pal[2], pal[5]), name = NULL, guide = 'none', labels = LETTERS[1:nclusters]) + 
+	geom_tile(color = 'black',  aes(alpha = alpha), size = 1) + 
+	geom_text(aes(
+		label = ifelse(p < 0.005, "", 100 * round(p,2))), size = 8) + 
+  scale_alpha_continuous(guide = 'none', limits = c(0, 1)) + 
+	theme_minimal() + 
+	theme(
+	  text=element_text(size=25),
+		  axis.ticks.x = element_blank(),
+		  axis.title = element_blank(),
+		  legend.position="bottom",
+		  legend.title = element_blank(),
+		  line = element_blank())
+
+ggsave('figs/example_clusters_detailed.pdf', width = 8, height = 6)
+embed_fonts("figs/example_clusters_detailed.pdf", outfile="figs/example_clusters_detailed.pdf")
+
+```
+
 # Session info

### 2016-11-11 figs.Rmd
updates to loss curve illustrations

@@
 
-cities <- c('Atlanta','Philadelphia', 'Detroit')
+cities <- c('Atlanta','Philadelphia', 'Detroit', 'Chicago')
 
@@
 	# geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
-	facet_wrap(~city) + 
+	facet_wrap(~city, ncol = 2) + 
   scale_fill_brewer(palette = 'Set1', name = NULL, guide = 'none') +
@@
 
-ggsave('figs/AUC_illustration.pdf', width = 10, height = 4)
+ggsave('figs/AUC_illustration.pdf', width = 7, height = 5.5)
 embed_fonts('figs/AUC_illustration.pdf', outfile = 'figs/AUC_illustration.pdf')
@@
 adjust_thresh = 0.01
-highlights  <- c('Philadelphia', 'Atlanta', 'Detroit')
+highlights  <- c('Philadelphia', 'Atlanta', 'Detroit', 'Chicago')
 curves <- binary_loss_curves %>% 
@@
 
-end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit'),
-                         nclust = c(50, 7, 8),
-                         I_XY = c(.34, .18, .37))
+end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit', 'Chicago'),
+                         nclust = c(50, 7, 8, 1000),
+                         I_XY = c(.35, .175, .38, .42))
 
@@
   geom_path(aes(group = city, color = city), data = highlighted) +
-  geom_text(aes(label = city), data = end_points) +
+  geom_label(aes(label = city, fill = city), data = end_points) +
   theme_light() +
@@
   scale_color_brewer(palette = 'Set1', name = NULL, guide = 'none') +
+  scale_fill_brewer(palette = 'Set1', name = NULL, guide = 'none') +
   xlab('Number of clusters (n)') + 
@@
 adjust_thresh = 0.01
-highlights  <- c('Philadelphia', 'Atlanta', 'Detroit')
+highlights  <- c('Philadelphia', 'Atlanta', 'Detroit', 'Chicago')
 curves <- loss_curves %>% 
@@
 
-end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit'),
-                         nclust = c(50, 7, 8),
-                         I_XY = c(.36, .19, .44))
+end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit', 'Chicago'),
+                         nclust = c(50, 7, 8, 1000),
+                         I_XY = c(.36, .19, .46, 0.66))
 
@@
 ```{r}
-
 city <- 'Detroit'
@@
 ```{r}
-
+for(city in c('Chicago', 'Atlanta', 'Detroit', 'Philadelphia')){
+  print(city)
+
+  g <- prominence_map(city, races)
+
+  out_dir <- paste0('figs/prominence_map_', city, '.pdf')
+
+  ggsave(out_dir, width = 14, height = 14, units = 'cm')
+  embed_fonts(out_dir, outfile=out_dir)
+
+  nclusters <- 1:6
+  g <- g %>% cluster_map(city, races, nclusters)
+
+  out_dir <- paste0('figs/cluster_map_', city, '.pdf')
+  ggsave(out_dir, width = 14, height = 14, units = 'cm')
+  embed_fonts(out_dir, outfile=out_dir)
+}
 ```

### 2016-11-11 figs.Rmd
loss curve updates

@@
 
+rects <- highlighted %>% group_by(city) %>% 
+	summarise(xmin = 1, ymin = 0, xmax = max(nclust), ymax = max(I_XY))
+
 # grey background
@@
 background %>% 
-  ggplot(aes(x = nclust, y = I_XY)) +
-  geom_path(aes(group = city), color = 'grey', alpha = 0.6) + 
-  geom_path(aes(group = city, color = city), data = highlighted) +
-  geom_label(aes(label = city, fill = city), data = end_points) +
+  ggplot() +
+  geom_path(aes(group = city, x = nclust, y = I_XY), color = 'grey', alpha = 0.6) + 
+  geom_path(aes(group = city, color = city, x = nclust, y = I_XY), data = highlighted) +
+  geom_label(aes(x = nclust, y = I_XY, label = city, fill = city), data = end_points) +
   theme_light() +
@@
   # geom_vline(xintercept=2, alpha = .2) + 
-  annotation_logticks(sides = 'b', size = .2)
+  annotation_logticks(sides = 'b', size = .2) +
+  geom_rect(data = rects, aes(xmin = xmin,
+								ymin = ymin,
+								xmax = xmax,
+								ymax = ymax,
+								group = city,
+								color = city),
+			  # color = 'grey',
+			  alpha = 0,
+			  size = 0.25)
 
@@
 
+rects <- highlighted %>% group_by(city) %>% 
+	summarise(xmin = 1, ymin = 0, xmax = max(nclust), ymax = max(I_XY))
+
 # grey background
 
+filled <- highlighted[,c('nclust','I_XY', 'city')] %>% 
+  select(city, I_XY, nclust) %>% 
+  ungroup()
+
+corner <- filled %>% group_by(city) %>%
+	summarise(I_XY = 0,
+	          nclust = max(nclust))
+
+filled <- rbind(corner, filled) %>%  # problem is right here, the columns don't add up
+	dplyr::arrange(city, nclust, desc(I_XY))
+
 background %>% 
-  ggplot(aes(x = nclust, y = I_XY)) +
-  geom_path(aes(group = city), color = 'grey', alpha = 0.6) + 
-  geom_path(aes(group = city, color = city), data = highlighted) +
-  geom_label(aes(label = city, fill = city), data = end_points) +
+  ggplot() +
+  geom_path(aes(group = city, x = nclust, y = I_XY), color = 'grey', alpha = 0.6) + 
+  geom_path(aes(group = city, color = city, x = nclust, y = I_XY), data = highlighted, size = 1) +
   theme_light() +
@@
   # geom_vline(xintercept=2, alpha = .2) + 
-  annotation_logticks(sides = 'b', size = .2)
+  annotation_logticks(sides = 'b', size = .2) +
+  # geom_polygon(aes(nclust, I_XY, fill = city, group = city), data = filled, alpha = .15) + 
+  geom_rect(data = rects, aes(xmin = xmin,
+								ymin = ymin,
+								xmax = xmax,
+								ymax = ymax,
+								group = city,
+								color = city),
+			  # color = 'grey',
+			  alpha = 0,
+			  size = 0.3) + 
+  geom_label(aes(x = nclust, y = I_XY, label = city, fill = city), data = end_points)
 

### 2016-11-11 figs.Rmd
more grids, multigroup
labelled detroit

@@
 
-cities <- c('Philadelphia', 'Atlanta')
-
-tab <- data.frame(city = character(), percent_black = numeric(), H_Y = numeric(), I_XY = numeric(), J_XY = numeric())
+cities <- c('Philadelphia', 'Atlanta', 'Chicago', 'Detroit')
+
+tab <- data.frame(city = character(), H_Y = numeric(), I_XY = numeric(), J_XY = numeric())
 
@@
 
-  columns = c('non_black', 'Black')
+  columns = races
   stuff <- compx::info_analysis(tracts, columns = columns, resolution = NULL, grid_tract = grid_tract, grid = grid)
@@
 
-  new_row <- data.frame(city = city, percent_black = percent_black, H_Y = stuff$H_Y, I_XY = stuff$I_XY, J_XY = stuff$J_XY)
+  new_row <- data.frame(city = city, H_Y = stuff$H_Y, I_XY = stuff$I_XY, J_XY = stuff$J_XY)
   tab <- rbind(tab, new_row)

### 2016-11-11 figs.Rmd
increased label size in detroit

@@
 g +
-  geom_point(aes(x = long, y = lat), data = labels, size = 6, pch=21, color = 'black', fill = 'white', alpha = .6) + 
+  geom_point(aes(x = long, y = lat), data = labels, size = 9, pch=21, color = 'black', fill = 'white', alpha = .6) + 
   geom_text(aes(x = long, y = lat, label = label), 
                   data = labels, 
-                  size = 4, fontface = 'bold')
-
-out_dir <- paste0('figs/cluster_map_Detroit.pdf')
+                  size = 7, fontface = 'bold')
+
+out_dir <- paste0('figs/cluster_map_Detroit_labeled.pdf')
 

### 2016-11-12 figs.Rmd
modified legends in cluster map

@@
 races <- c('Asian', 'Black', 'Hispanic', 'Other', 'White')
-city <- 'Atlanta'
 
@@
       ggthemes::theme_map() + 
-      scale_fill_brewer(palette = 'Set1') + 
       scale_alpha_continuous(guide = 'none', limits = c(0,1)) + 
       ggtitle(city) + 
-      theme(text=element_text(size=16),
-            plot.title = element_text(size=20, vjust = -5),
+      theme(plot.title = element_text(size=20, vjust = -5),
             legend.title = element_blank(),
-            legend.position="bottom")
+            legend.position = c(.7, .1))
   g
@@
   		g <- g + geom_polygon(aes(x = long, y = lat, group = group),data = overlay,
-  		                      alpha = 0, fill = NA, size = 5*weight, color = 'black') 
+  		                      alpha = 0, fill = NA, size = 8*weight, color = 'black') 
     }
@@
   g <- g + geom_polygon(aes(x = long, y = lat, group = group),data = overlay,
-  		                      alpha = 0, fill = NA, size = 5*max_weight, color = 'white')
+  		                      alpha = 0, fill = NA, size = 8*max_weight, color = 'white')
 
@@
   g <- prominence_map(city, races)
+  g <- g + scale_fill_brewer(palette = 'Set1')
 
@@
 g <- prominence_map(city, races)
+
+guide <- guide_legend(label.theme = element_text(size = 14, angle = 0))
+
+g <- g + scale_fill_brewer(palette = 'Set1', guide = guide)
+
 g <- g %>% cluster_map(city, races, nclusters)
@@
 
-ggsave(out_dir, width = 14, height = 14, units = 'cm')
+ggsave(out_dir, width = 14, height = 10, units = 'cm')
 embed_fonts(out_dir, outfile=out_dir)
+
+```
+```{r}
+
+g <- g + scale_fill_brewer(palette = 'Set1', guide = guide)
+g + theme(legend.position = c(.8, .1))
 
@@
 
+  if(city %in% c('Detroit', 'Philadelphia')){
+    guide <- guide_legend(label.theme = element_text(size = 14, angle = 0))
+
+  }else{
+    guide = FALSE
+  }
+
+  g <- g + scale_fill_brewer(palette = 'Set1', guide = guide)
+
   out_dir <- paste0('figs/prominence_map_', city, '.pdf')
 
-  ggsave(out_dir, width = 14, height = 14, units = 'cm')
+  ggsave(out_dir, width = 14, height = 10, units = 'cm')
   embed_fonts(out_dir, outfile=out_dir)
@@
   out_dir <- paste0('figs/cluster_map_', city, '.pdf')
-  ggsave(out_dir, width = 14, height = 14, units = 'cm')
+  ggsave(out_dir, width = 14, height = 10, units = 'cm')
   embed_fonts(out_dir, outfile=out_dir)

### 2016-11-12 figs.Rmd
resized figs

@@
 out_dir <- paste0('figs/prominence_map_Detroit.pdf')
-ggsave(out_dir, width = 14, height = 14, units = 'cm')
+ggsave(out_dir, width = 14, height = 10, units = 'cm')
 embed_fonts(out_dir, outfile=out_dir)
@@
   }
+  if(city == 'Detroit'){
+    height = 10
+  }else{
+    height = 14
+  }
 
@@
 
-  ggsave(out_dir, width = 14, height = 10, units = 'cm')
+  ggsave(out_dir, width = 14, height = height, units = 'cm')
   embed_fonts(out_dir, outfile=out_dir)
@@
   out_dir <- paste0('figs/cluster_map_', city, '.pdf')
-  ggsave(out_dir, width = 14, height = 10, units = 'cm')
+  ggsave(out_dir, width = 14, height = height, units = 'cm')
   embed_fonts(out_dir, outfile=out_dir)

### 2016-11-12 figs.Rmd
updated tables for multigroup

@@
 
-tab %>% 
-  mutate(percent_black = percent_black * 100) %>%   
+# tab %>% 
+#   mutate(percent_black = percent_black * 100) %>%   
+#   write_csv('figs/comparative_summary.csv')
+```
+```{r}
+cache %>% filter(city %in% c('Philadelphia', 'Atlanta')) %>% 
+  select(city, H_Y, I_XY, J_XY) %>%
+  mutate_at(.cols = c('H_Y', 'I_XY','J_XY'), 
+            function(x) round(x, 2)) %>%
   write_csv('figs/comparative_summary.csv')
-
 ```
@@
 ```{r}
-
 cluster_map_percent_black <- function(tracts_name, city, nclusters = 10){
@@
 tab %>% 
+  write_csv('figs/clustering_summary.csv')s
+```
+
+```{r}
+loss_curves %>% 
+  group_by(city) %>% 
+  filter(city %in% c('Atlanta', 'Philadelphia'),
+         nclust %in% c(2, 5, 10)) %>% 
+  select(-I_diff) %>% 
+  mutate(I_XY = round(I_XY,2)) %>% 
+  spread(key = city, value = I_XY, fill = NA) %>%
   write_csv('figs/clustering_summary.csv')
+  # slice(match(nclust, c('2', '5', '10', 'I(X,Y)')))
+
+loss_curves %>% 
+  group_by(city) %>% 
+  filter(city %in% c('Atlanta', 'Philadelphia')) %>% 
+  summarise(`I(X,Y)` = max(I_XY)) %>% 
+  spread(key = city, value = `I(X,Y)`, fill = NA) %>% 
+  write_csv('figs/max_information.csv')
 

### 2016-11-13 figs.Rmd
removed chicago from loss curves

@@
 
-columns = c('Asian', 'Black', 'Other', 'White', 'Hispanic')
+races   <- c('Asian', 'Black', 'Other', 'White', 'Hispanic')
+columns <- c('Asian', 'Black', 'Other', 'White', 'Hispanic')
 
@@
 
-cities <- c('Atlanta','Philadelphia', 'Detroit', 'Chicago')
+cities <- c('Atlanta','Philadelphia', 'Detroit')
 
@@
 	# geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
-	facet_wrap(~city, ncol = 2) + 
+	facet_wrap(~city) + 
   scale_fill_brewer(palette = 'Set1', name = NULL, guide = 'none') +
@@
 adjust_thresh = 0.01
-highlights  <- c('Philadelphia', 'Atlanta', 'Detroit', 'Chicago')
+highlights  <- c('Philadelphia', 'Atlanta', 'Detroit')
 curves <- binary_loss_curves %>% 
@@
 
-end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit', 'Chicago'),
-                         nclust = c(50, 7, 8, 1000),
-                         I_XY = c(.35, .175, .38, .42))
+end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit'),
+                         nclust = c(50, 7, 8),
+                         I_XY = c(.35, .175, .38))
 
@@
 adjust_thresh = 0.01
-highlights  <- c('Philadelphia', 'Atlanta', 'Detroit', 'Chicago')
+highlights  <- c('Philadelphia', 'Atlanta', 'Detroit')
 curves <- loss_curves %>% 
@@
 
-end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit', 'Chicago'),
-                         nclust = c(50, 7, 8, 1000),
-                         I_XY = c(.36, .19, .46, 0.66))
+end_points <- data.frame(city = c('Atlanta', 'Philadelphia', 'Detroit'),
+                         nclust = c(50, 7, 8),
+                         I_XY = c(.36, .19, .46))
 
@@
 tab %>% 
-  write_csv('figs/clustering_summary.csv')s
+  write_csv('figs/clustering_summary.csv')
 ```
@@
          nclust %in% c(2, 5, 10)) %>% 
-  select(-I_diff) %>% 
   mutate(I_XY = round(I_XY,2)) %>% 
@@
 
-loss_curves %>% 
-  group_by(city) %>% 
+cache %>% 
   filter(city %in% c('Atlanta', 'Philadelphia')) %>% 
-  summarise(`I(X,Y)` = max(I_XY)) %>% 
-  spread(key = city, value = `I(X,Y)`, fill = NA) %>% 
+  select(city, I_XY) %>% 
+  spread(key = city, value = I_XY, fill = NA) %>% 
   write_csv('figs/max_information.csv')

### 2016-12-15 figs.Rmd
figures redone after bugfix in clustering

@@
 
-cluster_map('Detroit', 'Detroit', nclusters = 6)
+cluster_map('Detroit', 'Detroit', nclusters = 5)
 
@@
 ```{r}
-make_cluster_groups('Detroit', 6) %>% 
+make_cluster_groups('Detroit', 5) %>% 
 	mutate(race = factor(race, levels = c('Other', 'Asian', 'Hispanic', 'Black', 'White'))) %>% 
@@
 print(city)
-nclusters <- 1:6
+nclusters <- 1:5
 g <- prominence_map(city, races)
@@
 
-labels <- data_frame(label = c('A', 'B', 'C', 'D', 'E', 'F'),
-                     long   = c(-83.35, -83.1558, -83.164, -82.905, -83.055, -83.1258),
-                     lat  = c(42.37, 42.39, 42.28, 42.4, 42.4, 42.3214))
+labels <- data_frame(label = c('A', 'B',  'E', 'C', 'D'),
+                     long   = c(-83.35, -83.1568, -83.132, -82.905, -83.055),
+                     lat  = c(42.37, 42.39, 42.32, 42.4, 42.4))
 
@@
 
-  nclusters <- 1:6
+  nclusters <- 1:5
   g <- g %>% cluster_map(city, races, nclusters)
@@
 
-make_cluster_groups('Detroit', 6) %>% 
+make_cluster_groups('Detroit', 5) %>% 
 	mutate(race = factor(race, levels = c('Asian', 'Other', 'Hispanic', 'Black', 'White'))) %>% 

### 2016-12-15 2_make_cities.R

@@ -15,7 +15,7 @@ cities <- cities[!is.element(cities, list.dirs('cities', full.names = FALSE))]
 
 states <- city_df$state %>% 
 	str_split('--') %>% 
-	do.call(c,.) %>% 
+	# do.call(c,.) %>% 
 	unique()
 
 shapes <- list()
@@ -49,7 +49,14 @@ for(city in cities){
 																	                                 "data.frame")))))
 	}
 	city_polys <- do.call(rbind, city_components)
+	city_polys@data <- city_polys@data %>%
+	  mutate(total = Asian + Black + Hispanic + Other + White)
 	city_polys <- city_polys[city_polys@data$total != 0,]
+	if(city == 'Philadelphia'){
+	  exclude <- c('421010098023','421010098022', '421010098021', '421010098011',  '421010098012')
+	  city_polys <- city_polys[!(city_polys@data$GEOID %in% exclude),]   
+	}
+	
 	tryCatch(writeOGR(city_polys, paste0("data/cities/", city),
 	                  'geo', 
 	                  driver = 'ESRI Shapefile', 

### 2016-12-15 3_local_info.R

@@ -4,13 +4,11 @@ library(compx, quietly = TRUE)
 library(plyr, quietly = TRUE)
 library(dplyr, quietly = TRUE)
 library(readr, quietly = TRUE)
-library(ggmap, quietly = TRUE)
 library(tidyr, quietly = TRUE)
 library(magrittr, quietly = TRUE)
-library(ggrepel, quietly = TRUE)
 library(grid, quietly = TRUE)
 library(rgdal, quietly = TRUE)
-library(data.table)
+library(data.table, quietly = TRUE)
 
 races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
 
@@ -45,7 +43,7 @@ if(file.exists('throughput/info_cache.csv')){
 		I_XY = numeric(),
 		J_XY = numeric())
 }
-city = 'Boston'
+
 for(city in cities){
 	tracts <- readOGR(dsn = paste0('data/cities/',city),
 	                  layer = 'geo',
@@ -56,7 +54,7 @@ for(city in cities){
 	if(file.exists(grid_tract_path) & dir.exists(grid_dir)){
 	  grid_tract <- read_csv(grid_tract_path)
 	  grid <- readOGR(dsn = paste0('throughput/grids/',city),
-	                  layer = 'geo',
+	                  layer = 'grid',
 	                  verbose = FALSE)
 
 	  print(city)

### 2016-12-15 4_cluster.R

@@ -18,9 +18,11 @@ cities <- summary$city
 
 if(file.exists('throughput/loss_curves.csv')){
 	cache <- read_csv('throughput/loss_curves.csv')
+	cache_binary <- read_csv('throughput/loss_curves_binary.csv')
 	cities <- cities[!is.element(cities, cache$city)]
 } else {
 	cache <- data.frame('I_XY' = numeric(), 'nclust' = numeric(), 'city' = character())
+	cache_binary <- data.frame('I_XY' = numeric(), 'nclust' = numeric(), 'city' = character())
 }
 
 if(!dir.exists('throughput/clusterings')){
@@ -31,12 +33,22 @@ for(city in cities){
 	print(city)
 	spdf <- readOGR(dsn = paste0('data/cities/', city), layer = 'geo',verbose = FALSE)
 	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
-	df <- spdf@data[,columns]
+	df   <- spdf@data[,columns]
 	row.names(df) <- 1:nrow(df)
-	constraint <- gRelate(spdf, byid = TRUE, pattern = '****1****')
-	a <- info_clust(df, constraint)
+	constraint    <- gRelate(spdf, byid = TRUE, pattern = '****1****')
+	a     <- info_clust(df, constraint)
 	saveRDS(a, file = paste0('throughput/clusterings/',city))
 	added <- data.frame('I_XY' = c(0,a$height), 'nclust' = 1:nrow(spdf), 'city' = city)
 	cache <- rbind(cache, added)
 	write_csv(cache,'throughput/loss_curves.csv')
+	
+	df <- df %>% 
+	  mutate(non_black = Asian + Hispanic + Other + White ) %>% 
+	  select(Black, non_black)
+	
+	b <- info_clust(df, constraint)
+	saveRDS(b, file = paste0('throughput/clusterings/',city, '_binary'))
+	added        <- data.frame('I_XY' = c(0,b$height), 'nclust' = 1:nrow(spdf), 'city' = city)
+	cache_binary <- rbind(cache_binary, added)
+	write_csv(cache_binary,'throughput/loss_curves_binary.csv')
 }

### 2016-12-15 figs.Rmd
fix to clustering summary =

@@
 
+# check in here: are these binary clustering summaries? If so, we need to fix and get the full 
+# clustering summaries instead.
+
 tab <- data.frame(city = character(), total_info = numeric(), info_4 = numeric(), info_20 = numeric())
@@
   mutate(I_XY = round(I_XY,2)) %>% 
+  select(city, I_XY, nclust) %>% 
   spread(key = city, value = I_XY, fill = NA) %>%

### 2016-12-15 figs.Rmd
added rounding to max info

@@
   select(city, I_XY) %>% 
+  mutate(I_XY = round(I_XY,2)) %>% 
   spread(key = city, value = I_XY, fill = NA) %>% 

### 2017-02-04 6_figs.Rmd
basic analysis, decent results

---
title: "Figures for The *Spatial Structure of Segregation*"
author: "Phil Chodrow"
date: "1/22/2017"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
---

# Introduction

The purpose of this document is to generate and lightly explain the figures used in "The Spatial Structure of Segregation." For more complete documentation, please see the technical report.

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4, fig.width = 6)
```

```{r}
library(tidyverse)
library(rgeos)
library(maptools)
library(scales)
library(ggthemes)
library(ggrepel)
library(igraph)
library(compx)
library(RColorBrewer)
library(gplots)
library(viridis)
library(ggmap)
```

```{r}
if(!dir.exists('figs')){
    dir.create('figs')
}
```

# Metric Structure

First we read in the data...
```{r}

city     <- 'Detroit'
input    <- readRDS(paste0('throughput/metric/', city, '.RDS'))
f_tracts <- input$tracts %>%
    fortify(region = 'GEOID')

metric_df <- input$metric_df

f_tracts <- f_tracts %>% 
    left_join(metric_df, by = c('id' = 'geoid')) 
```

And now we plot.

```{r}
# the tr/4 is from our existence theorem
f_tracts %>% 
    ggplot() + 
    aes(x = long, y = lat, fill = tr/4, group = group) + 
    geom_polygon() + 
    scale_fill_viridis(option = 'magma' , 
                                trans = 'log10',
                       limits = c(NA, 12e-1),
                                oob = squish) +
    theme_map() + 
    theme(legend.justification = c(0,0), 
          legend.position      = c(.8,.1),
          panel.background = element_rect(fill = 'grey80', color = NA),
          legend.background = element_rect(fill = NA)) + 
    guides(fill = guide_colorbar(ticks = F, title = expression(tr~italic(g[x]))))

ggsave('figs/local_density.pdf')

```

## Comparative Information Density

```{r}
scatterplot_theme <- function(){
     layout <- theme_bw() %+replace%
         theme(axis.ticks.y = element_line(size = 0, color = 'black'),
          axis.ticks.x = element_line(size = .2, color = 'black'),
          axis.ticks.length=unit(-0.25, "cm"), 
          axis.text.x = element_text(size = 11, margin=unit(c(0.5,0.5,0.5,0.5), "cm")), 
          axis.text.y = element_text(size = 11, margin=unit(c(0.5,0.5,0.1,0.5), "cm")),
          axis.title.y = element_text(margin=unit(c(0.1,-0.5,0.1,0.1), "cm"), angle = 90),
          axis.title.x = element_text(margin=unit(c(0,0,-0.1,-1), "cm")))

    list(layout)
}
```

```{r}

summary <- read_csv('throughput/metric.csv')

# summary <- summary %>% 
#     mutate(p_vol = p_vol / max(p_vol),
#            p_trace = p_trace / max(p_trace))

highlight <- summary %>% 
    filter(I_XY < 0 | I_XY > .35)

 summary %>% 
    ggplot() + 
    aes(x = p_vol, y = p_trace) + 
    scale_y_continuous(trans = 'log10') +
    scale_x_continuous(trans = 'log10') +
    geom_point(color = 'grey') + 
    geom_point(color = 'firebrick', data = highlight) + 
    geom_text_repel(aes(label = city), data = highlight) +
    scatterplot_theme() + 
    xlab(expression(italic(E[X]) ~ sqrt(~det~g[X]))) + 
    ylab(expression(italic(E[X]) ~ (tr ~~  g[X]))) + 
    annotation_logticks(scaled = TRUE, size = .2, sides = 'lb')

 ggsave('figs/average_geometry.pdf')

```

```{r}
summary <- read_csv('throughput/metric.csv')

highlight <- summary %>% 
    filter(I_XY < 0 | I_XY > .35)

summary %>% 
    ggplot() + 
    aes(x = I_XY, y = p_trace) + 
    scale_y_continuous(trans = 'log10') + 
    geom_point(color = 'grey') + 
    geom_point(color = 'firebrick', data = highlight) + 
    geom_text_repel(aes(label = city), data = highlight) +
    scatterplot_theme() + 
    ylab(expression(italic(E[X]) ~ (tr ~~  g[X]))) + 
    xlab(expression(italic(I(X,Y)))) + 
    annotation_logticks(scaled = TRUE, size = .2, sides = 'l') +

ggsave('figs/density_mutual.pdf')
```

# Network Structure

## Spectra
```{r}
spectra <- read_csv('throughput/spectra.csv')

cities <- c('Philadelphia', 'Detroit', 'Atlanta', 'Chicago', 'New York City')
highlight <- spectra %>% 
    filter(city %in% cities)
```

```{r}

spectra_theme <- function(){

    layout <- theme_bw() %+replace%
    theme(axis.ticks.y = element_line(size = .2, color = 'black'),
          axis.ticks.x = element_line(size = .2, color = 'black'),
          axis.ticks.length=unit(-0.25, "cm"),
          axis.text.x = element_text(size = 11, margin=unit(c(0.5,0.5,0.5,0.5), "cm")), 
          axis.text.y = element_text(size = 11, margin=unit(c(0.5,0.5,0.1,0.0), "cm")),
          axis.title.y = element_text(margin=unit(c(0.1,-0.1,0.5,0.5), "cm"), angle = 90)
          )

    list(layout, 
         scale_color_brewer(palette = 'Set1'),
         guides(color = guide_legend(title = element_blank())))
    # scale_color_brewer(palette = 'Set1') %+replace% 
    # guides(color = guide_legend(title = element_blank())) 
}

```

```{r}

spectra %>% 
    ggplot() + 
    aes(x = n, y = 1-eig, group = city) + 
    geom_line(color = 'grey', alpha = 1, size = .3) + 
    scale_y_continuous(trans = 'log10', limit = c(8e-2, NA), expand=c(0,0)) + 
    scale_x_continuous(limit = c(0, 100), expand=c(0,0)) + 
    geom_line(aes(color = city),  data = highlight, size = 1) + 
    geom_point(data = highlight, size = .1, color = 'black') + 
    ylab(expression(1-lambda[italic(n)])) + 
    xlab(expression(italic(n))) + 
    annotation_logticks(scaled = T, sides = 'l', size = .2) +
    spectra_theme() + 
    theme(axis.ticks.y = element_line(size = 0, color = 'black'))

ggsave('figs/spectra.pdf')

```

## Differenced Spectra
```{r}

spectra <- spectra %>% 
    mutate(diff = lead(eig) / eig)

highlight <- spectra %>%
    filter(city %in% cities)

spectra %>% 
    ggplot() + 
    aes(x = n, y = diff, group = city) + 
    geom_line(alpha = .5, color = 'grey') +
    geom_line(aes(color = city), data = highlight, size = .7) + 
    geom_point(data = highlight, size = .5, color = 'black') +
    scale_x_continuous(limits = c(2, 25)) +
    scale_y_continuous(limits = c(1, 3.5)) +
    ylab(expression(lambda[italic(n+1)] /lambda[italic(n)])) +
    xlab(expression(italic(n))) + 
    spectra_theme()

ggsave('figs/spectra_differenced.pdf')

```

What happens when we incorporate area?

```{r, echo = FALSE, eval = FALSE}

area_df <- summary %>% 
    select(city, total_area)

spectra <- spectra %>% 
    # left_join(area_df) %>% 
    group_by(city) %>%
    mutate(scale = n/max(n)) %>% 
    ungroup()

spectra <- spectra %>% 
    mutate(diff = (1-eig) / (1-lead(eig)))

highlight <- spectra %>%
    filter(city %in% cities, scale < .1)

spectra %>% 
    filter(scale <= .1) %>%
    ggplot() + 
    aes(x = scale, y = diff, group = city) + 
    # geom_point(size = 2) + 
    geom_line(color = 'grey') + 
    geom_line(aes(color = city),  data = highlight, size = 1) +
    # geom_point(data = highlight, size = 1, color = 'black') + 
    # scale_x_continuous(limits = c(1, 30)) +
    scale_y_continuous(limits = c(1, 3.7)) +
    theme_bw() + 
    # ylab(expression(1-lambda[italic(n)])) + 
    xlab(expression(italic(n))) + 
    # annotation_logticks(scaled = T, sides = 'l', size = .2) + 
    theme(axis.ticks.y = element_line(size = .2, color = 'black'),
          axis.ticks.x = element_line(size = .2, color = 'black'),
          legend.position = 'bottom') + 
    ggtitle('Spectra of selected cities') + 
    scale_color_brewer(palette = 'Set1') + 
    guides(color = guide_legend(title = element_blank()))

```

What about when we normalize n?

```{r}

normed_spectra <- spectra %>%
    group_by(city) %>% 
    mutate(n = n/max(n))

normed_highlight <- normed_spectra %>% 
    filter(city %in% cities)

normed_spectra %>% 
    ggplot() + 
    aes(x = n, y = 1-eig, group = city) + 
    geom_line(color = 'grey', alpha = 1, size = .3) + 
    scale_y_continuous(trans = 'log10', limit = c(1e-1, NA), expand=c(0,0)) + 
    scale_x_continuous(limit = c(0.001, .5), expand=c(0,0), trans = 'log10') +
    geom_line(aes(color = city),  data = normed_highlight, size = 1) + 
    geom_point(data = normed_highlight, size = .3, color = 'black') + 
    theme_bw() + 
    ylab(expression(1-lambda[italic(n)])) + 
    xlab(expression(italic(n))) + 
    annotation_logticks(scaled = T, sides = 'l', size = .2) + 
    theme(axis.ticks.y = element_line(size = .2, color = 'black'),
          axis.ticks.x = element_line(size = .2, color = 'black'),
          legend.position = 'bottom') + 
    ggtitle('Spectra of selected cities') + 
    scale_color_brewer(palette = 'Set1') + 
    guides(color = guide_legend(title = element_blank()))

ggsave('figs/normed_spectra.pdf')

```

```{r}

spectral_properties <- spectra %>% 
    group_by(city) %>% 
    filter(n >= 10 | eig  > 1e-10) %>% 
    filter(eig < 1) %>% 
    summarise(aff_trace =   sum( 1 - eig),
              aff_det   = - sum(log(1-eig))) %>%
    arrange(desc(aff_trace))

spectral_properties %>% 
    left_join(summary, by = c('city' = 'city')) %>% 
    ggplot() + 
    aes(x = aff_trace, y = total_vol) + 
    geom_point() +
    geom_text_repel(aes(label = city)) + 
    scale_y_continuous(trans = 'log10') + 
    scale_x_continuous(trans = 'log10') + 
    geom_smooth() + 
    ggtitle('Correlation between the trace of the affinity matrix and information volume.')

```

## Clustering

First we'll do ths clustering. 
```{r, fig.height = 6, fig.width = 10}
g <- readRDS(paste0('throughput/graphs/',city,'.RDS'))
A <- affinity_matrix(g, sigma = 1)
sigma <- 1
clust <- hclust(as.dist(-A),  method = 'complete')
pal   <- brewer.pal(9, name = "Set1") %>%
	colorRampPalette()

ncluster <- 7
g <- g %>% set_vertex_attr(name = 'cluster', index = colnames(A), value = clust %>% cutree(ncluster))

```

## Geodesics and Space Distortion

```{r}
k <- 600
center <- V(g)[[k]]$name
center_coords <- data_frame(x = V(g)[[k]]$x, y = V(g)[[k]]$y)

dists <- g %>% igraph::distances(weights = E(g)$dist)

dist_lookup <- dists[,center] %>% 
    as.data.frame() %>% 
    rownames_to_column(var = 'geoid')

names(dist_lookup) <- c('geoid', 'dist')

f_tracts %>% 
    left_join(dist_lookup, by = c('id' = 'geoid')) %>% 
    ggplot() + 
    aes(x = long, y = lat, group = group, fill = dist) %>% 
    geom_polygon() + 
    scale_fill_viridis(option = 'magma' , 
                                oob = squish) +
    theme_map() + 
    theme(legend.justification = c(0,0), 
          legend.position      = c(.8,.1),
          panel.background = element_rect(fill = 'grey80', color = NA),
          legend.background = element_rect(fill = NA)) + 
    guides(fill = guide_colorbar(ticks = F, title = expression(d[gamma]))) + 
    geom_point(aes(x,y), data = center_coords, color = 'white')

ggsave('figs/geo_distance.pdf')
```

## Bare network structure

Visualization of network structure in Detroit. The weight of each line segment is proportional to the affinity between the tracts it connects. 
```{r}
edges <- g %>% as_long_data_frame() %>% tbl_df()
nodes <- data_frame(x = V(g)$x, y = V(g)$y, cluster = as.character(V(g)$cluster))

ggplot() + 
    geom_polygon(aes(x = long, y = lat, group = group), fill = 'grey10', alpha = .7, data = f_tracts) +
    geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, size = exp(-7 * dist^2), alpha = exp(-7 * dist^2)), color = 'white', data = edges) +
    theme_map() +
    scale_size_continuous(range = c(0,.5)) + 
    guides(alpha = 'none', color = 'none', size = 'none') +
    theme(panel.background = element_rect(fill = 'grey80', color = NA))

ggsave('figs/network_structure.pdf')

```

Now we'll have the color of the edge proportional to its geodesic distance.

```{r}
edges <- g %>% as_long_data_frame() %>% tbl_df()
nodes <- data_frame(x = V(g)$x, y = V(g)$y, cluster = as.character(V(g)$cluster))

ggplot() + 
    geom_polygon(aes(x = long, y = lat, group = group), fill = 'grey80', alpha = .6, data = f_tracts) +
    geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, color = dist), data = edges) +
    theme_map() +
    scale_color_viridis(option = 'magma', trans = 'log10') + 
    guides(alpha = 'none', color = 'none', size = 'none') +
    theme(panel.background = element_rect(fill = 'grey80', color = NA))

ggsave('figs/network_structure_colored.pdf')

```

This one is a little bit busier, but shows fairly nicely how the network is most strongly separated at points of high curvature.

```{r}
edges <- g %>% as_long_data_frame() %>% tbl_df()
nodes <- data_frame(x = V(g)$x, y = V(g)$y, cluster = as.character(V(g)$cluster))

ggplot() + 
    geom_polygon(aes(x = long, y = lat, fill = vol/area, group = group), data = f_tracts) + 
    scale_fill_viridis(option = 'magma' , 
                                trans = 'log10', 
                                oob = squish) +
    geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, size = exp(-10 * dist)), color = 'white', data = edges) +
    theme_map() +
    scale_size_continuous(range = c(0,1)) + 
    guides(alpha = 'none', color = 'none', size = 'none') +
    theme(panel.background = element_rect(fill = 'grey80', color = NA),
          legend.justification = c(0,0), 
          legend.position      = c(.8,.1),
          legend.background = element_rect(fill = NA)) + 
    guides(fill = guide_colorbar(ticks = F, title = expression(rho)))

ggsave('figs/network_structure_curvature.pdf')
```

Now let's do it with spectral clustering.

```{r}

A <- affinity_matrix(g, sigma = 2)
L <- generalized_laplacian_matrix(A)
evL <- eigen(L, symmetric=TRUE)

# let's pick a number of clusters by checking out the spectrum. 
test_df <- data_frame(n = 1:nrow(A), eig = rev(evL$values)) %>% 
	mutate(eig = eig - min(eig),
		   eig = eig / max(eig)) %>% 
	mutate(val = lead(eig) / eig) %>% 
	filter(n < 50)

plot(test_df$val, log = 'y')

```

```{r}

make_clusters <- function(k, nreps = 1000){
    Z <- evL$vectors[,(ncol(evL$vectors)-k+1):(ncol(evL$vectors))]

    models <- data_frame(n = 1:nreps) %>% 
    	dplyr::mutate(model = map(n, ~ kmeans(Z, centers = k))) %>% 
    	mutate(perf = map_dbl(model, ~.$tot.withinss))

    model <- models %>% 
    	filter(perf == min(perf))
    km <- model$model[[1]]
    return(km$cluster)
}

get_node_df <- function(k){
    h <- g %>% set.vertex.attribute('cluster', value = make_clusters(k)) 
    data_frame(geoid = V(h)$name, 
               x = V(h)$x, 
               y = V(h)$y, 
               cluster = V(h)$cluster, 
               k = k)
}

```

```{r}

ks <- c(2, 6, 10, 15)

clusters <- ks %>% 
    map(get_node_df) %>% 
    reduce(rbind)

pal <- brewer_pal(palette = 'Set1')
pal <- colorRampPalette(pal(9))
pal <- pal(max(ks))

ggplot() + 
	geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, size = exp(-5 * dist^2)), color = 'black', alpha = .5, data = edges) +
	theme_map() +
	scale_size_continuous(range = c(0,1)) + 
	guides(alpha = 'none', color = 'none', size = 'none') +
	theme(panel.background = element_rect(fill = 'grey80', color = NA)) + 
	geom_point(aes(x, y, color = as.character(cluster)), data = clusters) + 
    scale_color_manual(values = pal) + 
    facet_wrap(~k, nrow = 3)

ggsave('figs/spectral_clustered.pdf')

```

```{r}

label_generator <- function(variable, value){
    return(paste0('k = ', value))
}

f_tracts %>% 
    left_join(clusters, by = c('id' = 'geoid')) %>% 
    ggplot() + 
    aes(x = long, y = lat, fill = as.character(cluster), group = id) + 
    facet_wrap(~k, nrow = 2, labeller = label_generator) + 
    geom_polygon() + 
    theme_map() +
    theme(panel.background = element_rect(fill = 'grey80', color = NA)) + 
    scale_fill_manual(values = pal) +
    guides(fill = FALSE)

ggsave('figs/clustered_maps.pdf')

```

## Matrix Structure

Now let's visualize the affinity matrix! We can use hierarchical clustering for this task.

```{r}
sigma = 1
affinity_matrix_viz <- function(city){
    g <- readRDS(paste0('throughput/graphs/',city,'.RDS'))
    A <- affinity_matrix(g, sigma = sigma)

    clust <- hclust(as.dist(-A),  method = 'complete')

    the_order <- data_frame(geoid = clust$labels[clust$order], order = 1:length(clust$labels))

    A %>% 
        as.data.frame() %>% 
        rownames_to_column('geoid_1') %>% 
        tbl_df() %>% 
        gather(geoid_2, a, -geoid_1) %>% 
        left_join(the_order, by = c('geoid_1' = 'geoid')) %>% 
        dplyr::rename(x = order) %>%
        left_join(the_order, by = c('geoid_2' = 'geoid')) %>% 
        dplyr::rename(y = order) %>% 
        ggplot() + 
        aes(x = x, y = y, fill = a) + 
        geom_raster() + 
        scale_fill_viridis(option = 'viridis') + 
        # theme_map() +
        theme(legend.position = 'right',
              axis.line=element_blank(),
              axis.text.x=element_blank(),
              axis.text.y=element_blank(),
              axis.ticks=element_blank(),
              axis.title.x=element_blank(),
              axis.title.y=element_blank()) + 
        scale_x_continuous(expand = c(0,0)) + 
        scale_y_continuous(expand = c(0,0)) + 
        guides(fill = guide_colorbar(ticks = F, title = expression(italic(A))))
}

```

```{r, fig.width = 5}
affinity_matrix_viz('Atlanta')
ggsave('figs/Atlanta_affinity.pdf')
```

```{r, fig.width = 5}
affinity_matrix_viz('Detroit')
ggsave('figs/Detroit_affinity.pdf')
```

```{r}

prominence_map <- function(tracts){
    tracts <- input$tracts

    demographics <- read_csv('data/demographics.csv') %>% 
        dplyr::mutate(tract = as.character(GEOID),
               tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
        select(-GEOID)

    data <- demographics %>% 
        filter(tract %in% tracts$GEOID)

    marginal <- data %>% 
        group_by(group) %>% 
        summarise(n = sum(n)) %>%
        mutate(n = n / sum(n))

    f <- function(df){
        df$group[which.max(df$n *  log(df$n / marginal$n))][1]
    }

    g <- function(df){
        max(df$n *  log(df$n / marginal$n), na.rm = T)
    }

    prominence_df <- data %>% 
        group_by(tract) %>% 
        mutate(n = n / sum(n)) %>% 
        nest(-tract, .key = df) %>% 
        mutate(prominent_group = map_chr(df, f),
               prominence = map_dbl(df, g))

    g <- tracts %>% fortify(region = 'GEOID') %>% 
          tbl_df() %>% 
          left_join(prominence_df, by = c('id' = 'tract')) %>% 
          ggplot() + 
          geom_polygon(aes(x = long, 
                           y = lat, 
                           fill = prominent_group, 
                           group = group, 
                           alpha = prominence))  +
          ggthemes::theme_map() + 
          scale_alpha_continuous(guide = 'none', limits = c(0,1)) +
          theme(plot.title = element_text(size=20, vjust = -5),
                legend.title = element_blank(),
                legend.position = c(.7, .1))
      g
}
```

Generate Detroit, with legend

```{r}
city  <- 'Detroit'
input <- readRDS(paste0('throughput/metric/', city, '.RDS'))

prominence_map(input$tracts) + 
    scale_fill_brewer(palette = 'Set1')

ggsave('figs/detroit_prominence.pdf')
```

Generate Atlanta, no legend
```{r}
city  <- 'Atlanta'
input <- readRDS(paste0('throughput/metric/', city, '.RDS'))

prominence_map(input$tracts) + 
    scale_fill_brewer(palette = 'Set1') + 
    guides(fill = FALSE)

ggsave('figs/atlanta_prominence.pdf')
```

```{r}
city  <- 'Newark'
input <- readRDS(paste0('throughput/metric/', city, '.RDS'))

prominence_map(input$tracts) + 
    scale_fill_brewer(palette = 'Set1') + 
    guides(fill = FALSE)

ggsave('figs/newark_prominence.pdf')

```

# Session Info
```{r}
sessionInfo()
```

### 2017-02-04 2_get_states.R

@@ -1,0 +1,124 @@
+# library(sp, quietly = TRUE)
+library(acs, quietly = TRUE)
+library(tidyverse, quietly = TRUE)
+library(rgdal, quietly = TRUE)
+
+level <- 'blockgroup'
+
+# -------------------
+
+clean_race <- function(race){
+    
+    
+    # transform into data frame
+    race <- cbind(data.frame(race@geography),
+                  data.frame(race@estimate)) %>%
+        tbl_df() 
+    
+    # aggregate into 5 categories
+    race_lookup <- read_csv('assumptions/race_lookup.csv')
+    
+    au_vars <- c('state', 'county', 'tract', 'blockgroup') %>% 
+        intersect(names(race))
+    
+    gather_cols <- names(race) %>% 
+        setdiff(au_vars) %>% 
+        setdiff('NAME')
+    
+    race <- race %>% 
+        gather_(key_col = 'original_race', value_col = 'n', gather_cols = gather_cols) %>% 
+        left_join(race_lookup, by = c('original_race' = 'original')) %>% 
+        dplyr::group_by_(.dots = au_vars %>% append(c('aggregated'))) %>% 
+        dplyr::summarise(n = sum(n)) %>% 
+        ungroup() %>%  
+        filter(!is.na(aggregated)) %>% 
+        spread(key = aggregated, value = n, fill = 0)
+    
+    # clean up the areal unit designations to give a single GEOID
+    
+    lengths <- list(state      = 2, 
+                    county     = 3, 
+                    tract      = 6, 
+                    blockgroup = 1)
+    
+    lengths <- lengths[au_vars]
+    
+    for(colname in names(lengths)){
+        race[[colname]] <- str_pad(race[[colname]], 
+                                   lengths[[colname]], 
+                                   'left', 
+                                   pad = 0)
+    }
+    
+    race <- race %>% 
+        tidyr::unite_('GEOID', au_vars, sep = '') %>% 
+        gather(key = group, value = n, -GEOID)
+    race
+}
+
+
+getter <- function(state){
+    print(state)
+    counties <- fips.county %>% 
+        filter(State == state) %>% 
+        select(County.ANSI) %>% 
+        unlist()
+
+    if(level == 'blockgroup'){
+        tracts <- tigris::block_groups(state = state, county = counties, cb=TRUE)
+        geo    <- geo.make(state = state, 
+                           county = counties, 
+                           tract = '*', 
+                           block.group = '*')
+    }else{
[…]

### 2017-02-04 3_make_cities.R

@@ -1,0 +1,61 @@
+library(sp, quietly = TRUE)
+library(grid, quietly = TRUE)
+library(rgdal, quietly = TRUE)
+library(stringr, quietly = TRUE)
+library(tidyverse, quietly = TRUE)
+
+city_df <- read_csv('assumptions/cities.csv')
+cities <- unique(city_df$name)
+cities <- cities %>% setdiff(list.dirs('data/cities', full.names = F))
+
+city_df <- city_df %>% 
+    filter(name %in% cities)
+
+v_lookup_code <- Vectorize(tigris::lookup_code)
+
+city_df <- city_df %>% 
+    dplyr::mutate(state_fips = v_lookup_code(state),
+           state_fips = stringr::str_sub(state_fips, start = -4, end = -3),
+           state_fips = as.integer(state_fips))
+
+states <- city_df$state %>% 
+	str_split('--') %>% 
+	unique()
+
+the_map <- states %>%
+    purrr::map(~readOGR(dsn = paste0('data/states/',.x), 
+                       layer = 'geo',
+                       verbose = FALSE)) %>% 
+    reduce(rbind, makeUniqueIDs = TRUE)
+
+if(!dir.exists('data/cities')){
+	dir.create('data/cities')
+}
+
+make_city <- function(city){
+    this_city <- city_df %>% filter(name == city)
+    
+    counties <- this_city$county
+    states   <- this_city$state_fips
+    states   <- str_pad(states, width = 2, pad = 0)
+    right_county <- as.integer(as.character(the_map@data$COUNTYFP)) %in% counties
+    right_state  <- the_map@data$STATEFP %in% states
+    
+    the_map[right_county & right_state,]
+}
+
+writer <- function(tracts, city){
+    writeOGR(tracts, 
+             paste0("data/cities/", city),
+             'geo', 
+             driver = 'ESRI Shapefile', 
+             morphToESRI = TRUE)
+}
+safe_writer <- safely(writer)
+
+cities %>% 
+    as.list() %>% 
+    map(make_city) %>% 
+    walk2(cities, safe_writer)
+
+

### 2017-02-04 4_compute_metric.R

@@ -1,0 +1,103 @@
+library(compx)
+library(rgdal, quietly = TRUE)
+library(stringr, quietly = TRUE)
+library(tidyverse, quietly = TRUE)
+
+source('R/smoother.R')
+
+# Prep data
+
+demographics <- read_csv('data/demographics.csv') %>% 
+    dplyr::mutate(tract = as.character(GEOID),
+           tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
+    select(-GEOID) 
+
+tract_list <- demographics %>% 
+    group_by(tract) %>% 
+    dplyr::summarise(n = sum(n)) %>% 
+    filter(n > 100) %>% 
+    select(-n)
+
+if(!dir.exists('throughput/metric')){
+    dir.create('throughput/metric')
+}
+
+analysis <- function(city){
+    
+    tracts <- readOGR(dsn = paste0('data/cities/',city), 
+                      layer = 'geo',
+                      verbose = FALSE)
+    print(paste(city, ':', nrow(tracts), 'tracts'))
+    
+    
+    tracts <- tracts[tracts@data$GEOID %in% tract_list$tract,] 
+    area_df <- tracts@data %>% 
+        dplyr::select(GEOID) %>% 
+        mutate(area = abs(raster::area(tracts) / 1e6))
+    
+    data <- demographics %>% 
+        dplyr::filter(tract %in% tracts@data$GEOID)
+    
+    data <- rbf_smoother(data, tracts, sigma = .5)
+    
+    metric_df <- compute_metric(tracts, 
+                                data, 
+                                km = T, 
+                                sigma = 10, 
+                                divergence = 'DKL', 
+                                smooth = T) %>% 
+        left_join(area_df, by = c('geoid' = 'GEOID'))
+    
+    metric_df <- metric_df %>% 
+        mutate(vol = map_dbl(local_g, . %>% det() %>% abs() %>% sqrt()),
+               tr  = map_dbl(local_g, . %>% diag() %>% sum())) 
+    
+
+    
+    
+    list(metric_df = metric_df, 
+         tracts    = tracts) %>% 
+        saveRDS(paste0('throughput/metric/', city, '.RDS'))
+    
+    mutual_info <- data %>% 
+        spread(key = group, value = n) %>% 
+        select(-tract) %>% 
+        compx::mutual_info()
+    
+    local_entropy <- data %>% 
+        nest(-tract) %>% 
[…]

### 2017-02-04 5_network_structure.R

@@ -1,0 +1,62 @@
+library(compx, quietly = T)
+library(tidyverse, quietly = T)
+library(rgdal)
+
+source('R/smoother.R')
+
+demographics <- read_csv('data/demographics.csv') %>% 
+    dplyr::mutate(tract = as.character(GEOID),
+           tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
+    select(-GEOID)
+
+tract_list <- demographics %>% 
+    group_by(tract) %>% 
+    dplyr::summarise(n = sum(n)) %>% 
+    filter(n > 0) %>% 
+    select(-n)
+
+if(!dir.exists('throughput/graphs')){
+    dir.create('throughput/graphs')
+}
+
+analysis <- function(city){
+    print(city)
+    
+    # get the tracts, compute area
+    tracts <- readOGR(dsn = paste0('data/cities/',city), 
+                      layer = 'geo',
+                      verbose = FALSE)
+    
+    tracts <- tracts[tracts@data$GEOID %in% tract_list$tract,] 
+    area_df <- tracts@data %>% 
+        select(GEOID) %>% 
+        mutate(area = map_dbl(1:nrow(tracts), ~ tracts@polygons[[.]]@area))
+    
+    # filter the demographics to the tracts
+    
+    geoids <- tracts@data$GEOID %>% as.character() %>% unique()
+    
+    data <- demographics %>% 
+        filter(tract %in% geoids)
+    
+    data <- rbf_smoother(data, tracts, sigma = .5)
+    
+    # do the analysis
+    g <- construct_information_graph(tracts, data, divergence = 'DKL')
+    A <- affinity_matrix(g, sigma = 5)
+    L <- generalized_laplacian_matrix(A)
+    eig <- eigen(L, symmetric = T)
+    
+    g %>% saveRDS(file = paste0('throughput/graphs/', city, '.RDS'))
+    
+    # return the spectra, might add more later
+    data_frame(city = city, n = 1:length(eig$values), eig = rev(eig$values)) %>% 
+        mutate(eig = eig - min(eig),
+               eig = eig / max(eig))
+}
+
+list.dirs('data/cities', full.names = FALSE)[-1] %>% 
+    map(analysis) %>% 
+    reduce(rbind) %>% 
+    write_csv('throughput/spectra.csv')
+

### 2017-02-04 R/smoother.R

@@ -1,0 +1,42 @@
+library(compx, quietly = T)
+library(tidyverse, quietly = T)
+library(rgdal)
+library(rgeos)
+
+rbf_smoother <- function(data, tracts, sigma){
+    
+    # compute geographic distances
+    ids <- compx::id_lookup(tracts, key_col = 'GEOID')	
+    coords <- coords_df(tracts, km = T)
+    centroids <- coords %>% 
+        select(-geoid) %>% 
+        as.matrix() 
+    rownames(centroids) <- coords$geoid
+    dists <- dist(centroids) %>% 
+        as.matrix() 
+    
+    # construct the kernel matrix
+    # sigma <- .5
+    K <- exp(-sigma * dists^2)
+    adj <- gRelate(tracts, byid = TRUE, pattern = '****1****')
+    K[adj == 0] <- 0
+    
+    # Convert data to matrix form
+    P_df <- data %>% 
+        spread(key = group, value = n, fill = 0) 
+    P <- P_df %>% select(-tract) %>% 
+        as.matrix()
+    row.names(P) <- P_df$tract
+    P <- P[row.names(K),] # to ensure the orders match
+    
+    # compute the smoother and normalize
+    smoothed_P <- K %*% P
+    smoothed_P <- (smoothed_P / rowSums(smoothed_P)) * rowSums(P)
+    
+    # convert back into data frame form
+    smoothed_df <- smoothed_P %>% 
+        as.data.frame() %>% 
+        rownames_to_column('tract') %>% 
+        gather(key = group, value = n, -tract) %>% 
+        tbl_df() 
+}

### 2017-02-04 run_all.R

@@ -1,0 +1,7 @@
+library(tidyverse)
+c('2_get_states.R',
+  '3_make_cities.R', 
+  '4_compute_metric.R',
+  '5_network_structure.R') %>% 
+    map(source)
+

### 2017-08-23 6_figs.Rmd
after sf refactor

@@
 ---
-title: "Figures for The *Spatial Structure of Segregation*"
+title: "Figures for The Spatial Structure of Segregation"
 author: "Phil Chodrow"
@@
   html_document:
-    theme: journal
+    theme: sandstone
     toc: true
     toc_float: true
+    code_folding: hide
 ---
@@
 ```{r}
+library(magrittr)
 library(tidyverse)
@@
 library(ggmap)
+library(sf)
+library(stringr)
 ```
@@
 }
+
+cities <- c('Philadelphia', 
+            'Detroit', 
+            'Atlanta', 
+            'Chicago', 
+            'New York City', 
+            'Los Angeles', 
+            'Seattle', 
+            'San Francisco', 
+            'New Orleans')
 ```
@@
 
-First we read in the data...
 ```{r}
 
-city     <- 'Detroit'
+metric_map <- function(city){
+
     input    <- readRDS(paste0('throughput/metric/', city, '.RDS'))
-f_tracts <- input$tracts %>%
-    fortify(region = 'GEOID')
 
@@
 
-f_tracts <- f_tracts %>% 
-    left_join(metric_df, by = c('id' = 'geoid')) 
+    input$tracts %>% 
+        left_join(metric_df, by = c('GEOID' = 'geoid')) %>% 
+        ggplot() + 
+        geom_sf(aes(fill = tr/(4 * pi)), size = 0) + 
+        scale_fill_viridis(option = 'magma' , 
+                           trans = 'log10',
+                           limits = c(.005, .7),
+                           oob = squish) +
+        ggthemes::theme_map() + 
+        theme(legend.justification = c(0,0), 
+              legend.position      = c(.8,.02),
+              # panel.background = element_rect(fill = 'grey80', color = NA),
+              legend.background = element_rect(fill = NA)) + 
+        guides(fill = guide_colorbar(ticks = T, title = expression(atop(italic(j(x)),(nats/km^2)))))
+        # guides(fill = guide_colorbar(ticks = F, title = expression(sqrt(group("|" , det~italic(g[x]), "|")))))
+
+}
 ```
@@
 ```{r}
-# the tr/4 is from our existence theorem
-f_tracts %>% 
-    ggplot() + 
-    aes(x = long, y = lat, fill = tr/4, group = group) + 
-    geom_polygon() + 
-    scale_fill_viridis(option = 'magma' , 
-                                trans = 'log10',
-                       limits = c(NA, 12e-1),
-                                oob = squish) +
-    theme_map() + 
-    theme(legend.justification = c(0,0), 
-          legend.position      = c(.8,.1),
-          panel.background = element_rect(fill = 'grey80', color = NA),
-          legend.background = element_rect(fill = NA)) + 
-    guides(fill = guide_colorbar(ticks = F, title = expression(tr~italic(g[x]))))
+map_width  <- 4
+map_height <- 3
+```
 
-ggsave('figs/local_density.pdf')
+```{r}
+city <- 'Detroit'
+metric_map(city) + theme(legend.text=element_text(size=10), legend.title = element_text(size=10))
+paste0('figs/spatial_trace_', city, '.pdf') %>% ggsave(width = map_width, height = map_height)
+```
 
+```{r}
+city <- 'Atlanta'
+metric_map(city) + guides(fill = FALSE) + theme(legend.text=element_text(size=10))
+paste0('figs/spatial_trace_', city, '.pdf') %>% ggsave(width = map_width, height = map_height)
+```
+
+```{r}
+city <- 'Chicago'
+metric_map(city) + guides(fill = FALSE) + theme(legend.text=element_text(size=10))
+paste0('figs/spatial_trace_', city, '.pdf') %>% ggsave(width = map_width, height = map_height)
+```
+
+```{r}
+city <- 'Philadelphia'
+metric_map(city) + theme(legend.text=element_text(size=10), legend.title = element_text(size=10)) + guides(fill = FALSE)
+paste0('figs/spatial_trace_', city, '.pdf') %>% ggsave(width = map_width, height = map_height)
+```
+
+```{r}
+city <- 'Dallas'
+metric_map(city) + theme(legend.text=element_text(size=10), legend.title = element_text(size=10))
+paste0('figs/spatial_trace_', city, '.pdf') %>% ggsave(width = map_width, height = map_height)
 ```
@@
 
-summary <- read_csv('throughput/metric.csv')
+demographics <- read_csv('data/demographics.csv') %>% 
+        dplyr::mutate(tract = as.character(GEOID),
+               tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
+        select(-GEOID)
 
-# summary <- summary %>% 
-#     mutate(p_vol = p_vol / max(p_vol),
-#            p_trace = p_trace / max(p_trace))
+```
+
+```{r}
+assumptions <- read_csv('assumptions/cities.csv')
+
+code_lookup <- tigris::fips_codes %>% 
+    select(state, state_code) %>% 
+    distinct(state, state_code, keepall = TRUE)
+
+assumptions <- assumptions %>% 
+    left_join(code_lookup, by = c('state' = 'state')) %>% 
+    mutate(county = stringr::str_pad(county, width = 3, side = 'left', pad = '0'))
+
+dem <- demographics %>% 
+    mutate(state = stringr::str_sub(tract, start = 1, end = 2), 
+           county = stringr::str_sub(tract, start = 3, end = 5)) %>% 
+    left_join(assumptions, by = c('state' = 'state_code', 'county' = 'county')) 
+```
+
+```{r}
+MIs <- dem %>% 
+    filter(!is.na(name)) %>% 
+    select(-state, -county, -state.y) %>% 
+    tidyr::spread(key = group, value = n, fill = 0) %>% 
+    select(-tract) %>% 
+    nest(-name) %>% 
+    mutate(I_XY = map_dbl(data, mutual_info)) %>% 
+    select(-data)
+
+```
+
+```{r}
+
+summary <- read_csv('throughput/metric.csv') %>% 
+    left_join(MIs, by = c('city' = 'name'))
 
 highlight <- summary %>%
-    filter(I_XY < 0 | I_XY > .35)
+    filter(city %in% cities)
+
+```
+
+```{r}
+summary %>% 
+    write_csv('figs/metric_summary.csv')
+```
+
+Unevenness profile:
+
+```{r}
+highlight <- summary %>% 
+    filter(city %in% cities)
 
@@
     ggplot() + 
-    aes(x = p_vol, y = p_trace) + 
-    scale_y_continuous(trans = 'log10') +
-    scale_x_continuous(trans = 'log10') +
+    aes(x = I_XY , y = trim_trace / (4 * pi) ) + 
+    scale_x_continuous(limits = c(0, NA)) + 
+    scale_y_continuous(limits = c(0, NA)) + 
+    geom_point(color = 'firebrick') + 
+    # geom_point(color = 'firebrick', data = highlight) + 
+    # geom_text_repel(aes(label = city), data = highlight) +
+    geom_text_repel(aes(label = city), size = 3, segment.size = .3) +
+    scatterplot_theme() + 
+    xlab(expression(italic(I[f](X,Y))~(nats))) + 
+    ylab(expression(italic(J(X,Y))~(nats/km^2))) + 
+    annotation_logticks(scaled = TRUE, size = .2, sides = 'l')
+
+ ggsave('figs/unevenness_density.pdf', width = 7, height = 5)
+
+```
+
+```{r}
+
+cities <- list.files(path = 'throughput/metric/') %>% 
+    map(str_sub, end = -5)
+
+metric_df <- list.files(path = 'throughput/metric/', full.names = TRUE) %>% 
+    map(readRDS) %>% 
+    map('metric_df') %>% 
+    map2(cities, ~mutate(.x, city = .y)) %>% 
+    reduce(rbind)
+
+metric_df %>% 
+    mutate(norm_tr = tr / H) %>% 
+    group_by(city) %>% 
+    summarise(trim_trace = mean(norm_tr, trim = .01, na.rm = T)) %>% 
+    left_join(summary, by = c('city' = 'city')) %>% 
+    ggplot() + 
+    aes(x = I_XY , y = trim_trace.x / (4 * pi)) + 
+    scale_x_continuous(limits = c(0, NA)) + 
+    scale_y_continuous(limits = c(0, .15)) + 
+    geom_point(color = 'firebrick') + 
+    geom_text_repel(aes(label = city), size = 3, segment.size = .3) +
+    scatterplot_theme() + 
+    xlab(expression(italic(I(X,Y))~(nats))) + 
+    ylab(expression(italic(J(X,Y))~(nats/km^2))) + 
+    annotation_logticks(scaled = TRUE, size = .2, sides = 'l')
+
+ggsave('figs/unevenness_normalized_density.pdf', width = 7, height = 5)
+
+```
+
+```{r}
+highlight <- summary %>% 
+    filter(city %in% cities)
+
+ summary %>% 
+    ggplot() + 
+    aes(x = I_XY , y = trim_trace * n_tracts / total_pop ) + 
+    # scale_y_continuous(trans = 'log10') +
     geom_point(color = 'grey') + 
@@
     scatterplot_theme() + 
-    xlab(expression(italic(E[X]) ~ sqrt(~det~g[X]))) + 
-    ylab(expression(italic(E[X]) ~ (tr ~~  g[X]))) + 
-    annotation_logticks(scaled = TRUE, size = .2, sides = 'lb')
+    xlab(expression(italic(I[f](X,Y)))) + 
+    ylab(expression(italic(J(X,Y)))) + 
+    annotation_logticks(scaled = TRUE, size = .2, sides = 'l')
 
- ggsave('figs/average_geometry.pdf')
+ ggsave('figs/unevenness_density_pop.pdf')
 
-```
-
-```{r}
-summary <- read_csv('throughput/metric.csv')
-
-highlight <- summary %>% 
-    filter(I_XY < 0 | I_XY > .35)
-
-summary %>% 
-    ggplot() + 
-    aes(x = I_XY, y = p_trace) + 
-    scale_y_continuous(trans = 'log10') + 
-    geom_point(color = 'grey') + 
-    geom_point(color = 'firebrick', data = highlight) + 
-    geom_text_repel(aes(label = city), data = highlight) +
-    scatterplot_theme() + 
-    ylab(expression(italic(E[X]) ~ (tr ~~  g[X]))) + 
-    xlab(expression(italic(I(X,Y)))) + 
-    annotation_logticks(scaled = TRUE, size = .2, sides = 'l') +
-
-ggsave('figs/density_mutual.pdf')
 ```
@@
 
-cities <- c('Philadelphia', 'Detroit', 'Atlanta', 'Chicago', 'New York City')
 highlight <- spectra %>% 
@@
           axis.text.y = element_text(size = 11, margin=unit(c(0.5,0.5,0.1,0.0), "cm")),
-          axis.title.y = element_text(margin=unit(c(0.1,-0.1,0.5,0.5), "cm"), angle = 90)
+          axis.title.y = element_text(angle = 90),
+          panel.grid.major = element_blank(), 
+          panel.grid.minor = element_blank()
           )
@@
     list(layout, 
-         scale_color_brewer(palette = 'Set1'),
+         scale_color_brewer(palette = 'Dark2'),
          guides(color = guide_legend(title = element_blank())))
@@
     geom_line(color = 'grey', alpha = 1, size = .3) + 
-    scale_y_continuous(trans = 'log10', limit = c(8e-2, NA), expand=c(0,0)) + 
-    scale_x_continuous(limit = c(0, 100), expand=c(0,0)) + 
+    scale_y_continuous(trans = 'log10', limit = c(5e-1, NA), expand=c(0,0)) + 
+    scale_x_continuous(limit = c(0, 50), expand=c(0,0)) + 
     geom_line(aes(color = city),  data = highlight, size = 1) + 
@@
 
-spectra <- spectra %>% 
-    mutate(diff = lead(eig) / eig)
+area_df <- summary %>% 
+    select(city, total_area, total_pop)
 
-highlight <- spectra %>%
+normed_spectra <- spectra %>% 
+    mutate(diff = abs(lead(eig)) / abs(eig) - 1) %>% 
+    group_by(city) %>% 
+    filter(!is.infinite(diff)) %>%
+    mutate(diff = diff / max(diff, na.rm = T)) %>%
+    filter(diff > 0, diff < 10)
+
+highlight <- normed_spectra %>%
     filter(city %in% cities)
 
-spectra %>% 
+normed_spectra %>% 
     ggplot() + 
@@
     geom_point(data = highlight, size = .5, color = 'black') +
-    scale_x_continuous(limits = c(2, 25)) +
-    scale_y_continuous(limits = c(1, 3.5)) +
+    scale_x_continuous(limits = c(2, 30)) +
+    scale_y_continuous(trans = 'log10', limits = c(1e-8, NA)) +
     ylab(expression(lambda[italic(n+1)] /lambda[italic(n)])) +
@@
 
-What happens when we incorporate area?
-
-```{r, echo = FALSE, eval = FALSE}
-
+```{r}
 area_df <- summary %>% 
-    select(city, total_area)
-
-spectra <- spectra %>% 
-    # left_join(area_df) %>% 
-    group_by(city) %>%
-    mutate(scale = n/max(n)) %>% 
-    ungroup()
-
-spectra <- spectra %>% 
-    mutate(diff = (1-eig) / (1-lead(eig)))
-
-highlight <- spectra %>%
-    filter(city %in% cities, scale < .1)
-
-spectra %>% 
-    filter(scale <= .1) %>%
-    ggplot() + 
-    aes(x = scale, y = diff, group = city) + 
-    # geom_point(size = 2) + 
-    geom_line(color = 'grey') + 
-    geom_line(aes(color = city),  data = highlight, size = 1) +
-    # geom_point(data = highlight, size = 1, color = 'black') + 
-    # scale_x_continuous(limits = c(1, 30)) +
-    scale_y_continuous(limits = c(1, 3.7)) +
-    theme_bw() + 
-    # ylab(expression(1-lambda[italic(n)])) + 
-    xlab(expression(italic(n))) + 
-    # annotation_logticks(scaled = T, sides = 'l', size = .2) + 
-    theme(axis.ticks.y = element_line(size = .2, color = 'black'),
-          axis.ticks.x = element_line(size = .2, color = 'black'),
-          legend.position = 'bottom') + 
-    ggtitle('Spectra of selected cities') + 
-    scale_color_brewer(palette = 'Set1') + 
-    guides(color = guide_legend(title = element_blank()))
-
-```
-
-What about when we normalize n?
-
-```{r}
+    select(city, total_area, total_pop)
 
 normed_spectra <- spectra %>% 
+    mutate(diff = lead(eig) - (eig)) %>% 
     group_by(city) %>% 
-    mutate(n = n/max(n))
+    filter(!is.infinite(diff)) %>%
+    mutate(diff = diff / max(diff, na.rm = T)) %>%
+    filter(diff > 0, diff < 10)
 
-normed_highlight <- normed_spectra %>% 
+highlight <- normed_spectra %>%
     filter(city %in% cities)
@@
     ggplot() + 
-    aes(x = n, y = 1-eig, group = city) + 
-    geom_line(color = 'grey', alpha = 1, size = .3) + 
-    scale_y_continuous(trans = 'log10', limit = c(1e-1, NA), expand=c(0,0)) + 
-    scale_x_continuous(limit = c(0.001, .5), expand=c(0,0), trans = 'log10') +
-    geom_line(aes(color = city),  data = normed_highlight, size = 1) + 
-    geom_point(data = normed_highlight, size = .3, color = 'black') + 
-    theme_bw() + 
-    ylab(expression(1-lambda[italic(n)])) + 
+    aes(x = n, y = diff, group = city) + 
+    geom_line(alpha = .5, color = 'grey') +
+    geom_line(aes(color = city), data = highlight, size = .7) + 
+    geom_point(data = highlight, size = .5, color = 'black') +
+    scale_x_continuous(limits = c(2, 30)) +
+    scale_y_continuous(trans = 'log10', limits = c(1e-8, NA)) +
+    ylab(expression(lambda[italic(n+1)] /lambda[italic(n)])) +
     xlab(expression(italic(n))) + 
-    annotation_logticks(scaled = T, sides = 'l', size = .2) + 
-    theme(axis.ticks.y = element_line(size = .2, color = 'black'),
-          axis.ticks.x = element_line(size = .2, color = 'black'),
-          legend.position = 'bottom') + 
-    ggtitle('Spectra of selected cities') + 
-    scale_color_brewer(palette = 'Set1') + 
-    guides(color = guide_legend(title = element_blank()))
+    spectra_theme()
 
-ggsave('figs/normed_spectra.pdf')
+```
+
+What happens when we incorporate area?
+
+```{r}
+
+spectra <- spectra %>% 
+    left_join(area_df) %>%
+    mutate(scale = total_pop / n )
+
+normed_spectra <- spectra %>% 
+    mutate(diff = lead(eig) - eig) %>% 
+    group_by(city) %>% 
+    mutate(diff = diff * max(n)) %>% 
+    filter(n < .03 * max(n))
+
+highlight <- normed_spectra %>%
+    filter(city %in% cities)
+
+normed_spectra %>% 
+    ggplot() + 
+    aes(x = scale, y = diff, group = city) + 
+    geom_hline(yintercept = 1) + 
+    geom_line(alpha = .3, color = 'grey') +
+    geom_line(aes(color = city), data = highlight) + 
+    # geom_point(data = highlight, size = .1, color = 'black') +
+    scale_x_continuous(limits = c(NA, NA), trans = 'log10') +
+    scale_y_continuous( trans = 'log10') +
+    ylab(expression(italic(N)~(lambda[italic(n+1)] - lambda[italic(n)]))) +
+    xlab(expression(Characteristic~Population:~italic(p==frac(P,n)))) +  
+    spectra_theme()
+
+ggsave('figs/spectra_pop.pdf')
 
@@
 
-spectral_properties <- spectra %>% 
+spectra <- spectra %>% 
+    mutate(scale = sqrt(total_area / n * pi))
+
+normed_spectra <- spectra %>% 
+    mutate(diff = lead(eig) - eig) %>% 
     group_by(city) %>% 
-    filter(n >= 10 | eig  > 1e-10) %>% 
-    filter(eig < 1) %>% 
-    summarise(aff_trace =   sum( 1 - eig),
-              aff_det   = - sum(log(1-eig))) %>%
-    arrange(desc(aff_trace))
+    mutate(diff = diff * max(n)) %>% 
+    filter(n < .03 * max(n))
 
-spectral_properties %>% 
+highlight <- normed_spectra %>%
+    filter(city %in% cities)
+
+normed_spectra %>% 
+    ggplot() + 
+    aes(x = scale, y = diff, group = city) + 
+    geom_hline(yintercept = 1) + 
+    geom_line(alpha = .3, color = 'grey') +
+    geom_line(aes(color = city), data = highlight) + 
+    # geom_point(data = highlight, size = .1, color = 'black') +
+    scale_x_continuous(limits = c(NA, NA), trans = 'log10') +
+    scale_y_continuous( trans = 'log10') +
+    ylab(expression(italic(N)~(lambda[italic(n+1)] - lambda[italic(n)]))) +
+    xlab(expression(Characteristic~Length~(km):~italic(L)==sqrt(frac(italic(A), italic(n)*pi)))) + 
+    spectra_theme()
+
+ggsave('figs/spectra_geo.pdf')
+
+```
+
+```{r}
+normed_spectra %>% 
+    group_by(city) %>% 
+    mutate(n_tracts = max(n)) %>% 
+    filter(diff > 1, n < .9 * max(n)) %>% 
+    summarise(max_n = max(n),
+              n_tracts = max(n_tracts)) %>% 
+    mutate(max_n = max_n / n_tracts) %>% 
     left_join(summary, by = c('city' = 'city')) %>%
     ggplot() + 
-    aes(x = aff_trace, y = total_vol) + 
+    aes(x = total_vol/total_area, y = max_n) + 
     geom_point() + 
     geom_text_repel(aes(label = city)) + 
-    scale_y_continuous(trans = 'log10') + 
+    scale_x_continuous(trans = 'log10') 
+    # scale_y_continuous(trans = 'log10') 
+```
+
+# Hierarchical Clustering
+
+```{r}
+loss_curves <- read_csv('throughput/hclust_loss.csv') %>% 
+    left_join(summary, by = c('city' = 'city'))
+
+loss_curves <- loss_curves %>% 
+    group_by(city) %>% 
+    mutate(value = info - lag(info))
+
+highlight <- loss_curves %>% 
+    filter(city %in% cities)
+
+loss_curves %>% 
+    ggplot() + 
+    aes(x = n, y = info, group = city) + 
+    geom_line(color = 'grey') + 
+    geom_line(aes(color = city), data = highlight) + 
+    scale_x_continuous(limits = c(1, NA), trans = 'log10') +
+    spectra_theme()
+```
+
+```{r}
+
+cities_h <- c('Atlanta', 'Chicago', 'Detroit', 'Milwaukee')
+
+loss_curves <- loss_curves %>% 
+    group_by(city) %>% 
+    mutate(value = info - lag(info))
+
+highlight <- loss_curves %>% 
+    filter(city %in% cities_h)
+
+loss_curves %>% 
+    ggplot() + 
+    aes(x = n, y = value, group = city) + 
+    geom_line(color = 'grey', alpha = .5) +
+    geom_line(aes(color = city), data = highlight) + 
+    geom_point(aes(color = city), data = highlight) + 
+    scale_x_continuous(limits = c(2, 12), breaks = 2*(1:8)) + 
+    scale_y_continuous(limits = c(0, NA), expand = c(0,.01)) + 
+    spectra_theme() + 
+    ylab(expression(italic(I(C[k],Y) - I(C[k-1],Y)))) + 
+    xlab(expression(italic(k))) + 
+    theme(legend.position = 'bottom',
+          axis.ticks.length=unit(-0.1, "cm"))
+
+ggsave('figs/hclust_differenced_loss_curves.pdf')
+```
+
+```{r}
+loss_curves <- loss_curves %>% 
+    mutate(length_scale = sqrt(total_area / (pi * n)))
+
+highlight <- loss_curves %>% 
+    filter(city %in% cities)
+point_highlight <- highlight %>% 
+    filter(length_scale > 2)
+
+loss_curves %>% 
+    ggplot() + 
+    aes(x = length_scale, y = value, group = city) + 
+    geom_line(color = 'grey', alpha = .3) + 
+    geom_line(aes(color = city), data = highlight) + 
+    geom_point(data = point_highlight, size = .2) + 
+    scale_x_continuous(limits = c(0, 20)) +
+    scale_y_continuous(limits = c(0, NA)) + 
+    spectra_theme()
+
+ggsave('figs/hclust_geo_scale_differenced.pdf')
+
+```
+
+```{r}
+loss_curves %>% 
+    group_by(city) %>% 
+    filter(info / max(info) > .5) %>% 
+    filter(max(info) > .3) %>%
+    filter(n == min(n)) %>% 
+    ggplot() + 
+    aes(y = n, x = trim_trace) + 
+    geom_smooth(se = F) + 
+    geom_text_repel(aes(label = city)) + 
+    geom_point()
+```
+
+# AOC calc?
+
+```{r}
+reg_df <- spectra %>% 
+    group_by(city) %>% 
+    filter(eig != 0, eig <= .9999) %>% 
+    # filter(n > 800) %>% 
+    summarise(log_det = mean((-log(eig)))) %>% 
+    left_join(summary)
+# %>% 
+#     filter(I_XY > .2)
+reg_df %>% 
+    ggplot() + 
+    aes(x = log_det, y = trim_trace) + 
+    geom_point() + 
+    geom_text_repel(aes(label = city)) + 
+    scatterplot_theme()
+```
+
+```{r}
+reg_df %>% 
+    lm(log_det ~ trim_trace + I_XY,  data = .) %>% 
+    summary()
+```
+
+```{r}
+reg_df <- loss_curves %>% 
+    filter(info - lag(info) > .01 | n > 10) %>%
+    mutate(n = row_number()) %>% 
+    # filter(row_number() < 10) %>% 
+    summarise(
+              trim_trace = mean(trim_trace),
+              I_XY = mean(I_XY),
+              AOC = 1 / (max(info) * log(max(n_tracts))) * sum(info * log((n + 1) / n)),
+              H = mean(H))
+
+reg_df %>% 
+    ggplot() + 
+    aes(x = trim_trace, y = AOC) +
+    geom_point() + 
+    geom_text_repel(aes(label = city)) + 
+    geom_smooth() + 
+    scatterplot_theme()
+```
+
+```{r}
+reg_df %>% 
+    lm(AOC ~ trim_trace + I_XY, data = .) %>%
+    summary()
+```
+
+```{r}
+loss_curves <- loss_curves %>% 
+    mutate(length_scale = sqrt(total_area / (pi * n)),
+           diff_value   = (info - lag(info)) / (log(n) - log(lag(n)))) %>% 
+    group_by(city) %>% 
+    mutate(baseline = max(info) / log(n()),
+           to_plot = diff_value - baseline) %>% 
+    filter(length_scale > 0)
+
+highlight <- loss_curves %>% 
+    filter(city %in% cities)
+point_highlight <- highlight %>% 
+    filter(length_scale > 2)
+
+loss_curves %>% 
+    ggplot() + 
+    aes(x = length_scale, y = value, group = city) + 
+    geom_hline(yintercept = 0) + 
+    geom_line(color = 'grey', alpha = .3) + 
+    geom_line(aes(color = city), data = highlight) + 
+    # geom_point(data = point_highlight, size = .2) + 
+    scale_x_continuous(limits = c(1, 100), trans = 'log10') +
+    spectra_theme()
+
+ggsave('figs/hclust_geo_scale_differenced.pdf')
+
+```
+
+```{r}
+loss_curves <- loss_curves %>% 
+    dplyr::group_by(city) %>% 
+    dplyr::mutate(rev_info = max(info) - info)
+
+highlight <- loss_curves %>% 
+    dplyr::filter(city %in% cities)
+
+loss_curves %>% 
+    ggplot() + 
+    aes(x = length_scale, y = rev_info, group = city) + 
+    geom_line(color = 'grey', alpha = .3) + 
+    geom_line(aes(color = city), data = highlight) + 
+    # geom_point(data = point_highlight, size = .2) + 
+    scale_x_continuous(limits = c(NA, 100), trans = 'log10') +
+    spectra_theme() + 
+    ylab(expression(italic(I(C[n],Y)))) +
+    xlab(expression(Characteristic~Length~(km):~italic(L)==sqrt(frac(italic(A), italic(n)*pi))))
+
+ggsave('figs/hclust_geo_scale_cumulative.pdf')
+
+```
+
+```{r}
+loss_curves <- loss_curves %>% 
+    dplyr::mutate(pop_scale = total_pop / n)
+highlight <- loss_curves %>% 
+    dplyr::filter(city %in% cities)
+loss_curves  %>% 
+    ggplot() + 
+    aes(x = pop_scale, y = rev_info, group = city) + 
+    geom_line(color = 'grey', alpha = .3) + 
+    geom_line(aes(color = city), data = highlight) + 
+    scale_x_continuous(limits = c(1e3, 6e6), trans = 'log10') +
+    spectra_theme() + 
+    ylab(expression(italic(I(C[n],Y)))) +
+    xlab(expression(Characteristic~Population:~italic(p==frac(P,n))))
+```
+
+# Gathered Loss Curves
+
+```{r}
+
+gathered_spectra <- spectra%>% 
+    select(-scale) %>% 
+    left_join(area_df) %>%
+    mutate(length_scale = sqrt(total_area / (pi * n)),
+           pop_scale    = total_pop / n) %>% 
+    group_by(city) %>% 
+    mutate(diff = log(max(n) * (lead(eig) - eig))) %>% 
+    ungroup() %>% 
+    gather(key = type, value = scale, length_scale, pop_scale) %>% 
+    group_by(city) %>% 
+    filter(n < .3 * max(n)) %>% 
+    mutate(measure_type = 'spectra')
+
+highlight <- gathered_spectra %>% 
+    filter(city %in% cities)
+
+gathered_spectra %>% 
+    ggplot() + 
+    aes(x = scale, y = diff, group = city) + 
+    geom_line(color = 'grey', alpha = .3) + 
+    geom_line(aes(color = city), data = highlight) + 
+    facet_wrap(~type, scales = 'free_x') + 
     scale_x_continuous(trans = 'log10') +
-    geom_smooth() + 
-    ggtitle('Correlation between the trace of the affinity matrix and information volume.')
+    # scale_y_continuous(trans = 'log10') +
+    spectra_theme()
+
+```
+
+```{r}
+
+gathered_hclust <- loss_curves %>% 
+    mutate(length_scale = sqrt(total_area / (pi * n)),
+           pop_scale = total_pop / n) %>% 
+    group_by(city) %>% 
+    mutate(baseline = max(info) / log10(max(n)),
+           info =(info - lag(info))/(log10(n) - log10(lag(n))) - baseline,
+           info = info
+           ) %>% 
+    gather(key = type, value = scale, length_scale, pop_scale) %>% 
+    mutate(measure_type = 'hierarchical') %>%
+    filter(n < .3 * max(n))
+
+highlight <- gathered_hclust %>% 
+    filter(city %in% cities)
+
+gathered_hclust %>% 
+    ggplot() + 
+    aes(x = scale, y = info, group = city) + 
+    geom_hline(yintercept = 1) + 
+    geom_line(color = 'grey', alpha = .3) + 
+    geom_line(aes(color = city), data = highlight) + 
+    facet_wrap(~type, scales = 'free_x') + 
+    scale_x_continuous(trans = 'log10') +
+    # scale_y_continuous(trans = 'log10') +
+    spectra_theme()
+```
+
+```{r, fig.height = 5}
+
+g_spectra <- gathered_spectra %>% 
+    select(city, scale, type, loss = diff, measure_type)
+
+g_hclust <- gathered_hclust %>% 
+    select(city, scale, type, loss = info, measure_type)
+
+gathered <- rbind(g_hclust, g_spectra)
+
+gathered <- gathered %>% 
+    mutate(y = 0)
+
+gathered <- gathered %>% 
+    mutate(measure_type = ifelse(measure_type == 'hierarchical', 'Hierarchical', 'Spectral'),
+           type = ifelse(type == 'length_scale', 'Length~Scale', 'Population~Scale'))
+
+highlight <- gathered %>% 
+    filter(city %in% cities)
+
+```
+
+```{r, fig.height = 6}
+
+gathered %>% 
+    ggplot() + 
+    aes(x = scale, y = loss, group = city) + 
+    geom_hline(aes(yintercept = y), size = .5) + 
+    geom_line(color = 'grey', alpha = 1, size = .2) + 
+    geom_line(aes(color = city), data = highlight) + 
+    facet_grid(measure_type~type, 
+               scales = 'free',
+               labeller = label_parsed) + 
+    scale_x_continuous(trans = 'log10') +
+    spectra_theme() + 
+    theme(legend.text=element_text(size=12),
+          legend.position="bottom",
+          strip.background = element_blank(),
+          strip.text = element_text(size = 12)) + 
+    xlab(expression(italic(S(k)))) + 
+    ylab(expression(italic(Q(k))))
+
+ggsave('figs/all_loss_curves.pdf')
 
@@
 ```{r, fig.height = 6, fig.width = 10}
+city <- 'Detroit'
 g <- readRDS(paste0('throughput/graphs/',city,'.RDS'))
-A <- affinity_matrix(g, sigma = 1)
-sigma <- 1
-clust <- hclust(as.dist(-A),  method = 'complete')
+sigma <- 2
+
 pal   <- brewer.pal(9, name = "Set1") %>%
 	colorRampPalette()
-
-ncluster <- 7
-g <- g %>% set_vertex_attr(name = 'cluster', index = colnames(A), value = clust %>% cutree(ncluster))
 
@@
 ```{r}
+city <- 'Detroit'
+input <- readRDS(paste0('throughput/metric/', city, '.RDS'))
+tracts <- input$tracts
+
 k <- 600
@@
 
-f_tracts %>% 
-    left_join(dist_lookup, by = c('id' = 'geoid')) %>% 
+tracts %>% 
+    left_join(dist_lookup, by = c('GEOID' = 'geoid')) %>%
     ggplot() + 
-    aes(x = long, y = lat, group = group, fill = dist) %>% 
-    geom_polygon() + 
+    geom_sf(aes(fill = dist), size = .1) +
     scale_fill_viridis(option = 'magma' ,
@@
           legend.position      = c(.8,.1),
-          panel.background = element_rect(fill = 'grey80', color = NA),
+          # panel.background = element_rect(fill = 'grey80', color = NA),
           legend.background = element_rect(fill = NA)) + 
@@
 edges <- g %>% as_long_data_frame() %>% tbl_df()
-nodes <- data_frame(x = V(g)$x, y = V(g)$y, cluster = as.character(V(g)$cluster))
+nodes <- data_frame(x = V(g)$x, y = V(g)$y)
 
 ggplot() + 
-    geom_polygon(aes(x = long, y = lat, group = group), fill = 'grey10', alpha = .7, data = f_tracts) +
-    geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, size = exp(-7 * dist^2), alpha = exp(-7 * dist^2)), color = 'white', data = edges) +
+    geom_sf(fill = 'grey10', alpha = .7, data = tracts) +
+    geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, size = exp(-50 * dist^2), alpha = exp(-50 * dist^2)), color = 'white', data = edges) +
     theme_map() +
-    scale_size_continuous(range = c(0,.5)) + 
-    guides(alpha = 'none', color = 'none', size = 'none') +
-    theme(panel.background = element_rect(fill = 'grey80', color = NA))
+    scale_size_continuous(range = c(0,2)) + 
+    guides(alpha = 'none', color = 'none', size = 'none') 
+# +
+#     theme(panel.background = element_rect(fill = 'grey80', color = NA))
 
 ggsave('figs/network_structure.pdf')
-
-```
-
-Now we'll have the color of the edge proportional to its geodesic distance.
-
-```{r}
-edges <- g %>% as_long_data_frame() %>% tbl_df()
-nodes <- data_frame(x = V(g)$x, y = V(g)$y, cluster = as.character(V(g)$cluster))
-
-ggplot() + 
-    geom_polygon(aes(x = long, y = lat, group = group), fill = 'grey80', alpha = .6, data = f_tracts) +
-    geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, color = dist), data = edges) +
-    theme_map() +
-    scale_color_viridis(option = 'magma', trans = 'log10') + 
-    guides(alpha = 'none', color = 'none', size = 'none') +
-    theme(panel.background = element_rect(fill = 'grey80', color = NA))
-
-ggsave('figs/network_structure_colored.pdf')
-
-```
-
-This one is a little bit busier, but shows fairly nicely how the network is most strongly separated at points of high curvature.
-
-```{r}
-edges <- g %>% as_long_data_frame() %>% tbl_df()
-nodes <- data_frame(x = V(g)$x, y = V(g)$y, cluster = as.character(V(g)$cluster))
-
-ggplot() + 
-    geom_polygon(aes(x = long, y = lat, fill = vol/area, group = group), data = f_tracts) + 
-    scale_fill_viridis(option = 'magma' , 
-                                trans = 'log10', 
-                                oob = squish) +
-    geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, size = exp(-10 * dist)), color = 'white', data = edges) +
-    theme_map() +
-    scale_size_continuous(range = c(0,1)) + 
-    guides(alpha = 'none', color = 'none', size = 'none') +
-    theme(panel.background = element_rect(fill = 'grey80', color = NA),
-          legend.justification = c(0,0), 
-          legend.position      = c(.8,.1),
-          legend.background = element_rect(fill = NA)) + 
-    guides(fill = guide_colorbar(ticks = F, title = expression(rho)))
-
-ggsave('figs/network_structure_curvature.pdf')
-```
-
-Now let's do it with spectral clustering.
-
-```{r}
-
-A <- affinity_matrix(g, sigma = 2)
-L <- generalized_laplacian_matrix(A)
-evL <- eigen(L, symmetric=TRUE)
-
-# let's pick a number of clusters by checking out the spectrum. 
-test_df <- data_frame(n = 1:nrow(A), eig = rev(evL$values)) %>% 
-	mutate(eig = eig - min(eig),
-		   eig = eig / max(eig)) %>% 
-	mutate(val = lead(eig) / eig) %>% 
-	filter(n < 50)
-
-plot(test_df$val, log = 'y')
-
-```
-
-```{r}
-
-make_clusters <- function(k, nreps = 1000){
-    Z <- evL$vectors[,(ncol(evL$vectors)-k+1):(ncol(evL$vectors))]
-
-    models <- data_frame(n = 1:nreps) %>% 
-    	dplyr::mutate(model = map(n, ~ kmeans(Z, centers = k))) %>% 
-    	mutate(perf = map_dbl(model, ~.$tot.withinss))
-
-    model <- models %>% 
-    	filter(perf == min(perf))
-    km <- model$model[[1]]
-    return(km$cluster)
-}
-
-get_node_df <- function(k){
-    h <- g %>% set.vertex.attribute('cluster', value = make_clusters(k)) 
-    data_frame(geoid = V(h)$name, 
-               x = V(h)$x, 
-               y = V(h)$y, 
-               cluster = V(h)$cluster, 
-               k = k)
-}
-
-```
-
-```{r}
-
-ks <- c(2, 6, 10, 15)
-
-clusters <- ks %>% 
-    map(get_node_df) %>% 
-    reduce(rbind)
-
-pal <- brewer_pal(palette = 'Set1')
-pal <- colorRampPalette(pal(9))
-pal <- pal(max(ks))
-
-ggplot() + 
-	geom_segment(aes(x = from_x, y = from_y, xend = to_x, yend = to_y, size = exp(-5 * dist^2)), color = 'black', alpha = .5, data = edges) +
-	theme_map() +
-	scale_size_continuous(range = c(0,1)) + 
-	guides(alpha = 'none', color = 'none', size = 'none') +
-	theme(panel.background = element_rect(fill = 'grey80', color = NA)) + 
-	geom_point(aes(x, y, color = as.character(cluster)), data = clusters) + 
-    scale_color_manual(values = pal) + 
-    facet_wrap(~k, nrow = 3)
-
-ggsave('figs/spectral_clustered.pdf')
-
-```
-
-```{r}
-
-label_generator <- function(variable, value){
-    return(paste0('k = ', value))
-}
-
-f_tracts %>% 
-    left_join(clusters, by = c('id' = 'geoid')) %>% 
-    ggplot() + 
-    aes(x = long, y = lat, fill = as.character(cluster), group = id) + 
-    facet_wrap(~k, nrow = 2, labeller = label_generator) + 
-    geom_polygon() + 
-    theme_map() +
-    theme(panel.background = element_rect(fill = 'grey80', color = NA)) + 
-    scale_fill_manual(values = pal) +
-    guides(fill = FALSE)
-
-ggsave('figs/clustered_maps.pdf')
-
-```
-
-## Matrix Structure
-
-Now let's visualize the affinity matrix! We can use hierarchical clustering for this task.
-
-```{r}
-sigma = 1
-affinity_matrix_viz <- function(city){
-    g <- readRDS(paste0('throughput/graphs/',city,'.RDS'))
-    A <- affinity_matrix(g, sigma = sigma)
-
-    clust <- hclust(as.dist(-A),  method = 'complete')
-
-    the_order <- data_frame(geoid = clust$labels[clust$order], order = 1:length(clust$labels))
-
-    A %>% 
-        as.data.frame() %>% 
-        rownames_to_column('geoid_1') %>% 
-        tbl_df() %>% 
-        gather(geoid_2, a, -geoid_1) %>% 
-        left_join(the_order, by = c('geoid_1' = 'geoid')) %>% 
-        dplyr::rename(x = order) %>%
-        left_join(the_order, by = c('geoid_2' = 'geoid')) %>% 
-        dplyr::rename(y = order) %>% 
-        ggplot() + 
-        aes(x = x, y = y, fill = a) + 
-        geom_raster() + 
-        scale_fill_viridis(option = 'viridis') + 
-        # theme_map() +
-        theme(legend.position = 'right',
-              axis.line=element_blank(),
-              axis.text.x=element_blank(),
-              axis.text.y=element_blank(),
-              axis.ticks=element_blank(),
-              axis.title.x=element_blank(),
-              axis.title.y=element_blank()) + 
-        scale_x_continuous(expand = c(0,0)) + 
-        scale_y_continuous(expand = c(0,0)) + 
-        guides(fill = guide_colorbar(ticks = F, title = expression(italic(A))))
-}
-
-```
-
-```{r, fig.width = 5}
-affinity_matrix_viz('Atlanta')
-ggsave('figs/Atlanta_affinity.pdf')
-```
-
-```{r, fig.width = 5}
-affinity_matrix_viz('Detroit')
-ggsave('figs/Detroit_affinity.pdf')
-```
-
-```{r}
-
-prominence_map <- function(tracts){
-    tracts <- input$tracts
-
-    demographics <- read_csv('data/demographics.csv') %>% 
-        dplyr::mutate(tract = as.character(GEOID),
-               tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
-        select(-GEOID)
-
-    data <- demographics %>% 
-        filter(tract %in% tracts$GEOID)
-
-    marginal <- data %>% 
-        group_by(group) %>% 
-        summarise(n = sum(n)) %>%
-        mutate(n = n / sum(n))
-
-    f <- function(df){
-        df$group[which.max(df$n *  log(df$n / marginal$n))][1]
-    }
-
-    g <- function(df){
-        max(df$n *  log(df$n / marginal$n), na.rm = T)
-    }
-
-    prominence_df <- data %>% 
-        group_by(tract) %>% 
-        mutate(n = n / sum(n)) %>% 
-        nest(-tract, .key = df) %>% 
-        mutate(prominent_group = map_chr(df, f),
-               prominence = map_dbl(df, g))
-
-    g <- tracts %>% fortify(region = 'GEOID') %>% 
-          tbl_df() %>% 
-          left_join(prominence_df, by = c('id' = 'tract')) %>% 
-          ggplot() + 
-          geom_polygon(aes(x = long, 
-                           y = lat, 
-                           fill = prominent_group, 
-                           group = group, 
-                           alpha = prominence))  +
-          ggthemes::theme_map() + 
-          scale_alpha_continuous(guide = 'none', limits = c(0,1)) +
-          theme(plot.title = element_text(size=20, vjust = -5),
-                legend.title = element_blank(),
-                legend.position = c(.7, .1))
-      g
-}
-```
-
-Generate Detroit, with legend
-
-```{r}
-city  <- 'Detroit'
-input <- readRDS(paste0('throughput/metric/', city, '.RDS'))
-
-prominence_map(input$tracts) + 
-    scale_fill_brewer(palette = 'Set1')
-
-ggsave('figs/detroit_prominence.pdf')
-```
-
-Generate Atlanta, no legend
-```{r}
-city  <- 'Atlanta'
-input <- readRDS(paste0('throughput/metric/', city, '.RDS'))
-
-prominence_map(input$tracts) + 
-    scale_fill_brewer(palette = 'Set1') + 
-    guides(fill = FALSE)
-
-ggsave('figs/atlanta_prominence.pdf')
-```
-
-```{r}
-city  <- 'Newark'
-input <- readRDS(paste0('throughput/metric/', city, '.RDS'))
-
-prominence_map(input$tracts) + 
-    scale_fill_brewer(palette = 'Set1') + 
-    guides(fill = FALSE)
-
-ggsave('figs/newark_prominence.pdf')
 

### 2017-08-23 7_cluster_figs.Rmd
after sf refactor

---
title: "7_cluster_figs"
author: "phil"
date: "7/17/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r}
map_width <- 4
map_height <- 3
```

```{r}
library(magrittr)
library(tidyverse)
library(scales)
library(ggthemes)
library(ggrepel)
library(igraph)
library(compx)
library(RColorBrewer)
library(gplots)
library(viridis)
library(ggmap)
library(sf)

knitr::read_chunk('R/utils.R')

# source('R/utils.R')
```

```{r utils}
```

```{r}
all_data <- read_csv('data/demographics.csv', 
                 col_types = cols(
                     GEOID = col_character()
                     )
                 ) %>% 
    rename(tract = GEOID) %>% 
    group_by(tract) %>% 
    filter(sum(n) > 0) %>% 
    ungroup() 
```

```{r}
df <- data_frame(city = character(), 
                 k = integer(), 
                 j = integer(),
                 hclust = numeric(), 
                 mixed = numeric())
```

```{r}
loss_curves <- data_frame(city = character(),
                          k = integer(),
                          h_val = numeric(), 
                          m_val = numeric())
```

```{r}
MI <- data_frame(city = character(), 
                 I_XY = numeric())
```

# Detroit

```{r}
city <- 'Detroit'
in_data <- read_data(city, all_data)
data    <- in_data$data
tracts  <- in_data$tracts
```

```{r}
g <- construct_information_graph(tracts = tracts, 
								 data = data, 
								 divergence = js_dist)

sigma <- 30
plot_spectrum(g, sigma)

ggsave(paste0('figs/spectrum_', city, '.pdf'))

```

```{r}
k     <- 34
g     <- spectral_cluster(g, sigma = sigma, k = k, nreps = 1000)
```

```{r}
input <- group_data(tracts, data, g) 
blocked_tracts <- input$blocked_tracts
blocked_data <- input$blocked_data
```

## Hierarchical Clustering

```{r}
b <- h_clust(tracts, data)
a <- h_clust(blocked_tracts, blocked_data)
```

```{r}
compare_hclusts(a, b)
```

```{r}

j <- 7

m <- prominence_map(tracts, data)

ggsave(paste0('figs/prominence_', city, '.pdf'), width = map_width, height = map_height)

s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = 1)
s

ggsave(paste0('figs/sclust_', city, '.pdf'), width = map_width, height = map_height)

h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = 1)
h

ggsave(paste0('figs/mixed_', city,'.pdf'), width = map_width, height = map_height)

```

```{r}
f <- h_clust_map(m, tracts, data, b, j, size_coef = 1)
f
ggsave(paste0('figs/hclust_', city,'.pdf'), width = map_width, height = map_height)
```

```{r}
# total MI
data %>% 
    spread(key = group, value = n) %>% 
    select(-tract) %>% 
    mutual_info()
```

```{r}
df <- df %>% 
    rbind(data_frame(
        city = city, 
        j = j,
        k = k, 
        hclust = cumsum(rev(b$height))[j],
        mixed = cumsum(rev(a$height))[j]
    ))

```

```{r}
loss_curves <- loss_curves %>%
    rbind(data_frame(
        city = city, 
        k = 1:length(a$height),
        h_val = rev(b$height)[1:length(a$height)],
        m_val = rev(a$height)
    ))

MI <- MI %>% 
    rbind(data_frame(
        city = city, 
        I_XY = data %>% 
            spread(key = group, value = n) %>% 
            select(-tract) %>% 
            mutual_info()
    ))
```

### Add place decomposition for first fig

```{r}
detroit <- st_union(tracts)

detroit_places <- tigris::places(c(26), cb = FALSE, year = 2015, refresh = FALSE) %>% 
	st_as_sf()

st_crs(detroit_places) <- st_crs(tracts)

places <- detroit_places %>% 
	mutate(include = st_intersects(detroit_places, detroit),
		   include = map_dbl(include, ~ifelse(length(.) == 0, 0, 1))) %>% 
	filter(include == 1) %>% 
	st_intersection(detroit) %>% 
    mutate(type = st_geometry_type(geometry)) %>%
    filter(type != 'GEOMETRYCOLLECTION')

m + 
	geom_sf(data = places, alpha = 0, size = .5, color = 'black') + 
    theme(legend.text = element_text(size = 8))

ggsave('figs/detroit_places.pdf', width = map_width, height = map_height)

```

```{r}
lookup <- st_join(tracts, detroit_places, st_within) %>% 
    select(GEOID.x, NAME.y) %>% 
    select(-geometry)

data %>% 
    left_join(lookup, by = c('tract' = 'GEOID.x')) %>% 
    mutate(NAME = ifelse(is.na(NAME.y), 'NA', NAME.y)) %>% 
    group_by(NAME, group) %>% 
    summarise(n = sum(n)) %>% 
    ungroup() %>% 
    spread(key = group, value = n) %>%
    select(-NAME) %>% 
    mutual_info()

```

```{r}

data %>% 
    spread(key = group, value = n) %>% 
    select(-tract) %>% 
    mutual_info()

```

# Chicago

```{r}
city    <- 'Chicago'
in_data <- read_data(city, all_data)
data    <- in_data$data
tracts  <- in_data$tracts

b <- h_clust(tracts, data)
```

```{r}
g <- construct_information_graph(tracts = tracts, 
								 data = data, 
								 divergence = js_dist)

sigma <- 30
plot_spectrum(g, sigma)
ggsave(paste0('figs/spectrum_', city, '.pdf'))
```

```{r}
k     <- 36
g     <- spectral_cluster(g, sigma = sigma, k = k, nreps = 1000)
```

```{r}
input <- group_data(tracts, data, g) 
blocked_tracts <- input$blocked_tracts
blocked_data <- input$blocked_data

blocked_tracts %>% 
    ggplot() + 
    geom_sf(aes(fill = as.character(cluster))) +
    guides(fill = FALSE)
```

## Hierarchical Clustering

```{r}
a <- h_clust(blocked_tracts, blocked_data)
compare_hclusts(a, b)
```

```{r}

j <- 11

m <- prominence_map(tracts, data) + 
        guides(fill = FALSE)

s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = 1)
s

ggsave(paste0('figs/sclust_', city, '.pdf'), width = map_width, height = map_height)

h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = 1)

h

ggsave(paste0('figs/mixed_', city,'.pdf'), width = map_width, height = map_height)
```

```{r}
f <- h_clust_map(m, tracts, data, b, j, size_coef = 1)
f
ggsave(paste0('figs/hclust_', city,'.pdf'), width = map_width, height = map_height)
```

```{r}
df <- df %>% 
    rbind(data_frame(
        city = city, 
        k = k, 
        j = j,
        hclust = cumsum(rev(b$height))[j],
        mixed = cumsum(rev(a$height))[j]
    ))
```

```{r}
loss_curves <- loss_curves %>%
    rbind(data_frame(
        city = city, 
        k = 1:length(a$height),
        h_val = rev(b$height)[1:length(a$height)],
        m_val = rev(a$height)
    ))

MI <- MI %>% 
    rbind(data_frame(
        city = city, 
        I_XY = data %>% 
            spread(key = group, value = n) %>% 
            select(-tract) %>% 
            mutual_info()
    ))
```

<!-- # Atlanta -->

<!-- ```{r} -->
<!-- city    <- 'Atlanta' -->
<!-- in_data <- read_data(city, all_data) -->
<!-- data    <- in_data$data -->
<!-- tracts  <- in_data$tracts -->

<!-- b <- h_clust(tracts, data) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- g <- construct_information_graph(tracts = tracts,  -->
<!-- 								 data = data,  -->
<!-- 								 divergence = js_dist) -->

<!-- sigma <- 30 -->
<!-- plot_spectrum(g, sigma) -->
<!-- ggsave(paste0('figs/spectrum_', city, '.pdf')) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- k     <- 26 -->
<!-- g     <- spectral_cluster(g, sigma = sigma, k = k, nreps = 1000) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- input <- group_data(tracts, data, g)  -->
<!-- blocked_tracts <- input$blocked_tracts -->
<!-- blocked_data <- input$blocked_data -->

<!-- blocked_tracts %>%  -->
<!--     ggplot() +  -->
<!--     geom_sf(aes(fill = as.character(cluster))) + -->
<!--     guides(fill = FALSE) -->
<!-- ``` -->

<!-- ## Hierarchical Clustering -->

<!-- ```{r} -->
<!-- a <- h_clust(blocked_tracts, blocked_data) -->
<!-- compare_hclusts(a, b) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- j <- 3 -->
<!-- m <- prominence_map(tracts, data) +  -->
<!--         guides(fill = FALSE)  -->

<!-- s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = 1) -->
<!-- s -->

<!-- ggsave(paste0('figs/sclust_', city, '.pdf'), width = map_width, height = map_height) -->

<!-- h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = 1) -->
<!-- h -->
<!-- ggsave(paste0('figs/mixed_', city,'.pdf'), width = map_width, height = map_height) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- f <- h_clust_map(m, tracts, data, b, j, size_coef = 1) -->
<!-- f -->
<!-- ggsave(paste0('figs/hclust_', city,'.pdf'), width = map_width, height = map_height) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df <- df %>%  -->
<!--     rbind(data_frame( -->
<!--         city = city,  -->
<!--         k = k,  -->
<!--         j = j, -->
<!--         hclust = cumsum(rev(b$height))[j], -->
<!--         mixed = cumsum(rev(a$height))[j] -->
<!--     )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- loss_curves <- loss_curves %>% -->
<!--     rbind(data_frame( -->
<!--         city = city,  -->
<!--         k = 1:length(a$height), -->
<!--         h_val = rev(b$height)[1:length(a$height)], -->
<!--         m_val = rev(a$height) -->
<!--     )) -->

<!-- MI <- MI %>%  -->
<!--     rbind(data_frame( -->
<!--         city = city,  -->
<!--         I_XY = data %>%  -->
<!--             spread(key = group, value = n) %>%  -->
<!--             select(-tract) %>%  -->
<!--             mutual_info() -->
<!--     )) -->
<!-- ``` -->

# Philadelphia

```{r}
city    <- 'Philadelphia'
in_data <- read_data(city, all_data)
data    <- in_data$data
tracts  <- in_data$tracts
```

## Filter out the small island
```{r}
g <- construct_information_graph(tracts = tracts, 
								 data = data, 
								 divergence = euc)

comps <- igraph::components(g)

largest_cluster <- which.max(table(comps$membership))
exclude <- which(comps$membership != largest_cluster) %>% names()

tracts <- tracts %>% 
    filter(!(GEOID %in% exclude))

data <- data %>%
    filter(tract %in% tracts$GEOID)
```

```{r}
b <- h_clust(tracts, data)
```

```{r}
g <- construct_information_graph(tracts = tracts, 
								 data = data, 
								 divergence = js_dist)

sigma <- 30
plot_spectrum(g, sigma)
ggsave(paste0('figs/spectrum_', city, '.pdf'))
```

```{r}
k     <- 11
g     <- spectral_cluster(g, sigma = sigma, k = k, nreps = 1000)
```

```{r}
input <- group_data(tracts, data, g) 
blocked_tracts <- input$blocked_tracts
blocked_data <- input$blocked_data

blocked_tracts %>% 
    ggplot() + 
    geom_sf(aes(fill = as.character(cluster))) +
    guides(fill = FALSE)
```

## Hierarchical Clustering

```{r}
a <- h_clust(blocked_tracts, blocked_data)
compare_hclusts(a, b)
```

```{r}
j <- 10
m <- prominence_map(tracts, data) + 
        guides(fill = FALSE)

s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = 1)
s

ggsave(paste0('figs/sclust_', city, '.pdf'), width = map_width, height = map_height)

h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = 1)
h
ggsave(paste0('figs/mixed_', city,'.pdf'), width = map_width, height = map_height)
```

```{r}
f <- h_clust_map(m, tracts, data, b, j, size_coef = 1)
f
ggsave(paste0('figs/hclust_', city,'.pdf'), width = map_width, height = map_height)
```

```{r}
df <- df %>% 
    rbind(data_frame(
        city = city, 
        k = k, 
        j = j,
        hclust = cumsum(rev(b$height))[j],
        mixed = cumsum(rev(a$height))[j]
    ))
```

```{r}
loss_curves <- loss_curves %>%
    rbind(data_frame(
        city = city, 
        k = 1:length(a$height),
        h_val = rev(b$height)[1:length(a$height)],
        m_val = rev(a$height)
    ))

MI <- MI %>% 
    rbind(data_frame(
        city = city, 
        I_XY = data %>% 
            spread(key = group, value = n) %>% 
            select(-tract) %>% 
            mutual_info()
    ))
```

<!-- # Dallas -->

<!-- ```{r} -->
<!-- city    <- 'Dallas' -->
<!-- in_data <- read_data(city, all_data) -->
<!-- data    <- in_data$data -->
<!-- tracts  <- in_data$tracts -->
<!-- ``` -->

<!-- ## Filter out the small island -->
<!-- ```{r} -->
<!-- g <- construct_information_graph(tracts = tracts,  -->
<!-- 								 data = data,  -->
<!-- 								 divergence = euc) -->

<!-- comps <- igraph::components(g) -->

<!-- largest_cluster <- which.max(table(comps$membership)) -->
<!-- exclude <- which(comps$membership != largest_cluster) %>% names() -->

<!-- tracts <- tracts %>%  -->
<!--     filter(!(GEOID %in% exclude)) -->

<!-- data <- data %>% -->
<!--     filter(tract %in% tracts$GEOID) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- b <- h_clust(tracts, data) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- g <- construct_information_graph(tracts = tracts,  -->
<!-- 								 data = data,  -->
<!-- 								 divergence = js_dist) -->

<!-- sigma <- 30 -->
<!-- plot_spectrum(g, sigma) -->
<!-- ggsave(paste0('figs/spectrum_', city, '.pdf')) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- k     <- 38 -->
<!-- g     <- spectral_cluster(g, sigma = sigma, k = k, nreps = 1000) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- input <- group_data(tracts, data, g)  -->
<!-- blocked_tracts <- input$blocked_tracts -->
<!-- blocked_data <- input$blocked_data -->

<!-- blocked_tracts %>%  -->
<!--     ggplot() +  -->
<!--     geom_sf(aes(fill = as.character(cluster))) + -->
<!--     guides(fill = FALSE) -->
<!-- ``` -->

<!-- ## Hierarchical Clustering -->

<!-- ```{r} -->
<!-- a <- h_clust(blocked_tracts, blocked_data) -->
<!-- compare_hclusts(a, b) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- j <- 10 -->

<!-- m <- prominence_map(tracts, data) +  -->
<!--         guides(fill = FALSE)  -->

<!-- s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = 1) -->
<!-- s -->

<!-- ggsave(paste0('figs/sclust_', city, '.pdf'), width = map_width, height = map_height) -->

<!-- h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = .7) -->
<!-- h -->
<!-- ggsave(paste0('figs/mixed_', city,'.pdf'), width = map_width, height = map_height) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- f <- h_clust_map(m, tracts, data, b, j, size_coef = .7) -->
<!-- f -->
<!-- ggsave(paste0('figs/hclust_', city,'.pdf'), width = map_width, height = map_height) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df <- df %>%  -->
<!--     rbind(data_frame( -->
<!--         city = city,  -->
<!--         k = k,  -->
<!--         j = j, -->
<!--         hclust = cumsum(rev(b$height))[j], -->
<!--         mixed = cumsum(rev(a$height))[j] -->
<!--     )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- loss_curves <- loss_curves %>% -->
<!--     rbind(data_frame( -->
<!--         city = city,  -->
<!--         k = 1:length(a$height), -->
<!--         h_val = rev(b$height)[1:length(a$height)], -->
<!--         m_val = rev(a$height) -->
<!--     )) -->

<!-- MI <- MI %>%  -->
<!--     rbind(data_frame( -->
<!--         city = city,  -->
<!--         I_XY = data %>%  -->
<!--             spread(key = group, value = n) %>%  -->
<!--             select(-tract) %>%  -->
<!--             mutual_info() -->
<!--     )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df %>%  -->
<!--     write_csv('figs/clustering_vals.csv') -->
<!-- loss_curves %>%  -->
<!--     write_csv('figs/loss_curves.csv') -->
<!-- MI %>%  -->
<!--     write_csv('figs/MI.csv') -->

<!-- ``` -->

<!-- ```{r} -->

<!-- gathered <- loss_curves %>%  -->
<!--     gather(key = type, value = val, -city, -k) %>%  -->
<!--     group_by(city, type) %>%  -->
<!--     mutate(val = cumsum(val)) %>%  -->
<!--     ungroup() -->

<!-- to_show <- c('Detroit', 'Chicago', 'Philadelphia') -->

<!-- MI_to_show <- MI %>%  -->
<!--     filter(city %in% to_show) -->

<!-- city <- NULL -->

<!-- labels <- gathered %>%  -->
<!--     filter(type == 'h_val', city %in% to_show) %>%  -->
<!--     group_by(city) %>%  -->
<!--     filter(val == max(val)) -->

<!-- labels$val <- labels$val + c(.04, -.05, -.06) -->
<!-- labels$k <- labels$k + c(-10, -2, 2) -->

<!-- gathered %>%  -->
<!--     filter(type == 'h_val', city %in% to_show) %>%  -->
<!--     ggplot() +  -->
<!--     geom_point(aes(x = k, y = val, color = city)) +  -->
<!--     geom_line(aes(x = k, y = val, color = city, group = city)) +  -->
<!--     geom_hline(aes(yintercept = I_XY, color = city), size = 1, data = MI_to_show, linetype = 'dashed') +  -->
<!--     theme_bw() +  -->
<!--     scale_color_brewer(palette = 'Accent') +  -->
<!--     ylab(expression(italic(I(C,Y)))) +  -->
<!--     xlab(expression(italic(k))) +  -->
<!--     ylim(c(0,NA)) +  -->
<!--     geom_label(aes(x = k, y = val, label = city, fill = city), data = labels) +  -->
<!--     scale_fill_brewer(palette = 'Accent') +  -->
<!--     guides(color = FALSE, fill = FALSE)  -->

<!-- ggsave('figs/loss_curves_h.pdf', height = map_height, width = map_width) -->

<!-- labels <- gathered %>%  -->
<!--     filter(type == 'm_val', city %in% to_show) %>%  -->
<!--     group_by(city) %>%  -->
<!--     filter(val == max(val)) -->

<!-- labels$val <- labels$val + c(.04, -.05, -.06) -->
<!-- labels$k <- labels$k + c(-10, -2, 2) -->

<!-- gathered %>%  -->
<!--     filter(type == 'm_val', city %in% to_show) %>%  -->
<!--      ggplot() +  -->
<!--     geom_point(aes(x = k, y = val, color = city)) +  -->
<!--     geom_line(aes(x = k, y = val, color = city, group = city)) +  -->
<!--     geom_hline(aes(yintercept = I_XY, color = city), size = 1, data = MI_to_show, linetype = 'dashed') +  -->
<!--     theme_bw() +  -->
<!--     scale_color_brewer(palette = 'Accent') +  -->
<!--     ylab(expression(italic(I(C,Y)))) +  -->
<!--     xlab(expression(italic(k))) +  -->
<!--     ylim(c(0,NA)) +  -->
<!--     geom_label(aes(x = k, y = val, label = city, fill = city), data = labels) +  -->
<!--     scale_fill_brewer(palette = 'Accent') +  -->
<!--     guides(color = FALSE, fill = FALSE)  -->

<!-- ggsave('figs/loss_curves_m.pdf', height = map_height, width = map_width) -->

<!-- ``` -->

### 2017-08-23 2_get_states.R

@@ -2,13 +2,14 @@
 library(acs, quietly = TRUE)
 library(tidyverse, quietly = TRUE)
 library(rgdal, quietly = TRUE)
+library(tigris)
+
 
 level <- 'blockgroup'
 
 # -------------------
 
 clean_race <- function(race){
-    
     
     # transform into data frame
     race <- cbind(data.frame(race@geography),

### 2017-08-23 4_compute_metric.R

@@ -2,8 +2,7 @@ library(compx)
 library(rgdal, quietly = TRUE)
 library(stringr, quietly = TRUE)
 library(tidyverse, quietly = TRUE)
-
-source('R/smoother.R')
+library(sf)
 
 # Prep data
 
@@ -15,46 +14,61 @@ demographics <- read_csv('data/demographics.csv') %>%
 tract_list <- demographics %>% 
     group_by(tract) %>% 
     dplyr::summarise(n = sum(n)) %>% 
-    filter(n > 100) %>% 
+    filter(n > 0) %>% 
     select(-n)
 
+if(!dir.create('throughput')){
+    dir.create('throughput')
+}
 if(!dir.exists('throughput/metric')){
     dir.create('throughput/metric')
 }
 
 analysis <- function(city){
     
-    tracts <- readOGR(dsn = paste0('data/cities/',city), 
-                      layer = 'geo',
-                      verbose = FALSE)
+    tracts <- sf::st_read(dsn = paste0('data/cities/', city), 
+                          layer = 'geo', stringsAsFactors = FALSE)
+    
     print(paste(city, ':', nrow(tracts), 'tracts'))
     
+    tracts <- tracts %>% 
+        filter(GEOID %in% tract_list$tract) %>% 
+        mutate(area = st_area(geometry) / 1000^2,
+               area = as.numeric(area)) # area in km
     
-    tracts <- tracts[tracts@data$GEOID %in% tract_list$tract,] 
-    area_df <- tracts@data %>% 
-        dplyr::select(GEOID) %>% 
-        mutate(area = abs(raster::area(tracts) / 1e6))
+    area_df <- tracts %>%  
+        as_data_frame() %>% 
+        select(GEOID, area) 
     
-    data <- demographics %>% 
-        dplyr::filter(tract %in% tracts@data$GEOID)
+    data      <- demographics %>% 
+        dplyr::filter(tract %in% tracts$GEOID) 
     
-    data <- rbf_smoother(data, tracts, sigma = .5)
+    # data <- data %>% 
+    #     mutate(n = ifelse(n < 1, 1, n))
+
+    data <- compx::rbf_smoother(data, tracts, sigma = c(.3, .3)) %>%
+        arrange(tract) # expensive, dense distance matrix
     
     metric_df <- compute_metric(tracts, 
                                 data, 
                                 km = T, 
-                                sigma = 10, 
-                                divergence = 'DKL', 
[…]

### 2017-08-23 5_network_structure.R

@@ -1,12 +1,11 @@
 library(compx, quietly = T)
 library(tidyverse, quietly = T)
-library(rgdal)
-
-source('R/smoother.R')
+library(sf)
+# source('R/hcluster.R')
 
 demographics <- read_csv('data/demographics.csv') %>% 
     dplyr::mutate(tract = as.character(GEOID),
-           tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
+                  tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
     select(-GEOID)
 
 tract_list <- demographics %>% 
@@ -18,45 +17,84 @@ tract_list <- demographics %>%
 if(!dir.exists('throughput/graphs')){
     dir.create('throughput/graphs')
 }
+if(!dir.exists('throughput/hclusts')){
+    dir.create('throughput/hclusts')
+}
 
 analysis <- function(city){
     print(city)
     
     # get the tracts, compute area
-    tracts <- readOGR(dsn = paste0('data/cities/',city), 
-                      layer = 'geo',
-                      verbose = FALSE)
+    tracts <- sf::st_read(dsn = paste0('data/cities/', city), 
+                          layer = 'geo', stringsAsFactors = FALSE)
     
-    tracts <- tracts[tracts@data$GEOID %in% tract_list$tract,] 
-    area_df <- tracts@data %>% 
-        select(GEOID) %>% 
-        mutate(area = map_dbl(1:nrow(tracts), ~ tracts@polygons[[.]]@area))
+    tracts <- tracts %>% 
+        filter(GEOID %in% tract_list$tract) %>% 
+        mutate(area = st_area(geometry) / 1000^2,
+               area = as.numeric(area)) # area in km
+    
+    area_df <- tracts %>% 
+        as_data_frame() %>% 
+        select(GEOID, area) 
     
     # filter the demographics to the tracts
     
-    geoids <- tracts@data$GEOID %>% as.character() %>% unique()
-    
     data <- demographics %>% 
-        filter(tract %in% geoids)
+        dplyr::filter(tract %in% tracts$GEOID) 
     
-    data <- rbf_smoother(data, tracts, sigma = .5)
+    # smoothed_data <- compx::rbf_smoother(data, tracts, sigma = c(.2, .2))
+    smoothed_data <- data
     
     # do the analysis
[…]

### 2017-08-23 R/smoother.R

@@ -1,42 +1,0 @@
-library(compx, quietly = T)
-library(tidyverse, quietly = T)
-library(rgdal)
-library(rgeos)
-
-rbf_smoother <- function(data, tracts, sigma){
-    
-    # compute geographic distances
-    ids <- compx::id_lookup(tracts, key_col = 'GEOID')	
-    coords <- coords_df(tracts, km = T)
-    centroids <- coords %>% 
-        select(-geoid) %>% 
-        as.matrix() 
-    rownames(centroids) <- coords$geoid
-    dists <- dist(centroids) %>% 
-        as.matrix() 
-    
-    # construct the kernel matrix
-    # sigma <- .5
-    K <- exp(-sigma * dists^2)
-    adj <- gRelate(tracts, byid = TRUE, pattern = '****1****')
-    K[adj == 0] <- 0
-    
-    # Convert data to matrix form
-    P_df <- data %>% 
-        spread(key = group, value = n, fill = 0) 
-    P <- P_df %>% select(-tract) %>% 
-        as.matrix()
-    row.names(P) <- P_df$tract
-    P <- P[row.names(K),] # to ensure the orders match
-    
-    # compute the smoother and normalize
-    smoothed_P <- K %*% P
-    smoothed_P <- (smoothed_P / rowSums(smoothed_P)) * rowSums(P)
-    
-    # convert back into data frame form
-    smoothed_df <- smoothed_P %>% 
-        as.data.frame() %>% 
-        rownames_to_column('tract') %>% 
-        gather(key = group, value = n, -tract) %>% 
-        tbl_df() 
-}

### 2017-08-23 8_time.rmd
cleanup

---
title: "8_time"
author: "phil"
date: "7/24/2017"
output: html_document
---

Overall goal: visualization in which we see boundaries shifting in time, and in which the importance of each boundary also changes in a way that we mark visually.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1234)
```

```{r}

library(compx)        # main analysis functions
library(tidyverse)    # for data manipulation and viz
library(igraph)       # for network analysis
library(RColorBrewer) # for plots
library(ggthemes)     # for theme_map
library(units)        # needed for sf
library(sf)           # needed for io and visualization in this vignette
library(scales)       # for ggplot2 viz
library(RSpectra)     # fast eigenvalues, also used within compx
library(tidygraph)
source('R/utils.R')
```

```{r}

tracts_to_use <- detroit_race %>% 
	group_by(tract, t) %>%
	summarise(n = sum(n)) %>%
	filter(n > 100)

tracts <- detroit_tracts %>% 
	filter(GEOID %in% tracts_to_use$tract)

```

```{r}
js_dist <- function(n,m){

	N <- sum(n)
	M <- sum(m)

	p <- n/N
	q <- m/M

	sqrt((N/(M+N))*DKL(p, p+q) + (M/(M+N))*DKL(q, p+q))
}

```

# Changing Boundaries

```{r}

race_temporal <- detroit_race %>%
	filter(t %in% c(1990, 2000, 2010))

g <- construct_information_graph(tracts, race_temporal, divergence = js_dist)

```

```{r}
k     <- 7
sigma <- 100
g     <- g %>% spectral_cluster(sigma = sigma, k = k, nreps = 1000)
```

```{r}

nodes    <- data_frame(geoid = V(g)$geoid, t = V(g)$t, cluster = V(g)$cluster)

pal <- scales::brewer_pal(palette = 'Set2')
pal <- colorRampPalette(pal(k))
pal <- pal(k)

tracts %>% 
	left_join(nodes, by = c('GEOID' = 'geoid')) %>% 
	ggplot() + 
	geom_sf(aes(fill = as.character(cluster)), size = .1) +
	ggthemes::theme_map() + 
	scale_fill_manual(values = pal) + 
	theme(legend.position = 'bottom',
		  legend.justification = c(.5, .0),
		  plot.margin=unit(c(0,0,0,0),"mm")) + 
	guides(fill = guide_legend(title = 'Cluster', nrow = 1)) + 
	facet_wrap(~t)
```

```{r}
clusters <- data_frame(GEOID = V(g)$geoid, t = V(g)$t, cluster = V(g)$cluster)

blocked_tracts <- tracts %>% 
    left_join(clusters) %>% 
    select(cluster, t) %>% 
    aggregate(by = list(cluster = .$cluster, t = .$t), FUN = identity) %>% 
    mutate(GEOID = as.character(cluster))

blocked_data <- race_temporal %>% 
    left_join(clusters, by = c('tract' = 'GEOID', 't' = 't')) %>% 
    group_by(cluster, group, t) %>% 
    summarise(n = sum(n)) %>% 
    ungroup() %>% 
    mutate(tract = as.character(cluster))
```

```{r}
h <- construct_information_graph(blocked_tracts, blocked_data, js_dist)
a <- h_clust(h)
```

```{r}
get_MI <- function(k, the_year){
    v <- a %>% cutree(k)

    cluster_df <- data_frame(small_cluster = names(v), h_cluster = as.character(v)) %>% 
        tidyr::separate(small_cluster, into = c('cluster', 't'), sep = '_')

    cluster_lookup <- data_frame(name = V(g)$name, cluster = V(g)$cluster) %>% 
        separate(name, into = c('geoid', 't'), sep = '_') %>% 
        mutate(cluster = as.character(cluster)) %>% 
        left_join(cluster_df, by = c('t' = 't', 'cluster' = 'cluster')) %>% 
        mutate(t = as.integer(t))

    clustered_data <- race_temporal %>% 
        left_join(cluster_lookup, by = c('tract' = 'geoid', 't' = 't'))

    clustered_data %>% 
        filter(t == the_year) %>% 
        select(-tract, cluster) %>% 
        group_by(group, h_cluster) %>% 
        summarise(n = sum(n)) %>% 
        filter(!is.na(h_cluster)) %>% 
        spread(key = h_cluster, value = n) %>% 
        ungroup() %>% 
        select(-group) %>% 
        mutual_info()
}
```

```{r}
MIs <- expand.grid(t = c(1990, 2000, 2010), k = 2:7) %>% 
    dplyr::as_data_frame() %>% 
    mutate(MI = map2_dbl(k, t, get_MI))

vals_1 <- MIs %>% 
    group_by(t) %>% 
    mutate(val = ifelse(k == 2, MI, MI - lag(MI))) %>% 
    ungroup() %>% 
    group_by(k) %>% 
    mutate(sign = val - lag(val)) %>% 
    ungroup()
```

```{r}

v <- a %>% cutree(k)

    cluster_df <- data_frame(small_cluster = names(v), h_cluster = as.character(v)) %>% 
        tidyr::separate(small_cluster, into = c('cluster', 't'), sep = '_')

    cluster_lookup <- data_frame(name = V(g)$name, cluster = V(g)$cluster) %>% 
        separate(name, into = c('geoid', 't'), sep = '_') %>% 
        mutate(cluster = as.character(cluster)) %>% 
        left_join(cluster_df, by = c('t' = 't', 'cluster' = 'cluster')) %>% 
        mutate(t = as.integer(t))

    clustered_data <- race_temporal %>% 
        left_join(cluster_lookup, by = c('tract' = 'geoid', 't' = 't'))

```

```{r}

boundaries_map <- function(the_year){

all <- tracts %>% 
        aggregate(by = list(rep(1, nrow(.))), FUN = identity)

all_bd <- st_boundary(all)

m <- detroit_race %>% 
    filter(t == the_year) %>% 
    prominence_map(tracts, .)

val <- vals_1 %>% 
    filter(t == the_year)

for(j in rev(2:k)){

    v <- a %>% cutree(j)

    cluster_df <- data_frame(small_cluster = names(v), h_cluster = as.character(v)) %>% 
        tidyr::separate(small_cluster, into = c('cluster', 't'), sep = '_')

    cluster_lookup <- data_frame(name = V(g)$name, cluster = V(g)$cluster) %>% 
        separate(name, into = c('geoid', 't'), sep = '_') %>% 
        mutate(cluster = as.character(cluster)) %>% 
        left_join(cluster_df, by = c('t' = 't', 'cluster' = 'cluster')) %>% 
        mutate(t = as.integer(t))

    clustered_data <- race_temporal %>% 
        left_join(cluster_lookup, by = c('tract' = 'geoid', 't' = 't'))

    plot_tracts <- tracts %>% 
        left_join(cluster_lookup, by = c('GEOID' = 'geoid')) %>% 
        aggregate(by = list(cluster = .$h_cluster, t = .$t), FUN = identity) %>% 
        filter(t == the_year)

        bd <- st_boundary(plot_tracts) %>% st_difference(all_bd)

        size <- val %>%
            filter(k == j) %>% 
            select(val) %>% 
            unlist()

        sign <- val %>% 
            filter(k == j) %>% 
            select(sign) %>% 
            unlist()

        if(the_year == 1990){
            sign = 1
        }
        if(sign > 0){
            color = 'black'
        }else{
            color = 'white'
        }

        m <- m + 
            geom_sf(size = exp(4 * size), data = bd, alpha = 0, color = color) 
}
m
}

```

```{r}
boundaries_map(1990) 
ggsave('figs/detroit_boundaries_1990.pdf', width = 4, height = 3)
```

```{r}
boundaries_map(2000) + guides(fill = FALSE)
ggsave('figs/detroit_boundaries_2000.pdf', width = 4, height = 3)
```

```{r}
boundaries_map(2010) + guides(fill = FALSE)
ggsave('figs/detroit_boundaries_2010.pdf', width = 4, height = 3)
```

```{r}
vals_1 %>% 
    ggplot() + 
    aes(x = t, y = val, color = as.factor(k), group = k) + 
    geom_point() + 
    geom_line()

```

```{r}
vals_1 %>%
    ggplot() + 
    aes(x = k, y = val, group = t, color = as.factor(t)) + 
    geom_point() + 
    geom_line() 
```

```{r}
I_CY <- vals_1 %>% 
    group_by(t) %>% 
    summarise(I_CY = sum(val))
```

```{r}
# Mutual information values
detroit_race %>% 
    spread(key = group, value = n) %>% 
    select(-tract) %>% 
    nest(-t) %>% 
    mutate(I_XY = map_dbl(data, mutual_info)) %>% 
    left_join(I_CY)
```

```{r}
detroit_race %>% 
    left_join(cluster_lookup, by = c('tract' = 'geoid', 't' = 't')) %>% 
    filter(!is.na(cluster)) %>% 
    group_by(t, group, cluster) %>% 
    summarise(n = sum(n)) %>% 
    ungroup() %>% 
    group_by(cluster,t) %>% 
    mutate(perc = n / sum(n)) %>% 
    ungroup() %>%
    mutate(cluster = chartr("123456789", "ABCDEFGHI", cluster)) %>% 
    ggplot() + 
    aes(x = t, y = perc, color = group) + 
    geom_line() + 
    geom_point() + 
    facet_wrap(~cluster, nrow = 1) + 
    theme_bw() + 
    scale_color_brewer(palette = 'Set1') + 
    guides(color = FALSE) + 
    scale_x_continuous(breaks = c(1990, 2000, 2010)) + 
    theme(axis.text.x = element_text(angle = 90),
          axis.title.x=element_blank()) + 
    scale_y_continuous(labels=percent) + 
    ylab('Population Share')

ggsave('figs/detroit_breakdown.pdf', width = 10, height = 2)
```

### 2017-08-23 1_get_states.R

@@ -1,70 +1,0 @@
-# library(sp, quietly = TRUE)
-library(acs, quietly = TRUE)
-library(dplyr, quietly = TRUE)
-library(tidyr, quietly = TRUE)
-library(stringr, quietly = TRUE)
-library(readr)
-
-# get list of states
-
-cities <- readr::read_csv('assumptions/cities.csv')
-cities$state <- word(cities$state, -1)
-
-if(!dir.exists('data')){
-	dir.create('data')
-}
-
-if(!dir.exists('data/states')){
-	dir.create('data/states')
-}
-
-# Define data-getter
-
-f <- function(state){
-	counties <- acs::acs.fetch(geography=acs::geo.make(state=state, county="*"), 
-				   endyear = 2014,
-				   table.number="B01003")
-	counties <- as.numeric(geography(counties)[[3]])
-	
-	race <- acs::acs.fetch(endyear = 2014,
-						   span = 5,
-						   geography = acs::geo.make(state = state, county = counties, tract = '*', block.group = '*'),
-						   table.number = 'B03002',
-						   col.names = "pretty")
-	
-	race <- cbind(data.frame(race@geography),
-				  data.frame(race@estimate)) %>%
-		tbl_df() 
-	
-	race_lookup <- read_csv('assumptions/race_lookup.csv')
-	
-	race <- race %>% 
-		gather(key = 'original_race', value = 'n', -(NAME:blockgroup)) %>% 
-		left_join(race_lookup, by = c('original_race' = 'original')) %>% 
-		group_by(NAME, state, county, tract, blockgroup, aggregated) %>% 
-		summarise(n = sum(n)) %>% 
-		ungroup() %>% 
-		filter(!is.na(aggregated)) %>% 
-		spread(key = aggregated, value = n, fill = 0)
-	
-	race <- race %>% 
-		mutate(GEOID = paste0(str_pad(state, 2, 'left', pad = '0'),
-							  str_pad(county, 3, 'left', pad = '0'),
-							  str_pad(tract, 6, 'left', pad = '0'),
-							  str_pad(blockgroup, 1, 'left', pad = '0')))
-	
-	tracts <- tigris::block_groups(state = state, county = counties, cb=TRUE)
-	tracts@data$area <- tracts@data$ALAND / 1000^2
-	
-	tracts <- tigris::geo_join(tracts, race, "GEOID", "GEOID")
[…]

### 2017-08-23 2_get_states.R

@@ -1,0 +1,125 @@
+# library(sp, quietly = TRUE)
+library(acs, quietly = TRUE)
+library(tidyverse, quietly = TRUE)
+library(rgdal, quietly = TRUE)
+library(tigris)
+
+
+level <- 'blockgroup'
+
+# -------------------
+
+clean_race <- function(race){
+    
+    # transform into data frame
+    race <- cbind(data.frame(race@geography),
+                  data.frame(race@estimate)) %>%
+        tbl_df() 
+    
+    # aggregate into 5 categories
+    race_lookup <- read_csv('assumptions/race_lookup.csv')
+    
+    au_vars <- c('state', 'county', 'tract', 'blockgroup') %>% 
+        intersect(names(race))
+    
+    gather_cols <- names(race) %>% 
+        setdiff(au_vars) %>% 
+        setdiff('NAME')
+    
+    race <- race %>% 
+        gather_(key_col = 'original_race', value_col = 'n', gather_cols = gather_cols) %>% 
+        left_join(race_lookup, by = c('original_race' = 'original')) %>% 
+        dplyr::group_by_(.dots = au_vars %>% append(c('aggregated'))) %>% 
+        dplyr::summarise(n = sum(n)) %>% 
+        ungroup() %>%  
+        filter(!is.na(aggregated)) %>% 
+        spread(key = aggregated, value = n, fill = 0)
+    
+    # clean up the areal unit designations to give a single GEOID
+    
+    lengths <- list(state      = 2, 
+                    county     = 3, 
+                    tract      = 6, 
+                    blockgroup = 1)
+    
+    lengths <- lengths[au_vars]
+    
+    for(colname in names(lengths)){
+        race[[colname]] <- str_pad(race[[colname]], 
+                                   lengths[[colname]], 
+                                   'left', 
+                                   pad = 0)
+    }
+    
+    race <- race %>% 
+        tidyr::unite_('GEOID', au_vars, sep = '') %>% 
+        gather(key = group, value = n, -GEOID)
+    race
+}
+
+
+getter <- function(state){
+    print(state)
+    counties <- fips.county %>% 
+        filter(State == state) %>% 
+        select(County.ANSI) %>% 
+        unlist()
+
+    if(level == 'blockgroup'){
+        tracts <- tigris::block_groups(state = state, county = counties, cb=TRUE)
+        geo    <- geo.make(state = state, 
+                           county = counties, 
+                           tract = '*', 
+                           block.group = '*')
[…]

### 2017-08-23 2_make_cities.R

@@ -1,65 +1,0 @@
-library(sp, quietly = TRUE)
-library(compx, quietly = TRUE)
-library(plyr, quietly = TRUE)
-library(dplyr, quietly = TRUE)
-library(readr, quietly = TRUE)
-library(tidyr, quietly = TRUE)
-library(magrittr, quietly = TRUE)
-library(grid, quietly = TRUE)
-library(rgdal, quietly = TRUE)
-library(stringr, quietly = TRUE)
-
-city_df <- read_csv('assumptions/cities.csv')
-cities <- unique(city_df$name)
-cities <- cities[!is.element(cities, list.dirs('cities', full.names = FALSE))]
-
-states <- city_df$state %>% 
-	str_split('--') %>% 
-	# do.call(c,.) %>% 
-	unique()
-
-shapes <- list()
-for(state in states){
-	print(paste('Loading', state))
-	shapes[state] <- readOGR(dsn = paste0('data/states/',state), 
-	                         layer = 'geo', 
-	                         verbose = FALSE)
-}
-
-if(!dir.exists('data/cities')){
-	dir.create('data/cities')
-}
-
-
-for(city in cities){
-	print(city)
-	city_info <- city_df %>%
-		filter(name == city) 
-	
-	city_components <- list()
-	for(i in 1:nrow(city_info)){
-		state <- city_info$state[i]
-		county <- city_info$county[i]
-		component <- shapes[[state]]
-		component <- component[as.integer(as.character(component@data$COUNTYFP)) == county,]
-		city_components[paste(state, county, sep = '_')] <- spChFIDs(component,
-																	  as.character(paste(state, 
-																	                     county, 
-																	                     rownames(as(component,
-																	                                 "data.frame")))))
-	}
-	city_polys <- do.call(rbind, city_components)
-	city_polys@data <- city_polys@data %>%
-	  mutate(total = Asian + Black + Hispanic + Other + White)
-	city_polys <- city_polys[city_polys@data$total != 0,]
-	if(city == 'Philadelphia'){
-	  exclude <- c('421010098023','421010098022', '421010098021', '421010098011',  '421010098012')
-	  city_polys <- city_polys[!(city_polys@data$GEOID %in% exclude),]   
-	}
-	
-	tryCatch(writeOGR(city_polys, paste0("data/cities/", city),
-	                  'geo', 
-	                  driver = 'ESRI Shapefile', 
-	                  morphToESRI = TRUE),
-			 error = function(e) NULL)
-}

### 2017-08-23 3_local_info.R

@@ -1,94 +1,0 @@
-
-library(sp, quietly = TRUE)
-library(compx, quietly = TRUE)
-library(plyr, quietly = TRUE)
-library(dplyr, quietly = TRUE)
-library(readr, quietly = TRUE)
-library(tidyr, quietly = TRUE)
-library(magrittr, quietly = TRUE)
-library(grid, quietly = TRUE)
-library(rgdal, quietly = TRUE)
-library(data.table, quietly = TRUE)
-
-races <- c('Black', 'Hispanic', 'Asian', 'White', 'Other')
-
-cities <- list.files('data/cities')
-resolution <- 1
-
-cities <- list.files('data/cities')
-# cities <- c('Boston')
-
-if(!dir.exists('throughput/grids')){
-	dir.create('throughput/grids')
-}
-
-if(!dir.exists('throughput/grid_tracts')){
-	dir.create('throughput/grid_tracts')
-}
-
-
-# MAIN COMPUTATION ---------------
-
-# Cache logic
-
-if(file.exists('throughput/info_cache.csv')){
-	cache <- read_csv('throughput/info_cache.csv')
-	cities <-  cities[!is.element(cities, cache$city)]
-}else{
-	cache <- data.frame(
-		city = character(),
-		area = numeric(),
-		population = numeric(),
-		density = numeric(),
-		I_XY = numeric(),
-		J_XY = numeric())
-}
-
-for(city in cities){
-	tracts <- readOGR(dsn = paste0('data/cities/',city),
-	                  layer = 'geo',
-	                  verbose = FALSE)
-	grid_dir        <- paste0('throughput/grids/',city)
-	grid_tract_path <- paste0('throughput/grid_tracts/', city, '.csv')
-
-	if(file.exists(grid_tract_path) & dir.exists(grid_dir)){
-	  grid_tract <- read_csv(grid_tract_path)
-	  grid <- readOGR(dsn = paste0('throughput/grids/',city),
-	                  layer = 'grid',
-	                  verbose = FALSE)
-
-	  print(city)
-	  analysis <- info_analysis(tracts, 
-	                            columns = races, 
-	                            resolution = resolution,
-	                            grid_polys = grid,
-	                            grid_tract = grid_tract)
-	}else{
-	  
-	  tryCatch(
-	    {print(city)
-	    analysis <- info_analysis(tracts, 
-	                              columns = races, 
-	                              resolution = resolution)
[…]

### 2017-08-23 4_cluster.R

@@ -1,54 +1,0 @@
-library(dplyr, quietly = TRUE)
-library(compx, quietly = TRUE)
-library(rgdal, quietly = TRUE)
-library(rgeos, quietly = TRUE)
-library(ggplot2, quietly = TRUE)
-library(Matrix, quietly = TRUE)
-library(maptools, quietly = TRUE)
-library(readr, quietly = TRUE)
-library(ggrepel, quietly = TRUE)
-library(caTools, quietly = TRUE)
-
-columns <- c('White', 'Black', 'Asian', 'Hispanic', 'Other')
-
-# --------------------
-
-summary <- read_csv('throughput/info_cache.csv')
-cities <- summary$city
-
-if(file.exists('throughput/loss_curves.csv')){
-	cache <- read_csv('throughput/loss_curves.csv')
-	cache_binary <- read_csv('throughput/loss_curves_binary.csv')
-	cities <- cities[!is.element(cities, cache$city)]
-} else {
-	cache <- data.frame('I_XY' = numeric(), 'nclust' = numeric(), 'city' = character())
-	cache_binary <- data.frame('I_XY' = numeric(), 'nclust' = numeric(), 'city' = character())
-}
-
-if(!dir.exists('throughput/clusterings')){
-	dir.create('throughput/clusterings')
-}
-
-for(city in cities){
-	print(city)
-	spdf <- readOGR(dsn = paste0('data/cities/', city), layer = 'geo',verbose = FALSE)
-	spdf <- spdf[sum(spdf@data[,columns]) != 0,]
-	df   <- spdf@data[,columns]
-	row.names(df) <- 1:nrow(df)
-	constraint    <- gRelate(spdf, byid = TRUE, pattern = '****1****')
-	a     <- info_clust(df, constraint)
-	saveRDS(a, file = paste0('throughput/clusterings/',city))
-	added <- data.frame('I_XY' = c(0,a$height), 'nclust' = 1:nrow(spdf), 'city' = city)
-	cache <- rbind(cache, added)
-	write_csv(cache,'throughput/loss_curves.csv')
-	
-	df <- df %>% 
-	  mutate(non_black = Asian + Hispanic + Other + White ) %>% 
-	  select(Black, non_black)
-	
-	b <- info_clust(df, constraint)
-	saveRDS(b, file = paste0('throughput/clusterings/',city, '_binary'))
[…]

### 2017-08-23 4_compute_metric.R

@@ -1,0 +1,120 @@
+library(compx)
+library(rgdal, quietly = TRUE)
+library(stringr, quietly = TRUE)
+library(tidyverse, quietly = TRUE)
+library(sf)
+
+# Prep data
+
+demographics <- read_csv('data/demographics.csv') %>% 
+    dplyr::mutate(tract = as.character(GEOID),
+           tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
+    select(-GEOID) 
+
+tract_list <- demographics %>% 
+    group_by(tract) %>% 
+    dplyr::summarise(n = sum(n)) %>% 
+    filter(n > 0) %>% 
+    select(-n)
+
+if(!dir.create('throughput')){
+    dir.create('throughput')
+}
+if(!dir.exists('throughput/metric')){
+    dir.create('throughput/metric')
+}
+
+analysis <- function(city){
+    
+    tracts <- sf::st_read(dsn = paste0('data/cities/', city), 
+                          layer = 'geo', stringsAsFactors = FALSE)
+    
+    print(paste(city, ':', nrow(tracts), 'tracts'))
+    
+    tracts <- tracts %>% 
+        filter(GEOID %in% tract_list$tract) %>% 
+        mutate(area = st_area(geometry) / 1000^2,
+               area = as.numeric(area)) # area in km
+    
+    area_df <- tracts %>%  
+        as_data_frame() %>% 
+        select(GEOID, area) 
+    
+    data      <- demographics %>% 
+        dplyr::filter(tract %in% tracts$GEOID) 
+    
+    # data <- data %>% 
+    #     mutate(n = ifelse(n < 1, 1, n))
+
+    data <- compx::rbf_smoother(data, tracts, sigma = c(.3, .3)) %>%
+        arrange(tract) # expensive, dense distance matrix
+    
+    metric_df <- compute_metric(tracts, 
+                                data, 
+                                km = T, 
+                                r_sigma = 10, 
+                                s_sigma = 10,
+                                smooth = T,
+                                hessian = DKL_) 
+    
+    metric_df <- metric_df %>% 
+        mutate(vol = map_dbl(g, . %>% det() %>% abs() %>% sqrt()),
+               tr  = map_dbl(g, . %>% diag() %>% sum())) 
+    
+    H_df <- data %>% 
+        group_by(tract) %>% 
+        mutate(p = n / sum(n)) %>% 
+        summarise(H = - sum(p * log(p)))
+    
+    metric_df <- metric_df %>% 
[…]

### 2017-08-23 5_make_figs.R

@@ -1,2 +1,0 @@
-Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio-server/bin/pandoc") 
-rmarkdown::render(input = "figs.Rmd")

### 2017-08-23 5_network_structure.R

@@ -1,0 +1,100 @@
+library(compx, quietly = T)
+library(tidyverse, quietly = T)
+library(sf)
+# source('R/hcluster.R')
+
+demographics <- read_csv('data/demographics.csv') %>% 
+    dplyr::mutate(tract = as.character(GEOID),
+                  tract = stringr::str_pad(tract, 12, side = 'left', pad = '0')) %>% 
+    select(-GEOID)
+
+tract_list <- demographics %>% 
+    group_by(tract) %>% 
+    dplyr::summarise(n = sum(n)) %>% 
+    filter(n > 0) %>% 
+    select(-n)
+
+if(!dir.exists('throughput/graphs')){
+    dir.create('throughput/graphs')
+}
+if(!dir.exists('throughput/hclusts')){
+    dir.create('throughput/hclusts')
+}
+
+analysis <- function(city){
+    print(city)
+    
+    # get the tracts, compute area
+    tracts <- sf::st_read(dsn = paste0('data/cities/', city), 
+                          layer = 'geo', stringsAsFactors = FALSE)
+    
+    tracts <- tracts %>% 
+        filter(GEOID %in% tract_list$tract) %>% 
+        mutate(area = st_area(geometry) / 1000^2,
+               area = as.numeric(area)) # area in km
+    
+    area_df <- tracts %>% 
+        as_data_frame() %>% 
+        select(GEOID, area) 
+    
+    # filter the demographics to the tracts
+    
+    data <- demographics %>% 
+        dplyr::filter(tract %in% tracts$GEOID) 
+    
+    # smoothed_data <- compx::rbf_smoother(data, tracts, sigma = c(.2, .2))
+    smoothed_data <- data
+    
+    # do the analysis
+    divergence <- function(n,m){
+        p <- n / sum(n)
+        q <- m / sum(m)
+        sqrt(.5 * DKL(p, p+q) + .5*DKL(q, p+q))
+    }
+    
+    # spectral
+    g <- construct_information_graph(tracts, smoothed_data, divergence = divergence)
+    A <- affinity_matrix(g, sigma = 50)
+    L <- normalized_laplacian(A)
+    eig <- RSpectra::eigs(L, 100, which = 'LM', sigma = 1e-10)         
+    
+    g %>% saveRDS(file = paste0('throughput/graphs/', city, '.RDS'))
+    
+    # return the spectra, might add more later
[…]

### 2017-08-23 R/utils.R

@@ -1,0 +1,298 @@
+
+## @knitr utils
+
+library(tidyverse)
+library(compx)
+library(sf)
+library(RColorBrewer)
+library(ggthemes)
+# 
+# h_clust <- function(input, data){
+#     
+#     if('sf' %in% class(input)){
+#         tract_lookup <- id_lookup(input)
+#         adj <- st_relate(input, pattern = '****1****', sparse = TRUE)
+#         
+#         data_lookup <- data %>% 
+#             nest(-tract) %>% 
+#             mutate(n = map(data, ~.$n)) %>% 
+#             select(-data) %>% 
+#             left_join(tract_lookup, by = c('tract' = 'geoid'))
+#         
+#         adj <- 1:length(adj) %>%
+#             map(~data_frame(from = as.character(.),
+#                             to = as.character(adj[[.]]))) %>%
+#             reduce(rbind) %>%
+#             filter(from != to) %>%
+#             left_join(data_lookup, by = c('from' = 'row')) %>%
+#             left_join(data_lookup, by = c('to' = 'row'), suffix = c('_from', '_to')) %>% 
+#             rename(from_n = n_from, to_n = n_to)
+#         n <- nrow(input)
+#         names <- input$cluster
+#         
+#         M <- data %>% 
+#             group_by(group) %>% 
+#             summarise(n = sum(n)) %>% 
+#             select(n) %>% 
+#             unlist()
+#         
+#     }else{
+#         adj <- input %>% 
+#             as.directed(mode = 'mutual') %>% 
+#             as_long_data_frame() %>% 
+#             dplyr::as_data_frame()
+#         
+#         M <- V(input)$n %>% reduce(`+`)
+#         n <- length(V(input))
+#         names <- V(input)$name
+#         }
+#     
+#     divergence <- function(n,m){
+#         p <- n / sum(n)
+#         q <- m / sum(m)
+#         p_bar <- sum(n) / sum(m + n) * p + sum(m) / sum(m + n) * q
+#         r <- M / sum(M)
+#         sum(n) / sum(M) * DKL(p, r) +
+#             sum(m) / sum(M)*DKL(q, r) -
+#             sum(m + n) / sum(M) * DKL(p_bar, r)
+#     }  
+#     
+#     adj <- adj %>%
+#         mutate(dist = map2_dbl(from_n, to_n, divergence)) %>%
+#         mutate(from = as.integer(from),
[…]

### 2017-08-23 make_figs.R

@@ -1,2 +1,0 @@
-Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio-server/bin/pandoc") 
-rmarkdown::render(input = "5_figs.Rmd")

### 2018-03-28 7_cluster_figs.Rmd
final cluster figs

@@
 
-knitr::read_chunk('R/utils.R')
+# knitr::read_chunk('R/utils.R')
 
-# source('R/utils.R')
-```
+source('R/utils.R')
 
-```{r utils}
 ```
@@
 
-s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = 1)
+s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = .75)
 s
@@
 
-h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = 1)
+h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = .75)
 h
@@
 ```{r}
-f <- h_clust_map(m, tracts, data, b, j, size_coef = 1)
+f <- h_clust_map(m, tracts, data, b, j, size_coef = .75)
 f
@@
 b <- h_clust(tracts, data)
+
 ```
@@
 ```{r}
+
 g <- construct_information_graph(tracts = tracts, 
@@
 ```{r}
+set.seed(1250)
 k     <- 36
@@
 
-h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = 1)
+h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = .75)
 
@@
 tracts  <- in_data$tracts
+
+set.seed(1234)
 ```
@@
 
-<!-- # Dallas -->
+# Loss Curves
 
-<!-- ```{r} -->
-<!-- city    <- 'Dallas' -->
-<!-- in_data <- read_data(city, all_data) -->
-<!-- data    <- in_data$data -->
-<!-- tracts  <- in_data$tracts -->
-<!-- ``` -->
+```{r}
 
-<!-- ## Filter out the small island -->
-<!-- ```{r} -->
-<!-- g <- construct_information_graph(tracts = tracts,  -->
-<!-- 								 data = data,  -->
-<!-- 								 divergence = euc) -->
+labels <- data_frame(
+    city = c('Detroit', 'Philadelphia', 'Chicago'),
+    k    = c(20, 12, 32),
+    I_XY   = c(.52, .3, .4)
+)
 
-<!-- comps <- igraph::components(g) -->
+loss_curves %>% 
+    group_by(city) %>% 
+    mutate(val = cumsum(h_val)) %>% 
+    ggplot() + 
+    geom_line(aes(x = k, y = val, color = city)) + 
+    geom_point(aes(x = k, y = val, color = city)) + 
+    theme_bw() + 
+    scale_color_brewer(palette = 'Accent') + 
+    scale_fill_brewer(palette = 'Accent') + 
+    guides(color = FALSE, fill = FALSE) + 
+    xlab(expression(italic(k))) + 
+    ylab(expression(italic(I(C[k], Y)))) + 
+    geom_hline(aes(yintercept = I_XY, color = city), data = MI, linetype = 'dashed', size = 1) + 
+    geom_label(aes(x = k, y = I_XY, label = city, fill = city), data = labels) + 
+    scale_y_continuous(lim = c(0, NA))
 
-<!-- largest_cluster <- which.max(table(comps$membership)) -->
-<!-- exclude <- which(comps$membership != largest_cluster) %>% names() -->
+ggsave('figs/loss_curves_h.pdf', width = map_width, height = map_height)
+```
 
-<!-- tracts <- tracts %>%  -->
-<!--     filter(!(GEOID %in% exclude)) -->
+```{r}
 
-<!-- data <- data %>% -->
-<!--     filter(tract %in% tracts$GEOID) -->
-<!-- ``` -->
+labels <- data_frame(
+    city = c('Detroit', 'Philadelphia', 'Chicago'),
+    k    = c(20, 12, 32),
+    I_XY   = c(.49, .25, .34)
+)
 
-<!-- ```{r} -->
-<!-- b <- h_clust(tracts, data) -->
-<!-- ``` -->
+loss_curves %>% 
+    group_by(city) %>% 
+    mutate(val = cumsum(m_val)) %>% 
+    ggplot() + 
+    geom_line(aes(x = k, y = val, color = city)) + 
+    geom_point(aes(x = k, y = val, color = city)) + 
+    theme_bw() + 
+    scale_color_brewer(palette = 'Accent') + 
+    scale_fill_brewer(palette = 'Accent') + 
+    guides(color = FALSE, fill = FALSE) + 
+    xlab(expression(italic(k))) + 
+    ylab(expression(italic(I(C[k], Y)))) + 
+    geom_hline(aes(yintercept = I_XY, color = city), data = MI, linetype = 'dashed', size = 1) + 
+    geom_label(aes(x = k, y = I_XY, label = city, fill = city), data = labels) +
+    scale_y_continuous(lim = c(0, NA))
 
-<!-- ```{r} -->
-<!-- g <- construct_information_graph(tracts = tracts,  -->
-<!-- 								 data = data,  -->
-<!-- 								 divergence = js_dist) -->
-
-<!-- sigma <- 30 -->
-<!-- plot_spectrum(g, sigma) -->
-<!-- ggsave(paste0('figs/spectrum_', city, '.pdf')) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- k     <- 38 -->
-<!-- g     <- spectral_cluster(g, sigma = sigma, k = k, nreps = 1000) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- input <- group_data(tracts, data, g)  -->
-<!-- blocked_tracts <- input$blocked_tracts -->
-<!-- blocked_data <- input$blocked_data -->
-
-<!-- blocked_tracts %>%  -->
-<!--     ggplot() +  -->
-<!--     geom_sf(aes(fill = as.character(cluster))) + -->
-<!--     guides(fill = FALSE) -->
-<!-- ``` -->
-
-<!-- ## Hierarchical Clustering -->
-
-<!-- ```{r} -->
-<!-- a <- h_clust(blocked_tracts, blocked_data) -->
-<!-- compare_hclusts(a, b) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- j <- 10 -->
-
-<!-- m <- prominence_map(tracts, data) +  -->
-<!--         guides(fill = FALSE)  -->
-
-<!-- s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = 1) -->
-<!-- s -->
-
-<!-- ggsave(paste0('figs/sclust_', city, '.pdf'), width = map_width, height = map_height) -->
-
-<!-- h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = .7) -->
-<!-- h -->
-<!-- ggsave(paste0('figs/mixed_', city,'.pdf'), width = map_width, height = map_height) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- f <- h_clust_map(m, tracts, data, b, j, size_coef = .7) -->
-<!-- f -->
-<!-- ggsave(paste0('figs/hclust_', city,'.pdf'), width = map_width, height = map_height) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- df <- df %>%  -->
-<!--     rbind(data_frame( -->
-<!--         city = city,  -->
-<!--         k = k,  -->
-<!--         j = j, -->
-<!--         hclust = cumsum(rev(b$height))[j], -->
-<!--         mixed = cumsum(rev(a$height))[j] -->
-<!--     )) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- loss_curves <- loss_curves %>% -->
-<!--     rbind(data_frame( -->
-<!--         city = city,  -->
-<!--         k = 1:length(a$height), -->
-<!--         h_val = rev(b$height)[1:length(a$height)], -->
-<!--         m_val = rev(a$height) -->
-<!--     )) -->
-
-<!-- MI <- MI %>%  -->
-<!--     rbind(data_frame( -->
-<!--         city = city,  -->
-<!--         I_XY = data %>%  -->
-<!--             spread(key = group, value = n) %>%  -->
-<!--             select(-tract) %>%  -->
-<!--             mutual_info() -->
-<!--     )) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- df %>%  -->
-<!--     write_csv('figs/clustering_vals.csv') -->
-<!-- loss_curves %>%  -->
-<!--     write_csv('figs/loss_curves.csv') -->
-<!-- MI %>%  -->
-<!--     write_csv('figs/MI.csv') -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- gathered <- loss_curves %>%  -->
-<!--     gather(key = type, value = val, -city, -k) %>%  -->
-<!--     group_by(city, type) %>%  -->
-<!--     mutate(val = cumsum(val)) %>%  -->
-<!--     ungroup() -->
-
-<!-- to_show <- c('Detroit', 'Chicago', 'Philadelphia') -->
-
-<!-- MI_to_show <- MI %>%  -->
-<!--     filter(city %in% to_show) -->
-
-<!-- city <- NULL -->
-
-<!-- labels <- gathered %>%  -->
-<!--     filter(type == 'h_val', city %in% to_show) %>%  -->
-<!--     group_by(city) %>%  -->
-<!--     filter(val == max(val)) -->
-
-<!-- labels$val <- labels$val + c(.04, -.05, -.06) -->
-<!-- labels$k <- labels$k + c(-10, -2, 2) -->
-
-<!-- gathered %>%  -->
-<!--     filter(type == 'h_val', city %in% to_show) %>%  -->
-<!--     ggplot() +  -->
-<!--     geom_point(aes(x = k, y = val, color = city)) +  -->
-<!--     geom_line(aes(x = k, y = val, color = city, group = city)) +  -->
-<!--     geom_hline(aes(yintercept = I_XY, color = city), size = 1, data = MI_to_show, linetype = 'dashed') +  -->
-<!--     theme_bw() +  -->
-<!--     scale_color_brewer(palette = 'Accent') +  -->
-<!--     ylab(expression(italic(I(C,Y)))) +  -->
-<!--     xlab(expression(italic(k))) +  -->
-<!--     ylim(c(0,NA)) +  -->
-<!--     geom_label(aes(x = k, y = val, label = city, fill = city), data = labels) +  -->
-<!--     scale_fill_brewer(palette = 'Accent') +  -->
-<!--     guides(color = FALSE, fill = FALSE)  -->
-
-<!-- ggsave('figs/loss_curves_h.pdf', height = map_height, width = map_width) -->
-
-<!-- labels <- gathered %>%  -->
-<!--     filter(type == 'm_val', city %in% to_show) %>%  -->
-<!--     group_by(city) %>%  -->
-<!--     filter(val == max(val)) -->
-
-<!-- labels$val <- labels$val + c(.04, -.05, -.06) -->
-<!-- labels$k <- labels$k + c(-10, -2, 2) -->
-
-<!-- gathered %>%  -->
-<!--     filter(type == 'm_val', city %in% to_show) %>%  -->
-<!--      ggplot() +  -->
-<!--     geom_point(aes(x = k, y = val, color = city)) +  -->
-<!--     geom_line(aes(x = k, y = val, color = city, group = city)) +  -->
-<!--     geom_hline(aes(yintercept = I_XY, color = city), size = 1, data = MI_to_show, linetype = 'dashed') +  -->
-<!--     theme_bw() +  -->
-<!--     scale_color_brewer(palette = 'Accent') +  -->
-<!--     ylab(expression(italic(I(C,Y)))) +  -->
-<!--     xlab(expression(italic(k))) +  -->
-<!--     ylim(c(0,NA)) +  -->
-<!--     geom_label(aes(x = k, y = val, label = city, fill = city), data = labels) +  -->
-<!--     scale_fill_brewer(palette = 'Accent') +  -->
-<!--     guides(color = FALSE, fill = FALSE)  -->
-
-<!-- ggsave('figs/loss_curves_m.pdf', height = map_height, width = map_width) -->
-
-<!-- ``` -->
+ggsave('figs/loss_curves_m.pdf',  width = map_width, height = map_height)
+```

### 2018-03-28 7_cluster_figs.Rmd

@@
 
-# knitr::read_chunk('R/utils.R')
-
 source('R/utils.R')
-
 ```
@@
 s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = .75)
+
 s
@@
 ```{r}
+
 f <- h_clust_map(m, tracts, data, b, j, size_coef = .75)
@@
 ```{r}
+
 set.seed(1250)
+
 k     <- 36
@@
 s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = .75)
-s
 
@@
 ```{r}
+
 f <- h_clust_map(m, tracts, data, b, j, size_coef = .75)
+
 f
@@
 
-<!-- # Atlanta -->
-
-<!-- ```{r} -->
-<!-- city    <- 'Atlanta' -->
-<!-- in_data <- read_data(city, all_data) -->
-<!-- data    <- in_data$data -->
-<!-- tracts  <- in_data$tracts -->
-
-<!-- b <- h_clust(tracts, data) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- g <- construct_information_graph(tracts = tracts,  -->
-<!-- 								 data = data,  -->
-<!-- 								 divergence = js_dist) -->
-
-<!-- sigma <- 30 -->
-<!-- plot_spectrum(g, sigma) -->
-<!-- ggsave(paste0('figs/spectrum_', city, '.pdf')) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- k     <- 26 -->
-<!-- g     <- spectral_cluster(g, sigma = sigma, k = k, nreps = 1000) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- input <- group_data(tracts, data, g)  -->
-<!-- blocked_tracts <- input$blocked_tracts -->
-<!-- blocked_data <- input$blocked_data -->
-
-<!-- blocked_tracts %>%  -->
-<!--     ggplot() +  -->
-<!--     geom_sf(aes(fill = as.character(cluster))) + -->
-<!--     guides(fill = FALSE) -->
-<!-- ``` -->
-
-<!-- ## Hierarchical Clustering -->
-
-<!-- ```{r} -->
-<!-- a <- h_clust(blocked_tracts, blocked_data) -->
-<!-- compare_hclusts(a, b) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- j <- 3 -->
-<!-- m <- prominence_map(tracts, data) +  -->
-<!--         guides(fill = FALSE)  -->
-
-<!-- s <- s_clust_map(m, tracts, data, blocked_tracts, size_coef = 1) -->
-<!-- s -->
-
-<!-- ggsave(paste0('figs/sclust_', city, '.pdf'), width = map_width, height = map_height) -->
-
-<!-- h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = 1) -->
-<!-- h -->
-<!-- ggsave(paste0('figs/mixed_', city,'.pdf'), width = map_width, height = map_height) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- f <- h_clust_map(m, tracts, data, b, j, size_coef = 1) -->
-<!-- f -->
-<!-- ggsave(paste0('figs/hclust_', city,'.pdf'), width = map_width, height = map_height) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- df <- df %>%  -->
-<!--     rbind(data_frame( -->
-<!--         city = city,  -->
-<!--         k = k,  -->
-<!--         j = j, -->
-<!--         hclust = cumsum(rev(b$height))[j], -->
-<!--         mixed = cumsum(rev(a$height))[j] -->
-<!--     )) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- loss_curves <- loss_curves %>% -->
-<!--     rbind(data_frame( -->
-<!--         city = city,  -->
-<!--         k = 1:length(a$height), -->
-<!--         h_val = rev(b$height)[1:length(a$height)], -->
-<!--         m_val = rev(a$height) -->
-<!--     )) -->
-
-<!-- MI <- MI %>%  -->
-<!--     rbind(data_frame( -->
-<!--         city = city,  -->
-<!--         I_XY = data %>%  -->
-<!--             spread(key = group, value = n) %>%  -->
-<!--             select(-tract) %>%  -->
-<!--             mutual_info() -->
-<!--     )) -->
-<!-- ``` -->
-
 # Philadelphia
@@
 set.seed(1234)
+
 ```
@@
 h <- h_clust_map(m, blocked_tracts, blocked_data, a, j, size_coef = .75)
+
 h
